### 2014-06-30 Quality_Reports/PCR_bottleneck.Rmd
First commit of PBC report

# PCR bottleneck report

## PCR bottleneck

According to the **ENCODE** definitions of [quality metrics](http://encodeproject.org/ENCODE/qualityMetrics.html#definitions), one of the metrics used is the *PCR bottleneck coefficient* or *PBC*.

The definition is:

A measure of library complexity, i.e. how skewed the distribution of read counts per location is towards 1 read per location.

PBC = N1/Nd

(where N1= number of genomic locations to which EXACTLY one unique mapping read maps, and Nd = the number of genomic locations to which **AT LEAST** one unique mapping read maps, i.e. the number of non-redundant, unique mapping reads).

In particular we can see that always N1 <= Nd, so 0<= PBC <= 1. ENCODE recomends the following classification:

| PBC range | Bottleneck class |
| :---:     | :---: |
|0 -0.5  | Severe |
| 0.5-0.8 | Moderate|
|0.8-0.9 | Mild |
|0.9 - 1| Non-existant|

For that purpose the following function was coded:

```{r PBC_fun,include = TRUE,results = "hide"}
PBC <- function(file)
{
  require(GenomicAlignments)
  message(file)
  rr = readGAlignmentsFromBam(file,param = NULL)
  ss1 = subset(rr,subset =strand(rr)=="+")
  ss2 = subset(rr,subset = strand(rr)=="-")
  N11 = length(unique(start(ss1)))
  N12 = length(unique(end(ss2)))
  N1 = N11+N12
  Nd1 = length(ss1)
  Nd2 = length(ss2)
  Nd = Nd1 + Nd2
  PBC1 = N11 / Nd1
  PBC2 = N12 / Nd2
  PBC = N1 /Nd
  message("PBC +:",round(PBC1,4))
  message("PBC -:",round(PBC2,4))
  message("PBC:",round(PBC,4))
  return(c(PBC_plus=PBC1,PBC_minus=PBC2,PBC=PBC))
}
```

### 2014-06-30 Quality_Reports/PCR_bottleneck.Rmd
test

@@
 
-According to the **ENCODE** definitions of [quality metrics](http://encodeproject.org/ENCODE/qualityMetrics.html#definitions), one of the metrics used is the *PCR bottleneck coefficient* or *PBC*.
+```{r package_options, include=FALSE}
+opts_knit$set(progress = TRUE, verbose = TRUE)
+```
+
+According to the **ENCODE** definitions of [quality metrics](http://encodeproject.org/ENCODE/qualityMetrics.html#definitions), one of the metrics used is the *PCR bottleneck coefficient* or **PBC**.
 
@@
 
-(where N1= number of genomic locations to which EXACTLY one unique mapping read maps, and Nd = the number of genomic locations to which **AT LEAST** one unique mapping read maps, i.e.
+(where N1= number of genomic locations to which **EXACTLY** one unique mapping read maps, and Nd = the number of genomic locations to which **AT LEAST** one unique mapping read maps, i.e.
 the number of non-redundant, unique mapping reads).
@@
 |0.9 - 1| Non-existant|
+
+```{r init,include = TRUE,echo = TRUE}
+library(parallel)
+library(GenomicAlignments)
+library(ggplot2)
+dr ="/NO_BACKUP/KelesGroup_DongjunChung/ChIP-exo/Landick_ChIP-exo3/rawdata"
+folder = c("ChIPexo","ChIPseq_PET","ChIPseq_SET")
+files = lapply(folder,function(x,dr){
+  ff = list.files(file.path(dr,x))
+  return(ff[!grepl("bai",ff) & !grepl("sam",ff)])},dr)
+names(files) = folder
+message(files[[1]])
+```
 

### 2014-06-30 Quality_Reports/PCR_bottleneck.Rmd
Loading data generated by PCR_bottleneck.R

@@
-# PCR bottleneck report
+## PCR bottleneck report
 
-## PCR bottleneck
+### PCR bottleneck
 
@@
 
-```{r init,include = TRUE,echo = TRUE}
+```{r init,include = FALSE,echo = TRUE,results = "hide"}
 library(parallel)
@@
 names(files) = folder
-message(files[[1]])
 ```
+### Code
 
-For that purpose the following function was coded:
+The following function was used to calculate the PBC:
 
-```{r PBC_fun,include = TRUE,results = "hide"}
+```{r PBC_fun,include = TRUE,results = "hide",echo = TRUE}
 PBC <- function(file)
@@
   require(GenomicAlignments)
-  message(file)
   rr = readGAlignmentsFromBam(file,param = NULL)
@@
   Nd = Nd1 + Nd2
-  PBC1 = N11 / Nd1
-  PBC2 = N12 / Nd2
-  PBC = N1 /Nd
-  message("PBC +:",round(PBC1,4))
-  message("PBC -:",round(PBC2,4))
-  message("PBC:",round(PBC,4))
+  PBC1 = round(N11 / Nd1,4)
+  PBC2 = round(N12 / Nd2,4)
+  PBC = round(N1 /Nd,4)
   return(c(PBC_plus=PBC1,PBC_minus=PBC2,PBC=PBC))  
@@
 ```
+And the following results where obtained:
+```{r load_data,include = FALSE,echo = TRUE,results = "hide"}
+load("../RData/PBC.RData")
+```
+
+```{r show,include = TRUE,echo = TRUE, results = "asis"}
+ls()
+```

### 2014-06-30 R/PCR_bottleneck.R

@@ -1,0 +1,49 @@
+
+rm(list = ls())
+
+
+library(knitr)
+library(parallel)
+library(GenomicAlignments)
+
+dr ="/NO_BACKUP/KelesGroup_DongjunChung/ChIP-exo/Landick_ChIP-exo3/rawdata"
+folder = c("ChIPexo","ChIPseq_PET","ChIPseq_SET")
+
+## Set bam files on list
+files = lapply(folder,function(x,dr){
+  ff = list.files(file.path(dr,x))
+  return(ff[!grepl("bai",ff) & !grepl("sam",ff)])},dr)
+names(files) = folder
+
+
+
+PBC <- function(file)
+{
+  require(GenomicAlignments)
+  rr = readGAlignmentsFromBam(file,param = NULL)
+  ss1 = subset(rr,subset =strand(rr)=="+")
+  ss2 = subset(rr,subset = strand(rr)=="-")
+  N11 = length(unique(start(ss1)))
+  N12 = length(unique(end(ss2)))
+  N1 = N11+N12
+  Nd1 = length(ss1)
+  Nd2 = length(ss2)
+  Nd = Nd1 + Nd2
+  PBC1 = round(N11 / Nd1,4)
+  PBC2 = round(N12 / Nd2,4)
+  PBC = round(N1 /Nd,4)
+  return(c(PBC_plus=PBC1,PBC_minus=PBC2,PBC=PBC))  
+}
+ 
+
+ff = lapply(folder,function(x,dr,files)file.path(dr,x,files[[x]]),dr,files)
+
+suppressWarnings(SET <- lapply(ff[[3]],FUN = PBC))
+names(SET) = c("Input +","Input -","ChIP -","ChIP +")
+
+
+suppressWarnings(PET <- lapply(ff[[2]],FUN = PBC))
+names(PET) = files[[2]]
+
+suppressWarnings(EXO <- lapply(ff[[1]],FUN = PBC))
+names(EXO) = files[[1]]

### 2014-06-30 Quality_Reports/PCR_bottleneck.Rmd
Added PBC tables

@@
 ```
+### Results
+
 And the following results where obtained:
@@
 
-```{r show,include = TRUE,echo = TRUE, results = "asis"}
-ls()
+#### ChIP - Seq - SET
+```{r show_set,include = TRUE,echo = FALSE, results = "asis"}
+kable(as.data.frame(t(do.call(cbind,SET))),format = "markdown")
 ```
+
+#### ChIP - Seq - PET
+```{r show_pet,include = TRUE,echo = FALSE, results = "asis"}
+kable(as.data.frame(t(do.call(cbind,PET))),format = "markdown")
+```
+#### ChIP - Exo
+```{r show_exo,include = TRUE,echo = FALSE, results = "asis"}
+kable(as.data.frame(t(do.call(cbind,EXO))),format = "markdown")
+```

### 2014-06-30 Quality_Reports/PCR_bottleneck.Rmd
Included overall boxplot

@@
 load("../RData/PBC.RData")
+tab = list()
+tab[[1]] = as.data.frame(t(do.call(cbind,SET)))
+tab[[1]]$type = "ChIP-Seq.SET"
+tab[[2]] = as.data.frame(t(do.call(cbind,PET)))
+tab[[2]]$type = "ChIP-Seq.PET"
+tab[[3]] = as.data.frame(t(do.call(cbind,EXO)))
+tab[[3]]$type = "ChIP-Exo"
 ```
+### Figures
+```{r overall_boxplot,include = TRUE,echo = FALSE,results = "hide",fig.height = 4,fig.width = 4}
+df = do.call(rbind,tab)
+p <- ggplot(df,aes(factor(type),PBC))+geom_boxplot(aes(fill = factor(type)))
+print(p)
+```
+
+### Tables
 
@@
 ```{r show_set,include = TRUE,echo = FALSE, results = "asis"}
-kable(as.data.frame(t(do.call(cbind,SET))),format = "markdown")
+kable(tab[[1]][,1:3],format = "markdown")
 ```
@@
 ```{r show_pet,include = TRUE,echo = FALSE, results = "asis"}
-kable(as.data.frame(t(do.call(cbind,PET))),format = "markdown")
+kable(tab[[2]][,1:3],format = "markdown")
 ```
+
 #### ChIP - Exo
 ```{r show_exo,include = TRUE,echo = FALSE, results = "asis"}
-kable(as.data.frame(t(do.call(cbind,EXO))),format = "markdown")
+kable(tab[[3]][,1:3],format = "markdown")
 ```

### 2014-06-30 Quality_Reports/PCR_bottleneck.Rmd
Changed overall_boxplot format

@@
 df = do.call(rbind,tab)
-p <- ggplot(df,aes(factor(type),PBC))+geom_boxplot(aes(fill = factor(type)))
+p <- ggplot(df,aes(factor(type),PBC))+geom_boxplot(aes(fill = factor(type)))+
+   theme(legend.position = "none",axis.text.x = element_text(angle = 90))
 print(p)

### 2014-06-30 Quality_Reports/PCR_bottleneck.Rmd
Added observations on figure 1

--- 2014-06-30 Quality_Reports/PCR_bottleneck.Rmd
@@
 ```
-### Results
 
-And the following results where obtained:
 ```{r load_data,include = FALSE,echo = TRUE,results = "hide"}
@@
 ### Figures
+
+**Figure 1**
 ```{r overall_boxplot,include = TRUE,echo = FALSE,results = "hide",fig.height = 4,fig.width = 4}
 df = do.call(rbind,tab)
-p <- ggplot(df,aes(factor(type),PBC))+geom_boxplot(aes(fill = factor(type)))
+df$type = factor(df$type)
+p <- ggplot(df,aes(type,PBC))+geom_boxplot(aes(fill = type))+
+   theme(legend.position = "none",axis.text.x = element_text(angle = 90))
 print(p)
 ```
+Some observations on the figure:
+- The overall level of the ChIP-Exo's PBC coefficients is lower than the level of the other two ChIP-Seq labels
+- The range of the ChIP-Exo's PBC coefficients is also smaller than other two ranges. In particular, the range of the PET coefficients is smaller than the range of the SET coefficients.
 

### 2014-06-30 Quality_Reports/PCR_bottleneck.Rmd
test on indentation

@@
 **Figure 1**
+
 ```{r overall_boxplot,include = TRUE,echo = FALSE,results = "hide",fig.height = 4,fig.width = 4}
@@
 ```
+
 Some observations on the figure:

### 2014-06-30 Quality_Reports/PCR_bottleneck.Rmd
Added new figure

@@
 
-**Figure 1**
+#### Figure 1
 
@@
 df$type = factor(df$type)
-p <- ggplot(df,aes(type,PBC))+geom_boxplot(aes(fill = type))+
+p <- ggplot(df,aes(type,PBC))+geom_boxplot()+
    theme(legend.position = "none",axis.text.x = element_text(angle = 90))
@@
 
-Some observations on the figure:
+We can see:
 - The overall level of the ChIP-Exo's PBC coefficients is lower than the level of the other two ChIP-Seq labels
-- The range of the ChIP-Exo's PBC coefficients is also smaller than other two ranges. In particular, the range of the PET coefficients is smaller than the range of the SET coefficients.
+- The range of the ChIP-Exo's PBC coefficients is also smaller than other two ranges.
+In particular, the interquartile range of the PET coefficients is smaller than the interquartile range of the SET coefficients.
+
+#### Figure 2
+
+```{r boxplot_strand,include = TRUE,echo = FALSE,results = "hide",fig.height = 4,fig.width = 12}
+tab1 = lapply(tab,function(x)rbind(data.frame(type = x$type,PBC = x$PBC_plus,strand = "+"),
+  data.frame(type = x$type,PBC = x$PBC_minus,strand = "-")))
+df1 = do.call(rbind,tab1)
+df1$type = factor(df1$type)
+df1$strand = factor(df1$strand)
+p1 <- ggplot(df1,aes(strand,PBC))+geom_boxplot()+facet_grid(.~type)+
+  theme(legend.position = "none",axis.text.x = element_text(angle =90))
+print(p1)
+```
 

### 2014-06-30 Quality_Reports/PCR_bottleneck.Rmd
Added observations of fig 2 and inverted order in facet

@@
 ```{r boxplot_strand,include = TRUE,echo = FALSE,results = "hide",fig.height = 4,fig.width = 12}
-tab1 = lapply(tab,function(x)rbind(data.frame(type = x$type,PBC = x$PBC_plus,strand = "+"),
+tab1 = lapply(tab[3:1],function(x)rbind(data.frame(type = x$type,PBC = x$PBC_plus,strand = "+"),
   data.frame(type = x$type,PBC = x$PBC_minus,strand = "-")))
@@
 ```
+
+- For this plot, the PBC coefficents was calculated using only the reads with certain strand. In particular, we can see that for all 3 types, the boxplots for both strands are very similar.
 

### 2014-06-30 Quality_Reports/PCR_bottleneck.Rmd
Added observation, changed orientation of x axis ticks

@@
 p1 <- ggplot(df1,aes(strand,PBC))+geom_boxplot()+facet_grid(.~type)+
-  theme(legend.position = "none",axis.text.x = element_text(angle =90))
+  theme(legend.position = "none")
 print(p1)
@@
 - For this plot, the PBC coefficents was calculated using only the reads with certain strand. In particular, we can see that for all 3 types, the boxplots for both strands are very similar.
+- The observations of figure 1 hold for both strands.
 

### 2014-06-30 Quality_Reports/PCR_bottleneck.Rmd
test about building an index
removed index test
width test strand boxplot

@@
 
-```{r boxplot_strand,include = TRUE,echo = FALSE,results = "hide",fig.height = 4,fig.width = 12}
+```{r boxplot_strand,include = TRUE,echo = FALSE,results = "hide",fig.height = 4}
 tab1 = lapply(tab[3:1],function(x)rbind(data.frame(type = x$type,PBC = x$PBC_plus,strand = "+"),

### 2014-07-01 Quality_Reports/PCR_bottleneck.Rmd
Changed the PBC code, to be based on table and include a cap

@@
 ```{r PBC_fun,include = TRUE,results = "hide",echo = TRUE}
-PBC <- function(file)
+PBC <- function(file,cap = Inf)
 {
   require(GenomicAlignments)
+  message(file)
   rr = readGAlignmentsFromBam(file,param = NULL)
@@
   ss2 = subset(rr,subset = strand(rr)=="-")
-  N11 = length(unique(start(ss1)))
-  N12 = length(unique(end(ss2)))
-  N1 = N11+N12
-  Nd1 = length(ss1)
-  Nd2 = length(ss2)
-  Nd = Nd1 + Nd2
-  PBC1 = round(N11 / Nd1,4)
-  PBC2 = round(N12 / Nd2,4)
-  PBC = round(N1 /Nd,4)
+  tab1 = table(start(ss1))
+  tab2 = table(end(ss2))
+  tab = table(c(start(ss1),end(ss2)))
+  PBC1 = round(sum(tab1 == 1)/sum(tab1 >= 1 & tab1 <= cap),4)
+  PBC2 = round(sum(tab2 == 1)/sum(tab2 >= 1 & tab2 <= cap),4)
+  PBC = round(sum(tab == 1) / sum(tab >= 1 & tab <= 4),4)
+  message("PBC +:",PBC1)
+  message("PBC -:",PBC2)
+  message("PBC:",PBC)
   return(c(PBC_plus=PBC1,PBC_minus=PBC2,PBC=PBC))  
@@
 - The overall level of the ChIP-Exo's PBC coefficients is lower than the level of the other two ChIP-Seq labels
-- The range of the ChIP-Exo's PBC coefficients is also smaller than other two ranges.
-In particular, the interquartile range of the PET coefficients is smaller than the interquartile range of the SET coefficients.
+- The level of ChIP-Seq PET is higher that the level of both SET generated y ChIP-Seq and ChIP-Exo
 

### 2014-07-01 Quality_Reports/PCR_bottleneck.Rmd
Fixed figure 2 and added observations

@@
 tab1 = lapply(tab[3:1],function(x)rbind(data.frame(type = x$type,PBC = x$PBC_plus,strand = "+"),
-  data.frame(type = x$type,PBC = x$PBC_minus,strand = "-")))
+  data.frame(type = x$type,PBC = x$PBC_minus,strand = "-"),data.frame(type = x$type, PBC = x$PBC,strand = "both")))
 df1 = do.call(rbind,tab1)
@@
 - The observations of figure 1 hold for both strands.
+- Also, we can see that for both ChIP-Seq types, the overall level of the PBC calculated using both strands is lower than the PBC level calculating only one strand (in particular the level of the PBC score is approximately the same for both strands).
+However, for ChIP-Exo, we can see that the level of the PBC score increases when both strands are joined.
 

### 2014-07-01 R/PCR_bottleneck.R

@@ -1,8 +1,6 @@
 
 
 rm(list = ls())
-
-
 library(knitr)
 library(parallel)
 library(GenomicAlignments)
@@ -18,24 +16,21 @@ names(files) = folder
 
 
 
-PBC <- function(file)
+PBC <- function(file,cap = Inf)
 {
   require(GenomicAlignments)
   message(file)
   rr = readGAlignmentsFromBam(file,param = NULL)
   ss1 = subset(rr,subset =strand(rr)=="+")
   ss2 = subset(rr,subset = strand(rr)=="-")
-  N11 = length(unique(start(ss1)))
-  N12 = length(unique(end(ss2)))
-  N1 = N11+N12
-  Nd1 = length(ss1)
-  Nd2 = length(ss2)
-  Nd = Nd1 + Nd2
-  PBC1 = round(N11 / Nd1,4)
+  tab1 = table(start(ss1))
+  tab2 = table(end(ss2))
+  tab = table(c(start(ss1),end(ss2)))
+  PBC1 = round(sum(tab1 == 1)/sum(tab1 >= 1 & tab1 <= cap),4)
+  PBC2 = round(sum(tab2 == 1)/sum(tab2 >= 1 & tab2 <= cap),4)
+  PBC = round(sum(tab == 1) / sum(tab >= 1 & tab <= 4),4)
   message("PBC +:",PBC1)
-  PBC2 = round(N12 / Nd2,4)
   message("PBC -:",PBC2)
-  PBC = round(N1 /Nd,4)
   message("PBC:",PBC)
   return(c(PBC_plus=PBC1,PBC_minus=PBC2,PBC=PBC))  
 }
@@ -43,11 +38,13 @@ PBC <- function(file)
 
 ff = lapply(folder,function(x,dr,files)file.path(dr,x,files[[x]]),dr,files)
 
-suppressWarnings(SET <- lapply(ff[[3]],FUN = PBC))
+cap = Inf
+
+suppressWarnings(SET <- mclapply(ff[[3]],FUN = PBC,cap=cap,mc.cores =8))
 names(SET) = files[[3]]
-suppressWarnings(PET <- lapply(ff[[2]],FUN = PBC))
+suppressWarnings(PET <- mclapply(ff[[2]],FUN = PBC,cap=cap,mc.cores = 8))
 names(PET) = files[[2]]
-suppressWarnings(EXO <- lapply(ff[[1]],FUN = PBC))
[…]

### 2014-07-14 Quality_Reports/PCR_bottleneck.Rmd
Change calculation of PCR to use spp package

@@
   require(GenomicAlignments)
+  require(spp)
   message(file)
@@
   PBC2 = round(sum(tab2 == 1)/sum(tab2 >= 1 & tab2 <= cap),4)
-  PBC = round(sum(tab == 1) / sum(tab >= 1 & tab <= 4),4)
+  chip.data = read.bam.tags(file)
+  table.chip.data = lapply(chip.data$tags,table)
+  nUniq = sum(sapply(table.chip.data,function(x) sum(x==1)))
+  nTotal =  sum(sapply(table.chip.data,length))
+  PBC = round(nUniq/nTotal ,4)
   message("PBC +:",PBC1)

### 2014-07-14 R/PCR_bottleneck.R

@@ -4,6 +4,8 @@ rm(list = ls())
 library(knitr)
 library(parallel)
 library(GenomicAlignments)
+library(spp)
+
 
 dr ="/NO_BACKUP/KelesGroup_DongjunChung/ChIP-exo/Landick_ChIP-exo3/rawdata"
 folder = c("ChIPexo","ChIPseq_PET","ChIPseq_SET")
@@ -28,7 +30,11 @@ PBC <- function(file,cap = Inf)
   tab = table(c(start(ss1),end(ss2)))
   PBC1 = round(sum(tab1 == 1)/sum(tab1 >= 1 & tab1 <= cap),4)
   PBC2 = round(sum(tab2 == 1)/sum(tab2 >= 1 & tab2 <= cap),4)
-  PBC = round(sum(tab == 1) / sum(tab >= 1 & tab <= 4),4)
+  chip.data = read.bam.tags(file)
+  table.chip.data = lapply(chip.data$tags,table)
+  nUniq = sum(sapply(table.chip.data,function(x) sum(x==1)))
+  nTotal =  sum(sapply(table.chip.data,length))
+  PBC = round(nUniq/nTotal ,4)
   message("PBC +:",PBC1)
   message("PBC -:",PBC2)
   message("PBC:",PBC)
@@ -48,3 +54,8 @@ suppressWarnings(EXO <- mclapply(ff[[1]],FUN = PBC,cap=cap,mc.cores =8))
 names(EXO) = files[[1]]
 
 save(list = c("SET","PET","EXO"),file = "../RData/PBC.RData")
+
+
+
+
+

### 2014-07-14 Quality_Reports/PCR_bottleneck.Rmd
Changed final observation in figure 2

@@
 - The observations of figure 1 hold for both strands.
-- Also, we can see that for both ChIP-Seq types, the overall level of the PBC calculated using both strands is lower than the PBC level calculating only one strand (in particular the level of the PBC score is approximately the same for both strands).
-However, for ChIP-Exo, we can see that the level of the PBC score increases when both strands are joined.
+- We can see that for all data types: The PBC level for each individual strand is very close to the one calculated using both strands.
+- After inspecting the calculated values in the tables below, we can see that for both ChIP-Seq tags, the level of the PBC calculated with both strands is in between of the PBC's calculated using only one strand.
+However, for most of the ChIP-Exo samples, we can see that the PBC calculated with both strand is slightly higher than the PBC calculated for both individual strands.
 

### 2014-07-22 Quality_Reports/Strand_exploration.Rmd
First commit cross - correlation per strand

--- 2014-07-01 Quality_Reports/PCR_bottleneck.Rmd
@@
-## PCR bottleneck report
+## Exploratory analysis of the strand
 
-### PCR bottleneck
+### Description of the experiment
 
@@
 
-According to the **ENCODE** definitions of [quality metrics](http://encodeproject.org/ENCODE/qualityMetrics.html#definitions), one of the metrics used is the *PCR bottleneck coefficient* or **PBC**.
-
-The definition is:
-
-A measure of library complexity, i.e. how skewed the distribution of read counts per location is towards 1 read per location.
-
-PBC = N1/Nd
-
-(where N1= number of genomic locations to which **EXACTLY** one unique mapping read maps, and Nd = the number of genomic locations to which **AT LEAST** one unique mapping read maps, i.e.
-the number of non-redundant, unique mapping reads).
-
-In particular we can see that always N1 <= Nd, so 0<= PBC <= 1. ENCODE recomends the following classification:
-
-| PBC range | Bottleneck class |
-| :---:     | :---: |
-|0 -0.5  | Severe |
-| 0.5-0.8 | Moderate|
-|0.8-0.9 | Mild |
-|0.9 - 1| Non-existant|
-
-```{r init,include = FALSE,echo = TRUE,results = "hide"}
-library(parallel)
-library(GenomicAlignments)
-library(ggplot2)
-dr ="/NO_BACKUP/KelesGroup_DongjunChung/ChIP-exo/Landick_ChIP-exo3/rawdata"
-folder = c("ChIPexo","ChIPseq_PET","ChIPseq_SET")
-files = lapply(folder,function(x,dr){
-  ff = list.files(file.path(dr,x))
-  return(ff[!grepl("bai",ff) & !grepl("sam",ff)])},dr)
-names(files) = folder
-```
-### Code
-
-The following function was used to calculate the PBC:
-
-```{r PBC_fun,include = TRUE,results = "hide",echo = TRUE}
-PBC <- function(file,cap = Inf)
-{
-  require(GenomicAlignments)
-  message(file)
-  rr = readGAlignmentsFromBam(file,param = NULL)
-  ss1 = subset(rr,subset =strand(rr)=="+")
-  ss2 = subset(rr,subset = strand(rr)=="-")
-  tab1 = table(start(ss1))
-  tab2 = table(end(ss2))
-  tab = table(c(start(ss1),end(ss2)))
-  PBC1 = round(sum(tab1 == 1)/sum(tab1 >= 1 & tab1 <= cap),4)
-  PBC2 = round(sum(tab2 == 1)/sum(tab2 >= 1 & tab2 <= cap),4)
-  PBC = round(sum(tab == 1) / sum(tab >= 1 & tab <= 4),4)
-  message("PBC +:",PBC1)
-  message("PBC -:",PBC2)
-  message("PBC:",PBC)
-  return(c(PBC_plus=PBC1,PBC_minus=PBC2,PBC=PBC))  
-}
-```
-
-```{r load_data,include = FALSE,echo = TRUE,results = "hide"}
-load("../RData/PBC.RData")
-tab = list()
-tab[[1]] = as.data.frame(t(do.call(cbind,SET)))
-tab[[1]]$type = "ChIP-Seq.SET"
-tab[[2]] = as.data.frame(t(do.call(cbind,PET)))
-tab[[2]]$type = "ChIP-Seq.PET"
-tab[[3]] = as.data.frame(t(do.call(cbind,EXO)))
-tab[[3]]$type = "ChIP-Exo"
-```
-### Figures
-
-#### Figure 1
-
-```{r overall_boxplot,include = TRUE,echo = FALSE,results = "hide",fig.height = 4,fig.width = 4}
-df = do.call(rbind,tab)
-df$type = factor(df$type)
-p <- ggplot(df,aes(type,PBC))+geom_boxplot()+
-   theme(legend.position = "none",axis.text.x = element_text(angle = 90))
-print(p)
-```
-
-We can see:
-- The overall level of the ChIP-Exo's PBC coefficients is lower than the level of the other two ChIP-Seq labels
-- The level of ChIP-Seq PET is higher that the level of both SET generated y ChIP-Seq and ChIP-Exo
-
-#### Figure 2
-
-```{r boxplot_strand,include = TRUE,echo = FALSE,results = "hide",fig.height = 4}
-tab1 = lapply(tab[3:1],function(x)rbind(data.frame(type = x$type,PBC = x$PBC_plus,strand = "+"),
-  data.frame(type = x$type,PBC = x$PBC_minus,strand = "-"),data.frame(type = x$type, PBC = x$PBC,strand = "both")))
-df1 = do.call(rbind,tab1)
-df1$type = factor(df1$type)
-df1$strand = factor(df1$strand)
-p1 <- ggplot(df1,aes(strand,PBC))+geom_boxplot()+facet_grid(.~type)+
-  theme(legend.position = "none")
-print(p1)
-```
-
-- For this plot, the PBC coefficents was calculated using only the reads with certain strand. In particular, we can see that for all 3 types, the boxplots for both strands are very similar.
-- The observations of figure 1 hold for both strands.
-- Also, we can see that for both ChIP-Seq types, the overall level of the PBC calculated using both strands is lower than the PBC level calculating only one strand (in particular the level of the PBC score is approximately the same for both strands).
-However, for ChIP-Exo, we can see that the level of the PBC score increases when both strands are joined.
-
-### Tables
-
-#### ChIP - Seq - SET
-```{r show_set,include = TRUE,echo = FALSE, results = "asis"}
-kable(tab[[1]][,1:3],format = "markdown")
-```
-
-#### ChIP - Seq - PET
-```{r show_pet,include = TRUE,echo = FALSE, results = "asis"}
-kable(tab[[2]][,1:3],format = "markdown")
-```
-
-#### ChIP - Exo
-```{r show_exo,include = TRUE,echo = FALSE, results = "asis"}
-kable(tab[[3]][,1:3],format = "markdown")
-```
+![fig](figure/fromPaper/Kharchenko-Park-2008.jpg)

### 2014-07-22 Quality_Reports/Strand_exploration.Rmd
First paragraph, removed picture

@@
-## Exploratory analysis of the strand
+## Strand exploratory analysis
+
+### What is the idea behind the strand cross - correlation
+
+According to [ChIP-seq guidelines and practices of the ENCODE and modENCODE consortia](http://www.ncbi.nlm.nih.gov/pmc/articles/PMC3431496/), a useful quality metric is the use of strand cross - correlation.
+Which is based in the fact that a high-quality ChIP-Seq experiment produces significant clustering of enriched DNA sequence tags at locations bound by the protein of interest, and that the sequence tag density acumulates in both strand centered around the binding site.
 
@@
 ```
-
-![fig](figure/fromPaper/Kharchenko-Park-2008.jpg)

### 2014-07-23 Quality_Reports/Strand_exploration.Rmd
Added experiment design table

@@
 ```
+
+```{r init,include = FALSE,echo = TRUE,results = "hide"}
+library(GenomicAlignments)
+library(parallel)
+library(knitr)
+dr ="/NO_BACKUP/KelesGroup_DongjunChung/ChIP-exo/Landick_ChIP-exo3/rawdata"
+folder = c("ChIPexo","ChIPseq_PET","ChIPseq_SET")
+files = lapply(folder,function(x,dr){
+  ff = list.files(file.path(dr,x))
+  return(ff[!grepl("bai",ff) & !grepl("sam",ff)])},dr)
+names(files) = folder
+ff = lapply(folder,function(x,dr,files)file.path(dr,x,files[[x]]),dr,files)
+names(ff) = folder
+df = data.frame(Rep = c(rep(1,4),rep(2,4)),Rif = rep(c(0,0,20,20),2),IP = rep(c("Sigma70","BetaPrimeFlag"),4),
+  exo.edns = c(1311,1312,1314,1315,1317,1318,1320,1321),pet.edsn = seq(1396,1403))
+```
+```{r show_df,include = TRUE,echo = FALSE,results = "asis"}
+kable(df,format = "markdown")
+```

### 2014-07-23 Quality_Reports/Strand_exploration.Rmd
Added betaF table to R code

--- 2014-07-22 Quality_Reports/Strand_exploration.Rmd
@@
 ```
+
+```{r init,include = FALSE,echo = TRUE,results = "hide"}
+library(GenomicAlignments)
+library(parallel)
+library(knitr)
+dr ="/NO_BACKUP/KelesGroup_DongjunChung/ChIP-exo/Landick_ChIP-exo3/rawdata"
+folder = c("ChIPexo","ChIPseq_PET","ChIPseq_SET")
+files = lapply(folder,function(x,dr){
+  ff = list.files(file.path(dr,x))
+  return(ff[!grepl("bai",ff) & !grepl("sam",ff)])},dr)
+names(files) = folder
+ff = lapply(folder,function(x,dr,files)file.path(dr,x,files[[x]]),dr,files)
+names(ff) = folder
+design = data.frame(Rep = c(rep(1,4),rep(2,4)),Rif = rep(c(0,0,20,20),2),IP = rep(c("Sigma70","BetaPrimeFlag"),4),
+  exo.edns = c(1311,1312,1314,1315,1317,1318,1320,1321),pet.edsn = seq(1396,1403))
+betaF = data.frame(Rep = c(1,1,2,2),Rif = c(0,20,0,20),IP = rep("Beta",4),exo.edns = c(1310,1313,1316,1319))
+```
+
+```{r show_df,include = TRUE,echo = FALSE,results = "asis"}
+kable(design,format = "markdown")
+```

### 2014-07-23 Quality_Reports/Strand_exploration.Rmd
added figure 5 from carrol et al 2014 to document

--- 2014-07-22 Quality_Reports/Strand_exploration.Rmd
@@
 
+A graphical explanation of this effect is given in the figure 5 of [Carroll et. al. 2014's paper](http://journal.frontiersin.org/Journal/10.3389/fgene.2014.00075/full):
+
+![plot of fraglen est. by cross - corr.](figure/fromPaper/figure5-carroll-2014.jpg)
+
+In this figure we can see:
+
 ### Description of the experiment
@@
 ```
+
+```{r init,include = FALSE,echo = TRUE,results = "hide"}
+library(GenomicAlignments)
+library(parallel)
+library(knitr)
+dr ="/NO_BACKUP/KelesGroup_DongjunChung/ChIP-exo/Landick_ChIP-exo3/rawdata"
+folder = c("ChIPexo","ChIPseq_PET","ChIPseq_SET")
+files = lapply(folder,function(x,dr){
+  ff = list.files(file.path(dr,x))
+  return(ff[!grepl("bai",ff) & !grepl("sam",ff)])},dr)
+names(files) = folder
+ff = lapply(folder,function(x,dr,files)file.path(dr,x,files[[x]]),dr,files)
+names(ff) = folder
+design = data.frame(Rep = c(rep(1,4),rep(2,4)),Rif = rep(c(0,0,20,20),2),IP = rep(c("Sigma70","BetaPrimeFlag"),4),
+  exo.edns = c(1311,1312,1314,1315,1317,1318,1320,1321),pet.edsn = seq(1396,1403))
+betaF = data.frame(Rep = c(1,1,2,2),Rif = c(0,20,0,20),IP = rep("Beta",4),exo.edns = c(1310,1313,1316,1319))
+start = c(576875,798921,1152674,1192831,1820005,2310506,2996662,3351853,3973436,4262028)
+end = c(577492,799434,1153291,1193448,1820696,2311153,2997175,3352366,3973949,4262541)
+peaks = GRanges(seqnames = "U00096",ranges = IRanges(start = start,end = end),strand = "*")
+width(peaks) = 1000
+```
+
+```{r show_df,include = TRUE,echo = FALSE,results = "asis"}
+kable(design,format = "markdown")
+```

### 2014-07-23 Quality_Reports/Strand_exploration.Rmd
peaks table and first figure observations added

--- 2014-07-23 Quality_Reports/Strand_exploration.Rmd
@@
 Which is based in the fact that a high-quality ChIP-Seq experiment produces significant clustering of enriched DNA sequence tags at locations bound by the protein of interest, and that the sequence tag density acumulates in both strand centered around the binding site.
+
+A graphical explanation of this effect is given in the figure 5 of [Carroll et. al. 2014's paper](http://journal.frontiersin.org/Journal/10.3389/fgene.2014.00075/full):
+
+![plot of fraglen est. by cross - corr.](figure/fromPaper/figure5-carroll-2014.jpg)
+
+In this figure we can see an ideal case of the fragment length estimation by the use of the cross correlation function.
+On the top panel we can see a very smooth cross correlation function that shows both the an artifact peak and a fragment peak.
+In the bottom panels we can see (from left to right):
+- An under estimation of the shift between both strands, which makes the pooled peak of both strands wider and therefore harder to recognize the binding event signal.
+In this case the fragment length is under-estimated as a the shift size where the artifact peak is ocurring.
+- Just right, the shift is estimated at the peak of the cross correlation function. Which makes the signal to be higher and more recognizable when both strand reads are pooled.
+- The shift is over-estimated. Which makes to have a "bimodal" peak when both strands are pooled.
 
@@
 names(ff) = folder
-df = data.frame(Rep = c(rep(1,4),rep(2,4)),Rif = rep(c(0,0,20,20),2),IP = rep(c("Sigma70","BetaPrimeFlag"),4),
+design = data.frame(Rep = c(rep(1,4),rep(2,4)),Rif = rep(c(0,0,20,20),2),IP = rep(c("Sigma70","BetaPrimeFlag"),4),
   exo.edns = c(1311,1312,1314,1315,1317,1318,1320,1321),pet.edsn = seq(1396,1403))
+betaF = data.frame(Rep = c(1,1,2,2),Rif = c(0,20,0,20),IP = rep("Beta",4),exo.edns = c(1310,1313,1316,1319))
+start = c(576875,798921,1152674,1192831,1820005,2310506,2996662,3351853,3973436,4262028)
+end = c(577492,799434,1153291,1193448,1820696,2311153,2997175,3352366,3973949,4262541)
+peaks = GRanges(seqnames = "U00096",ranges = IRanges(start = start,end = end),strand = "*")
+width(peaks) = 1000
 ```
+In this experiment we are trying to check how the effect explained above occurrs for reads generated by the ChIP - exo protocol.
+For that purpose, the following regions where visually located using the e.
+coli genome viewer Mochi-view:
+
+```{r show_peaks,include= TRUE,echo = FALSE,results = "asis"}
+kable(as.data.frame(peaks),format = "markdown")
+```
+
+For the inspection the following structures where considered:
+
 ```{r show_df,include = TRUE,echo = FALSE,results = "asis"}
-kable(df,format = "markdown")
+kable(design,format = "markdown")
+kable(betaF,format = "markdown")
 ```

### 2014-07-23 Quality_Reports/Strand_exploration.Rmd
Added function of how to the coverage was calculated

@@
 kable(design,format = "markdown")
-kable(betaF,format = "markdown")
 ```
+
+Using the following function, we generated the strand coverage across for all peaks, for all data sets. The coverage was calculated using the function:
+
+```{r coverage_fun,include = TRUE,results = "hide",echo = TRUE}
+strand_coverage <- function(reads,peak,st, fragLen = 0)
+{
+  reads = as(reads,"GRanges")
+  reads = subset(reads, subset = strand(reads) == st)
+  if(fragLen >0){
+    reads = resize(reads,fragLen)
+  }
+  match = subjectHits(findOverlaps(peak,reads))
+  cover = coverage(reads[match])[[1]]
+  x = seq(start(peak),end(peak))
+  if(nrun(cover) == 1){
+    y = rep(runValue(cover),length(x))
+  }else{
+    xp = cumsum(runLength(cover)[1:(nrun(cover)-1)])
+    yp = runValue(cover)
+    y = stepfun(xp,yp)(x)
+  }
+  return(y)
+}
+```

### 2014-10-14 inst/rscripts/Script_conditionalDensity.R
Added section to vignette on conditional densities

@@ -7,15 +7,136 @@ library(knitr)
 
 load("data/chip.exo.RData")
 load("data/chip.seq.pet.RData")
+load("data/chip.seq.set.from.pet.RData")
 load("data/sample.summary.RData")
 
 source("R/density_functions.R")
 source("R/conditional_density_functions.R")
 
 # param
-binsizes = c(200,500,750,1000)
-probs = c(0,0.5,.75,0.9,0.95,0.975)
 figsdir = "inst/figs/cond_densities"
+binsizes = c(300,500,750)
+bins = lapply(binsizes,create.bins,genomeLength=seqlengths(exo[[1]]))
+names(bins) = as.character(binsizes)
+
+
+counts_strand <- function(bins,reads)
+{
+  counts_F = countOverlaps(bins,subset(reads,subset = strand(reads) == "+"))
+  counts_R = countOverlaps(bins,subset(reads,subset = strand(reads) == "-"))
+  return(DataFrame(counts_F,counts_R))
+}
+
+
+exo_strand_counts = mclapply(exo,function(x,bins)lapply(bins,counts_strand,x),bins,mc.cores=6)
+pet_strand_counts = mclapply(pet,function(x,bins)lapply(bins,counts_strand,x),bins,mc.cores=6)
+set_strand_counts = mclapply(set,function(x,bins)lapply(bins,counts_strand,x),bins,mc.cores=6)
+
+exo_strand_counts = unlist(exo_strand_counts)
+pet_strand_counts = unlist(pet_strand_counts)
+set_strand_counts = unlist(set_strand_counts)
+
+exo_counts = mclapply(exo_strand_counts,function(x)do.call("+",x),mc.cores =6)
+pet_counts = mclapply(pet_strand_counts,function(x)do.call("+",x),mc.cores =6)
+set_counts = mclapply(set_strand_counts,function(x)do.call("+",x),mc.cores =6)
+
+probs = c(0.9,0.95,0.975,.99)
+exo_quantiles = mclapply(exo_counts,function(x)floor(quantile(x,probs)),mc.cores = 6)
+
+pet_quantiles = mclapply(pet_counts,function(x)floor(quantile(x,probs)),mc.cores = 6)
+
+set_quantiles = mclapply(set_counts,function(x)floor(quantile(x,probs)),mc.cores = 6)
+
+
+fwd_strand_ratio <- function(f,r)return((f+1)/(f+r+2))
+
[…]

### 2014-10-14 vignettes/Summary_ChIPexo.R

@@ -71,3 +71,26 @@ idd = do.call(c,lapply(row$edsn,function(x)grep(x,names(pet_density))))
   ggplot(df,aes(x.x,x.y,colour = bin))+geom_line()+theme(legend.position = "bottom",aspect.ratio = 1,axis.title = element_text(size = rel(2.5)),axis.text = element_text(size = rel(2),angle=90),legend.text = element_text(size = rel(2)),legend.title = element_text(size = rel(2)))+ scale_color_brewer(palette ="Dark2")+facet_grid(edsn~seq)+scale_x_continuous(limits = c(0,1))+xlab("fwd. strand ratio")+ylab("density")
   ggsave("seqDens.pdf",width = 6,height = 6)  
 
+## ----include=FALSE,echo=FALSE,eval=TRUE-------------------------------------------------
+load(file = "../data/cond_densities.RData")
+fix_set <- function(cond_densities)
+{
+  cond_density = do.call(rbind,cond_densities)
+  cond_density$edsn = factor(cond_density$edsn)
+  cond_density$bin  = factor(cond_density$bin )
+  cond_density$prob = factor(cond_density$prob)
+  return(cond_density)
+}
+exo_cond_densities = lapply(exo_cond_densities,as.data.frame)
+exo_cond_density = fix_set(exo_cond_densities)
+
+## ----include=FALSE,echo=FALSE,eval=TRUE-------------------------------------------------
+  df = subset(exo_cond_density,edsn == "edsn1310")
+  ggplot(df,aes(x.x,x.y,colour = prob))+geom_line()+theme(legend.position = "bottom",aspect.ratio = 1,axis.title = element_text(size = rel(2.5)),axis.text = element_text(size = rel(2),angle=90),legend.text = element_text(size = rel(2)),legend.title = element_text(size = rel(2)))+ scale_color_brewer(palette ="Dark2")+facet_grid(edsn ~ bin)+scale_x_continuous(limits = c(0,1))+xlab("fwd. strand ratio")+ylab("density") 
+  ggsave("exo1310.pdf",width = 12,height = 6)  
+
+## ----include=FALSE,echo=FALSE,eval=TRUE-------------------------------------------------
+  df = subset(exo_cond_density,edsn == "edsn1319")
[…]

### 2014-10-20 vignettes/Summary_ChIPexo.R
Last version, modified plots

@@ -53,7 +53,7 @@ rm(rows)
 ## ----include=FALSE,echo=FALSE,eval=TRUE-------------------------------------------------
   df = do.call(rbind, lapply(exo_density[idd],as.data.frame))
   idd
-  ggplot(df,aes(x.x,x.y,colour = bin))+geom_line()+theme(legend.position = "bottom",aspect.ratio = 1,axis.title = element_text(size = rel(2.5)),axis.text = element_text(size = rel(2),angle=90),legend.text = element_text(size = rel(2)),legend.title = element_text(size = rel(2)))+ scale_color_brewer(palette ="Dark2")+facet_grid(.~edsn)+scale_x_continuous(limits = c(0,1))+xlab("fwd. strand ratio")+ylab("density")
+  ggplot(df,aes(x.x,x.y,colour = bin))+geom_line()+theme(legend.position = "bottom",aspect.ratio = 1,axis.title = element_text(size = rel(2.5)),axis.text = element_text(size = rel(2),angle=90),legend.text = element_text(size = rel(2)),legend.title = element_text(size = rel(2)),strip.text = element_text(size = rel(2)))+ scale_color_brewer(palette ="Dark2")+facet_grid(.~edsn)+scale_x_continuous(limits = c(0,1))+xlab("fwd. strand ratio")+ylab("density")
   ggsave("exoDens.pdf",width = 6,height = 6)
   rm(idd)
 
@@ -68,7 +68,7 @@ idd = do.call(c,lapply(row$edsn,function(x)grep(x,names(pet_density))))
   df2$seq = "SET"
   df = rbind(df1,df2)
   df$seq = factor(df$seq)
-  ggplot(df,aes(x.x,x.y,colour = bin))+geom_line()+theme(legend.position = "bottom",aspect.ratio = 1,axis.title = element_text(size = rel(2.5)),axis.text = element_text(size = rel(2),angle=90),legend.text = element_text(size = rel(2)),legend.title = element_text(size = rel(2)))+ scale_color_brewer(palette ="Dark2")+facet_grid(edsn~seq)+scale_x_continuous(limits = c(0,1))+xlab("fwd. strand ratio")+ylab("density")
[…]

### 2015-02-20 vignettes/chipexo_figure_summary.R
This files are the first draft of the chip-exo poster figure summary.

@@ -1,0 +1,122 @@
+## ----<style-knitr, eval=TRUE, echo=FALSE, results="asis"--------------------------------
+  BiocStyle::latex()
+
+## ----init,include=FALSE,echo=FALSE,eval=TRUE,results='hide'-----------------------------
+
+  library(knitr)
+  library(data.table)
+  library(devtools)
+
+  load_all("/p/keles/ChIPexo/volume4/ChIPexoQual")
+
+
+
+## ----depth,include=FALSE,echo=FALSE,eval=TRUE,results='hide'----------------------------
+
+  datadir = "/p/keles/ChIPexo/volume3/Analysis"
+  rensamples = c("H3k27ac","H3k4me1-rep1","H3k4me1-rep2")
+  rendir = file.path(datadir,"Ren",rensamples,"data")
+  carrollsamples = c(paste0("ER-rep",1:3),paste0("FoxA1-rep",1:3))
+  carrolldir = file.path(datadir,"Carroll",rep(c("human","mouse"),each=3),
+    carrollsamples,"data")
+  rifsamples = paste0(rep(c("BetaPrimeFlag","Beta","Sig70"),each=4),"-rif",
+    rep(c(0,20),each=2),"-rep",1:2)
+  statsamples = c("Input",paste0(rep(c("Sig70","SigmaS"),each=4),rep(c("-exp-","-stat-"),each=2),
+    "rep",rep(1:2,4)))
+  landicksamples = c(rifsamples,statsamples)
+  landickdir = file.path(datadir,"Landick",c(rep("rif-treatment",12),rep("stat-vs-exp",9)),
+    landicksamples,"data")
+
+  load_depth <- function(direc,samp){
+    load(file.path(direc,paste0(samp,"_depth.RData")))
+    return(depth)}
+
+  rendepths = mapply(load_depth,rendir,rensamples)
+  names(rendepths) = rensamples
+
+  carrolldepths = mapply(load_depth,carrolldir,carrollsamples)
+  names(carrolldepths) = carrollsamples
+
+  landickdepths = mapply(load_depth,landickdir,landicksamples)
+  names(landickdepths) = landicksamples
+
+
+## ----pbc,include=FALSE,echo=FALSE,eval=TRUE,results='hide'------------------------------
+
+  load_PBC <- function(direc,samp){    
+    load(file.path(direc,paste0(samp,"_PBC.RData")))
+  return(pcr_coeff)}
+
[…]

### 2015-09-08 vignettes/chipexo_figure_summary.R
Last vignette, adding for completeness

@@ -14,6 +14,8 @@
 ## ----depth,include=FALSE,echo=FALSE,eval=TRUE,results='hide'----------------------------
 
   datadir = "/p/keles/ChIPexo/volume3/Analysis"
+  pughsamples = "CTCF"
+  pughdir = file.path(datadir,"Pugh",pughsamples,"data")
   rensamples = c("H3k27ac","H3k4me1-rep1","H3k4me1-rep2")
   rendir = file.path(datadir,"Ren",rensamples,"data")
   carrollsamples = c(paste0("ER-rep",1:3),paste0("FoxA1-rep",1:3))
@@ -31,6 +33,9 @@
     load(file.path(direc,paste0(samp,"_depth.RData")))
     return(depth)}
 
+  pughdepths = mapply(load_depth,pughdir,pughsamples)
+  names(pughdepths) = pughsamples
+
   rendepths = mapply(load_depth,rendir,rensamples)
   names(rendepths) = rensamples
 
@@ -46,6 +51,9 @@
   load_PBC <- function(direc,samp){    
     load(file.path(direc,paste0(samp,"_PBC.RData")))
   return(pcr_coeff)}
+
+  pughpbc = mapply(load_PBC,pughdir,pughsamples)
+  names(pughpbc) =pughsamples
 
   renpbc= mapply(load_PBC,rendir,rensamples)
   names(renpbc) = rensamples
@@ -64,6 +72,9 @@
     load(file.path(direc,paste0(samp,"_cross.corr.RData")))
   return(cc)}
 
+  pughcc = mapply(load_cc,pughdir,pughsamples,SIMPLIFY=FALSE)
+  names(pughcc) = pughsamples
+
   rencc= mapply(load_cc,rendir,rensamples,SIMPLIFY=FALSE)
   names(rencc) = rensamples
 
@@ -73,8 +84,9 @@
   landickcc = mapply(load_cc,landickdir,landicksamples,SIMPLIFY=FALSE)
   names(landickcc) = landicksamples
 
-  nsc <- function(cc)cc[,max(cross.corr)/min(cross.corr)]
+  nsc <- function(cc)cc[between(shift,1,300),max(cross.corr)/min(cross.corr)]
 
+  pughnsc = sapply(pughcc,nsc)
   rennsc = sapply(rencc,nsc)
   carrollnsc = sapply(carrollcc,nsc)
   landicknsc = sapply(landickcc,nsc)
@@ -84,11 +96,12 @@
[…]

### 2015-10-22 rmakdown/ChIPexo_figures_list.Rmd
list of current figures and explanation

---
title: "High Resolution Identification of Protein-DNA Binding Events using ChIP-exo and dPeak"
author: "Rene Welch"
output: html_document
---

The idea of this, is to list the figures that we currently have in the manuscript and try to figure it out the idea behind each figure.

# Abstract

- Background in different in ChIPseq from ChIPexo, but sequence biases exists in ChIPseq and they exists also in ChIPexo

- In ChIPexo reads are allocated more tightly around binding sites and hence, it has more potential for high resolution

- Peak pair assumption does not hold well in ChIPexo

- dPeak makes a striking balance between sensitivity, specificity and spatial resolution for ChIPexo

- Spatial resolution is comparable to ChIPseq PET and it is significantly better than resolution of ChIPseq SET

- ChIPexo provides higher sensitivity, specificity and spatial resolution any fixed depth

## List of figures

### Figure 1 - Comparison of resolution between ChIPexo and ChIPseq data

A. This figure compares the sensitivity of all three protocols. For peaks with more than one binding site, the average distance between binding sites is calculated and plotted again the proportion of binding sites that is finding in each peak.
It is expected to see that as the average distance is increased, the methods is going to be more sensitive i.e. is gonna be more capable of differentiating all the binding events.

B. Peak resolution by dataset. For each peak, the resolution is defined as the distance between a regulonDB annotation and its closed prediction. Two points: The mean resolution value is ordered as exo < pet < set and the resolution's range is ordered as pet < exo < set. In both cases, ChIPseq SET shows a higher resolution in both moments.

### Figure 2 - Recommendations for the design of ChIPexo experiments

In all this four figures we are sampling  $N$ reads (or $N$ pairs for the case of ChIPseq PET) and running those sampled datasets across the whole pipeline (i.e. binning, call peaks and then estimate binding events). Using those outputs (both peaks and binding sites) we compare the average behaviour as the $N$ is increasing and between the different technologies.

A. Candidate regions. Overall in this plot as the number of fragments increase, the number of regions is increasing too. With both ChIPseq methods, the number of regions is roughly the same and in higher than with ChIPexo. This implies that with ChIPexo it may be necessary to generate datasets with higher depth.

B. Predicted events. The conclussion of this plot is analogous to the one before, with a fixed depth more targets are being identified by the ChIPseq protocols. Therefore it seems that ChIPexo experiments need to be more deeply sequenced to be able to predict more events.

C. Number of identified targets. With a fixed depth ChIPexo datasets, the number of identified targets is always higher than with both ChIPseq methods. This means, that even with the fact that less candidate regions are being obtained we are identifying more targets with ChIPexo, therefore the FDR may be lower with ChIPexo.

D. Resolution. ChIPexo's resolution is much smaller than ChIPseq resolution's for any given depth.

### Figure 3 - Comparison of signal, balance of read strands, and peak shape betwen ChIPexo and ChIPseq

A. ChIPexo tagcounts vs ChIPseq PET tagcounts. For this plot, the genome was partitioned into bins and the number of fragments that overlap each bin was calculated. Overall for high count regions, it seems that both protocols are linearly related but in the case for low count regions it seems that ChIPexo's background is mucho lower than ChIPseq.

B. Foward strand ratio. When comparing the forward strand ratio between both single end protocols (ChIPexo and ChIPseq), it seems that both densities are centered around $0.5$. However for the ChIPexo case, the tails seem to be heavier, therefore several regions are going to show a degree of strand imbalance which may bias the binding event estimation. However, the density seems to be symmetric, therefore it seems that the $\lambda$ exonuclease enzyme doesn't show preference to diggest one side more than the other. Also, for ChIPseq, the density is tighly concentrated around the mode, which may makes the pair assumption suitable for ChIPseq, which is not the case for ChIPexo.

C. $\delta$ comparison. The $\delta$ parameter of dPeak is being compared between the estimation of binding events for ChIPexo and a ChIPseq SET dataset. This plot shows that the distance to the binding events is higher for the SET case , therefore for datasets where the TF's are more tighly concentrated this may make the model to confuse binding events that are close among each other.

D. $\sigma$ comparison. The $\sigma$ parameter of dPeak's model is being compared among SET methods. ChIPexo's variance seems to be smaller than ChIPseq SET's. This agrees with the previous conclusion and also shows that for ChIPexo reads are less scattered around their binding sites compared to ChIPseq SET's case.

### Figure 4 -  Mappability and GC content biases in CTCF ChIPexo data

For this figure, the genome was binned and for each bins the following quantities were calculated:

- TagCounts. Number of fragments that overlap each bin
- Mappability score. How mappable is the bin by following Kharchenko et al.
- GC content. Proportion of bases that are G or C

A. Mappability score vs tagCounts. It seems to be an increasing trend of the ChIPseq counts as a function of the mappability score.

B. GC content vs tagCounts. The average ChIPcount increases until GC content is $0.6$ and then decreases.

### Figure 5 - Resolution by algorithm.

A. Resolution. The resolution is calculated as the minimun distance between a gold stanrd event and a prediction. Roughly speaking  dpeak improves over GEM and it seems to perform equally good as Apex.

### 2015-10-27 rmarkdown/ChIPexo_QC_figures_list.Rmd
renamed rmakdown folder to rmarkdown and add outline of the paper

---
title: "ChIPexo QC figures"
author: "Rene Welch"
output: html_document
---

## Outline and ideas

The objective of the project was to develop a quality control pipeline for ChIPexo data. Therefore we want to:

1. Why did we decided to make a pipeline specific for ChIPexo, how does the current quality guidelines for ChIPseq behave in ChIPexo. Here we are going to show that those are not enough to asses the quality of ChIPexo data

2. Show how the pipeline works. The first part is to explain what do we want to detect and why, then we need to show how are we detecting those and finally how did we evaluated that we are effectively finding or illustrating those effects

## About ChIPexo data in general

1. Show how does it compare respect to chipseq (PET) and (SET)

2. Show how does the biases of ChIPseq data are behaving in ChIPexo

3. We want to

### 2015-10-27 rmarkdown/ChIPexo_paper.Rmd
renamed rmakdown folder to rmarkdown and add outline of the paper

---
title: "A quality control pipeline for ChIP-exo"
author: "Rene Welch"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
	html document:
		fig_caption: yes
		number_sections: yes
	md document:
		variant: markdown_github
	pdf document:
		fig_caption: yes
		number_sections: yes
		toc: yes
---

# Introduction

## What is ChIPexo ?

<!-- Here I am thinking to write the first section of the poster, which was -->
<!-- based on the Mendenhall and Bernstein note on Genome Biology -->

<!-- this paper need to include the analysis of the paper that dongjun made -->

### Literature review

#### ChIPexo related methods

<!-- About ChIPexo, there are: -->
<!-- CexoR -->
<!-- MACE -->
<!-- GEM (I think this two are the same) -->

#### ChIPseq methods used with ChIPexo

<!-- Some ChIPseq peak callers are used with ChIPexo regardless of they -->
<!-- work or not. Need to check which ones, if they work and if not perhaps -->
<!-- propose -->

### Biases present in ChIPexo data

#### Strand imbalance

#### Sequence digestion bias

### Reproducibility analysis

### Between protocols (ChIPseq vs ChIPexo)
### Between biological replicates

### Discussion

<!-- Here we are supposed to write the analysis that we made, the quality -->
<!-- control pipeline and how our claims were validated -->

### Conclusions

### Bibliography

### 2015-10-27 rmarkdown/ChIPexo_QC_figures_list.Rmd
Update the QC figures list

@@
 Why did we decided to make a pipeline specific for ChIPexo, how does the current quality guidelines for ChIPseq behave in ChIPexo.
-Here we are going to show that those are not enough to asses the quality of ChIPexo data
+Here we are going to show that current guidelines for ChIPseq are not enough for ChIPexo, and that our ideas are useful to asses the quality of this datasets.
+
+    a. Diagram of how the pipeline works, rough explanation of biases and effects that are being measured
+    b.
+Comparison of all datasets under current quality measures and comparison with the measures for the respective chipseq datasets.
+For the case of strand cross correlation it may be necessary to emphasize the absence of input reads for chipexo, which makes measures as RSC and NSC unavailable.
 
@@
 
+    a. Biases that we are detecting with the pipeline
+    b. Determine good vs bad datasets, by comparing respect to a gold standard.
+    c.
+
 ## About ChIPexo data in general
@@
 
+    a. Compare without model, the tagcount scatter plot, forward strand ratio density, etc. This section analysis may be done with the construction of bins
+    b. Compare with model, i.e. resolution, sensitivity and saturation analyses. The last part may link with experimental design issues
+
 2. Show how does the biases of ChIPseq data are behaving in ChIPexo
 
-3. We want to
+    a. This part may be GC and mappability plots. We have the case for CTCF which show that this quantities behave in the same fashion as the chip seq case.
+
+## Algorithmic issues...
+
+I haven't paid to much attention to this, are we going to make comparisons with more datasets or only use the one that we have ?
+
+3. We want to asses experimental design considerations by comparing all protocols against a gold standard set

### 2015-10-27 rscripts/condition_analysis.R

@@ -21,7 +21,7 @@ set_char <- edsn_tab("set")
 ## Initial parameters
 
 tf <- "Sig70"
-rif <- "rif0min"
+rif <- "rif20min"
 bs <- 150
 fl <- 150
 fdr <- .1
@@ -30,7 +30,7 @@ mc <- 16
 maxComp <- 1
 
 flag_bins <- FALSE
-flag_peaks <- FALSE
+flag_peaks <- TRUE
 flag_binding <- TRUE
 flag_input <- FALSE
 

### 2015-10-27 rscripts/saturation_analysis.R

@@ -34,10 +34,10 @@ maxComp <- 5
 seedset <- c( 23456, 34567, 45678, 56789,45987 )
 
 ### we have 5 different seeds
-k <- 2
-flag_sample <- FALSE
-flag_bins <- FALSE
-flag_peaks <- FALSE
+k <- 3
+flag_sample <- TRUE
+flag_bins <- TRUE
+flag_peaks <- TRUE
 flag_binding <- TRUE
 
 ## notes on seeds:

### 2015-10-29 rmarkdown/ChIPexo_QC_figures_list.Rmd
Added scripts to work with the old aerobic exponential ChIPexo Sig70 data

@@
     a. Diagram of how the pipeline works, rough explanation of biases and effects that are being measured
+
+        a1. The diagram may consist of 3 parts: Coverage plot, vector of stats per island, summary measures
+
     b.
@@
 
-I haven't paid to much attention to this, are we going to make comparisons with more datasets or only use the one that we have ?
+1.
+Since we only published this comparison for ChIPseq PET and SET, it is gonna be necessary to make comparisons of ChIPexo.
+We already have the one speaking about resolution, but it may be worth look at different measures.
 

### 2015-10-29 R/base_edsn.R

@@ -6,17 +6,11 @@
 edsn_tab <- function(what){
   stopifnot(what %in% c("exo","pet","set"))
   if(what == "exo"){
-    edsn <- as.character(931:938)
-    ip <- rep(c("Sig70","SigmaS"),4)
-    condition <- rep(c("exp","stat"),each = 4)
-    repl <- rep( rep(1:2,each =2),2)
-    dt1 <- data.table(edsn,ip,condition,repl)
     edsn <- as.character(1310:1321)
     ip <- rep(c("Beta","Sig70","BetaPF"),4)
     condition <- rep( rep(c("rif0min","rif20min"),each = 3),2)
     repl <- rep(rep(1:2,each = 6))
-    dt2 <- data.table(edsn,ip,condition,repl)
-    dt <- rbind(dt1,dt2)    
+    dt <- data.table(edsn,ip,condition,repl)
   }else{
     edsn <- as.character(1396:1403)
     ip <- rep(c("Sig70","BetaPF"),4)
@@ -26,3 +20,37 @@ edsn_tab <- function(what){
   }
   return(dt)
 }
+
+
+edsn_tab_old <- function(what){
+  stopifnot(what %in% c("exo","pet","set"))
+  if(what == "exo"){
+    ## exo
+    edsn <- as.character(c("931","933"))
+    ip <- rep("Sig70",2)
+    condition <- rep("aerobic",2)
+    growth <- rep("exp",2)
+    repl <- 1:2
+    dt <- data.table(edsn,ip,condition,growth,repl)    
+  }else if(what == "pet"){    
+    ## pet
+    edsn <- "788"
+    ip <- "Sig70"
+    condition <- "aerobic"
+    growth <- "exp"
+    repl <- 1
+    dt <- data.table(edsn,ip,condition,growth,repl)    
+  }else{
+    ## set
+    edsn <- "80"
+    ip <- "Sig70"
+    condition <- "aerobic"
+    growth <- "exp"
+    repl <- 1
+    dt <- data.table(edsn,ip,condition,growth,repl)        
+  }
+  return(dt)
+
+}
+
+  

### 2015-10-29 rscripts/condition_analysis_aerobic.R

@@ -1,4 +1,3 @@
-
 rm(list = ls())
 
 library(mosaics)
@@ -12,37 +11,34 @@ library(dpeak)
 
 source("R/base_edsn.R")
 
-exo_char <- edsn_tab("exo")
-pet_char <- edsn_tab("pet")
-set_char <- edsn_tab("set")
+what <- c("exo","pet","set")
+char <- lapply(what,edsn_tab_old)
+names(char) <- what
 
 #############1#########################################################################
 
 ## Initial parameters
 
 tf <- "Sig70"
-rif <- "rif20min"
+cond <- "aerobic"
 bs <- 150
 fl <- 150
-fdr <- .1
+fdr <- .001
 thresh <- 10
 mc <- 16
-maxComp <- 5
+maxComp <- 1
 
 flag_bins <- FALSE
 flag_peaks <- FALSE
 flag_binding <- TRUE
 flag_input <- FALSE
 
-exo <- exo_char[ip == tf & condition == rif ]
-pet <- pet_char[ip == tf & condition == rif ]
-set <- set_char[ip == tf & condition == rif ]
+char <- lapply(char,function(x)x[ip == tf & condition == cond])
 
 check_create <- function(dr)if(!dir.exists(dr))dir.create(dr)
 
 base_dir <- "/p/keles/ChIPexo/volume6/condition"
-folder <- paste(tf,rif,sep = "_")
-what <- c("exo","pet","set")
+folder <- paste(tf,cond,sep = "_")
 
 base_dir <- file.path(base_dir,folder)
 check_create(base_dir)
@@ -53,13 +49,19 @@ check_create(base_dir)
 
 inputs_dir <- "/p/keles/ChIPexo/volume6/inputs"
 inputs_dirs <- lapply(what[-1],function(x)file.path(inputs_dir,x))
-  
+
+chip_dirs <- list()
+chip_dirs[["exo"]] <- "/p/keles/ChIPexo/volume3/LandickData/ChIPexo"
+chip_dirs[["pet"]] <- "/p/keles/ChIPexo/volume3/LandickData/ChIPseq_PET"
+chip_dirs[["set"]] <- "/p/keles/ChIPexo/volume3/LandickData/ChIPseq_SET"
+
+
 if(flag_input){
   lapply(inputs_dirs,check_create)
-  constructBins(file.path(pet_dir,"edsn1369_042814_qc.filter.bam"),
[…]

### 2015-10-29 rscripts/condition_analysis_results.R

@@ -22,13 +22,13 @@ names(char) <- what
 ## Initial parameters
 
 tf <- "Sig70"
-rif <- "rif20min"
+rif <- "aerobic"
 bs <- 150
 fl <- 150
 mc <- detectCores()
 figs_dir <- "figs/condition"
-fdr <- .01
-G <- 1
+fdr <- .001
+G <- 5
 
 char <- lapply(char,function(x)x[ip ==tf & condition == rif])
 
@@ -37,7 +37,10 @@ folder <- paste(tf,rif,sep = "_")
 
 check_create <- function(dr)if(!dir.exists(dr))dir.create(dr)
 
-figs_dir <- file.path(figs_dir,folder,paste0("FDR",fdr*100))
+figs_dir <- file.path(figs_dir,folder)
+check_create(figs_dir)
+
+figs_dir <- file.path(figs_dir,paste0("FDR",fdr*100))
 check_create(figs_dir)
 
 figs_dir <- file.path(figs_dir,paste0("G_",G))
@@ -94,7 +97,6 @@ names(peaks) <- what
 
 load_binding <- function(dir,what,fdr,G,char)
 {
-  browser()
   dir <- file.path(dir,"binding",what,paste0("FDR",fdr*100),paste0("G_",G))
   files <- list.files(dir)
   edsn <- char[[what]][,(edsn)]
@@ -226,7 +228,7 @@ sens <- mapply(sensitivity ,
            results[["set"]],1:2,MoreArgs = list(ext = 20),SIMPLIFY = FALSE)
 sens <- do.call(rbind,sens)
 
-xl <- 500
+xl <- 400
 pdf(file = file.path(figs_dir,"sensitivity.pdf"))
 ggplot(sens[nAnnot > 1 & between(aveDist,0, xl,incbounds = FALSE) ],
        aes(aveDist,frac,colour = seq,shape = seq))+geom_point()+

### 2015-10-29 rscripts/saturation_analysis_aerobic.R

@@ -30,14 +30,14 @@ fl <- 150
 fdr <- .001
 thresh <- 10
 mc <- 16
-maxComp <- 5
+maxComp <- 3
 seedset <- c( 23456, 34567, 45678, 56789,45987 )
 
 ### we have 5 different seeds
 k <- 3
 flag_sample <- FALSE
 flag_bins <- FALSE
-flag_peaks <- TRUE
+flag_peaks <- FALSE
 flag_binding <- TRUE
 
 ## notes on seeds:
@@ -213,19 +213,16 @@ mosaics_peaks_wrap <- function(in_dir,out_dir,what,extra,fdr,thresh,bs,mc)
   }else{
     opt <- c("chip","input")
   }
-  browser()
 
   files <- file.path(in_dir,list.files(in_dir))
-  peaks <- mclapply(files,call_peaks,opt,extra,fdr,bs,thresh,
-    mc.cores = mc,mc.preschedule = FALSE)
+  peaks <- mclapply(files,call_peaks,opt,extra,fdr,bs,thresh,mc.cores = mc)
 
   files <- list.files(in_dir)
   files <- sapply(strsplit(files,".",fixed = TRUE),function(x)x[1])
   files <- file.path(out_dir,paste0(files,"_peaks.txt"))
 
   out <- mcmapply(write.table,peaks,files,MoreArgs = list(
-    sep = "\t",quote = FALSE,row.names = FALSE,col.names = FALSE),
-    SIMPLIFY = FALSE,mc.cores = mc,mc.preschedule =FALSE)  
+    sep = "\t",quote = FALSE,row.names = FALSE,col.names = FALSE),SIMPLIFY = FALSE,mc.cores = mc)  
 
   return(out)
 

### 2015-10-29 rscripts/saturation_analysis_results.R

@@ -22,14 +22,14 @@ names(char) <- what
 ## Initial parameters
 
 tf <- "Sig70"
-rif <- "rif20min"
+rif <- "rif0min"
 bs <- 150
 fl <- 150
 mc <- 8
 figs_dir <- "figs/saturation"
 fdr <- .001
 G <- 5
-k <- 1
+k <- 3
 ext <- 20
 
 char <- lapply(char,function(x)x[ip ==tf & condition == rif])
@@ -197,12 +197,12 @@ nr_peaks_by_sample <- function(results)
 nr_cand_region <- nr_peaks_by_sample(results)
 
 pdf(file = file.path(figs_dir,"Nr_candidate_regions.pdf"),width = 6,height = 5)
-ggplot(nr_cand_region[,mean(nr_peaks),by = .(size,seq)],aes(size , V1,colour = seq))+
-  geom_line(size = 1)+theme(legend.position = "top")+ylim(140,400)+
+ggplot(nr_cand_region[!is.na(edsn),mean(nr_peaks),by = .(size,seq)],aes(size , V1,colour = seq))+
+  geom_line(size = 1)+theme(legend.position = "top")+
   scale_color_brewer(name = "",palette = "Set1")+ylab("Number of candidate regions")+xlab("Number of reads")
-ggplot(nr_cand_region,aes(size , nr_peaks,colour = seq,linetype = as.factor(repl)))+geom_line(size = 1)+
+ggplot(nr_cand_region[!is.na(edsn)],aes(size , nr_peaks,colour = seq,linetype = as.factor(repl)))+geom_line(size = 1)+
   scale_color_brewer(name = "",palette = "Set1",guide = "none")+scale_linetype_manual(values = c(1,2),name = "rep")+
-  ylab("Number of candidate regions")+xlab("Number of reads")+theme(legend.position = "top")+ylim(140,400)
+  ylab("Number of candidate regions")+xlab("Number of reads")+theme(legend.position = "top")
 dev.off()
 
 ######################################################################################
@@ -231,12 +231,12 @@ nr_predicted_events_by_sample <- function(results)
 nr_pred_events <- nr_predicted_events_by_sample(results)
 
 pdf(file = file.path(figs_dir,"Nr_predicted_events.pdf"),width = 6,height = 5)
-ggplot(nr_pred_events[,mean(nr_events),by = .(size,seq)],aes(size , V1,colour = seq))+
-  geom_line(size = 1)+theme(legend.position = "top")+ylim(250,1e3)+
[…]

### 2015-11-02 rmarkdown/chip_exo_paper.Rnw
added paper start

%\VignetteEngine{knitr::knitr}
\documentclass{article}

\title{High Resolution Identification of Protein-DNA Binding Events
  using ChIP-exo}

\begin{document}

\maketitle

\end{document}

### 2015-11-02 rmarkdown/chip_exo_paper.tex
added abstract to paper

%\VignetteEngine{knitr::knitr}
\documentclass{article}\usepackage[]{graphicx}\usepackage[]{color}
%% maxwidth is the original width if it is less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%

\usepackage{framed}
\makeatletter
\newenvironment{kframe}{%
 \def\at@end@of@kframe{}%
 \ifinner\ifhmode%
  \def\at@end@of@kframe{\end{minipage}}%
  \begin{minipage}{\columnwidth}%
 \fi\fi%
 \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
 \colorbox{shadecolor}{##1}\hskip-\fboxsep
     % There is no \\@totalrightmargin, so:
     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
 \MakeFramed {\advance\hsize-\width
   \@totalleftmargin\z@ \linewidth\hsize
   \@setminipage}}%
 {\par\unskip\endMakeFramed%
 \at@end@of@kframe}
\makeatother

\definecolor{shadecolor}{rgb}{.97, .97, .97}
\definecolor{messagecolor}{rgb}{0, 0, 0}
\definecolor{warningcolor}{rgb}{1, 0, 1}
\definecolor{errorcolor}{rgb}{1, 0, 0}
\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX

\usepackage{alltt}

\usepackage{xcolor}
\usepackage{url}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows}
\usepackage{float}
\usepackage{verbatim}

\usepackage{authblk}

\title{High Resolution Identification of Protein-DNA Binding Events
  using ChIP-exo}
\date{}
\author[6]{Dongjun Chung}
\author[1]{Ren\'e Welch}
\author[3]{Irene Ong}
\author[3,4]{Jeffrey Grass}
\author[3,4,5]{Robert Landick}
\author[1,2]{S\"und\"uz Kele\c{s}}

\affil[1]{Department of Statistics, University of Wisconsin - Madison}
\affil[2]{Department of Biostatistics and Medical Informatics, University of Wisconsin - Madison}
\affil[3]{Great Lakes Bioenergy Research Center}
\affil[4]{Department of Biochemistry, University of Wisconsin - Madison}
\affil[5]{Department of Bacteriology, University of Wisconsin - Madison}
\affil[6]{Department of Public Health Sciences, Medical University of South Carolina}

\addtolength{\oddsidemargin}{-.875in}
\addtolength{\evensidemargin}{-.875in}
\addtolength{\textwidth}{1.75in}

\addtolength{\topmargin}{-.875in}
\addtolength{\textheight}{1.75in}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}

\maketitle

\begin{abstract}

  Recently, ChIP-exo has been developed to investigate protein-DNA
  interaction in higher resolution compared to popularly used
  ChIP-Seq. Although ChIP-exo has drawn much attention and is
  considered as powerful assay, currently, no systematic studies have
  yet been conducted to determine optimal strategies for experimental
  design and analysis of ChIP-exo. In order to address these
  questions, we evaluated diverse aspects of ChIP-exo and found the
  following characteristics of ChIP-exo data. First, Background of
  ChIP-exo data is qute different from that of ChIP-Seq data. However,
  sequence biases inherently present in ChIP-Seq data still exist in
  ChIP-exo data. Second, in ChIP-exo data, reads are located around
  binding sites much more tightly and hence, it has potential for high
  resolution identification of protein-DNA interaction sites, hence
  the space to allocate the reads is greatly reduced. Third, although
  often assumed in the ChIP-exo data analysis methods, the peak pair
  assumption does not hold well in real ChIP-exo data. Fourth, spatial
  resolution of ChIP-exo is comparable to that of PET ChIP-Seq and
  both of them are significantly better than resolution of SET
  ChIP-Seq. Finally, for given fixed sequencing depth, ChIP-exo
  provides higher sensitivity, specificity, and spatial resolution
  than PET ChIP-seq.

  In this article, we provide a quality control pipeline which
  visually asses ChIP-exo biases and calculates a signal-to-noise
  measure. Also, we updated dPeak, which makes a striking
  balance in sensitivity, specificity, and spatial resolution for
  ChIP-exo data analysis.

\end{abstract}

\newpage

\section{Introduction}
\label{sec:intro}

ChIP-exo (Chromatin Immunopecipitation followed by exonuclease
digestion and next generation sequencing) Rhee and Pugh (\cite{exo1})
is the state-of-the-art experiment developed to attain single
base-pair resolution of protein binding site identification and it is
considered as a powerful alternative to popularly used ChIP-seq
(Chromatin Immunoprecipitation coupled with next generation sequencing
) assay. ChIP-exo experiments first capture millions of DNA fragments
(150 - 250 bp in length) that the protein under study interacts with
using random fragmentation of DNA and a protein-specific
antibody. Then, exonuclease is introducted to trim 5' end of each DNA
fragment to a fixed distance from the bound protein. As a result,
boundaries around the protein of interest constructed with 5' ends of
fragments are located much closed to bound protein compared to
ChIP-seq. This is the step unique to ChIP-exo that could potentially
provide significantly higher spatial resolution compared to
ChIP-seq. Finally, high throughput sequencing of a small region (25 to
100 bp) at 5' end of each fragment generates millions of reads or
tags.

While the number of produced ChIP-exo data keeps increasing,
characteristicos of ChIP-exo data and optimal strategies for
experimental design and analysis of ChIP-exo data are not fully
investigated yet, including issues of sequence biases inherent to
ChIP-exo data, choice of optimal statistical methods, and
determination of optimal sequencing deoth. However, currently, the
number of available ChIP-exo data is still limited and their
sequencing depths are still insufficient for such investigation.

\section{Results}
\label{sec:results}

\section{Discussion}
\label{sec:disc}

\section*{Methods}
\label{sec:methods}

\bibliography{chip_exo_paper}
\bibliographystyle{plain}

\nocite{exo_gb}

\end{document}

### 2015-11-03 rmarkdown/chip_exo_paper.Rnw
added genome biology journal format

@@
 %\VignetteEngine{knitr::knitr}
-\documentclass{article}
+\documentclass{bmcart}
+
+\usepackage{xcolor}
+\usepackage{url}
+\usepackage{amsmath}
+\usepackage{amsthm}
+\usepackage{amssymb}
+\usepackage{graphicx}
+\usepackage{tikz}
+\usetikzlibrary{shapes,arrows}
+\usepackage{float}
+\usepackage{verbatim}
+
+% \usepackage{authblk}
+\usepackage[utf8]{inputenc} %unicode support
+
+\begin{document}
+
+%\maketitle
+
+\begin{frontmatter}
+
+\begin{fmbox}
+\dochead{Draft}
 
@@
 
-\begin{document}
+\author[
+   addressref={aff6},                   % id's of addresses, e.g. {aff1,aff2}
+   email={chungd@musc.edu}   % email address
+]{\inits{DC}\fnm{Dongjun} \snm{Chung}}
+\author[
+   addressref={aff1},
+   email={welch@stat.wisc.edu}
+]{\inits{RW}\fnm{Rene} \snm{Welch}}
+\author[
+   addressref={aff3},
+   email={ong@cs.wisc.edu}
+]{\inits{IO}\fnm{Irene} \snm{Ong}}
+\author[
+   addressref={aff3,aff4},
+   email={jagrass@wisc.edu}
+]{\inits{JG}\fnm{Jeffrey} \snm{Grass}}
+\author[
+   addressref={aff3,aff4,aff5},
+   email={landick@bact.wisc.edu}
+]{\inits{RL}\fnm{Robert} \snm{Landick}}
+\author[
+   addressref={aff1,aff2},
+   corref={aff1},
+   email={keles@stat.wisc.edu}
+]{\inits{SK}\fnm{S\"und\"uz} \snm{Kele\c{s}}}
 
-\maketitle
+\address[id=aff6]{
+  \orgname{Department of Public Health Sciences, Medical University of South Carolina},
+  \street{135 Cannon Street},
+  \city{Charleston},
+  \cny{SC}
+}
+\address[id=aff1]{%                           % unique id
+  \orgname{Department of Statistics, University of Wisconsin Madison}, % university, etc
+  \street{1300 University Avenue},                     %
+  %\postcode{}                                % post or zip code
+  \city{Madison},                              % city
+  \cny{WI}                                    % country
+}
+
+\address[id=aff2]{%
+  \orgname{Department of Biostatistics and Medical Informatics, University of Wisconsin Madison},
+  \street{600 Highland Avenue},
+%  \postcode{24105}
+  \city{Madison},
+  \cny{WI}
+}
+\address[id=aff3]{
+  \orgname{Great Lakes Bioenergy Research Center, University of Wisconsin Madison},
+  \street{1552 University Avenue},
+  \city{Madison},
+  \cny{WI}
+}
+\address[id=aff4]{
+  \orgname{Department of Biochemistry, University of Wisconsin Madison},
+  \street{433 Babcock Drive},
+  \city{Madison},
+  \cny{WI}
+}
+\address[id=aff5]{
+  \orgname{Department of Bacteriology, University of Wisconsin Madison},
+  \street{1550 Linden Drive},
+  \city{Madison},
+  \cny{WI}
+}
+
+% \begin{artnotes}
+% %\note{Sample of title note}     % note to the article
+% \note[id=n1]{Equal contributor} % note, connected to author
+% \end{artnotes}
+
+\end{fmbox}
+
+\begin{abstractbox}
+
+  \begin{abstract}
+
+  Recently, ChIP-exo has been developed to investigate protein-DNA
+  interaction in higher resolution compared to popularly used
+  ChIP-Seq. Although ChIP-exo has drawn much attention and is
+  considered as powerful assay, currently, no systematic studies have
+  yet been conducted to determine optimal strategies for experimental
+  design and analysis of ChIP-exo. In order to address these
+  questions, we evaluated diverse aspects of ChIP-exo and found the
+  following characteristics of ChIP-exo data. First, Background of
+  ChIP-exo data is qute different from that of ChIP-Seq data. However,
+  sequence biases inherently present in ChIP-Seq data still exist in
+  ChIP-exo data. Second, in ChIP-exo data, reads are located around
+  binding sites much more tightly and hence, it has potential for high
+  resolution identification of protein-DNA interaction sites, hence
+  the space to allocate the reads is greatly reduced. Third, although
+  often assumed in the ChIP-exo data analysis methods, the peak pair
+  assumption does not hold well in real ChIP-exo data. Fourth, spatial
+  resolution of ChIP-exo is comparable to that of PET ChIP-Seq and
+  both of them are significantly better than resolution of SET
+  ChIP-Seq. Finally, for given fixed sequencing depth, ChIP-exo
+  provides higher sensitivity, specificity, and spatial resolution
+  than PET ChIP-Seq.
+
+  In this article, we provide a quality control pipeline which
+  visually asses ChIP-exo biases and calculates a signal-to-noise
+  measure. Also, we updated dPeak, which makes a striking
+  balance in sensitivity, specificity, and spatial resolution for
+  ChIP-exo data analysis.
+
+\end{abstract}
+\begin{keyword}
+  \kwd{ChIP-exo}
+  \kwd{Quality control}
+\end{keyword}
+
+\end{abstractbox}
+
+\end{frontmatter}
+
+\newpage
+
+* to remove later
+
+\tableofcontents
+
+\newpage
+
+\section{Background}
+\label{sec:intro}
+
+ChIP-exo (Chromatin Immunopecipitation followed by exonuclease
+digestion and next generation sequencing) Rhee and Pugh (\cite{exo1})
+is the state-of-the-art experiment developed to attain single
+base-pair resolution of protein binding site identification and it is
+considered as a powerful alternative to popularly used ChIP-Seq
+(Chromatin Immunoprecipitation coupled with next generation sequencing
+) assay. ChIP-exo experiments first capture millions of DNA fragments
+(150 - 250 bp in length) that the protein under study interacts with
+using random fragmentation of DNA and a protein-specific
+antibody. Then, exonuclease is introducted to trim 5' end of each DNA
+fragment to a fixed distance from the bound protein. As a result,
+boundaries around the protein of interest constructed with 5' ends of
+fragments are located much closed to bound protein compared to
+ChIP-Seq. This is the step unique to ChIP-exo that could potentially
+provide significantly higher spatial resolution compared to
+ChIP-Seq. Finally, high throughput sequencing of a small region (25 to
+100 bp) at 5' end of each fragment generates millions of reads or
+tags.
+
+While the number of produced ChIP-exo data keeps increasing,
+characteristicos of ChIP-exo data and optimal strategies for
+experimental design and analysis of ChIP-exo data are not fully
+investigated yet, including issues of sequence biases inherent to
+ChIP-exo data, choice of optimal statistical methods, and
+determination of optimal sequencing deoth. However, currently, the
+number of available ChIP-exo data is still limited and their
+sequencing depths are still insufficient for such investigation.
+
+\section{Results and discussion}
+\label{sec:results}
+
+\section*{Methods}
+\label{sec:methods}
+
+\bibliographystyle{bmc-mathphys} % Style BST file (bmc-mathphys, vancouver, spbasic).
+\bibliography{chip_exo_paper}
+
+\nocite{exo_gb}
 

### 2015-11-03 rmarkdown/chip_exo_paper.tex
added genome biology journal format

@@
 %\VignetteEngine{knitr::knitr}
-\documentclass{article}\usepackage[]{graphicx}\usepackage[]{color}
+\documentclass{bmcart}\usepackage[]{graphicx}\usepackage[]{color}
 %% maxwidth is the original width if it is less than linewidth
@@
 
-\usepackage{authblk}
+% \usepackage{authblk}
+\usepackage[utf8]{inputenc} %unicode support
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+\begin{document}
+
+%\maketitle
+
+\begin{frontmatter}
+
+\begin{fmbox}
+\dochead{Draft}
 
@@
   using ChIP-exo}
-\date{}
-\author[6]{Dongjun Chung}
-\author[1]{Ren\'e Welch}
-\author[3]{Irene Ong}
-\author[3,4]{Jeffrey Grass}
-\author[3,4,5]{Robert Landick}
-\author[1,2]{S\"und\"uz Kele\c{s}}
 
-\affil[1]{Department of Statistics, University of Wisconsin - Madison}
-\affil[2]{Department of Biostatistics and Medical Informatics, University of Wisconsin - Madison}
-\affil[3]{Great Lakes Bioenergy Research Center}
-\affil[4]{Department of Biochemistry, University of Wisconsin - Madison}
-\affil[5]{Department of Bacteriology, University of Wisconsin - Madison}
-\affil[6]{Department of Public Health Sciences, Medical University of South Carolina}
+\author[
+   addressref={aff6},                   % id's of addresses, e.g. {aff1,aff2}
+   email={chungd@musc.edu}   % email address
+]{\inits{DC}\fnm{Dongjun} \snm{Chung}}
+\author[
+   addressref={aff1},
+   email={welch@stat.wisc.edu}
+]{\inits{RW}\fnm{Rene} \snm{Welch}}
+\author[
+   addressref={aff3},
+   email={ong@cs.wisc.edu}
+]{\inits{IO}\fnm{Irene} \snm{Ong}}
+\author[
+   addressref={aff3,aff4},
+   email={jagrass@wisc.edu}
+]{\inits{JG}\fnm{Jeffrey} \snm{Grass}}
+\author[
+   addressref={aff3,aff4,aff5},
+   email={landick@bact.wisc.edu}
+]{\inits{RL}\fnm{Robert} \snm{Landick}}
+\author[
+   addressref={aff1,aff2},
+   corref={aff1},
+   email={keles@stat.wisc.edu}
+]{\inits{SK}\fnm{S\"und\"uz} \snm{Kele\c{s}}}
 
-\addtolength{\oddsidemargin}{-.875in}
-\addtolength{\evensidemargin}{-.875in}
-\addtolength{\textwidth}{1.75in}
+\address[id=aff6]{
+  \orgname{Department of Public Health Sciences, Medical University of South Carolina},
+  \street{135 Cannon Street},
+  \city{Charleston},
+  \cny{SC}
+}
+\address[id=aff1]{%                           % unique id
+  \orgname{Department of Statistics, University of Wisconsin Madison}, % university, etc
+  \street{1300 University Avenue},                     %
+  %\postcode{}                                % post or zip code
+  \city{Madison},                              % city
+  \cny{WI}                                    % country
+}
 
-\addtolength{\topmargin}{-.875in}
-\addtolength{\textheight}{1.75in}
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-\begin{document}
+\address[id=aff2]{%
+  \orgname{Department of Biostatistics and Medical Informatics, University of Wisconsin Madison},
+  \street{600 Highland Avenue},
+%  \postcode{24105}
+  \city{Madison},
+  \cny{WI}
+}
+\address[id=aff3]{
+  \orgname{Great Lakes Bioenergy Research Center, University of Wisconsin Madison},
+  \street{1552 University Avenue},
+  \city{Madison},
+  \cny{WI}
+}
+\address[id=aff4]{
+  \orgname{Department of Biochemistry, University of Wisconsin Madison},
+  \street{433 Babcock Drive},
+  \city{Madison},
+  \cny{WI}
+}
+\address[id=aff5]{
+  \orgname{Department of Bacteriology, University of Wisconsin Madison},
+  \street{1550 Linden Drive},
+  \city{Madison},
+  \cny{WI}
+}
 
-\maketitle
+% \begin{artnotes}
+% %\note{Sample of title note}     % note to the article
+% \note[id=n1]{Equal contributor} % note, connected to author
+% \end{artnotes}
+
+\end{fmbox}
+
+\begin{abstractbox}
 
@@
   provides higher sensitivity, specificity, and spatial resolution
-  than PET ChIP-seq.
+  than PET ChIP-Seq.
 
@@
 \end{abstract}
+\begin{keyword}
+  \kwd{ChIP-exo}
+  \kwd{Quality control}
+\end{keyword}
+
+\end{abstractbox}
+
+\end{frontmatter}
 
@@
 
-\section{Introduction}
+* to remove later
+
+\tableofcontents
+
+\newpage
+
+\section{Background}
 \label{sec:intro}
@@
 base-pair resolution of protein binding site identification and it is
-considered as a powerful alternative to popularly used ChIP-seq
+considered as a powerful alternative to popularly used ChIP-Seq
 (Chromatin Immunoprecipitation coupled with next generation sequencing
@@
 fragments are located much closed to bound protein compared to
-ChIP-seq. This is the step unique to ChIP-exo that could potentially
+ChIP-Seq. This is the step unique to ChIP-exo that could potentially
 provide significantly higher spatial resolution compared to
-ChIP-seq. Finally, high throughput sequencing of a small region (25 to
+ChIP-Seq. Finally, high throughput sequencing of a small region (25 to
 100 bp) at 5' end of each fragment generates millions of reads or
@@
 
-\section{Results}
+\section{Results and discussion}
 \label{sec:results}
-
-\section{Discussion}
-\label{sec:disc}
 
@@
 
+\bibliographystyle{bmc-mathphys} % Style BST file (bmc-mathphys, vancouver, spbasic).
 \bibliography{chip_exo_paper}
-\bibliographystyle{plain}
 

### 2015-11-03 rmarkdown/chip_exo_paper.Rnw
added new bibliography and a couple paragraphs on background section

@@
 
-\title{High Resolution Identification of Protein-DNA Binding Events
-  using ChIP-exo}
+\title{ChIP-exo: High Resolution Identification of Protein-DNA Binding
+  Events and Quality Control}
 
@@
 
-\address[id=aff6]{
-  \orgname{Department of Public Health Sciences, Medical University of South Carolina},
-  \street{135 Cannon Street},
-  \city{Charleston},
-  \cny{SC}
-}
 \address[id=aff1]{%                           % unique id
@@
   \cny{WI}
+}
+\address[id=aff6]{
+  \orgname{Department of Public Health Sciences, Medical University of South Carolina},
+  \street{135 Cannon Street},
+  \city{Charleston},
+  \cny{SC}
 }
@@
   visually asses ChIP-exo biases and calculates a signal-to-noise
-  measure. Also, we updated dPeak, which makes a striking
+  measure. Also, we updated dPeak \cite{dpeak}, which makes a striking
   balance in sensitivity, specificity, and spatial resolution for
@@
   \kwd{ChIP-exo}
-  \kwd{Quality control}
+  \kwd{QC}
+  % \kwd{Transcription Factor}
+  % \kwd{Binding Event}
 \end{keyword}
@@
 number of available ChIP-exo data is still limited and their
-sequencing depths are still insufficient for such investigation.
+sequencing depths are still insufficient for such investigation. To
+address this limitation we gathered ChIP exo data from diverse
+organisms: CTCF factor in human \cite{exo1}; ER factor in human and
+FoxA1 factor in mouse from \cite{exo_illumnina}; and generated
+$\sigma^{70}$ factor in Escherichia coli (E. Coli) under aerobic ($ +
+O_2$) condition, and treated by rifampicin by 0 and 20 minutes.
 
@@
 \label{sec:results}
+
+\subsection{Characteristics of ChIP exo data}
+\label{sec:charac}
+
+\subsection{ChIP exo Quality Control pipeline}
+\label{sec:QC}
 

### 2015-11-03 rmarkdown/chip_exo_paper.Rnw
run several cases of condition and saturation analyiss and updated background part of paper

@@
 organisms: CTCF factor in human \cite{exo1}; ER factor in human and
-FoxA1 factor in mouse from \cite{exo_illumnina}; and generated
+FoxA1 factor in mouse from \cite{exoillumina}; and generated
 $\sigma^{70}$ factor in Escherichia coli (E. Coli) under aerobic ($ +
 O_2$) condition, and treated by rifampicin by 0 and 20 minutes.
+
+In order to archieve potential benefits of ChIP exo on protein binding
+site identification, it is critical to understand which are the
+important characteristics of ChIP exo data and to use algorithms that
+could fully utilize information available in ChIP exo data. Rhee and
+Pugh \cite{exo1} discussed that reads in the forward and reverse
+strand might construct peak apirs around bound protein, of which
+heights were implicitly assumed to be symmetric. Hence, they used the
+``peak pair method'' taht preducts the midpoint of two modes of pak
+pairs as potential binding site. Apex, a recetly developed ChIP-exo
+data analysis method, is also based on this peak pair
+assumption. However, appropriatness of such assumption was not fully
+evaluated in the literature yet. Furthermore, it is still unknown
+which factors could affect protein binding site ientification using
+ChIP exo data. In order to address this problem, we investigated
+various aspects of ChIP exo data by contrasting them with their
+respective ChIP Seq experiments.
 

### 2015-11-03 rscripts/condition_analysis_aerobic.R

@@ -23,13 +23,13 @@ tf <- "Sig70"
 cond <- "aerobic"
 bs <- 150
 fl <- 150
-fdr <- .001
+fdr <- .01
 thresh <- 10
-mc <- 16
+mc <- 24
 maxComp <- 1
 
 flag_bins <- FALSE
-flag_peaks <- FALSE
+flag_peaks <- TRUE
 flag_binding <- TRUE
 flag_input <- FALSE
 

### 2015-11-03 rscripts/saturation_analysis_aerobic.R

@@ -37,10 +37,10 @@ maxComp <- 5
 seedset <- c( 23456, 34567, 45678, 56789,45987 ,521324)
 
 ### we have 5 different seeds
-k <- 1
-flag_sample <- FALSE
-flag_bins <- FALSE
-flag_peaks <- FALSE
+k <- 5
+flag_sample <- TRUE
+flag_bins <- TRUE
+flag_peaks <- TRUE
 flag_binding <- TRUE
 
 ## notes on seeds:

### 2015-11-05 rmarkdown/chip_exo_paper.Rnw
added more references to bibliography and finished draft of background section

@@
   \kwd{QC}
-  % \kwd{Transcription Factor}
-  % \kwd{Binding Event}
+  \kwd{TFBS}
+  \kwd{BS identification on High-Res}
 \end{keyword}
@@
 
-In order to archieve potential benefits of ChIP exo on protein binding
-site identification, it is critical to understand which are the
-important characteristics of ChIP exo data and to use algorithms that
-could fully utilize information available in ChIP exo data. Rhee and
-Pugh \cite{exo1} discussed that reads in the forward and reverse
-strand might construct peak apirs around bound protein, of which
+DNA libraries generated by the ChIP-exo protocol seem to be less
+complex than the libraries generated by ChIP-Seq
+\cite{exo_review}. Hence, most of current guidelines \cite{encode_qc}
+may not be aplicable on ChIP-exo. To address this challenge, we
+suggest a series of quality control visualizations to understand which
+biases are present in ChIP-exo data. Related to quality control,
+Previous ChIP-exo analysis used ChIP-Seq samples to compare the
+resolution between experiments (\cite{exo1}, \cite{exo2},
+\cite{exoillumina}). In \cite{carroll.qc}, Carroll et. al. studied the
+use of the Strand-Cross Correlation (SCC) \cite{strandcc}. and showed
+that by filtering blacklisted regions the estimation of the SCC is
+improved. However, this method requires to know blacklisted regions in
+advance which may not be available. In our pipeline we propose two
+out-the-shelf metrics equivalent to RSC and NSC for the estimation of
+the signal-to-noise ratio for a ChIP-exo sample.
+
+In order to archieve the potential benefits of ChIP-exo on protein
+binding site identification, it is critical to understand which are
+the important characteristics of ChIP exo data and to use algorithms
+that could fully utilize information available in ChIP exo data. Rhee
+and Pugh \cite{exo1} discussed that reads in the forward and reverse
+strand might construct peak pairs around bound protein, of which
 heights were implicitly assumed to be symmetric. Hence, they used the
-``peak pair method'' taht preducts the midpoint of two modes of pak
-pairs as potential binding site. Apex, a recetly developed ChIP-exo
-data analysis method, is also based on this peak pair
+``peak pair method'' that predicts the midpoint of two modes of peak
+pairs as potential binding site. Mace \cite{mace}, CexoR \cite{cexor}
+and peakzilla \cite{peakzilla}, recently developed ChIP-exo data
+analysis methods, are also based on this peak pair
 assumption. However, appropriatness of such assumption was not fully
@@
 
+Currently, research on statistical methods for ChIP-exo data is still
+in its very early stage. Although many methods have been proposed to
+identify protein binding sites from ChIP Seq data (reviewed in
+\cite{evaluation} and \cite{computation}), such as MACS \cite{macs},
+CisGenome \cite{cisgenome} and MOSAiCS \cite{mosaics}, these
+approaches reveal protein binding sites in lower resolution, i.e., at
+an interval of hundreds to thousands of base pairs. Furthermore, they
+report only one ``mode'' or ``predicted binding location'' per
+peak. Hence, these methods are not appropiate to evaluate the
+potential of ChIP-exo data for high resolution identification of
+protein binding sites. More recently, deconvolution algorithms such as
+CSDeconv \cite{csdeconv}, GEM \cite{gem} (an improved version of
+\cite{gps}) and PICS \cite{pics} have been proposed to identify
+binding sites in higher resolution using ChIP-Seq data. However, most
+of them are still not tailored for ChIP-exo and PET and SET ChIP-Seq
+data in a unified framework and as a result, currently available
+methods are not appropiate for fair comparison between ChIP-exo and
+ChIP-Seq. To address these limitations, we developed an improved of
+dPeak \cite{dpeak}, a high resolution binding site identification
+(deconvolution) algorithm that we previously developed for PET and SET
+ChIP-Seq data, so that it can also handle ChIP-exo data. The dPeak
+algorithm implements a probabilistic model that accurately describes
+the ChIP-exo and ChIP-Seq data generation process.
+
+In this paper, we demonstrate that the peak pair assumption of Rhee
+and Pugh \cite{exo1} does not hold well in real ChIP-exo
+data. Furthermore, we found that when we analyze ChIP-exo data from
+eukaryotic genomes, it is important to consider sequence biases
+inherent to ChIP-exo data, such as mappability and GC content in order
+to improve sensitivity and specificity of binding site
+identification. dPeak outperforms or performs competitively with
+ChIP-exo data analysis such as GEM and MACE. More importantly, when
+comparable number of reads is used for both ChIP-exo and ChIP-Seq,
+dPeak couple with ChIP-exo data provides resolution comparable to PET
+ChIP-Seq and both significantly improve the resolution of protein
+binding site identification compared to SET-based analysis with any of
+the available methods.
+
 \section{Results and discussion}
 \label{sec:results}
+
+\subsection{Data considered}
+\label{sec:alldata}
+
+\subsubsection{Deeply sequenced E.Coli $\sigma^{70}$ ChIP-exo and ChIP-Seq data}
+\label{sec:ourdata}
 

### 2015-11-05 rmarkdown/chip_exo_paper.Rnw
aaded more references and started results part

@@
 
-\subsection{Data considered}
-\label{sec:alldata}
+\subsection{Deeply sequenced E.Coli $\sigma^{70}$ ChIP-exo and ChIP-Seq data}
+\label{sec:ourdata}
 
-\subsubsection{Deeply sequenced E.Coli $\sigma^{70}$ ChIP-exo and ChIP-Seq data}
-\label{sec:ourdata}
+$\sigma^{70}$ factor is a transcription initiation factor of
+housekeeping genesin E. coli. In E. coli genomes, many promoters
+contain multiple transcription start sites (TSS) and these TSS are
+often closely space (10 ~ 150 bp). These closely spaced binding sites
+are considered to be multiple ``switches'' that differentially
+regulate gene expression under diverse growth conditions
+\cite{regdb_old}, \cite{regulondb}. Therefore, investigation of
+ChIP-exo's potential for identification and differentiation of closely
+spaced binding sites are invaluable for elucidating the
+transcriptional networks of prokaryotic genomes.
+
+\subsection{Current guidelines and quality metrics}
+\label{sec:qcnow}
 

### 2015-11-09 rscripts/sig70_saturation_all.R
Finished saturation analysis for Sig70 aerobic sample

@@ -1,0 +1,219 @@
+
+rm(list = ls())
+
+library(data.table)
+library(GenomicAlignments)
+library(ggplot2)
+library(grid)
+library(gridExtra)
+library(RColorBrewer)
+library(scales)
+
+######################################################################################
+
+## Condition table
+
+source("R/base_edsn.R")
+
+what <- c("exo","pet","set")
+char <- lapply(what,edsn_tab)
+names(char) <- what
+
+######################################################################################
+
+## Initial parameters
+
+tf <- "Sig70"
+rif <- "aerobic"
+mc <- 8
+figs_dir <- "figs/saturation"
+ext <- 20
+
+char <- lapply(char,function(x)x[ip ==tf & condition == rif])
+
+base_dir <- "/p/keles/ChIPexo/volume6/saturation"
+folder <- paste(tf,rif,sep = "_")
+
+check_create <- function(dr)if(!dir.exists(dr))dir.create(dr)
+
+base_dir <- file.path(base_dir,folder)
+
+seeds <- list.files(base_dir)
+
+######################################################################################
+
+## get promoter locations table
+
+prom_dir <- "/p/keles/ChIPexo/volume6/gold_standard/Landick"
+pfiles <- list.files(prom_dir)
+pfiles <- pfiles[grep(".bed",pfiles)]
+promoters <- lapply(file.path(prom_dir,pfiles),read.table)
+promoters <- lapply(promoters,data.table)
+
+promoters[[1]][,str := "R"]
+promoters[[2]][,str := "F"]
+
+sites <- do.call(rbind,promoters)
+sites[, id := 1:nrow(sites)]
+setnames(sites,names(sites),c("seqnames","start","end","strand","id"))
+
+sites[,seqnames := "U00096"]
+sites[,strand := ifelse(strand == "F","+","-")]
+
+######################################################################################
+
+load_all <- function(dir,sites)
+{
+  all_files <- list.files(dir,recursive = TRUE)
+
+  peak_files <- all_files[grep("peaks",all_files)]
+  peaks <- lapply(file.path(dir,peak_files),read.table)
[…]

### 2015-11-09 rmarkdown/chip_exo_paper.Rnw
Finished saturation analysis for Sig70 aerobic sample
Fixed typos in bibliography, updated figures to include the number of element in the whole paper figure, added a segment about recomendations for design in article

@@
 
-\title{ChIP-exo: High Resolution Identification of Protein-DNA Binding
-  Events and Quality Control}
+\title{High Resolution Identification of Protein-DNA Binding
+  Events and Quality Control for ChIP-exo data}
 
@@
 sequencing depths are still insufficient for such investigation. To
-address this limitation we gathered ChIP exo data from diverse
+address this limitation we gathered ChIP-exo data from diverse
 organisms: CTCF factor in human \cite{exo1}; ER factor in human and
@@
 binding site identification, it is critical to understand which are
-the important characteristics of ChIP exo data and to use algorithms
-that could fully utilize information available in ChIP exo data. Rhee
+the important characteristics of ChIP-exo data and to use algorithms
+that could fully utilize information available in ChIP-exo data. Rhee
 and Pugh \cite{exo1} discussed that reads in the forward and reverse
@@
 which factors could affect protein binding site ientification using
-ChIP exo data. In order to address this problem, we investigated
+ChIP-exo data. In order to address this problem, we investigated
 various aspects of ChIP exo data by contrasting them with their
@@
 
+To have a first view on the data, we calculated a series of the most
+commonly used quality indicators for ChIP-Seq data, which are listed
+in table \ref{tab:qcbase}. The PBC, NSC and RSC were calculated as
+defined in \cite{encode_qc} and the SSD was calculated as defined in
+\cite{htseq}. Additionally, we calculated the FSR to compare the
+number of fragments in each strand. A few conclussions that we may get
+by observing the PBC in table \ref{tab:qcbase} are all the samples
+have a quite low library complexity, hence it would be suggested to
+repeat the experiment.
+
+\begin{table}[h]
+  \centering
+\begin{tabular}{l|r|r|r|r|r|r|r}
+\hline\hline
+Condition & Rep. & Nr. reads & PBC & SSD & FSR & NSC & RSC \\
+\hline\hline
+Aerobic & 1 & 1,454,566 & 0.2156 & 0.0830 & 0.5012 & 114.14 & -1.670\\
+\hline
+Aerobic & 2 & 864,714 & 0.2471 & 0.0488 & 0.4923 & 160.93 & -1.503 \\
+\hline
+Rif-0min & 1 & 960,256 & 0.2823 & 0.0361 & 0.4998 & 10.29 & -0.092\\
+\hline
+Rif-0min & 2 & 2,247,295 & 0.2656 & 0.1091 & 0.5047 &  25.08 & -0.2279 \\
+\hline
+Rif-20min & 1 & 1,940,387 & 0.2698 & 0.0820 & 0.5054 & 17.69 & -0.450 \\
+\hline
+Rif-20min & 2 & 4,229,574 & 0.2153 & 0.1647 & 0.5071 &  14.11 & -0.533\\
+\hline
+\end{tabular}  
+\caption{Usual quality control indicators applied to our $\sigma^{70}$ samples.
+PBC stands PCR-bottleneck coefficient, SSD for standarized standard deviation, FSR for forward strand ratio, NSC for normalized strand cross-correlation coefficient and RSC for relative strand cross-correlation coefficient.}
+  \label{tab:qcbase}
+\end{table}
+
+\subsection{ChIP exo Quality Control pipeline}
+\label{sec:QC}
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width = .9\textwidth]{../figs/for_paper/coverage_diagram.pdf}
+  \caption{Diagram of the ChIP-exo Quality control pipeline(this is a
+    place-holder). The genome is partitioned into several regions by
+    removing the empty segments and for each regions we calculate a
+    collection of qc indicators.}
+  \label{fig:qcdiagram}
+\end{figure}
+
 \subsection{Characteristics of ChIP exo data}
@@
 
-\subsection{ChIP exo Quality Control pipeline}
-\label{sec:QC}
+\subsection{Comparison with ChIP-Seq data}
+\label{sec:comp}
+
+\subsection{Recomendations for the desgin of ChIP-exo experiments}
+\label{sec:reco}
+
+<<design_param , include = FALSE , echo = FALSE , eval = TRUE>>=
+
+  K <- 10
+  ext <- 15
+
+@
+
+We sampled a fixed amoung of fragments for each of the ChIP-exo, PET
+ChIP-Seq and SET ChIP-Seq datasets of the $\sigma^{70}$ sample in
+aerobic conditions. For each sampled dataset we applied our
+lower-to-higher resolution pipeline by calling peaks with MOSAiCS
+\cite{mosaics} and then deconvoling the binding events by using dPeak
+\cite{dpeak}. For the ChIP-exo datasets we called peaks by using
+GC-content and mappability with MOSAiCS, and for the ChIP-Seq datasets
+we used their respective Input samples.
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width = .8\textwidth]{../figs/for_paper/Sig70_aerobic_saturation.pdf}
+  \caption{Comparison of the number of candidate regions (A),
+    predicted events (B), identified targets (C) and resolution (D)
+    among ChIP-exo, PET ChIP-Seq and SET ChIP-Seq. RegulonDB
+    annotations are considered as a gold standard. A gold standard was
+    considered identified if a BS was estimated inside a \Sexpr{ext}
+    bp window around it}
+  \label{fig:design}
+\end{figure}
+
+In figure \ref{fig:design} we compared the behaviour of each data type
+when their depth is fixed.
+
+\newpage
 

### 2015-11-09 rmarkdown/chip_exo_paper.tex
Finished saturation analysis for Sig70 aerobic sample
Fixed typos in bibliography, updated figures to include the number of element in the whole paper figure, added a segment about recomendations for design in article

@@
 
-\subsection{ChIP exo Quality Control pipeline}
-\label{sec:QC}
+\subsection{Comparison with ChIP-Seq data}
+\label{sec:comp}
+
+\subsection{Recomendations for the desgin of ChIP-exo experiments}
+\label{sec:reco}
+
+We sampled a fixed amoung of fragments for each of the ChIP-exo, PET
+ChIP-Seq and SET ChIP-Seq datasets of the $\sigma^{70}$ sample in
+aerobic conditions. For each sampled dataset we applied our
+lower-to-higher resolution pipeline by calling peaks with MOSAiCS
+\cite{mosaics} and then deconvoling the binding events by using dPeak
+\cite{dpeak}. For the ChIP-exo datasets we called peaks by using
+GC-content and mappability with MOSAiCS, and for the ChIP-Seq datasets
+we used their respective Input samples.
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width = .8\textwidth]{../figs/for_paper/Sig70_aerobic_saturation.pdf}
+  \caption{Comparison of the number of candidate regions (A),
+    predicted events (B), identified targets (C) and resolution (D)
+    among ChIP-exo, PET ChIP-Seq and SET ChIP-Seq. RegulonDB
+    annotations are considered as a gold standard. A gold standard was
+    considered identified if a BS was estimated inside a 15
+    bp window around it}
+  \label{fig:design}
+\end{figure}
+
+In figure \ref{fig:design} we compared the behaviour of each data type
+when their depth is fixed.
+
+\newpage
 

### 2015-11-09 rscripts/ChIPQC_diagram_coverage_plot.R

@@ -50,13 +50,13 @@ dt <- do.call(rbind,dt)
 p <- ggplot(dt,aes(coords,tagCounts,colour = what))+geom_line()+
   scale_colour_manual(values = c("black","blue","red"))+xlab("Genomic coordinates")+
   theme_bw()+theme(legend.position = "none",axis.text = element_text(size = 0),
-                   axis.ticks = element_blank())+
+                   axis.ticks = element_blank(),plot.title = element_text(hjust = 0))+
   geom_abline(slope = 0,intercept = 5,colour = "cornsilk4",linetype =2)+ylim(0,500)+
   ylab("Number of reads by position")
 
 
-pdf(file = "figs/for_paper/coverage_diagram.pdf",width = 12 ,height = 5)
-print(p)
+pdf(file = "figs/for_paper/coverage_diagram.pdf",width = 9 ,height = 4)
+print(p + ggtitle("A"))
 dev.off()
 
 

### 2015-11-09 rscripts/sig70_saturation_all.R

@@ -166,16 +166,16 @@ npeaks <- summaries[,length(unique(peakId)), by = .(seed,what,sample)]
 npeaks[ ,sample := as.numeric(gsub("K","",sample)) ]
 p1 <- ggplot(npeaks,aes(sample,V1,colour = what))+geom_point(size = 1)+
   geom_smooth(method = "loess",se = FALSE,size = 1.5)+
-  ylim(150,375)+scale_color_brewer(palette = "Set1",name = "Protocol")+
-  theme_bw()+theme(legend.position = "top")+
+  ylim(150,375)+scale_color_brewer(palette = "Set1",name = "")+
+  theme_bw()+theme(legend.position = "top",plot.title = element_text(hjust = 0))+
   xlab("Nr. reads (in thousands)")+ylab("Nr. of candidate peaks")
 
 npred <- summaries[,length(peakId),by = .(seed,what,sample)]
 npred[ ,sample := as.numeric(gsub("K","",sample)) ]
 p2 <- ggplot(npred,aes(sample,V1,colour = what))+geom_point(size = 1)+
   geom_smooth(method = "loess",se = FALSE,size = 1.5)+
-  scale_color_brewer(palette = "Set1",name = "Protocol")+
-  theme_bw()+theme(legend.position = "top")+ylim(300,1100)+
+  scale_color_brewer(palette = "Set1",name = "")+
+  theme_bw()+theme(legend.position = "top",plot.title = element_text(hjust = 0))+ylim(300,1100)+
   xlab("Nr. reads (in thousands)")+ylab("Nr. of predicted events")
 
 ext <- 15
@@ -183,8 +183,8 @@ niden <- summaries[ ,sum(dist <= ext),by = .(seed,what,sample)]
 niden[ ,sample := as.numeric(gsub("K","",sample)) ]
 p3 <- ggplot(niden,aes(sample,V1,colour = what))+geom_point(size = 1)+
   geom_smooth(method = "loess",se = FALSE,size = 1.5)+
-  scale_color_brewer(palette = "Set1",name = "Protocol")+
-  theme_bw()+theme(legend.position = "top")+
+  scale_color_brewer(palette = "Set1",name = "")+
+  theme_bw()+theme(legend.position = "top",plot.title = element_text(hjust = 0))+
   xlab("Nr. reads (in thousands)")+ylab("Nr. of identificed targets")
 
[…]

### 2015-11-09 rmarkdown/chip_exo_paper.R
fixed typo in section title and adding chunks of R code in paper

@@ -1,0 +1,6 @@
+## ----design_param , include = FALSE , echo = FALSE , eval = TRUE---------
+
+  K <- 10
+  ext <- 15
+  
+

### 2015-11-10 rmarkdown/chip_exo_paper.Rnw
fixed typo in section title and adding chunks of R code in paper
Added comparison with chipseq figure on paper

@@
 ChIP-exo data. In order to address this problem, we investigated
-various aspects of ChIP exo data by contrasting them with their
-respective ChIP Seq experiments.
+various aspects of ChIP-exo data by contrasting them with their
+respective ChIP-Seq experiments.
 
@@
 in its very early stage. Although many methods have been proposed to
-identify protein binding sites from ChIP Seq data (reviewed in
+identify protein binding sites from ChIP-Seq data (reviewed in
 \cite{evaluation} and \cite{computation}), such as MACS \cite{macs},
@@
 
-\subsection{ChIP exo Quality Control pipeline}
+\subsection{ChIP-exo Quality Control pipeline}
 \label{sec:QC}
+
+Figure \ref{fig:qcdiagram} shows a flowchart for the ChIPexoQC
+pipeline. Which basically partitions genome by keeping the
+non-diggested ChIP-exo regions. Then, for each region it calculates a
+series of independent statistics. Finally, it creates several
+visualizations that were designed to asses the overall quality levels
+of A ChIP-exo sample.
 
@@
 
-\subsection{Characteristics of ChIP exo data}
-\label{sec:charac}
+\subsubsection{Assessing the strand imbalance in ChIP-exo data}
+\label{sec:strand_imbalance}
+
+The strand imbalance assesment is based in the observation that the
+enriched regions usually have a higher concentration of fragments,
+therefore we examine the FSR (defined as the ratio of the number of
+forward stranded reads divided by the total number of reads in a given
+region) as the region with lower depth are being filtered out. This
+indicator is of particular importance, since several methods rely on
+the ``peak-pair'' assumption. In table \ref{tab:qcbase}, we calculated
+the FSR and noticed that for all the samples, it's value is close to
+0.5, which means that there are roughly the same amount of reads in
+both strands. However, as we can see in figure \ref{fig:fsr}
+
+\begin{figure}[h]
+  \centering
+
+  \caption{This is a place-holder figure by now. The idea of this segments it }
+  \label{fig:fsr}
+\end{figure}
 
@@
 
-\subsection{Recomendations for the desgin of ChIP-exo experiments}
+We compared
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width = .6\textwidth,page = 3 ]{../figs/for_paper/ChIPseqPET_ChIPexo_tagCount_comparison.pdf}
+  \includegraphics[width = .9\textwidth ]{../figs/for_paper/MA_plot_fwd_bwd_by_seq.pdf}
+  \caption{Comparison}
+  \label{fig:comp}
+\end{figure}
+
+% \[
+% \mbox{FSR} = \frac{\mbox{Number of forward strand reads in region}}{\mbox{Total number of reads in region}}
+% \]
+
+\subsection{Recomendations for the design of ChIP-exo experiments}
 \label{sec:reco}
@@
 In figure \ref{fig:design} we compared the behaviour of each data type
-when their depth is fixed.
+when their depth is fixed. It is remarkable that even when the number
+of candiate peaks or the number of predicted events in quite lower
+than for both ChIP-Seq cases, it outperforms them in the number of
+identified targets and in resolution. This may suggest that with
+ChIP-exo less false positive peaks are being called and that when the
+targets are being identified, dpeak is able to estimate binding
+locations closer to the actual true cases. Additionally, we can see
+that as the read depth increases all four indicators do so as well,
+which may indicate that with ChIP-exo a smaller amount of reads is
+necessary to identify a higher number of targets, but it may also be
+possible that this is an artifact occurring due to ChIP-exo's lower
+library complexity.
 
@@
 
-\section*{Methods}
+\section{Methods}
 \label{sec:methods}
+
+\subsection{Growth conditions}
+\label{sec:growth}
+
+\subsection{ChIP-exo experiments}
+\label{sec:experiments}
+
+\subsection{Library preparation, sequencing and mapping of sequencing reads}
+\label{sec:library}
+
+\subsection{Method comparison for ChIP-exo}
+\label{sec:method_comp}
 

### 2015-11-10 rscripts/ChIPQC_diagram_coverage_plot.R

@@ -28,7 +28,7 @@ covers <- lapply(frags,coverage)
 
 chr <- "U00096"
 start <-  5000
-end <- 15000
+end <- 12000
 
 reg <- GRanges(seqnames = chr,ranges = IRanges(
   start = start, end= end),strand = "*")

### 2015-11-10 rscripts/sig70_library_complexity.R

@@ -94,14 +94,203 @@ create_regions <- function(reads,lower,rangesOnly=TRUE)
   stopifnot(lower > 0)
   cover <- coverage(reads)
   islands <- slice(cover,lower = lower,rangesOnly = rangesOnly)
+  islands <- as(islands,"GRanges")
   return(islands)
 }
 
 gr <- lapply(reads,function(x)dt2gr(rbind(readsF(x)[[1]],readsR(x)[[1]])))
-
 regs <- lapply(gr,create_regions,lower  = 1)
-regs <- lapply(gr,as,"GRanges")
 
-## ovs <- mcmapply(findOverlaps,gr,regs,SIMPLIFY = FALSE,mc.cores = 6)
- 
+build_stats <- function(region,reads)
+{
+  ## fix formats and stuff
+  ov <- findOverlaps(region,reads)
+  reads <- gr2dt(reads)
+  w <- width(region)    
+  region <- gr2dt(region)
+  region[ , width := w]
+  region[, match := paste0(seqnames,":",start,"_",end)]
+  reads[  subjectHits(ov), match := region[queryHits(ov), (match)] ]  
+  reads[,strand := ifelse(strand == "+", "F","R")]
 
+  ## get base statistics
+  f <- reads[,sum(strand == "F"),by = match]
+  setnames(f,names(f),c("match","f"))
+  setkey(f,match)
+  r <- reads[,sum(strand == "R"),by = match]
+  setkey(r,match)
+  setnames(r,names(r),c("match","r"))
+  f_uniq <- reads[strand == "F",length(unique(start)),by = match]
+  setnames(f_uniq,names(f_uniq),c("match","f_pos"))
+  setkey(f_uniq,match)
+  r_uniq <- reads[strand == "R",length(unique(end)),by = match]
+  setnames(r_uniq,names(r_uniq),c("match","r_pos"))
+  setkey(r_uniq,match)
+
+  ## merge statistics
+  stats <- merge(region,f,by = "match",allow.cartesian = TRUE)
+  stats <- merge(stats,r,by = "match",allow.cartesian = TRUE)
+  stats <- merge(stats,f_uniq,by = "match",allow.cartesian = TRUE,all = TRUE)
+  stats <- merge(stats,r_uniq,by = "match",allow.cartesian = TRUE, all = TRUE)
+  stats[is.na(f_pos), f_pos := 0]
+  stats[is.na(r_pos), r_pos := 0]
[…]

### 2015-11-10 rscripts/sig70_saturation_all.R

@@ -185,7 +185,7 @@ p3 <- ggplot(niden,aes(sample,V1,colour = what))+geom_point(size = 1)+
   geom_smooth(method = "loess",se = FALSE,size = 1.5)+
   scale_color_brewer(palette = "Set1",name = "")+
   theme_bw()+theme(legend.position = "top",plot.title = element_text(hjust = 0))+
-  xlab("Nr. reads (in thousands)")+ylab("Nr. of identificed targets")
+  xlab("Nr. reads (in thousands)")+ylab("Nr. of identified targets")
 
 reso <- summaries[, min(dist), by = .(seed,what,sample,peakId)]
 reso <- reso[,median(V1), by = .(seed,what,sample)]

### 2015-11-10 rmarkdown/chip_exo_paper.Rnw
Finished analysis of local SCC for Sig70 data

@@
 
-\begin{figure}[h]
-  \centering
+% \begin{figure}[h]
+%   \centering
 
-  \caption{This is a place-holder figure by now. The idea of this segments it }
-  \label{fig:fsr}
-\end{figure}
+%   \caption{This is a place-holder figure by now. The idea of this segments it }
+%   \label{fig:fsr}
+% \end{figure}
 
@@
 
-We compared
+<<comp_param,include = FALSE,echo = FALSE,eval = TRUE>>=
+
+  bin_size <- 150
+  frag_len <- 150
+
+@
+
+We first compared various factors that could affect binding site
+identification between ChIP-exo and ChIP-Seq data. in order to compare
+distribution of signal and background between ChIP-exo and ChIP-Seq
+data, we calculated ChIP tag counts accross the genome by counting the
+number of reads mapping to each of \Sexpr{bin_size} non-overlapping
+window after extending reads by \Sexpr{frag_len} to their 3' end
+directions. ChIP tag counts in ChIP-exo data were linearly related to
+ChIP tag counts in ChIP-Seq data for the regions with high ChIP tag
+counts (Figure \ref{fig:comp}A). This implies that signals for
+pontential binding sites are well reproducible between ChIP-exo and
+ChIP-Seq data. On the other hand, there was clear difference in the
+background distribution between them. In CHIP-Seq datam reads were
+almost uniformly distributed over background (non-binding) regions and
+the ChIP tag counts in there regions were significantly larger than
+zero. In contrast, in ChIP-exo data, there was larger variation in
+ChIP tag counts among background regions and ChIP tag counts were much
+lower in these regions compared to ChIP-Seq data. There were also
+large proportion of regions without any read in ChIP-exo data. There
+results indicate that background distribution of CHIP-exo data is less
+homogeneous than that of ChIP-Seq data.
 

### 2015-11-10 rmarkdown/chip_exo_paper.tex
Finished analysis of local SCC for Sig70 data

@@
 
-We compared
+We first compared various factors that could affect binding site
+identification between ChIP-exo and ChIP-Seq data. in order to compare
+distribution of signal and background between ChIP-exo and ChIP-Seq
+data, we calculated ChIP tag counts accross the genome by counting the
+number of reads mapping to each of 150 non-overlapping
+window after extending reads by 150 to their 3' end
+directions. ChIP tag counts in ChIP-exo data were linearly related to
+ChIP tag counts in ChIP-Seq data for the regions with high ChIP tag
+counts (Figure \ref{fig:comp}A). This implies that signals for
+pontential binding sites are well reproducible between ChIP-exo and
+ChIP-Seq data. On the other hand, there was clear difference in the
+background distribution between them. In CHIP-Seq datam reads were
+almost uniformly distributed over background (non-binding) regions and
+the ChIP tag counts in there regions were significantly larger than
+zero. In contrast, in ChIP-exo data, there was larger variation in
+ChIP tag counts among background regions and ChIP tag counts were much
+lower in these regions compared to ChIP-Seq data. There were also
+large proportion of regions without any read in ChIP-exo data. There
+results indicate that background distribution of CHIP-exo data is less
+homogeneous than that of ChIP-Seq data.
 

### 2015-11-10 rmarkdown/chip_exo_paper.R

@@ -1,3 +1,9 @@
+## ----comp_param,include = FALSE,echo = FALSE,eval = TRUE-----------------
+
+  bin_size <- 150
+  frag_len <- 150
+  
+
 ## ----design_param , include = FALSE , echo = FALSE , eval = TRUE---------
 
   K <- 10

### 2015-11-12 rmarkdown/chip_exo_paper.Rnw
finished complexity analysis, and added comments to paper

@@
 \cite{exo_review}. Hence, most of current guidelines \cite{encode_qc}
-may not be aplicable on ChIP-exo. To address this challenge, we
-suggest a series of quality control visualizations to understand which
-biases are present in ChIP-exo data. Related to quality control,
-Previous ChIP-exo analysis used ChIP-Seq samples to compare the
-resolution between experiments (\cite{exo1}, \cite{exo2},
+may not be aplicable on ChIP-exo, additionally to our knowledge there
+are not established quality control pipelines for ChIP-exo. To address
+this challenge, we suggest a series of quality control visualizations
+to understand which biases are present in ChIP-exo data. Related to
+quality control, Previous ChIP-exo analysis used ChIP-Seq samples to
+compare the resolution between experiments (\cite{exo1}, \cite{exo2},
 \cite{exoillumina}). In \cite{carroll.qc}, Carroll et. al. studied the
@@
   \centering
-\begin{tabular}{l|r|r|r|r|r|r|r}
+\begin{tabular}{l|r|r|r|r|r|r}
 \hline\hline
-Condition & Rep. & Nr. reads & PBC & SSD & FSR & NSC & RSC \\
+Condition & Rep. & Nr. reads & PBC & SSD & FSR & NSC  \\
 \hline\hline
-Aerobic & 1 & 1,454,566 & 0.2156 & 0.0830 & 0.5012 & 114.14 & -1.670\\
+Aerobic & 1 & 1,454,566 & 0.2156 & 0.0830 & 0.5012 & 114.14 \\
 \hline
-Aerobic & 2 & 864,714 & 0.2471 & 0.0488 & 0.4923 & 160.93 & -1.503 \\
+Aerobic & 2 & 864,714 & 0.2471 & 0.0488 & 0.4923 & 160.93  \\
 \hline
-Rif-0min & 1 & 960,256 & 0.2823 & 0.0361 & 0.4998 & 10.29 & -0.092\\
+Rif-0min & 1 & 960,256 & 0.2823 & 0.0361 & 0.4998 & 10.29 \\
 \hline
-Rif-0min & 2 & 2,247,295 & 0.2656 & 0.1091 & 0.5047 &  25.08 & -0.2279 \\
+Rif-0min & 2 & 2,247,295 & 0.2656 & 0.1091 & 0.5047 &  25.08  \\
 \hline
-Rif-20min & 1 & 1,940,387 & 0.2698 & 0.0820 & 0.5054 & 17.69 & -0.450 \\
+Rif-20min & 1 & 1,940,387 & 0.2698 & 0.0820 & 0.5054 & 17.69  \\
 \hline
-Rif-20min & 2 & 4,229,574 & 0.2153 & 0.1647 & 0.5071 &  14.11 & -0.533\\
+Rif-20min & 2 & 4,229,574 & 0.2153 & 0.1647 & 0.5071 &  14.11 \\
 \hline
 \end{tabular}  
-\caption{Usual quality control indicators applied to our $\sigma^{70}$ samples.
-PBC stands PCR-bottleneck coefficient, SSD for standarized standard deviation, FSR for forward strand ratio, NSC for normalized strand cross-correlation coefficient and RSC for relative strand cross-correlation coefficient.}
+  \caption{Usual quality control indicators applied to our
+    $\sigma^{70}$ samples. PBC stands PCR-bottleneck coefficient, SSD
+    for standarized standard deviation, FSR for forward strand ratio,
+    NSC for normalized strand cross-correlation coefficient. We
+    ommited the RSC (relative strand cross-correlation since ChIP-exo
+    experiments don't usually have an acomnying input file)}
+
   \label{tab:qcbase}
 \end{table}
+
+In a good ChIP-Seq experiment the strand cross-correlation is
+maximized at the fragment length and there is also a ``phantom peak''
+at the sample's read length (\cite{encode_qc} and \cite{strandcc}).
+However, in ChIP-exo's case this two peaks are confounded. Hence an
+enrichment measure based in the strand cross-correlation such as NSC
+or RSC is harder to interpret. In figure \ref{figsup:scc} we plotted
+the SCC curves for $\sigma^{70}$ ChIP-exo experiments used to
+calculate the NSC and RSC values in table \ref{tab:qcbase}. According
+to ENCODE's quality metrics, NSC values close to one corresponds to
+low enrichment samples or bad quality datasets, hence in our case the
+last four samples would be considered to have higher quality than they
+actually have. On the other hand, all PBC values are considerably low
+and correspond to samples with severe bottlenecking issues, in
+particular we are going to show that for the first two samples we are
+obtaining a smaller resolution and identifying a comparable number of
+targets as in ChIP-Seq PET experiments.
 
@@
 series of independent statistics. Finally, it creates several
-visualizations that were designed to asses the overall quality levels
-of A ChIP-exo sample.
+visualizations designed to asses the overall quality levels of A
+ChIP-exo sample.
 
@@
 
-\subsubsection{Assessing the strand imbalance in ChIP-exo data}
+\color{blue}
+
+\subsubsection{Strand imbalance in ChIP-exo data}
 \label{sec:strand_imbalance}
@@
 0.5, which means that there are roughly the same amount of reads in
-both strands. However, as we can see in figure \ref{fig:fsr}
+both strands. To asses
+
+\color{black}
+
+\subsubsection{Enrichment and library complexity in ChIP-exo data}
+\label{sec:enri}
 
@@
 ChIP-Seq data. On the other hand, there was clear difference in the
-background distribution between them. In CHIP-Seq datam reads were
+background distribution between them. In ChIP-Seq data reads were
 almost uniformly distributed over background (non-binding) regions and
@@
 lower in these regions compared to ChIP-Seq data. There were also
-large proportion of regions without any read in ChIP-exo data. There
-results indicate that background distribution of CHIP-exo data is less
+large proportion of regions without any read in ChIP-exo data. These
+results indicate that background distribution of ChIP-exo data is less
 homogeneous than that of ChIP-Seq data.
+
+\color{blue}
+
+We next evaluated the ``peak pair'' assumption of Rhee and Pugh
+\cite{exo1}, MACE \cite{MACE}, CexoR, \cite{cexor} and peakzilla
+\cite{peakzilla}, i.e. a peak of reads in the forward strand is
+usually paired with a peak of reads in the reverse strand that is
+located in the other site of the binding site.
+
+\color{black}
 
@@
   \includegraphics[width = .6\textwidth,page = 3 ]{../figs/for_paper/ChIPseqPET_ChIPexo_tagCount_comparison.pdf}
-  \includegraphics[width = .9\textwidth ]{../figs/for_paper/MA_plot_fwd_bwd_by_seq.pdf}
-  \caption{Comparison}
+  % \includegraphics[width = .9\textwidth ,page = 3]{../figs/for_paper/MA_plot_fwd_bwd_by_seq.pdf}
+  \caption{ A) ChIP signal in ChIP-exo was linearly related to that in
+    PET ChIP-Seq data in the region with high ChIP tag counts. In
+    contrast, there wer clear differences in their background
+    distribution, where several of the background ChIP-exo regions
+    were empty.}
   \label{fig:comp}
@@
 
+\color{red}
+
 \subsection{Growth conditions}
@@
 
+\color{black}
+
 \subsection{Method comparison for ChIP-exo}
 \label{sec:method_comp}
+
+\newpage
+
+\section*{Supplement figures}
+
+\subsection*{Strand cross-correlation for ChIP-exo data}
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width = .8\textwidth]{../figs/for_paper/EColi_strand_cross_corr.pdf}
+  \caption{Strand cross-correlation curves for $\sigma^{70}$
+    samples.}
+  \label{figsup:scc}
+\end{figure}
+
+\newpage
 

### 2015-11-12 rscripts/sig70_library_complexity.R

@@ -176,7 +176,7 @@ p <- ggplot(aux , aes( ave_reads,cover_rate))+stat_binhex(bins = 50)+
     labels = trans_format("log10",math_format(10^.x)))+xlim(0,7)+ylim(0,.6)+
   theme_bw()+theme(legend.position = "top",plot.title = element_text(hjust = 0))+
   xlab("Average read coverage")+
-  ylab("Uniquue read coverage rate")
+  ylab("Unique read coverage rate")
 print(p + ggtitle("All regions"))
 print( p  %+% aux[ npos > 10] + ggtitle("Npos > 10")) 
 print( p  %+% aux[ npos > 30] + ggtitle("Npos > 30")) 
@@ -358,8 +358,6 @@ grid.arrange(strata_plots[[1]]+ggtitle("A"),
              strata_plots[[2]]+ggtitle("B"),
              strata_plots[[3]]+ggtitle("C"),nrow = 3)
 dev.off()
-
-
 
 plots <- list()
 plots[[1]] <- ggplot(dat , aes( edsn , noise , colour = edsn))+geom_boxplot()+

### 2015-11-12 rmarkdown/chip_exo_paper.Rnw
Changed the size of enrichment plots and added labels to plots

@@
 
+\newpage
+
 \section{Results and discussion}
@@
 
+\newpage
+
 \subsection{ChIP-exo Quality Control pipeline}
@@
 \label{sec:enri}
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width = .8\textwidth]{../figs/for_paper/Sig70_enrichment.pdf}
+  \includegraphics[width = .8\textwidth,page =3]{../figs/for_paper/Local_SCC_indicator_by_strata.pdf}
+  \caption{Enrichment and else}
+  \label{fig:enrich}
+\end{figure}
 
@@
     contrast, there wer clear differences in their background
-    distribution, where several of the background ChIP-exo regions
-    were empty.}
+    distribution, where several ChIP-exo background regions were
+    empty.}
   \label{fig:comp}
 \end{figure}
-
-% \[
-% \mbox{FSR} = \frac{\mbox{Number of forward strand reads in region}}{\mbox{Total number of reads in region}}
-% \]
 

### 2015-11-12 rscripts/sig70_library_complexity.R

@@ -1,4 +1,4 @@
-
+p
 rm(list = ls())
 
 library(GenomicAlignments)
@@ -177,6 +177,7 @@ p <- ggplot(aux , aes( ave_reads,cover_rate))+stat_binhex(bins = 50)+
   theme_bw()+theme(legend.position = "top",plot.title = element_text(hjust = 0))+
   xlab("Average read coverage")+
   ylab("Unique read coverage rate")
+print(p + ggtitle("A"))
 print(p + ggtitle("All regions"))
 print( p  %+% aux[ npos > 10] + ggtitle("Npos > 10")) 
 print( p  %+% aux[ npos > 30] + ggtitle("Npos > 30")) 
@@ -195,7 +196,6 @@ p <-ggplot(aux[f > 0 & r > 0 ] , aes( M , A))+stat_binhex(bins = 50)+
   facet_wrap( ~ edsn)+scale_fill_gradientn(colours = r,trans = "log10",
     labels = trans_format("log10",math_format(10^.x)))+
   theme_bw()+theme(legend.position = "top",plot.title = element_text(hjust = 0))
-dev.off()
 print(p + ggtitle("All regions"))
 print( p  %+% aux[ npos > 10] + ggtitle("Npos > 10")) 
 print( p  %+% aux[ npos > 30] + ggtitle("Npos > 30")) 
@@ -341,12 +341,12 @@ strata_plots[[2]] <- ggplot(dat , aes( strata , max,colour = edsn))+geom_boxplot
 strata_plots[[3]] <- ggplot(dat , aes( strata , nsc,colour = edsn))+geom_boxplot()+facet_grid( . ~ edsn) +
   scale_color_brewer(palette = "Dark2")+theme_bw()+
   theme(legend.position = "none", axis.text.x = element_text(angle = 90),plot.title = element_text(hjust = 0))+
-  xlab("")+ylab("Local NSC")+ylim(-1,6)+geom_abline( slope = 0 , intercept = 0, linetype = 2)
+  xlab("")+ylab("Local NSC")+ylim(-1,6)+geom_abline( slope = 0 , intercept = 0, linetype = 2)+ggtitle("B")
 
 
 save(strata_plots,file = "data/for_paper/Local_SCC_indicator_by_strata.RData")
 
-pdf(file = "figs/for_paper/Local_SCC_indicator_by_strata.pdf",width = 9,height = 6)
[…]

### 2015-11-12 rmarkdown/chip_exo_paper.Rnw
added comments in the enrichment plot

@@
 While the number of produced ChIP-exo data keeps increasing,
-characteristicos of ChIP-exo data and optimal strategies for
+characteristics of ChIP-exo data and optimal strategies for
 experimental design and analysis of ChIP-exo data are not fully
@@
 
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width = .8\textwidth]{../figs/for_paper/EColi_strand_cross_corr.pdf}
+  \caption{Strand cross-correlation curves for $\sigma^{70}$
+    samples.}
+  \label{figsup:scc}
+\end{figure}
+
 In a good ChIP-Seq experiment the strand cross-correlation is
@@
 targets as in ChIP-Seq PET experiments.
-
-\newpage
 
@@
 
+In ChIP-exo experiments and input samples is often unavailable,
+therefore to determine the quality of a ChIP-exo experiment is
+necesary to address the balance between the enrichment and the library
+complexity of the experiment. To diagnose this balance we considered
+the Average Read Coverage (ARC) and the Unique Read Coverage Rate (URCR) which are defined as:
+
+\begin{align}
+  \mbox{ARC} &= \frac{\text{Nr. of reads in the region}}{\text{Width of the region}} \nonumber \\
+  \mbox{URCR} &= \frac{\text{Nr. of reads mapped to only one position
+      in the region}}{\text{Nr. of reads in the region}} \nonumber
+\end{align}
+
 \begin{figure}[h!]
@@
   \includegraphics[width = .8\textwidth,page =3]{../figs/for_paper/Local_SCC_indicator_by_strata.pdf}
-  \caption{Enrichment and else}
+  \caption{A) For all $\sigma^{70}$ ChIP-exo experiments we are
+    comparing the ARC against the URCR.}
   \label{fig:enrich}
 \end{figure}
-
-% \begin{figure}[h]
-%   \centering
-
-%   \caption{This is a place-holder figure by now. The idea of this segments it }
-%   \label{fig:fsr}
-% \end{figure}
 
@@
 
-\newpage
-
-\section*{Supplement figures}
-
-\subsection*{Strand cross-correlation for ChIP-exo data}
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = .8\textwidth]{../figs/for_paper/EColi_strand_cross_corr.pdf}
-  \caption{Strand cross-correlation curves for $\sigma^{70}$
-    samples.}
-  \label{figsup:scc}
-\end{figure}
+% \newpage
+% \section*{Supplement figures}
+% \subsection*{Strand cross-correlation for ChIP-exo data}
 

### 2015-11-12 rscripts/sig70_library_complexity.R

@@ -1,4 +1,4 @@
-p
+
 rm(list = ls())
 
 library(GenomicAlignments)
@@ -157,8 +157,8 @@ stats <- mcmapply(build_stats,regs,gr,SIMPLIFY = FALSE,mc.cores = 6)
 
 aux <- mapply(function(x,y)x[,edsn := y],stats,exo[,(edsn)],SIMPLIFY =FALSE)
 aux <- do.call(rbind,aux)
-aux[ , edsn := factor(edsn , levels  = c(1311,1317, 931 , 1314,1320 , 933))]
-aux[ , edsn := plyr::mapvalues(edsn , from = c(1311,1317, 931 , 1314,1320 , 933),
+aux[ , edsn := factor(edsn , levels  = c(1311,1314, 931 , 1317,1320 , 933))]
+aux[ , edsn := plyr::mapvalues(edsn , from = c(1311,1314, 931 , 1317,1320 , 933),
           to = c("Rif0_rep1","Rif20_rep1","Aerobic_rep1","Rif0_rep2","Rif20_rep2","Aerobic_rep2"))]
          
 
@@ -327,6 +327,12 @@ dat[ , strata := plyr::mapvalues(strata ,
            "(10,15)"             
            ))]
 
+dat[ , edsn := factor(edsn , levels  = c(1311,1317,1314,1320, 931 ,  933))]
+dat[ , edsn := plyr::mapvalues(edsn , from = c(1311,1314, 931 , 1317,1320 , 933),
+          to = c("Rif0_rep1","Rif20_rep1","Aerobic_rep1","Rif0_rep2","Rif20_rep2","Aerobic_rep2"))]
+
+
+            
 
 strata_plots <- list()
 strata_plots[[1]] <- ggplot(dat , aes( strata , noise,colour = edsn))+geom_boxplot()+facet_grid( . ~ edsn) +

### 2015-11-14 rmarkdown/chip_exo_paper.Rnw
Added segment to paper about enrichment and library complexity and fixed figures to include labels

@@
   \caption{Strand cross-correlation curves for $\sigma^{70}$
-    samples.}
+    samples. The ``phantom peak'' and the summit that corresponds to
+    the read and fragment length respectively are confounded, due to
+    the exonuclease enzyme digestion.}
   \label{figsup:scc}
@@
 
+\subsubsection{Enrichment and library complexity in ChIP-exo data}
+\label{sec:enri}
+
+In ChIP-exo experiments, the background is often digested by the
+exonuclease enzyme, therefore to determine the sample's quality is
+necesary to address the balance between the enrichment and library
+complexity. To diagnose this, we considered the Average Read Coverage
+(ARC) and the Unique Read Coverage Rate (URCR) which are defined as:
+
+\begin{align}
+  \mbox{ARC} &= \frac{\text{Nr. of reads in the region}}{\text{Width of the region}} \nonumber \\
+  \mbox{URCR} &= \frac{\text{Nr. of reads mapped to only one position
+      in the region}}{\text{Nr. of reads in the region}} \nonumber
+\end{align}
+
+Figure \ref{fig:enrich}A shows hexbin plots illustrating the
+interaction between these two quantities. Notice how there are two
+strong arms in each panel: The first one corresponds to regions with
+low $\mbox{ARC}$ values and $\mbox{URCR}$ values that can vary accross
+the whole $[0,1]$ interval, while the second one shows a decreasing
+trend in $\mbox{URCR}$ as $\mbox{ARC}$ increases. When an experiment
+shows a higher degree of enrichment, then the separation of this two
+arms is more noticeable, since the second arm corresponds to possibly
+enriched regions (\ref{fig:enrich2}A and \ref{fig:enrich2}B).
+
+<<strata_param , include =FALSE,echo = FALSE, eval = TRUE>>=
+
+  nstrata <- 100
+
+@
+
+Figure \ref{fig:enrich}B shows boxplots of the $\mbox{local-NSC}$
+stratified by the number of reads mapped exactly to one position for
+all $\sigma^{70}$ ChIP-exo experiments; for each stratum, we sampled
+\Sexpr{nstrata} regions that whenever possible (in the opposite case,
+all the regions in the category were considered). There is an
+increasing trend as the number of unique position increases, which
+means that this is an effective indicator of library complexity since
+regions where the reads are mapped to more positions are usually more
+complex.
+
+%% perhaps do two other slots about how this works... a good experiment vs a bad experiment
+%% of the same region
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width = .8\textwidth]{../figs/for_paper/Sig70_enrichment.pdf}
+  \includegraphics[width = .8\textwidth,page =3]{../figs/for_paper/Local_SCC_indicator_by_strata.pdf}
+  \caption{A) Hexbin plots of $\mbox{ARC}$ against $\mbox{URCR}$ and
+    B) boxplots of the $\mbox{local-NSC}$ stratified by nr. of reads
+    mapped to only one position in the regions for the ChIP-exo
+    $\sigma^{70}$ experiments divided by condition and biological
+    replicate}
+  \label{fig:enrich}
+\end{figure}
+
 \color{blue}
@@
 \color{black}
-
-\subsubsection{Enrichment and library complexity in ChIP-exo data}
-\label{sec:enri}
-
-In ChIP-exo experiments and input samples is often unavailable,
-therefore to determine the quality of a ChIP-exo experiment is
-necesary to address the balance between the enrichment and the library
-complexity of the experiment. To diagnose this balance we considered
-the Average Read Coverage (ARC) and the Unique Read Coverage Rate (URCR) which are defined as:
-
-\begin{align}
-  \mbox{ARC} &= \frac{\text{Nr. of reads in the region}}{\text{Width of the region}} \nonumber \\
-  \mbox{URCR} &= \frac{\text{Nr. of reads mapped to only one position
-      in the region}}{\text{Nr. of reads in the region}} \nonumber
-\end{align}
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = .8\textwidth]{../figs/for_paper/Sig70_enrichment.pdf}
-  \includegraphics[width = .8\textwidth,page =3]{../figs/for_paper/Local_SCC_indicator_by_strata.pdf}
-  \caption{A) For all $\sigma^{70}$ ChIP-exo experiments we are
-    comparing the ARC against the URCR.}
-  \label{fig:enrich}
-\end{figure}
 
@@
 We next evaluated the ``peak pair'' assumption of Rhee and Pugh
-\cite{exo1}, MACE \cite{MACE}, CexoR, \cite{cexor} and peakzilla
+\cite{exo1}, MACE \cite{mace}, CexoR, \cite{cexor} and peakzilla
 \cite{peakzilla}, i.e. a peak of reads in the forward strand is
@@
   \includegraphics[width = .6\textwidth,page = 3 ]{../figs/for_paper/ChIPseqPET_ChIPexo_tagCount_comparison.pdf}
-  % \includegraphics[width = .9\textwidth ,page = 3]{../figs/for_paper/MA_plot_fwd_bwd_by_seq.pdf}
   \caption{ A) ChIP signal in ChIP-exo was linearly related to that in
@@
 \nocite{exo_gb}
+\nocite{maplot1}
+\nocite{maplot2}
+
+\newpage
+
+\section*{Supplement}
+\label{sec:supp}
+
+\subsection*{Additional Enrichment plots for $\sigma^{70}$}
+\label{sec:enrichsup}
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width = .8\textwidth,page = 3]{../figs/for_paper/Sig70_enrichment.pdf}
+  \includegraphics[width = .8\textwidth,page = 4]{../figs/for_paper/Sig70_enrichment.pdf}
+  \caption{Hexbin plots of $\mbox{ARC}$ vs $\mbox{URCR}$ for each
+    region after partitioning the genome. In A and B the regions with
+    reads mapped to at most 10 and 30 positions respectively where not
+    considered.}
+  \label{fig:enrich2}
+\end{figure}
+
+\subsection{Local-NSC}
+\label{sec:localnsc}
+
+The local strand cross-correlaiton is defined as:
+
+\begin{align}
+  \mbox{local-NSC} &= \frac{max_{x_\delta} f(x_\delta)}{\sigma} \nonumber \\
+ y_\delta &= f(x_\delta) + \sigma \epsilon_\delta \nonumber
+\end{align}
+
+where $f$ is calculated by fitting a non-parametric regression model
+over the strand cross-correlation calculated using exclusively the
+fragments that overlap a pre-defined genomic region, $y_\delta$ is the
+correlation of the forward and backward coverages when one is shifted
+$x_\delta$ bp towards their 3' end.
 

### 2015-11-14 rmarkdown/chip_exo_paper.tex
Added segment to paper about enrichment and library complexity and fixed figures to include labels

@@
 
+\subsubsection{Enrichment and library complexity in ChIP-exo data}
+\label{sec:enri}
+
+In ChIP-exo experiments, the background is often digested by the
+exonuclease enzyme, therefore to determine the sample's quality is
+necesary to address the balance between the enrichment and library
+complexity. To diagnose this, we considered the Average Read Coverage
+(ARC) and the Unique Read Coverage Rate (URCR) which are defined as:
+
+\begin{align}
+  \mbox{ARC} &= \frac{\text{Nr. of reads in the region}}{\text{Width of the region}} \nonumber \\
+  \mbox{URCR} &= \frac{\text{Nr. of reads mapped to only one position
+      in the region}}{\text{Nr. of reads in the region}} \nonumber
+\end{align}
+
+Figure \ref{fig:enrich}A shows hexbin plots illustrating the
+interaction between these two quantities. Notice how there are two
+strong arms in each panel: The first one corresponds to regions with
+low $\mbox{ARC}$ values and $\mbox{URCR}$ values that can vary accross
+the whole $[0,1]$ interval, while the second one shows a decreasing
+trend in $\mbox{URCR}$ as $\mbox{ARC}$ increases. When an experiment
+shows a higher degree of enrichment, then the separation of this two
+arms is more noticeable, since the second arm corresponds to possibly
+enriched regions (\ref{fig:enrich2}A and \ref{fig:enrich2}B).
+
+Figure \ref{fig:enrich}B shows boxplots of the $\mbox{local-NSC}$
+stratified by the number of reads mapped exactly to one position for
+all $\sigma^{70}$ ChIP-exo experiments; for each stratum, we sampled
+100 regions that whenever possible (in the opposite case,
+all the regions in the category were considered). There is an
+increasing trend as the number of unique position increases, which
+means that this is an effective indicator of library complexity since
+regions where the reads are mapped to more positions are usually more
+complex.
+
+%% perhaps do two other slots about how this works... a good experiment vs a bad experiment
+%% of the same region
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width = .8\textwidth]{../figs/for_paper/Sig70_enrichment.pdf}
+  \includegraphics[width = .8\textwidth,page =3]{../figs/for_paper/Local_SCC_indicator_by_strata.pdf}
+  \caption{A) Hexbin plots of $\mbox{ARC}$ against $\mbox{URCR}$ and
+    B) boxplots of the $\mbox{local-NSC}$ stratified by nr. of reads
+    mapped to only one position in the regions for the ChIP-exo
+    $\sigma^{70}$ experiments divided by condition and biological
+    replicate}
+  \label{fig:enrich}
+\end{figure}
+
 \color{blue}

### 2015-11-14 rmarkdown/chip_exo_paper.R

@@ -1,3 +1,8 @@
+## ----strata_param , include =FALSE,echo = FALSE, eval = TRUE-------------
+
+  nstrata <- 100
+  
+
 ## ----comp_param,include = FALSE,echo = FALSE,eval = TRUE-----------------
 
   bin_size <- 150

### 2015-11-14 rscripts/bin_analysis.R

@@ -44,14 +44,15 @@ read_files <- mapply(filter_reads,
 reads <- mapply(create_reads,read_files,c(FALSE,TRUE,FALSE),SIMPLIFY = FALSE)
 
 
-pdf(file = file.path("figs/for_paper","ChIPseqPET_ChIPexo_tagCount_comparison.pdf"))
+pdf(file = file.path("figs/for_paper","ChIPseqPET_ChIPexo_tagCount_comparison.pdf"),width = 6,height = 6)
 p <- hexbin_plot(reads[[2]],reads[[1]],150,frag_len = 150)+xlab("ChIP-seq (PET) tag counts")+ylab("ChIP-exo tag counts")+theme_bw()
 dt <- copy(p$data)
 dt[,x := 1 + x]
 dt[,y := 1 + y]
 print(p)    
 print(p + xlim(0,500)+ylim(0,500)+coord_fixed())
-z <- p %+% dt +coord_fixed() + scale_x_log10() + scale_y_log10()
+z <- p %+% dt +coord_fixed() + scale_x_log10() + scale_y_log10()+theme(plot.title = element_text(hjust = 0))+
+  ggtitle("A")
 print(z)
 dev.off()
 

### 2015-11-14 rscripts/peak_pair_assumption_new.R

@@ -2,12 +2,12 @@
 rm(list = ls())
 
 library(mosaics)
-library(dpeak)
 library(ChIPUtils)
 
+library(dpeak)
+
 exo_file <- "/p/keles/ChIPexo/volume3/LandickData/ChIPexo/edsn931_042814_qc.sorted.bam"
-pet_file1 <- "/p/keles/ChIPexo/volume3/LandickData/ChIPseq_PET/edsn788_042814_qc.filter.bam"
-pet_file2 <- "/p/keles/ChIPexo/volume3/LandickData/ChIPseq_PET/edsn788_042814_qc.sorted.bam"
+pet_file1 <- "/p/keles/ChIPexo/volume3/LandickData/ChIPseq_PET/edsn790_042814_qc.filter.bam"
 set_file <- "/p/keles/ChIPexo/volume3/LandickData/ChIPseq_SET/edsn80_042814_qc.sorted.bam"
 
 frag_len <- 150
@@ -54,7 +54,7 @@ if(bin_flag){
 
 
 
-peak_flag <- TRUE
+peak_flag <- FALSE
 
 if(peak_flag){
   exo_bin <- file.path(bin_dir,paste0(basename(exo_file),"_fragL",frag_len,"_bin",bin_size,".txt"))
@@ -115,12 +115,24 @@ if(peak_flag){
   write.table(set_peak , file = set_peak_file, sep = "\t",row.names = FALSE,quote = FALSE,col.names=FALSE)
 }
 
+  peak_dir <- file.path(out_dir,"peaks")
+
+  ## exo_peak_file <- file.path(peak_dir,gsub("_qc.sorted.bam","_peaks.txt",basename(exo_file)))
+  pet_peak_file1 <-  file.path(peak_dir,gsub("_qc.filter.bam", "_filter_peaks.txt",basename(pet_file1)))
+  ## pet_peak_file2 <- file.path(peak_dir, gsub("_qc.sorted.bam", "_all_peaks.txt",basename(pet_file2)))
+  ## set_peak_file <-  file.path(peak_dir,gsub("_qc.sorted.bam", "_peaks.txt",basename(set_file)))
+
+
 ## estimate binding events
 
 bs_dir <- file.path(out_dir,"binding")
 check_create(bs_dir)
 
 bind_flag <- TRUE
+
+## load_all("~/Desktop/Docs/Code/dpeak")
+
+
 
 if(bind_flag){
 
@@ -131,20 +143,20 @@ if(bind_flag){
[…]

### 2015-11-14 rscripts/sig70_library_complexity.R

@@ -82,9 +82,15 @@ exo[,which.max := SCC[,which.max(cross.corr),by = sample][,(V1)]]
 save(exo,file = "data/for_paper/sig70_summary.RData")
 
 
+SCC[ , sample := factor(sample , levels  = c(1311,1317,1314,1320, 931 ,  933))]
+SCC[ , sample := plyr::mapvalues(sample , from = c(1311,1314, 931 , 1317,1320 , 933),
+          to = c("Rif:0-rep1","Rif:20-rep1","Aerobic-rep1","Rif:0-rep2","Rif:20-rep2","Aerobic-rep2"))]
+
+
 pdf(file = "figs/for_paper/EColi_strand_cross_corr.pdf",width = 9 , height = 5)
 ggplot(SCC,aes(shift,cross.corr,colour = sample))+geom_line(size = .8)+
-  scale_color_brewer(palette = "Dark2")+theme_bw()+theme(legend.position = "top")
+  scale_color_brewer(name = "",palette = "Dark2")+theme_bw()+theme(legend.position = "top")+
+  ylab("Strand cross-correlation")
 dev.off()
 
 
@@ -159,7 +165,7 @@ aux <- mapply(function(x,y)x[,edsn := y],stats,exo[,(edsn)],SIMPLIFY =FALSE)
 aux <- do.call(rbind,aux)
 aux[ , edsn := factor(edsn , levels  = c(1311,1314, 931 , 1317,1320 , 933))]
 aux[ , edsn := plyr::mapvalues(edsn , from = c(1311,1314, 931 , 1317,1320 , 933),
-          to = c("Rif0_rep1","Rif20_rep1","Aerobic_rep1","Rif0_rep2","Rif20_rep2","Aerobic_rep2"))]
+          to = c("Rif:0-rep1","Rif:20-rep1","Aerobic-rep1","Rif:0-rep2","Rif:20-rep2","Aerobic-rep2"))]
          
 
 pdf(file = "figs/for_paper/Sig70_number_unique_positions_perSample.pdf")
@@ -179,8 +185,8 @@ p <- ggplot(aux , aes( ave_reads,cover_rate))+stat_binhex(bins = 50)+
   ylab("Unique read coverage rate")
 print(p + ggtitle("A"))
 print(p + ggtitle("All regions"))
-print( p  %+% aux[ npos > 10] + ggtitle("Npos > 10")) 
[…]

### 2015-11-16 rmarkdown/chip_exo_paper.Rnw
Replaced saturation, resolution and sensitivity analysis of the paper with new plots using the old data

@@
 \usepackage{verbatim}
+\usepackage{hyperref}
+
+\usepackage{hyperref}
+\hypersetup{
+     colorlinks   = true,
+     citecolor    = black
+}
+\hypersetup{linkcolor=black}
+\hypersetup{linkcolor=black}
+\hypersetup{linkcolor=black}
 
@@
 
-\color{blue}
-
 \subsubsection{Strand imbalance in ChIP-exo data}
@@
 0.5, which means that there are roughly the same amount of reads in
-both strands. To asses
-
-\color{black}
+both strands. However, figure \ref{fig:comp}B shows that this
+assumption don't hold in practice.
 
@@
 
-\color{blue}
-
 We next evaluated the ``peak pair'' assumption of Rhee and Pugh
@@
 usually paired with a peak of reads in the reverse strand that is
-located in the other site of the binding site.
+located in the other site of the binding site. In order to evaluate
+this assumptiom, we reviewed the proportion of reads in the forward
+streand in candidate regions (i.e. regions with at least one binding
+site) in $\sigma^{70}$ ChIP-exo data. We found that strands of reads
+were much less balanced in ChIP-exo data than in ChIP-Seq data in
+these regions with potential binding sites (Fig. \ref{fig:comp}B) and
+htis indicates that the peak pair assumptuon might not hold in real
+ChIP-exo data.
 
-\color{black}
+%  We obtained esentially the same conclusion from the
+% ChIP-exo data for CTCF factir from human sample Rhee and Pugh
+% (\cite{exo1}) as well.
 
@@
   \centering
-  \includegraphics[width = .6\textwidth,page = 3 ]{../figs/for_paper/ChIPseqPET_ChIPexo_tagCount_comparison.pdf}
+  \includegraphics[width = .46\textwidth,page = 3 ]{../figs/for_paper/ChIPseqPET_ChIPexo_tagCount_comparison.pdf}
+  \includegraphics[width = .46\textwidth]{../figs/for_paper/forward_strand_ratio_comp_old.pdf}
   \caption{ A) ChIP signal in ChIP-exo was linearly related to that in
@@
     distribution, where several ChIP-exo background regions were
-    empty.}
+    empty. B) Density of the proportion of the ratio of forward strand
+    reads in a ChIP-exo and ChIP-Seq SET data.}
   \label{fig:comp}
+\end{figure}
+
+\subsubsection{Comparison with ChIP-Seq data using dPeak}
+\label{sec:dpeak_analysis}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width = .4\textwidth]{../figs/for_paper/sensitivity_exo_olda_data.pdf}
+  \includegraphics[width = .4\textwidth]{../figs/for_paper/resolution_by_dataset_old_data.pdf}
+  \caption{Comparison of (A) sensitivity and (B) resolution between
+    ChIP-exo and ChIP-Seq data. Sensitivity is defined as the
+    propotion of regulonDB annotations identificed using each
+    data. Resolution is defined as the distance between regulonDB
+    annotation and its closest prediction.}
+  \label{fig:reso_all}
 \end{figure}
@@
 lower-to-higher resolution pipeline by calling peaks with MOSAiCS
-\cite{mosaics} and then deconvoling the binding events by using dPeak
+\cite{mosaics} and then deconvolving the binding events by using dPeak
 \cite{dpeak}. For the ChIP-exo datasets we called peaks by using
@@
   \centering
-  \includegraphics[width = .8\textwidth]{../figs/for_paper/Sig70_aerobic_saturation.pdf}
+%  \includegraphics[width = .8\textwidth]{../figs/for_paper/Sig70_aerobic_saturation.pdf}
+  \includegraphics[width = .8\textwidth]{../figs/for_paper/saturation_analysis_old.pdf}
   \caption{Comparison of the number of candidate regions (A),
@@
 
-In figure \ref{fig:design} we compared the behaviour of each data type
-when their depth is fixed. It is remarkable that even when the number
-of candiate peaks or the number of predicted events in quite lower
-than for both ChIP-Seq cases, it outperforms them in the number of
+Figure \ref{fig:design} shows the behaviour of each data type when
+their depth is fixed. It is remarkable that even when the number of
+candiate peaks or the number of predicted events in quite lower than
+for both ChIP-Seq cases, it outperforms them in the number of
 identified targets and in resolution. This may suggest that with
@@
 
-\subsection{Method comparison for ChIP-exo}
-\label{sec:method_comp}
+\subsection{Method comparison with ChIP-exo}
+\label{sec:suppcomp}
 
-% \newpage
-% \section*{Supplement figures}
-% \subsection*{Strand cross-correlation for ChIP-exo data}
+We considered dPeak Chung et al. \cite{dpeak}, GEM Guo et
+al. \cite{gem} and MACE Wang et al. \cite{mace} for ChIP-exo data
+analysis. For the dPeak algorithm, we used the R package \emph{dPeak}
+version 1.0.0, which is available from
+\url{http://dongjunchung.github.io/dpeak/}. For the GEM algorithm we
+used its Java implementation version 0.9, which is available from
+\url{http://cgs.csail.mit.edu/gem/}. For the MACE algorithm, we used
+ity Python implementation version 1.0, which is available from
+\url{http://dldcc-web.brc.bcm.edu/lilab/MACE/docs/html/}. Candidate
+regions for dPeak were identified for each replicate of ChIP-exo data
+using the MOSAiCS algorithm Kuan et al. \cite{mosaics} (one sample
+analysis using false discovery rate of 0.0001) implemented as an R
+package \emph{mosaics} 2.4 (available from \emph{Bioconductor}
+\url{https://www.bioconductor.org/packages/release/bioc/html/mosaics.html}). We
+further filtered out candidate region with average ChIP tag count less
+than 3,000 to avoid potential false positives based in exploratory
+analysis. These regions were also explicitly provided to the GEM
+algorithm as candidate regions. Default tuning parameters were used
+duging model fitting for all methods.
+
+\subsection{Local-NSC}
+\label{sec:localnsc}
+
+The local strand cross-correlation is defined as:
+
+\begin{align}
+  \mbox{local-NSC} &= \frac{max_{x_\delta} f(x_\delta)}{\sigma} \nonumber \\
+ y_\delta &= f(x_\delta) + \sigma \epsilon_\delta \nonumber
+\end{align}
+
+where $f$ is calculated by fitting a local polynomial regression model
+over the strand cross-correlation calculated using exclusively the
+fragments that overlap a pre-defined genomic region, $y_\delta$ is the
+correlation of the forward and backward coverages when one is shifted
+$x_\delta$ bp towards their 3' end.
+
+The local polynomial model was fitted using the \emph{loess} function
+from the R software version 3.2.1 using the default tunning parameters.
+
+%% not sure if here or before it would be nice to add an example of how this 
+%% method works. probably here
 
@@
 \nocite{maplot2}
+% \nocite{esl}
 
@@
 
+In figure \ref{fig:enrich2} it is shown the same relationshtip as in
+figure \ref{fig:enrich}, but considering only regions formed where the
+reads are being allocated in more than 10 (A) and 30 (B) unique
+positions respectively. Both plots show that the vertical arms formed
+by regions with low $\mbox{ARC}$ is formed by low complexity regions,
+that way suggesting that this segment correspond to the background of
+a ChIP-exo experiment.
+
 \begin{figure}[h!]
@@
     reads mapped to at most 10 and 30 positions respectively where not
-    considered.}
+    considered. $\mbox{ARC}$ is defined as the ratio of the nr. of
+    reads and the width of a region and $\mbox{URCR}$ is the ratio of
+    the number of unique position where the reads are being allocated
+    and the number of reads in a region.}
   \label{fig:enrich2}
@@
 
-\subsection{Local-NSC}
-\label{sec:localnsc}
-
-The local strand cross-correlaiton is defined as:
-
-\begin{align}
-  \mbox{local-NSC} &= \frac{max_{x_\delta} f(x_\delta)}{\sigma} \nonumber \\
- y_\delta &= f(x_\delta) + \sigma \epsilon_\delta \nonumber
-\end{align}
-
-where $f$ is calculated by fitting a non-parametric regression model
-over the strand cross-correlation calculated using exclusively the
-fragments that overlap a pre-defined genomic region, $y_\delta$ is the
-correlation of the forward and backward coverages when one is shifted
-$x_\delta$ bp towards their 3' end.
-
 \end{document}

### 2015-11-16 rscripts/bin_analysis.R

@@ -44,14 +44,15 @@ read_files <- mapply(filter_reads,
 reads <- mapply(create_reads,read_files,c(FALSE,TRUE,FALSE),SIMPLIFY = FALSE)
 
 
-pdf(file = file.path("figs/for_paper","ChIPseqPET_ChIPexo_tagCount_comparison.pdf"),width = 6,height = 6)
+pdf(file = file.path("figs/for_paper","ChIPseqPET_ChIPexo_tagCount_comparison.pdf"),width = 4,height = 4)
 p <- hexbin_plot(reads[[2]],reads[[1]],150,frag_len = 150)+xlab("ChIP-seq (PET) tag counts")+ylab("ChIP-exo tag counts")+theme_bw()
 dt <- copy(p$data)
 dt[,x := 1 + x]
 dt[,y := 1 + y]
 print(p)    
 print(p + xlim(0,500)+ylim(0,500)+coord_fixed())
-z <- p %+% dt +coord_fixed() + scale_x_log10() + scale_y_log10()+theme(plot.title = element_text(hjust = 0))+
+z <- p %+% dt +coord_fixed() + scale_x_log10() + scale_y_log10()+
+  theme(legend.position = "top",plot.title = element_text(hjust = 0))+
   ggtitle("A")
 print(z)
 dev.off()

### 2015-11-16 rscripts/peak_pair_assumption.R

@@ -51,7 +51,7 @@ setnames(sites,names(sites),c("seqnames","start","end","strand","id"))
 sites[,seqnames := "U00096"]
 sites[,strand := ifelse(strand == "F","+","-")]
 
-zites <- data.table(seqnames = "U00096",start = loc.prom, end = loc.prom , strand = "*")
+## zites <- data.table(seqnames = "U00096",start = loc.prom, end = loc.prom , strand = "*")
 
 ## ######################################################################################
 
@@ -60,56 +60,56 @@ bs_dir <- file.path(out_dir,"binding")
 check_create(bs_dir)
 
   load(file = file.path(bs_dir, "dpeakFit_exo_sig70aerobic.RData"))
-  load( file = file.path(bs_dir, "dpeakFit_pet1_sig70aerobic.RData"))
+  load( file = file.path(bs_dir, "dpeakFit_pet3_sig70aerobic.RData"))
   load( file = file.path(bs_dir, "dpeakFit_set_sig70aerobic.RData"))
 
-fit_comparison <- function(fit1 , fit2 , mm)
-{
-  mu1 <- lapply(fit1@optFit,function(x)x$mu)
-  mu2 <- lapply(fit2@optFit,function(x)x$mu)
+## fit_comparison <- function(fit1 , fit2 , mm)
+## {
+##   mu1 <- lapply(fit1@optFit,function(x)x$mu)
+##   mu2 <- lapply(fit2@optFit,function(x)x$mu)
 
-  G1 <- sapply(mu1,length)
-  G2 <- sapply(mu2,length)
+##   G1 <- sapply(mu1,length)
+##   G2 <- sapply(mu2,length)
 
-  print(identical(mu1,mu2))
-  print(table(G1))
-  print(table(G2))
-  print(table(G1 , G2))
+##   print(identical(mu1,mu2))
+##   print(table(G1))
+##   print(table(G2))
+##   print(table(G1 , G2))
 
-  min_dist <- function(m1,m2){
+##   min_dist <- function(m1,m2){
 
-    out <- sapply(m1 , function(x) min(abs(x - m2)))
-    out <- min(out)
+##     out <- sapply(m1 , function(x) min(abs(x - m2)))
+##     out <- min(out)
 
-    return(out)
+##     return(out)
 
-  }
+##   }
   
[…]

### 2015-11-16 rscripts/peak_pair_assumption_new.R

@@ -7,7 +7,7 @@ library(ChIPUtils)
 library(dpeak)
 
 exo_file <- "/p/keles/ChIPexo/volume3/LandickData/ChIPexo/edsn931_042814_qc.sorted.bam"
-pet_file1 <- "/p/keles/ChIPexo/volume3/LandickData/ChIPseq_PET/edsn790_042814_qc.filter.bam"
+pet_file <- "/p/keles/ChIPexo/volume3/LandickData/ChIPseq_PET/edsn790_042814_qc.filter.bam"
 set_file <- "/p/keles/ChIPexo/volume3/LandickData/ChIPseq_SET/edsn80_042814_qc.sorted.bam"
 
 frag_len <- 150
@@ -33,7 +33,7 @@ out_dir <- "/p/keles/ChIPexo/volume6/peak_pair"
 ### create bins
 
 
-bin_flag <- FALSE
+bin_flag <- TRUE
 
 bin_dir <- file.path(out_dir,"bins")
 check_create(bin_dir)
@@ -42,9 +42,7 @@ check_create(bin_dir)
 if(bin_flag){
   constructBins(exo_file,fileFormat = "bam",outfileLoc = bin_dir,PET = FALSE,
                 fragLen = frag_len,binSize = bin_size)
-  constructBins(pet_file1,fileFormat = "bam",outfileLoc = bin_dir, PET =  TRUE,
-               binSize = bin_size)
-  constructBins(pet_file2,fileFormat = "bam",outfileLoc = bin_dir, PET =  TRUE,
+  constructBins(pet_file,fileFormat = "bam",outfileLoc = bin_dir, PET =  TRUE,
                binSize = bin_size)
   constructBins(set_file,fileFormat = "bam",outfileLoc = bin_dir, PET =  FALSE,
                fragLen = frag_len , binSize = bin_size)  
@@ -52,31 +50,25 @@ if(bin_flag){
 
 ### call peaks
 
-
-
-peak_flag <- FALSE
+peak_flag <- TRUE
 
 if(peak_flag){
   exo_bin <- file.path(bin_dir,paste0(basename(exo_file),"_fragL",frag_len,"_bin",bin_size,".txt"))
-  pet_bin1 <- file.path(bin_dir,paste0(basename(pet_file1),"_bin",bin_size,".txt"))
-  pet_bin2 <- file.path(bin_dir,paste0(basename(pet_file2),"_bin",bin_size,".txt"))
[…]

### 2015-11-16 rscripts/peak_pair_assumption_old_data.R

@@ -1,0 +1,358 @@
+rm( list=ls() )
+graphics.off()
+library(dpeak)
+library(hexbin)
+library(IRanges)
+library(data.table)
+library(ggplot2)
+library(RColorBrewer)
+library(reshape2)
+
+dir.out <- "/p/keles/Dongjun/ChIP-exo/paper/figures/"
+
+# load peak lists
+
+peak.R1 <- read.table( "/NO_BACKUP/KelesGroup_DongjunChung/ChIP-exo/Landick_ChIP-exo/analysis_sigma70/analysis_sigma70_mosaics/ChIP-exo_sigma70_exp_phase_R1_mosaics_1S_peak.bed", sep="\t" )
+peak.R2 <- read.table( "/NO_BACKUP/KelesGroup_DongjunChung/ChIP-exo/Landick_ChIP-exo/analysis_sigma70/analysis_sigma70_mosaics/ChIP-exo_sigma70_exp_phase_R2_mosaics_1S_peak.bed", sep="\t" )
+
+# load predictions
+
+load( "/NO_BACKUP/KelesGroup_DongjunChung/ChIP-exo/Landick_ChIP-exo/analysis_sigma70/analysis_sigma70_R1/ChIP-exo_sigma70_exp_phase_dpeak_fit.RData" )
+fit.exo1 <- fitSET
+rm(fitSET)
+load( "/NO_BACKUP/KelesGroup_DongjunChung/ChIP-exo/Landick_ChIP-exo/analysis_sigma70/analysis_sigma70_R2/ChIP-exo_sigma70_exp_phase_dpeak_fit.RData" )
+fit.exo2 <- fitSET
+rm(fitSET)
+
+load( "/NO_BACKUP/KelesGroup_DongjunChung/ChIP-exo/Landick_ChIP-exo/chipseq/Kiley_PET/sigma70+O2/mosaics_PET/sigma70+O2_PET_fit.RData" )
+fit.PET <- fitPET
+rm(fitPET)
+
+load( "/NO_BACKUP/KelesGroup_DongjunChung/ChIP-exo/Landick_ChIP-exo/chipseq/Kiley_SET/sigma70+O2/mosaics_SET/fit_sigma70+O2.RData" )
+fit.SET <- fitSET
+rm(fitSET)
+
+# extract delta & sigma
+
+delta.exo1 <- unlist(fit.exo1@optDelta)
[…]

### 2015-11-16 rscripts/peak_pair_real.R

@@ -1,0 +1,131 @@
+
+rm(list = ls())
+
+library(data.table)
+library(GenomicRanges)
+library(GenomicAlignments)
+library(viridis)
+library(scales)
+library(ggplot2)
+library(ChIPUtils)
+
+base_dir <- "/p/keles/ChIPexo/volume3/LandickData"
+
+exo_dir <- file.path(base_dir,"ChIPexo")
+pet_dir <- file.path(base_dir,"ChIPseq_PET")
+set_dir <- file.path(base_dir,"ChIPseq_PET")
+
+
+## peaks
+exo_peaks <- "/p/keles/ChIPexo/volume6/peak_pair/peaks/edsn931_042814_peaks.txt"
+pet_peaks <- "/p/keles/ChIPexo/volume6/peak_pair/peaks/edsn790_042814_filter_peaks.txt"
+set_peaks <- "/p/keles/ChIPexo/volume6/peak_pair/peaks/edsn80_042814_peaks.txt"
+
+exo_peaks <- data.table(read.table(exo_peaks))
+pet_peaks <- data.table(read.table(pet_peaks))
+set_peaks <- data.table(read.table(set_peaks))
+
+exo_gr <- exo_peaks[,1:3,with = FALSE]
+pet_gr <- pet_peaks[,1:3,with = FALSE]
+set_gr <- set_peaks[,1:3,with = FALSE]
+
+setnames(exo_gr,names(exo_gr),c("seqnames","start","end"))
+setnames(pet_gr,names(pet_gr),c("seqnames","start","end"))
+setnames(set_gr,names(set_gr),c("seqnames","start","end"))
+
+exo_gr <- dt2gr(exo_gr)
+pet_gr <- dt2gr(pet_gr)
+set_gr <- dt2gr(set_gr)
+
+common <- subsetByOverlaps(exo_gr, subsetByOverlaps(pet_gr,set_gr))
+
+exo_file <- "/p/keles/ChIPexo/volume3/LandickData/ChIPexo/edsn931_042814_qc.sorted.bam"
+pet_file <- "/p/keles/ChIPexo/volume3/LandickData/ChIPseq_PET/edsn790_042814_qc.filter.bam"
+set_file <- "/p/keles/ChIPexo/volume3/LandickData/ChIPseq_SET/edsn80_042814_qc.sorted.bam"
+ 
+prm <- NULL
+
+exo <- readGAlignments(exo_file, param = prm)
[…]

### 2015-11-16 rscripts/resolution_sensitivity_R1_old_data.R

@@ -1,0 +1,322 @@
+rm( list=ls() )
+graphics.off()
+library(dpeak)
+library(IRanges)
+library(data.table)
+library(ggplot2)
+library(reshape2)
+library(RColorBrewer)
+
+# options
+
+extension <- 20
+
+#dir.peak <- "/scratch/deconvolution/comparison_qSET/"
+dir.out <- "/p/keles/Dongjun/ChIP-exo/paper/figures/"
+
+# read promoter regions for sigma70 from regulonDB
+
+prom <- read.table(
+    "/NO_BACKUP/KelesGroup_DongjunChung/ChIP-exo/Landick_ChIP-exo/database_regulonDB/PromoterSigma70Set.txt",
+  skip=33, sep="\t")
+loc.prom <- prom[,4]
+
+# load peak region & motif
+
+peakList.exo <- read.table( 
+    "/NO_BACKUP/KelesGroup_DongjunChung/ChIP-exo/Landick_ChIP-exo/analysis_sigma70/analysis_sigma70_mosaics/ChIP-exo_sigma70_exp_phase_R1_mosaics_1S_peak.bed",
+    sep="\t", stringsAsFactors=FALSE )
+
+peakList.PET <- read.table( 
+    "/NO_BACKUP/KelesGroup_DongjunChung/ChIP-exo/Landick_ChIP-exo/chipseq/Kiley_PET/sigma70+O2/mosaics_PET/peaklist_IO.bed",
+    sep="\t", stringsAsFactors=FALSE, skip=1 )
+
+peakList.SET <- read.table( 
+    "/NO_BACKUP/KelesGroup_DongjunChung/ChIP-exo/Landick_ChIP-exo/chipseq/Kiley_SET/sigma70+O2/mosaics_SET/peaklist_IO.bed",
+    sep="\t", stringsAsFactors=FALSE, skip=1 )
+
+# load predictions
+
+load( "/NO_BACKUP/KelesGroup_DongjunChung/ChIP-exo/Landick_ChIP-exo/analysis_sigma70/analysis_sigma70_R1/ChIP-exo_sigma70_exp_phase_dpeak_fit.RData" )
+#load( "/NO_BACKUP/KelesGroup_DongjunChung/ChIP-exo/Landick_ChIP-exo/analysis_sigma70/analysis_sigma70_R2/ChIP-exo_sigma70_exp_phase_dpeak_fit.RData" )
+fit.exo <- fitSET
+rm(fitSET)
+
[…]

### 2015-11-16 rscripts/saturation_analysis_old.R

@@ -1,0 +1,352 @@
+rm( list=ls() )
+graphics.off()
+library(mosaics)
+library(IRanges)
+library(data.table)
+library(ggplot2)
+library(reshape2)
+library(RColorBrewer)
+
+# option
+
+maxM <- 10
+ext_iden <- 20
+max_dist <- 1000
+seedset <- c( 12345, 23456, 34567, 45678, 56789 )
+ntop <- 500
+
+dir_out <- "/p/keles/Dongjun/ChIP-exo/paper/figures/"
+
+dir_mosaics_exo_R1 <- "/NO_BACKUP/KelesGroup_DongjunChung/ChIP-exo/Landick_ChIP-exo/saturation_analysis/mosaics_ChIPexo_R1/"
+dir_mosaics_exo_R2 <- "/NO_BACKUP/KelesGroup_DongjunChung/ChIP-exo/Landick_ChIP-exo/saturation_analysis/mosaics_ChIPexo_R2/"
+dir_mosaics_seq_PET <- "/NO_BACKUP/KelesGroup_DongjunChung/ChIP-exo/Landick_ChIP-exo/saturation_analysis/mosaics_ChIPseq_PET/"
+dir_mosaics_seq_SET <- "/NO_BACKUP/KelesGroup_DongjunChung/ChIP-exo/Landick_ChIP-exo/saturation_analysis/mosaics_ChIPseq_SET/"
+
+dir_dpeak_exo_R1 <- "/NO_BACKUP/KelesGroup_DongjunChung/ChIP-exo/Landick_ChIP-exo/saturation_analysis/dpeak_ChIPexo_R1/"
+dir_dpeak_seq_PET <- "/NO_BACKUP/KelesGroup_DongjunChung/ChIP-exo/Landick_ChIP-exo/saturation_analysis/dpeak_ChIPseq_PET/"
+dir_dpeak_seq_SET <- "/NO_BACKUP/KelesGroup_DongjunChung/ChIP-exo/Landick_ChIP-exo/saturation_analysis/dpeak_ChIPseq_SET/"
+
+# read promoter regions for sigma70 from regulonDB
+
+prom <- read.table(
+    "/NO_BACKUP/KelesGroup_DongjunChung/ChIP-exo/Landick_ChIP-exo/database_regulonDB/PromoterSigma70Set.txt",
+  skip=33, sep="\t")
+loc.prom <- prom[,4]
+
+# read peak summary
+
[…]

### 2015-11-16 rscripts/sig70_forward_strand_ratio_analysis.R

@@ -1,0 +1,9 @@
+
+rm(list = ls())
+
+library(ggplot2)
+library(data.table)
+library(ChIPUtils)
+library(reshape2)
+library(scales)
+library(RColorBrewer)

### 2015-11-17 rmarkdown/chip_exo_paper.Rnw
Added section to paper on peak shape comparison between chip seq and chip exo

@@
     distribution, where several ChIP-exo background regions were
-    empty. B) Density of the proportion of the ratio of forward strand
-    reads in a ChIP-exo and ChIP-Seq SET data.}
+    empty. B) In ChIP-exo data, strands of reads were significantly
+    less balanced in the regions with potential binding sites compared
+    to SET ChIP-Seq data. }
   \label{fig:comp}
@@
 
-\begin{figure}[h]
+<<reso_param, include = FALSE , eval  = TRUE , echo = FALSE>>=
+
+  ext <- 20
+
+@
+
+Figure \ref{fig:reso_all} shows different comparisons among ChIP-exo,
+PET ChIP-Seq and SET ChIP-Seq. A regulonDB annotation was considered
+identified if the distance between it and dPeak binding site estimate
+was at most of \Sexpr{ext} bp. That way, the sensitivity is defined as
+the proportion of regulonDB annotations identified in a peak. In
+figure \ref{fig:reso_all}A can be seen that the sensitivity of all
+protocols increases as the average distance between binding sites
+does. Despite that when the binding events in a peak are closer to
+each other both ChIP-exo and PET ChIP-Seq are comparable, as the
+distance increases ChIP-exo identifies a higher proportion of the
+regulonDB annotations; additionally SET ChIP-Seq is significantly less
+sensible than both ChIP-exo and PET ChIP-Seq. In figure
+\ref{fig:reso_all}B the distance between a regulonDB annotation to its
+closest prediction is compared for ChIP-exo, PET and SET ChIP-Seq;
+while the first two are comparable, both outperform SET
+ChIP-Seq.
+
+In figures \ref{fig:reso_all}C and \ref{fig:reso_all}D, we observe the
+behaviour of dPeak's estimated parameters for single end data (both
+ChIP-exo and ChIP-Seq). The $\delta$ parameter measures the distance
+from the 5' end of the fragment to its respective binding event and
+the $\sigma$ parameter indicater the fragments distribution
+variability around their respective binding sites (more information in
+\cite{dpeak}). For both parameters, it can be seen that the ChIP-exo
+estimates are smaller than the SET ChIP-Seq estimates in average,
+which agrees with the fact that the sample reads are allocated more
+tihgtly around the binding events in ChIP-exo data. Hence we can
+conclude that the peaks shape is quite different between ChIP-exo and
+SET ChIP-Seq.
+
+\begin{figure}[h!]
   \centering
-  \includegraphics[width = .4\textwidth]{../figs/for_paper/sensitivity_exo_olda_data.pdf}
-  \includegraphics[width = .4\textwidth]{../figs/for_paper/resolution_by_dataset_old_data.pdf}
+  \includegraphics[width = .46\textwidth]{../figs/for_paper/sensitivity_exo_olda_data.pdf}
+  \includegraphics[width = .46\textwidth]{../figs/for_paper/resolution_by_dataset_old_data.pdf}
+   \includegraphics[width = .46\textwidth,page = 1]{../figs/for_paper/sigma_delta_old_densities.pdf}
+   \includegraphics[width = .46\textwidth,page = 2]{../figs/for_paper/sigma_delta_old_densities.pdf}
+
    \caption{Comparison of (A) sensitivity and (B) resolution between
@@
      data. Resolution is defined as the distance between regulonDB
-    annotation and its closest prediction.}
+     annotation and its closest prediction. $\quad$ C) $\delta$ parameter in
+     dPeak measures average distance of the reads to their respective
+     binding site. In ChIP-exo data, reads were located much closer to
+     the binding site than in SET ChIP-Seq. D) $\sigma$ parameter
+     measure the dispersion of reads around each binding site. In
+     ChIP-exo data, reads showed less variation around the their
+     respective binding sites compared to SET ChIP-Seq.}
   \label{fig:reso_all}

### 2015-11-17 rmarkdown/chip_exo_paper.tex
Added section to paper on peak shape comparison between chip seq and chip exo

@@
 
-\begin{figure}[h]
+Figure \ref{fig:reso_all} shows different comparisons among ChIP-exo,
+PET ChIP-Seq and SET ChIP-Seq. A regulonDB annotation was considered
+identified if the distance between it and dPeak binding site estimate
+was at most of 20 bp. That way, the sensitivity is defined as
+the proportion of regulonDB annotations identified in a peak. In
+figure \ref{fig:reso_all}A can be seen that the sensitivity of all
+protocols increases as the average distance between binding sites
+does. Despite that when the binding events in a peak are closer to
+each other both ChIP-exo and PET ChIP-Seq are comparable, as the
+distance increases ChIP-exo identifies a higher proportion of the
+regulonDB annotations; additionally SET ChIP-Seq is significantly less
+sensible than both ChIP-exo and PET ChIP-Seq. In figure
+\ref{fig:reso_all}B the distance between a regulonDB annotation to its
+closest prediction is compared for ChIP-exo, PET and SET ChIP-Seq;
+while the first two are comparable, both outperform SET
+ChIP-Seq.
+
+In figures \ref{fig:reso_all}C and \ref{fig:reso_all}D, we observe the
+behaviour of dPeak's estimated parameters for single end data (both
+ChIP-exo and ChIP-Seq). The $\delta$ parameter measures the distance
+from the 5' end of the fragment to its respective binding event and
+the $\sigma$ parameter indicater the fragments distribution
+variability around their respective binding sites (more information in
+\cite{dpeak}). For both parameters, it can be seen that the ChIP-exo
+estimates are smaller than the SET ChIP-Seq estimates in average,
+which agrees with the fact that the sample reads are allocated more
+tihgtly around the binding events in ChIP-exo data. Hence we can
+conclude that the peaks shape is quite different between ChIP-exo and
+SET ChIP-Seq.
+
+\begin{figure}[h!]
   \centering
-  \includegraphics[width = .4\textwidth]{../figs/for_paper/sensitivity_exo_olda_data.pdf}
-  \includegraphics[width = .4\textwidth]{../figs/for_paper/resolution_by_dataset_old_data.pdf}
+  \includegraphics[width = .46\textwidth]{../figs/for_paper/sensitivity_exo_olda_data.pdf}
+  \includegraphics[width = .46\textwidth]{../figs/for_paper/resolution_by_dataset_old_data.pdf}
+   \includegraphics[width = .46\textwidth,page = 1]{../figs/for_paper/sigma_delta_old_densities.pdf}
+   \includegraphics[width = .46\textwidth,page = 2]{../figs/for_paper/sigma_delta_old_densities.pdf}
+
    \caption{Comparison of (A) sensitivity and (B) resolution between

### 2015-11-17 rmarkdown/chip_exo_paper.R

@@ -9,6 +9,11 @@
   frag_len <- 150
   
 
+## ----reso_param, include = FALSE , eval  = TRUE , echo = FALSE-----------
+
+  ext <- 20
+
+
 ## ----design_param , include = FALSE , echo = FALSE , eval = TRUE---------
 
   K <- 10

### 2015-11-17 rscripts/peak_pair_assumption_old_data.R

@@ -168,23 +168,30 @@ dt[ , Dataset := plyr::mapvalues( Dataset ,
         from = c("ChIP-exo","ChIP-seq (PET)","ChIP-seq (SET)"),
         to = c("exo","pet","set"))]
 
-brewer.pal(3,"Set1")
-
 p1 = ggplot(dt,aes(delta,colour = Dataset))+geom_density(size=1)+facet_grid(Dataset~.)+
-  theme(legend.position = "none")+scale_color_manual( values = brewer.pal(3,"Set1")[c(1,3)])+
-  ylab("")+xlab(expression(delta))
-u = print(p1)
-dev.off()
-
+  theme_bw()+theme(legend.position = "none",plot.title = element_text(hjust = 0))+
+  scale_color_manual( values = brewer.pal(3,"Set1")[c(1,3)])+
+  ylab("")+xlab("delta estimate")
 dt1 = data.table(sigma = sigma.exo1)
 dt1[,Dataset :="ChIP-exo"]
 dt2 = data.table(sigma = sigma.SET)
 dt2[,Dataset:= "ChIP-seq (SET)"]
 dt = rbind(dt1,dt2)
 
-p1 = ggplot(dt,aes(sigma,colour = Dataset))+geom_density(size=1)+facet_grid(Dataset~.)+
-  theme(legend.position = "none")+scale_color_brewer(palette = "Dark2")+ylab("")+
-  xlab(expression(sigma))
+dt[ , Dataset := plyr::mapvalues( Dataset ,
+        from = c("ChIP-exo","ChIP-seq (PET)","ChIP-seq (SET)"),
+        to = c("exo","pet","set"))]
+
+p2 = ggplot(dt,aes(sigma,colour = Dataset))+geom_density(size=1)+facet_grid(Dataset~.)+
+  theme_bw()+theme(legend.position = "none",plot.title = element_text(hjust = 0))+
+  scale_color_manual(values = brewer.pal(3,"Set1")[c(1,3)])+ylab("")+
+  xlab("sigma estimate")
+
+pdf(file = "figs/for_paper/sigma_delta_old_densities.pdf",width = 4 ,height  = 4)
+print(p1 + ggtitle("C"))
+print(p2 + ggtitle("D"))
+dev.off()
+
 pdf(file = file.path(dr,"sigma_estimate_exo_seq_R1.pdf"))
 u = print(p1)
 dev.off()

### 2015-11-17 rscripts/sig70_forward_strand_ratio_analysis.R

@@ -7,3 +7,87 @@ library(ChIPUtils)
 library(reshape2)
 library(scales)
 library(RColorBrewer)
+library(grid)
+library(gridExtra)
+
+## load data from pipeline
+
+dr <- "../Analysis/Landick/stat-vs-exp/Sig70-exp-rep2/data"
+
+files <- list.files(dr)
+
+load( file.path(dr, files[grep("summ",files)]))
+
+stats <- mcmapply(function(x,y)x[,chr := y],summary_stats,names(summary_stats),SIMPLIFY = FALSE,mc.cores = 20)
+stats <- do.call(rbind,stats)
+
+forward_strand_ratio_plot <- function(stat,probs = c(0,.25,.5,.75,1),values = 1:750, mc = 8)
+{
+  stopifnot(is.numeric(values))
+  stopifnot(length(probs) <= 9)
+  stopifnot(length(values) > 1)
+  to_use <- stat[,.(depth, prob)]
+  rows <- mclapply(values,function(i)to_use[depth > i,quantile(prob, probs)],mc.cores = mc)
+  rows <- mapply(function(x,i)data.table(depth = i,quantiles = probs,fsr = x ),rows,values,SIMPLIFY = FALSE)
+  rows <- do.call(rbind,rows)
+  rows[,quantiles := factor(quantiles,levels = sort(probs))]
+  p <- ggplot(rows,aes(depth , fsr , colour = quantiles))+geom_line(size = 1)+
+    scale_color_brewer(name = "quantiles",palette = "Dark2")+ylim(0,1)+xlim(1,max(values))+theme_bw()+
+    theme(legend.position = "top",plot.title = element_text(hjust =0))+
+    ylab("fwd strand ratio")+xlab("least amount of fragments in region")
+  return(p)
+}
+
+
+label_plot <- function(stat,values = 1:750,mc = 8,prop = FALSE)
+{
+  stopifnot(is.numeric(values))
+  stopifnot(length(values) > 1)
+
+  to_use <- stat[,.(depth,label)]
+  setkey(to_use,label)
+  fwd <- to_use["fwd"]
+  fwd <- do.call(c,mclapply(values,function(i)nrow(fwd[depth > i]),mc.cores = mc))
+  bwd <- to_use["bwd"]
+  bwd <- do.call(c,mclapply(values,function(i)nrow(bwd[depth > i]),mc.cores = mc))
+  both <- to_use["both"]
[…]

### 2015-11-17 rmarkdown/chip_exo_paper.Rnw
added section on strand imbalance

@@
 
+In order to asses the strand imbalance we create the following
+visualization in figure \ref{fig:strand}. A) presents the FSR's
+behaviour as lower depth regions are being filtered out. while B)
+shows which percentage of the regions are composed by fragments of
+both strand or only one. Roughly speaking in a lower quality dataset
+there is going to be a higher amount of single stranded-regions. To
+asses this, we divided the ChIP-exo regions into two sets, the oens
+that overlap with a known set of peaks and the opposite and performed
+$\chi^2$ tests.
+
+\begin{figure}[h!]
+  \centering  
+  \includegraphics[width = .8\textwidth]{../figs/fsr/draft_fsr_plot.pdf}
+  \includegraphics[width = .8\textwidth]{../figs/fsr/draft_label_prop_plot.pdf}
+  \caption{ABCD}
+  \label{fig:strand}
+\end{figure}
+
 \subsection{Comparison with ChIP-Seq data}
@@
 %  We obtained esentially the same conclusion from the
-% ChIP-exo data for CTCF factir from human sample Rhee and Pugh
+% ChIP-exo data for CTCF factor from human sample Rhee and Pugh
 % (\cite{exo1}) as well.
@@
 
+\section{Conclusions}
+\label{sec:conclusions}
+
 \newpage

### 2015-11-17 rmarkdown/chip_exo_paper.Rnw
Corrected orthography with ispell

@@
 \usepackage{hyperref}
-
-\usepackage{hyperref}
 \hypersetup{
@@
 \hypersetup{linkcolor=black}
-\hypersetup{linkcolor=black}
-\hypersetup{linkcolor=black}
+% \hypersetup{linkcolor=black}
+% \hypersetup{linkcolor=black}
 
@@
   following characteristics of ChIP-exo data. First, Background of
-  ChIP-exo data is qute different from that of ChIP-Seq data. However,
+  ChIP-exo data is quite different from that of ChIP-Seq data. However,
   sequence biases inherently present in ChIP-Seq data still exist in
@@
 
+\listoffigures
+
+\newpage
+
 \section{Background}
@@
 using random fragmentation of DNA and a protein-specific
-antibody. Then, exonuclease is introducted to trim 5' end of each DNA
+antibody. Then, exonuclease is introduced to trim 5' end of each DNA
 fragment to a fixed distance from the bound protein. As a result,
@@
 ChIP-exo data, choice of optimal statistical methods, and
-determination of optimal sequencing deoth. However, currently, the
+determination of optimal sequencing depth. However, currently, the
 number of available ChIP-exo data is still limited and their
@@
 \cite{exo_review}. Hence, most of current guidelines \cite{encode_qc}
-may not be aplicable on ChIP-exo, additionally to our knowledge there
+may not be applicable on ChIP-exo, additionally to our knowledge there
 are not established quality control pipelines for ChIP-exo. To address
@@
 
-In order to archieve the potential benefits of ChIP-exo on protein
+In order to archive the potential benefits of ChIP-exo on protein
 binding site identification, it is critical to understand which are
@@
 analysis methods, are also based on this peak pair
-assumption. However, appropriatness of such assumption was not fully
+assumption. However, appropriateness of such assumption was not fully
 evaluated in the literature yet.  Furthermore, it is still unknown
-which factors could affect protein binding site ientification using
+which factors could affect protein binding site identification using
 ChIP-exo data. In order to address this problem, we investigated
@@
 report only one ``mode'' or ``predicted binding location'' per
-peak. Hence, these methods are not appropiate to evaluate the
+peak. Hence, these methods are not appropriate to evaluate the
 potential of ChIP-exo data for high resolution identification of
 protein binding sites. More recently, deconvolution algorithms such as
-CSDeconv \cite{csdeconv}, GEM \cite{gem} (an improved version of
+Deconvolution \cite{csdeconv}, GEM \cite{gem} (an improved version of
 \cite{gps}) and PICS \cite{pics} have been proposed to identify
@@
 data in a unified framework and as a result, currently available
-methods are not appropiate for fair comparison between ChIP-exo and
+methods are not appropriate for fair comparison between ChIP-exo and
 ChIP-Seq. To address these limitations, we developed an improved of
@@
 \cite{htseq}. Additionally, we calculated the FSR to compare the
-number of fragments in each strand. A few conclussions that we may get
+number of fragments in each strand. A few conclusions that we may get
 by observing the PBC in table \ref{tab:qcbase} are all the samples
@@
     $\sigma^{70}$ samples. PBC stands PCR-bottleneck coefficient, SSD
-    for standarized standard deviation, FSR for forward strand ratio,
+    for standardized standard deviation, FSR for forward strand ratio,
     NSC for normalized strand cross-correlation coefficient. We
-    ommited the RSC (relative strand cross-correlation since ChIP-exo
-    experiments don't usually have an acomnying input file)}
+    omitted the RSC (relative strand cross-correlation since ChIP-exo
+    experiments don't usually have an accompanying input file)}
 
@@
 pipeline. Which basically partitions genome by keeping the
-non-diggested ChIP-exo regions. Then, for each region it calculates a
+non-digested ChIP-exo regions. Then, for each region it calculates a
 series of independent statistics. Finally, it creates several
@@
 exonuclease enzyme, therefore to determine the sample's quality is
-necesary to address the balance between the enrichment and library
+necessary to address the balance between the enrichment and library
 complexity. To diagnose this, we considered the Average Read Coverage
@@
 strong arms in each panel: The first one corresponds to regions with
-low $\mbox{ARC}$ values and $\mbox{URCR}$ values that can vary accross
+low $\mbox{ARC}$ values and $\mbox{URCR}$ values that can vary across
 the whole $[0,1]$ interval, while the second one shows a decreasing
@@
 
-The strand imbalance assesment is based in the observation that the
+The strand imbalance assessment is based in the observation that the
 enriched regions usually have a higher concentration of fragments,
@@
 visualization in figure \ref{fig:strand}. A) presents the FSR's
-behaviour as lower depth regions are being filtered out. while B)
+behavior as lower depth regions are being filtered out. while B)
 shows which percentage of the regions are composed by fragments of
@@
 there is going to be a higher amount of single stranded-regions. To
-asses this, we divided the ChIP-exo regions into two sets, the oens
+asses this, we divided the ChIP-exo regions into two sets, the ones
 that overlap with a known set of peaks and the opposite and performed
@@
 distribution of signal and background between ChIP-exo and ChIP-Seq
-data, we calculated ChIP tag counts accross the genome by counting the
+data, we calculated ChIP tag counts across the genome by counting the
 number of reads mapping to each of \Sexpr{bin_size} non-overlapping
@@
 counts (Figure \ref{fig:comp}A). This implies that signals for
-pontential binding sites are well reproducible between ChIP-exo and
+potential binding sites are well reproducible between ChIP-exo and
 ChIP-Seq data. On the other hand, there was clear difference in the
@@
 located in the other site of the binding site. In order to evaluate
-this assumptiom, we reviewed the proportion of reads in the forward
-streand in candidate regions (i.e. regions with at least one binding
+this assumption, we reviewed the proportion of reads in the forward
+strand in candidate regions (i.e. regions with at least one binding
 site) in $\sigma^{70}$ ChIP-exo data. We found that strands of reads
@@
 these regions with potential binding sites (Fig. \ref{fig:comp}B) and
-htis indicates that the peak pair assumptuon might not hold in real
+this indicates that the peak pair assumption might not hold in real
 ChIP-exo data.
@@
     PET ChIP-Seq data in the region with high ChIP tag counts. In
-    contrast, there wer clear differences in their background
+    contrast, there were clear differences in their background
     distribution, where several ChIP-exo background regions were
@@
 In figures \ref{fig:reso_all}C and \ref{fig:reso_all}D, we observe the
-behaviour of dPeak's estimated parameters for single end data (both
+behavior of dPeak's estimated parameters for single end data (both
 ChIP-exo and ChIP-Seq). The $\delta$ parameter measures the distance
 from the 5' end of the fragment to its respective binding event and
-the $\sigma$ parameter indicater the fragments distribution
+the $\sigma$ parameter indicate the fragments distribution
 variability around their respective binding sites (more information in
@@
 which agrees with the fact that the sample reads are allocated more
-tihgtly around the binding events in ChIP-exo data. Hence we can
+tightly around the binding events in ChIP-exo data. Hence we can
 conclude that the peaks shape is quite different between ChIP-exo and
@@
      ChIP-exo and ChIP-Seq data. Sensitivity is defined as the
-     propotion of regulonDB annotations identificed using each
+     proportion of regulonDB annotations identified using each
      data. Resolution is defined as the distance between regulonDB
@@
 
-\subsection{Recomendations for the design of ChIP-exo experiments}
+\subsection{Recommendations for the design of ChIP-exo experiments}
 \label{sec:reco}
@@
 
-We sampled a fixed amoung of fragments for each of the ChIP-exo, PET
+We sampled a fixed among of fragments for each of the ChIP-exo, PET
 ChIP-Seq and SET ChIP-Seq datasets of the $\sigma^{70}$ sample in
@@
 
-Figure \ref{fig:design} shows the behaviour of each data type when
+Figure \ref{fig:design} shows the behavior of each data type when
 their depth is fixed. It is remarkable that even when the number of
-candiate peaks or the number of predicted events in quite lower than
+candidate peaks or the number of predicted events in quite lower than
 for both ChIP-Seq cases, it outperforms them in the number of
@@
 ChIP-exo less false positive peaks are being called and that when the
-targets are being identified, dpeak is able to estimate binding
+targets are being identified, dPeak is able to estimate binding
 locations closer to the actual true cases. Additionally, we can see
@@
 \url{http://cgs.csail.mit.edu/gem/}. For the MACE algorithm, we used
-ity Python implementation version 1.0, which is available from
-\url{http://dldcc-web.brc.bcm.edu/lilab/MACE/docs/html/}. Candidate
+its Python implementation version 1.0, which is available from
+\url{http://dldcc-web.brc.bcm.edu/lilab/MACE/docs/html/}. For the
+PEAKZILLA algorithm, we used its Python implementation, which is
+available in \url{http://www.starklab.org/data/peakzilla/}. Candidate
 regions for dPeak were identified for each replicate of ChIP-exo data
@@
 algorithm as candidate regions. Default tuning parameters were used
-duging model fitting for all methods.
+during model fitting for all methods.
+
+% And for
+% the CexoR algorithm, we used its R implementation version 1.8.0 which
+% is available from
+% \url{http://bioconductor.org/packages/release/bioc/html/CexoR.html}
 
@@
 The local polynomial model was fitted using the \emph{loess} function
-from the R software version 3.2.1 using the default tunning parameters.
+from the R software version 3.2.1 using the default tuning parameters.
 
@@
 
-In figure \ref{fig:enrich2} it is shown the same relationshtip as in
+In figure \ref{fig:enrich2} it is shown the same relationship as in
 figure \ref{fig:enrich}, but considering only regions formed where the

### 2015-11-17 rmarkdown/chip_exo_paper.tex
Corrected orthography with ispell

@@
 distribution of signal and background between ChIP-exo and ChIP-Seq
-data, we calculated ChIP tag counts accross the genome by counting the
+data, we calculated ChIP tag counts across the genome by counting the
 number of reads mapping to each of 150 non-overlapping
@@
 
-\subsection{Recomendations for the design of ChIP-exo experiments}
+\subsection{Recommendations for the design of ChIP-exo experiments}
 \label{sec:reco}
 
-We sampled a fixed amoung of fragments for each of the ChIP-exo, PET
+We sampled a fixed among of fragments for each of the ChIP-exo, PET
 ChIP-Seq and SET ChIP-Seq datasets of the $\sigma^{70}$ sample in

### 2015-11-19 rmarkdown/chip_exo_paper.Rnw
Added plots on enrichment and strand imbalance

@@
   omitted the RSC (relative strand cross-correlation since ChIP-exo
-    experiments don't usually have an accompanying input file)}
+  experiments don't usually have an accompanying input file) 
+  [note: I need to change the table, due the alignment issue]}
 
@@
 
-Figure \ref{fig:enrich}A shows hexbin plots illustrating the
-interaction between these two quantities. Notice how there are two
-strong arms in each panel: The first one corresponds to regions with
-low $\mbox{ARC}$ values and $\mbox{URCR}$ values that can vary across
-the whole $[0,1]$ interval, while the second one shows a decreasing
-trend in $\mbox{URCR}$ as $\mbox{ARC}$ increases. When an experiment
-shows a higher degree of enrichment, then the separation of this two
-arms is more noticeable, since the second arm corresponds to possibly
-enriched regions (\ref{fig:enrich2}A and \ref{fig:enrich2}B).
+Using the mouse-FoxA1 experiment from \cite{exoillumina}, the
+relationship between this two quantities, library complexty and sample
+enrichment was explored. Figure \ref{fig:enrich}A shows hexbin plots
+illustrating the interaction between these two quantities. Notice how
+there are two strong arms in each panel: The first one corresponds to
+regions with low $\mbox{ARC}$ values and $\mbox{URCR}$ values that can
+vary across the whole $[0,1]$ interval, while the second one shows a
+decreasing trend in $\mbox{URCR}$ as $\mbox{ARC}$ increases. When an
+experiment shows a higher degree of enrichment, then the separation of
+this two arms is more noticeable, since the second arm corresponds to
+possibly enriched regions (\ref{fig:enrich2}A and \ref{fig:enrich2}B).
 
@@
 
-  nstrata <- 100
+  nstrata <- 400
+  high <- 100
+  med <- 50
+  low <- 20
 
@@
 
-Figure \ref{fig:enrich}B shows boxplots of the $\mbox{local-NSC}$
-stratified by the number of reads mapped exactly to one position for
-all $\sigma^{70}$ ChIP-exo experiments; for each stratum, we sampled
-\Sexpr{nstrata} regions that whenever possible (in the opposite case,
-all the regions in the category were considered). There is an
-increasing trend as the number of unique position increases, which
-means that this is an effective indicator of library complexity since
-regions where the reads are mapped to more positions are usually more
-complex.
-
-%% perhaps do two other slots about how this works... a good experiment vs a bad experiment
-%% of the same region
+Figure \ref{fig:enrich}B shows boxplots of the $\mbox{local-NSC}$ on
+regions common to the 3 replicates and stratified by the number of
+reads mapped exactly to one position for all $\sigma^{70}$ ChIP-exo
+experiments; The high stratum is defined by regions consising of more
+than \Sexpr{high} unique positions, the medium stratum for regions
+where the number of unique position is on the
+(\Sexpr{med},\Sexpr{high}) range and the low stratum in the
+(\Sexpr{low},\Sexpr{med}) range. For each stratum and biological
+replicate, we sampled \Sexpr{nstrata} regions whenever possible (in
+the opposite case, all the regions in the category were considered)
+and calculated the $\mbox{local-NSC}$ for those regions. For all three
+replicates, it is shown an increasing trend as the number of unique
+position increases, which means that the $\mbox{local-NSC}$ is an
+effective indicator of library complexity. Figure \ref{fig:enrich}C
+shows the $\mbox{local-NSC}$ coefficient calculated for the same set
+of regions, stratified by the number of unique positions in the
+regions of ERR336942 using the same criteria as before. In the high
+panel it is shown that the ERR336942 sample performs slightly better
+than ERR336956 and both outperform ERR336935, this agrees with
+\ref{fig:enrich}A, where only in the left-most panel doesn't seem to
+be a separation between the low $\mbox{ARC}$ arm and the other.
 
@@
   \centering
-  \includegraphics[width = .8\textwidth]{../figs/for_paper/Sig70_enrichment.pdf}
-  \includegraphics[width = .8\textwidth,page =3]{../figs/for_paper/Local_SCC_indicator_by_strata.pdf}
-  \caption{A) Hexbin plots of $\mbox{ARC}$ against $\mbox{URCR}$ and
-    B) boxplots of the $\mbox{local-NSC}$ stratified by nr. of reads
-    mapped to only one position in the regions for the ChIP-exo
-    $\sigma^{70}$ experiments divided by condition and biological
-    replicate}
+  \includegraphics[width = .8\textwidth]{../figs/Carroll_mice_for_paper/FoxA1_enrichment.pdf}
+  \includegraphics[width = .8\textwidth,page =3]{../figs/Carroll_mice_for_paper/Local_SCC_indicator_by_strata.pdf}
+  \includegraphics[width = .8\textwidth,page =4]{../figs/Carroll_mice_for_paper/Local_SCC_indicator_by_strata_sameRegion.pdf}
+  \caption{Using the mouse-FoxA1 experiment from \cite{exoillumina}:
+    A) Hexbin plots of $\mbox{ARC}$ against $\mbox{URCR}$, B) Boxplots
+    of the $\mbox{local-NSC}$ stratified by nr. of reads mapped to
+    only one position in the regions for the ChIP-exo experiments
+    divided by condition and biological replicate and C) Boxplots of
+    the a set of ERR336942 regions, stratified as in (B) with the
+    $\mbox{local-NSC}$ calculated with the different biological
+    replicates.}
   \label{fig:enrich}
@@
 visualization in figure \ref{fig:strand}. A) presents the FSR's
-behavior as lower depth regions are being filtered out. while B)
-shows which percentage of the regions are composed by fragments of
-both strand or only one. Roughly speaking in a lower quality dataset
-there is going to be a higher amount of single stranded-regions. To
-asses this, we divided the ChIP-exo regions into two sets, the ones
-that overlap with a known set of peaks and the opposite and performed
-$\chi^2$ tests.
+behavior as lower depth regions are being filtered out, while B) shows
+which percentage of the regions are composed by fragments of both
+strand or only one. Roughly speaking in a lower quality dataset there
+is going to be a higher amount of single stranded-regions. To evaluate
+this statement, for each replicate of the FoxA1 dataset from
+\cite{exoillumina}, we divided the regions considered by the QC
+pipeline into two sets by reviewing if they overlap any of the
+sample's peaks and then calculate the Kolmogorov-Smirnov test.
 
@@
   \centering  
-  \includegraphics[width = .8\textwidth]{../figs/fsr/draft_fsr_plot.pdf}
-  \includegraphics[width = .8\textwidth]{../figs/fsr/draft_label_prop_plot.pdf}
-  \caption{ABCD}
+  \includegraphics[width = .8\textwidth,page = 3]{../figs/Carroll_mice_for_paper/Strand_imbalance.pdf} 
+  \caption{Strand imbalance QC plots for the same data as in figure \ref{fig:enrich}.}
   \label{fig:strand}
@@
 algorithm as candidate regions. Default tuning parameters were used
-during model fitting for all methods.
-
-% And for
-% the CexoR algorithm, we used its R implementation version 1.8.0 which
-% is available from
-% \url{http://bioconductor.org/packages/release/bioc/html/CexoR.html}
+during model fitting for all methods. We tried to use the CexoR
+algorithm with its R implementation version 1.8.0 available from
+\url{http://bioconductor.org/packages/release/bioc/html/CexoR.html},
+but it returned an error everytime we used it.
 
@@
 from the R software version 3.2.1 using the default tuning parameters.
-
-%% not sure if here or before it would be nice to add an example of how this 
-%% method works. probably here
 
@@
   \centering
-  \includegraphics[width = .8\textwidth,page = 3]{../figs/for_paper/Sig70_enrichment.pdf}
-  \includegraphics[width = .8\textwidth,page = 4]{../figs/for_paper/Sig70_enrichment.pdf}
+  \includegraphics[width = .8\textwidth,page = 3]{../figs/Carroll_mice_for_paper/FoxA1_enrichment.pdf}
+  \includegraphics[width = .8\textwidth,page = 4]{../figs/Carroll_mice_for_paper/FoxA1_enrichment.pdf}
   \caption{Hexbin plots of $\mbox{ARC}$ vs $\mbox{URCR}$ for each

### 2015-11-19 rmarkdown/chip_exo_paper.tex
Added plots on enrichment and strand imbalance

@@
 
-Figure \ref{fig:enrich}A shows hexbin plots illustrating the
-interaction between these two quantities. Notice how there are two
-strong arms in each panel: The first one corresponds to regions with
-low $\mbox{ARC}$ values and $\mbox{URCR}$ values that can vary across
-the whole $[0,1]$ interval, while the second one shows a decreasing
-trend in $\mbox{URCR}$ as $\mbox{ARC}$ increases. When an experiment
-shows a higher degree of enrichment, then the separation of this two
-arms is more noticeable, since the second arm corresponds to possibly
-enriched regions (\ref{fig:enrich2}A and \ref{fig:enrich2}B).
+Using the mouse-FoxA1 experiment from \cite{exoillumina}, the
+relationship between this two quantities, library complexty and sample
+enrichment was explored. Figure \ref{fig:enrich}A shows hexbin plots
+illustrating the interaction between these two quantities. Notice how
+there are two strong arms in each panel: The first one corresponds to
+regions with low $\mbox{ARC}$ values and $\mbox{URCR}$ values that can
+vary across the whole $[0,1]$ interval, while the second one shows a
+decreasing trend in $\mbox{URCR}$ as $\mbox{ARC}$ increases. When an
+experiment shows a higher degree of enrichment, then the separation of
+this two arms is more noticeable, since the second arm corresponds to
+possibly enriched regions (\ref{fig:enrich2}A and \ref{fig:enrich2}B).
 
-Figure \ref{fig:enrich}B shows boxplots of the $\mbox{local-NSC}$
-stratified by the number of reads mapped exactly to one position for
-all $\sigma^{70}$ ChIP-exo experiments; for each stratum, we sampled
-100 regions that whenever possible (in the opposite case,
-all the regions in the category were considered). There is an
-increasing trend as the number of unique position increases, which
-means that this is an effective indicator of library complexity since
-regions where the reads are mapped to more positions are usually more
-complex.
-
-%% perhaps do two other slots about how this works... a good experiment vs a bad experiment
-%% of the same region
+Figure \ref{fig:enrich}B shows boxplots of the $\mbox{local-NSC}$ on
+regions common to the 3 replicates and stratified by the number of
+reads mapped exactly to one position for all $\sigma^{70}$ ChIP-exo
+experiments; The high stratum is defined by regions consising of more
+than 100 unique positions, the medium stratum for regions
+where the number of unique position is on the
+(50,100) range and the low stratum in the
+(20,50) range. For each stratum and biological
+replicate, we sampled 400 regions whenever possible (in
+the opposite case, all the regions in the category were considered)
+and calculated the $\mbox{local-NSC}$ for those regions. For all three
+replicates, it is shown an increasing trend as the number of unique
+position increases, which means that the $\mbox{local-NSC}$ is an
+effective indicator of library complexity. Figure \ref{fig:enrich}C
+shows the $\mbox{local-NSC}$ coefficient calculated for the same set
+of regions, stratified by the number of unique positions in the
+regions of ERR336942 using the same criteria as before. In the high
+panel it is shown that the ERR336942 sample performs slightly better
+than ERR336956 and both outperform ERR336935, this agrees with
+\ref{fig:enrich}A, where only in the left-most panel doesn't seem to
+be a separation between the low $\mbox{ARC}$ arm and the other.
 

### 2015-11-19 rmarkdown/chip_exo_paper.R

@@ -1,6 +1,9 @@
 ## ----strata_param , include =FALSE,echo = FALSE, eval = TRUE-------------
 
-  nstrata <- 100
+  nstrata <- 400
+  high <- 100
+  med <- 50
+  low <- 20
   
 
 ## ----comp_param,include = FALSE,echo = FALSE,eval = TRUE-----------------

### 2015-11-19 rscripts/carrol_mice_library_complexity_and_enrichment.R

@@ -1,0 +1,510 @@
+
+rm(list = ls())
+
+library(reshape2)
+library(ChIPUtils)
+library(ggplot2)
+library(data.table)
+library(GenomicAlignments)
+library(parallel)
+library(viridis)
+
+dr <- "/p/keles/ChIPexo/volume3/CarrollData/mouse"
+files <- list.files(dr)
+
+files <- files[grep("bam",files)]
+files <- files[grep("bai",files,invert = TRUE)]
+
+files <- file.path(dr,files)
+
+reads <- mclapply(files,create_reads,mc.cores = 3)
+names(reads) <- basename(files)
+
+pbc <- sapply(reads,PBC)
+depth <- sapply(reads,nreads)
+ssd <- sapply(reads,SSD) ## there is a small issue with this function. sometimes can get numeric overflow
+## Warning messages:
+## 1: In sum(w) : integer overflow - use sum(as.numeric(.))
+## 2: In sum(w) : integer overflow - use sum(as.numeric(.))
+## 3: In sum(w) : integer overflow - use sum(as.numeric(.))
+
+
+
+FSR <- function(reads)
+{
+  summ <- summary(reads)
+  rF <- summ[,sum(readsF)]
+  rR <- summ[,sum(readsR)]
+
+  return(rF / (rF + rR))
+}
+
+fsr <- sapply(reads,FSR)
+
+exo <- data.table(sample  = names(reads))
+
+exo[,nreads := depth]
+exo[,pbc := pbc]
+exo[,ssd := ssd] 
+exo[,fsr := fsr]
+
+mouse.size <- data.table(read.table(
+ system.file( "extdata","chrom.sizes","mm9.chrom.sizes",package = "ChIPUtils")))
+
+scc <- lapply(reads,strand_cross_corr,shift = 1:300,chrom.sizes = mouse.size,parallel = TRUE)
+
+nsc <- sapply(scc,function(x)x[ ,max(cross.corr) / min(cross.corr)])
+exo[, nsc := nsc]
+
+## Carroll mouse
+## ERR336942.bam FoxA1-rep1 
+## ERR336956.bam FoxA1-rep2
+## ERR336935.bam FoxA1-rep3
+
+bamfiles <-c("ERR336942.bam","ERR336956.bam","ERR336935.bam")
+repl <- paste0("rep-",1:3)
+
+scc <- mapply(function(x,y)x[,sample := y],scc,exo[,(sample)],SIMPLIFY = FALSE)
+SCC <- do.call(rbind,scc)
[…]

### 2015-11-19 rscripts/carroll_mice_strand_imbalance.R

@@ -1,0 +1,221 @@
+
+rm(list = ls())
+
+library(reshape2)
+library(ChIPUtils)
+library(ggplot2)
+library(data.table)
+library(GenomicAlignments)
+library(parallel)
+library(viridis)
+
+dr <- "/p/keles/ChIPexo/volume3/CarrollData/mouse"
+files <- list.files(dr)
+
+files <- files[grep("bam",files)]
+files <- files[grep("bai",files,invert = TRUE)]
+
+files <- file.path(dr,files)
+
+reads <- mclapply(files,create_reads,mc.cores = 3)
+names(reads) <- basename(files)
+
+
+create_regions <- function(reads,lower,rangesOnly=TRUE)
+{
+  stopifnot(class(reads) == "GRanges")
+  stopifnot(lower > 0)
+  sreads <- split(reads,seqnames(reads))
+  sreads <- as.list(sreads)
+  cover <- mcmapply(function(x,y)coverage(x)[[y]],sreads,names(sreads),mc.cores = detectCores())
+  islands <- lapply(cover,slice,lower = lower, rangesOnly = rangesOnly)
+  islands <- mapply(function(x,y)GRanges(seqnames = y,ranges = x),islands,names(islands),SIMPLIFY =FALSE)
+  islands <- do.call(rbind,lapply(islands,gr2dt))
+  return(islands)
+}
+
+gr <- lapply(reads,function(x){
+  byChr <- mapply(rbind,readsF(x),readsR(x),SIMPLIFY =FALSE)
+  out <- do.call(rbind,byChr)
+  return(dt2gr(out))
+})
+
+regs <- lapply(gr,create_regions,lower  = 1)
+
+build_stats <- function(region,reads)
+{
+  ## fix formats and stuff
+  region <- dt2gr(region)
+  
+  ov <- findOverlaps(region,reads)
+  reads <- gr2dt(reads)
+  w <- width(region)    
+  region <- gr2dt(region)
+  region[ , width := w]
+  region[, match := paste0(seqnames,":",start,"_",end)]
+  reads[  subjectHits(ov), match := region[queryHits(ov), (match)] ]  
+  reads[,strand := ifelse(strand == "+", "F","R")]
+  reads <- reads[!is.na(match)]
+
+  ## get base statistics
[…]

### 2015-11-19 rmarkdown/chip_exo_paper.Rnw
Added plots on enrichment and strand imbalance

@@
 
+<<fsr,include = FALSE,eval = TRUE,echo = FALSE>>=
+
+  de <- 100
+
+pval <- list()
+pval[[1]] <- 0.1039
+pval[[2]] <- 0.0064
+pval[[3]] <- 0.0871
+
+@
+
 In order to asses the strand imbalance we create the following
@@
 this statement, for each replicate of the FoxA1 dataset from
-\cite{exoillumina}, we divided the regions considered by the QC
-pipeline into two sets by reviewing if they overlap any of the
-sample's peaks and then calculate the Kolmogorov-Smirnov test.
+\cite{exoillumina}, we divided the regions formed by at least
+\Sexpr{de} fragments into into the ones that overlap with any of the
+sample's peaks. The p.value for ERR36935 is \Sexpr{pval[[1]]}, for
+ERR336942 is \Sexpr{pval[[2]]} and for ERR336956 is \Sexpr{pval[[3]]}.
 

### 2015-11-19 rmarkdown/chip_exo_paper.Rnw
Added ChIPnexus paper to bibliography and added section of biases comparison of ChIPexo for eukaryotic genomes

@@
 are not established quality control pipelines for ChIP-exo. To address
-this challenge, we suggest a series of quality control visualizations
-to understand which biases are present in ChIP-exo data. Related to
-quality control, Previous ChIP-exo analysis used ChIP-Seq samples to
-compare the resolution between experiments (\cite{exo1}, \cite{exo2},
-\cite{exoillumina}). In \cite{carroll.qc}, Carroll et. al. studied the
-use of the Strand-Cross Correlation (SCC) \cite{strandcc}. and showed
-that by filtering blacklisted regions the estimation of the SCC is
-improved. However, this method requires to know blacklisted regions in
-advance which may not be available. In our pipeline we propose two
-out-the-shelf metrics equivalent to RSC and NSC for the estimation of
-the signal-to-noise ratio for a ChIP-exo sample.
+this challenge, we suggest a collection of quality control
+visualizations to understand which biases are present in ChIP-exo
+data. Related to quality control, Previous ChIP-exo analysis used
+ChIP-Seq samples to compare the resolution between experiments
+(\cite{exo1}, \cite{exo2}, \cite{exoillumina}). In \cite{carroll.qc},
+Carroll et. al. studied the use of the Strand-Cross Correlation (SCC)
+\cite{strandcc}. and showed that by filtering blacklisted regions the
+estimation of the SCC is improved. However, this method requires to
+know blacklisted regions in advance which may not be available. In our
+pipeline we propose two out-the-shelf metrics equivalent to RSC and
+NSC for the estimation of the signal-to-noise ratio for a ChIP-exo
+sample.
 
@@
 
-To have a first view on the data, we calculated a series of the most
-commonly used quality indicators for ChIP-Seq data, which are listed
-in table \ref{tab:qcbase}. The PBC, NSC and RSC were calculated as
-defined in \cite{encode_qc} and the SSD was calculated as defined in
+To have a first view on the data, we calculated the most commonly used
+quality indicators for ChIP-Seq data, which are listed in table
+\ref{tab:qcbase}. The PBC, NSC and RSC were calculated as defined in
+\cite{encode_qc} and the SSD was calculated as defined in
 \cite{htseq}. Additionally, we calculated the FSR to compare the
@@
   \includegraphics[width = .9\textwidth]{../figs/for_paper/coverage_diagram.pdf}
-  \caption{Diagram of the ChIP-exo Quality control pipeline(this is a
-    place-holder). The genome is partitioned into several regions by
+  \caption{Diagram of the ChIP-exo Quality control pipeline [this is a
+    place-holder]. The genome is partitioned into several regions by
     removing the empty segments and for each regions we calculate a
@@
   \caption{Using the mouse-FoxA1 experiment from \cite{exoillumina}:
-    A) Hexbin plots of $\mbox{ARC}$ against $\mbox{URCR}$, B) Boxplots
-    of the $\mbox{local-NSC}$ stratified by nr. of reads mapped to
-    only one position in the regions for the ChIP-exo experiments
-    divided by condition and biological replicate and C) Boxplots of
-    the a set of ERR336942 regions, stratified as in (B) with the
-    $\mbox{local-NSC}$ calculated with the different biological
-    replicates.}
+    A) Hexbin plots of $\mbox{ARC}$ against $\mbox{URCR}$, in general
+    we can see a slight separation into two strong arms, one
+    corresponds to low $\mbox{ARC}$ and varying $\mbox{URCR}$, and for
+    the other $\mbox{URCR}$ decreases as $\mbox{ARC}$ increases B)
+    Boxplots of the $\mbox{local-NSC}$ stratified by nr. of reads
+    mapped to only one position in the regions for the ChIP-exo
+    experiments divided by condition and biological replicate and C)
+    Boxplots of the a set of ERR336942 regions, stratified as in (B)
+    with the $\mbox{local-NSC}$ calculated with the different
+    biological replicates.}
   \label{fig:enrich}
@@
 sample's peaks. The p.value for ERR36935 is \Sexpr{pval[[1]]}, for
-ERR336942 is \Sexpr{pval[[2]]} and for ERR336956 is \Sexpr{pval[[3]]}.
+ERR336942 is \Sexpr{pval[[2]]} and for ERR336956 is
+\Sexpr{pval[[3]]}. This shows that only the middle sample from figure
+\ref{fig:strand} is the least imbalance sample from the the three,
+while the other two still show a higher imbalance.
 
@@
   \includegraphics[width = .8\textwidth,page = 3]{../figs/Carroll_mice_for_paper/Strand_imbalance.pdf} 
-  \caption{Strand imbalance QC plots for the same data as in figure \ref{fig:enrich}.}
+  \caption{Strand imbalance QC plots for the same data as in figure
+    \ref{fig:enrich}. A) FSR distribution quantiles as the lower depth
+    regions are being filtered out. B) Stacked histogram with the
+    proportion of regions that are formed by two strands or only one.}
   \label{fig:strand}
@@
     less balanced in the regions with potential binding sites compared
-    to SET ChIP-Seq data. }
+    to SET ChIP-Seq data. [C and D correspond to fig4 from manuscript]
+    C) ChIP tag counts increase linearly as mappability scores
+    increases. D) ChIP tag counts increase linearly as GC content
+    score increases when GC content is less than 0.6 and then ChIP tag
+    counts decrease as GC content increases. }
   \label{fig:comp}
 \end{figure}
+
+\subsubsection{ChIP-exo data from eukaryotic genome shows clear sequence biases}
+\label{sec:eukaryotic}
+
+We evaluated ChIP-exo data for CTCF factor from human genome
+\cite{exo1} to investigate issues specific to eukaryotic genomes for
+binding sites identification. Figures \ref{fig:comp}C and
+\ref{fig:comp}D display the bin-level average read counts against
+mappability and GC content. Each data point is obtained by averaging
+the read counts across bins with the same mappability of GC
+content. These results indicate that binding site identification in
+ChIP-exo sample might also benefic from the use of methods that take
+into account apparent sequence biases such as mappability and GC
+content.
 
@@
 library complexity.
+
+\subsection{dPeak outperforms compeing methods in discovering closely
+  space binding event from ChIP-exo and ChIP-Seq data}
+\label{sec:dpeak_comp}
+
+\emph{I plan to update this analysis, since the dpeak's new strategy
+  seems to decrease resolution. Also, Apex was updated into MACE, so I
+  think it could be better to add that method instead}
 

### 2015-11-19 rmarkdown/chip_exo_paper.tex
Added ChIPnexus paper to bibliography and added section of biases comparison of ChIPexo for eukaryotic genomes

@@
 this statement, for each replicate of the FoxA1 dataset from
-\cite{exoillumina}, we divided the regions considered by the QC
-pipeline into two sets by reviewing if they overlap any of the
-sample's peaks and then calculate the Kolmogorov-Smirnov test.
+\cite{exoillumina}, we divided the regions formed by at least
+100 fragments into into the ones that overlap with any of the
+sample's peaks. The p.value for ERR36935 is 0.1039, for
+ERR336942 is 0.0064 and for ERR336956 is
+0.0871. This shows that only the middle sample from figure
+\ref{fig:strand} is the least imbalance sample from the the three,
+while the other two still show a higher imbalance.
 

### 2015-11-19 rmarkdown/chip_exo_paper.R

@@ -6,6 +6,16 @@
   low <- 20
   
 
+## ----fsr,include = FALSE,eval = TRUE,echo = FALSE------------------------
+
+  de <- 100
+
+  pval <- list()
+  pval[[1]] <- 0.1039
+  pval[[2]] <- 0.0064
+  pval[[3]] <- 0.0871
+
+
 ## ----comp_param,include = FALSE,echo = FALSE,eval = TRUE-----------------
 
   bin_size <- 150

### 2015-11-19 rmarkdown/chip_exo_paper.Rnw
Added mappability and gc score plot to paper

@@
   \includegraphics[width = .46\textwidth]{../figs/for_paper/forward_strand_ratio_comp_old.pdf}
+  \includegraphics[width = .46\textwidth,page = 1]{../figs/for_paper/eukaryotic_bias_CTCF.pdf}
+  \includegraphics[width = .46\textwidth,page = 2]{../figs/for_paper/eukaryotic_bias_CTCF.pdf}
   \caption{ A) ChIP signal in ChIP-exo was linearly related to that in
@@
 
+We provide a ChIP-exo QC pipeline capable of asses the balance between
+enriched samples and low complexity regions. It is shown that the
+``peak-pair'' assumption doesn't hold in practice and we provide two
+out-of-the-box visualization capable to asses the strand imbalance in
+a ChIP-exo experiment.
+
+We updated the dPeak algorithm from \cite{dpeak} and showed that
+ChIP-exo is comparable in resolution and sensitivity with PET ChIP-Seq
+and outperforms SET ChIP-Seq. ChIP-exo compared with PET and SET
+ChIP-Seq at fixed depth sample is able to identify more targets at a
+lower resolution.
+
 \newpage

### 2015-11-19 rscripts/CTCF_eukaryotic_biases.R

@@ -1,0 +1,94 @@
+rm( list=ls() )
+graphics.off()
+library(mosaics)
+library(data.table)
+library(ggplot2)
+library(RColorBrewer)
+library(scales)
+library(reshape2)
+
+Rlist <- system( "ls /u/w/e/welch/Desktop/Docs/packages/mosaics/R/*.R", intern=TRUE )
+for ( i in 1:length(Rlist) ) {
+	source( Rlist[i] )
+}
+
+
+#constructBins( infile="CTCF.bowtie", fileFormat="bowtie", outfileLoc="./", 
+#	byChr=FALSE, useChrfile=FALSE, chrfile=NULL, excludeChr=NULL, 
+#	PET=FALSE, fragLen=200, binSize=200, capping=0, perl = "perl" )
+         
+dir_out <- "/p/keles/Dongjun/ChIP-exo/paper/figures/"
+
+bin.exo <- readBins( c("chip","M","GC","N"),
+	c( "/NO_BACKUP/KelesGroup_DongjunChung/ChIP-exo/CTCF/raw_data/ChIP-exo/CTCF.bowtie_fragL200_bin200.txt",
+	"/NO_BACKUP/KelesGroup_DongjunChung/ChIP-exo/CTCF/raw_data/map50/map50_fragL200_bin200.txt",
+	"/NO_BACKUP/KelesGroup_DongjunChung/ChIP-exo/CTCF/raw_data/GC/GC_fragL200_bin200.txt",
+	"/NO_BACKUP/KelesGroup_DongjunChung/ChIP-exo/CTCF/raw_data/N/N_fragL200_bin200.txt" ) )
+
+bin.seq <- readBins( c("chip","input","M","GC","N"),
+	c( "/NO_BACKUP/KelesGroup_DongjunChung/ChIP-exo/CTCF/raw_data/ChIP-seq_Crawford/CTCF_Crawford.bowtie_fragL200_bin200.txt",
+	"/NO_BACKUP/KelesGroup_DongjunChung/ChIP-exo/CTCF/raw_data/Input_Crawford/Input_Crawford.bowtie_fragL200_bin200.txt",
+	"/NO_BACKUP/KelesGroup_DongjunChung/ChIP-exo/CTCF/raw_data/map50/map50_fragL200_bin200.txt",
+	"/NO_BACKUP/KelesGroup_DongjunChung/ChIP-exo/CTCF/raw_data/GC/GC_fragL200_bin200.txt",
[…]

### 2015-11-20 rmarkdown/chip_exo_paper.Rnw
first draft of paper

@@
-%\VignetteEngine{knitr::knitr}
+% \VignetteEngine{knitr::knitr}
 \documentclass{bmcart}
@@
     ChIP-Seq. Although ChIP-exo has drawn much attention and is
-  considered as powerful assay, currently, no systematic studies have
-  yet been conducted to determine optimal strategies for experimental
-  design and analysis of ChIP-exo. In order to address these
-  questions, we evaluated diverse aspects of ChIP-exo and found the
-  following characteristics of ChIP-exo data. First, Background of
-  ChIP-exo data is quite different from that of ChIP-Seq data. However,
-  sequence biases inherently present in ChIP-Seq data still exist in
-  ChIP-exo data. Second, in ChIP-exo data, reads are located around
-  binding sites much more tightly and hence, it has potential for high
-  resolution identification of protein-DNA interaction sites, hence
-  the space to allocate the reads is greatly reduced. Third, although
-  often assumed in the ChIP-exo data analysis methods, the peak pair
-  assumption does not hold well in real ChIP-exo data. Fourth, spatial
-  resolution of ChIP-exo is comparable to that of PET ChIP-Seq and
-  both of them are significantly better than resolution of SET
-  ChIP-Seq. Finally, for given fixed sequencing depth, ChIP-exo
-  provides higher sensitivity, specificity, and spatial resolution
-  than PET ChIP-Seq.
+    considered as powerful assay, currently, no systematic studies
+    have yet been conducted to determine optimal strategies for
+    experimental design and analysis of ChIP-exo. In order to address
+    these questions, we evaluated diverse aspects of ChIP-exo and
+    found the following characteristics of ChIP-exo data. First, the
+    background of ChIP-exo data is quite different from that of
+    ChIP-Seq data. However, sequence biases inherently present in
+    ChIP-Seq data still exist in ChIP-exo data. Second, in ChIP-exo
+    data, reads are located around binding sites much more tightly and
+    hence, it has potential for high resolution identification of
+    protein-DNA interaction sites, and also the space to allocate the
+    reads is greatly reduced. Third, although often assumed in the
+    ChIP-exo data analysis methods, the ``peak pair'' assumption does
+    not hold well in real ChIP-exo data. Fourth, spatial resolution of
+    ChIP-exo is comparable to that of PET ChIP-Seq and both of them
+    are significantly better than resolution of SET ChIP-Seq. Finally,
+    for given fixed sequencing depth, ChIP-exo provides higher
+    sensitivity, specificity and spatial resolution than PET
+    ChIP-Seq.
 
-  In this article, we provide a quality control pipeline which
+    In this article we provide a quality control pipeline which
     visually asses ChIP-exo biases and calculates a signal-to-noise
-  measure. Also, we updated dPeak \cite{dpeak}, which makes a striking
-  balance in sensitivity, specificity, and spatial resolution for
-  ChIP-exo data analysis.
+    measure. Also, we updated dPeak \cite{dpeak}, which makes a
+    striking balance in sensitivity, specificity and spatial
+    resolution for ChIP-exo data analysis.
 
 \end{abstract}
+
 \begin{keyword}
@@
 \end{frontmatter}
-
-\newpage
-
-* to remove later
-
-\tableofcontents
 
@@
 fragments are located much closed to bound protein compared to
-ChIP-Seq. This is the step unique to ChIP-exo that could potentially
+ChIP-Seq. This step is unique to ChIP-exo and could potentially
 provide significantly higher spatial resolution compared to
@@
 ChIP-exo data, choice of optimal statistical methods, and
-determination of optimal sequencing depth. However, currently, the
+determination of optimal sequencing depth. However, currently the
 number of available ChIP-exo data is still limited and their
@@
 organisms: CTCF factor in human \cite{exo1}; ER factor in human and
-FoxA1 factor in mouse from \cite{exoillumina}; and generated
-$\sigma^{70}$ factor in Escherichia coli (E. Coli) under aerobic ($ +
-O_2$) condition, and treated by rifampicin by 0 and 20 minutes.
+FoxA1 factor in mouse \cite{exoillumina}; and generated $\sigma^{70}$
+factor in Escherichia coli (E. Coli) under aerobic ($ + O_2$)
+condition, and treated by rifampicin by 0 and 20 minutes.
 
@@
 complex than the libraries generated by ChIP-Seq
-\cite{exo_review}. Hence, most of current guidelines \cite{encode_qc}
-may not be applicable on ChIP-exo, additionally to our knowledge there
-are not established quality control pipelines for ChIP-exo. To address
-this challenge, we suggest a collection of quality control
-visualizations to understand which biases are present in ChIP-exo
-data. Related to quality control, Previous ChIP-exo analysis used
-ChIP-Seq samples to compare the resolution between experiments
-(\cite{exo1}, \cite{exo2}, \cite{exoillumina}). In \cite{carroll.qc},
-Carroll et. al. studied the use of the Strand-Cross Correlation (SCC)
-\cite{strandcc}. and showed that by filtering blacklisted regions the
-estimation of the SCC is improved. However, this method requires to
-know blacklisted regions in advance which may not be available. In our
-pipeline we propose two out-the-shelf metrics equivalent to RSC and
-NSC for the estimation of the signal-to-noise ratio for a ChIP-exo
-sample.
+\cite{exo_review}. Hence, most of current QC guidelines
+\cite{encode_qc} may not be applicable on ChIP-exo, additionally to
+our knowledge there are not established quality control pipelines for
+ChIP-exo. To address this challenge, we suggest a collection of
+quality control visualizations to understand which biases are present
+in ChIP-exo data. Previous ChIP-exo analysis used ChIP-Seq samples to
+compare the resolution between experiments (\cite{exo1}, \cite{exo2},
+\cite{exoillumina}); in \cite{carroll.qc}, Carroll et. al. studied the
+use of the Strand-Cross Correlation (SCC) \cite{strandcc} and showed
+that by filtering blacklisted regions the estimation of the SCC is
+improved. However, this method requires to know blacklisted regions in
+advance which may not be available, additionally using SCC may not be
+useful since the peaks that are attained at the read and fragment
+length are confused in a typical ChIP-exo SCC curve.  In our pipeline
+we propose two out-the-shelf analysis: an enrichment plot and the
+local normalized SCC cofficient.
 
@@
 
-In this paper, we demonstrate that the peak pair assumption of Rhee
-and Pugh \cite{exo1} does not hold well in real ChIP-exo
+In this paper, we demonstrate that the ``peak pair'' assumption of
+Rhee and Pugh \cite{exo1} does not hold well in real ChIP-exo
 data. Furthermore, we found that when we analyze ChIP-exo data from
@@
 to improve sensitivity and specificity of binding site
-identification. dPeak outperforms or performs competitively with
-ChIP-exo data analysis such as GEM and MACE. More importantly, when
-comparable number of reads is used for both ChIP-exo and ChIP-Seq,
-dPeak couple with ChIP-exo data provides resolution comparable to PET
-ChIP-Seq and both significantly improve the resolution of protein
-binding site identification compared to SET-based analysis with any of
-the available methods.
+identification. We evaluated several method to identify binding events
+and dPeak outperforms or performs competitively respecto to GEM and
+MACE when analyzing ChIP-exo data. More importantly, when comparable
+number of reads is used for both ChIP-exo and ChIP-Seq, dPeak coupled
+with ChIP-exo data provides resolution comparable to PET ChIP-Seq and
+both significantly improve the resolution of protein binding site
+identification compared to SET-based analysis with any of the
+available methods.
 
@@
 
-To have a first view on the data, we calculated the most commonly used
-quality indicators for ChIP-Seq data, which are listed in table
-\ref{tab:qcbase}. The PBC, NSC and RSC were calculated as defined in
-\cite{encode_qc} and the SSD was calculated as defined in
-\cite{htseq}. Additionally, we calculated the FSR to compare the
-number of fragments in each strand. A few conclusions that we may get
-by observing the PBC in table \ref{tab:qcbase} are all the samples
-have a quite low library complexity, hence it would be suggested to
-repeat the experiment.
+We calculated the most commonly used quality indicators for ChIP-Seq
+data, which are listed in table \ref{tab:qcbase}. The PBC and NSC were
+calculated as defined in \cite{encode_qc} and the SSD was calculated
+as defined in \cite{htseq}. Additionally, we calculated the FSR to
+compare the number of fragments in each strand. A few conclusions that
+we may get by observing the PBC in table \ref{tab:qcbase} are all the
+samples have a quite low library complexity, hence it would be
+suggested to repeat the experiment.
 
@@
 
-In a good ChIP-Seq experiment the strand cross-correlation is
-maximized at the fragment length and there is also a ``phantom peak''
-at the sample's read length (\cite{encode_qc} and \cite{strandcc}).
-However, in ChIP-exo's case this two peaks are confounded. Hence an
-enrichment measure based in the strand cross-correlation such as NSC
-or RSC is harder to interpret. In figure \ref{figsup:scc} we plotted
-the SCC curves for $\sigma^{70}$ ChIP-exo experiments used to
-calculate the NSC and RSC values in table \ref{tab:qcbase}. According
-to ENCODE's quality metrics, NSC values close to one corresponds to
-low enrichment samples or bad quality datasets, hence in our case the
-last four samples would be considered to have higher quality than they
-actually have. On the other hand, all PBC values are considerably low
-and correspond to samples with severe bottlenecking issues, in
-particular we are going to show that for the first two samples we are
-obtaining a smaller resolution and identifying a comparable number of
-targets as in ChIP-Seq PET experiments.
+It is of special importance to notice that the vlaues of PBC or are
+quite low, hence these experiment would be repited if there were
+considered as ChIP-Seq experiments. On the other hand, in a good
+ChIP-Seq experiment the strand cross-correlation is maximized at the
+fragment length and there is also a ``phantom peak'' at the sample's
+read length (\cite{encode_qc} and \cite{strandcc}).  However, in
+ChIP-exo's case this two peaks are confounded. Hence an enrichment
+measure based in the strand cross-correlation such as NSC or RSC is
+harder to interpret. In figure \ref{figsup:scc} we plotted the SCC
+curves for $\sigma^{70}$ ChIP-exo experiments used to calculate the
+NSC and RSC values in table \ref{tab:qcbase}. According to ENCODE's
+quality metrics, NSC values close to one corresponds to low enrichment
+samples or bad quality datasets, hence in our case the last four
+samples would be considered to have higher quality than they actually
+have. On the other hand, all PBC values are considerably low and
+correspond to samples with severe bottlenecking issues, in particular
+we are going to show that for the first two samples we are obtaining a
+smaller resolution and identifying a comparable number of targets as
+in ChIP-Seq PET experiments.
 
@@
 Figure \ref{fig:qcdiagram} shows a flowchart for the ChIPexoQC
-pipeline. Which basically partitions genome by keeping the
+pipeline. Which basically partitions the genome by keeping the
 non-digested ChIP-exo regions. Then, for each region it calculates a
-series of independent statistics. Finally, it creates several
-visualizations designed to asses the overall quality levels of A
-ChIP-exo sample.
+series of statistics. Finally, it creates several visualizations
+designed to asses the quality levels of a ChIP-exo sample.
 
@@
     removing the empty segments and for each regions we calculate a
-    collection of qc indicators.}
+    collection of QC indicators.}
   \label{fig:qcdiagram}
@@
 
-Using the mouse-FoxA1 experiment from \cite{exoillumina}, the
-relationship between this two quantities, library complexty and sample
-enrichment was explored. Figure \ref{fig:enrich}A shows hexbin plots
-illustrating the interaction between these two quantities. Notice how
-there are two strong arms in each panel: The first one corresponds to
-regions with low $\mbox{ARC}$ values and $\mbox{URCR}$ values that can
-vary across the whole $[0,1]$ interval, while the second one shows a
-decreasing trend in $\mbox{URCR}$ as $\mbox{ARC}$ increases. When an
-experiment shows a higher degree of enrichment, then the separation of
-this two arms is more noticeable, since the second arm corresponds to
-possibly enriched regions (\ref{fig:enrich2}A and \ref{fig:enrich2}B).
+Using the mouse-FoxA1 experiment from \cite{exoillumina} and the
+relationship between this two quantities, library complexity and
+sample enrichment was explored. Figure \ref{fig:enrich}A shows hexbin
+plots illustrating the interaction between these two
+quantities. Notice how there are two strong arms in each panel: The
+first one corresponds to regions with low $\mbox{ARC}$ values and
+varying $\mbox{URCR}$ values across the $(0,1]$ interval, while the
+second one shows a decreasing trend in $\mbox{URCR}$ as $\mbox{ARC}$
+increases. When an experiment shows a higher degree of enrichment,
+then the separation of this two arms is more noticeable, since the
+second arm corresponds to possibly enriched regions
+(\ref{fig:enrich2}A and \ref{fig:enrich2}B).
 
-<<strata_param , include =FALSE,echo = FALSE, eval = TRUE>>=
+<<strata_para , include =FALSE,echo = FALSE, veal = TRUE>>=
 
-  nstrata <- 400
+  strata <- 400
   high <- 100
@@
 
-Figure \ref{fig:enrich}B shows boxplots of the $\mbox{local-NSC}$ on
+Figure \ref{fig:enrich}B shows box plots of the $\mbox{local-NSC}$ on
 regions common to the 3 replicates and stratified by the number of
 reads mapped exactly to one position for all $\sigma^{70}$ ChIP-exo
-experiments; The high stratum is defined by regions consising of more
+experiments; The high stratum is defined by regions consisting of more
 than \Sexpr{high} unique positions, the medium stratum for regions
@@
 (\Sexpr{low},\Sexpr{med}) range. For each stratum and biological
-replicate, we sampled \Sexpr{nstrata} regions whenever possible (in
+replicate, we sampled \Sexpr{strata} regions whenever possible (in
 the opposite case, all the regions in the category were considered)
@@
 be a separation between the low $\mbox{ARC}$ arm and the other.
+
+\emph{Comment: I think I need to change the name of the FoxA1 samples
+  to Rep1 to Rep3, but I wasn't sure about this. Therefore I used the
+  name of the bam files}
 
@@
 enriched regions usually have a higher concentration of fragments,
-therefore we examine the FSR (defined as the ratio of the number of
+therefore we examined the FSR (defined as the ratio of the number of
 forward stranded reads divided by the total number of reads in a given
@@
 0.5, which means that there are roughly the same amount of reads in
-both strands. However, figure \ref{fig:comp}B shows that this
-assumption don't hold in practice.
+both strands. However, figure \ref{fig:comp}B shows that this value is
+not representative of the sample locally, therefore the assumption
+doesn't hold in practice.
 
@@
 
-In order to asses the strand imbalance we create the following
-visualization in figure \ref{fig:strand}. A) presents the FSR's
+In order to asses the strand imbalance we created the following
+visualization presented in figure \ref{fig:strand}. A) shows the FSR's
 behavior as lower depth regions are being filtered out, while B) shows
 which percentage of the regions are composed by fragments of both
-strand or only one. Roughly speaking in a lower quality dataset there
-is going to be a higher amount of single stranded-regions. To evaluate
-this statement, for each replicate of the FoxA1 dataset from
-\cite{exoillumina}, we divided the regions formed by at least
-\Sexpr{de} fragments into into the ones that overlap with any of the
-sample's peaks. The p.value for ERR36935 is \Sexpr{pval[[1]]}, for
-ERR336942 is \Sexpr{pval[[2]]} and for ERR336956 is
-\Sexpr{pval[[3]]}. This shows that only the middle sample from figure
-\ref{fig:strand} is the least imbalance sample from the the three,
-while the other two still show a higher imbalance.
+strands or only one (forward or backward). As the number of reads per
+region's lower bound increases, the quantiles tend to approach the
+median. To evaluate this statement, for each replicate of the FoxA1
+dataset, we divided the regions formed by at least \Sexpr{de}
+fragments into into the ones that overlap with any of the sample's
+peaks. The p.value for ERR36935 is \Sexpr{pval[[1]]}, for ERR336942 is
+\Sexpr{pval[[2]]} and for ERR336956 is \Sexpr{pval[[3]]}. This shows
+that only the middle sample from figure \ref{fig:strand} is the least
+imbalance sample from the the three, while the other two still show a
+higher imbalance at the \Sexpr{de} level.
 
@@
     \ref{fig:enrich}. A) FSR distribution quantiles as the lower depth
-    regions are being filtered out. B) Stacked histogram with the
-    proportion of regions that are formed by two strands or only one.}
+    regions are being filtered out, all quantiles approach to the
+    median as the lower bound increases. B) Stacked histogram with the
+    proportion of regions that are formed by two strands or only one,
+    in a good sample the single-stranded regions are going to be
+    filtered out quickly as in the middle row.}
   \label{fig:strand}
@@
 We first compared various factors that could affect binding site
-identification between ChIP-exo and ChIP-Seq data. in order to compare
+identification between ChIP-exo and ChIP-Seq data. In order to compare
 distribution of signal and background between ChIP-exo and ChIP-Seq
@@
     less balanced in the regions with potential binding sites compared
-    to SET ChIP-Seq data. [C and D correspond to fig4 from manuscript]
-    C) ChIP tag counts increase linearly as mappability scores
-    increases. D) ChIP tag counts increase linearly as GC content
-    score increases when GC content is less than 0.6 and then ChIP tag
-    counts decrease as GC content increases. }
+    to SET ChIP-Seq data. C) ChIP tag counts increase linearly as
+    mappability scores increases. D) ChIP tag counts increase linearly
+    as GC content score increases when GC content is less than 0.6 and
+    then ChIP tag counts decrease as GC content increases. }
   \label{fig:comp}
@@
 content. These results indicate that binding site identification in
-ChIP-exo sample might also benefic from the use of methods that take
-into account apparent sequence biases such as mappability and GC
+ChIP-exo sample might also benefit from the use of methods that take
+into account of apparent sequence biases such as mappability and GC
 content.
@@
 Figure \ref{fig:reso_all} shows different comparisons among ChIP-exo,
-PET ChIP-Seq and SET ChIP-Seq. A regulonDB annotation was considered
+PET ChIP-Seq and SET ChIP-Seq. A RegulonDB annotation was considered
 identified if the distance between it and dPeak binding site estimate
 was at most of \Sexpr{ext} bp. That way, the sensitivity is defined as
-the proportion of regulonDB annotations identified in a peak. In
+the proportion of RegulonDB annotations identified in a peak. In
 figure \ref{fig:reso_all}A can be seen that the sensitivity of all
@@
 does. Despite that when the binding events in a peak are closer to
-each other both ChIP-exo and PET ChIP-Seq are comparable, as the
+each other, both ChIP-exo and PET ChIP-Seq are comparable, as the
 distance increases ChIP-exo identifies a higher proportion of the
-regulonDB annotations; additionally SET ChIP-Seq is significantly less
+RegulonDB annotations; additionally SET ChIP-Seq is significantly less
 sensible than both ChIP-exo and PET ChIP-Seq. In figure
-\ref{fig:reso_all}B the distance between a regulonDB annotation to its
+\ref{fig:reso_all}B the distance between a RegulonDB annotation to its
 closest prediction is compared for ChIP-exo, PET and SET ChIP-Seq;
-while the first two are comparable, both outperform SET
-ChIP-Seq.
+while the first two are comparable, both outperform SET ChIP-Seq.
 
@@
 from the 5' end of the fragment to its respective binding event and
-the $\sigma$ parameter indicate the fragments distribution
+the $\sigma$ parameter indicates the fragments distribution
 variability around their respective binding sites (more information in
@@
      ChIP-exo and ChIP-Seq data. Sensitivity is defined as the
-     proportion of regulonDB annotations identified using each
-     data. Resolution is defined as the distance between regulonDB
+     proportion of RegulonDB annotations identified using each
+     data. Resolution is defined as the distance between RegulonDB
      annotation and its closest prediction. $\quad$ C) $\delta$ parameter in
@@
 
-We sampled a fixed among of fragments for each of the ChIP-exo, PET
+We sampled a fixed amount of fragments for each of the ChIP-exo, PET
 ChIP-Seq and SET ChIP-Seq datasets of the $\sigma^{70}$ sample in
@@
 their depth is fixed. It is remarkable that even when the number of
-candidate peaks or the number of predicted events in quite lower than
-for both ChIP-Seq cases, it outperforms them in the number of
-identified targets and in resolution. This may suggest that with
-ChIP-exo less false positive peaks are being called and that when the
-targets are being identified, dPeak is able to estimate binding
-locations closer to the actual true cases. Additionally, we can see
-that as the read depth increases all four indicators do so as well,
-which may indicate that with ChIP-exo a smaller amount of reads is
-necessary to identify a higher number of targets, but it may also be
-possible that this is an artifact occurring due to ChIP-exo's lower
-library complexity.
+candidate peaks or the number of predicted events is lower for
+ChIP-exo, it outperforms both PET and SET ChIP-Seq in the number of
+identified targets and resolution. This may suggest that with ChIP-exo
+less false positive peaks are being called and that when the targets
+are being identified, dPeak estimates binding locations closer to the
+true location. Additionally, we can see that as the read depth
+increases all four indicators do so as well, which may indicate that
+with ChIP-exo a smaller amount of reads is necessary to identify a
+higher number of targets, but it may also be possible that this is an
+artifact occurring due to ChIP-exo's lower library complexity.
 
-\subsection{dPeak outperforms compeing methods in discovering closely
+\subsection{dPeak outperforms competing methods in discovering closely
   space binding event from ChIP-exo and ChIP-Seq data}
@@
   seems to decrease resolution. Also, Apex was updated into MACE, so I
-  think it could be better to add that method instead}
+  think it could be better to add that method instead.}
+
+\emph{Also need to make an equivalente to Fig5 in manuscript}
 
@@
 ChIP-Seq at fixed depth sample is able to identify more targets at a
-lower resolution.
-
-\newpage
+lower resolution. We compared dPeak with another algorithms to
+estimate binding locations in ChIPexo data, dPeak is comparable to
+MACE and outperforms GEM in resolution. dPeak provided a striking
+balance in sensitivity, specificity and spatial resolution for
+ChIP-exo analysis.
 
@@
 \url{http://bioconductor.org/packages/release/bioc/html/CexoR.html},
-but it returned an error everytime we used it.
+but it returned an error every time we used it.
 
@@
 \end{document}
+
+% LocalWords:  ChIP exo URCR

### 2015-11-20 rmarkdown/chip_exo_paper.tex
first draft of paper

@@
-%\VignetteEngine{knitr::knitr}
+% \VignetteEngine{knitr::knitr}
 \documentclass{bmcart}\usepackage[]{graphicx}\usepackage[]{color}
@@
 
-Using the mouse-FoxA1 experiment from \cite{exoillumina}, the
-relationship between this two quantities, library complexty and sample
-enrichment was explored. Figure \ref{fig:enrich}A shows hexbin plots
-illustrating the interaction between these two quantities. Notice how
-there are two strong arms in each panel: The first one corresponds to
-regions with low $\mbox{ARC}$ values and $\mbox{URCR}$ values that can
-vary across the whole $[0,1]$ interval, while the second one shows a
-decreasing trend in $\mbox{URCR}$ as $\mbox{ARC}$ increases. When an
-experiment shows a higher degree of enrichment, then the separation of
-this two arms is more noticeable, since the second arm corresponds to
-possibly enriched regions (\ref{fig:enrich2}A and \ref{fig:enrich2}B).
+Using the mouse-FoxA1 experiment from \cite{exoillumina} and the
+relationship between this two quantities, library complexity and
+sample enrichment was explored. Figure \ref{fig:enrich}A shows hexbin
+plots illustrating the interaction between these two
+quantities. Notice how there are two strong arms in each panel: The
+first one corresponds to regions with low $\mbox{ARC}$ values and
+varying $\mbox{URCR}$ values across the $(0,1]$ interval, while the
+second one shows a decreasing trend in $\mbox{URCR}$ as $\mbox{ARC}$
+increases. When an experiment shows a higher degree of enrichment,
+then the separation of this two arms is more noticeable, since the
+second arm corresponds to possibly enriched regions
+(\ref{fig:enrich2}A and \ref{fig:enrich2}B).
 
-Figure \ref{fig:enrich}B shows boxplots of the $\mbox{local-NSC}$ on
+Figure \ref{fig:enrich}B shows box plots of the $\mbox{local-NSC}$ on
 regions common to the 3 replicates and stratified by the number of
 reads mapped exactly to one position for all $\sigma^{70}$ ChIP-exo
-experiments; The high stratum is defined by regions consising of more
+experiments; The high stratum is defined by regions consisting of more
 than 100 unique positions, the medium stratum for regions
@@
 0.5, which means that there are roughly the same amount of reads in
-both strands. However, figure \ref{fig:comp}B shows that this
-assumption don't hold in practice.
+both strands. However, figure \ref{fig:comp}B shows that this value is
+not representative of the sample locally, therefore the assumption
+doesn't hold in practice.
 
-In order to asses the strand imbalance we create the following
-visualization in figure \ref{fig:strand}. A) presents the FSR's
+In order to asses the strand imbalance we created the following
+visualization presented in figure \ref{fig:strand}. A) shows the FSR's
 behavior as lower depth regions are being filtered out, while B) shows
 which percentage of the regions are composed by fragments of both
-strand or only one. Roughly speaking in a lower quality dataset there
-is going to be a higher amount of single stranded-regions. To evaluate
-this statement, for each replicate of the FoxA1 dataset from
-\cite{exoillumina}, we divided the regions formed by at least
-100 fragments into into the ones that overlap with any of the
-sample's peaks. The p.value for ERR36935 is 0.1039, for
-ERR336942 is 0.0064 and for ERR336956 is
-0.0871. This shows that only the middle sample from figure
-\ref{fig:strand} is the least imbalance sample from the the three,
-while the other two still show a higher imbalance.
+strands or only one (forward or backward). As the number of reads per
+region's lower bound increases, the quantiles tend to approach the
+median. To evaluate this statement, for each replicate of the FoxA1
+dataset, we divided the regions formed by at least 100
+fragments into into the ones that overlap with any of the sample's
+peaks. The p.value for ERR36935 is 0.1039, for ERR336942 is
+0.0064 and for ERR336956 is 0.0871. This shows
+that only the middle sample from figure \ref{fig:strand} is the least
+imbalance sample from the the three, while the other two still show a
+higher imbalance at the 100 level.
 
@@
 Figure \ref{fig:reso_all} shows different comparisons among ChIP-exo,
-PET ChIP-Seq and SET ChIP-Seq. A regulonDB annotation was considered
+PET ChIP-Seq and SET ChIP-Seq. A RegulonDB annotation was considered
 identified if the distance between it and dPeak binding site estimate
 was at most of 20 bp. That way, the sensitivity is defined as
-the proportion of regulonDB annotations identified in a peak. In
+the proportion of RegulonDB annotations identified in a peak. In
 figure \ref{fig:reso_all}A can be seen that the sensitivity of all

### 2015-11-20 rmarkdown/chip_exo_paper.R

@@ -1,6 +1,6 @@
-## ----strata_param , include =FALSE,echo = FALSE, eval = TRUE-------------
+## ----strata_para , include =FALSE,echo = FALSE, veal = TRUE--------------
 
-  nstrata <- 400
+  strata <- 400
   high <- 100
   med <- 50
   low <- 20

### 2015-11-21 rmarkdown/prelims_paper.Rnw
first step of prelim paper

--- 2015-11-19 rmarkdown/chip_exo_paper.tex
@@
-%\VignetteEngine{knitr::knitr}
-\documentclass{bmcart}\usepackage[]{graphicx}\usepackage[]{color}
-%% maxwidth is the original width if it is less than linewidth
-%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
-\makeatletter
-\def\maxwidth{ %
-  \ifdim\Gin@nat@width>\linewidth
-    \linewidth
-  \else
-    \Gin@nat@width
-  \fi
-}
-\makeatother
+\documentclass[11pt]{article}
+%\documentclass[11pt]{scrartcl}
+\usepackage{amsmath, amsfonts, amssymb}
+\usepackage{adjustbox}
+\usepackage{fullpage}
+\usepackage{epsfig}
+%\setkomafont{disposition}{\normalfont\bfseries}
+% \usepackage[round, sort]{natbib}
+\renewcommand{\baselinestretch}{1.4}
+\usepackage{float}
 
-\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
-\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
-\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
-\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
-\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
-\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
-\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
-\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
-\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
-\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
+\title{High Resolution Identification of Protein-DNA Binding Events
+  and Quality Control for ChIP-exo data\vspace*{\fill}}
 
-\usepackage{framed}
-\makeatletter
-\newenvironment{kframe}{%
- \def\at@end@of@kframe{}%
- \ifinner\ifhmode%
-  \def\at@end@of@kframe{\end{minipage}}%
-  \begin{minipage}{\columnwidth}%
- \fi\fi%
- \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
- \colorbox{shadecolor}{##1}\hskip-\fboxsep
-     % There is no \\@totalrightmargin, so:
-     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
- \MakeFramed {\advance\hsize-\width
-   \@totalleftmargin\z@ \linewidth\hsize
-   \@setminipage}}%
- {\par\unskip\endMakeFramed%
- \at@end@of@kframe}
-\makeatother
+\author{Rene Welch\\Preliminary Examination\\Department of Statistics, University of Wisconsin-Madison}
 
-\definecolor{shadecolor}{rgb}{.97, .97, .97}
-\definecolor{messagecolor}{rgb}{0, 0, 0}
-\definecolor{warningcolor}{rgb}{1, 0, 1}
-\definecolor{errorcolor}{rgb}{1, 0, 0}
-\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX
+\date{December 1st, 2015}
 
-\usepackage{alltt}
-
-\usepackage{xcolor}
-\usepackage{url}
-\usepackage{amsmath}
-\usepackage{amsthm}
-\usepackage{amssymb}
-\usepackage{graphicx}
-\usepackage{tikz}
-\usetikzlibrary{shapes,arrows}
-\usepackage{float}
-\usepackage{verbatim}
-\usepackage{hyperref}
-\hypersetup{
-     colorlinks   = true,
-     citecolor    = black
-}
-\hypersetup{linkcolor=black}
-% \hypersetup{linkcolor=black}
-% \hypersetup{linkcolor=black}
-
-% \usepackage{authblk}
-\usepackage[utf8]{inputenc} %unicode support
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+%\usepackage{Sweave}
 \begin{document}
 
-%\maketitle
+\maketitle
 
-\begin{frontmatter}
+\vspace*{\fill}
 
-\begin{fmbox}
-\dochead{Draft}
+\textbf{Committee Members:}
 
-\title{High Resolution Identification of Protein-DNA Binding
-  Events and Quality Control for ChIP-exo data}
+\textbf{Professor S\"und\"uz Kele\c{s}}, Department of Statistics,
+Department of Biostatistics and Medical Informatics
 
-\author[
-   addressref={aff6},                   % id's of addresses, e.g. {aff1,aff2}
-   email={chungd@musc.edu}   % email address
-]{\inits{DC}\fnm{Dongjun} \snm{Chung}}
-\author[
-   addressref={aff1},
-   email={welch@stat.wisc.edu}
-]{\inits{RW}\fnm{Rene} \snm{Welch}}
-\author[
-   addressref={aff3},
-   email={ong@cs.wisc.edu}
-]{\inits{IO}\fnm{Irene} \snm{Ong}}
-\author[
-   addressref={aff3,aff4},
-   email={jagrass@wisc.edu}
-]{\inits{JG}\fnm{Jeffrey} \snm{Grass}}
-\author[
-   addressref={aff3,aff4,aff5},
-   email={landick@bact.wisc.edu}
-]{\inits{RL}\fnm{Robert} \snm{Landick}}
-\author[
-   addressref={aff1,aff2},
-   corref={aff1},
-   email={keles@stat.wisc.edu}
-]{\inits{SK}\fnm{S\"und\"uz} \snm{Kele\c{s}}}
+\textbf{Professor Karl Broman}, Department of Biostatistics and
+Medical Informatics
 
-\address[id=aff1]{%                           % unique id
-  \orgname{Department of Statistics, University of Wisconsin Madison}, % university, etc
-  \street{1300 University Avenue},                     %
-  %\postcode{}                                % post or zip code
-  \city{Madison},                              % city
-  \cny{WI}                                    % country
-}
+\textbf{Professor Colin Dewey}, Department of Computer Sciences,
+Department of Biostatistics and Medical Informatics
 
-\address[id=aff2]{%
-  \orgname{Department of Biostatistics and Medical Informatics, University of Wisconsin Madison},
-  \street{600 Highland Avenue},
-%  \postcode{24105}
-  \city{Madison},
-  \cny{WI}
-}
-\address[id=aff3]{
-  \orgname{Great Lakes Bioenergy Research Center, University of Wisconsin Madison},
-  \street{1552 University Avenue},
-  \city{Madison},
-  \cny{WI}
-}
-\address[id=aff4]{
-  \orgname{Department of Biochemistry, University of Wisconsin Madison},
-  \street{433 Babcock Drive},
-  \city{Madison},
-  \cny{WI}
-}
-\address[id=aff5]{
-  \orgname{Department of Bacteriology, University of Wisconsin Madison},
-  \street{1550 Linden Drive},
-  \city{Madison},
-  \cny{WI}
-}
-\address[id=aff6]{
-  \orgname{Department of Public Health Sciences, Medical University of South Carolina},
-  \street{135 Cannon Street},
-  \city{Charleston},
-  \cny{SC}
-}
+\textbf{Professor Christina Kendziorski}, Department of Biostatistics
+and Medical Informatics
 
-% \begin{artnotes}
-% %\note{Sample of title note}     % note to the article
-% \note[id=n1]{Equal contributor} % note, connected to author
-% \end{artnotes}
+\textbf{Professor Ming Yuan}, Department of Statistics
 
-\end{fmbox}
-
-\begin{abstractbox}
-
-  \begin{abstract}
-
-  Recently, ChIP-exo has been developed to investigate protein-DNA
-  interaction in higher resolution compared to popularly used
-  ChIP-Seq. Although ChIP-exo has drawn much attention and is
-  considered as powerful assay, currently, no systematic studies have
-  yet been conducted to determine optimal strategies for experimental
-  design and analysis of ChIP-exo. In order to address these
-  questions, we evaluated diverse aspects of ChIP-exo and found the
-  following characteristics of ChIP-exo data. First, Background of
-  ChIP-exo data is quite different from that of ChIP-Seq data. However,
-  sequence biases inherently present in ChIP-Seq data still exist in
-  ChIP-exo data. Second, in ChIP-exo data, reads are located around
-  binding sites much more tightly and hence, it has potential for high
-  resolution identification of protein-DNA interaction sites, hence
-  the space to allocate the reads is greatly reduced. Third, although
-  often assumed in the ChIP-exo data analysis methods, the peak pair
-  assumption does not hold well in real ChIP-exo data. Fourth, spatial
-  resolution of ChIP-exo is comparable to that of PET ChIP-Seq and
-  both of them are significantly better than resolution of SET
-  ChIP-Seq. Finally, for given fixed sequencing depth, ChIP-exo
-  provides higher sensitivity, specificity, and spatial resolution
-  than PET ChIP-Seq.
-
-  In this article, we provide a quality control pipeline which
-  visually asses ChIP-exo biases and calculates a signal-to-noise
-  measure. Also, we updated dPeak \cite{dpeak}, which makes a striking
-  balance in sensitivity, specificity, and spatial resolution for
-  ChIP-exo data analysis.
-
-\end{abstract}
-\begin{keyword}
-  \kwd{ChIP-exo}
-  \kwd{QC}
-  \kwd{TFBS}
-  \kwd{BS identification on High-Res}
-\end{keyword}
-
-\end{abstractbox}
-
-\end{frontmatter}
+\thispagestyle{empty}
 
 \newpage
-
-* to remove later
 
@@
 
-\listoffigures
+\section*{Abstract}
+
+    Recently, ChIP-exo has been developed to investigate protein-DNA
+    interaction in higher resolution compared to popularly used
+    ChIP-Seq. Although ChIP-exo has drawn much attention and is
+    considered as powerful assay, currently, no systematic studies
+    have yet been conducted to determine optimal strategies for
+    experimental design and analysis of ChIP-exo. In order to address
+    these questions, we evaluated diverse aspects of ChIP-exo and
+    found the following characteristics of ChIP-exo data. First, the
+    background of ChIP-exo data is quite different from that of
+    ChIP-Seq data. However, sequence biases inherently present in
+    ChIP-Seq data still exist in ChIP-exo data. Second, in ChIP-exo
+    data, reads are located around binding sites much more tightly and
+    hence, it has potential for high resolution identification of
+    protein-DNA interaction sites, and also the space to allocate the
+    reads is greatly reduced. Third, although often assumed in the
+    ChIP-exo data analysis methods, the ``peak pair'' assumption does
+    not hold well in real ChIP-exo data. Fourth, spatial resolution of
+    ChIP-exo is comparable to that of PET ChIP-Seq and both of them
+    are significantly better than resolution of SET ChIP-Seq. Finally,
+    for given fixed sequencing depth, ChIP-exo provides higher
+    sensitivity, specificity and spatial resolution than PET
+    ChIP-Seq.
+
+    In this article we provide a quality control pipeline which
+    visually asses ChIP-exo biases and calculates a signal-to-noise
+    measure. Also, we updated dPeak \cite{dpeak}, which makes a
+    striking balance in sensitivity, specificity and spatial
+    resolution for ChIP-exo data analysis.
+
+\section{Introduction}
+\label{sec:intro}
+
+\section{Statistical Framework for ChIP-exo}
+\label{sec:stat}
+
+\section{Results}
+\label{sec:results}
+
+\section{Planned work}
+\label{sec:future}
 
@@
 
-\section{Background}
-\label{sec:intro}
-
-ChIP-exo (Chromatin Immunopecipitation followed by exonuclease
-digestion and next generation sequencing) Rhee and Pugh (\cite{exo1})
-is the state-of-the-art experiment developed to attain single
-base-pair resolution of protein binding site identification and it is
-considered as a powerful alternative to popularly used ChIP-Seq
-(Chromatin Immunoprecipitation coupled with next generation sequencing
-) assay. ChIP-exo experiments first capture millions of DNA fragments
-(150 - 250 bp in length) that the protein under study interacts with
-using random fragmentation of DNA and a protein-specific
-antibody. Then, exonuclease is introduced to trim 5' end of each DNA
-fragment to a fixed distance from the bound protein. As a result,
-boundaries around the protein of interest constructed with 5' ends of
-fragments are located much closed to bound protein compared to
-ChIP-Seq. This is the step unique to ChIP-exo that could potentially
-provide significantly higher spatial resolution compared to
-ChIP-Seq. Finally, high throughput sequencing of a small region (25 to
-100 bp) at 5' end of each fragment generates millions of reads or
-tags.
-
-While the number of produced ChIP-exo data keeps increasing,
-characteristics of ChIP-exo data and optimal strategies for
-experimental design and analysis of ChIP-exo data are not fully
-investigated yet, including issues of sequence biases inherent to
-ChIP-exo data, choice of optimal statistical methods, and
-determination of optimal sequencing depth. However, currently, the
-number of available ChIP-exo data is still limited and their
-sequencing depths are still insufficient for such investigation. To
-address this limitation we gathered ChIP-exo data from diverse
-organisms: CTCF factor in human \cite{exo1}; ER factor in human and
-FoxA1 factor in mouse from \cite{exoillumina}; and generated
-$\sigma^{70}$ factor in Escherichia coli (E. Coli) under aerobic ($ +
-O_2$) condition, and treated by rifampicin by 0 and 20 minutes.
-
-DNA libraries generated by the ChIP-exo protocol seem to be less
-complex than the libraries generated by ChIP-Seq
-\cite{exo_review}. Hence, most of current guidelines \cite{encode_qc}
-may not be applicable on ChIP-exo, additionally to our knowledge there
-are not established quality control pipelines for ChIP-exo. To address
-this challenge, we suggest a collection of quality control
-visualizations to understand which biases are present in ChIP-exo
-data. Related to quality control, Previous ChIP-exo analysis used
-ChIP-Seq samples to compare the resolution between experiments
-(\cite{exo1}, \cite{exo2}, \cite{exoillumina}). In \cite{carroll.qc},
-Carroll et. al. studied the use of the Strand-Cross Correlation (SCC)
-\cite{strandcc}. and showed that by filtering blacklisted regions the
-estimation of the SCC is improved. However, this method requires to
-know blacklisted regions in advance which may not be available. In our
-pipeline we propose two out-the-shelf metrics equivalent to RSC and
-NSC for the estimation of the signal-to-noise ratio for a ChIP-exo
-sample.
-
-In order to archive the potential benefits of ChIP-exo on protein
-binding site identification, it is critical to understand which are
-the important characteristics of ChIP-exo data and to use algorithms
-that could fully utilize information available in ChIP-exo data. Rhee
-and Pugh \cite{exo1} discussed that reads in the forward and reverse
-strand might construct peak pairs around bound protein, of which
-heights were implicitly assumed to be symmetric. Hence, they used the
-``peak pair method'' that predicts the midpoint of two modes of peak
-pairs as potential binding site. Mace \cite{mace}, CexoR \cite{cexor}
-and peakzilla \cite{peakzilla}, recently developed ChIP-exo data
-analysis methods, are also based on this peak pair
-assumption. However, appropriateness of such assumption was not fully
-evaluated in the literature yet.  Furthermore, it is still unknown
-which factors could affect protein binding site identification using
-ChIP-exo data. In order to address this problem, we investigated
-various aspects of ChIP-exo data by contrasting them with their
-respective ChIP-Seq experiments.
-
-Currently, research on statistical methods for ChIP-exo data is still
-in its very early stage. Although many methods have been proposed to
-identify protein binding sites from ChIP-Seq data (reviewed in
-\cite{evaluation} and \cite{computation}), such as MACS \cite{macs},
-CisGenome \cite{cisgenome} and MOSAiCS \cite{mosaics}, these
-approaches reveal protein binding sites in lower resolution, i.e., at
-an interval of hundreds to thousands of base pairs. Furthermore, they
-report only one ``mode'' or ``predicted binding location'' per
-peak. Hence, these methods are not appropriate to evaluate the
-potential of ChIP-exo data for high resolution identification of
-protein binding sites. More recently, deconvolution algorithms such as
-Deconvolution \cite{csdeconv}, GEM \cite{gem} (an improved version of
-\cite{gps}) and PICS \cite{pics} have been proposed to identify
-binding sites in higher resolution using ChIP-Seq data. However, most
-of them are still not tailored for ChIP-exo and PET and SET ChIP-Seq
-data in a unified framework and as a result, currently available
-methods are not appropriate for fair comparison between ChIP-exo and
-ChIP-Seq. To address these limitations, we developed an improved of
-dPeak \cite{dpeak}, a high resolution binding site identification
-(deconvolution) algorithm that we previously developed for PET and SET
-ChIP-Seq data, so that it can also handle ChIP-exo data. The dPeak
-algorithm implements a probabilistic model that accurately describes
-the ChIP-exo and ChIP-Seq data generation process.
-
-In this paper, we demonstrate that the peak pair assumption of Rhee
-and Pugh \cite{exo1} does not hold well in real ChIP-exo
-data. Furthermore, we found that when we analyze ChIP-exo data from
-eukaryotic genomes, it is important to consider sequence biases
-inherent to ChIP-exo data, such as mappability and GC content in order
-to improve sensitivity and specificity of binding site
-identification. dPeak outperforms or performs competitively with
-ChIP-exo data analysis such as GEM and MACE. More importantly, when
-comparable number of reads is used for both ChIP-exo and ChIP-Seq,
-dPeak couple with ChIP-exo data provides resolution comparable to PET
-ChIP-Seq and both significantly improve the resolution of protein
-binding site identification compared to SET-based analysis with any of
-the available methods.
-
-\newpage
-
-\section{Results and discussion}
-\label{sec:results}
-
-\subsection{Deeply sequenced E.Coli $\sigma^{70}$ ChIP-exo and ChIP-Seq data}
-\label{sec:ourdata}
-
-$\sigma^{70}$ factor is a transcription initiation factor of
-housekeeping genesin E. coli. In E. coli genomes, many promoters
-contain multiple transcription start sites (TSS) and these TSS are
-often closely space (10 ~ 150 bp). These closely spaced binding sites
-are considered to be multiple ``switches'' that differentially
-regulate gene expression under diverse growth conditions
-\cite{regdb_old}, \cite{regulondb}. Therefore, investigation of
-ChIP-exo's potential for identification and differentiation of closely
-spaced binding sites are invaluable for elucidating the
-transcriptional networks of prokaryotic genomes.
-
-\subsection{Current guidelines and quality metrics}
-\label{sec:qcnow}
-
-To have a first view on the data, we calculated the most commonly used
-quality indicators for ChIP-Seq data, which are listed in table
-\ref{tab:qcbase}. The PBC, NSC and RSC were calculated as defined in
-\cite{encode_qc} and the SSD was calculated as defined in
-\cite{htseq}. Additionally, we calculated the FSR to compare the
-number of fragments in each strand. A few conclusions that we may get
-by observing the PBC in table \ref{tab:qcbase} are all the samples
-have a quite low library complexity, hence it would be suggested to
-repeat the experiment.
-
-\begin{table}[h]
-  \centering
-\begin{tabular}{l|r|r|r|r|r|r}
-\hline\hline
-Condition & Rep. & Nr. reads & PBC & SSD & FSR & NSC  \\
-\hline\hline
-Aerobic & 1 & 1,454,566 & 0.2156 & 0.0830 & 0.5012 & 114.14 \\
-\hline
-Aerobic & 2 & 864,714 & 0.2471 & 0.0488 & 0.4923 & 160.93  \\
-\hline
-Rif-0min & 1 & 960,256 & 0.2823 & 0.0361 & 0.4998 & 10.29 \\
-\hline
-Rif-0min & 2 & 2,247,295 & 0.2656 & 0.1091 & 0.5047 &  25.08  \\
-\hline
-Rif-20min & 1 & 1,940,387 & 0.2698 & 0.0820 & 0.5054 & 17.69  \\
-\hline
-Rif-20min & 2 & 4,229,574 & 0.2153 & 0.1647 & 0.5071 &  14.11 \\
-\hline
-\end{tabular}  
-\caption{Usual quality control indicators applied to our
-  $\sigma^{70}$ samples. PBC stands PCR-bottleneck coefficient, SSD
-  for standardized standard deviation, FSR for forward strand ratio,
-  NSC for normalized strand cross-correlation coefficient. We
-  omitted the RSC (relative strand cross-correlation since ChIP-exo
-  experiments don't usually have an accompanying input file) 
-  [note: I need to change the table, due the alignment issue]}
-
-  \label{tab:qcbase}
-\end{table}
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = .8\textwidth]{../figs/for_paper/EColi_strand_cross_corr.pdf}
-  \caption{Strand cross-correlation curves for $\sigma^{70}$
-    samples. The ``phantom peak'' and the summit that corresponds to
-    the read and fragment length respectively are confounded, due to
-    the exonuclease enzyme digestion.}
-  \label{figsup:scc}
-\end{figure}
-
-In a good ChIP-Seq experiment the strand cross-correlation is
-maximized at the fragment length and there is also a ``phantom peak''
-at the sample's read length (\cite{encode_qc} and \cite{strandcc}).
-However, in ChIP-exo's case this two peaks are confounded. Hence an
-enrichment measure based in the strand cross-correlation such as NSC
-or RSC is harder to interpret. In figure \ref{figsup:scc} we plotted
-the SCC curves for $\sigma^{70}$ ChIP-exo experiments used to
-calculate the NSC and RSC values in table \ref{tab:qcbase}. According
-to ENCODE's quality metrics, NSC values close to one corresponds to
-low enrichment samples or bad quality datasets, hence in our case the
-last four samples would be considered to have higher quality than they
-actually have. On the other hand, all PBC values are considerably low
-and correspond to samples with severe bottlenecking issues, in
-particular we are going to show that for the first two samples we are
-obtaining a smaller resolution and identifying a comparable number of
-targets as in ChIP-Seq PET experiments.
-
-\subsection{ChIP-exo Quality Control pipeline}
-\label{sec:QC}
-
-Figure \ref{fig:qcdiagram} shows a flowchart for the ChIPexoQC
-pipeline. Which basically partitions genome by keeping the
-non-digested ChIP-exo regions. Then, for each region it calculates a
-series of independent statistics. Finally, it creates several
-visualizations designed to asses the overall quality levels of A
-ChIP-exo sample.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = .9\textwidth]{../figs/for_paper/coverage_diagram.pdf}
-  \caption{Diagram of the ChIP-exo Quality control pipeline [this is a
-    place-holder]. The genome is partitioned into several regions by
-    removing the empty segments and for each regions we calculate a
-    collection of qc indicators.}
-  \label{fig:qcdiagram}
-\end{figure}
-
-\subsubsection{Enrichment and library complexity in ChIP-exo data}
-\label{sec:enri}
-
-In ChIP-exo experiments, the background is often digested by the
-exonuclease enzyme, therefore to determine the sample's quality is
-necessary to address the balance between the enrichment and library
-complexity. To diagnose this, we considered the Average Read Coverage
-(ARC) and the Unique Read Coverage Rate (URCR) which are defined as:
-
-\begin{align}
-  \mbox{ARC} &= \frac{\text{Nr. of reads in the region}}{\text{Width of the region}} \nonumber \\
-  \mbox{URCR} &= \frac{\text{Nr. of reads mapped to only one position
-      in the region}}{\text{Nr. of reads in the region}} \nonumber
-\end{align}
-
-Using the mouse-FoxA1 experiment from \cite{exoillumina}, the
-relationship between this two quantities, library complexty and sample
-enrichment was explored. Figure \ref{fig:enrich}A shows hexbin plots
-illustrating the interaction between these two quantities. Notice how
-there are two strong arms in each panel: The first one corresponds to
-regions with low $\mbox{ARC}$ values and $\mbox{URCR}$ values that can
-vary across the whole $[0,1]$ interval, while the second one shows a
-decreasing trend in $\mbox{URCR}$ as $\mbox{ARC}$ increases. When an
-experiment shows a higher degree of enrichment, then the separation of
-this two arms is more noticeable, since the second arm corresponds to
-possibly enriched regions (\ref{fig:enrich2}A and \ref{fig:enrich2}B).
-
-Figure \ref{fig:enrich}B shows boxplots of the $\mbox{local-NSC}$ on
-regions common to the 3 replicates and stratified by the number of
-reads mapped exactly to one position for all $\sigma^{70}$ ChIP-exo
-experiments; The high stratum is defined by regions consising of more
-than 100 unique positions, the medium stratum for regions
-where the number of unique position is on the
-(50,100) range and the low stratum in the
-(20,50) range. For each stratum and biological
-replicate, we sampled 400 regions whenever possible (in
-the opposite case, all the regions in the category were considered)
-and calculated the $\mbox{local-NSC}$ for those regions. For all three
-replicates, it is shown an increasing trend as the number of unique
-position increases, which means that the $\mbox{local-NSC}$ is an
-effective indicator of library complexity. Figure \ref{fig:enrich}C
-shows the $\mbox{local-NSC}$ coefficient calculated for the same set
-of regions, stratified by the number of unique positions in the
-regions of ERR336942 using the same criteria as before. In the high
-panel it is shown that the ERR336942 sample performs slightly better
-than ERR336956 and both outperform ERR336935, this agrees with
-\ref{fig:enrich}A, where only in the left-most panel doesn't seem to
-be a separation between the low $\mbox{ARC}$ arm and the other.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = .8\textwidth]{../figs/Carroll_mice_for_paper/FoxA1_enrichment.pdf}
-  \includegraphics[width = .8\textwidth,page =3]{../figs/Carroll_mice_for_paper/Local_SCC_indicator_by_strata.pdf}
-  \includegraphics[width = .8\textwidth,page =4]{../figs/Carroll_mice_for_paper/Local_SCC_indicator_by_strata_sameRegion.pdf}
-  \caption{Using the mouse-FoxA1 experiment from \cite{exoillumina}:
-    A) Hexbin plots of $\mbox{ARC}$ against $\mbox{URCR}$, in general
-    we can see a slight separation into two strong arms, one
-    corresponds to low $\mbox{ARC}$ and varying $\mbox{URCR}$, and for
-    the other $\mbox{URCR}$ decreases as $\mbox{ARC}$ increases B)
-    Boxplots of the $\mbox{local-NSC}$ stratified by nr. of reads
-    mapped to only one position in the regions for the ChIP-exo
-    experiments divided by condition and biological replicate and C)
-    Boxplots of the a set of ERR336942 regions, stratified as in (B)
-    with the $\mbox{local-NSC}$ calculated with the different
-    biological replicates.}
-  \label{fig:enrich}
-\end{figure}
-
-\subsubsection{Strand imbalance in ChIP-exo data}
-\label{sec:strand_imbalance}
-
-The strand imbalance assessment is based in the observation that the
-enriched regions usually have a higher concentration of fragments,
-therefore we examine the FSR (defined as the ratio of the number of
-forward stranded reads divided by the total number of reads in a given
-region) as the region with lower depth are being filtered out. This
-indicator is of particular importance, since several methods rely on
-the ``peak-pair'' assumption. In table \ref{tab:qcbase}, we calculated
-the FSR and noticed that for all the samples, it's value is close to
-0.5, which means that there are roughly the same amount of reads in
-both strands. However, figure \ref{fig:comp}B shows that this
-assumption don't hold in practice.
-
-In order to asses the strand imbalance we create the following
-visualization in figure \ref{fig:strand}. A) presents the FSR's
-behavior as lower depth regions are being filtered out, while B) shows
-which percentage of the regions are composed by fragments of both
-strand or only one. Roughly speaking in a lower quality dataset there
-is going to be a higher amount of single stranded-regions. To evaluate
-this statement, for each replicate of the FoxA1 dataset from
-\cite{exoillumina}, we divided the regions formed by at least
-100 fragments into into the ones that overlap with any of the
-sample's peaks. The p.value for ERR36935 is 0.1039, for
-ERR336942 is 0.0064 and for ERR336956 is
-0.0871. This shows that only the middle sample from figure
-\ref{fig:strand} is the least imbalance sample from the the three,
-while the other two still show a higher imbalance.
-
-\begin{figure}[h!]
-  \centering  
-  \includegraphics[width = .8\textwidth,page = 3]{../figs/Carroll_mice_for_paper/Strand_imbalance.pdf} 
-  \caption{Strand imbalance QC plots for the same data as in figure
-    \ref{fig:enrich}. A) FSR distribution quantiles as the lower depth
-    regions are being filtered out. B) Stacked histogram with the
-    proportion of regions that are formed by two strands or only one.}
-  \label{fig:strand}
-\end{figure}
-
-\subsection{Comparison with ChIP-Seq data}
-\label{sec:comp}
-
-We first compared various factors that could affect binding site
-identification between ChIP-exo and ChIP-Seq data. in order to compare
-distribution of signal and background between ChIP-exo and ChIP-Seq
-data, we calculated ChIP tag counts across the genome by counting the
-number of reads mapping to each of 150 non-overlapping
-window after extending reads by 150 to their 3' end
-directions. ChIP tag counts in ChIP-exo data were linearly related to
-ChIP tag counts in ChIP-Seq data for the regions with high ChIP tag
-counts (Figure \ref{fig:comp}A). This implies that signals for
-potential binding sites are well reproducible between ChIP-exo and
-ChIP-Seq data. On the other hand, there was clear difference in the
-background distribution between them. In ChIP-Seq data reads were
-almost uniformly distributed over background (non-binding) regions and
-the ChIP tag counts in there regions were significantly larger than
-zero. In contrast, in ChIP-exo data, there was larger variation in
-ChIP tag counts among background regions and ChIP tag counts were much
-lower in these regions compared to ChIP-Seq data. There were also
-large proportion of regions without any read in ChIP-exo data. These
-results indicate that background distribution of ChIP-exo data is less
-homogeneous than that of ChIP-Seq data.
-
-We next evaluated the ``peak pair'' assumption of Rhee and Pugh
-\cite{exo1}, MACE \cite{mace}, CexoR, \cite{cexor} and peakzilla
-\cite{peakzilla}, i.e. a peak of reads in the forward strand is
-usually paired with a peak of reads in the reverse strand that is
-located in the other site of the binding site. In order to evaluate
-this assumption, we reviewed the proportion of reads in the forward
-strand in candidate regions (i.e. regions with at least one binding
-site) in $\sigma^{70}$ ChIP-exo data. We found that strands of reads
-were much less balanced in ChIP-exo data than in ChIP-Seq data in
-these regions with potential binding sites (Fig. \ref{fig:comp}B) and
-this indicates that the peak pair assumption might not hold in real
-ChIP-exo data.
-
-%  We obtained esentially the same conclusion from the
-% ChIP-exo data for CTCF factor from human sample Rhee and Pugh
-% (\cite{exo1}) as well.
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width = .46\textwidth,page = 3 ]{../figs/for_paper/ChIPseqPET_ChIPexo_tagCount_comparison.pdf}
-  \includegraphics[width = .46\textwidth]{../figs/for_paper/forward_strand_ratio_comp_old.pdf}
-  \caption{ A) ChIP signal in ChIP-exo was linearly related to that in
-    PET ChIP-Seq data in the region with high ChIP tag counts. In
-    contrast, there were clear differences in their background
-    distribution, where several ChIP-exo background regions were
-    empty. B) In ChIP-exo data, strands of reads were significantly
-    less balanced in the regions with potential binding sites compared
-    to SET ChIP-Seq data. [C and D correspond to fig4 from manuscript]
-    C) ChIP tag counts increase linearly as mappability scores
-    increases. D) ChIP tag counts increase linearly as GC content
-    score increases when GC content is less than 0.6 and then ChIP tag
-    counts decrease as GC content increases. }
-  \label{fig:comp}
-\end{figure}
-
-\subsubsection{ChIP-exo data from eukaryotic genome shows clear sequence biases}
-\label{sec:eukaryotic}
-
-We evaluated ChIP-exo data for CTCF factor from human genome
-\cite{exo1} to investigate issues specific to eukaryotic genomes for
-binding sites identification. Figures \ref{fig:comp}C and
-\ref{fig:comp}D display the bin-level average read counts against
-mappability and GC content. Each data point is obtained by averaging
-the read counts across bins with the same mappability of GC
-content. These results indicate that binding site identification in
-ChIP-exo sample might also benefic from the use of methods that take
-into account apparent sequence biases such as mappability and GC
-content.
-
-\subsubsection{Comparison with ChIP-Seq data using dPeak}
-\label{sec:dpeak_analysis}
-
-Figure \ref{fig:reso_all} shows different comparisons among ChIP-exo,
-PET ChIP-Seq and SET ChIP-Seq. A regulonDB annotation was considered
-identified if the distance between it and dPeak binding site estimate
-was at most of 20 bp. That way, the sensitivity is defined as
-the proportion of regulonDB annotations identified in a peak. In
-figure \ref{fig:reso_all}A can be seen that the sensitivity of all
-protocols increases as the average distance between binding sites
-does. Despite that when the binding events in a peak are closer to
-each other both ChIP-exo and PET ChIP-Seq are comparable, as the
-distance increases ChIP-exo identifies a higher proportion of the
-regulonDB annotations; additionally SET ChIP-Seq is significantly less
-sensible than both ChIP-exo and PET ChIP-Seq. In figure
-\ref{fig:reso_all}B the distance between a regulonDB annotation to its
-closest prediction is compared for ChIP-exo, PET and SET ChIP-Seq;
-while the first two are comparable, both outperform SET
-ChIP-Seq.
-
-In figures \ref{fig:reso_all}C and \ref{fig:reso_all}D, we observe the
-behavior of dPeak's estimated parameters for single end data (both
-ChIP-exo and ChIP-Seq). The $\delta$ parameter measures the distance
-from the 5' end of the fragment to its respective binding event and
-the $\sigma$ parameter indicate the fragments distribution
-variability around their respective binding sites (more information in
-\cite{dpeak}). For both parameters, it can be seen that the ChIP-exo
-estimates are smaller than the SET ChIP-Seq estimates in average,
-which agrees with the fact that the sample reads are allocated more
-tightly around the binding events in ChIP-exo data. Hence we can
-conclude that the peaks shape is quite different between ChIP-exo and
-SET ChIP-Seq.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = .46\textwidth]{../figs/for_paper/sensitivity_exo_olda_data.pdf}
-  \includegraphics[width = .46\textwidth]{../figs/for_paper/resolution_by_dataset_old_data.pdf}
-   \includegraphics[width = .46\textwidth,page = 1]{../figs/for_paper/sigma_delta_old_densities.pdf}
-   \includegraphics[width = .46\textwidth,page = 2]{../figs/for_paper/sigma_delta_old_densities.pdf}
-
-   \caption{Comparison of (A) sensitivity and (B) resolution between
-     ChIP-exo and ChIP-Seq data. Sensitivity is defined as the
-     proportion of regulonDB annotations identified using each
-     data. Resolution is defined as the distance between regulonDB
-     annotation and its closest prediction. $\quad$ C) $\delta$ parameter in
-     dPeak measures average distance of the reads to their respective
-     binding site. In ChIP-exo data, reads were located much closer to
-     the binding site than in SET ChIP-Seq. D) $\sigma$ parameter
-     measure the dispersion of reads around each binding site. In
-     ChIP-exo data, reads showed less variation around the their
-     respective binding sites compared to SET ChIP-Seq.}
-  \label{fig:reso_all}
-\end{figure}
-
-\subsection{Recommendations for the design of ChIP-exo experiments}
-\label{sec:reco}
-
-We sampled a fixed among of fragments for each of the ChIP-exo, PET
-ChIP-Seq and SET ChIP-Seq datasets of the $\sigma^{70}$ sample in
-aerobic conditions. For each sampled dataset we applied our
-lower-to-higher resolution pipeline by calling peaks with MOSAiCS
-\cite{mosaics} and then deconvolving the binding events by using dPeak
-\cite{dpeak}. For the ChIP-exo datasets we called peaks by using
-GC-content and mappability with MOSAiCS, and for the ChIP-Seq datasets
-we used their respective Input samples.
-
-\begin{figure}[h!]
-  \centering
-%  \includegraphics[width = .8\textwidth]{../figs/for_paper/Sig70_aerobic_saturation.pdf}
-  \includegraphics[width = .8\textwidth]{../figs/for_paper/saturation_analysis_old.pdf}
-  \caption{Comparison of the number of candidate regions (A),
-    predicted events (B), identified targets (C) and resolution (D)
-    among ChIP-exo, PET ChIP-Seq and SET ChIP-Seq. RegulonDB
-    annotations are considered as a gold standard. A gold standard was
-    considered identified if a BS was estimated inside a 15
-    bp window around it}
-  \label{fig:design}
-\end{figure}
-
-Figure \ref{fig:design} shows the behavior of each data type when
-their depth is fixed. It is remarkable that even when the number of
-candidate peaks or the number of predicted events in quite lower than
-for both ChIP-Seq cases, it outperforms them in the number of
-identified targets and in resolution. This may suggest that with
-ChIP-exo less false positive peaks are being called and that when the
-targets are being identified, dPeak is able to estimate binding
-locations closer to the actual true cases. Additionally, we can see
-that as the read depth increases all four indicators do so as well,
-which may indicate that with ChIP-exo a smaller amount of reads is
-necessary to identify a higher number of targets, but it may also be
-possible that this is an artifact occurring due to ChIP-exo's lower
-library complexity.
-
-\subsection{dPeak outperforms compeing methods in discovering closely
-  space binding event from ChIP-exo and ChIP-Seq data}
-\label{sec:dpeak_comp}
-
-\emph{I plan to update this analysis, since the dpeak's new strategy
-  seems to decrease resolution. Also, Apex was updated into MACE, so I
-  think it could be better to add that method instead}
-
-\section{Conclusions}
-\label{sec:conclusions}
-
-\newpage
-
-\section{Methods}
-\label{sec:methods}
-
-\color{red}
-
-\subsection{Growth conditions}
-\label{sec:growth}
-
-\subsection{ChIP-exo experiments}
-\label{sec:experiments}
-
-\subsection{Library preparation, sequencing and mapping of sequencing reads}
-\label{sec:library}
-
-\color{black}
-
-\subsection{Method comparison with ChIP-exo}
-\label{sec:suppcomp}
-
-We considered dPeak Chung et al. \cite{dpeak}, GEM Guo et
-al. \cite{gem} and MACE Wang et al. \cite{mace} for ChIP-exo data
-analysis. For the dPeak algorithm, we used the R package \emph{dPeak}
-version 1.0.0, which is available from
-\url{http://dongjunchung.github.io/dpeak/}. For the GEM algorithm we
-used its Java implementation version 0.9, which is available from
-\url{http://cgs.csail.mit.edu/gem/}. For the MACE algorithm, we used
-its Python implementation version 1.0, which is available from
-\url{http://dldcc-web.brc.bcm.edu/lilab/MACE/docs/html/}. For the
-PEAKZILLA algorithm, we used its Python implementation, which is
-available in \url{http://www.starklab.org/data/peakzilla/}. Candidate
-regions for dPeak were identified for each replicate of ChIP-exo data
-using the MOSAiCS algorithm Kuan et al. \cite{mosaics} (one sample
-analysis using false discovery rate of 0.0001) implemented as an R
-package \emph{mosaics} 2.4 (available from \emph{Bioconductor}
-\url{https://www.bioconductor.org/packages/release/bioc/html/mosaics.html}). We
-further filtered out candidate region with average ChIP tag count less
-than 3,000 to avoid potential false positives based in exploratory
-analysis. These regions were also explicitly provided to the GEM
-algorithm as candidate regions. Default tuning parameters were used
-during model fitting for all methods. We tried to use the CexoR
-algorithm with its R implementation version 1.8.0 available from
-\url{http://bioconductor.org/packages/release/bioc/html/CexoR.html},
-but it returned an error everytime we used it.
-
-\subsection{Local-NSC}
-\label{sec:localnsc}
-
-The local strand cross-correlation is defined as:
-
-\begin{align}
-  \mbox{local-NSC} &= \frac{max_{x_\delta} f(x_\delta)}{\sigma} \nonumber \\
- y_\delta &= f(x_\delta) + \sigma \epsilon_\delta \nonumber
-\end{align}
-
-where $f$ is calculated by fitting a local polynomial regression model
-over the strand cross-correlation calculated using exclusively the
-fragments that overlap a pre-defined genomic region, $y_\delta$ is the
-correlation of the forward and backward coverages when one is shifted
-$x_\delta$ bp towards their 3' end.
-
-The local polynomial model was fitted using the \emph{loess} function
-from the R software version 3.2.1 using the default tuning parameters.
-
-\newpage
-
-\bibliographystyle{bmc-mathphys} % Style BST file (bmc-mathphys, vancouver, spbasic).
+\bibliographystyle{plain} % Style BST file (bmc-mathphys, vancouver, spbasic).
 \bibliography{chip_exo_paper}
 
-\nocite{exo_gb}
-\nocite{maplot1}
-\nocite{maplot2}
-% \nocite{esl}
-
-\newpage
-
-\section*{Supplement}
-\label{sec:supp}
-
-\subsection*{Additional Enrichment plots for $\sigma^{70}$}
-\label{sec:enrichsup}
-
-In figure \ref{fig:enrich2} it is shown the same relationship as in
-figure \ref{fig:enrich}, but considering only regions formed where the
-reads are being allocated in more than 10 (A) and 30 (B) unique
-positions respectively. Both plots show that the vertical arms formed
-by regions with low $\mbox{ARC}$ is formed by low complexity regions,
-that way suggesting that this segment correspond to the background of
-a ChIP-exo experiment.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = .8\textwidth,page = 3]{../figs/Carroll_mice_for_paper/FoxA1_enrichment.pdf}
-  \includegraphics[width = .8\textwidth,page = 4]{../figs/Carroll_mice_for_paper/FoxA1_enrichment.pdf}
-  \caption{Hexbin plots of $\mbox{ARC}$ vs $\mbox{URCR}$ for each
-    region after partitioning the genome. In A and B the regions with
-    reads mapped to at most 10 and 30 positions respectively where not
-    considered. $\mbox{ARC}$ is defined as the ratio of the nr. of
-    reads and the width of a region and $\mbox{URCR}$ is the ratio of
-    the number of unique position where the reads are being allocated
-    and the number of reads in a region.}
-  \label{fig:enrich2}
-\end{figure}
-
 \end{document}

### 2015-11-21 rmarkdown/prelims_paper.tex
first step of prelim paper

--- 2015-11-19 rmarkdown/chip_exo_paper.tex
@@
-%\VignetteEngine{knitr::knitr}
-\documentclass{bmcart}\usepackage[]{graphicx}\usepackage[]{color}
+\documentclass[11pt]{article}\usepackage[]{graphicx}\usepackage[]{color}
 %% maxwidth is the original width if it is less than linewidth
@@
 \usepackage{alltt}
+%\documentclass[11pt]{scrartcl}
+\usepackage{amsmath, amsfonts, amssymb}
+\usepackage{adjustbox}
+\usepackage{fullpage}
+\usepackage{epsfig}
+%\setkomafont{disposition}{\normalfont\bfseries}
+% \usepackage[round, sort]{natbib}
+\renewcommand{\baselinestretch}{1.4}
+\usepackage{float}
 
-\usepackage{xcolor}
-\usepackage{url}
-\usepackage{amsmath}
-\usepackage{amsthm}
-\usepackage{amssymb}
-\usepackage{graphicx}
-\usepackage{tikz}
-\usetikzlibrary{shapes,arrows}
-\usepackage{float}
-\usepackage{verbatim}
-\usepackage{hyperref}
-\hypersetup{
-     colorlinks   = true,
-     citecolor    = black
-}
-\hypersetup{linkcolor=black}
-% \hypersetup{linkcolor=black}
-% \hypersetup{linkcolor=black}
+\title{High Resolution Identification of Protein-DNA Binding Events
+  and Quality Control for ChIP-exo data\vspace*{\fill}}
 
-% \usepackage{authblk}
-\usepackage[utf8]{inputenc} %unicode support
+\author{Rene Welch\\Preliminary Examination\\Department of Statistics, University of Wisconsin-Madison}
+
+\date{December 1st, 2015}
+
+%\usepackage{Sweave}
 \IfFileExists{upquote.sty}{\usepackage{upquote}}{}
@@
 
-%\maketitle
+\maketitle
 
-\begin{frontmatter}
+\vspace*{\fill}
 
-\begin{fmbox}
-\dochead{Draft}
+\textbf{Committee Members:}
 
-\title{High Resolution Identification of Protein-DNA Binding
-  Events and Quality Control for ChIP-exo data}
+\textbf{Professor S\"und\"uz Kele\c{s}}, Department of Statistics,
+Department of Biostatistics and Medical Informatics
 
-\author[
-   addressref={aff6},                   % id's of addresses, e.g. {aff1,aff2}
-   email={chungd@musc.edu}   % email address
-]{\inits{DC}\fnm{Dongjun} \snm{Chung}}
-\author[
-   addressref={aff1},
-   email={welch@stat.wisc.edu}
-]{\inits{RW}\fnm{Rene} \snm{Welch}}
-\author[
-   addressref={aff3},
-   email={ong@cs.wisc.edu}
-]{\inits{IO}\fnm{Irene} \snm{Ong}}
-\author[
-   addressref={aff3,aff4},
-   email={jagrass@wisc.edu}
-]{\inits{JG}\fnm{Jeffrey} \snm{Grass}}
-\author[
-   addressref={aff3,aff4,aff5},
-   email={landick@bact.wisc.edu}
-]{\inits{RL}\fnm{Robert} \snm{Landick}}
-\author[
-   addressref={aff1,aff2},
-   corref={aff1},
-   email={keles@stat.wisc.edu}
-]{\inits{SK}\fnm{S\"und\"uz} \snm{Kele\c{s}}}
+\textbf{Professor Karl Broman}, Department of Biostatistics and
+Medical Informatics
 
-\address[id=aff1]{%                           % unique id
-  \orgname{Department of Statistics, University of Wisconsin Madison}, % university, etc
-  \street{1300 University Avenue},                     %
-  %\postcode{}                                % post or zip code
-  \city{Madison},                              % city
-  \cny{WI}                                    % country
-}
+\textbf{Professor Colin Dewey}, Department of Computer Sciences,
+Department of Biostatistics and Medical Informatics
 
-\address[id=aff2]{%
-  \orgname{Department of Biostatistics and Medical Informatics, University of Wisconsin Madison},
-  \street{600 Highland Avenue},
-%  \postcode{24105}
-  \city{Madison},
-  \cny{WI}
-}
-\address[id=aff3]{
-  \orgname{Great Lakes Bioenergy Research Center, University of Wisconsin Madison},
-  \street{1552 University Avenue},
-  \city{Madison},
-  \cny{WI}
-}
-\address[id=aff4]{
-  \orgname{Department of Biochemistry, University of Wisconsin Madison},
-  \street{433 Babcock Drive},
-  \city{Madison},
-  \cny{WI}
-}
-\address[id=aff5]{
-  \orgname{Department of Bacteriology, University of Wisconsin Madison},
-  \street{1550 Linden Drive},
-  \city{Madison},
-  \cny{WI}
-}
-\address[id=aff6]{
-  \orgname{Department of Public Health Sciences, Medical University of South Carolina},
-  \street{135 Cannon Street},
-  \city{Charleston},
-  \cny{SC}
-}
+\textbf{Professor Christina Kendziorski}, Department of Biostatistics
+and Medical Informatics
 
-% \begin{artnotes}
-% %\note{Sample of title note}     % note to the article
-% \note[id=n1]{Equal contributor} % note, connected to author
-% \end{artnotes}
+\textbf{Professor Ming Yuan}, Department of Statistics
 
-\end{fmbox}
-
-\begin{abstractbox}
-
-  \begin{abstract}
-
-  Recently, ChIP-exo has been developed to investigate protein-DNA
-  interaction in higher resolution compared to popularly used
-  ChIP-Seq. Although ChIP-exo has drawn much attention and is
-  considered as powerful assay, currently, no systematic studies have
-  yet been conducted to determine optimal strategies for experimental
-  design and analysis of ChIP-exo. In order to address these
-  questions, we evaluated diverse aspects of ChIP-exo and found the
-  following characteristics of ChIP-exo data. First, Background of
-  ChIP-exo data is quite different from that of ChIP-Seq data. However,
-  sequence biases inherently present in ChIP-Seq data still exist in
-  ChIP-exo data. Second, in ChIP-exo data, reads are located around
-  binding sites much more tightly and hence, it has potential for high
-  resolution identification of protein-DNA interaction sites, hence
-  the space to allocate the reads is greatly reduced. Third, although
-  often assumed in the ChIP-exo data analysis methods, the peak pair
-  assumption does not hold well in real ChIP-exo data. Fourth, spatial
-  resolution of ChIP-exo is comparable to that of PET ChIP-Seq and
-  both of them are significantly better than resolution of SET
-  ChIP-Seq. Finally, for given fixed sequencing depth, ChIP-exo
-  provides higher sensitivity, specificity, and spatial resolution
-  than PET ChIP-Seq.
-
-  In this article, we provide a quality control pipeline which
-  visually asses ChIP-exo biases and calculates a signal-to-noise
-  measure. Also, we updated dPeak \cite{dpeak}, which makes a striking
-  balance in sensitivity, specificity, and spatial resolution for
-  ChIP-exo data analysis.
-
-\end{abstract}
-\begin{keyword}
-  \kwd{ChIP-exo}
-  \kwd{QC}
-  \kwd{TFBS}
-  \kwd{BS identification on High-Res}
-\end{keyword}
-
-\end{abstractbox}
-
-\end{frontmatter}
+\thispagestyle{empty}
 
 \newpage
-
-* to remove later
 

### 2015-11-21 rmarkdown/chip_exo_paper.Rnw
Changed figure in paper

@@
 Figure \ref{fig:enrich}B shows box plots of the $\mbox{local-NSC}$ on
-regions common to the 3 replicates and stratified by the number of
-reads mapped exactly to one position for all $\sigma^{70}$ ChIP-exo
+regions of the three replicates, stratified by the number of reads
+mapped exactly to one position for all $\sigma^{70}$ ChIP-exo
 experiments; The high stratum is defined by regions consisting of more
@@
 (\Sexpr{low},\Sexpr{med}) range. For each stratum and biological
-replicate, we sampled \Sexpr{strata} regions whenever possible (in
-the opposite case, all the regions in the category were considered)
-and calculated the $\mbox{local-NSC}$ for those regions. For all three
+replicate, we sampled \Sexpr{strata} regions whenever possible (in the
+opposite case, all the regions in the category were considered) and
+calculated the $\mbox{local-NSC}$ for those regions. For all three
 replicates, it is shown an increasing trend as the number of unique

### 2015-11-21 rmarkdown/chip_exo_paper.tex
Changed figure in paper

@@
 (20,50) range. For each stratum and biological
-replicate, we sampled 400 regions whenever possible (in
-the opposite case, all the regions in the category were considered)
-and calculated the $\mbox{local-NSC}$ for those regions. For all three
+replicate, we sampled 400 regions whenever possible (in the
+opposite case, all the regions in the category were considered) and
+calculated the $\mbox{local-NSC}$ for those regions. For all three
 replicates, it is shown an increasing trend as the number of unique

### 2015-11-21 rscripts/carrol_mice_library_complexity_and_enrichment.R

@@ -160,7 +160,11 @@ build_stats <- function(region,reads)
   return(stats)
 }
 
-stats <- mclapply(gr,function(x)build_stats(common,x),mc.cores = 6)
+regs <- lapply(regs,dt2gr)
+
+stats <- mcmapply(build_stats,regs,gr,mc.cores = 6 ,SIMPLIFY = FALSE)
+
+stats_common <- mclapply(gr,function(x)build_stats(common,x),mc.cores = 6)
 
 aux <- mapply(function(x,y)x[,sample := y],stats,names(stats),SIMPLIFY =FALSE)
 aux <- do.call(rbind,aux)
@@ -346,6 +350,12 @@ strata_plots[[4]] <- ggplot(dat , aes( file , nsc,colour = strata))+geom_boxplot
   scale_color_brewer(palette = "Dark2")+theme_bw()+
   theme(legend.position = "none", axis.text.x = element_text(angle = 90),plot.title = element_text(hjust = 0))+
   xlab("")+ylab("Local NSC")+ylim(0,3)+geom_abline( slope = 0 , intercept = 0, linetype = 2)+ggtitle("B")
+
+pdf(file = file.path(figs_dir,"Local_SCC_indicator_by_strata.pdf"),width = 9,height = 4)
+u <- lapply(strata_plots,print)
+dev.off()
+
+
 
 anchor_stats <- filter_stats[[2]]
 

### 2015-11-21 rmarkdown/chip_exo_paper.Rnw
Changed asses to assess in paper

@@
     In this article we provide a quality control pipeline which
-    visually asses ChIP-exo biases and calculates a signal-to-noise
+    visually assess ChIP-exo biases and calculates a signal-to-noise
     measure. Also, we updated dPeak \cite{dpeak}, which makes a
@@
 series of statistics. Finally, it creates several visualizations
-designed to asses the quality levels of a ChIP-exo sample.
+designed to assess the quality levels of a ChIP-exo sample.
 
@@
 
-In order to asses the strand imbalance we created the following
+In order to assess the strand imbalance we created the following
 visualization presented in figure \ref{fig:strand}. A) shows the FSR's
@@
 
-We provide a ChIP-exo QC pipeline capable of asses the balance between
+We provide a ChIP-exo QC pipeline capable of assess the balance between
 enriched samples and low complexity regions. It is shown that the
 ``peak-pair'' assumption doesn't hold in practice and we provide two
-out-of-the-box visualization capable to asses the strand imbalance in
+out-of-the-box visualization capable to assess the strand imbalance in
 a ChIP-exo experiment.

### 2015-11-22 rmarkdown/prelims_paper.Rnw
Second version of prelims paper, added example about chip exo and diagram of chip exo's protocol

@@
 \usepackage{epsfig}
+
+\usepackage{multicol}
 %\setkomafont{disposition}{\normalfont\bfseries}
@@
 \usepackage{float}
+
+\usepackage{wrapfig}
 
@@
 
+%% commmands
 %\usepackage{Sweave}
+
 \begin{document}
+
+\newcommand{\sig}{\sigma^{70}}
 
@@
 
+\listoffigures
+
+\newpage
+
 \section*{Abstract}
@@
     ChIP-Seq. Although ChIP-exo has drawn much attention and is
-    considered as powerful assay, currently, no systematic studies
+    considered as powerful assay, currently no systematic studies
     have yet been conducted to determine optimal strategies for
@@
 
-    In this article we provide a quality control pipeline which
-    visually asses ChIP-exo biases and calculates a signal-to-noise
-    measure. Also, we updated dPeak \cite{dpeak}, which makes a
-    striking balance in sensitivity, specificity and spatial
-    resolution for ChIP-exo data analysis.
+    We provide a quality control pipeline which visually assess
+    ChIP-exo biases, library complexity and enrichment; and calculates
+    a signal-to-noise measure. Also, we updated dPeak (Chung et al.,
+    2012 \cite{dpeak}), which makes a striking balance in sensitivity,
+    specificity and spatial resolution for ChIP-exo data analysis.
+
+\newpage
 
@@
 
-\section{Statistical Framework for ChIP-exo}
-\label{sec:stat}
+ChIP-exo (Chromatin Immunopecipitation followed by exonuclease
+digestion and next generation sequencing) Rhee and Pugh, 2011
+(\cite{exo1}) is the state-of-the-art experiment developed to attain
+single base-pair resolution of protein binding site identification and
+it is considered as a powerful alternative to popularly used ChIP-Seq
+(Chromatin Immunoprecipitation coupled with next generation sequencing
+) assay.
+
+While the number of produced ChIP-exo data keeps increasing,
+characteristics of ChIP-exo data and optimal strategies for
+experimental design and analysis of ChIP-exo data are not fully
+investigated yet, including issues of sequence biases inherent to
+ChIP-exo data, choice of optimal statistical methods, and
+determination of optimal sequencing depth. However, currently the
+number of available ChIP-exo data is still limited and their
+sequencing depths are still insufficient for such investigation. To
+address this limitation we gathered ChIP-exo data from diverse
+organisms: CTCF factor in human \cite{exo1}; ER factor in human and
+FoxA1 factor in mouse (Serandour et al., 2013 \cite{exoillumina}); and
+generated $\sig$ factor in Escherichia coli (E. Coli) under
+aerobic ($ + O_2$) condition, and treated by rifampicin by 0 and 20
+minutes.
+
+DNA libraries generated by the ChIP-exo protocol seem to be less
+complex than the libraries generated by ChIP-Seq (Mahony et al., 2015
+\cite{exo_review}). Hence, most of current QC guidelines (Landt et
+al., 2012 \cite{encode_qc}) may not be applicable on ChIP-exo,
+additionally to our knowledge there are not established quality
+control pipelines for ChIP-exo. To address this challenge, we suggest
+a collection of quality control visualizations to understand which
+biases are present in ChIP-exo data. Previous ChIP-exo analysis used
+ChIP-Seq samples to compare the resolution between experiments
+(\cite{exo1}, \cite{exo2}, \cite{exoillumina}); Carroll et al., 2014
+\cite{carroll.qc} studied the use of the Strand-Cross Correlation
+(SCC) (Kharchenko et al., 2008 \cite{strandcc}) and showed that by
+filtering blacklisted regions the estimation of the SCC is
+improved. However, this method requires to know blacklisted regions in
+advance which may not be available, additionally using SCC may not be
+useful since the peaks that are attained at the read and fragment
+length are confused in a typical ChIP-exo SCC curve.  In our pipeline
+we propose two out-the-shelf analysis: an enrichment plot and the
+local normalized SCC cofficient.
+
+In order to archive the potential benefits of ChIP-exo on protein
+binding site identification, it is critical to understand which are
+the important characteristics of ChIP-exo data and to use algorithms
+that could fully utilize information available in ChIP-exo data. Rhee
+and Pugh, 2011 \cite{exo1} discussed that reads in the forward and
+reverse strand might construct peak pairs around bound protein, of
+which heights were implicitly assumed to be symmetric. Hence, they
+used the ``peak pair method'' that predicts the midpoint of two modes
+of peak pairs as potential binding site. Mace (Wang et al., 2014
+\cite{mace}), CexoR (Madrigal, 2015 \cite{cexor}) and peakzilla
+(Bardet et al., 2013 \cite{peakzilla}), recently developed ChIP-exo
+data analysis methods, are also based on this peak pair
+assumption. However, appropriateness of such assumption was not fully
+evaluated in the literature yet.  Furthermore, it is still unknown
+which factors could affect protein binding site identification using
+ChIP-exo data. In order to address this problem, we investigated
+various aspects of ChIP-exo data by contrasting them with their
+respective ChIP-Seq experiments.
+
+Currently, research on statistical methods for ChIP-exo data is still
+in its very early stage. Although many methods have been proposed to
+identify protein binding sites from ChIP-Seq data (reviewed by
+Wilbanks and Facciotti, 2012 \cite{evaluation} and Pepke and Wold,
+2009 \cite{computation}), such as MACS (Zhang et al., 2008
+\cite{macs}), CisGenome (Hongkai et al., 2008 \cite{cisgenome}) and
+MOSAiCS (Kuan et al., 2009 \cite{mosaics}), these approaches reveal
+protein binding sites in lower resolution, i.e., at an interval of
+hundreds to thousands of base pairs. Furthermore, they report only one
+``mode'' or ``predicted binding location'' per peak. Hence, these
+methods are not appropriate to evaluate the potential of ChIP-exo data
+for high resolution identification of protein binding sites. More
+recently, deconvolution algorithms such as Deconvolution (Lun et al.,
+2009 \cite{csdeconv}), GEM (Guo et al., 2012 \cite{gem}, an improved
+version of Guo et al., 2010 \cite{gps} ) and PICS (Zhang et al., 2010
+\cite{pics}) have been proposed to identify binding sites in higher
+resolution using ChIP-Seq data. However, most of them are still not
+tailored for ChIP-exo and PET and SET ChIP-Seq data in a unified
+framework and as a result, currently available methods are not
+appropriate for fair comparison between ChIP-exo and ChIP-Seq. To
+address these limitations, we developed an improved of dPeak (Chung et
+al., 2013 \cite{dpeak}), a high resolution binding site identification
+(deconvolution) algorithm that we previously developed for PET and SET
+ChIP-Seq data, so that it can also handle ChIP-exo data. The dPeak
+algorithm implements a probabilistic model that accurately describes
+the ChIP-exo and ChIP-Seq data generation process.
+
+In this work, we demonstrate that the ``peak pair'' assumption of
+Rhee and Pugh \cite{exo1} does not hold well in real ChIP-exo
+data. Furthermore, we found that when we analyze ChIP-exo data from
+eukaryotic genomes, it is important to consider sequence biases
+inherent to ChIP-exo data, such as mappability and GC content in order
+to improve sensitivity and specificity of binding site
+identification. We evaluated several method to identify binding events
+and dPeak outperforms or performs competitively respecto to GEM and
+MACE when analyzing ChIP-exo data. More importantly, when comparable
+number of reads is used for both ChIP-exo and ChIP-Seq, dPeak coupled
+with ChIP-exo data provides resolution comparable to PET ChIP-Seq and
+both significantly improve the resolution of protein binding site
+identification compared to SET-based analysis with any of the
+available methods.
+
+\section{ChIP-exo's review}
+\label{sec:exo}
+
+\begin{wrapfigure}{l}{.4\textwidth}
+\centering
+\includegraphics{../figs/diagrams/ChIPexo_diagram.png}
+\caption{ChIP-exo diagram. The only difference with ChIP-Seq is the
+  second step, where an exonuclease enzyme is added and it diggest the
+  DNA fragments starting from the 5' end until it finds a crosslink
+  protein (this figure is taken from Furey, 2012 \cite{chipbeyond}).}
+\label{fig:exo}
+\end{wrapfigure}
+
+ChIP-exo is based in a modification to the ChIP-seq protocol, where an
+exonuclease enzyme is added and it diggests the fragments starting by
+the 5' ends and stops until it finds a protein that is crosslinked to
+the DNA.
+
+ChIP-exo experiments first capture millions of DNA fragments
+(150 - 250 bp in length) that the protein under study interacts with
+using random fragmentation of DNA and a protein-specific
+antibody. Then, exonuclease is introduced to trim 5' end of each DNA
+fragment to a fixed distance from the bound protein. As a result,
+boundaries around the protein of interest constructed with 5' ends of
+fragments are located much closed to bound protein compared to
+ChIP-Seq. This step is unique to ChIP-exo and could potentially
+provide significantly higher spatial resolution compared to
+ChIP-Seq. Finally, high throughput sequencing of a small region (25 to
+100 bp) at 5' end of each fragment generates millions of reads or
+tags.
+
+Since this protocol is a modification to ChIP-Seq, some of its
+characteristics are still maintained in ChIP-exo while several new are
+being found. Rhee and Pugh, 2011 \cite{exo1}, showed that ChIP-exo
+generated fragments are located around the binding events more tightly
+than ChIP-Seq generated fragments. Hence, it is of utmost importance
+to understand how this new protocol works, specifically we focused on
+understanding which biases are mantained from ChIP-Seq to ChIP-exo,
+which biases are specific to ChIP-exo, how the current ChIP-Seq QC
+guidelines behave in ChIP-exo and we developed a new QC pipeline
+specific to ChIP-exo. A typical example is shown in figure
+\ref{fig:exo_example}, which shows the $\sig$ and $\beta'_f$ factors
+(and the $\beta$ factor only on top) on a E. Coli genome's region:
+First, both regions show that the $\sig$ sample is enriched; second
+the $\sig$ sample in the ChIP-exo panel shows that its respective form
+in the ChIP-Seq panel may be form of more binding events; and third,
+either of the $\beta$ factors show that the complexity of the ChIP-exo
+libraries may be lower in some cases, i.e. the peak is mapped with
+fewer positions.
+
+\subsection{Current Quality Control Measures for ChIP-Seq}
+\label{sec:QC_seq}
+
+To our knowledge it doesn't exists a QC pipeline for ChIP-exo.
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width = .8 \textwidth,page = 107]{/p/keles/ChIPexo/volume5/OpenClosedComplexes/figs/peaks/ChIPseq_ChIPexo_peaks_rif20.pdf}
+  \caption{An example of a PET ChIP-Seq peak compared to
+    ChIP-exo. $\sig$, $\beta$ and $\beta_f'$ are subunits of
+    the transcription initiation complex in E. Coli. Gene regions on
+    the reverse strand are highlighted in light blue and the region
+    was centered respect to ChIP-Seq $\sig$ summit. The library
+    complexity is lower in ChIP-exo than in ChIP-Seq but the
+    resolution is increased, this can be observed at the $\sig$
+    peak that for ChIP-exo was created as the combination of at least
+    three ChIP-exo ``sub-peaks''.}
+  \label{fig:exo_example}
+\end{figure}
 
@@
 \bibliography{chip_exo_paper}
+\nocite{exo_gb}
+\nocite{maplot1}
+\nocite{maplot2}
 

### 2015-11-22 rmarkdown/prelims_paper.tex
Second version of prelims paper, added example about chip exo and diagram of chip exo's protocol

@@
 
+%% commmands
 %\usepackage{Sweave}
@@
 \begin{document}
+
+\newcommand{\sig}{\sigma^{70}}
 
@@
 
-\section{Statistical Framework for ChIP-exo}
-\label{sec:stat}
+ChIP-exo (Chromatin Immunopecipitation followed by exonuclease
+digestion and next generation sequencing) Rhee and Pugh, 2011
+(\cite{exo1}) is the state-of-the-art experiment developed to attain
+single base-pair resolution of protein binding site identification and
+it is considered as a powerful alternative to popularly used ChIP-Seq
+(Chromatin Immunoprecipitation coupled with next generation sequencing
+) assay.
+
+While the number of produced ChIP-exo data keeps increasing,
+characteristics of ChIP-exo data and optimal strategies for
+experimental design and analysis of ChIP-exo data are not fully
+investigated yet, including issues of sequence biases inherent to
+ChIP-exo data, choice of optimal statistical methods, and
+determination of optimal sequencing depth. However, currently the
+number of available ChIP-exo data is still limited and their
+sequencing depths are still insufficient for such investigation. To
+address this limitation we gathered ChIP-exo data from diverse
+organisms: CTCF factor in human \cite{exo1}; ER factor in human and
+FoxA1 factor in mouse (Serandour et al., 2013 \cite{exoillumina}); and
+generated $\sig$ factor in Escherichia coli (E. Coli) under
+aerobic ($ + O_2$) condition, and treated by rifampicin by 0 and 20
+minutes.
+
+DNA libraries generated by the ChIP-exo protocol seem to be less
+complex than the libraries generated by ChIP-Seq (Mahony et al., 2015
+\cite{exo_review}). Hence, most of current QC guidelines (Landt et
+al., 2012 \cite{encode_qc}) may not be applicable on ChIP-exo,
+additionally to our knowledge there are not established quality
+control pipelines for ChIP-exo. To address this challenge, we suggest
+a collection of quality control visualizations to understand which
+biases are present in ChIP-exo data. Previous ChIP-exo analysis used
+ChIP-Seq samples to compare the resolution between experiments
+(\cite{exo1}, \cite{exo2}, \cite{exoillumina}); Carroll et al., 2014
+\cite{carroll.qc} studied the use of the Strand-Cross Correlation
+(SCC) (Kharchenko et al., 2008 \cite{strandcc}) and showed that by
+filtering blacklisted regions the estimation of the SCC is
+improved. However, this method requires to know blacklisted regions in
+advance which may not be available, additionally using SCC may not be
+useful since the peaks that are attained at the read and fragment
+length are confused in a typical ChIP-exo SCC curve.  In our pipeline
+we propose two out-the-shelf analysis: an enrichment plot and the
+local normalized SCC cofficient.
+
+In order to archive the potential benefits of ChIP-exo on protein
+binding site identification, it is critical to understand which are
+the important characteristics of ChIP-exo data and to use algorithms
+that could fully utilize information available in ChIP-exo data. Rhee
+and Pugh, 2011 \cite{exo1} discussed that reads in the forward and
+reverse strand might construct peak pairs around bound protein, of
+which heights were implicitly assumed to be symmetric. Hence, they
+used the ``peak pair method'' that predicts the midpoint of two modes
+of peak pairs as potential binding site. Mace (Wang et al., 2014
+\cite{mace}), CexoR (Madrigal, 2015 \cite{cexor}) and peakzilla
+(Bardet et al., 2013 \cite{peakzilla}), recently developed ChIP-exo
+data analysis methods, are also based on this peak pair
+assumption. However, appropriateness of such assumption was not fully
+evaluated in the literature yet.  Furthermore, it is still unknown
+which factors could affect protein binding site identification using
+ChIP-exo data. In order to address this problem, we investigated
+various aspects of ChIP-exo data by contrasting them with their
+respective ChIP-Seq experiments.
+
+Currently, research on statistical methods for ChIP-exo data is still
+in its very early stage. Although many methods have been proposed to
+identify protein binding sites from ChIP-Seq data (reviewed by
+Wilbanks and Facciotti, 2012 \cite{evaluation} and Pepke and Wold,
+2009 \cite{computation}), such as MACS (Zhang et al., 2008
+\cite{macs}), CisGenome (Hongkai et al., 2008 \cite{cisgenome}) and
+MOSAiCS (Kuan et al., 2009 \cite{mosaics}), these approaches reveal
+protein binding sites in lower resolution, i.e., at an interval of
+hundreds to thousands of base pairs. Furthermore, they report only one
+``mode'' or ``predicted binding location'' per peak. Hence, these
+methods are not appropriate to evaluate the potential of ChIP-exo data
+for high resolution identification of protein binding sites. More
+recently, deconvolution algorithms such as Deconvolution (Lun et al.,
+2009 \cite{csdeconv}), GEM (Guo et al., 2012 \cite{gem}, an improved
+version of Guo et al., 2010 \cite{gps} ) and PICS (Zhang et al., 2010
+\cite{pics}) have been proposed to identify binding sites in higher
+resolution using ChIP-Seq data. However, most of them are still not
+tailored for ChIP-exo and PET and SET ChIP-Seq data in a unified
+framework and as a result, currently available methods are not
+appropriate for fair comparison between ChIP-exo and ChIP-Seq. To
+address these limitations, we developed an improved of dPeak (Chung et
+al., 2013 \cite{dpeak}), a high resolution binding site identification
+(deconvolution) algorithm that we previously developed for PET and SET
+ChIP-Seq data, so that it can also handle ChIP-exo data. The dPeak
+algorithm implements a probabilistic model that accurately describes
+the ChIP-exo and ChIP-Seq data generation process.
+
+In this work, we demonstrate that the ``peak pair'' assumption of
+Rhee and Pugh \cite{exo1} does not hold well in real ChIP-exo
+data. Furthermore, we found that when we analyze ChIP-exo data from
+eukaryotic genomes, it is important to consider sequence biases
+inherent to ChIP-exo data, such as mappability and GC content in order
+to improve sensitivity and specificity of binding site
+identification. We evaluated several method to identify binding events
+and dPeak outperforms or performs competitively respecto to GEM and
+MACE when analyzing ChIP-exo data. More importantly, when comparable
+number of reads is used for both ChIP-exo and ChIP-Seq, dPeak coupled
+with ChIP-exo data provides resolution comparable to PET ChIP-Seq and
+both significantly improve the resolution of protein binding site
+identification compared to SET-based analysis with any of the
+available methods.
+
+\section{ChIP-exo's review}
+\label{sec:exo}
+
+\begin{wrapfigure}{l}{.4\textwidth}
+\centering
+\includegraphics{../figs/diagrams/ChIPexo_diagram.png}
+\caption{ChIP-exo diagram. The only difference with ChIP-Seq is the
+  second step, where an exonuclease enzyme is added and it diggest the
+  DNA fragments starting from the 5' end until it finds a crosslink
+  protein (this figure is taken from Furey, 2012 \cite{chipbeyond}).}
+\label{fig:exo}
+\end{wrapfigure}
+
+ChIP-exo is based in a modification to the ChIP-seq protocol, where an
+exonuclease enzyme is added and it diggests the fragments starting by
+the 5' ends and stops until it finds a protein that is crosslinked to
+the DNA.
+
+ChIP-exo experiments first capture millions of DNA fragments
+(150 - 250 bp in length) that the protein under study interacts with
+using random fragmentation of DNA and a protein-specific
+antibody. Then, exonuclease is introduced to trim 5' end of each DNA
+fragment to a fixed distance from the bound protein. As a result,
+boundaries around the protein of interest constructed with 5' ends of
+fragments are located much closed to bound protein compared to
+ChIP-Seq. This step is unique to ChIP-exo and could potentially
+provide significantly higher spatial resolution compared to
+ChIP-Seq. Finally, high throughput sequencing of a small region (25 to
+100 bp) at 5' end of each fragment generates millions of reads or
+tags.
+
+Since this protocol is a modification to ChIP-Seq, some of its
+characteristics are still maintained in ChIP-exo while several new are
+being found. Rhee and Pugh, 2011 \cite{exo1}, showed that ChIP-exo
+generated fragments are located around the binding events more tightly
+than ChIP-Seq generated fragments. Hence, it is of utmost importance
+to understand how this new protocol works, specifically we focused on
+understanding which biases are mantained from ChIP-Seq to ChIP-exo,
+which biases are specific to ChIP-exo, how the current ChIP-Seq QC
+guidelines behave in ChIP-exo and we developed a new QC pipeline
+specific to ChIP-exo. A typical example is shown in figure
+\ref{fig:exo_example}, which shows the $\sig$ and $\beta'_f$ factors
+(and the $\beta$ factor only on top) on a E. Coli genome's region:
+First, both regions show that the $\sig$ sample is enriched; second
+the $\sig$ sample in the ChIP-exo panel shows that its respective form
+in the ChIP-Seq panel may be form of more binding events; and third,
+either of the $\beta$ factors show that the complexity of the ChIP-exo
+libraries may be lower in some cases, i.e. the peak is mapped with
+fewer positions.
+
+\subsection{Current Quality Control Measures for ChIP-Seq}
+\label{sec:QC_seq}
+
+Despite that not specific QC pipeline exists for ChIP-exo.
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width = .8 \textwidth,page = 107]{/p/keles/ChIPexo/volume5/OpenClosedComplexes/figs/peaks/ChIPseq_ChIPexo_peaks_rif20.pdf}
+  \caption{An example of a PET ChIP-Seq peak compared to
+    ChIP-exo. $\sig$, $\beta$ and $\beta_f'$ are subunits of
+    the transcription initiation complex in E. Coli. Gene regions on
+    the reverse strand are highlighted in light blue and the region
+    was centered respect to ChIP-Seq $\sig$ summit. The library
+    complexity is lower in ChIP-exo than in ChIP-Seq but the
+    resolution is increased, this can be observed at the $\sig$
+    peak that for ChIP-exo was created as the combination of at least
+    three ChIP-exo ``sub-peaks''.}
+  \label{fig:exo_example}
+\end{figure}
 

### 2015-11-22 rmarkdown/chip_exo_paper.Rnw
Second version of prelims paper, added example about chip exo and diagram of chip exo's protocol
Added conclusion and sections about comparisons realized with dpeak

@@
 length are confused in a typical ChIP-exo SCC curve.  In our pipeline
-we propose two out-the-shelf analysis: an enrichment plot and the
+we propose two out-the-shelf analysis: An enrichment plot and the
 local normalized SCC cofficient.
@@
 \hline\hline
-Aerobic & 1 & 1,454,566 & 0.2156 & 0.0830 & 0.5012 & 114.14 \\
-\hline
-Aerobic & 2 & 864,714 & 0.2471 & 0.0488 & 0.4923 & 160.93  \\
-\hline
 Rif-0min & 1 & 960,256 & 0.2823 & 0.0361 & 0.4998 & 10.29 \\
@@
 
-We next evaluated the ``peak pair'' assumption of Rhee and Pugh
+We next evaluated the ``peak pair'' assumption from Rhee and Pugh
 \cite{exo1}, MACE \cite{mace}, CexoR, \cite{cexor} and peakzilla
@@
 package \emph{mosaics} 2.4 (available from \emph{Bioconductor}
-\url{https://www.bioconductor.org/packages/release/bioc/html/mosaics.html}). We
-further filtered out candidate region with average ChIP tag count less
-than 3,000 to avoid potential false positives based in exploratory
-analysis. These regions were also explicitly provided to the GEM
-algorithm as candidate regions. Default tuning parameters were used
-during model fitting for all methods. We tried to use the CexoR
+\url{https://www.bioconductor.org/packages/release/bioc/html/mosaics.html}).
+We further filtered out candidate region with average ChIP tag count
+less than 3,000 to avoid potential false positives based in
+exploratory analysis. These regions were also explicitly provided to
+the GEM algorithm as candidate regions. Default tuning parameters were
+used during model fitting for all methods. We tried to use the CexoR
 algorithm with its R implementation version 1.8.0 available from
@@
 \nocite{maplot2}
+\nocite{chipbeyond}
 % \nocite{esl}

### 2015-11-22 rmarkdown/prelims_paper.Rnw
Added conclusion and sections about comparisons realized with dpeak

@@
 \usepackage{epsfig}
+\usepackage{hyperref}
+\hypersetup{
+     colorlinks   = true,
+     citecolor    = black
+}
+\hypersetup{linkcolor=black}
 
 \usepackage{multicol}
-%\setkomafont{disposition}{\normalfont\bfseries}
-% \usepackage[round, sort]{natbib}
+
 \renewcommand{\baselinestretch}{1.4}
 \usepackage{float}
-
 \usepackage{wrapfig}
@@
 
-\section{ChIP-exo's review}
+\section{ChIP-exo and ChIP-Seq review}
 \label{sec:exo}
@@
 
-\subsection{Current Quality Control Measures for ChIP-Seq}
-\label{sec:QC_seq}
-
-To our knowledge it doesn't exists a QC pipeline for ChIP-exo.
-
 \begin{figure}[h!]
@@
 
+\subsection{Current Quality Control Measures for ChIP-Seq}
+\label{sec:QC_chipseq}
+
+To our knowledge it doesn't exists a QC pipeline for ChIP-exo. Hence,
+we analyzed how the current ChIP-Seq QC measures behave on ChIP-exo
+data.
+
+\begin{table}[h]
+  \centering
+  \caption{Usual quality control indicators applied to the $\sigma^{70}$ samples.
+PBC stands PCR-bottleneck coefficient, SSD for standardized standard deviation, NSC for normalized strand cross-correlation coefficient (Another measure related to the strand cross-correlation curve is the RSC was ommited since input sample usually don't exists for ChIP-exo).
+$\sig$ samples were given by Robert Landick, FoxA1 and ER samples are from Serandour et al., 2013 \cite{exoillumina} and CTCF sample is from Rhee and Pugh, 2011 \cite{exo1}.}
+
+\begin{tabular}{l|l|r|r|r|r|r|r}
+\hline\hline
+IP & Organism  &Condition & Rep. & Nr. reads & PBC & SSD &  NSC  \\
+\hline\hline
+$\sig$ & E.Coli & Rif-0min & 1 & 960,256 & 0.2823 & 0.0361 &  10.29 \\
+\hline
+$\sig$ &  E.Coli & Rif-0min & 2 & 2,247,295 & 0.2656 & 0.1091  &  25.08  \\
+\hline
+$\sig$ &  E.Coli & Rif-20min & 1 & 1,940,387 & 0.2698 & 0.0820  & 17.69  \\
+\hline
+$\sig$ &  E.Coli & Rif-20min & 2 & 4,229,574 & 0.2153 & 0.1647  &  14.11 \\
+\hline
+FoxA1 &  Mouse &  NA & 1 & 22,210,461 & 0.6562 & 9.12 $\times 10^{-5}$ & 21.452 \\
+\hline
+FoxA1 &  Mouse &  NA & 2 & 22,307,557 & 0.7996 & 7.94 $\times 10^{-5}$ & 60.661 \\
+\hline
+FoxA1 &  Mouse &  NA & 3 & 22,421,729 & 0.1068 & 1.31 $\times 10^{-4}$ & 72.312 \\
+\hline
+ER & Human & NA & 1 & 9,289,835 & 0.8082 & 3.64 $\times 10^{-5}$ & 19.843 \\
+\hline
+ER & Human & NA & 2 & 11,041,833 & 0.8024 & 4.6 $\times 10^{-5}$ & 21.422 \\
+\hline
+ER & Human & NA & 3 & 12,464,836 & 0.8203 & 4.89 $\times 10^{-5}$ & 19.699 \\
+\hline
+CTCF & Human & NA & 1 &   48,478,450 & 0.4579 & 1.29 $\times 10^{-4}$ & 15.977 \\
+\hline
+\end{tabular}  
+  \label{tab:qcbase}
+\end{table}
+
+Table \ref{tab:qcbase} contains measures that are commonly used with
+ChIP-Seq data. PBC stands for PCR bottleneck amplification and is
+defined as the ratio between the number of positions to which exactly
+one mappping read maps divided by the number of position to which at
+least one mapping reads maps, the ENCODE project proposed a series of
+ranges to interpret this quantity (which can be found in
+\url{https://www.encodeproject.org/data-standards/2012-quality-metrics/}),
+but it doesn't apply to ChIP-exo, if it did all the samples with
+$\mbox{PBC} < 0.5$ should be reject due to be showing severe
+bottlenecking. Planet et al., 2011 proposed the Standarized Standard
+Deviation which defined as:
+
+\[
+\mbox{SSD} = \mbox{SD}_N / \sqrt{N} 
+\]
+
+where $\mbox{SD}_N$ is the standard deviation of the coverage of the
+sample and $N$ is the number of reads. This measure is smaller for
+ChIP-exo since the reads are located more tightly around binding
+sites, but it may be biased since several fragments are being
+diggested by exonuclease enzyme, hence there are several position in
+the genome where no fragment is being mapped. Finally the normalized
+strand cross-correlation is defined as:
+
+\[
+\mbox{NSC} = \frac{ \max }{\min}
+\]
+
+\subsection{Comparison with ChIP-Seq data}
+\label{sec:comp_chipseq}
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width = .46\textwidth,page = 3 ]{../figs/for_paper/ChIPseqPET_ChIPexo_tagCount_comparison.pdf}
+  \includegraphics[width = .46\textwidth]{../figs/for_paper/forward_strand_ratio_comp_old.pdf}
+  \includegraphics[width = .46\textwidth,page = 1]{../figs/for_paper/eukaryotic_bias_CTCF.pdf}
+  \includegraphics[width = .46\textwidth,page = 2]{../figs/for_paper/eukaryotic_bias_CTCF.pdf}
+  \caption{ A) ChIP signal in ChIP-exo was linearly related to that in
+    PET ChIP-Seq data in the region with high ChIP tag counts. In
+    contrast, there were clear differences in their background
+    distribution, where several ChIP-exo background regions were
+    empty. B) In ChIP-exo data, strands of reads were significantly
+    less balanced in the regions with potential binding sites compared
+    to SET ChIP-Seq data. C) ChIP tag counts increase linearly as
+    mappability scores increases. D) ChIP tag counts increase linearly
+    as GC content score increases when GC content is less than 0.6 and
+    then ChIP tag counts decrease as GC content increases. }
+  \label{fig:comp}
+\end{figure}
+
+<<comp_param,include = FALSE,echo = FALSE,eval = TRUE>>=
+
+  bin_size <- 150
+  frag_len <- 150
+
+@
+
+We first compared various factors that could affect binding site
+identification between ChIP-exo and ChIP-Seq data. In order to compare
+distribution of signal and background between ChIP-exo and ChIP-Seq
+data, we calculated ChIP tag counts across the genome by counting the
+number of reads mapping to each of \Sexpr{bin_size} non-overlapping
+window after extending reads by \Sexpr{frag_len} to their 3' end
+directions. ChIP tag counts in ChIP-exo data were linearly related to
+ChIP tag counts in ChIP-Seq data for the regions with high ChIP tag
+counts (Figure \ref{fig:comp}A). This implies that signals for
+potential binding sites are well reproducible between ChIP-exo and
+ChIP-Seq data. On the other hand, there was clear difference in the
+background distribution between them. In ChIP-Seq data reads were
+almost uniformly distributed over background (non-binding) regions and
+the ChIP tag counts in there regions were significantly larger than
+zero. In contrast, in ChIP-exo data, there was larger variation in
+ChIP tag counts among background regions and ChIP tag counts were much
+lower in these regions compared to ChIP-Seq data. There were also
+large proportion of regions without any read in ChIP-exo data. These
+results indicate that background distribution of ChIP-exo data is less
+homogeneous than that of ChIP-Seq data.
+
+We next evaluated the ``peak pair'' assumption from Rhee and Pugh,
+2011 \cite{exo1}, i.e. a peak of reads in the forward strand is
+usually paired with a peak of reads in the reverse strand that is
+located in the other site of the binding site. Wang et al., 2014
+\cite{mace}, Madrigal 2015 \cite{cexor} and Bardet et al., 2013
+\cite{peakzilla} proposed method rely in this assumption. In order to
+evaluate this assumption, we reviewed the proportion of reads in the
+forward strand in candidate regions (i.e. regions with at least one
+binding site) in $\sig$ ChIP-exo data. We found that strands of reads
+were much less balanced in ChIP-exo data than in ChIP-Seq data in
+these regions with potential binding sites (Fig. \ref{fig:comp}B) and
+this indicates that the peak pair assumption might not hold in real
+ChIP-exo data.
+
+We evaluated ChIP-exo data for CTCF factor from human genome
+\cite{exo1} to investigate issues specific to eukaryotic genomes for
+binding sites identification. Figures \ref{fig:comp}C and
+\ref{fig:comp}D display the bin-level average read counts against
+mappability and GC content. Each data point is obtained by averaging
+the read counts across bins with the same mappability of GC
+content. These results indicate that binding site identification in
+ChIP-exo sample might also benefit from the use of methods that take
+into account of apparent sequence biases such as mappability and GC
+content.
+
 \section{Results}
 \label{sec:results}
+
+\subsubsection{Comparison with ChIP-Seq data using dPeak}
+\label{sec:dpeak_analysis}
+
+<<reso_param, include = FALSE , eval  = TRUE , echo = FALSE>>=
+
+  ext <- 20
+
+@
+
+Figure \ref{fig:reso_all} shows different comparisons among ChIP-exo,
+PET ChIP-Seq and SET ChIP-Seq. A RegulonDB annotation was considered
+identified if the distance between it and dPeak binding site estimate
+was at most of \Sexpr{ext} bp. That way, the sensitivity is defined as
+the proportion of RegulonDB annotations identified in a peak. In
+figure \ref{fig:reso_all}A can be seen that the sensitivity of all
+protocols increases as the average distance between binding sites
+does. Despite that when the binding events in a peak are closer to
+each other, both ChIP-exo and PET ChIP-Seq are comparable, as the
+distance increases ChIP-exo identifies a higher proportion of the
+RegulonDB annotations; additionally SET ChIP-Seq is significantly less
+sensible than both ChIP-exo and PET ChIP-Seq. In figure
+\ref{fig:reso_all}B the distance between a RegulonDB annotation to its
+closest prediction is compared for ChIP-exo, PET and SET ChIP-Seq;
+while the first two are comparable, both outperform SET ChIP-Seq.
+
+In figures \ref{fig:reso_all}C and \ref{fig:reso_all}D, we observe the
+behavior of dPeak's estimated parameters for single end data (both
+ChIP-exo and ChIP-Seq). The $\delta$ parameter measures the distance
+from the 5' end of the fragment to its respective binding event and
+the $\sigma$ parameter indicates the fragments distribution
+variability around their respective binding sites (more information in
+\cite{dpeak}). For both parameters, it can be seen that the ChIP-exo
+estimates are smaller than the SET ChIP-Seq estimates in average,
+which agrees with the fact that the sample reads are allocated more
+tightly around the binding events in ChIP-exo data. Hence we can
+conclude that the peaks shape is quite different between ChIP-exo and
+SET ChIP-Seq.
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width = .46\textwidth]{../figs/for_paper/sensitivity_exo_olda_data.pdf}
+  \includegraphics[width = .46\textwidth]{../figs/for_paper/resolution_by_dataset_old_data.pdf}
+   \includegraphics[width = .46\textwidth,page = 1]{../figs/for_paper/sigma_delta_old_densities.pdf}
+   \includegraphics[width = .46\textwidth,page = 2]{../figs/for_paper/sigma_delta_old_densities.pdf}
+
+   \caption{Comparison of (A) sensitivity and (B) resolution between
+     ChIP-exo and ChIP-Seq data. Sensitivity is defined as the
+     proportion of RegulonDB annotations identified using each
+     data. Resolution is defined as the distance between RegulonDB
+     annotation and its closest prediction. $\quad$ C) $\delta$ parameter in
+     dPeak measures average distance of the reads to their respective
+     binding site. In ChIP-exo data, reads were located much closer to
+     the binding site than in SET ChIP-Seq. D) $\sigma$ parameter
+     measure the dispersion of reads around each binding site. In
+     ChIP-exo data, reads showed less variation around the their
+     respective binding sites compared to SET ChIP-Seq.}
+  \label{fig:reso_all}
+\end{figure}
+
+\subsection{Recommendations for the design of ChIP-exo experiments}
+\label{sec:reco}
+
+<<design_param , include = FALSE , echo = FALSE , eval = TRUE>>=
+
+  K <- 10
+  ext <- 15
+
+@
+
+We sampled a fixed amount of fragments for each of the ChIP-exo, PET
+ChIP-Seq and SET ChIP-Seq datasets of the $\sigma^{70}$ sample in
+aerobic conditions. For each sampled dataset we applied our
+lower-to-higher resolution pipeline by calling peaks with MOSAiCS
+\cite{mosaics} and then deconvolving the binding events by using dPeak
+\cite{dpeak}. For the ChIP-exo datasets we called peaks by using
+GC-content and mappability with MOSAiCS, and for the ChIP-Seq datasets
+we used their respective Input samples.
+
+\begin{figure}[h!]
+  \centering
+%  \includegraphics[width = .8\textwidth]{../figs/for_paper/Sig70_aerobic_saturation.pdf}
+  \includegraphics[width = .8\textwidth]{../figs/for_paper/saturation_analysis_old.pdf}
+  \caption{Comparison of the number of candidate regions (A),
+    predicted events (B), identified targets (C) and resolution (D)
+    among ChIP-exo, PET ChIP-Seq and SET ChIP-Seq. RegulonDB
+    annotations are considered as a gold standard. A gold standard was
+    considered identified if a BS was estimated inside a \Sexpr{ext}
+    bp window around it}
+  \label{fig:design}
+\end{figure}
+
+Figure \ref{fig:design} shows the behavior of each data type when
+their depth is fixed. It is remarkable that even when the number of
+candidate peaks or the number of predicted events is lower for
+ChIP-exo, it outperforms both PET and SET ChIP-Seq in the number of
+identified targets and resolution. This may suggest that with ChIP-exo
+less false positive peaks are being called and that when the targets
+are being identified, dPeak estimates binding locations closer to the
+true location. Additionally, we can see that as the read depth
+increases all four indicators do so as well, which may indicate that
+with ChIP-exo a smaller amount of reads is necessary to identify a
+higher number of targets, but it may also be possible that this is an
+artifact occurring due to ChIP-exo's lower library complexity.
+
+\section{Conclusions}
+\label{sec:conclusions}
+
+We provide a ChIP-exo QC pipeline capable of assess the balance between
+enriched samples and low complexity regions. It is shown that the
+``peak-pair'' assumption doesn't hold in practice and we provide two
+out-of-the-box visualization capable to assess the strand imbalance in
+a ChIP-exo experiment.
+
+We updated the dPeak algorithm from \cite{dpeak} and showed that
+ChIP-exo is comparable in resolution and sensitivity with PET ChIP-Seq
+and outperforms SET ChIP-Seq. ChIP-exo compared with PET and SET
+ChIP-Seq at fixed depth sample is able to identify more targets at a
+lower resolution. We compared dPeak with another algorithms to
+estimate binding locations in ChIPexo data, dPeak is comparable to
+MACE and outperforms GEM in resolution. dPeak provided a striking
+balance in sensitivity, specificity and spatial resolution for
+ChIP-exo analysis.
+
+\newpage
 

### 2015-11-22 rmarkdown/prelims_paper.tex
Added conclusion and sections about comparisons realized with dpeak

@@
 
-\subsection{Current Quality Control Measures for ChIP-Seq}
-\label{sec:QC_seq}
-
-Despite that not specific QC pipeline exists for ChIP-exo.
-
 \begin{figure}[h!]
@@
 
+\subsection{Current Quality Control Measures for ChIP-Seq}
+\label{sec:QC_chipseq}
+
+To our knowledge it doesn't exists a QC pipeline for ChIP-exo. Hence,
+we analyzed how the current ChIP-Seq QC measures behave on ChIP-exo
+data.
+
+\begin{table}[h]
+  \centering
+  \caption{Usual quality control indicators applied to the $\sigma^{70}$ samples.
+PBC stands PCR-bottleneck coefficient, SSD for standardized standard deviation, NSC for normalized strand cross-correlation coefficient (Another measure related to the strand cross-correlation curve is the RSC was ommited since input sample usually don't exists for ChIP-exo).
+$\sig$ samples were given by Robert Landick, FoxA1 and ER samples are from Serandour et al., 2013 \cite{exoillumina} and CTCF sample is from Rhee and Pugh, 2011 \cite{exo1}.}
+
+\begin{tabular}{l|l|r|r|r|r|r|r}
+\hline\hline
+IP & Organism  &Condition & Rep. & Nr. reads & PBC & SSD &  NSC  \\
+\hline\hline
+$\sig$ & E.Coli & Rif-0min & 1 & 960,256 & 0.2823 & 0.0361 &  10.29 \\
+\hline
+$\sig$ &  E.Coli & Rif-0min & 2 & 2,247,295 & 0.2656 & 0.1091  &  25.08  \\
+\hline
+$\sig$ &  E.Coli & Rif-20min & 1 & 1,940,387 & 0.2698 & 0.0820  & 17.69  \\
+\hline
+$\sig$ &  E.Coli & Rif-20min & 2 & 4,229,574 & 0.2153 & 0.1647  &  14.11 \\
+\hline
+FoxA1 &  Mouse &  NA & 1 & 22,210,461 & 0.6562 & 9.12 $\times 10^{-5}$ & 21.452 \\
+\hline
+FoxA1 &  Mouse &  NA & 2 & 22,307,557 & 0.7996 & 7.94 $\times 10^{-5}$ & 60.661 \\
+\hline
+FoxA1 &  Mouse &  NA & 3 & 22,421,729 & 0.1068 & 1.31 $\times 10^{-4}$ & 72.312 \\
+\hline
+ER & Human & NA & 1 & 9,289,835 & 0.8082 & 3.64 $\times 10^{-5}$ & 19.843 \\
+\hline
+ER & Human & NA & 2 & 11,041,833 & 0.8024 & 4.6 $\times 10^{-5}$ & 21.422 \\
+\hline
+ER & Human & NA & 3 & 12,464,836 & 0.8203 & 4.89 $\times 10^{-5}$ & 19.699 \\
+\hline
+CTCF & Human & NA & 1 &   48,478,450 & 0.4579 & 1.29 $\times 10^{-4}$ & 15.977 \\
+\hline
+\end{tabular}  
+  \label{tab:qcbase}
+\end{table}
+
+Table \ref{tab:qcbase} contains measures that are commonly used with
+ChIP-Seq data. PBC stands for PCR bottleneck amplification and is
+defined as the ratio between the number of positions to which exactly
+one mappping read maps divided by the number of position to which at
+least one mapping reads maps, the ENCODE project proposed a series of
+ranges to interpret this quantity (which can be found in
+\url{https://www.encodeproject.org/data-standards/2012-quality-metrics/}),
+but it doesn't apply to ChIP-exo, if it did all the samples with
+$\mbox{PBC} < 0.5$ should be reject due to be showing severe
+bottlenecking. Planet et al., 2011 proposed the Standarized Standard
+Deviation which defined as:
+
+\[
+\mbox{SSD} = \mbox{SD}_N / \sqrt{N} 
+\]
+
+where $\mbox{SD}_N$ is the standard deviation of the coverage of the
+sample and $N$ is the number of reads. This measure is smaller for
+ChIP-exo since the reads are located more tightly around binding
+sites, but it may be biased since several fragments are being
+diggested by exonuclease enzyme, hence there are several position in
+the genome where no fragment is being mapped. Finally the normalized
+strand cross-correlation is defined as:
+
+\[
+\mbox{NSC} = \frac{ \max }{\min}
+\]
+
+\subsection{Comparison with ChIP-Seq data}
+\label{sec:comp_chipseq}
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width = .46\textwidth,page = 3 ]{../figs/for_paper/ChIPseqPET_ChIPexo_tagCount_comparison.pdf}
+  \includegraphics[width = .46\textwidth]{../figs/for_paper/forward_strand_ratio_comp_old.pdf}
+  \includegraphics[width = .46\textwidth,page = 1]{../figs/for_paper/eukaryotic_bias_CTCF.pdf}
+  \includegraphics[width = .46\textwidth,page = 2]{../figs/for_paper/eukaryotic_bias_CTCF.pdf}
+  \caption{ A) ChIP signal in ChIP-exo was linearly related to that in
+    PET ChIP-Seq data in the region with high ChIP tag counts. In
+    contrast, there were clear differences in their background
+    distribution, where several ChIP-exo background regions were
+    empty. B) In ChIP-exo data, strands of reads were significantly
+    less balanced in the regions with potential binding sites compared
+    to SET ChIP-Seq data. C) ChIP tag counts increase linearly as
+    mappability scores increases. D) ChIP tag counts increase linearly
+    as GC content score increases when GC content is less than 0.6 and
+    then ChIP tag counts decrease as GC content increases. }
+  \label{fig:comp}
+\end{figure}
+
+We first compared various factors that could affect binding site
+identification between ChIP-exo and ChIP-Seq data. In order to compare
+distribution of signal and background between ChIP-exo and ChIP-Seq
+data, we calculated ChIP tag counts across the genome by counting the
+number of reads mapping to each of 150 non-overlapping
+window after extending reads by 150 to their 3' end
+directions. ChIP tag counts in ChIP-exo data were linearly related to
+ChIP tag counts in ChIP-Seq data for the regions with high ChIP tag
+counts (Figure \ref{fig:comp}A). This implies that signals for
+potential binding sites are well reproducible between ChIP-exo and
+ChIP-Seq data. On the other hand, there was clear difference in the
+background distribution between them. In ChIP-Seq data reads were
+almost uniformly distributed over background (non-binding) regions and
+the ChIP tag counts in there regions were significantly larger than
+zero. In contrast, in ChIP-exo data, there was larger variation in
+ChIP tag counts among background regions and ChIP tag counts were much
+lower in these regions compared to ChIP-Seq data. There were also
+large proportion of regions without any read in ChIP-exo data. These
+results indicate that background distribution of ChIP-exo data is less
+homogeneous than that of ChIP-Seq data.
+
+We next evaluated the ``peak pair'' assumption from Rhee and Pugh,
+2011 \cite{exo1}, i.e. a peak of reads in the forward strand is
+usually paired with a peak of reads in the reverse strand that is
+located in the other site of the binding site. Wang et al., 2014
+\cite{mace}, Madrigal 2015 \cite{cexor} and Bardet et al., 2013
+\cite{peakzilla} proposed method rely in this assumption. In order to
+evaluate this assumption, we reviewed the proportion of reads in the
+forward strand in candidate regions (i.e. regions with at least one
+binding site) in $\sig$ ChIP-exo data. We found that strands of reads
+were much less balanced in ChIP-exo data than in ChIP-Seq data in
+these regions with potential binding sites (Fig. \ref{fig:comp}B) and
+this indicates that the peak pair assumption might not hold in real
+ChIP-exo data.
+
+We evaluated ChIP-exo data for CTCF factor from human genome
+\cite{exo1} to investigate issues specific to eukaryotic genomes for
+binding sites identification. Figures \ref{fig:comp}C and
+\ref{fig:comp}D display the bin-level average read counts against
+mappability and GC content. Each data point is obtained by averaging
+the read counts across bins with the same mappability of GC
+content. These results indicate that binding site identification in
+ChIP-exo sample might also benefit from the use of methods that take
+into account of apparent sequence biases such as mappability and GC
+content.
+
 \section{Results}
 \label{sec:results}
+
+\subsubsection{Comparison with ChIP-Seq data using dPeak}
+\label{sec:dpeak_analysis}
+
+Figure \ref{fig:reso_all} shows different comparisons among ChIP-exo,
+PET ChIP-Seq and SET ChIP-Seq. A RegulonDB annotation was considered
+identified if the distance between it and dPeak binding site estimate
+was at most of 20 bp. That way, the sensitivity is defined as
+the proportion of RegulonDB annotations identified in a peak. In
+figure \ref{fig:reso_all}A can be seen that the sensitivity of all
+protocols increases as the average distance between binding sites
+does. Despite that when the binding events in a peak are closer to
+each other, both ChIP-exo and PET ChIP-Seq are comparable, as the
+distance increases ChIP-exo identifies a higher proportion of the
+RegulonDB annotations; additionally SET ChIP-Seq is significantly less
+sensible than both ChIP-exo and PET ChIP-Seq. In figure
+\ref{fig:reso_all}B the distance between a RegulonDB annotation to its
+closest prediction is compared for ChIP-exo, PET and SET ChIP-Seq;
+while the first two are comparable, both outperform SET ChIP-Seq.
+
+In figures \ref{fig:reso_all}C and \ref{fig:reso_all}D, we observe the
+behavior of dPeak's estimated parameters for single end data (both
+ChIP-exo and ChIP-Seq). The $\delta$ parameter measures the distance
+from the 5' end of the fragment to its respective binding event and
+the $\sigma$ parameter indicates the fragments distribution
+variability around their respective binding sites (more information in
+\cite{dpeak}). For both parameters, it can be seen that the ChIP-exo
+estimates are smaller than the SET ChIP-Seq estimates in average,
+which agrees with the fact that the sample reads are allocated more
+tightly around the binding events in ChIP-exo data. Hence we can
+conclude that the peaks shape is quite different between ChIP-exo and
+SET ChIP-Seq.
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width = .46\textwidth]{../figs/for_paper/sensitivity_exo_olda_data.pdf}
+  \includegraphics[width = .46\textwidth]{../figs/for_paper/resolution_by_dataset_old_data.pdf}
+   \includegraphics[width = .46\textwidth,page = 1]{../figs/for_paper/sigma_delta_old_densities.pdf}
+   \includegraphics[width = .46\textwidth,page = 2]{../figs/for_paper/sigma_delta_old_densities.pdf}
+
+   \caption{Comparison of (A) sensitivity and (B) resolution between
+     ChIP-exo and ChIP-Seq data. Sensitivity is defined as the
+     proportion of RegulonDB annotations identified using each
+     data. Resolution is defined as the distance between RegulonDB
+     annotation and its closest prediction. $\quad$ C) $\delta$ parameter in
+     dPeak measures average distance of the reads to their respective
+     binding site. In ChIP-exo data, reads were located much closer to
+     the binding site than in SET ChIP-Seq. D) $\sigma$ parameter
+     measure the dispersion of reads around each binding site. In
+     ChIP-exo data, reads showed less variation around the their
+     respective binding sites compared to SET ChIP-Seq.}
+  \label{fig:reso_all}
+\end{figure}
+
+\subsection{Recommendations for the design of ChIP-exo experiments}
+\label{sec:reco}
+
+We sampled a fixed amount of fragments for each of the ChIP-exo, PET
+ChIP-Seq and SET ChIP-Seq datasets of the $\sigma^{70}$ sample in
+aerobic conditions. For each sampled dataset we applied our
+lower-to-higher resolution pipeline by calling peaks with MOSAiCS
+\cite{mosaics} and then deconvolving the binding events by using dPeak
+\cite{dpeak}. For the ChIP-exo datasets we called peaks by using
+GC-content and mappability with MOSAiCS, and for the ChIP-Seq datasets
+we used their respective Input samples.
+
+\begin{figure}[h!]
+  \centering
+%  \includegraphics[width = .8\textwidth]{../figs/for_paper/Sig70_aerobic_saturation.pdf}
+  \includegraphics[width = .8\textwidth]{../figs/for_paper/saturation_analysis_old.pdf}
+  \caption{Comparison of the number of candidate regions (A),
+    predicted events (B), identified targets (C) and resolution (D)
+    among ChIP-exo, PET ChIP-Seq and SET ChIP-Seq. RegulonDB
+    annotations are considered as a gold standard. A gold standard was
+    considered identified if a BS was estimated inside a 15
+    bp window around it}
+  \label{fig:design}
+\end{figure}
+
+Figure \ref{fig:design} shows the behavior of each data type when
+their depth is fixed. It is remarkable that even when the number of
+candidate peaks or the number of predicted events is lower for
+ChIP-exo, it outperforms both PET and SET ChIP-Seq in the number of
+identified targets and resolution. This may suggest that with ChIP-exo
+less false positive peaks are being called and that when the targets
+are being identified, dPeak estimates binding locations closer to the
+true location. Additionally, we can see that as the read depth
+increases all four indicators do so as well, which may indicate that
+with ChIP-exo a smaller amount of reads is necessary to identify a
+higher number of targets, but it may also be possible that this is an
+artifact occurring due to ChIP-exo's lower library complexity.
+
+\section{Conclusions}
+\label{sec:conclusions}
+
+We provide a ChIP-exo QC pipeline capable of assess the balance between
+enriched samples and low complexity regions. It is shown that the
+``peak-pair'' assumption doesn't hold in practice and we provide two
+out-of-the-box visualization capable to assess the strand imbalance in
+a ChIP-exo experiment.
+
+We updated the dPeak algorithm from \cite{dpeak} and showed that
+ChIP-exo is comparable in resolution and sensitivity with PET ChIP-Seq
+and outperforms SET ChIP-Seq. ChIP-exo compared with PET and SET
+ChIP-Seq at fixed depth sample is able to identify more targets at a
+lower resolution. We compared dPeak with another algorithms to
+estimate binding locations in ChIPexo data, dPeak is comparable to
+MACE and outperforms GEM in resolution. dPeak provided a striking
+balance in sensitivity, specificity and spatial resolution for
+ChIP-exo analysis.
+
+\newpage
 

### 2015-11-22 rscripts/carroll_human_library_complexity.R

@@ -15,7 +15,7 @@ files <- list.files(dr)
 files <- files[grep("bam",files)]
 files <-files[grep("bai",files,invert = TRUE)]
 
-files <- files[sapply(c("41","52","59"),grep,files)]
+files <- files[sapply(c("933","950","938"),grep,files)]
 
 files <- file.path(dr,files)
 

### 2015-11-22 rscripts/sig70_resolution_sensivity_all.R

@@ -1,0 +1,302 @@
+
+rm(list = ls())
+
+library(data.table)
+library(GenomicAlignments)
+library(ggplot2)
+library(MASS)
+library(RColorBrewer)
+
+######################################################################################
+
+## Initial parameters
+
+mc <- detectCores()
+figs_dir <- "figs/condition"
+
+base_dir <- "/p/keles/ChIPexo/volume6/condition"
+
+check_create <- function(dr)if(!dir.exists(dr))dir.create(dr)
+
+
+figs_dir <- file.path(figs_dir,"all_together")
+check_create(figs_dir)
+
+######################################################################################
+
+## get promoter locations table
+
+prom_dir <- "/p/keles/ChIPexo/volume6/gold_standard/Landick"
+pfiles <- list.files(prom_dir)
+pfiles <- pfiles[grep(".bed",pfiles)]
+promoters <- lapply(file.path(prom_dir,pfiles),read.table)
+promoters <- lapply(promoters,data.table)
+
+promoters[[1]][,str := "R"]
+promoters[[2]][,str := "F"]
+
+sites <- do.call(rbind,promoters)
+sites[, id := 1:nrow(sites)]
+setnames(sites,names(sites),c("seqnames","start","end","strand","id"))
+
+sites[,seqnames := "U00096"]
+sites[,strand := ifelse(strand == "F","+","-")]
+
+######################################################################################
+
+conditions <-list.files(base_dir)
+dirs <- file.path(base_dir,conditions)
+
+fdr <- c(.1,1,10,5)
+what <- c("exo","pet","set")
+
+load_all <- function(dir,sites)
+{
+  all_files <- list.files(dir,recursive = TRUE)
+
+  peak_files <- all_files[grep("peaks",all_files)]
+  peaks <- lapply(file.path(dir,peak_files),read.table)
+  peaks <- lapply(peaks,data.table)
+  peaks <- lapply(peaks,function(x){
+    x[,peakId := paste0(V1,":",V2,"-",V3)]
+    return(x)})
+
+  bs_files <- all_files[grep("binding",all_files)]
[…]

### 2015-11-22 rmarkdown/prelims_paper.Rnw
Added section on strand cc

@@
 \[
-\mbox{NSC} = \frac{ \max }{\min}
+\mbox{NSC} = \frac{ \max_\delta y(\delta)}{\min_\delta y(\delta)}
 \]
+
+where $y(\delta)$ is strand cross-correlation defined as in (\ref{scc}).
+
+\subsubsection{Strand cross-correlation}
+\label{sec:scc}
+
+The strand cross-correlation (SCC) was proposed by Kharchenko et al.,
+2008 \cite{strandcc} and it may be one of the most used of the
+ChIP-Seq QC metrics. The SCC curve is defined as:
+
+\begin{align}
+  y(\delta) = \sum_c w_c r\left[ n_c^+ \left(x + \frac{\delta}{2}
+    \right), n_c^- \left( x- \frac{\delta}{2} \right)\right]
+\label{scc}
+\end{align}
+
+where $y(\delta)$ is strand cross-correlation for a strand shift
+$\delta$, $r$ is the Pearson correlation, $w_c$ is the propotion of
+tags mapped to chromosome $c$ and $n_c^S$ is the tag count vector for
+strand $S$ and chromsome $c$.
 

### 2015-11-23 rmarkdown/prelims_paper.Rnw
Added dpeak section to prelims paper and updated figures to include rep numbers instead of bam files name

@@
 
-\author{Rene Welch\\Preliminary Examination\\Department of Statistics, University of Wisconsin-Madison}
+\author{Rene Welch\\Preliminary Examination\\Department of Statistics,
+  University of Wisconsin-Madison}
 
@@
 \caption{ChIP-exo diagram. The only difference with ChIP-Seq is the
-  second step, where an exonuclease enzyme is added and it diggest the
+  second step, where an exonuclease enzyme is added and it digest the
   DNA fragments starting from the 5' end until it finds a crosslink
@@
 ChIP-exo is based in a modification to the ChIP-seq protocol, where an
-exonuclease enzyme is added and it diggests the fragments starting by
+exonuclease enzyme is added and it digests the fragments starting by
 the 5' ends and stops until it finds a protein that is crosslinked to
@@
 
-\begin{figure}[h!]
+\begin{figure}[H]
   \centering
@@
 sites, but it may be biased since several fragments are being
-diggested by exonuclease enzyme, hence there are several position in
+digested by exonuclease enzyme, hence there are several position in
 the genome where no fragment is being mapped. Finally the normalized
@@
 
-where $y(\delta)$ is strand cross-correlation for a strand shift
-$\delta$, $r$ is the Pearson correlation, $w_c$ is the propotion of
-tags mapped to chromosome $c$ and $n_c^S$ is the tag count vector for
-strand $S$ and chromsome $c$.
+where $y(\delta)$ is the SCC for a strand shift $\delta$, $r$ is the
+Pearson correlation, $w_c$ is the propotion of tags mapped to
+chromosome $c$ and $n_c^S$ is the tag count vector for strand $S$ and
+chromsome $c$.
+
+For ChIP-Seq data, the SCC will have two peaks: One when $\delta$ is
+equal to reads length and one when $\delta$ is equal to the unobserved
+fragment length, ideally the SCC function is maximized at this
+point. Landt et al., 2012 \cite{encode_qc} explain that a succesful if
+this happens, it is marginally succesful if its maximized at the read
+length peak but still shows the fragment length summit and failed if
+is maximized at the read length peak but the second summit is not
+present.
+
+\begin{figure}[H]
+  \centering  
+\includegraphics[width = .7\textwidth]{../figs/for_paper/scc_ctcf.pdf}
+  \caption{SCC for CTCF factor in HeLa cell line for ChIP-exo (obtained from \cite{exo1}) and SET ChIP-Seq (obtained from \cite{encode1}) }
+  \label{fig:scc}
+\end{figure}
+
+Figure \ref{fig:scc} shows the SCC function for both a ChIP-exo (in
+red) and a SET ChIP-Seq (in green) for CTCF in HeLa cell line. The
+ChIP-Seq dataset behaves as the succesful case from \cite{encode_qc},
+but the ChIP-exo doesn't. Additionally, since the exonuclease enzyme
+digests the DNA fragments, then the shifts where the SCC may reach a
+locall or a global maximum are confounded. Hence, new QC measures are
+necessary for ChIP-exo data.
+
+\subsection{The dPeak model}
+\label{sec:dpeak}
+
+Chung et al., 2013 proposed the dpeak model to identifity
+transcription factor binding sites in high resolution. Two models are
+proposed: One to deconvolve the signal of single end tags (SET) data,
+i.e. when only one end of the DNA fragment is sequenced; and another
+to deconvolve the signal of paired end tags (PET) data, i.e. when both
+ends of the DNA fragment are sequenced. For ChIP-exo, both ends are
+sequenced but only the 5' end is considered, since is the side where
+the exonuclease enzyme digests the DNA.
+
+Given a peak region with $m$ positions, with $n$ DNA fragments and that
+the region is generated by a known $g^*$ amount of binding
+events. Each fragment is described by the start position where the
+starting end is aligned and the strand from which the fragment is
+sequenced. Since only one end is sequenced, then the fragment length
+is not observed. dPeak assumes that the strand $D_i$ follows a
+Bernoulli distribution with known parameter $p_D$, and the start
+position $R_i$ for the $i$-th fragment is generated according to the
+following procedure:
+
+\begin{itemize}
+\item If the fragment was sequenced from the forward strand ($D_i =
+  1$):
+  \begin{itemize}
+  \item The reads belongs to the background: $R_i | Z_i = 0, D_i = 1
+    \sim \mbox{Unif}(1 - \beta + 1, m)$
+  \item The read belong to the $g$-th binding event: $R_i | Z_i = g,
+    D_i = 1 \sim \mbox{N}(\mu_g - \delta , \sigma^2)$
+  \end{itemize}
+\item If the fragment was sequenced from the backward strand ($D_i =
+  0$):
+  \begin{itemize}
+  \item The reads belongs to the background: $R_i | Z_i = 0, D_i = 1
+    \sim \mbox{Unif}(1 , m + \beta  - 1)$
+  \item The read belong to the $g$-th binding event: $R_i | Z_i = g,
+    D_i = 1 \sim \mbox{N}(\mu_g + \delta , \sigma^2)$
+  \end{itemize}
+
+\end{itemize}
+
+and the binding event to which the $i$-th fragment belongs $Z_i = g$
+is generated from a
+
+$\mbox{Multinomial}(\pi_0,\pi_1,\cdots,\pi_{g^*})$.
 
@@
 
+\subsection{ChIP-exo Quality Control pipeline}
+\label{sec:QC}
+
+Figure \ref{fig:qcdiagram} shows a flowchart for the ChIPexoQC
+pipeline. Which basically partitions the genome by keeping the
+non-digested ChIP-exo regions. Then, for each region it calculates a
+series of statistics. Finally, it creates several visualizations
+designed to assess the quality levels of a ChIP-exo sample.
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width = .9\textwidth]{../figs/for_paper/coverage_diagram.pdf}
+  \caption{Diagram of the ChIP-exo Quality control pipeline [this is a
+    place-holder]. The genome is partitioned into several regions by
+    removing the empty segments and for each regions we calculate a
+    collection of QC indicators.}
+  \label{fig:qcdiagram}
+\end{figure}
+
+\subsubsection{Enrichment and library complexity in ChIP-exo data}
+\label{sec:enri}
+
+In ChIP-exo experiments, the background is often digested by the
+exonuclease enzyme, therefore to determine the sample's quality is
+necessary to address the balance between the enrichment and library
+complexity. To diagnose this, we considered the Average Read Coverage
+(ARC) and the Unique Read Coverage Rate (URCR) which are defined as:
+
+\begin{align}
+  \mbox{ARC} &= \frac{\text{Nr. of reads in the region}}{\text{Width of the region}} \nonumber \\
+  \mbox{URCR} &= \frac{\text{Nr. of reads mapped to only one position
+      in the region}}{\text{Nr. of reads in the region}} \nonumber
+\end{align}
+
+Using the mouse-FoxA1 experiment from \cite{exoillumina} and the
+relationship between this two quantities, library complexity and
+sample enrichment was explored. Figure \ref{fig:enrich}A shows hexbin
+plots illustrating the interaction between these two
+quantities. Notice how there are two strong arms in each panel: The
+first one corresponds to regions with low $\mbox{ARC}$ values and
+varying $\mbox{URCR}$ values across the $(0,1]$ interval, while the
+second one shows a decreasing trend in $\mbox{URCR}$ as $\mbox{ARC}$
+increases. When an experiment shows a higher degree of enrichment,
+then the separation of this two arms is more noticeable, since the
+second arm corresponds to possibly enriched regions
+(\ref{fig:enrich2}A and \ref{fig:enrich2}B).
+
+<<strata_para , include =FALSE,echo = FALSE, veal = TRUE>>=
+
+  strata <- 400
+  high <- 100
+  med <- 50
+  low <- 20
+
+@
+
+Figure \ref{fig:enrich}B shows box plots of the $\mbox{local-NSC}$ on
+regions of the three replicates, stratified by the number of reads
+mapped exactly to one position for all $\sigma^{70}$ ChIP-exo
+experiments; The high stratum is defined by regions consisting of more
+than \Sexpr{high} unique positions, the medium stratum for regions
+where the number of unique position is on the
+(\Sexpr{med},\Sexpr{high}) range and the low stratum in the
+(\Sexpr{low},\Sexpr{med}) range. For each stratum and biological
+replicate, we sampled \Sexpr{strata} regions whenever possible (in the
+opposite case, all the regions in the category were considered) and
+calculated the $\mbox{local-NSC}$ for those regions. For all three
+replicates, it is shown an increasing trend as the number of unique
+position increases, which means that the $\mbox{local-NSC}$ is an
+effective indicator of library complexity. Figure \ref{fig:enrich}C
+shows the $\mbox{local-NSC}$ coefficient calculated for the same set
+of regions, stratified by the number of unique positions in the
+regions of ERR336942 using the same criteria as before. In the high
+panel it is shown that the ERR336942 sample performs slightly better
+than ERR336956 and both outperform ERR336935, this agrees with
+\ref{fig:enrich}A, where only in the left-most panel doesn't seem to
+be a separation between the low $\mbox{ARC}$ arm and the other.
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width = .8\textwidth]{../figs/Carroll_mice_for_paper/FoxA1_enrichment.pdf}
+  \includegraphics[width = .8\textwidth,page =3]{../figs/Carroll_mice_for_paper/Local_SCC_indicator_by_strata.pdf}
+  \includegraphics[width = .8\textwidth,page =4]{../figs/Carroll_mice_for_paper/Local_SCC_indicator_by_strata_sameRegion.pdf}
+  \caption{Using the mouse-FoxA1 experiment from \cite{exoillumina}:
+    A) Hexbin plots of $\mbox{ARC}$ against $\mbox{URCR}$, in general
+    we can see a slight separation into two strong arms, one
+    corresponds to low $\mbox{ARC}$ and varying $\mbox{URCR}$, and for
+    the other $\mbox{URCR}$ decreases as $\mbox{ARC}$ increases B)
+    Boxplots of the $\mbox{local-NSC}$ stratified by nr. of reads
+    mapped to only one position in the regions for the ChIP-exo
+    experiments divided by condition and biological replicate and C)
+    Boxplots of the a set of ERR336942 regions, stratified as in (B)
+    with the $\mbox{local-NSC}$ calculated with the different
+    biological replicates.}
+  \label{fig:enrich}
+\end{figure}
+
+\subsubsection{Strand imbalance in ChIP-exo data}
+\label{sec:strand_imbalance}
+
+The strand imbalance assessment is based in the observation that the
+enriched regions usually have a higher concentration of fragments,
+therefore we examined the FSR (defined as the ratio of the number of
+forward stranded reads divided by the total number of reads in a given
+region) as the region with lower depth are being filtered out. This
+indicator is of particular importance, since several methods rely on
+the ``peak-pair'' assumption. In table \ref{tab:qcbase}, we calculated
+the FSR and noticed that for all the samples, it's value is close to
+0.5, which means that there are roughly the same amount of reads in
+both strands. However, figure \ref{fig:comp}B shows that this value is
+not representative of the sample locally, therefore the assumption
+doesn't hold in practice.
+
+<<fsr,include = FALSE,eval = TRUE,echo = FALSE>>=
+
+  de <- 100
+
+  pval <- list()
+  pval[[1]] <- 0.1039
+  pval[[2]] <- 0.0064
+  pval[[3]] <- 0.0871
+
+@
+
+In order to assess the strand imbalance we created the following
+visualization presented in figure \ref{fig:strand}. A) shows the FSR's
+behavior as lower depth regions are being filtered out, while B) shows
+which percentage of the regions are composed by fragments of both
+strands or only one (forward or backward). As the number of reads per
+region's lower bound increases, the quantiles tend to approach the
+median. To evaluate this statement, for each replicate of the FoxA1
+dataset, we divided the regions formed by at least \Sexpr{de}
+fragments into into the ones that overlap with any of the sample's
+peaks. The p.value for ERR36935 is \Sexpr{pval[[1]]}, for ERR336942 is
+\Sexpr{pval[[2]]} and for ERR336956 is \Sexpr{pval[[3]]}. This shows
+that only the middle sample from figure \ref{fig:strand} is the least
+imbalance sample from the the three, while the other two still show a
+higher imbalance at the \Sexpr{de} level.
+
+\begin{figure}[H]
+  \centering  
+  \includegraphics[width = .8\textwidth,page = 3]{../figs/Carroll_mice_for_paper/Strand_imbalance.pdf} 
+  \caption{Strand imbalance QC plots for the same data as in figure
+    \ref{fig:enrich}. A) FSR distribution quantiles as the lower depth
+    regions are being filtered out, all quantiles approach to the
+    median as the lower bound increases. B) Stacked histogram with the
+    proportion of regions that are formed by two strands or only one,
+    in a good sample the single-stranded regions are going to be
+    filtered out quickly as in the middle row.}
+  \label{fig:strand}
+\end{figure}
+
 \subsubsection{Comparison with ChIP-Seq data using dPeak}
@@
 
+\subsection{GC content and mappability quality indicators}
+\label{sec:gc_map_qc}
+
+Figure \ref{fig:comp} shows a relationship between the ChIP tag counts
+of a ChIP-exo experiment and both mappability and GC content
+scores. Furthermore, in sections \ref{sec:reco} and
+\ref{sec:dpeak_analysis} we called peaks by using MOSAiCS which
+considers them as necessary information when there is not input
+sample. Hence, quality measures we may consider to add indicators to
+the QC pipeline that considers this biases.
+
+\subsection{Quality control for ChIP-Nexus}
+\label{sec:nexus}
+
+He et al., 2014 \cite{chipnexus} proposed a modification to the
+ChIP-exo protocol that may fix its low library complexity issues by
+extending the fragments by a random barcode prior to the exonuclease
+digestion. This last statement is briefly mentioned by Mahony and
+Pugh, 2015 \cite{exo_review}. We want to apply our QC pipeline to
+ChIP-Nexus. Depending if ChIP-Nexus data shows background we may need
+to modify the QC pipeline in the step where it paritions the reads
+into regions, but the $\mbox{local-NSC}$ may be used to prove this
+statement.
+
 \newpage

### 2015-11-23 rmarkdown/prelims_paper.tex
Added dpeak section to prelims paper and updated figures to include rep numbers instead of bam files name

@@
 
+\subsection{ChIP-exo Quality Control pipeline}
+\label{sec:QC}
+
+Figure \ref{fig:qcdiagram} shows a flowchart for the ChIPexoQC
+pipeline. Which basically partitions the genome by keeping the
+non-digested ChIP-exo regions. Then, for each region it calculates a
+series of statistics. Finally, it creates several visualizations
+designed to assess the quality levels of a ChIP-exo sample.
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width = .9\textwidth]{../figs/for_paper/coverage_diagram.pdf}
+  \caption{Diagram of the ChIP-exo Quality control pipeline [this is a
+    place-holder]. The genome is partitioned into several regions by
+    removing the empty segments and for each regions we calculate a
+    collection of QC indicators.}
+  \label{fig:qcdiagram}
+\end{figure}
+
+\subsubsection{Enrichment and library complexity in ChIP-exo data}
+\label{sec:enri}
+
+In ChIP-exo experiments, the background is often digested by the
+exonuclease enzyme, therefore to determine the sample's quality is
+necessary to address the balance between the enrichment and library
+complexity. To diagnose this, we considered the Average Read Coverage
+(ARC) and the Unique Read Coverage Rate (URCR) which are defined as:
+
+\begin{align}
+  \mbox{ARC} &= \frac{\text{Nr. of reads in the region}}{\text{Width of the region}} \nonumber \\
+  \mbox{URCR} &= \frac{\text{Nr. of reads mapped to only one position
+      in the region}}{\text{Nr. of reads in the region}} \nonumber
+\end{align}
+
+Using the mouse-FoxA1 experiment from \cite{exoillumina} and the
+relationship between this two quantities, library complexity and
+sample enrichment was explored. Figure \ref{fig:enrich}A shows hexbin
+plots illustrating the interaction between these two
+quantities. Notice how there are two strong arms in each panel: The
+first one corresponds to regions with low $\mbox{ARC}$ values and
+varying $\mbox{URCR}$ values across the $(0,1]$ interval, while the
+second one shows a decreasing trend in $\mbox{URCR}$ as $\mbox{ARC}$
+increases. When an experiment shows a higher degree of enrichment,
+then the separation of this two arms is more noticeable, since the
+second arm corresponds to possibly enriched regions
+(\ref{fig:enrich2}A and \ref{fig:enrich2}B).
+
+Figure \ref{fig:enrich}B shows box plots of the $\mbox{local-NSC}$ on
+regions of the three replicates, stratified by the number of reads
+mapped exactly to one position for all $\sigma^{70}$ ChIP-exo
+experiments; The high stratum is defined by regions consisting of more
+than 100 unique positions, the medium stratum for regions
+where the number of unique position is on the
+(50,100) range and the low stratum in the
+(20,50) range. For each stratum and biological
+replicate, we sampled 400 regions whenever possible (in the
+opposite case, all the regions in the category were considered) and
+calculated the $\mbox{local-NSC}$ for those regions. For all three
+replicates, it is shown an increasing trend as the number of unique
+position increases, which means that the $\mbox{local-NSC}$ is an
+effective indicator of library complexity. Figure \ref{fig:enrich}C
+shows the $\mbox{local-NSC}$ coefficient calculated for the same set
+of regions, stratified by the number of unique positions in the
+regions of ERR336942 using the same criteria as before. In the high
+panel it is shown that the ERR336942 sample performs slightly better
+than ERR336956 and both outperform ERR336935, this agrees with
+\ref{fig:enrich}A, where only in the left-most panel doesn't seem to
+be a separation between the low $\mbox{ARC}$ arm and the other.
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width = .8\textwidth]{../figs/Carroll_mice_for_paper/FoxA1_enrichment.pdf}
+  \includegraphics[width = .8\textwidth,page =3]{../figs/Carroll_mice_for_paper/Local_SCC_indicator_by_strata.pdf}
+  \includegraphics[width = .8\textwidth,page =4]{../figs/Carroll_mice_for_paper/Local_SCC_indicator_by_strata_sameRegion.pdf}
+  \caption{Using the mouse-FoxA1 experiment from \cite{exoillumina}:
+    A) Hexbin plots of $\mbox{ARC}$ against $\mbox{URCR}$, in general
+    we can see a slight separation into two strong arms, one
+    corresponds to low $\mbox{ARC}$ and varying $\mbox{URCR}$, and for
+    the other $\mbox{URCR}$ decreases as $\mbox{ARC}$ increases B)
+    Boxplots of the $\mbox{local-NSC}$ stratified by nr. of reads
+    mapped to only one position in the regions for the ChIP-exo
+    experiments divided by condition and biological replicate and C)
+    Boxplots of the a set of ERR336942 regions, stratified as in (B)
+    with the $\mbox{local-NSC}$ calculated with the different
+    biological replicates.}
+  \label{fig:enrich}
+\end{figure}
+
+\subsubsection{Strand imbalance in ChIP-exo data}
+\label{sec:strand_imbalance}
+
+The strand imbalance assessment is based in the observation that the
+enriched regions usually have a higher concentration of fragments,
+therefore we examined the FSR (defined as the ratio of the number of
+forward stranded reads divided by the total number of reads in a given
+region) as the region with lower depth are being filtered out. This
+indicator is of particular importance, since several methods rely on
+the ``peak-pair'' assumption. In table \ref{tab:qcbase}, we calculated
+the FSR and noticed that for all the samples, it's value is close to
+0.5, which means that there are roughly the same amount of reads in
+both strands. However, figure \ref{fig:comp}B shows that this value is
+not representative of the sample locally, therefore the assumption
+doesn't hold in practice.
+
+In order to assess the strand imbalance we created the following
+visualization presented in figure \ref{fig:strand}. A) shows the FSR's
+behavior as lower depth regions are being filtered out, while B) shows
+which percentage of the regions are composed by fragments of both
+strands or only one (forward or backward). As the number of reads per
+region's lower bound increases, the quantiles tend to approach the
+median. To evaluate this statement, for each replicate of the FoxA1
+dataset, we divided the regions formed by at least 100
+fragments into into the ones that overlap with any of the sample's
+peaks. The p.value for ERR36935 is 0.1039, for ERR336942 is
+0.0064 and for ERR336956 is 0.0871. This shows
+that only the middle sample from figure \ref{fig:strand} is the least
+imbalance sample from the the three, while the other two still show a
+higher imbalance at the 100 level.
+
+\begin{figure}[H]
+  \centering  
+  \includegraphics[width = .8\textwidth,page = 3]{../figs/Carroll_mice_for_paper/Strand_imbalance.pdf} 
+  \caption{Strand imbalance QC plots for the same data as in figure
+    \ref{fig:enrich}. A) FSR distribution quantiles as the lower depth
+    regions are being filtered out, all quantiles approach to the
+    median as the lower bound increases. B) Stacked histogram with the
+    proportion of regions that are formed by two strands or only one,
+    in a good sample the single-stranded regions are going to be
+    filtered out quickly as in the middle row.}
+  \label{fig:strand}
+\end{figure}
+
 \subsubsection{Comparison with ChIP-Seq data using dPeak}

### 2015-11-23 rmarkdown/prelims_paper.R

@@ -4,6 +4,24 @@
   frag_len <- 150
   
 
+## ----strata_para , include =FALSE,echo = FALSE, veal = TRUE--------------
+
+  strata <- 400
+  high <- 100
+  med <- 50
+  low <- 20
+  
+
+## ----fsr,include = FALSE,eval = TRUE,echo = FALSE------------------------
+
+  de <- 100
+
+  pval <- list()
+  pval[[1]] <- 0.1039
+  pval[[2]] <- 0.0064
+  pval[[3]] <- 0.0871
+
+
 ## ----reso_param, include = FALSE , eval  = TRUE , echo = FALSE-----------
 
   ext <- 20

### 2015-11-23 rmarkdown/prelims_paper.Rnw
Added section about software

@@
 (Chromatin Immunoprecipitation coupled with next generation sequencing
-) assay.
+ assay.
 
@@
 
+\subsubsection{dPeak's algorithm initialization strategy}
+\label{sec:algo}
+
+dPeak's model likelihood is optimized by the stochastic EM
+algorithm. Hence, it is of utmost importance how the parameters are
+initialized in order to obtain a global maximum.
+
+Chung et al., 2013 \cite{dpeak} derived an initialization strategy for
+the stochastic EM algorithm where each region is initialized by
+separate independently and the binding sites locations $\mu_g,
+g=1,\cdots,g^*$ are initialized by a uniform partition.
+
+For this work, we considered a new strategy where the all the regions
+are ranked by a statistical measure (as the $-\log(p)$ where $p$ is
+the empirical p.value of the test $H_0: \mathcal{R} \text{ is
+  enriched}$, where $\mathcal{R}$ is a genomic region) or a empirical
+quantity (as the number of ChIP tag counts in the region) and use as
+initial estimates of the $\delta$ and $\sigma$ parameters the average
+of the estimates from the highest ranked $N_{\text{top}}$ peaks.
+
 \subsection{Comparison with ChIP-Seq data}
 \label{sec:comp_chipseq}
-
-\begin{figure}[H]
-  \centering
-  \includegraphics[width = .46\textwidth,page = 3 ]{../figs/for_paper/ChIPseqPET_ChIPexo_tagCount_comparison.pdf}
-  \includegraphics[width = .46\textwidth]{../figs/for_paper/forward_strand_ratio_comp_old.pdf}
-  \includegraphics[width = .46\textwidth,page = 1]{../figs/for_paper/eukaryotic_bias_CTCF.pdf}
-  \includegraphics[width = .46\textwidth,page = 2]{../figs/for_paper/eukaryotic_bias_CTCF.pdf}
-  \caption{ A) ChIP signal in ChIP-exo was linearly related to that in
-    PET ChIP-Seq data in the region with high ChIP tag counts. In
-    contrast, there were clear differences in their background
-    distribution, where several ChIP-exo background regions were
-    empty. B) In ChIP-exo data, strands of reads were significantly
-    less balanced in the regions with potential binding sites compared
-    to SET ChIP-Seq data. C) ChIP tag counts increase linearly as
-    mappability scores increases. D) ChIP tag counts increase linearly
-    as GC content score increases when GC content is less than 0.6 and
-    then ChIP tag counts decrease as GC content increases. }
-  \label{fig:comp}
-\end{figure}
 
@@
 homogeneous than that of ChIP-Seq data.
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width = .46\textwidth,page = 3 ]{../figs/for_paper/ChIPseqPET_ChIPexo_tagCount_comparison.pdf}
+  \includegraphics[width = .46\textwidth]{../figs/for_paper/forward_strand_ratio_comp_old.pdf}
+\includegraphics[width = .46\textwidth,page = 1]{../figs/for_paper/eukaryotic_bias_CTCF.pdf}
+  \includegraphics[width = .46\textwidth,page = 2]{../figs/for_paper/eukaryotic_bias_CTCF.pdf}
+  \caption{ A) ChIP signal in ChIP-exo was linearly related to that in
+    PET ChIP-Seq data in the region with high ChIP tag counts. In
+    contrast, there were clear differences in their background
+    distribution, where several ChIP-exo background regions were
+    empty. B) In ChIP-exo data, strands of reads were significantly
+    less balanced in the regions with potential binding sites compared
+    to SET ChIP-Seq data. C) ChIP tag counts increase linearly as
+    mappability scores increases. D) ChIP tag counts increase linearly
+    as GC content score increases when GC content is less than 0.6 and
+    then ChIP tag counts decrease as GC content increases. }
+  \label{fig:comp}
+\end{figure}
 
@@
 \end{figure}
+
+\subsubsection{Local Normalized Strand cross-correlation}
+\label{sec:locNSC}
+
+We considered the creation of a novel QC measure for ChIP-exo called
+local normalized SCC or $\mbox{local-NSC}$.
 
@@
 of regions, stratified by the number of unique positions in the
-regions of ERR336942 using the same criteria as before. In the high
-panel it is shown that the ERR336942 sample performs slightly better
-than ERR336956 and both outperform ERR336935, this agrees with
-\ref{fig:enrich}A, where only in the left-most panel doesn't seem to
-be a separation between the low $\mbox{ARC}$ arm and the other.
+regions of the first replicate using the same criteria as before. In
+the high panel it is shown that the first replicate sample performs
+slightly better than the second and both outperform the third one,
+this agrees with \ref{fig:enrich}A, where only in the left-most panel
+doesn't seem to be a separation between the low $\mbox{ARC}$ arm and
+the other.
 
@@
   \centering
-  \includegraphics[width = .8\textwidth]{../figs/Carroll_mice_for_paper/FoxA1_enrichment.pdf}
-  \includegraphics[width = .8\textwidth,page =3]{../figs/Carroll_mice_for_paper/Local_SCC_indicator_by_strata.pdf}
-  \includegraphics[width = .8\textwidth,page =4]{../figs/Carroll_mice_for_paper/Local_SCC_indicator_by_strata_sameRegion.pdf}
+  \includegraphics[width = .7\textwidth]{../figs/Carroll_mice_for_paper/FoxA1_enrichment.pdf}
+  \includegraphics[width = .7\textwidth,page =3]{../figs/Carroll_mice_for_paper/Local_SCC_indicator_by_strata.pdf}
+  \includegraphics[width = .7\textwidth,page =4]{../figs/Carroll_mice_for_paper/Local_SCC_indicator_by_strata_sameRegion.pdf}
   \caption{Using the mouse-FoxA1 experiment from \cite{exoillumina}:
@@
     experiments divided by condition and biological replicate and C)
-    Boxplots of the a set of ERR336942 regions, stratified as in (B)
-    with the $\mbox{local-NSC}$ calculated with the different
-    biological replicates.}
+    Boxplots of the a set of the first replicate regions, stratified
+    as in (B) with the $\mbox{local-NSC}$ calculated with the
+    different biological replicates.}
   \label{fig:enrich}
@@
 fragments into into the ones that overlap with any of the sample's
-peaks. The p.value for ERR36935 is \Sexpr{pval[[1]]}, for ERR336942 is
-\Sexpr{pval[[2]]} and for ERR336956 is \Sexpr{pval[[3]]}. This shows
-that only the middle sample from figure \ref{fig:strand} is the least
-imbalance sample from the the three, while the other two still show a
-higher imbalance at the \Sexpr{de} level.
+peaks. The p.value for the third replicate is \Sexpr{pval[[1]]}, for
+first replicate is \Sexpr{pval[[2]]} and for the second is
+\Sexpr{pval[[3]]}. This shows that only the middle sample from figure
+\ref{fig:strand} is the least imbalance sample from the the three,
+while the other two still show a higher imbalance at the \Sexpr{de}
+level.
 
@@
 
-\subsubsection{Comparison with ChIP-Seq data using dPeak}
+\subsection{Comparison with ChIP-Seq data using dPeak}
 \label{sec:dpeak_analysis}
@@
 
+\section{Software packages}
+\label{sec:software}
+
+For this project several R packages were created or updated:
+
+\begin{itemize}
+\item \textbf{dPeak}: We updated the initialization strategy. The
+  latest version is currently available from
+  \url{http://dongjunchung.github.io/dpeak/}.
+\item \textbf{ChIPexoQual}: This package contains the QC pipeline for
+  ChIP-exo. The last version is available in
+  \url{https://github.com/welch16/ChIPexoQual}.
+\item \textbf{Segvis}: The goal of this package is to visualize
+  genomic regions by using aligned reads such as in figures
+  \ref{fig:exo} or \ref{fig:qcdiagram}.  The latest version is
+  available in \url{https://github.com/keleslab/Segvis}.
+\item \textbf{ChIPUtils}: This package attempts to gather the most
+  commonly used ChIP-Seq QC (as in table \ref{tab:qcbase} or figure
+  \ref{fig:scc}) indicators and additional utilities (as figure
+  \ref{fig:comp}A) for ChIP-Seq. The latest available version is in
+  \url{https://github.com/welch16/ChIPUtils}.
+\end{itemize}
+
 \section{Conclusions}
@@
 
+\subsection{Analysis of E. Coli's Transcription Initiation Complexes}
+\label{sec:ecoli}
+
 \newpage
@@
 
+\section*{Supplement}
+\label{sec:supp}
+
+\subsection*{Additional Enrichment plots for $\sigma^{70}$}
+\label{sec:enrichsup}
+
+In figure \ref{fig:enrich2} it is shown the same relationship as in
+figure \ref{fig:enrich}, but considering only regions formed where the
+reads are being allocated in more than 10 (A) and 30 (B) unique
+positions respectively. Both plots show that the vertical arms formed
+by regions with low $\mbox{ARC}$ is formed by low complexity regions,
+that way suggesting that this segment correspond to the background of
+a ChIP-exo experiment.
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width = .8\textwidth,page = 3]{../figs/Carroll_mice_for_paper/FoxA1_enrichment.pdf}
+  \includegraphics[width = .8\textwidth,page = 4]{../figs/Carroll_mice_for_paper/FoxA1_enrichment.pdf}
+  \caption{Hexbin plots of $\mbox{ARC}$ vs $\mbox{URCR}$ for each
+    region after partitioning the genome. In A and B the regions with
+    reads mapped to at most 10 and 30 positions respectively where not
+    considered. $\mbox{ARC}$ is defined as the ratio of the nr. of
+    reads and the width of a region and $\mbox{URCR}$ is the ratio of
+    the number of unique position where the reads are being allocated
+    and the number of reads in a region.}
+  \label{fig:enrich2}
+\end{figure}
+
 \end{document}

### 2015-11-23 rmarkdown/prelims_paper.tex
Added section about software

@@
 
+\subsubsection{Local Normalized Strand cross-correlation}
+\label{sec:locNSC}
+
+We considered the creation of a novel QC measure for ChIP-exo called
+local normalized SCC or $\mbox{local-NSC}$.
+
 \subsubsection{Enrichment and library complexity in ChIP-exo data}
@@
 fragments into into the ones that overlap with any of the sample's
-peaks. The p.value for ERR36935 is 0.1039, for ERR336942 is
-0.0064 and for ERR336956 is 0.0871. This shows
-that only the middle sample from figure \ref{fig:strand} is the least
-imbalance sample from the the three, while the other two still show a
-higher imbalance at the 100 level.
+peaks. The p.value for the third replicate is 0.1039, for
+first replicate is 0.0064 and for the second is
+0.0871. This shows that only the middle sample from figure
+\ref{fig:strand} is the least imbalance sample from the the three,
+while the other two still show a higher imbalance at the 100
+level.
 

### 2015-11-23 rscripts/carrol_mice_library_complexity_and_enrichment.R

@@ -169,8 +169,12 @@ stats_common <- mclapply(gr,function(x)build_stats(common,x),mc.cores = 6)
 aux <- mapply(function(x,y)x[,sample := y],stats,names(stats),SIMPLIFY =FALSE)
 aux <- do.call(rbind,aux)
 
+aux[, sample := plyr::mapvalues(sample,from = bamfiles,to =  repl)]
+
+
 pdf(file = file.path(figs_dir,"FoxA1_number_unique_positions_perSample.pdf"))
-ggplot(aux[npos > 10] , aes(npos))+  geom_histogram()+scale_x_log10()+xlim(0,100)+facet_wrap( ~  sample)
+ggplot(aux[npos > 10] , aes(npos))+  geom_histogram()+scale_x_log10()+xlim(0,100)+
+  facet_wrap( ~  sample)
 dev.off()
 
 library(hexbin)
@@ -315,6 +319,8 @@ dat[ , nsc := max / noise ]
 
 dat[ , strata := factor(strata, levels = rev(strati) ) ]
 
+dat[ , file := plyr::mapvalues(file, from = bamfiles, to = repl)]
+
 ## dat[ , strata := plyr::mapvalues(strata ,
 ##          from = c(
 ##            "high",
@@ -431,7 +437,7 @@ dat2 <- merge(ss,nsc,by = "match",allow.cartesian = TRUE)
 dat2[ , nsc := max / noise ]
 dat2[ , strata := factor(strata, levels = rev(strati) ) ]
 
-dat2[  , file := gsub(".bam","",file)]
+dat2[  , file := plyr::mapvalues(file,from = bamfiles, to = repl)]
 
 strata_plots <- list()
 strata_plots[[1]] <- ggplot(dat2 , aes( strata , noise,colour = file))+geom_boxplot()+facet_grid( . ~ file) +

### 2015-11-24 rmarkdown/prelims_paper.Rnw
made some minor corrections in the paper, and fixed the figures to be named as reps instead of the bam files

@@
 \usepackage{wrapfig}
+
+\usepackage{tikz}
+\usetikzlibrary{shapes,arrows}
 
@@
 FoxA1 factor in mouse (Serandour et al., 2013 \cite{exoillumina}); and
-generated $\sig$ factor in Escherichia coli (E. Coli) under
+generated $\sig$ factor in Escherichia Coli (E. Coli) under
 aerobic ($ + O_2$) condition, and treated by rifampicin by 0 and 20
@@
 
-\begin{table}[h]
-  \centering
-  \caption{Usual quality control indicators applied to the $\sigma^{70}$ samples.
-PBC stands PCR-bottleneck coefficient, SSD for standardized standard deviation, NSC for normalized strand cross-correlation coefficient (Another measure related to the strand cross-correlation curve is the RSC was ommited since input sample usually don't exists for ChIP-exo).
-$\sig$ samples were given by Robert Landick, FoxA1 and ER samples are from Serandour et al., 2013 \cite{exoillumina} and CTCF sample is from Rhee and Pugh, 2011 \cite{exo1}.}
-
-\begin{tabular}{l|l|r|r|r|r|r|r}
-\hline\hline
-IP & Organism  &Condition & Rep. & Nr. reads & PBC & SSD &  NSC  \\
-\hline\hline
-$\sig$ & E.Coli & Rif-0min & 1 & 960,256 & 0.2823 & 0.0361 &  10.29 \\
-\hline
-$\sig$ &  E.Coli & Rif-0min & 2 & 2,247,295 & 0.2656 & 0.1091  &  25.08  \\
-\hline
-$\sig$ &  E.Coli & Rif-20min & 1 & 1,940,387 & 0.2698 & 0.0820  & 17.69  \\
-\hline
-$\sig$ &  E.Coli & Rif-20min & 2 & 4,229,574 & 0.2153 & 0.1647  &  14.11 \\
-\hline
-FoxA1 &  Mouse &  NA & 1 & 22,210,461 & 0.6562 & 9.12 $\times 10^{-5}$ & 21.452 \\
-\hline
-FoxA1 &  Mouse &  NA & 2 & 22,307,557 & 0.7996 & 7.94 $\times 10^{-5}$ & 60.661 \\
-\hline
-FoxA1 &  Mouse &  NA & 3 & 22,421,729 & 0.1068 & 1.31 $\times 10^{-4}$ & 72.312 \\
-\hline
-ER & Human & NA & 1 & 9,289,835 & 0.8082 & 3.64 $\times 10^{-5}$ & 19.843 \\
-\hline
-ER & Human & NA & 2 & 11,041,833 & 0.8024 & 4.6 $\times 10^{-5}$ & 21.422 \\
-\hline
-ER & Human & NA & 3 & 12,464,836 & 0.8203 & 4.89 $\times 10^{-5}$ & 19.699 \\
-\hline
-CTCF & Human & NA & 1 &   48,478,450 & 0.4579 & 1.29 $\times 10^{-4}$ & 15.977 \\
-\hline
-\end{tabular}  
-  \label{tab:qcbase}
-\end{table}
-
 Table \ref{tab:qcbase} contains measures that are commonly used with
@@
 necessary for ChIP-exo data.
+
+\begin{table}[H]
+  \centering
+  \caption{Usual quality control indicators applied to the $\sigma^{70}$ samples.
+PBC stands PCR-bottleneck coefficient, SSD for standardized standard deviation, NSC for normalized strand cross-correlation coefficient (Another measure related to the strand cross-correlation curve is the RSC was ommited since input sample usually don't exists for ChIP-exo).
+$\sig$ samples were given by Robert Landick, FoxA1 and ER samples are from Serandour et al., 2013 \cite{exoillumina} and CTCF sample is from Rhee and Pugh, 2011 \cite{exo1}.}
+
+\begin{tabular}{l|l|r|r|r|r|r|r}
+\hline\hline
+IP & Organism  &Condition & Rep. & Nr. reads & PBC & SSD &  NSC  \\
+\hline\hline
+$\sig$ & E.Coli & Rif-0min & 1 & 960,256 & 0.2823 & 0.0361 &  10.29 \\
+\hline
+$\sig$ &  E.Coli & Rif-0min & 2 & 2,247,295 & 0.2656 & 0.1091  &  25.08  \\
+\hline
+$\sig$ &  E.Coli & Rif-20min & 1 & 1,940,387 & 0.2698 & 0.0820  & 17.69  \\
+\hline
+$\sig$ &  E.Coli & Rif-20min & 2 & 4,229,574 & 0.2153 & 0.1647  &  14.11 \\
+\hline
+FoxA1 &  Mouse &  NA & 1 & 22,210,461 & 0.6562 & 9.12 $\times 10^{-5}$ & 21.452 \\
+\hline
+FoxA1 &  Mouse &  NA & 2 & 22,307,557 & 0.7996 & 7.94 $\times 10^{-5}$ & 60.661 \\
+\hline
+FoxA1 &  Mouse &  NA & 3 & 22,421,729 & 0.1068 & 1.31 $\times 10^{-4}$ & 72.312 \\
+\hline
+ER & Human & NA & 1 & 9,289,835 & 0.8082 & 3.64 $\times 10^{-5}$ & 19.843 \\
+\hline
+ER & Human & NA & 2 & 11,041,833 & 0.8024 & 4.6 $\times 10^{-5}$ & 21.422 \\
+\hline
+ER & Human & NA & 3 & 12,464,836 & 0.8203 & 4.89 $\times 10^{-5}$ & 19.699 \\
+\hline
+CTCF & Human & NA & 1 &   48,478,450 & 0.4579 & 1.29 $\times 10^{-4}$ & 15.977 \\
+\hline
+\end{tabular}  
+  \label{tab:qcbase}
+\end{table}
+
+\vspace*{\fill}
 
@@
   \centering
-  \includegraphics[width = .9\textwidth]{../figs/for_paper/coverage_diagram.pdf}
-  \caption{Diagram of the ChIP-exo Quality control pipeline [this is a
-    place-holder]. The genome is partitioned into several regions by
-    removing the empty segments and for each regions we calculate a
-    collection of QC indicators.}
+  \includegraphics[width = .7\textwidth]{../figs/for_paper/coverage_diagram2.pdf}
+  \caption{Diagram of the ChIP-exo Quality control pipeline. The
+    genome is partitioned into several regions by removing the empty
+    segments and for each regions we calculate a collection of QC
+    indicators, then for each region $R_k$ a vector of summary
+    statistics $T(R_k)$ is calculated.}
   \label{fig:qcdiagram}
@@
 We considered the creation of a novel QC measure for ChIP-exo called
-local normalized SCC or $\mbox{local-NSC}$.
+local normalized SCC or $\mbox{local-NSC}$. This measure
+
+\begin{align}
+  y(\delta) = r\left[ n^+ \left(x + \frac{\delta}{2}
+    \right), n^- \left( x- \frac{\delta}{2} \right)\right]
+\nonumber
+\end{align}
+
+where again $r$ is the Pearson correlation but $n^+$ and $n^-$ are the
+coverage vectors of a given region in the genome. Figure
+\ref{fig:locSCC} shows that as the library complexity of the region is
+lower, then the local SCC seems to show higher variance. Hence, we
+fit a nonparametric regression model to smooth the SCC signal:
+
+\begin{align}
+ y(\delta) &= f(x_\delta) + \epsilon_\delta \nonumber
+\end{align}
+
+where $y(\delta)$ is the observed SCC for shift $x_\delta$,
+$\epsilon_\delta \sim N(0,\sigma_f^2)$. Finally the $\mbox{local-NSC}$
+is defined as a signal-to-noise ratio by using:
+
+\begin{align}
+    \mbox{local-NSC} &= \frac{max_{x_\delta} \hat{f}(x_\delta)}{\hat{\sigma}_f} \label{locNSC}
+\end{align}
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width = .8\textwidth]{../figs/for_paper/local_SCC_example.pdf}
+  \caption{Using the FoxA1 sample from \cite{exoillumina}. A) shows an
+    example region for the 3 replicates. The library complexity seem
+    to be comparable between the two top rows and both outperfom the
+    last row. The number of reads for the first replicate is higher
+    than the number of reads for the other replicates. B) Shows the
+    local SCC for the 3 replicates, this measure is comparable among
+    the 3 replicates. The first replicate shows the lowest residual
+    error, while the third one shows the highest. In blue, a ``loess''
+    regression model is fitted.}
+\label{fig:locSCC}
+\end{figure}
 
@@
 necessary to address the balance between the enrichment and library
-complexity. To diagnose this, we considered the Average Read Coverage
+complexity. To diagnose this, we consider the Average Read Coverage
 (ARC) and the Unique Read Coverage Rate (URCR) which are defined as:
@@
   \includegraphics[width = .7\textwidth,page =3]{../figs/Carroll_mice_for_paper/Local_SCC_indicator_by_strata.pdf}
-  \includegraphics[width = .7\textwidth,page =4]{../figs/Carroll_mice_for_paper/Local_SCC_indicator_by_strata_sameRegion.pdf}
+%  \includegraphics[width = .7\textwidth,page =4]{../figs/Carroll_mice_for_paper/Local_SCC_indicator_by_strata_sameRegion.pdf}
   \caption{Using the mouse-FoxA1 experiment from \cite{exoillumina}:
@@
     mapped to only one position in the regions for the ChIP-exo
-    experiments divided by condition and biological replicate and C)
-    Boxplots of the a set of the first replicate regions, stratified
-    as in (B) with the $\mbox{local-NSC}$ calculated with the
-    different biological replicates.}
+    experiments divided by condition and biological replicate.}
   \label{fig:enrich}
 \end{figure}
+
+% and C) Boxplots of the a set of the first replicate regions,
+% stratified as in (B) with the $\mbox{local-NSC}$ calculated with the
+% different biological replicates.
 
@@
 fragments into into the ones that overlap with any of the sample's
-peaks. The p.value for the third replicate is \Sexpr{pval[[1]]}, for
-first replicate is \Sexpr{pval[[2]]} and for the second is
+peaks. Then we considered a Kolmogorov-Smirnov tests to assess if the
+distribution of both groups were the same. The p.value for the third
+replicate's test is \Sexpr{pval[[1]]}, for the first replicate's test
+is \Sexpr{pval[[2]]} and for the second test is
 \Sexpr{pval[[3]]}. This shows that only the middle sample from figure
@@
   \includegraphics[width = .46\textwidth]{../figs/for_paper/resolution_by_dataset_old_data.pdf}
+\end{figure}
+
+\begin{figure}[H]
+\centering
    \includegraphics[width = .46\textwidth,page = 1]{../figs/for_paper/sigma_delta_old_densities.pdf}
@@
 
-\subsection{GC content and mappability quality indicators}
+\subsection{Some additional thoughts for this analysis}
+
+\subsubsection{GC content and mappability quality indicators}
 \label{sec:gc_map_qc}
@@
 
-\subsection{Quality control for ChIP-Nexus}
+\subsubsection{Quality control for ChIP-Nexus}
 \label{sec:nexus}
@@
 
-\subsection{Analysis of E. Coli's Transcription Initiation Complexes}
+\subsection{Analysis of E. Coli Transcription Initiation Complexes}
 \label{sec:ecoli}
+
+The data showed in figure \ref{fig:exo_example} is actually part of a
+more interesting experiment. We have 2 biological replicates of the
+$\sig$, $\beta'_f$ and $\beta$ transcription factors under two
+conditions, one where rifampicin was applied by 20 minutes and another
+where it wasn't applied. As seen in the figure, there is ChIP-exo data
+for the 3 factors and PET ChIP-Seq for $\sig$ and $\beta'_f$ (in the
+figure both replicates were pooled).
+
+$\sig$ factor is a transcription initiation factor of housekeeping
+genes in E. Coli, and both $\beta$ factors transcribe the DNA into
+mRNA. Hence, we want to use this data to have a better understanding
+of the transcription in E. Coli.
+
+The steps in formation of productive transcription complexes are:
+
+\begin{enumerate}
+\item Binding of RNAP to rpomoter DNA to form a closed complex.
+\item Rearrangement of RNAP-DNA contacts to form an open complex.
+\item Binding of nucleoside triphosphates in the active site and
+  synthesis of a initial RNA to yield an initial transcription complex
+  (ITC) - this is sometimes called a ``scrunched complex''.
+\item Promoter escape to form a elongated complex (EC$_1$) when the
+  promoter contacts are broken, and the $\sig$ factor is realeased.
+\item Conversion of EC$_1$ into EC$_2$ by loading of the ribosome and
+  elongation factors to enable efficient transcript elongation.
+\end{enumerate}
+
+% \begin{figure}[H]
+% \centering
+% \begin{tikzpicture}[node distance = 3cm]
+% \tikzstyle{region} = [draw,thin,fill = blue!20]
+% \tikzstyle{line} = [draw,thick]
+% %\path[use as bounding box] (-1,0) rectangle (10,-2);
+% \path[-] node[region] (step1) {RNAP + Promoters};
+% \path[-] node[region,right of =step1,fill = red!20] (step2) {Closed};
+% \path[-] node[region,right of =step2,fill = green!20] (step3) {Intermediate};
+% \path[-] node[region,right of =step3,fill = yellow!20] (step4) {Open};
+% \path[line][<->] (step1) -- (step2);
+% \path[line][<->] (step2) -- (step3);
+% \path[line][<->] (step3) -- (step4);
+% \end{tikzpicture}
+% \caption{Zooming into transcription}
+% \end{figure}
+
+As a proof of concept we were able to identify open and closed regions
+by centering the binding the peaks respecto to the highest $\sig$
+summit (among the rif0 and rif20 samples), and then using hierachical
+clustering on the $\beta'_f$ signal of the rif0 samples to produce:
+
+\begin{figure}[H]
+  \centering
+ \includegraphics[width = \textwidth]{/p/keles/ChIPexo/volume5/OpenClosedComplexes/figs/filtered/heatmaps_centroid.png}
+ \caption{Centered and normalized heatmaps for PET ChIP-Seq peaks of
+   $\sig$ and $\beta'_f$ under rif0 and rif20. The regions are
+   organized by the centroids clustering of the $\beta'_f$-rif0
+   profiles.}
+  \label{fig:hm1}
+\end{figure}
+
+Figure \ref{fig:hm1} shows heatmaps of the normalized coverages of PET
+ChIP-Seq peaks for $\beta'_f$ and $\sig70$ under rif0 and rif20
+treatments. Each row represents the same region, under the four
+different samples. As expected the signal for both $\sig$ samples is
+centered around the middle, for $\beta'_f$-rif20, the majority of he
+signal seems to be centered, hence the trascription was interrupted
+and for $\beta'_f$-rif0 we can see clusters formed by open regions.
+
+Then, using a set of high quality PET ChIP-Seq peaks and clustering
+with the mclust model by Fraley and Raftery, 2007 \cite{mclust} we
+were able to label the a set of regions as closed or open
+complexes. Hence, we expect that by using ChIP-exo data we may find
+different patterns that represent transcription units whose output was
+limited at differemt steps in the formation of EC$_2$.
+
+Additionally if we find several cases of those patterns, then we may
+be able to determine which sequences may be causing this behaviour at
+the different steps.
+
+\subsection{Enhancer prediction and active learning}
+\label{sec:enhancer}
+
+Last summer, we entered to a contest from ENCODE's functional
+characterization group to fit a model to classify enhancers. Roughly
+speaks, the data set consided of a set of regions
+$\mathcal{R}_1,\cdots,\mathcal{R}_n$ and a set of labels
+$Y_1,\cdots,Y_n \in \{0,1,\}$. Hence, we built a series of features
+$X_1,\cdots,X_p$ and fitted a model by dividing the $n$ observations
+into a training set and a test set, etc. However, there are in total
+another $m >> n$ regions in the experiment, and the cost of labelling
+those $m$ regions may be high (in either the actual monetary cost of
+the experiment or the time the scientists spends labelling those
+regions). Those $n+m$ regions are H3K27ac peaks, hence they were
+ranked and the first $n$ were selected randomly from groups with high,
+medium and low signal. This raises the following questions:
+
+\begin{itemize}
+\item Is there a better way of ranking this $n+m$ regions ?
+\item What is the best way to select the regions that are going to be
+  labelled, i.e. we would want to label the regions that gives us the
+  best model.
+\end{itemize}
+
+Hence, it may be worth to investigate which active learning strategies
+may be the best to apply in this particular problem. Settles, 2012
+\cite{al} is a survey of acive learning techniques that we want to
+explore in order to solve this problems.
 
@@
 \nocite{maplot2}
+\nocite{esl}
+\nocite{anp}
+\nocite{al}
+\nocite{oc_review}
+\nocite{exo4}
 

### 2015-11-24 rmarkdown/prelims_paper.tex
made some minor corrections in the paper, and fixed the figures to be named as reps instead of the bam files

@@
 fragments into into the ones that overlap with any of the sample's
-peaks. The p.value for the third replicate is 0.1039, for
-first replicate is 0.0064 and for the second is
+peaks. Then we considered a Kolmogorov-Smirnov tests to assess if the
+distribution of both groups were the same. The p.value for the third
+replicate's test is 0.1039, for the first replicate's test
+is 0.0064 and for the second test is
 0.0871. This shows that only the middle sample from figure

### 2015-11-24 rscripts/carrol_mice_library_complexity_and_enrichment.R

@@ -169,8 +169,8 @@ stats_common <- mclapply(gr,function(x)build_stats(common,x),mc.cores = 6)
 aux <- mapply(function(x,y)x[,sample := y],stats,names(stats),SIMPLIFY =FALSE)
 aux <- do.call(rbind,aux)
 
-aux[, sample := plyr::mapvalues(sample,from = bamfiles,to =  repl)]
-
+aux[, sample := plyr::mapvalues(sample , from = c("ERR336935.bam","ERR336942.bam","ERR336956.bam"),
+    to = c("rep-3","rep-1","rep-2"))]
 
 pdf(file = file.path(figs_dir,"FoxA1_number_unique_positions_perSample.pdf"))
 ggplot(aux[npos > 10] , aes(npos))+  geom_histogram()+scale_x_log10()+xlim(0,100)+
@@ -319,7 +319,8 @@ dat[ , nsc := max / noise ]
 
 dat[ , strata := factor(strata, levels = rev(strati) ) ]
 
-dat[ , file := plyr::mapvalues(file, from = bamfiles, to = repl)]
+dat[ , file := plyr::mapvalues(file, from = c("ERR336942.bam","ERR336956.bam","ERR336935.bam"),
+  to = c("rep-1","rep-2","rep-3"))]         
 
 ## dat[ , strata := plyr::mapvalues(strata ,
 ##          from = c(

### 2015-11-24 rscripts/carroll_mice_strand_imbalance.R

@@ -163,6 +163,15 @@ library(grid)
 library(gridExtra)
 
 
+fsr_dt[, sample := plyr::mapvalues(sample,
+    from = c("ERR336935.bam","ERR336942.bam","ERR336956.bam"),
+    to = c("rep-3","rep-1","rep-2"))]
+
+label_dt[, sample := plyr::mapvalues(sample,
+    from = c("ERR336935.bam","ERR336942.bam","ERR336956.bam"),
+    to = c("rep-3","rep-1","rep-2"))]
+
+
 
 pdf(file = file.path(figs_dir,"Strand_imbalance.pdf"),width = 9 , height = 5)
 p1 <- ggplot(fsr_dt,aes(depth , fsr , colour = quantiles))+geom_line(size = 1)+

### 2015-11-24 rmarkdown/slides.Rnw
Added instruction to Makefile to make slides for prelim presentation

\documentclass[table,xcolor=dvipsnames]{beamer}

\usepackage{xcolor}
\usepackage{url}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows}
\usepackage{float}
\usepackage{verbatim}
\usepackage{hyperref}

\usepackage{lmodern,tabularx,ragged2e,booktabs}
\newcolumntype{Y}{>{\arraybackslash\RaggedRight}X}
\newcolumntype{P}[1]{>{\arraybackslash\RaggedRight}p{#1}}

\usetheme{Szeged}
\usecolortheme[named=RoyalBlue]{structure}

% Personal Data

\title[High Resolution Identification of Protein-DNA Binding Events]{High Resolution Identification of Protein-DNA Binding Events
  and Quality Control for ChIP-exo data}

\author{Rene Welch\\Preliminary Examination}

\institute[UW-STAT]{Department of Statistics\\University of Wisconsin - Madison}

\date{December 1st, 2015}

\newcommand{\sig}{\sigma^{70}}

\begin{document}

\begin{frame}
  \maketitle
\end{frame}

\begin{frame}
\frametitle{Outline}
  \tableofcontents
\end{frame}

\section{ChIP-exo procedure}

\begin{frame}
\frametitle{ChIP-exo procedure}
  \begin{figure}[H]
    \centering
\includegraphics{../figs/diagrams/ChIPexo_diagram1.png}
\includegraphics{../figs/diagrams/ChIPexo_diagram2.png}\\
\includegraphics{../figs/diagrams/ChIPexo_diagram3.png}
    \caption{ChIP-exo procedure, the diagram is from Furey, 2012 \cite{chipbeyond}}
    \label{fig:exo_diagram}
  \end{figure}
\end{frame}

\section{ChIP-Seq QC measures}
\begin{frame}
  \frametitle{ChIP-Seq QC measures}

\only<1>{
\setlength\tabcolsep{3pt}  % default value: 6pt
\scriptsize  %%  command to change the font size
\rowcolors{1}{}{RoyalBlue!20} %-- this indicates the change in odd and pair rows
\begin{tabularx}{\textwidth}{ @{} c|   Y @{}}
  \toprule
  \textbf{QC measure} & \textbf{Definition} \\
  \midrule
  Nr. reads & Self-explanatory. The higher the better...  \\
  PCR bottleneck Coeff. &  Ratio of number of pos. to which EXACTLY one read maps and number  of pos. to which AT LEAST one read maps \\
  Standarized Std. Dev. &  Normalized Std. Deviation of the sequencing coverage \\
  Strand Cross-Corr. &  $  y(\delta) = \sum_c w_c r\left[ n_c^+ \left(x + \frac{\delta}{2}
    \right), n_c^- \left( x- \frac{\delta}{2} \right)\right]$ \\
  Normalized SCC &  Ratio of max value of SCC and min value of SCC$^*$  \\
  \bottomrule
\end{tabularx}
}

\only<1>{
where $n_c^S$ is the coverage for chromosome $c$ and strand $S$. $r$ is
the Pearson correlation and $w_c$ is the proportion of reads in the
experiment for chromosome $c$
}

\only<2>{
\setlength\tabcolsep{3pt}  % default value: 6pt
\scriptsize  %%  command to change the font size
% \rowcolor{RoyalBlue!20} %-- this indicates the change in odd and pair rows
\begin{tabularx}{\textwidth}{ @{} c|c|c|c|c|c|c| Y @{}}
  \toprule
IP & Organism  &Condition & Rep. & Nr. reads & PBC & SSD &  NSC  \\
\midrule
\rowcolor{RoyalBlue!20}$\sig$ & E.Coli & Rif-0min & 1 & 960,256 & 0.2823 & 0.0361 &  10.29 \\
\hline
\rowcolor{RoyalBlue!20}$\sig$ &  E.Coli & Rif-0min & 2 & 2,247,295 & 0.2656 & 0.1091  &  25.08  \\
\hline
\rowcolor{RoyalBlue!20}$\sig$ &  E.Coli & Rif-20min & 1 & 1,940,387 & 0.2698 & 0.0820  & 17.69  \\
\hline
\rowcolor{RoyalBlue!20}$\sig$ &  E.Coli & Rif-20min & 2 & 4,229,574 & 0.2153 & 0.1647  &  14.11 \\
\hline
\rowcolor{Red!20}FoxA1 &  Mouse &  - & 1 & 22,210,461 & 0.6562 & 9.12 $\times 10^{-5}$ & 21.452 \\
\hline
\rowcolor{Red!20}FoxA1 &  Mouse &  - & 2 & 22,307,557 & 0.7996 & 7.94 $\times 10^{-5}$ & 60.661 \\
\hline
\rowcolor{Red!20}FoxA1 &  Mouse &  - & 3 & 22,421,729 & 0.1068 & 1.31 $\times 10^{-4}$ & 72.312 \\
\hline
\rowcolor{Green!20}ER & Human & - & 1 & 9,289,835 & 0.8082 & 3.64 $\times 10^{-5}$ & 19.843 \\
\hline
\rowcolor{Green!20}ER & Human & - & 2 & 11,041,833 & 0.8024 & 4.6 $\times 10^{-5}$ & 21.422 \\
\hline
\rowcolor{Green!20}ER & Human & - & 3 & 12,464,836 & 0.8203 & 4.89 $\times 10^{-5}$ & 19.699 \\
\hline
\rowcolor{Yellow!20}CTCF & Human & - & 1 &   48,478,450 & 0.4579 & 1.29 $\times 10^{-4}$ & 15.977 \\
\bottomrule
\end{tabularx}  
}

\end{frame}

\section{Comparison of ChIP-exo and ChIP-seq}

\begin{frame}[plain]
  \frametitle{Comparison of ChIP-exo and ChIP-seq - SCC}  
\begin{figure}[H]
  \centering  
\includegraphics[width = .7\textwidth]{../figs/for_paper/scc_ctcf.pdf}
\caption{SCC for CTCF factor in HeLa cell line for ChIP-exo and
  SET-ChIP-Seq }
  \label{fig:scc}
\end{figure}

\begin{itemize}
\item There is a \emph{``phantom peak''} at read length.
\item In ChIP-Seq SCC is maximized at the unobserved fragment length.
\item In ChIP-exo, the \emph{``phantom peak''} and the frag. length
  summit are confounded.
\end{itemize}

\end{frame}

\begin{frame}
  \frametitle{Comparison of ChIP-exo and ChIP-Seq}

\begin{figure}[H]
  \centering
  \includegraphics[width = .5\textwidth,page = 3 ]{../figs/for_paper/ChIPseqPET_ChIPexo_tagCount_comparison.pdf}
  \includegraphics[width = .5\textwidth]{../figs/for_paper/forward_strand_ratio_comp_old.pdf}
\end{figure}

\begin{itemize}
\item
\item 
\end{itemize}

% \only<2>{
% \begin{figure}[H]
% \centering
% \includegraphics[width = .46\textwidth,page = 1]{../figs/for_paper/eukaryotic_bias_CTCF.pdf}
%   \includegraphics[width = .46\textwidth,page = 2]{../figs/for_paper/eukaryotic_bias_CTCF.pdf}
% \end{figure}
% }
\end{frame}

\begin{frame}[t]
\frametitle{Software}  
\begin{itemize}
\item \textbf{dPeak}: We updated the initialization strategy. The
  latest version is currently available from
  \url{http://dongjunchung.github.io/dpeak/}.
\item \textbf{ChIPexoQual}: This package contains the QC pipeline for
  ChIP-exo. The last version is available in
  \url{https://github.com/welch16/ChIPexoQual}.
\item \textbf{Segvis}: The goal of this package is to visualize
  genomic regions by using aligned reads. The latest version is
  available in \url{https://github.com/keleslab/Segvis}.
\item \textbf{ChIPUtils}: This package attempts to gather the most
  commonly used ChIP-Seq QC.The latest available version is in
  \url{https://github.com/welch16/ChIPUtils}.
\end{itemize}

\end{frame}

\begin{frame}
  \frametitle{References}

{\tiny

\nocite{exo1}
\nocite{dpeak}

\bibliographystyle{plain} % Style BST file (bmc-mathphys, vancouver, spbasic)
\bibliography{chip_exo_paper}

}

\end{frame}

\end{document}

### 2015-11-24 rmarkdown/slides.Rnw
finished slides

@@
-\documentclass[table,xcolor=dvipsnames]{beamer}
+\documentclass[compress,table,xcolor=dvipsnames]{beamer}
 
@@
 \usecolortheme[named=RoyalBlue]{structure}
+\setbeamertemplate{headline}{}
+\setbeamertemplate{navigation symbols}{}
 
@@
 
-\begin{frame}[plain]
-  \frametitle{Comparison of ChIP-exo and ChIP-seq - SCC}  
+\begin{frame}
+  \frametitle{Comparison of ChIP-exo and ChIP-seq I}  
 \begin{figure}[H]
@@
 \begin{frame}
-  \frametitle{Comparison of ChIP-exo and ChIP-Seq}
+  \frametitle{Comparison of ChIP-exo and ChIP-Seq II}
 
@@
 
+{\small
 \begin{itemize}
-\item
-\item 
+\item A shows that high dnesity regions are similar between ChIP-Seq
+  and ChIP-exo but background regions are not.
+\item The peak-pair assumption doesn't hold in ChIP-exo data, some
+  regions show strand-imbalance
 \end{itemize}
+}
 
@@
 
+\section{The ChIP-exo QC pipeline}
+
+\begin{frame}[plain]
+  \frametitle{The ChIP-exo QC pipeline}
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width = \textwidth]{../figs/for_paper/coverage_diagram2.pdf}
+\end{figure}
+
+{\small
+  \begin{itemize}
+  \item $\mbox{ARC} = \frac{\text{Nr. of reads in the
+        region}}{\text{Width of the region}}$
+  \item $\mbox{URCR} = \frac{\text{Nr. of reads mapped to only one
+        position in the region}}{\text{Nr. of reads in the region}}$
+  \item $\mbox{local-NSC}$
+  \item $\mbox{FSR} = \frac{\text{Nr. of fwd. strand reads in region}}{\text{Nr. of reads in region}}$
+  \end{itemize}  
+}
+
+\end{frame}
+
+\begin{frame}[t,plain]
+  \frametitle{Library complexity and enrichment}
+\only<1>{
+\begin{figure}[H]
+  \centering
+  \includegraphics[width = \textwidth]{../figs/Carroll_mice_for_paper/FoxA1_enrichment.pdf}
+\end{figure}
+}
+
+\only<2>{
+{\small
+\begin{figure}[H]
+  \centering
+  \includegraphics[width = \textwidth]{../figs/for_paper/local_SCC_example.pdf}
+\end{figure}
+}}
+\only<2>{
+\begin{align}
+ y(\delta) = f(x_\delta) + \epsilon_\delta \quad 
+\mbox{local-NSC} = \frac{max_{x_\delta} \hat{f}(x_\delta)}{\hat{\sigma}_f}
+\nonumber
+\end{align}
+}
+
+\only<3>{
+\begin{figure}[H]
+\centering
+  \includegraphics[width = \textwidth,page =3]{../figs/Carroll_mice_for_paper/Local_SCC_indicator_by_strata.pdf}
+\end{figure}
+}
+
+\only<3>{
+{\small
+where:
+\begin{itemize}
+\item high - regions with nr. of unique positions $ > 100$
+\item med - regions with nr. of unique positions in $(50,100)$
+\item low - regions with nr. of unique positions in $(20,100)$
+\end{itemize}
+}}
+
+\end{frame}
+
+\begin{frame}[t]
+  \frametitle{Strand imbalance}
+
+\begin{figure}[H]
+  \centering  
+  \includegraphics[width = \textwidth,page = 3]{../figs/Carroll_mice_for_paper/Strand_imbalance.pdf} 
+  \label{fig:strand}
+\end{figure}
+
+\end{frame}
+
+\section{Comparison with ChIP-Seq using dPeak}
+\begin{frame}[t]
+  \frametitle{dPeak model for SET case}
+
+\only<1>{
+  We consider a region with $n$ reads and $m$ positions, for the
+  $i$-th read:
+  \begin{itemize}
+  \item $Z_i \sim \mbox{Multi}(\pi_0,\pi_1,\cdots,\pi_{g^*})$
+  \item $D_i \sim \mbox{Ber}(p_D)$
+    \begin{itemize}
+    \item The read is in the forward strand ($D_i = 1$):
+      \begin{itemize}
+      \item The reads belongs to the background: $R_i | Z_i = 0, D_i =
+        1 \sim \mbox{Unif}(1 - \beta + 1, m)$
+      \item The read belong to the $g$-th binding event: $R_i | Z_i =
+        g, D_i = 1 \sim \mbox{N}(\mu_g - \delta , \sigma^2)$
+      \end{itemize}
+    \item The read is in the backward strand ($D_i = 0$):
+      \begin{itemize}
+      \item The reads belongs to the background: $R_i | Z_i = 0, D_i =
+        0 \sim \mbox{Unif}(1 , m + \beta - 1)$
+      \item The read belong to the $g$-th binding event: $R_i | Z_i =
+        g, D_i = 0 \sim \mbox{N}(\mu_g + \delta , \sigma^2)$
+      \end{itemize}
+    \end{itemize}
+  \end{itemize}
+}
+
+\only<2>{
+      \begin{figure}[H]
+        \centering
+        \includegraphics[width =
+        .8\textwidth]{/p/keles/ChIPexo/volume3/ChIPexo/poster/dpeak.png}
+      \end{figure}
+}
+\end{frame}
+
+\begin{frame}
+\frametitle{Comparison with ChIP-Seq using dPeak}
+\only<1>{
+  \begin{figure}[H]
+  \centering
+  \includegraphics[width = .46\textwidth]{../figs/for_paper/sensitivity_exo_olda_data.pdf}
+  \includegraphics[width = .46\textwidth]{../figs/for_paper/resolution_by_dataset_old_data.pdf}
+\end{figure}
+}
+\only<1>{
+{\small
+  \begin{itemize}
+  \item Sensitivity is the defined as the proportion of identified
+    peaks (regulonDoB \cite{regulonDB} is used as gold-stantard)
+  \item Resolution is defined as the min. absolute distance of a
+    regulonDB annotation to an est. binding location.
+  \end{itemize}
+}}
+
+\only<2>{
+\begin{figure}[H]
+\centering
+   \includegraphics[width = .46\textwidth,page = 1]{../figs/for_paper/sigma_delta_old_densities.pdf}
+   \includegraphics[width = .46\textwidth,page = 2]{../figs/for_paper/sigma_delta_old_densities.pdf} 
+ \end{figure}
+}
+\only<2>{
+{\small
+\begin{itemize}
+\item $\delta$ measures the average distance of reads to their
+  respective binding sites
+\item $\sigma$ measures the dispersion of reads around their
+  respective binding sites
+\end{itemize}
+}}
+\end{frame}
+
+\begin{frame}[t]
+  \frametitle{ChIP-Seq comparison at fixed depth}
+
+  We sampled $n$ fragment reads of each dataset ($2n$ for PET
+  ChIP-Seq), and applied the MOSAiCS / dPeak pipeline:
+
+\only<1>{
+\begin{figure}[H]
+  \centering
+  \includegraphics[width = .4\textwidth,page = 1]{../figs/for_paper/saturation_analysis_old_frames.pdf}
+  \includegraphics[width = .4\textwidth,page = 2]{../figs/for_paper/saturation_analysis_old_frames.pdf}
+\end{figure}
+}
+\only<1>{
+{\small
+  \begin{itemize}
+  \item ChIP-exo and PET ChIP-Seq are comparable and outperform SET
+    ChIP-Seq
+  \end{itemize}
+}}
+
+\only<2>{
+\begin{figure}[H]
+\centering
+\includegraphics[width = .4\textwidth,page = 3]{../figs/for_paper/saturation_analysis_old_frames.pdf}
+  \includegraphics[width = .4\textwidth,page = 4]{../figs/for_paper/saturation_analysis_old_frames.pdf}
+\end{figure}
+}
+
+\end{frame}
+
+\section{Conclusions and future work}
+
+\begin{frame}
+  \frametitle{Conclusions}
+
+  \begin{itemize}
+  \item Our pipeline is capable of assessing the balance between
+    sample enrichment and library complexity.
+  \item We shown that the ``peak-pair'' assumption doesn't hold well
+    in practice, and implemented a visualization capable of detecting
+    strand imbalance.
+  \item We updated dPeak, which makes a striking balance in
+    sensitivity, specificity and spatial resolution.
+  \item ChIP-exo and PET ChIP-Seq are comparable in resolution and
+    sensitivity, and both outperform SET ChIP-Seq.
+  \item We showed that with a fixed number of reads, ChIP-exo
+    outperforms PET and SET ChIP-Seq.
+  \end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{Future work}
+
+\begin{itemize}
+\item In the paper, we showed that there is a relationship between
+  ChIP-exo tag counts and both mappability and GC content scores. We
+  want to add a QC measure to the pipeline based on them.
+\item We want to assess if ChIP-Nexus library complexity is actually higher than ChIP-exo's by using the $\mbox{local-NSC}$.
+\item We have been studying E. Coli's transcription initiation
+  complexes with PET ChIP-Seq, being able to label regions as open or
+  closed complexes. We want to improve this analysis by using ChIP-exo
+  data, and hopefully detecting intermediate step between this two
+  states.
+\item Find a optimal strategy for labelling enhancer out of a
+  predetermined list of regions in the genome by the use of active
+  learning techniques.
+\end{itemize}
+
+\end{frame}
+
 \begin{frame}[t]
@@
 \begin{itemize}
-\item \textbf{dPeak}: We updated the initialization strategy. The
-  latest version is currently available from
+\item {\color{RoyalBlue}\textbf{dPeak}}: We updated the initialization
+  strategy. The latest version is currently available from
   \url{http://dongjunchung.github.io/dpeak/}.
-\item \textbf{ChIPexoQual}: This package contains the QC pipeline for
-  ChIP-exo. The last version is available in
+\item {\color{RoyalBlue}\textbf{ChIPexoQual}}: This package contains
+  the QC pipeline for ChIP-exo. The last version is available in
   \url{https://github.com/welch16/ChIPexoQual}.
-\item \textbf{Segvis}: The goal of this package is to visualize
-  genomic regions by using aligned reads. The latest version is
-  available in \url{https://github.com/keleslab/Segvis}.
-\item \textbf{ChIPUtils}: This package attempts to gather the most
-  commonly used ChIP-Seq QC.The latest available version is in
-  \url{https://github.com/welch16/ChIPUtils}.
+\item {\color{RoyalBlue}\textbf{Segvis}}: The goal of this package is
+  to visualize genomic regions by using aligned reads. The latest
+  version is available in \url{https://github.com/keleslab/Segvis}.
+\item {\color{RoyalBlue}\textbf{ChIPUtils}}: This package attempts to
+  gather the most commonly used ChIP-Seq QC.The latest available
+  version is in \url{https://github.com/welch16/ChIPUtils}.
 \end{itemize}
+
+\end{frame}
+
+\begin{frame}[plain]
+
+{\Huge
+{\color{RoyalBlue}
+\begin{center}
+  \textbf{Thank you very much!}
+\end{center}
+}
+}
 

### 2015-11-24 rmarkdown/slides.tex
Added instruction to Makefile to make slides for prelim presentation
finished slides

--- 2015-11-24 rmarkdown/slides.Rnw
@@
-\documentclass[table,xcolor=dvipsnames]{beamer}
+\documentclass[compress,table,xcolor=dvipsnames]{beamer}\usepackage[]{graphicx}\usepackage[]{color}
+%% maxwidth is the original width if it is less than linewidth
+%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
+\makeatletter
+\def\maxwidth{ %
+  \ifdim\Gin@nat@width>\linewidth
+    \linewidth
+  \else
+    \Gin@nat@width
+  \fi
+}
+\makeatother
+
+\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
+\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
+\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
+\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
+\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
+\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
+\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
+\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
+\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
+\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
+
+\usepackage{framed}
+\makeatletter
+\newenvironment{kframe}{%
+ \def\at@end@of@kframe{}%
+ \ifinner\ifhmode%
+  \def\at@end@of@kframe{\end{minipage}}%
+  \begin{minipage}{\columnwidth}%
+ \fi\fi%
+ \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
+ \colorbox{shadecolor}{##1}\hskip-\fboxsep
+     % There is no \\@totalrightmargin, so:
+     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
+ \MakeFramed {\advance\hsize-\width
+   \@totalleftmargin\z@ \linewidth\hsize
+   \@setminipage}}%
+ {\par\unskip\endMakeFramed%
+ \at@end@of@kframe}
+\makeatother
+
+\definecolor{shadecolor}{rgb}{.97, .97, .97}
+\definecolor{messagecolor}{rgb}{0, 0, 0}
+\definecolor{warningcolor}{rgb}{1, 0, 1}
+\definecolor{errorcolor}{rgb}{1, 0, 0}
+\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX
+
+\usepackage{alltt}
 
@@
 \newcommand{\sig}{\sigma^{70}}
-
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
 \begin{document}

### 2015-11-26 rmarkdown/prelims_paper.Rnw
Changed section and subsection titles

@@
 
-\title{High Resolution Identification of Protein-DNA Binding Events
-  and Quality Control for ChIP-exo data\vspace*{\fill}}
+\title{Data exploration, quality control and statistical analyses of
+  ChIP-exo experiments\vspace*{\fill}}
 
@@
 2009 \cite{computation}), such as MACS (Zhang et al., 2008
-\cite{macs}), CisGenome (Hongkai et al., 2008 \cite{cisgenome}) and
+\cite{macs}), CisGenome (Ji et al., 2008 \cite{cisgenome}) and
 MOSAiCS (Kuan et al., 2009 \cite{mosaics}), these approaches reveal
@@
 
-\subsubsection{Strand cross-correlation}
+\subsubsection{Strand Cross-Correlation}
 \label{sec:scc}
@@
 
-\subsection{The dPeak model}
+\subsection{The dPeak Model}
 \label{sec:dpeak}
@@
 
-\subsubsection{dPeak's algorithm initialization strategy}
+\subsubsection{A New Initialization Strategy for dPeak}
 \label{sec:algo}
@@
 
-\subsection{ChIP-exo Quality Control pipeline}
+\subsection{ChIP-exo Quality Control Pipeline}
 \label{sec:QC}
@@
 
-\subsubsection{Local Normalized Strand cross-correlation}
+\subsubsection{Normalized Local Strand Cross-Correlation}
 \label{sec:locNSC}
@@
 
-\subsubsection{Enrichment and library complexity in ChIP-exo data}
+\subsubsection{Enrichment and Library Complexity in ChIP-exo data}
 \label{sec:enri}
@@
 
-\subsubsection{Strand imbalance in ChIP-exo data}
+\subsubsection{Strand Imbalance in ChIP-exo data}
 \label{sec:strand_imbalance}
@@
 
-\subsection{Recommendations for the design of ChIP-exo experiments}
+\subsection{Systematic Comparison of ChIP-Seq vs ChIP-exo under
+  Varying Sequencing Depths}
 \label{sec:reco}
@@
 
-\subsection{Some additional thoughts for this analysis}
-
-\subsubsection{GC content and mappability quality indicators}
+\subsection{Additional Thoughts for this Analysis}
+
+\subsubsection{GC Content and Mappability Quality Indicators}
 \label{sec:gc_map_qc}
@@
 
-\subsubsection{Quality control for ChIP-Nexus}
+\subsubsection{Quality Control for ChIP-Nexus}
 \label{sec:nexus}
@@
 
-\subsection{Analysis of E. Coli Transcription Initiation Complexes}
+\subsection{Analysis of \emph{E. Coli} Transcription Initiation Complexes}
 \label{sec:ecoli}
@@
 
-\subsection{Enhancer prediction and active learning}
+\subsection{Enhancer Prediction and Active Learning}
 \label{sec:enhancer}

### 2015-11-26 rmarkdown/prelims_paper.Rnw
fixed introduction and removed SSD from table 1

@@
 
-    Recently, ChIP-exo has been developed to investigate protein-DNA
-    interaction in higher resolution compared to popularly used
-    ChIP-Seq. Although ChIP-exo has drawn much attention and is
-    considered as powerful assay, currently no systematic studies
-    have yet been conducted to determine optimal strategies for
-    experimental design and analysis of ChIP-exo. In order to address
-    these questions, we evaluated diverse aspects of ChIP-exo and
-    found the following characteristics of ChIP-exo data. First, the
-    background of ChIP-exo data is quite different from that of
-    ChIP-Seq data. However, sequence biases inherently present in
-    ChIP-Seq data still exist in ChIP-exo data. Second, in ChIP-exo
-    data, reads are located around binding sites much more tightly and
-    hence, it has potential for high resolution identification of
-    protein-DNA interaction sites, and also the space to allocate the
-    reads is greatly reduced. Third, although often assumed in the
-    ChIP-exo data analysis methods, the ``peak pair'' assumption does
-    not hold well in real ChIP-exo data. Fourth, spatial resolution of
-    ChIP-exo is comparable to that of PET ChIP-Seq and both of them
-    are significantly better than resolution of SET ChIP-Seq. Finally,
-    for given fixed sequencing depth, ChIP-exo provides higher
-    sensitivity, specificity and spatial resolution than PET
-    ChIP-Seq.
-
-    We provide a quality control pipeline which visually assess
-    ChIP-exo biases, library complexity and enrichment; and calculates
-    a signal-to-noise measure. Also, we updated dPeak (Chung et al.,
-    2012 \cite{dpeak}), which makes a striking balance in sensitivity,
-    specificity and spatial resolution for ChIP-exo data analysis.
+% Recently, ChIP-exo has been developed to investigate protein-DNA
+% interaction in higher resolution compared to popularly used
+% ChIP-Seq. Although ChIP-exo has drawn much attention and is considered
+% as powerful assay, currently no systematic studies have yet been
+% conducted to determine which biases impact in a ChIP-exo's dataset
+% quality.
+
+% optimal strategies for
+%     experimental design and analysis of ChIP-exo.
+
+%     In order to address these questions, we evaluated diverse aspects
+%     of ChIP-exo and found the following characteristics of ChIP-exo
+%     data. First, the background of ChIP-exo data is quite different
+%     from that of ChIP-Seq data. However, sequence biases inherently
+%     present in ChIP-Seq data still exist in ChIP-exo data. Second, in
+%     ChIP-exo data, reads are located around binding sites much more
+%     tightly and hence, it has potential for high resolution
+%     identification of protein-DNA interaction sites, and also the
+%     space to allocate the reads is greatly reduced. Third, although
+%     often assumed in the ChIP-exo data analysis methods, the ``peak
+%     pair'' assumption does not hold well in real ChIP-exo
+%     data. Fourth, spatial resolution of ChIP-exo is comparable to that
+%     of PET ChIP-Seq and both of them are significantly better than
+%     resolution of SET ChIP-Seq. Finally, for given fixed sequencing
+%     depth, ChIP-exo provides higher sensitivity, specificity and
+%     spatial resolution than PET ChIP-Seq.
+
+%     We provide a quality control pipeline which visually assess
+%     ChIP-exo biases, library complexity and enrichment; and calculates
+%     a signal-to-noise measure. Also, we updated dPeak (Chung et al.,
+%     2012 \cite{dpeak}), which makes a striking balance in sensitivity,
+%     specificity and spatial resolution for ChIP-exo data analysis.
 
@@
 ChIP-exo (Chromatin Immunopecipitation followed by exonuclease
-digestion and next generation sequencing) Rhee and Pugh, 2011
-(\cite{exo1}) is the state-of-the-art experiment developed to attain
+digestion and next generation sequencing, Rhee and Pugh, 2011
+\cite{exo1}) is the state-of-the-art experiment developed to attain
 single base-pair resolution of protein binding site identification and
 it is considered as a powerful alternative to popularly used ChIP-Seq
-(Chromatin Immunoprecipitation coupled with next generation sequencing
- assay.
+(Chromatin Immunoprecipitation coupled with next generation
+sequencing) assay.
 
 While the number of produced ChIP-exo data keeps increasing,
-characteristics of ChIP-exo data and optimal strategies for
-experimental design and analysis of ChIP-exo data are not fully
-investigated yet, including issues of sequence biases inherent to
-ChIP-exo data, choice of optimal statistical methods, and
-determination of optimal sequencing depth. However, currently the
-number of available ChIP-exo data is still limited and their
-sequencing depths are still insufficient for such investigation. To
-address this limitation we gathered ChIP-exo data from diverse
-organisms: CTCF factor in human \cite{exo1}; ER factor in human and
-FoxA1 factor in mouse (Serandour et al., 2013 \cite{exoillumina}); and
-generated $\sig$ factor in Escherichia Coli (E. Coli) under
-aerobic ($ + O_2$) condition, and treated by rifampicin by 0 and 20
-minutes.
-
-DNA libraries generated by the ChIP-exo protocol seem to be less
-complex than the libraries generated by ChIP-Seq (Mahony et al., 2015
-\cite{exo_review}). Hence, most of current QC guidelines (Landt et
-al., 2012 \cite{encode_qc}) may not be applicable on ChIP-exo,
-additionally to our knowledge there are not established quality
-control pipelines for ChIP-exo. To address this challenge, we suggest
-a collection of quality control visualizations to understand which
-biases are present in ChIP-exo data. Previous ChIP-exo analysis used
-ChIP-Seq samples to compare the resolution between experiments
-(\cite{exo1}, \cite{exo2}, \cite{exoillumina}); Carroll et al., 2014
-\cite{carroll.qc} studied the use of the Strand-Cross Correlation
-(SCC) (Kharchenko et al., 2008 \cite{strandcc}) and showed that by
-filtering blacklisted regions the estimation of the SCC is
-improved. However, this method requires to know blacklisted regions in
-advance which may not be available, additionally using SCC may not be
-useful since the peaks that are attained at the read and fragment
-length are confused in a typical ChIP-exo SCC curve.  In our pipeline
-we propose two out-the-shelf analysis: an enrichment plot and the
-local normalized SCC cofficient.
-
-In order to archive the potential benefits of ChIP-exo on protein
+characteristics of ChIP-exo data are not fully investigated
+yet. First, DNA libraries generated by the ChIP-exo protocol seem to
+be less complex than the libraries generated by ChIP-Seq (Mahony et
+al., 2015 \cite{exo_review}), i.e. the number of positions to which
+the reads can be aligned has been reduced due to the exonuclease
+digestion. Second, although the are roughly the same amount of reads
+in both strands, locally there may be more reads in one strand than in
+the other. To address this challenges, we suggest a collection of
+quality control visualizations to understand which of this biases are
+present in ChIP-exo data and globally assess the enrichment and
+library complexity of a ChIP-exo sample.We gathered ChIP-exo data from
+diverse organisms: CTCF factor in human \cite{exo1}; ER factor in
+human and FoxA1 factor in mouse (Serandour et al., 2013
+\cite{exoillumina}); and generated $\sig$ factor in Escherichia Coli
+(E. Coli) under aerobic ($ + O_2$) condition, and treated by
+rifampicin by 0 and 20 minutes (courtesy of Professor Robert Landick's
+lab).
+% While the number of produced ChIP-exo data keeps increasing,
+% characteristics of ChIP-exo data and its analysis are not fully
+% investigated yet, including issues of sequence biases inherent to
+% ChIP-exo data, choice of optimal statistical methods, and
+% determination of optimal sequencing depth. However, currently the
+% number of available ChIP-exo data is still limited and their
+% sequencing depths are still insufficient for such investigation.
+
+Most of current ChIP-Seq QC guidelines (Landt et al., 2012
+\cite{encode_qc}) may not be applicable on ChIP-exo, additionally to
+our knowledge there are not established quality control pipelines for
+ChIP-exo.Previous ChIP-exo analysis used ChIP-Seq samples to compare
+the resolution between experiments (\cite{exo1}, \cite{exo2},
+\cite{exoillumina}); Carroll et al., 2014 \cite{carroll.qc} studied
+the use of the Strand-Cross Correlation (SCC) (Kharchenko et al., 2008
+\cite{strandcc}) and showed that by filtering blacklisted regions the
+estimation of the SCC is improved. However, this method requires to
+know blacklisted regions in advance which may not be available,
+additionally using SCC may not be useful since the peaks that are
+attained at the read and fragment length are confused in a typical
+ChIP-exo SCC curve.  In our pipeline we propose two out-the-shelf
+analysis: an enrichment plot and the normalized local SCC coefficient.
+
+In order to obtain the potential benefits of ChIP-exo on protein
 binding site identification, it is critical to understand which are
@@
 
-In this work, we demonstrate that the ``peak pair'' assumption of
-Rhee and Pugh \cite{exo1} does not hold well in real ChIP-exo
+In this work, we demonstrate that the ``peak pair'' assumption of Rhee
+and Pugh \cite{exo1} does not hold well in real ChIP-exo
 data. Furthermore, we found that when we analyze ChIP-exo data from
@@
 identification. We evaluated several method to identify binding events
-and dPeak outperforms or performs competitively respecto to GEM and
+and dPeak outperforms or performs competitively respect to GEM and
 MACE when analyzing ChIP-exo data. More importantly, when comparable
 number of reads is used for both ChIP-exo and ChIP-Seq, dPeak coupled
-with ChIP-exo data provides resolution comparable to PET ChIP-Seq and
+with ChIP-exo data provides outperform Paired End (PE) ChIP-Seq and
 both significantly improve the resolution of protein binding site
-identification compared to SET-based analysis with any of the
-available methods.
+identification compared to Single End (SE) ChIP-Seq.
 
@@
 to understand how this new protocol works, specifically we focused on
-understanding which biases are mantained from ChIP-Seq to ChIP-exo,
+understanding which biases are maintained from ChIP-Seq to ChIP-exo,
 which biases are specific to ChIP-exo, how the current ChIP-Seq QC
@@
   \caption{An example of a PET ChIP-Seq peak compared to
-    ChIP-exo. $\sig$, $\beta$ and $\beta_f'$ are subunits of
-    the transcription initiation complex in E. Coli. Gene regions on
-    the reverse strand are highlighted in light blue and the region
+    ChIP-exo. $\sig$, $\beta$ and $\beta_f'$ are subunits of the
+    transcription initiation complex in \emph{E. Coli}. Gene regions
+    on the reverse strand are highlighted in light blue and the region
     was centered respect to ChIP-Seq $\sig$ summit. The library
     complexity is lower in ChIP-exo than in ChIP-Seq but the
-    resolution is increased, this can be observed at the $\sig$
-    peak that for ChIP-exo was created as the combination of at least
-    three ChIP-exo ``sub-peaks''.}
+    resolution is increased, this can be observed at the $\sig$ peak
+    that for ChIP-exo was created as the combination of at least three
+    ChIP-exo ``sub-peaks''.}
   \label{fig:exo_example}
@@
 defined as the ratio between the number of positions to which exactly
-one mappping read maps divided by the number of position to which at
+one mapping read maps divided by the number of position to which at
 least one mapping reads maps, the ENCODE project proposed a series of
@@
 $\mbox{PBC} < 0.5$ should be reject due to be showing severe
-bottlenecking. Planet et al., 2011 proposed the Standarized Standard
+bottlenecking. Planet et al., 2011 proposed the Standardized Standard
 Deviation which defined as:
@@
 where $y(\delta)$ is the SCC for a strand shift $\delta$, $r$ is the
-Pearson correlation, $w_c$ is the propotion of tags mapped to
+Pearson correlation, $w_c$ is the proportion of tags mapped to
 chromosome $c$ and $n_c^S$ is the tag count vector for strand $S$ and
-chromsome $c$.
+chromosome $c$.
 
@@
 fragment length, ideally the SCC function is maximized at this
-point. Landt et al., 2012 \cite{encode_qc} explain that a succesful if
-this happens, it is marginally succesful if its maximized at the read
+point. Landt et al., 2012 \cite{encode_qc} explain that a successful if
+this happens, it is marginally successful if its maximized at the read
 length peak but still shows the fragment length summit and failed if
@@
 red) and a SET ChIP-Seq (in green) for CTCF in HeLa cell line. The
-ChIP-Seq dataset behaves as the succesful case from \cite{encode_qc},
+ChIP-Seq dataset behaves as the successful case from \cite{encode_qc},
 but the ChIP-exo doesn't. Additionally, since the exonuclease enzyme
 digests the DNA fragments, then the shifts where the SCC may reach a
-locall or a global maximum are confounded. Hence, new QC measures are
+local or a global maximum are confounded. Hence, new QC measures are
 necessary for ChIP-exo data.
@@
   \caption{Usual quality control indicators applied to the $\sigma^{70}$ samples.
-PBC stands PCR-bottleneck coefficient, SSD for standardized standard deviation, NSC for normalized strand cross-correlation coefficient (Another measure related to the strand cross-correlation curve is the RSC was ommited since input sample usually don't exists for ChIP-exo).
+PBC stands PCR-bottleneck coefficient, SSD for standardized standard deviation, NSC for normalized strand cross-correlation coefficient (Another measure related to the strand cross-correlation curve is the RSC was omitted since input sample usually don't exists for ChIP-exo).
 $\sig$ samples were given by Robert Landick, FoxA1 and ER samples are from Serandour et al., 2013 \cite{exoillumina} and CTCF sample is from Rhee and Pugh, 2011 \cite{exo1}.}
 
-\begin{tabular}{l|l|r|r|r|r|r|r}
+\begin{tabular}{l|l|r|r|r|r|r}
 \hline\hline
-IP & Organism  &Condition & Rep. & Nr. reads & PBC & SSD &  NSC  \\
+IP & Organism  &Condition & Rep. & Nr. reads & PBC &  NSC  \\
 \hline\hline
-$\sig$ & E.Coli & Rif-0min & 1 & 960,256 & 0.2823 & 0.0361 &  10.29 \\
+$\sig$ & E.Coli & Rif-0min & 1 & 960,256 & 0.2823 &   10.29 \\
 \hline
-$\sig$ &  E.Coli & Rif-0min & 2 & 2,247,295 & 0.2656 & 0.1091  &  25.08  \\
+$\sig$ &  E.Coli & Rif-0min & 2 & 2,247,295 & 0.2656 &  25.08  \\
 \hline
-$\sig$ &  E.Coli & Rif-20min & 1 & 1,940,387 & 0.2698 & 0.0820  & 17.69  \\
+$\sig$ &  E.Coli & Rif-20min & 1 & 1,940,387 & 0.2698 &  17.69  \\
 \hline
-$\sig$ &  E.Coli & Rif-20min & 2 & 4,229,574 & 0.2153 & 0.1647  &  14.11 \\
+$\sig$ &  E.Coli & Rif-20min & 2 & 4,229,574 & 0.2153 &   14.11 \\
 \hline
-FoxA1 &  Mouse &  NA & 1 & 22,210,461 & 0.6562 & 9.12 $\times 10^{-5}$ & 21.452 \\
+FoxA1 &  Mouse &  NA & 1 & 22,210,461 & 0.6562 & 21.452 \\
 \hline
-FoxA1 &  Mouse &  NA & 2 & 22,307,557 & 0.7996 & 7.94 $\times 10^{-5}$ & 60.661 \\
+FoxA1 &  Mouse &  NA & 2 & 22,307,557 & 0.7996 & 60.661 \\
 \hline
-FoxA1 &  Mouse &  NA & 3 & 22,421,729 & 0.1068 & 1.31 $\times 10^{-4}$ & 72.312 \\
+FoxA1 &  Mouse &  NA & 3 & 22,421,729 & 0.1068 & 72.312 \\
 \hline
-ER & Human & NA & 1 & 9,289,835 & 0.8082 & 3.64 $\times 10^{-5}$ & 19.843 \\
+ER & Human & NA & 1 & 9,289,835 & 0.8082 & 19.843 \\
 \hline
-ER & Human & NA & 2 & 11,041,833 & 0.8024 & 4.6 $\times 10^{-5}$ & 21.422 \\
+ER & Human & NA & 2 & 11,041,833 & 0.8024 & 21.422 \\
 \hline
-ER & Human & NA & 3 & 12,464,836 & 0.8203 & 4.89 $\times 10^{-5}$ & 19.699 \\
+ER & Human & NA & 3 & 12,464,836 & 0.8203 & 19.699 \\
 \hline
-CTCF & Human & NA & 1 &   48,478,450 & 0.4579 & 1.29 $\times 10^{-4}$ & 15.977 \\
+CTCF & Human & NA & 1 &   48,478,450 & 0.4579 & 15.977 \\
 \hline
@@
 
-Chung et al., 2013 proposed the dpeak model to identifity
+Chung et al., 2013 proposed the dPeak model to identify the
 transcription factor binding sites in high resolution. Two models are
@@
     example region for the 3 replicates. The library complexity seem
-    to be comparable between the two top rows and both outperfom the
+    to be comparable between the two top rows and both outperform the
     last row. The number of reads for the first replicate is higher
@@
 ChIP-Nexus. Depending if ChIP-Nexus data shows background we may need
-to modify the QC pipeline in the step where it paritions the reads
+to modify the QC pipeline in the step where it partitions the reads
 into regions, but the $\mbox{local-NSC}$ may be used to prove this
@@
 \begin{enumerate}
-\item Binding of RNAP to rpomoter DNA to form a closed complex.
+\item Binding of RNAP to promoter DNA to form a closed complex.
 \item Rearrangement of RNAP-DNA contacts to form an open complex.
-\item Binding of nucleoside triphosphates in the active site and
+\item Binding of nucleotide triphosphates in the active site and
   synthesis of a initial RNA to yield an initial transcription complex
@@
 \item Promoter escape to form a elongated complex (EC$_1$) when the
-  promoter contacts are broken, and the $\sig$ factor is realeased.
+  promoter contacts are broken, and the $\sig$ factor is released.
 \item Conversion of EC$_1$ into EC$_2$ by loading of the ribosome and
@@
 As a proof of concept we were able to identify open and closed regions
-by centering the binding the peaks respecto to the highest $\sig$
-summit (among the rif0 and rif20 samples), and then using hierachical
+by centering the binding the peaks respect to the highest $\sig$
+summit (among the rif0 and rif20 samples), and then using hierarchical
 clustering on the $\beta'_f$ signal of the rif0 samples to produce:
@@
 centered around the middle, for $\beta'_f$-rif20, the majority of he
-signal seems to be centered, hence the trascription was interrupted
+signal seems to be centered, hence the transcription was interrupted
 and for $\beta'_f$-rif0 we can see clusters formed by open regions.
@@
 different patterns that represent transcription units whose output was
-limited at differemt steps in the formation of EC$_2$.
+limited at different steps in the formation of EC$_2$.
 
 Additionally if we find several cases of those patterns, then we may
-be able to determine which sequences may be causing this behaviour at
+be able to determine which sequences may be causing this behavior at
 the different steps.
@@
 characterization group to fit a model to classify enhancers. Roughly
-speaks, the data set consided of a set of regions
+speaking, the dataset consisted of a set of regions
 $\mathcal{R}_1,\cdots,\mathcal{R}_n$ and a set of labels
@@
 into a training set and a test set, etc. However, there are in total
-another $m >> n$ regions in the experiment, and the cost of labelling
+another $m >> n$ regions in the experiment, and the cost of labeling
 those $m$ regions may be high (in either the actual monetary cost of
-the experiment or the time the scientists spends labelling those
+the experiment or the time the scientists spends labeling those
 regions). Those $n+m$ regions are H3K27ac peaks, hence they were
@@
 \item What is the best way to select the regions that are going to be
-  labelled, i.e. we would want to label the regions that gives us the
+  labeled, i.e. we would want to label the regions that gives us the
   best model.
@@
 may be the best to apply in this particular problem. Settles, 2012
-\cite{al} is a survey of acive learning techniques that we want to
+\cite{al} is a survey of active learning techniques that we want to
 explore in order to solve this problems.
@@
 \end{document}
+
+% LocalWords:  deconvolve

### 2015-11-27 rmarkdown/prelims_paper.Rnw
Second version of the paper finished

@@
 
-% Recently, ChIP-exo has been developed to investigate protein-DNA
-% interaction in higher resolution compared to popularly used
-% ChIP-Seq. Although ChIP-exo has drawn much attention and is considered
-% as powerful assay, currently no systematic studies have yet been
-% conducted to determine which biases impact in a ChIP-exo's dataset
-% quality.
+ChIP-exo is a modification of the ChIP-seq protocol for high
+resolution mapping of transcription factor binding sites. Although
+many aspect of the ChIP-exo data analysis are similar to those of
+ChIP-Seq, ChIP-exo presents a number of unique challenges. We present
+a quality control pipeline that analyzes ChIP-exo's strand imbalance
+and library complexity. Assessment of these biases and artifacts are
+facilitated through diagnostic plots and summary statistical that
+compare a large portion of the genome as partitioned into island with
+high depth regions.
+
+We systematically evaluated diverse aspects of ChIP-exo and found the
+following characteristics: First, ChIP-exo's background is quite
+different than ChIP-Seq's. Second, although often assumed in ChIP-exo
+data analysis methods, the ``peak pair'' assumption doesn't hold
+locally in real ChIP-exo data. Third, we compared Paired End (PE)
+ChIP-Seq with ChIP-exo and found that both protocols are comparable in
+resolution and sensitivity for closely located binding event, but as
+the distance between binding event increases ChIP-exo shows higher
+sensitivity than PE ChIP-Seq. Finally, for given fixed sequencing
+depth, ChIP-exo provides higher sensitivity, specificity and spatial
+resolution than PE ChIP-Seq.
 
@@
 %     experimental design and analysis of ChIP-exo.
+
+% We also systematically compare multiple ChIP-seq and ChIP-exo
+% datasets to quantify differences between these two protocols. Our
+% analysis indicate that spatial resolution of ChIP-exo is comparable to
+% that of paired-end ChIP-seq and both of them are significantly better
+% than resolution of single-end ChIP-seq. Furthermore, for a given fixed
+% sequencing depth, ChIP-exo provides higher sensitivity, specificity,
+% and spatial resolution than pairedend ChIP-seq.
 
@@
 \cite{exoillumina}); and generated $\sig$ factor in Escherichia Coli
-(E. Coli) under aerobic ($ + O_2$) condition, and treated by
+(\emph{E. Coli}) under aerobic ($ + O_2$) condition, and treated by
 rifampicin by 0 and 20 minutes (courtesy of Professor Robert Landick's
@@
 resolution using ChIP-Seq data. However, most of them are still not
-tailored for ChIP-exo and PET and SET ChIP-Seq data in a unified
+tailored for ChIP-exo and PE and SE ChIP-Seq data in a unified
 framework and as a result, currently available methods are not
@@
 al., 2013 \cite{dpeak}), a high resolution binding site identification
-(deconvolution) algorithm that we previously developed for PET and SET
+(deconvolution) algorithm that we previously developed for PE and SE
 ChIP-Seq data, so that it can also handle ChIP-exo data. The dPeak
@@
   DNA fragments starting from the 5' end until it finds a crosslink
-  protein (this figure is taken from Furey, 2012 \cite{chipbeyond}).}
+  protein (this Figure is taken from Furey, 2012 \cite{chipbeyond}).}
 \label{fig:exo}
@@
 
-ChIP-exo experiments first capture millions of DNA fragments
-(150 - 250 bp in length) that the protein under study interacts with
-using random fragmentation of DNA and a protein-specific
-antibody. Then, exonuclease is introduced to trim 5' end of each DNA
-fragment to a fixed distance from the bound protein. As a result,
-boundaries around the protein of interest constructed with 5' ends of
-fragments are located much closed to bound protein compared to
-ChIP-Seq. This step is unique to ChIP-exo and could potentially
-provide significantly higher spatial resolution compared to
-ChIP-Seq. Finally, high throughput sequencing of a small region (25 to
-100 bp) at 5' end of each fragment generates millions of reads or
-tags.
+ChIP-exo experiments first capture millions of DNA fragments (150 -
+250 bp in length) that the protein under study interacts with using
+random fragmentation of DNA and a protein-specific antibody. Then,
+exonuclease is introduced to trim 5' end of each DNA fragment to a
+fixed distance from the bound protein. As a result, boundaries around
+the protein of interest constructed with 5' ends of reads are located
+much closed to bound protein compared to ChIP-Seq. This step is unique
+to ChIP-exo and could potentially provide significantly higher spatial
+resolution compared to ChIP-Seq. Finally, high throughput sequencing
+of a small region (25 to 100 bp) at 5' end of each fragment generates
+millions of reads or tags.
 
@@
 guidelines behave in ChIP-exo and we developed a new QC pipeline
-specific to ChIP-exo. A typical example is shown in figure
+specific to ChIP-exo. A typical example is shown in Figure
 \ref{fig:exo_example}, which shows the $\sig$ and $\beta'_f$ factors
-(and the $\beta$ factor only on top) on a E. Coli genome's region:
-First, both regions show that the $\sig$ sample is enriched; second
-the $\sig$ sample in the ChIP-exo panel shows that its respective form
-in the ChIP-Seq panel may be form of more binding events; and third,
-either of the $\beta$ factors show that the complexity of the ChIP-exo
-libraries may be lower in some cases, i.e. the peak is mapped with
-fewer positions.
+(and the $\beta$ factor only on top) on a \emph{E. Coli} genome's
+region: First, both regions show that the $\sig$ sample is enriched;
+second the $\sig$ sample in the ChIP-exo panel shows that its
+respective form in the ChIP-Seq panel may be form of more binding
+events; and third, either of the $\beta$ factors show that the
+complexity of the ChIP-exo libraries may be lower in some cases,
+i.e. the peak is mapped with fewer positions. In \emph{E. Coli}
+multiple promoters may be located close to each other, hence we may
+suspect that the ChIP-Seq peak is occasioned by several promoter and
+in the ChIP-exo panel we can observe that such is the case.
 
@@
   \includegraphics[width = .8 \textwidth,page = 107]{/p/keles/ChIPexo/volume5/OpenClosedComplexes/figs/peaks/ChIPseq_ChIPexo_peaks_rif20.pdf}
-  \caption{An example of a PET ChIP-Seq peak compared to
+  \caption{An example of a PE ChIP-Seq peak compared to
     ChIP-exo. $\sig$, $\beta$ and $\beta_f'$ are subunits of the
@@
 
-To our knowledge it doesn't exists a QC pipeline for ChIP-exo. Hence,
-we analyzed how the current ChIP-Seq QC measures behave on ChIP-exo
-data.
+We start our exploration by investigating the suitability of the
+current state of the art QC pipeline for ChIP-Seq to ChIP-exo.
 
@@
 $\mbox{PBC} < 0.5$ should be reject due to be showing severe
-bottlenecking. Planet et al., 2011 proposed the Standardized Standard
-Deviation which defined as:
-
-\[
-\mbox{SSD} = \mbox{SD}_N / \sqrt{N} 
-\]
-
-where $\mbox{SD}_N$ is the standard deviation of the coverage of the
-sample and $N$ is the number of reads. This measure is smaller for
-ChIP-exo since the reads are located more tightly around binding
-sites, but it may be biased since several fragments are being
-digested by exonuclease enzyme, hence there are several position in
-the genome where no fragment is being mapped. Finally the normalized
-strand cross-correlation is defined as:
+bottlenecking. The normalized strand cross-correlation (NSC) is
+defined as:
 
@@
 where $y(\delta)$ is the SCC for a strand shift $\delta$, $r$ is the
-Pearson correlation, $w_c$ is the proportion of tags mapped to
-chromosome $c$ and $n_c^S$ is the tag count vector for strand $S$ and
+Pearson correlation, $w_c$ is the proportion of reads mapped to
+chromosome $c$ and $n_c^S$ is the read count vector for strand $S$ and
 chromosome $c$.
 
+\begin{figure}[H]
+  \centering
+  \includegraphics[width = \textwidth]{../figs/for_paper/SCC.png}
+  \caption{SCC interpretation for ChIP-Seq from \cite{encode_qc}. In
+    all panels, we observe shift $\delta$ vs SCC. The quality of
+    ChIP-Seq is determined by the relative heights of the ``phantom
+    peak'' and the local maximum when the shift is equal to the
+    fragment length.}
+  \label{fig:scc_landt}
+\end{figure}
+
 For ChIP-Seq data, the SCC will have two peaks: One when $\delta$ is
-equal to reads length and one when $\delta$ is equal to the unobserved
-fragment length, ideally the SCC function is maximized at this
-point. Landt et al., 2012 \cite{encode_qc} explain that a successful if
-this happens, it is marginally successful if its maximized at the read
-length peak but still shows the fragment length summit and failed if
-is maximized at the read length peak but the second summit is not
-present.
+equal to read length and one when $\delta$ is equal to the unobserved
+fragment length. The local maximum at the read length may be
+occasioned by the duplicated reads or unmappable regions
+\cite{carroll.qc}, or simply by reads in both strand that are mapped
+in the same position. Figure \ref{fig:scc_landt} shows how to
+interpret the SCC, the quality of a ChIP-Seq sample is going to be
+determined by the relative height of the ``phantom peak'' and the
+local maximum at the fragment length.
 
@@
 \includegraphics[width = .7\textwidth]{../figs/for_paper/scc_ctcf.pdf}
-  \caption{SCC for CTCF factor in HeLa cell line for ChIP-exo (obtained from \cite{exo1}) and SET ChIP-Seq (obtained from \cite{encode1}) }
+  \caption{SCC for CTCF factor in HeLa cell line for ChIP-exo (obtained from \cite{exo1}) and SE ChIP-Seq (obtained from \cite{encode1}) }
   \label{fig:scc}
@@
 Figure \ref{fig:scc} shows the SCC function for both a ChIP-exo (in
-red) and a SET ChIP-Seq (in green) for CTCF in HeLa cell line. The
+red) and a SE ChIP-Seq (in green) for CTCF in HeLa cell line. The
 ChIP-Seq dataset behaves as the successful case from \cite{encode_qc},
@@
   \caption{Usual quality control indicators applied to the $\sigma^{70}$ samples.
-PBC stands PCR-bottleneck coefficient, SSD for standardized standard deviation, NSC for normalized strand cross-correlation coefficient (Another measure related to the strand cross-correlation curve is the RSC was omitted since input sample usually don't exists for ChIP-exo).
-$\sig$ samples were given by Robert Landick, FoxA1 and ER samples are from Serandour et al., 2013 \cite{exoillumina} and CTCF sample is from Rhee and Pugh, 2011 \cite{exo1}.}
+PBC stands PCR-bottleneck coefficient (0-0.5 is severe bottlenecking, 0.5-0.8 is moderate bottlenecking, 0.8-0.9 is mild bottlenecking, while 0.9-1.0 is no bottlenecking), NSC for normalized strand cross-correlation coefficient (Another measure related to the strand cross-correlation curve is the RSC was omitted since input sample usually don't exists for ChIP-exo).
+$\sig$ samples were given by Professor Robert Landick's Lab, FoxA1 and ER samples are from Serandour et al., 2013 \cite{exoillumina} and CTCF sample is from Rhee and Pugh, 2011 \cite{exo1}.
+}
 
@@
 
-\vspace*{\fill}
-
 \subsection{The dPeak Model}
@@
 transcription factor binding sites in high resolution. Two models are
-proposed: One to deconvolve the signal of single end tags (SET) data,
+proposed: One to deconvolve the signal of single end reads (SE) data,
 i.e. when only one end of the DNA fragment is sequenced; and another
-to deconvolve the signal of paired end tags (PET) data, i.e. when both
-ends of the DNA fragment are sequenced. For ChIP-exo, both ends are
-sequenced but only the 5' end is considered, since is the side where
-the exonuclease enzyme digests the DNA.
-
-Given a peak region with $m$ positions, with $n$ DNA fragments and that
+to deconvolve the signal of paired end reads (PE) data, i.e. when both
+ends of the DNA fragment are sequenced. For ChIP-exo, only the 5' end
+is considered, since is the side where the exonuclease enzyme digests
+the DNA.
+
+\newpage
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width = .6\textwidth]{/p/keles/ChIPexo/volume3/ChIPexo/poster/dpeak.png}
+  \caption{A schematic illustration of dPeak's model when the read is
+    located around binding site $\mu_g$}
+\end{figure}
+
+Given a peak region with $m$ positions, with $n$ DNA reads and that
 the region is generated by a known $g^*$ amount of binding
-events. Each fragment is described by the start position where the
-starting end is aligned and the strand from which the fragment is
-sequenced. Since only one end is sequenced, then the fragment length
-is not observed. dPeak assumes that the strand $D_i$ follows a
-Bernoulli distribution with known parameter $p_D$, and the start
-position $R_i$ for the $i$-th fragment is generated according to the
-following procedure:
+events. Each read is described by the start position where the 5' end
+is aligned and the strand from which the fragment is sequenced. Since
+only one end is sequenced, then the fragment length is not
+observed. dPeak assumes that the strand $D_i$ follows a Bernoulli
+distribution with known parameter $p_D$, the binding event is not
+observed and it follows a
+$\mbox{Multinomial}(\pi_0,\pi_1,\cdots,\pi_{g^*})$. And the 5' end
+$R_i$ for the $i$-th fragment is generated according to the following
+procedure:
 
 \begin{itemize}
-\item If the fragment was sequenced from the forward strand ($D_i =
-  1$):
+\item If the read was sequenced from the forward strand ($D_i = 1$):
   \begin{itemize}
-  \item The reads belongs to the background: $R_i | Z_i = 0, D_i = 1
-    \sim \mbox{Unif}(1 - \beta + 1, m)$
+  \item The read belongs to the background: $R_i | Z_i = 0, D_i = 1
+    \sim \mbox{Unif}(1, m)$
   \item The read belong to the $g$-th binding event: $R_i | Z_i = g,
@@
   \end{itemize}
-\item If the fragment was sequenced from the backward strand ($D_i =
+\item If the read was sequenced from the backward strand ($D_i =
   0$):
   \begin{itemize}
-  \item The reads belongs to the background: $R_i | Z_i = 0, D_i = 1
-    \sim \mbox{Unif}(1 , m + \beta  - 1)$
+  \item The read belongs to the background: $R_i | Z_i = 0, D_i = 1
+    \sim \mbox{Unif}(1 , m)$
   \item The read belong to the $g$-th binding event: $R_i | Z_i = g,
@@
 
-and the binding event to which the $i$-th fragment belongs $Z_i = g$
-is generated from a
-
-$\mbox{Multinomial}(\pi_0,\pi_1,\cdots,\pi_{g^*})$.
-
 \subsubsection{A New Initialization Strategy for dPeak}
@@
 Chung et al., 2013 \cite{dpeak} derived an initialization strategy for
-the stochastic EM algorithm where each region is initialized by
-separate independently and the binding sites locations $\mu_g,
-g=1,\cdots,g^*$ are initialized by a uniform partition.
-
-For this work, we considered a new strategy where the all the regions
-are ranked by a statistical measure (as the $-\log(p)$ where $p$ is
-the empirical p.value of the test $H_0: \mathcal{R} \text{ is
-  enriched}$, where $\mathcal{R}$ is a genomic region) or a empirical
-quantity (as the number of ChIP tag counts in the region) and use as
-initial estimates of the $\delta$ and $\sigma$ parameters the average
-of the estimates from the highest ranked $N_{\text{top}}$ peaks.
+the stochastic EM algorithm where for each region, the binding sites
+locations $\mu_g, g=1,\cdots,g^*$ were initialized equally spaced
+around the region and the parameters $\delta$ and $\sigma$ were
+estimated by separate for all regions. This is problematic since this
+initialization strategy does not depend on the data, specially in
+regions with low signal to noise ration and with a low amount of
+reads.
+
+For this work, we considered a new strategy where the $mu_g$'s are
+initialized by partitioning the region into $g^*$ bins and finding the
+coverage's local maxima in each bin and we initialize $\delta$ and
+$\sigma$ by ranking the regions by the signal strength or an empirical
+quantity (as the number of ChIP tag counts in the region), fit the
+model for the top $N_{\text{top}}$ regions and then use as initial
+values the average of the $\sigma$ and $\delta$ estimates for those
+regions.
 
@@
 large proportion of regions without any read in ChIP-exo data. These
-results indicate that background distribution of ChIP-exo data is less
-homogeneous than that of ChIP-Seq data.
+results indicate that for ChIP-exo data a much smaller portion of the
+genome is expected to be background.
 
@@
   \includegraphics[width = .46\textwidth,page = 2]{../figs/for_paper/eukaryotic_bias_CTCF.pdf}
-  \caption{ A) ChIP signal in ChIP-exo was linearly related to that in
-    PET ChIP-Seq data in the region with high ChIP tag counts. In
-    contrast, there were clear differences in their background
-    distribution, where several ChIP-exo background regions were
-    empty. B) In ChIP-exo data, strands of reads were significantly
-    less balanced in the regions with potential binding sites compared
-    to SET ChIP-Seq data. C) ChIP tag counts increase linearly as
+  \caption{ A) Hexbin plot of PE ChIP-Seq bin counts VS ChIP-exo bin
+    counts. B) Forward Strand Ratio densities for SE ChIP-Seq and
+    ChIP-exo peaks. C) ChIP tag counts increase linearly as
     mappability scores increases. D) ChIP tag counts increase linearly
@@
 
-\begin{figure}[H]
+\begin{figure}[h!]
   \centering
   \includegraphics[width = .7\textwidth]{../figs/for_paper/coverage_diagram2.pdf}
-  \caption{Diagram of the ChIP-exo Quality control pipeline. The
-    genome is partitioned into several regions by removing the empty
-    segments and for each regions we calculate a collection of QC
-    indicators, then for each region $R_k$ a vector of summary
-    statistics $T(R_k)$ is calculated.}
+  \caption{Flowchart of the ChIP-exo Quality control pipeline.}
+  % The genome is partitioned into several regions by removing the empty
+  % segments and for each regions we calculate a collection of QC
+  % indicators, then for each region $R_k$ a vector of summary
+  % statistics $T(R_k)$ is calculated.
   \label{fig:qcdiagram}
@@
 We considered the creation of a novel QC measure for ChIP-exo called
-local normalized SCC or $\mbox{local-NSC}$. This measure
+local normalized SCC or $\mbox{local-NSC}$. This measure is built by
+calculating the SCC for a given region:
 
@@
 where $y(\delta)$ is the observed SCC for shift $x_\delta$,
-$\epsilon_\delta \sim N(0,\sigma_f^2)$. Finally the $\mbox{local-NSC}$
-is defined as a signal-to-noise ratio by using:
+$\epsilon_\delta \sim N(0,\sigma_f^2)$.
+
+Finally the $\mbox{local-NSC}$ is defined as:
 
 \begin{align}
-    \mbox{local-NSC} &= \frac{max_{x_\delta} \hat{f}(x_\delta)}{\hat{\sigma}_f} \label{locNSC}
+    \mbox{local-NSC} &= \frac{\mbox{max}_{x_\delta} \hat{f}(x_\delta)}{\hat{\sigma}_f} 
+\label{locNSC}
 \end{align}
 
-\begin{figure}[H]
-  \centering
-  \includegraphics[width = .8\textwidth]{../figs/for_paper/local_SCC_example.pdf}
+Local NSC gives a signal-to-noise ratio for a given region in a
+ChIP-exo sample.
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width = .9\textwidth]{../figs/for_paper/local_SCC_example.pdf}
   \caption{Using the FoxA1 sample from \cite{exoillumina}. A) shows an
@@
 
-Using the mouse-FoxA1 experiment from \cite{exoillumina} and the
-relationship between this two quantities, library complexity and
-sample enrichment was explored. Figure \ref{fig:enrich}A shows hexbin
-plots illustrating the interaction between these two
-quantities. Notice how there are two strong arms in each panel: The
-first one corresponds to regions with low $\mbox{ARC}$ values and
-varying $\mbox{URCR}$ values across the $(0,1]$ interval, while the
-second one shows a decreasing trend in $\mbox{URCR}$ as $\mbox{ARC}$
-increases. When an experiment shows a higher degree of enrichment,
-then the separation of this two arms is more noticeable, since the
-second arm corresponds to possibly enriched regions
-(\ref{fig:enrich2}A and \ref{fig:enrich2}B).
+Figure \ref{fig:enrich}A shows hexbin plots illustrating the
+relationship between library complexity and sample enrichment for the
+FoxA1 in mouse cells from \cite{exoillumina}. Notice how there are two
+strong arms in each panel: The first one corresponds to regions with
+low $\mbox{ARC}$ values and varying $\mbox{URCR}$ values across the
+$(0,1]$ interval, while the second one shows a decreasing trend in
+$\mbox{URCR}$ as $\mbox{ARC}$ increases. When an experiment shows a
+higher degree of enrichment, then the separation of this two arms is
+more noticeable, since as Figures \ref{fig:enrich2}A and
+\ref{fig:enrich2}B show the ARC VS URCR plots but filtering out the
+regions formed by a low number of unique positions (10 and 30
+respectively), hence this second arm corresponds to possibly enriched
+regions.
 
@@
 regions of the three replicates, stratified by the number of reads
-mapped exactly to one position for all $\sigma^{70}$ ChIP-exo
-experiments; The high stratum is defined by regions consisting of more
-than \Sexpr{high} unique positions, the medium stratum for regions
-where the number of unique position is on the
-(\Sexpr{med},\Sexpr{high}) range and the low stratum in the
-(\Sexpr{low},\Sexpr{med}) range. For each stratum and biological
-replicate, we sampled \Sexpr{strata} regions whenever possible (in the
-opposite case, all the regions in the category were considered) and
-calculated the $\mbox{local-NSC}$ for those regions. For all three
-replicates, it is shown an increasing trend as the number of unique
-position increases, which means that the $\mbox{local-NSC}$ is an
-effective indicator of library complexity. Figure \ref{fig:enrich}C
-shows the $\mbox{local-NSC}$ coefficient calculated for the same set
-of regions, stratified by the number of unique positions in the
-regions of the first replicate using the same criteria as before. In
-the high panel it is shown that the first replicate sample performs
-slightly better than the second and both outperform the third one,
-this agrees with \ref{fig:enrich}A, where only in the left-most panel
-doesn't seem to be a separation between the low $\mbox{ARC}$ arm and
-the other.
+mapped exactly to one position for all Fox-A1 ChIP-exo experiments;
+The high stratum is defined by regions consisting of more than
+\Sexpr{high} unique positions, the medium stratum for regions where
+the number of unique position is on the (\Sexpr{med},\Sexpr{high})
+range and the low stratum in the (\Sexpr{low},\Sexpr{med}) range. For
+each stratum and biological replicate, we sampled \Sexpr{strata}
+regions whenever possible (in the opposite case, all the regions in
+the category were considered) and calculated the $\mbox{local-NSC}$
+for those regions. On one side, we can see that for all three
+replicates, as the number of unique positions increases, then the
+$\mbox{local-NSC}$ shows higher values. In the low panel, all three
+replicates are comparable, in the medium panel, the first two
+replicates are comparable and both outperform the third one, and
+finally in the high panel we can see that despite that the first two
+replicates are comparable, the first one slightly outperforms the
+second. We can see that all three panels coincide with the conclusion
+we get from Figure \ref{fig:enrich}A: First, the first replicate seems
+to have the highest library complexity and the third replicate the
+lowest. Second, despite that both first two replicates show a similar
+level of bottlenecking, the regions in the first replicate seems to
+have a higher ARC, those regions seem to be contributing to have a
+higher local-NSC in the high panel. Third, despite that the ARC of the
+first and third replicates is comparable, the third replicate has a
+higher bottlenecking level which cause to $\mbox{local-NSC}$ to have
+lower levels overall.
 
@@
   \includegraphics[width = .7\textwidth]{../figs/Carroll_mice_for_paper/FoxA1_enrichment.pdf}
-  \includegraphics[width = .7\textwidth,page =3]{../figs/Carroll_mice_for_paper/Local_SCC_indicator_by_strata.pdf}
+  \includegraphics[width = .7\textwidth,page =4]{../figs/Carroll_mice_for_paper/Local_SCC_indicator_by_strata.pdf}
 %  \includegraphics[width = .7\textwidth,page =4]{../figs/Carroll_mice_for_paper/Local_SCC_indicator_by_strata_sameRegion.pdf}
@@
 The strand imbalance assessment is based in the observation that the
-enriched regions usually have a higher concentration of fragments,
+enriched regions usually have a higher concentration of reads,
 therefore we examined the FSR (defined as the ratio of the number of
@@
 indicator is of particular importance, since several methods rely on
-the ``peak-pair'' assumption. In table \ref{tab:qcbase}, we calculated
+the ``peak-pair'' assumption. In Table \ref{tab:qcbase}, we calculated
 the FSR and noticed that for all the samples, it's value is close to
 0.5, which means that there are roughly the same amount of reads in
-both strands. However, figure \ref{fig:comp}B shows that this value is
-not representative of the sample locally, therefore the assumption
-doesn't hold in practice.
+both strands. However, Figure \ref{fig:comp}B shows that this value
+doesn't represent the sample locally, hence the assumption doesn't
+hold well in practice.
 
@@
 In order to assess the strand imbalance we created the following
-visualization presented in figure \ref{fig:strand}. A) shows the FSR's
+visualization presented in Figure \ref{fig:strand}. A) shows the FSR's
 behavior as lower depth regions are being filtered out, while B) shows
-which percentage of the regions are composed by fragments of both
-strands or only one (forward or backward). As the number of reads per
-region's lower bound increases, the quantiles tend to approach the
-median. To evaluate this statement, for each replicate of the FoxA1
-dataset, we divided the regions formed by at least \Sexpr{de}
-fragments into into the ones that overlap with any of the sample's
-peaks. Then we considered a Kolmogorov-Smirnov tests to assess if the
-distribution of both groups were the same. The p.value for the third
-replicate's test is \Sexpr{pval[[1]]}, for the first replicate's test
-is \Sexpr{pval[[2]]} and for the second test is
-\Sexpr{pval[[3]]}. This shows that only the middle sample from figure
+which percentage of the regions are composed by reads of both strands
+or only one (forward or backward). As the number of reads per region's
+lower bound increases, the quantiles tend to approach the median. To
+evaluate this statement, for each replicate of FoxA1, we divided the
+regions formed by at least \Sexpr{de} reads into the ones that overlap
+with any of the sample's peaks. Then we considered Kolmogorov-Smirnov
+tests to assess if the FSR's distribution was the same for the islands
+that overlap with a peak and the islands that didn't. The p.value for
+the third replicate's test is \Sexpr{pval[[1]]}, for the first
+replicate's test is \Sexpr{pval[[2]]} and for the second test is
+\Sexpr{pval[[3]]}. This shows that only the middle sample from Figure
 \ref{fig:strand} is the least imbalance sample from the the three,
@@
   \includegraphics[width = .8\textwidth,page = 3]{../figs/Carroll_mice_for_paper/Strand_imbalance.pdf} 
-  \caption{Strand imbalance QC plots for the same data as in figure
+  \caption{Strand imbalance QC plots for the same data as in Figure
     \ref{fig:enrich}. A) FSR distribution quantiles as the lower depth
@@
 Figure \ref{fig:reso_all} shows different comparisons among ChIP-exo,
-PET ChIP-Seq and SET ChIP-Seq. A RegulonDB annotation was considered
+PE ChIP-Seq and SE ChIP-Seq. A RegulonDB annotation was considered
 identified if the distance between it and dPeak binding site estimate
 was at most of \Sexpr{ext} bp. That way, the sensitivity is defined as
-the proportion of RegulonDB annotations identified in a peak. In
-figure \ref{fig:reso_all}A can be seen that the sensitivity of all
-protocols increases as the average distance between binding sites
-does. Despite that when the binding events in a peak are closer to
-each other, both ChIP-exo and PET ChIP-Seq are comparable, as the
-distance increases ChIP-exo identifies a higher proportion of the
-RegulonDB annotations; additionally SET ChIP-Seq is significantly less
-sensible than both ChIP-exo and PET ChIP-Seq. In figure
-\ref{fig:reso_all}B the distance between a RegulonDB annotation to its
-closest prediction is compared for ChIP-exo, PET and SET ChIP-Seq;
-while the first two are comparable, both outperform SET ChIP-Seq.
-
-In figures \ref{fig:reso_all}C and \ref{fig:reso_all}D, we observe the
+the proportion of RegulonDB annotations identified in a peak (Salgado
+et al., 2012 \cite{regulonDB}). Figure \ref{fig:reso_all}A illustrates
+that the sensitivity of all protocols increases as the average
+distance between binding sites does. Despite that when the binding
+events in a peak are closer to each other, both ChIP-exo and PE
+ChIP-Seq are comparable, as the distance increases ChIP-exo identifies
+a higher proportion of the RegulonDB annotations; additionally SE
+ChIP-Seq is significantly less sensitive than both ChIP-exo and PE
+ChIP-Seq. In Figure \ref{fig:reso_all}B the distance between a
+RegulonDB annotation to its closest prediction is compared for
+ChIP-exo, PE and SE ChIP-Seq; while the first two are comparable,
+both outperform SE ChIP-Seq.
+
+In Figures \ref{fig:reso_all}C and \ref{fig:reso_all}D, we observe the
 behavior of dPeak's estimated parameters for single end data (both
 ChIP-exo and ChIP-Seq). The $\delta$ parameter measures the distance
-from the 5' end of the fragment to its respective binding event and
-the $\sigma$ parameter indicates the fragments distribution
-variability around their respective binding sites (more information in
-\cite{dpeak}). For both parameters, it can be seen that the ChIP-exo
-estimates are smaller than the SET ChIP-Seq estimates in average,
-which agrees with the fact that the sample reads are allocated more
-tightly around the binding events in ChIP-exo data. Hence we can
-conclude that the peaks shape is quite different between ChIP-exo and
-SET ChIP-Seq.
-
+from the 5' end of the read's 5' end to its respective binding event
+and the $\sigma$ parameter indicates the read's starting position
+distribution variability around their respective binding sites (more
+information in \cite{dpeak}). For both parameters, it can be seen that
+the ChIP-exo estimates are smaller than the SE ChIP-Seq estimates in
+average, which agrees with the fact that the sample reads are
+allocated more tightly around the binding events in ChIP-exo
+data. This confirms that the dPeak model is able to recover the same
+conclusions that have been done when comparing ChIP-exo and SE
+ChIP-Seq.
+
+% Hence we can conclude that the peaks shape is quite different
+% between ChIP-exo and SE ChIP-Seq.
 \begin{figure}[H]
@@
      binding site. In ChIP-exo data, reads were located much closer to
-     the binding site than in SET ChIP-Seq. D) $\sigma$ parameter
+     the binding site than in SE ChIP-Seq. D) $\sigma$ parameter
      measure the dispersion of reads around each binding site. In
      ChIP-exo data, reads showed less variation around the their
-     respective binding sites compared to SET ChIP-Seq.}
+     respective binding sites compared to SE ChIP-Seq.}
   \label{fig:reso_all}
@@
 
-We sampled a fixed amount of fragments for each of the ChIP-exo, PET
-ChIP-Seq and SET ChIP-Seq datasets of the $\sigma^{70}$ sample in
-aerobic conditions. For each sampled dataset we applied our
+Previously, ChIP-exo and SE ChIP-Seq have been compared at a fixed
+depth level in the literature, but this comparisons didn't included PE
+ChIP-Seq. Hence, we sampled a fixed amount of reads for each of the
+ChIP-exo, PE ChIP-Seq and SE ChIP-Seq datasets of the $\sigma^{70}$
+sample in aerobic conditions. For each sampled dataset we applied our
 lower-to-higher resolution pipeline by calling peaks with MOSAiCS
@@
   \includegraphics[width = .8\textwidth]{../figs/for_paper/saturation_analysis_old.pdf}
-  \caption{Comparison of the number of candidate regions (A),
-    predicted events (B), identified targets (C) and resolution (D)
-    among ChIP-exo, PET ChIP-Seq and SET ChIP-Seq. RegulonDB
-    annotations are considered as a gold standard. A gold standard was
-    considered identified if a BS was estimated inside a \Sexpr{ext}
-    bp window around it}
+  \caption{\textbf{ChIP-exo, PE ChIP-Seq and SE ChIP-Seq comparison at
+      varying sequencing depths}Comparison of the number of candidate
+    regions (A), predicted events (B), identified targets (C) and
+    resolution (D) among ChIP-exo, PE ChIP-Seq and SE
+    ChIP-Seq. RegulonDB annotations are considered as a gold
+    standard. A gold standard binding events was deemed identified if
+    a binding event was estimated at a $\pm$ \Sexpr{ext} vicinity of
+    it.}
   \label{fig:design}
@@
 candidate peaks or the number of predicted events is lower for
-ChIP-exo, it outperforms both PET and SET ChIP-Seq in the number of
+ChIP-exo, it outperforms both PE and SE ChIP-Seq in the number of
 identified targets and resolution. This may suggest that with ChIP-exo
@@
 higher number of targets, but it may also be possible that this is an
-artifact occurring due to ChIP-exo's lower library complexity.
+artifact occurring due to ChIP-exo's lower library
+complexity. Additionally, it is worth noting that for PE ChIP we
+sampled both ends of the fragment, hence for each sequencing depth
+considered we are sampling twice as many reads for PE ChIP-Seq than
+for ChIP-exo or SE ChIP-seq.
 
@@
 \item \textbf{Segvis}: The goal of this package is to visualize
-  genomic regions by using aligned reads such as in figures
+  genomic regions by using aligned reads such as in Figures
   \ref{fig:exo} or \ref{fig:qcdiagram}.  The latest version is
@@
 \item \textbf{ChIPUtils}: This package attempts to gather the most
-  commonly used ChIP-Seq QC (as in table \ref{tab:qcbase} or figure
-  \ref{fig:scc}) indicators and additional utilities (as figure
+  commonly used ChIP-Seq QC (as in Table \ref{tab:qcbase} or Figure
+  \ref{fig:scc}) indicators and additional utilities (as Figure
   \ref{fig:comp}A) for ChIP-Seq. The latest available version is in
@@
 
-We provide a ChIP-exo QC pipeline capable of assess the balance between
-enriched samples and low complexity regions. It is shown that the
-``peak-pair'' assumption doesn't hold in practice and we provide two
-out-of-the-box visualization capable to assess the strand imbalance in
-a ChIP-exo experiment.
-
-We updated the dPeak algorithm from \cite{dpeak} and showed that
-ChIP-exo is comparable in resolution and sensitivity with PET ChIP-Seq
-and outperforms SET ChIP-Seq. ChIP-exo compared with PET and SET
-ChIP-Seq at fixed depth sample is able to identify more targets at a
-lower resolution. We compared dPeak with another algorithms to
-estimate binding locations in ChIPexo data, dPeak is comparable to
-MACE and outperforms GEM in resolution. dPeak provided a striking
-balance in sensitivity, specificity and spatial resolution for
-ChIP-exo analysis.
+We provide a ChIP-exo QC pipeline capable of assess the balance
+between enriched samples and low complexity regions. It is shown that
+the ``peak-pair'' assumption doesn't hold locally in practice and we
+provide two out-of-the-box visualization capable to assess the strand
+imbalance in a ChIP-exo experiment.
+
+We showed that ChIP-exo is comparable in resolution with PE ChIP-Seq
+and both outperforms SE ChIP-Seq. ChIP-exo shows higher sensitivity
+than ChIP-Seq (both PE and SET), and this effect increases when the
+average distance between binding events is larger.
+
+We demonstrated for the first time that when ChIP-exo is compared with
+PE ChIP-Seq (and SE ChIP-Seq) at a fixed sequencing depth, ChIP-exo is
+able to identify more targets at a lower resolution. dPeak provides a
+striking balance in sensitivity, specificity and spatial resolution
+for ChIP-exo analysis.
+
+% We compared dPeak with another
+% algorithms to estimate binding locations in ChIP-exo data, dPeak is
+% comparable to MACE and outperforms GEM in resolution. dPeak provided a
+% striking balance in sensitivity, specificity and spatial resolution
+% for ChIP-exo analysis.
 
@@
 \ref{sec:dpeak_analysis} we called peaks by using MOSAiCS which
-considers them as necessary information when there is not input
-sample. Hence, quality measures we may consider to add indicators to
-the QC pipeline that considers this biases.
+considers them as necessary information when there is not control (the
+GOF plot is in Figure \ref{fig:mosaics_exo}) sample. Hence, we may add
+QC measures that consider the GC content and mappability score biases
+into the QC pipeline.
 
@@
 
+% For example, the recently described ChIP-nexus protocol ligates both
+% sequencing adaptors onto one end of ChIP fragments in a single step
+% (as opposed to two separate ligation steps in the original ChIP-exo
+% protocol) (He et al., 2015). Exonuclease digestion, DNA
+% self-circularization with circLigase, and restriction enzyme cutting
+% between the two adaptors creates the final library. By removing one
+% of the relatively inefficient ligation steps, ChIP-nexus yields
+% higher complexity libraries with the same resolution as
+% ChIP-exo. One concern is that circLigase activity might have
+% sequence specificity (Kwok et al., 2013), potentially creating bias
+% in the identification of binding locations.
+
 He et al., 2014 \cite{chipnexus} proposed a modification to the
 ChIP-exo protocol that may fix its low library complexity issues by
-extending the fragments by a random barcode prior to the exonuclease
-digestion. This last statement is briefly mentioned by Mahony and
-Pugh, 2015 \cite{exo_review}. We want to apply our QC pipeline to
-ChIP-Nexus. Depending if ChIP-Nexus data shows background we may need
-to modify the QC pipeline in the step where it partitions the reads
-into regions, but the $\mbox{local-NSC}$ may be used to prove this
+extending the reads by a random barcode prior to the exonuclease
+digestion and ligating both sequencing adaptors onto one end of the
+ChIP DNA fragment (as opposed to two separate ligation steps in
+ChIP-exo). We want to apply our QC pipeline to ChIP-Nexus. Depending
+on the depth of the regions formed by ChIP-Nexus data, we may need to
+modify the QC pipeline in the step where it partitions the reads into
+regions, but the $\mbox{local-NSC}$ may be used to prove this
 statement.
@@
 
-The data showed in figure \ref{fig:exo_example} is actually part of a
-more interesting experiment. We have 2 biological replicates of the
-$\sig$, $\beta'_f$ and $\beta$ transcription factors under two
-conditions, one where rifampicin was applied by 20 minutes and another
-where it wasn't applied. As seen in the figure, there is ChIP-exo data
-for the 3 factors and PET ChIP-Seq for $\sig$ and $\beta'_f$ (in the
-figure both replicates were pooled).
+The data showed in Figure \ref{fig:exo_example} is actually part of a
+more interesting experiment from Professor Landick's lab. We have 2
+biological replicates of the $\sig$, $\beta'_f$ and $\beta$
+transcription factors under two conditions, one where rifampicin was
+applied by 20 minutes and another where it wasn't applied. As seen in
+the Figure, there is ChIP-exo data for the 3 factors and PE ChIP-Seq
+for $\sig$ and $\beta'_f$ (in the Figure both replicates were pooled).
 
 $\sig$ factor is a transcription initiation factor of housekeeping
-genes in E. Coli, and both $\beta$ factors transcribe the DNA into
+genes in \emph{E. Coli}, and both $\beta$ factors transcribe the DNA into
 mRNA. Hence, we want to use this data to have a better understanding
-of the transcription in E. Coli.
+of the transcription in \emph{E. Coli}.
 
@@
  \includegraphics[width = \textwidth]{/p/keles/ChIPexo/volume5/OpenClosedComplexes/figs/filtered/heatmaps_centroid.png}
- \caption{Centered and normalized heatmaps for PET ChIP-Seq peaks of
+ \caption{Centered and normalized heatmaps for PE ChIP-Seq peaks of
    $\sig$ and $\beta'_f$ under rif0 and rif20. The regions are
@@
 
-Figure \ref{fig:hm1} shows heatmaps of the normalized coverages of PET
+Figure \ref{fig:hm1} shows heatmaps of the normalized coverages of PE
 ChIP-Seq peaks for $\beta'_f$ and $\sig70$ under rif0 and rif20
@@
 different samples. As expected the signal for both $\sig$ samples is
-centered around the middle, for $\beta'_f$-rif20, the majority of he
+centered around the middle, for $\beta'_f$-rif20, the majority of the
 signal seems to be centered, hence the transcription was interrupted
@@
 
-Then, using a set of high quality PET ChIP-Seq peaks and clustering
+Then, using a set of high quality PE ChIP-Seq peaks and clustering
 with the mclust model by Fraley and Raftery, 2007 \cite{mclust} we
@@
 $Y_1,\cdots,Y_n \in \{0,1,\}$. Hence, we built a series of features
-$X_1,\cdots,X_p$ and fitted a model by dividing the $n$ observations
-into a training set and a test set, etc. However, there are in total
-another $m >> n$ regions in the experiment, and the cost of labeling
-those $m$ regions may be high (in either the actual monetary cost of
-the experiment or the time the scientists spends labeling those
-regions). Those $n+m$ regions are H3K27ac peaks, hence they were
+$X_1,\cdots,X_p$ and fitted a model. The model was trained by using
+10-fold CV and we selected the tunned parameters that maximized the
+area under the ROC curve. Then, an independent test set was provided
+for which we delivered a set of predictions. However, there are in
+total another $m >> n$ regions in the experiment, and the cost of
+labeling those $m$ regions may be high (in either the actual monetary
+cost of the experiment or the time the scientists spends labeling
+those regions). Those $n+m$ regions are H3K27ac peaks, hence they were
 ranked and the first $n$ were selected randomly from groups with high,
@@
 \begin{itemize}
-\item Is there a better way of ranking this $n+m$ regions ?
+\item Is there a better way of ranking these $n+m$ regions ?
 \item What is the best way to select the regions that are going to be
-  labeled, i.e. we would want to label the regions that gives us the
-  best model.
+  labeled, i.e. we would want to label the region that gives the
+  best model fit.
 \end{itemize}
@@
 \cite{al} is a survey of active learning techniques that we want to
-explore in order to solve this problems.
+explore in order to solve these problems.
 
@@
 \bibliography{chip_exo_paper}
+\nocite{exo_review}
 \nocite{exo_gb}
@@
 
-\section*{Supplement}
+\newpage
+
+\section*{Supplementary materials}
 \label{sec:supp}
 
-\subsection*{Additional Enrichment plots for $\sigma^{70}$}
+\subsection*{Additional Enrichment Plots for FoxA1}
 \label{sec:enrichsup}
 
-In figure \ref{fig:enrich2} it is shown the same relationship as in
-figure \ref{fig:enrich}, but considering only regions formed where the
+In Figure \ref{fig:enrich2} it is shown the same relationship as in
+Figure \ref{fig:enrich}, but considering only regions formed where the
 reads are being allocated in more than 10 (A) and 30 (B) unique
@@
 
+\subsection*{MOSAiCS GOF Plots for $\sig$ ChIP-exo}
+
+Figure \ref{fig:mosaics_exo} shows the GOF plot when the MOSAiCS model
+is used to call peaks on $\sig$ ChIP-exo data under aerobic condition.
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width = .5\textwidth]{../figs/sig70/gof_plots_sig70.pdf}
+  \caption{GOF plot for MOSAiCS model when applied to $\sig$ data under aerobic conditions}
+  \label{fig:mosaics_exo}
+\end{figure}
+
 \end{document}

### 2015-11-27 rmarkdown/prelims_paper.tex
Second version of the paper finished

@@
 regions of the three replicates, stratified by the number of reads
-mapped exactly to one position for all $\sigma^{70}$ ChIP-exo
-experiments; The high stratum is defined by regions consisting of more
-than 100 unique positions, the medium stratum for regions
-where the number of unique position is on the
-(50,100) range and the low stratum in the
-(20,50) range. For each stratum and biological
-replicate, we sampled 400 regions whenever possible (in the
-opposite case, all the regions in the category were considered) and
-calculated the $\mbox{local-NSC}$ for those regions. For all three
-replicates, it is shown an increasing trend as the number of unique
-position increases, which means that the $\mbox{local-NSC}$ is an
-effective indicator of library complexity. Figure \ref{fig:enrich}C
-shows the $\mbox{local-NSC}$ coefficient calculated for the same set
-of regions, stratified by the number of unique positions in the
-regions of the first replicate using the same criteria as before. In
-the high panel it is shown that the first replicate sample performs
-slightly better than the second and both outperform the third one,
-this agrees with \ref{fig:enrich}A, where only in the left-most panel
-doesn't seem to be a separation between the low $\mbox{ARC}$ arm and
-the other.
+mapped exactly to one position for all Fox-A1 ChIP-exo experiments;
+The high stratum is defined by regions consisting of more than
+100 unique positions, the medium stratum for regions where
+the number of unique position is on the (50,100)
+range and the low stratum in the (20,50) range. For
+each stratum and biological replicate, we sampled 400
+regions whenever possible (in the opposite case, all the regions in
+the category were considered) and calculated the $\mbox{local-NSC}$
+for those regions. On one side, we can see that for all three
+replicates, as the number of unique positions increases, then the
+$\mbox{local-NSC}$ shows higher values. In the low panel, all three
+replicates are comparable, in the medium panel, the first two
+replicates are comparable and both outperform the third one, and
+finally in the high panel we can see that despite that the first two
+replicates are comparable, the first one slightly outperforms the
+second. We can see that all three panels coincide with the conclusion
+we get from Figure \ref{fig:enrich}A: First, the first replicate seems
+to have the highest library complexity and the third replicate the
+lowest. Second, despite that both first two replicates show a similar
+level of bottlenecking, the regions in the first replicate seems to
+have a higher ARC, those regions seem to be contributing to have a
+higher local-NSC in the high panel. Third, despite that the ARC of the
+first and third replicates is comparable, the third replicate has a
+higher bottlenecking level which cause to $\mbox{local-NSC}$ to have
+lower levels overall.
 
@@
 indicator is of particular importance, since several methods rely on
-the ``peak-pair'' assumption. In table \ref{tab:qcbase}, we calculated
+the ``peak-pair'' assumption. In Table \ref{tab:qcbase}, we calculated
 the FSR and noticed that for all the samples, it's value is close to
 0.5, which means that there are roughly the same amount of reads in
-both strands. However, figure \ref{fig:comp}B shows that this value is
-not representative of the sample locally, therefore the assumption
-doesn't hold in practice.
+both strands. However, Figure \ref{fig:comp}B shows that this value
+doesn't represent the sample locally, hence the assumption doesn't
+hold well in practice.
 
 In order to assess the strand imbalance we created the following
-visualization presented in figure \ref{fig:strand}. A) shows the FSR's
+visualization presented in Figure \ref{fig:strand}. A) shows the FSR's
 behavior as lower depth regions are being filtered out, while B) shows
-which percentage of the regions are composed by fragments of both
-strands or only one (forward or backward). As the number of reads per
-region's lower bound increases, the quantiles tend to approach the
-median. To evaluate this statement, for each replicate of the FoxA1
-dataset, we divided the regions formed by at least 100
-fragments into into the ones that overlap with any of the sample's
-peaks. Then we considered a Kolmogorov-Smirnov tests to assess if the
-distribution of both groups were the same. The p.value for the third
-replicate's test is 0.1039, for the first replicate's test
-is 0.0064 and for the second test is
-0.0871. This shows that only the middle sample from figure
+which percentage of the regions are composed by reads of both strands
+or only one (forward or backward). As the number of reads per region's
+lower bound increases, the quantiles tend to approach the median. To
+evaluate this statement, for each replicate of FoxA1, we divided the
+regions formed by at least 100 reads into the ones that overlap
+with any of the sample's peaks. Then we considered Kolmogorov-Smirnov
+tests to assess if the FSR's distribution was the same for the islands
+that overlap with a peak and the islands that didn't. The p.value for
+the third replicate's test is 0.1039, for the first
+replicate's test is 0.0064 and for the second test is
+0.0871. This shows that only the middle sample from Figure
 \ref{fig:strand} is the least imbalance sample from the the three,
@@
 Figure \ref{fig:reso_all} shows different comparisons among ChIP-exo,
-PET ChIP-Seq and SET ChIP-Seq. A RegulonDB annotation was considered
+PE ChIP-Seq and SE ChIP-Seq. A RegulonDB annotation was considered
 identified if the distance between it and dPeak binding site estimate
 was at most of 20 bp. That way, the sensitivity is defined as
-the proportion of RegulonDB annotations identified in a peak. In
-figure \ref{fig:reso_all}A can be seen that the sensitivity of all
-protocols increases as the average distance between binding sites
-does. Despite that when the binding events in a peak are closer to
-each other, both ChIP-exo and PET ChIP-Seq are comparable, as the
-distance increases ChIP-exo identifies a higher proportion of the
-RegulonDB annotations; additionally SET ChIP-Seq is significantly less
-sensible than both ChIP-exo and PET ChIP-Seq. In figure
-\ref{fig:reso_all}B the distance between a RegulonDB annotation to its
-closest prediction is compared for ChIP-exo, PET and SET ChIP-Seq;
-while the first two are comparable, both outperform SET ChIP-Seq.
-
-In figures \ref{fig:reso_all}C and \ref{fig:reso_all}D, we observe the
+the proportion of RegulonDB annotations identified in a peak (Salgado
+et al., 2012 \cite{regulonDB}). Figure \ref{fig:reso_all}A illustrates
+that the sensitivity of all protocols increases as the average
+distance between binding sites does. Despite that when the binding
+events in a peak are closer to each other, both ChIP-exo and PE
+ChIP-Seq are comparable, as the distance increases ChIP-exo identifies
+a higher proportion of the RegulonDB annotations; additionally SE
+ChIP-Seq is significantly less sensitive than both ChIP-exo and PE
+ChIP-Seq. In Figure \ref{fig:reso_all}B the distance between a
+RegulonDB annotation to its closest prediction is compared for
+ChIP-exo, PE and SE ChIP-Seq; while the first two are comparable,
+both outperform SE ChIP-Seq.
+
+In Figures \ref{fig:reso_all}C and \ref{fig:reso_all}D, we observe the
 behavior of dPeak's estimated parameters for single end data (both
 ChIP-exo and ChIP-Seq). The $\delta$ parameter measures the distance
-from the 5' end of the fragment to its respective binding event and
-the $\sigma$ parameter indicates the fragments distribution
-variability around their respective binding sites (more information in
-\cite{dpeak}). For both parameters, it can be seen that the ChIP-exo
-estimates are smaller than the SET ChIP-Seq estimates in average,
-which agrees with the fact that the sample reads are allocated more
-tightly around the binding events in ChIP-exo data. Hence we can
-conclude that the peaks shape is quite different between ChIP-exo and
-SET ChIP-Seq.
-
+from the 5' end of the read's 5' end to its respective binding event
+and the $\sigma$ parameter indicates the read's starting position
+distribution variability around their respective binding sites (more
+information in \cite{dpeak}). For both parameters, it can be seen that
+the ChIP-exo estimates are smaller than the SE ChIP-Seq estimates in
+average, which agrees with the fact that the sample reads are
+allocated more tightly around the binding events in ChIP-exo
+data. This confirms that the dPeak model is able to recover the same
+conclusions that have been done when comparing ChIP-exo and SE
+ChIP-Seq.
+
+% Hence we can conclude that the peaks shape is quite different
+% between ChIP-exo and SE ChIP-Seq.
 \begin{figure}[H]
@@
   \includegraphics[width = .8\textwidth]{../figs/for_paper/saturation_analysis_old.pdf}
-  \caption{Comparison of the number of candidate regions (A),
-    predicted events (B), identified targets (C) and resolution (D)
-    among ChIP-exo, PET ChIP-Seq and SET ChIP-Seq. RegulonDB
-    annotations are considered as a gold standard. A gold standard was
-    considered identified if a BS was estimated inside a 15
-    bp window around it}
+  \caption{\textbf{ChIP-exo, PE ChIP-Seq and SE ChIP-Seq comparison at
+      varying sequencing depths}Comparison of the number of candidate
+    regions (A), predicted events (B), identified targets (C) and
+    resolution (D) among ChIP-exo, PE ChIP-Seq and SE
+    ChIP-Seq. RegulonDB annotations are considered as a gold
+    standard. A gold standard binding events was deemed identified if
+    a binding event was estimated at a $\pm$ 15 vicinity of
+    it.}
   \label{fig:design}

### 2015-11-27 rscripts/carrol_mice_library_complexity_and_enrichment.R

@@ -353,7 +353,7 @@ strata_plots[[3]] <- ggplot(dat , aes( strata , nsc,colour = file))+geom_boxplot
   scale_color_brewer(palette = "Dark2")+theme_bw()+
   theme(legend.position = "none", axis.text.x = element_text(angle = 90),plot.title = element_text(hjust = 0))+
   xlab("")+ylab("Local NSC")+ylim(0,3)+geom_abline( slope = 0 , intercept = 0, linetype = 2)+ggtitle("B")
-strata_plots[[4]] <- ggplot(dat , aes( file , nsc,colour = strata))+geom_boxplot()+facet_grid( . ~ strata) +
+strata_plots[[4]] <- ggplot(dat , aes( file , nsc,colour = file))+geom_boxplot()+facet_grid( . ~ strata) +
   scale_color_brewer(palette = "Dark2")+theme_bw()+
   theme(legend.position = "none", axis.text.x = element_text(angle = 90),plot.title = element_text(hjust = 0))+
   xlab("")+ylab("Local NSC")+ylim(0,3)+geom_abline( slope = 0 , intercept = 0, linetype = 2)+ggtitle("B")

### 2015-11-27 rscripts/local_NSC_example.R

@@ -1,0 +1,211 @@
+
+rm(list = ls())
+
+library(reshape2)
+library(ChIPUtils)
+library(ggplot2)
+library(data.table)
+library(GenomicAlignments)
+library(parallel)
+library(viridis)
+library(grid)
+library(gridExtra)
+
+dr <- "/p/keles/ChIPexo/volume3/CarrollData/mouse"
+files <- list.files(dr)
+
+files <- files[grep("bam",files)]
+files <- files[grep("bai",files,invert = TRUE)]
+
+files <- file.path(dr,files)
+
+reads <- mclapply(files,create_reads,mc.cores = 3)
+names(reads) <- files
+
+
+bamfiles <-c("ERR336942.bam","ERR336956.bam","ERR336935.bam")
+repl <- paste0("rep-",1:3)
+
+
+create_regions <- function(reads,lower,rangesOnly=TRUE)
+{
+  stopifnot(class(reads) == "GRanges")
+  stopifnot(lower > 0)
+  sreads <- split(reads,seqnames(reads))
+  sreads <- as.list(sreads)
+  cover <- mcmapply(function(x,y)coverage(x)[[y]],sreads,names(sreads),mc.cores = detectCores())
+  islands <- lapply(cover,slice,lower = lower, rangesOnly = rangesOnly)
+  islands <- mapply(function(x,y)GRanges(seqnames = y,ranges = x),islands,names(islands),SIMPLIFY =FALSE)
+  islands <- do.call(rbind,lapply(islands,gr2dt))
+  return(islands)
+}
+
+gr <- lapply(reads,function(x){
+  byChr <- mapply(rbind,readsF(x),readsR(x),SIMPLIFY =FALSE)
+  out <- do.call(rbind,byChr)
+  return(dt2gr(out))
+})
+
+regs <- lapply(gr,create_regions,lower  = 1)
+
+build_stats <- function(region,reads)
+{
+  ## fix formats and stuff
+  
+  ov <- findOverlaps(region,reads)
+  reads <- gr2dt(reads)
+  w <- width(region)    
+  region <- gr2dt(region)
+  region[ , width := w]
+  region[, match := paste0(seqnames,":",start,"_",end)]
+  reads[  subjectHits(ov), match := region[queryHits(ov), (match)] ]  
[…]

### 2015-11-27 rscripts/peak_pair_assumption_old_data.R

@@ -166,7 +166,7 @@ dt2[,Dataset:= "ChIP-seq (SET)"]
 dt = rbind(dt1,dt2)
 dt[ , Dataset := plyr::mapvalues( Dataset ,
         from = c("ChIP-exo","ChIP-seq (PET)","ChIP-seq (SET)"),
-        to = c("exo","pet","set"))]
+        to = c("ChIP-exo","ChIP-seq (PE)","ChIP-seq (SE)"))]
 
 p1 = ggplot(dt,aes(delta,colour = Dataset))+geom_density(size=1)+facet_grid(Dataset~.)+
   theme_bw()+theme(legend.position = "none",plot.title = element_text(hjust = 0))+
@@ -180,7 +180,7 @@ dt = rbind(dt1,dt2)
 
 dt[ , Dataset := plyr::mapvalues( Dataset ,
         from = c("ChIP-exo","ChIP-seq (PET)","ChIP-seq (SET)"),
-        to = c("exo","pet","set"))]
+        to = c("ChIP-exo","ChIP-seq (PE)","ChIP-seq (SE)"))]
 
 p2 = ggplot(dt,aes(sigma,colour = Dataset))+geom_density(size=1)+facet_grid(Dataset~.)+
   theme_bw()+theme(legend.position = "none",plot.title = element_text(hjust = 0))+
@@ -227,6 +227,9 @@ balanceSet = balanceSet[nfrag > 3e3]
 dt1 = data.table(seq = "exo", fratio = balance1)
 dt2 = data.table(seq = "set",fratio = balanceSet)
 dt = rbind(dt1,dt2)
+
+dt[ ,seq := plyr::mapvalues(seq , from = c("exo","set"),
+       to = c("ChIP-exo","ChIP-Seq (SE)"))]
 
 pdf(file = "figs/for_paper/forward_strand_ratio_comp_old.pdf",width = 4 ,height  = 4)
 ggplot(dt , aes(fratio, colour = seq))+geom_density()+

### 2015-11-27 rscripts/saturation_analysis_old.R

@@ -240,6 +240,12 @@ for ( i in 1:10 ) {
 
 # plot number of candidate regions
 
+## result_rene[, Seq := plyr::mapvalues(Seq,
+##    from = c("ChIP-exo","ChIP-Seq (PET)", "ChIP-Seq (SET)"),
+##    to = c("ChIP-exo","ChIP-Seq (PE)","ChIP-Seq (SE)"))]
+
+
+
 dr = "/p/keles/ChIPexo/volume3/ChIPexo/figs/for_paper"
 
 what <- c( "exo", "pet", "set")
@@ -268,27 +274,49 @@ dt4 = melt(dt4)
 dt4[,nreads:= 0.01 * 1:maxM]
 setnames(dt4,names(dt4),c("Dataset","resol","nreads"))
 
+dt1[, Dataset := plyr::mapvalues(Dataset,
+   from = c("exo","pet", "set"),
+   to = c("ChIP-exo","ChIP-Seq (PE)","ChIP-Seq (SE)"))]
+
+dt2[, Dataset := plyr::mapvalues(Dataset,
+   from = c("exo","pet", "set"),
+   to = c("ChIP-exo","ChIP-Seq (PE)","ChIP-Seq (SE)"))]
+
+dt3[, Dataset := plyr::mapvalues(Dataset,
+   from = c("exo","pet", "set"),
+   to = c("ChIP-exo","ChIP-Seq (PE)","ChIP-Seq (SE)"))]
+
+dt4[, Dataset := plyr::mapvalues(Dataset,
+   from = c("exo","pet", "set"),
+   to = c("ChIP-exo","ChIP-Seq (PE)","ChIP-Seq (SE)"))]
+
+
+
 library(grid)
 library(gridExtra)
 library(RColorBrewer)
 
 p1 <- ggplot(dt1,aes(nreads,candre, colour = Dataset))+
-  geom_line(size = 1.2)+theme_bw()+scale_color_brewer(palette = "Set1",name="")+
+  geom_line(size = 1.2)+theme_bw()+
+  scale_color_brewer(palette = "Set1",name="",guide = guide_legend(nrow = 2))+
   theme(legend.position = "top",plot.title = element_text(hjust = 0))+
   xlab("Number of reads (million)")+ylab("Number of candidate regions")+
   ylim(0,600)
 p2 <- ggplot(dt2,aes(nreads,predre, colour = Dataset))+
[…]

### 2015-11-28 rmarkdown/prelims_paper.Rnw
fixed first part of slides

@@
 
-\title{Data exploration, quality control and statistical analyses of
+\title{Data exploration, quality control and statistical analysis of
   ChIP-exo experiments\vspace*{\fill}}
@@
 the distance between binding event increases ChIP-exo shows higher
-sensitivity than PE ChIP-Seq. Finally, for given fixed sequencing
-depth, ChIP-exo provides higher sensitivity, specificity and spatial
+sensitivity than PE ChIP-Seq. Finally, for a given sequencing depth,
+ChIP-exo provides higher sensitivity, specificity and spatial
 resolution than PE ChIP-Seq.
@@
 be less complex than the libraries generated by ChIP-Seq (Mahony et
-al., 2015 \cite{exo_review}), i.e. the number of positions to which
-the reads can be aligned has been reduced due to the exonuclease
-digestion. Second, although the are roughly the same amount of reads
+al., 2015 \cite{exo_review}), i.e. the possible number of positions to
+which the reads can be aligned has been reduced due to the exonuclease
+digestion, therefore it is . Second, although the are roughly the same amount of reads
 in both strands, locally there may be more reads in one strand than in
@@
 present in ChIP-exo data and globally assess the enrichment and
-library complexity of a ChIP-exo sample.We gathered ChIP-exo data from
+library complexity of a ChIP-exo sample. We gathered ChIP-exo data from
 diverse organisms: CTCF factor in human \cite{exo1}; ER factor in
@@
 our knowledge there are not established quality control pipelines for
-ChIP-exo.Previous ChIP-exo analysis used ChIP-Seq samples to compare
+ChIP-exo. Previous ChIP-exo analysis used ChIP-Seq samples to compare
 the resolution between experiments (\cite{exo1}, \cite{exo2},
@@
 \caption{ChIP-exo diagram. The only difference with ChIP-Seq is the
-  second step, where an exonuclease enzyme is added and it digest the
+  second step, where an exonuclease enzyme is added and it digests the
   DNA fragments starting from the 5' end until it finds a crosslink
@@
 the protein of interest constructed with 5' ends of reads are located
-much closed to bound protein compared to ChIP-Seq. This step is unique
-to ChIP-exo and could potentially provide significantly higher spatial
-resolution compared to ChIP-Seq. Finally, high throughput sequencing
-of a small region (25 to 100 bp) at 5' end of each fragment generates
-millions of reads or tags.
+much closer to the bound protein compared to ChIP-Seq. This step is
+unique to ChIP-exo and could potentially provide significantly higher
+spatial resolution compared to ChIP-Seq. Finally, high throughput
+sequencing of a small region (25 to 100 bp) at 5' end of each fragment
+generates millions of reads or tags.
 
@@
     on the reverse strand are highlighted in light blue and the region
-    was centered respect to ChIP-Seq $\sig$ summit. The library
+    was centered respect to the ChIP-Seq $\sig$ summits. The library
     complexity is lower in ChIP-exo than in ChIP-Seq but the
@@
   \centering
-  \includegraphics[width = .7\textwidth]{../figs/Carroll_mice_for_paper/FoxA1_enrichment.pdf}
+  \includegraphics[width = .7\textwidth]{../figs/for_paper/FoxA1_enrichment.pdf}
   \includegraphics[width = .7\textwidth,page =4]{../figs/Carroll_mice_for_paper/Local_SCC_indicator_by_strata.pdf}
@@
   \caption{\textbf{ChIP-exo, PE ChIP-Seq and SE ChIP-Seq comparison at
-      varying sequencing depths}Comparison of the number of candidate
+      varying sequencing depths.} Comparison of the number of candidate
     regions (A), predicted events (B), identified targets (C) and
@@
 $X_1,\cdots,X_p$ and fitted a model. The model was trained by using
-10-fold CV and we selected the tunned parameters that maximized the
+10-fold CV and we selected the tuned parameters that maximized the
 area under the ROC curve. Then, an independent test set was provided
@@
 by regions with low $\mbox{ARC}$ is formed by low complexity regions,
-that way suggesting that this segment correspond to the background of
-a ChIP-exo experiment.
+hence it suggests that this arm may correspond to the background of a
+ChIP-exo experiment.
 

### 2015-11-28 rmarkdown/slides.Rnw
fixed first part of slides

@@
 
+\usepackage{multirow}
+
 \usepackage{lmodern,tabularx,ragged2e,booktabs}
@@
 
-\title[High Resolution Identification of Protein-DNA Binding Events]{High Resolution Identification of Protein-DNA Binding Events
-  and Quality Control for ChIP-exo data}
+\title[QC and analysis of ChIP-exo]{Data exploration, quality control
+  and statistical analysis of ChIP-exo experiments}
 
@@
 \institute[UW-STAT]{Department of Statistics\\University of Wisconsin - Madison}
+
+\setbeamertemplate{section in toc}[sections numbered]
+\setbeamertemplate{subsection in toc}[subsections numbered]
+
+\setbeamertemplate{section in toc}[square]
+\setbeamertemplate{subsection in toc}[square]
 
@@
 
-\section{ChIP-exo procedure}
+\section{Background}
 
@@
 \includegraphics{../figs/diagrams/ChIPexo_diagram3.png}
-    \caption{ChIP-exo procedure, the diagram is from Furey, 2012 \cite{chipbeyond}}
+    \caption{ChIP-exo procedure, Furey, 2012 \cite{chipbeyond}}
     \label{fig:exo_diagram}
@@
 
-\section{ChIP-Seq QC measures}
-\begin{frame}
+\section{ChIP-exo Data Exploration and Qualiy Control}
+
+\subsection{ChIP-Seq QC measures}
+
+\begin{frame}[t,plain]
   \frametitle{ChIP-Seq QC measures}
 
+  \begin{itemize}
 \only<1>{
-\setlength\tabcolsep{3pt}  % default value: 6pt
-\scriptsize  %%  command to change the font size
-\rowcolors{1}{}{RoyalBlue!20} %-- this indicates the change in odd and pair rows
-\begin{tabularx}{\textwidth}{ @{} c|   Y @{}}
-  \toprule
-  \textbf{QC measure} & \textbf{Definition} \\
-  \midrule
-  Nr. reads & Self-explanatory. The higher the better...  \\
-  PCR bottleneck Coeff. &  Ratio of number of pos. to which EXACTLY one read maps and number  of pos. to which AT LEAST one read maps \\
-  Standarized Std. Dev. &  Normalized Std. Deviation of the sequencing coverage \\
-  Strand Cross-Corr. &  $  y(\delta) = \sum_c w_c r\left[ n_c^+ \left(x + \frac{\delta}{2}
-    \right), n_c^- \left( x- \frac{\delta}{2} \right)\right]$ \\
-  Normalized SCC &  Ratio of max value of SCC and min value of SCC$^*$  \\
-  \bottomrule
-\end{tabularx}
+  \item Number of reads. Self-explanatory, the higher the better
 }
+  \item<2-> Number of reads.
+\only<2>{
+\item PCR bottleneck Coeff. Ratio of number of pos. to which
+  {\color{RoyalBlue}EXACTLY} one reads maps and number of pos. to
+  which {\color{RoyalBlue}AT LEAST} one reads maps }
+\only<2>{
+\begin{figure}[H]
+  \centering
+  \includegraphics[width = .75\textwidth]{../figs/for_paper/PBC_example.pdf}
+\end{figure}
+}
+\item<3-> PCR bottleneck Coeff.
+\only<3,4>{
+\item Strand Cross-Correlation. 
+\begin{align}
+  y(\delta) = \sum_c w_c r\left[ n_c^+ \left(x + \frac{\delta}{2}
+    \right), n_c^- \left( x- \frac{\delta}{2} \right)\right]
+\nonumber
+\end{align}
+}
+\only<3>{
+  \begin{figure}[H]
+    \centering
+    \includegraphics[width = .4\textwidth]{../figs/for_paper/Kharchenko.png}
+\caption{SCC explanation. Kharchenko et al., 2008 \cite{strandcc}}
+  \end{figure}
+}
+\only<4>{
+\begin{figure}[H]
+  \centering
+  \includegraphics[width = \textwidth]{../figs/for_paper/SCC.png}
+\caption{SCC as QC. Landt et al., 2012 \cite{encode_qc}}
+\end{figure}
+}
+\item<5-> Strand Cross-Correlation.
+\only<5>{
+\item Normalized Strand Cross-Correlation. Ratio between the SCC when
+  the shift is the fragment length and min. value of the SCC.
+}
+\end{itemize}
+\end{frame}
 
-\only<1>{
-where $n_c^S$ is the coverage for chromosome $c$ and strand $S$. $r$ is
-the Pearson correlation and $w_c$ is the proportion of reads in the
-experiment for chromosome $c$
-}
+\subsection{Evaluation of ChIP-Seq QC Metrics for ChIP-exo}
+\begin{frame}[plain]
+\frametitle{Evaluation of ChIP-Seq QC Metrics for ChIP-exo}
 
-\only<2>{
 \setlength\tabcolsep{3pt}  % default value: 6pt
@@
 % \rowcolor{RoyalBlue!20} %-- this indicates the change in odd and pair rows
+\begin{table}[H]
 \begin{tabularx}{\textwidth}{ @{} c|c|c|c|c|c|c| Y @{}}
   \toprule
-IP & Organism  &Condition & Rep. & Nr. reads & PBC & SSD &  NSC  \\
+\textbf{IP} & \textbf{Organism}  & \textbf{Condition} & \textbf{Rep.} & \textbf{Nr. reads} & \textbf{PBC} &  \textbf{NSC} & \textbf{Source} \\
 \midrule
-\rowcolor{RoyalBlue!20}$\sig$ & E.Coli & Rif-0min & 1 & 960,256 & 0.2823 & 0.0361 &  10.29 \\
+$\sig$ & E.Coli & Rif-0min & 1 & 960,256 & 0.2823 &   10.29   & \multirow{4}{\hsize}{Courtesy of Prof. Landick's lab} \\
+$\sig$ &  E.Coli & Rif-0min & 2 & 2,247,295 & 0.2656 &  25.08  \\
+$\sig$ &  E.Coli & Rif-20min & 1 & 1,940,387 & 0.2698 &  17.69  \\
+$\sig$ &  E.Coli & Rif-20min & 2 & 4,229,574 & 0.2153 &   14.11 \\
 \hline
-\rowcolor{RoyalBlue!20}$\sig$ &  E.Coli & Rif-0min & 2 & 2,247,295 & 0.2656 & 0.1091  &  25.08  \\
+FoxA1 &  Mouse &  - & 1 & 22,210,461 & 0.6562 & 21.452 & \multirow{3}{\hsize}{From Serandour et al., 2013 \cite{exoillumina}} \\
+FoxA1 &  Mouse &  - & 2 & 22,307,557 & 0.7996 & 60.661 \\
+FoxA1 &  Mouse &  - & 3 & 22,421,729 & 0.1068 & 72.312 \\
 \hline
-\rowcolor{RoyalBlue!20}$\sig$ &  E.Coli & Rif-20min & 1 & 1,940,387 & 0.2698 & 0.0820  & 17.69  \\
+ER & Human & - & 1 & 9,289,835 & 0.8082 & 19.843 & \multirow{3}{\hsize}{From Serandour et al., 2013 \cite{exoillumina}} \\
+ER & Human & - & 2 & 11,041,833 & 0.8024 & 21.422 \\
+ER & Human & - & 3 & 12,464,836 & 0.8203 & 19.699 \\
 \hline
-\rowcolor{RoyalBlue!20}$\sig$ &  E.Coli & Rif-20min & 2 & 4,229,574 & 0.2153 & 0.1647  &  14.11 \\
-\hline
-\rowcolor{Red!20}FoxA1 &  Mouse &  - & 1 & 22,210,461 & 0.6562 & 9.12 $\times 10^{-5}$ & 21.452 \\
-\hline
-\rowcolor{Red!20}FoxA1 &  Mouse &  - & 2 & 22,307,557 & 0.7996 & 7.94 $\times 10^{-5}$ & 60.661 \\
-\hline
-\rowcolor{Red!20}FoxA1 &  Mouse &  - & 3 & 22,421,729 & 0.1068 & 1.31 $\times 10^{-4}$ & 72.312 \\
-\hline
-\rowcolor{Green!20}ER & Human & - & 1 & 9,289,835 & 0.8082 & 3.64 $\times 10^{-5}$ & 19.843 \\
-\hline
-\rowcolor{Green!20}ER & Human & - & 2 & 11,041,833 & 0.8024 & 4.6 $\times 10^{-5}$ & 21.422 \\
-\hline
-\rowcolor{Green!20}ER & Human & - & 3 & 12,464,836 & 0.8203 & 4.89 $\times 10^{-5}$ & 19.699 \\
-\hline
-\rowcolor{Yellow!20}CTCF & Human & - & 1 &   48,478,450 & 0.4579 & 1.29 $\times 10^{-4}$ & 15.977 \\
+CTCF & Human & - & 1 &   48,478,450 & 0.4579 & 15.977 & From Rhee and Pugh 2011, \cite{exo1} \\
 \bottomrule
 \end{tabularx}  
+\end{table}
+
+{\small
+\begin{itemize}
+\item For PBC (human and mouse): 0 - 0.5 severe bottlenecking , 0.5 -
+  0.8 moderate bottlenecking, 0.8 - 0.9 mild bottlenecking, 0.9 - 1 no
+  bottlenecking.
+\item For NSC (human and mouse): $ < 1.1$ is relatively low.
+\end{itemize}
 }
-
 \end{frame}
 
-\section{Comparison of ChIP-exo and ChIP-seq}
+\subsection{Comparison of ChIP-exo and ChIP-seq}
 
 \begin{frame}
-  \frametitle{Comparison of ChIP-exo and ChIP-seq I}  
-\begin{figure}[H]
-  \centering  
-\includegraphics[width = .7\textwidth]{../figs/for_paper/scc_ctcf.pdf}
-\caption{SCC for CTCF factor in HeLa cell line for ChIP-exo and
-  SET-ChIP-Seq }
-  \label{fig:scc}
-\end{figure}
-
-\begin{itemize}
-\item There is a \emph{``phantom peak''} at read length.
-\item In ChIP-Seq SCC is maximized at the unobserved fragment length.
-\item In ChIP-exo, the \emph{``phantom peak''} and the frag. length
-  summit are confounded.
-\end{itemize}
-
-\end{frame}
-
-\begin{frame}
-  \frametitle{Comparison of ChIP-exo and ChIP-Seq II}
+  \frametitle{Comparison of ChIP-exo and ChIP-Seq}
 
@@
 \begin{itemize}
-\item A shows that high dnesity regions are similar between ChIP-Seq
+\item A shows that high density regions are similar between ChIP-Seq
   and ChIP-exo but background regions are not.
-\item The peak-pair assumption doesn't hold in ChIP-exo data, some
-  regions show strand-imbalance
+\item The peak-pair assumption does hold in ChIP-exo data, but not
+  locally since some regions show strand-imbalance.
 \end{itemize}
@@
 
-% \only<2>{
-% \begin{figure}[H]
-% \centering
-% \includegraphics[width = .46\textwidth,page = 1]{../figs/for_paper/eukaryotic_bias_CTCF.pdf}
-%   \includegraphics[width = .46\textwidth,page = 2]{../figs/for_paper/eukaryotic_bias_CTCF.pdf}
-% \end{figure}
-% }
 \end{frame}
 
-\section{The ChIP-exo QC pipeline}
+\begin{frame}[t]
+  \frametitle{Comparison of ChIP-exo and ChIP-seq}  
+\begin{columns}
+\column{.7\textwidth}
+\begin{figure}[H]
+  \centering  
+\includegraphics[width = .9\textwidth]{../figs/for_paper/scc_ctcf.pdf}
+\end{figure}
 
-\begin{frame}[plain]
-  \frametitle{The ChIP-exo QC pipeline}
+\column{.3\textwidth}
+  \begin{minipage}[c][.6\textheight][c]{\linewidth}
+{\small
+  {\color{RoyalBlue}Figure:} SCC for CTCF factor in HeLa cell line for
+    ChIP-exo and SET-ChIP-Seq.\\
+    - {\color{ForestGreen} PBC for rep1 is 0.56}
+    - {\color{ForestGreen} PBC for rep2 is 0.94}
+  }
+\end{minipage}
+\end{columns}
+
+{\small
+\begin{itemize}
+\item There is a \emph{``phantom peak''} at read length.
+\item In ChIP-Seq SCC is maximized at the unobserved fragment length.
+\item In ChIP-exo, the \emph{``phantom peak''} and the fragment length
+  summit are confounded.
+\end{itemize}
+}
+\end{frame}
+
+\subsection{Quality Control pipeline}
+
+\begin{frame}[plain,t]
 
@@
   \centering
-  \includegraphics[width = \textwidth]{../figs/for_paper/coverage_diagram2.pdf}
+  \includegraphics[width = .75\textwidth]{../figs/for_paper/coverage_diagram2.pdf}
 \end{figure}
@@
 \end{frame}
+
+\section{Systematic Evaluation of ChIP-Seq and ChIP-exo}
 
@@
   \item Sensitivity is the defined as the proportion of identified
-    peaks (regulonDoB \cite{regulonDB} is used as gold-stantard)
+    peaks (regulonDB \cite{regulonDB} is used as gold-standard)
   \item Resolution is defined as the min. absolute distance of a
@@
   states.
-\item Find a optimal strategy for labelling enhancer out of a
+\item Find a optimal strategy for labeling enhancer out of a
   predetermined list of regions in the genome by the use of active
@@
 \end{document}
+
+% LocalWords:  regulonDoB

### 2015-11-28 rscripts/Strand_cross_corr_comp_CTCF.R

@@ -13,19 +13,35 @@ sizes <- data.table(read.table(hg19))
 shift <- 1:200
 
 exo_file <- "/p/keles/ChIPexo/volume3/PughData/CTCF.bam"
-set_file <- "/p/keles/ChIPexo/volume3/PughData/encode-Uw-Helas3-Ctcf-rep1.bam"
+set_file1 <- "/p/keles/ChIPexo/volume3/PughData/encode-Uw-Helas3-Ctcf-rep1.bam"
+set_file2 <- "/p/keles/ChIPexo/volume3/PughData/encode-Uw-Helas3-Ctcf-rep2.bam"
 
 exo <- create_reads(exo_file)
-set <- create_reads(set_file)
+set1 <- create_reads(set_file1)
+set2 <- create_reads(set_file2)
 
 
 scc_exo <- strand_cross_corr(exo,chrom.sizes = sizes,parallel = TRUE,shift = 1:300)
-scc_set <- strand_cross_corr(set,chrom.sizes = sizes,parallel = TRUE,shift = 1:300) 
+scc_set1 <- strand_cross_corr(set1,chrom.sizes = sizes,parallel = TRUE,shift = 1:300) 
+scc_set2 <- strand_cross_corr(set2,chrom.sizes = sizes,parallel = TRUE,shift = 1:300) 
 
-dt <- rbind(scc_exo[,seq := "exo"],scc_set[,seq := "set"])
+scc_exo[,seq := "ChIP-exo"]
+scc_set1[,seq := "ChIP-Seq (SE)"]
+scc_set2[,seq := "ChIP-Seq (SE)"]
 
-pdf(file ="figs/for_paper/scc_ctcf.pdf",width = 8,height = 4)
-ggplot(dt,aes(shift , cross.corr,colour = seq))+geom_line(size = 1)+
-  theme_bw()+scale_color_manual(name = "",values = brewer.pal(3,"Set1")[c(1,3)])+
+
+dt <- rbind(scc_exo[, repl := 1],
+            scc_set1[,repl := 1],
+            scc_set2[,repl := 2])
+
+pdf(file ="figs/for_paper/scc_ctcf.pdf",width = 8,height = 5)
+ggplot(dt,
+  aes(shift , cross.corr,colour = seq,linetype = as.factor(repl)))+
+  geom_line(size = 1)+
+  theme_bw()+
+  scale_color_manual(name = "Protocol",values = brewer.pal(3,"Set1")[c(1,3,3)],
[…]

### 2015-11-28 rscripts/local_NSC_example.R

@@ -106,6 +106,31 @@ regs <- lapply(regs,dt2gr)
 
 stats <- mcmapply(build_stats,regs,gr,mc.cores = 6 ,SIMPLIFY = FALSE)
 
+aux <- mapply(function(x,y)x[,sample := y],stats,basename(names(stats)),SIMPLIFY =FALSE)
+aux <- do.call(rbind,aux)
+aux[, sample := plyr::mapvalues(sample , from = c("ERR336935.bam","ERR336942.bam","ERR336956.bam"),
+    to = c("rep-3","rep-1","rep-2"))]
+
+library(hexbin)
+library(scales)
+r <- viridis::viridis(100,option = "D")
+
+figs_dir <- "figs/for_paper"
+
+pdf(file = file.path(figs_dir,"FoxA1_enrichment.pdf"),width = 9 , height = 5 )
+p <- ggplot(aux , aes( ave_reads,cover_rate))+stat_binhex(bins = 50)+
+  facet_wrap( ~ sample)+scale_fill_gradientn(colours = r,trans = "log10",
+    labels = trans_format("log10",math_format(10^.x)))+xlim(0,4)+ylim(0,1)+
+  theme_bw()+theme(legend.position = "top",plot.title = element_text(hjust = 0))+
+  xlab("Average read coverage (ARC)")+
+  ylab("Unique read coverage rate (URCR)")
+print(p + ggtitle("All regions"))
+print( p  %+% aux[ npos > 10] + ggtitle("A"))  ## npos > 10 
+print( p  %+% aux[ npos > 30] + ggtitle("B")) ## npos > 30
+dev.off()
+
+
+
 all_stats <- stats
 
 ## candidates are high complexity regions
@@ -198,10 +223,11 @@ make_plots <- function(reg,regs,reads,nms)
 
 nms <- c("rep-3","rep-1","rep-2")
 Z <- make_plots(regs[[2]][1],regs,reads,nms)
-
 vplayout <- function(x, y) viewport(layout.pos.row = x, layout.pos.col = y)
 
 pdf(file = "figs/for_paper/local_SCC_example.pdf",width = 9,height = 5)
+
+pdf(width = 9,height = 5)
 grid.newpage()
 pushViewport(viewport(layout = grid.layout(1, 5)))
 print(Z[[1]]+ggtitle("A"), vp = vplayout(1,1:3))

### 2015-11-29 rmarkdown/prelims_paper.Rnw
finished first set of corrections to prelim paper

@@
 and library complexity. Assessment of these biases and artifacts are
-facilitated through diagnostic plots and summary statistical that
-compare a large portion of the genome as partitioned into island with
-high depth regions.
+facilitated through diagnostic plots and summary statistics calculated
+over regions of the genome with varying levels of coverage.
 
@@
 different than ChIP-Seq's. Second, although often assumed in ChIP-exo
-data analysis methods, the ``peak pair'' assumption doesn't hold
-locally in real ChIP-exo data. Third, we compared Paired End (PE)
-ChIP-Seq with ChIP-exo and found that both protocols are comparable in
-resolution and sensitivity for closely located binding event, but as
-the distance between binding event increases ChIP-exo shows higher
-sensitivity than PE ChIP-Seq. Finally, for a given sequencing depth,
-ChIP-exo provides higher sensitivity, specificity and spatial
-resolution than PE ChIP-Seq.
+data analysis methods, the ``peak pair'' assumption does not hold
+locally in actual ChIP-exo data. Third, we for the first time compared
+Paired End (PE) ChIP-Seq with ChIP-exo and found that both protocols
+are comparable in resolution and sensitivity for closely located
+binding event, but as the distance between binding event increases
+ChIP-exo shows higher sensitivity than PE ChIP-Seq. Finally, at fixed
+sequencing depths, ChIP-exo provides higher sensitivity, specificity
+and spatial resolution than PE ChIP-Seq.
 
@@
 \[
-\mbox{NSC} = \frac{ \max_\delta y(\delta)}{\min_\delta y(\delta)}
+\mbox{NSC} = \frac{ \max_\delta y(\delta)}{\min_\delta y(\delta)},
 \]
-
 where $y(\delta)$ is strand cross-correlation defined as in (\ref{scc}).
@@
   y(\delta) = \sum_c w_c r\left[ n_c^+ \left(x + \frac{\delta}{2}
-    \right), n_c^- \left( x- \frac{\delta}{2} \right)\right]
+    \right), n_c^- \left( x- \frac{\delta}{2} \right)\right],
 \label{scc}
@@
 observed and it follows a
-$\mbox{Multinomial}(\pi_0,\pi_1,\cdots,\pi_{g^*})$. And the 5' end
-$R_i$ for the $i$-th fragment is generated according to the following
-procedure:
+$\mbox{Multinomial}(\pi_0,\pi_1,\cdots,\pi_{g^*})$. Then the observed
+data, i.e., the 5' end $R_i$ for the $i$-th fragment is generated
+according to the following procedure:
 
 \begin{itemize}
-\item If the read was sequenced from the forward strand ($D_i = 1$):
+\item If the read is sequenced from the forward strand ($D_i = 1$):
   \begin{itemize}
@@
   \end{itemize}
-\item If the read was sequenced from the backward strand ($D_i =
+\item If the read is sequenced from the backward strand ($D_i =
   0$):
@@
 
-For this work, we considered a new strategy where the $mu_g$'s are
+For this work, we considered a new strategy where the $\mu_g$'s are
 initialized by partitioning the region into $g^*$ bins and finding the
-coverage's local maxima in each bin and we initialize $\delta$ and
+coverage's local maxima in each bin we initialize $\delta$ and
 $\sigma$ by ranking the regions by the signal strength or an empirical
@@
   \includegraphics[width = .7\textwidth]{../figs/for_paper/coverage_diagram2.pdf}
-  \caption{Flowchart of the ChIP-exo Quality control pipeline.}
+  \caption{Flowchart of the ChIP-exo quality control pipeline.}
   % The genome is partitioned into several regions by removing the empty
@@
   y(\delta) = r\left[ n^+ \left(x + \frac{\delta}{2}
-    \right), n^- \left( x- \frac{\delta}{2} \right)\right]
+    \right), n^- \left( x- \frac{\delta}{2} \right)\right],
 \nonumber

### 2015-11-29 rmarkdown/slides.Rnw
finished first set of corrections to prelim paper

@@
 
-\subsection{Quality Control pipeline}
+\subsection{Quality Control Pipeline}
 
-\begin{frame}[plain,t]
+\begin{frame}[t]
+  \frametitle{Quality Control Pipeline}
 
+  \begin{enumerate}
+\only<1>{
+  \item Partition the genome and generate ChIP-exo islands.
 \begin{figure}[H]
 \centering
-  \includegraphics[width = .75\textwidth]{../figs/for_paper/coverage_diagram2.pdf}
+\includegraphics[width = \textwidth]{../figs/for_paper/coverage_diagram2_part1.pdf}
 \end{figure}
+}
+\only<2>{
+\addtocounter{enumi}{1}
+\item Calculate a vector of summary statistics for each islands.
+\begin{figure}[H]
+\centering
+\includegraphics[width = \textwidth]{../figs/for_paper/coverage_diagram2_part2.pdf}
+\end{figure}
+}
+\only<3>{
+\addtocounter{enumi}{2}
+\item Visualize all islands together.
+\begin{figure}[H]
+\centering
+\includegraphics[width = \textwidth]{../figs/for_paper/coverage_diagram2_part3.pdf}
+\end{figure}
+}
+  \end{enumerate}
+\end{frame}
 
-{\small
-  \begin{itemize}
-  \item $\mbox{ARC} = \frac{\text{Nr. of reads in the
-        region}}{\text{Width of the region}}$
-  \item $\mbox{URCR} = \frac{\text{Nr. of reads mapped to only one
-        position in the region}}{\text{Nr. of reads in the region}}$
-  \item $\mbox{local-NSC}$
-  \item $\mbox{FSR} = \frac{\text{Nr. of fwd. strand reads in region}}{\text{Nr. of reads in region}}$
-  \end{itemize}  
-}
-
-\end{frame}
+  % \begin{itemize}
+  % \item $\mbox{ARC} = \frac{\text{Nr. of reads in the
+  %       region}}{\text{Width of the region}}$
+  % \item $\mbox{URCR} = \frac{\text{Nr. of reads mapped to only one
+  %       position in the region}}{\text{Nr. of reads in the region}}$
+  % \item $\mbox{local-NSC}$
+  % \item $\mbox{FSR} = \frac{\text{Nr. of fwd. strand reads in region}}{\text{Nr. of reads in region}}$
+  % \end{itemize}
 
@@
 \item med - regions with nr. of unique positions in $(50,100)$
-\item low - regions with nr. of unique positions in $(20,100)$
+\item low - regions with nr. of unique positions in $(20,50)$
 \end{itemize}
@@
 
-\section{Comparison with ChIP-Seq using dPeak}
 \begin{frame}[t]
@@
 
+\section{Comparison with Other Algorithms}
+\begin{frame}
+\frametitle{Comparison with Other Algorithms}
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width = .7\textwidth]{../figs/for_paper/algorithm_resolution.pdf}
+\end{figure}
+
+\end{frame}
+
 \section{Conclusions and future work}
@@
 
-\begin{frame}
+\begin{frame}[allowframebreaks]
   \frametitle{References}

### 2015-11-29 rscripts/sig70_methods_resolution.R

@@ -74,16 +74,18 @@ apex_res <- sapply(sites[,(start)],resolution,start(apex))
 gem_res <- sapply(sites[,(start)],resolution , start(gem))
 
 res1 <- data.table(reso = dpeak_res,method = "dPeak")
-res2 <- data.table(reso = apex_res, method = "Apex")
+res2 <- data.table(reso = apex_res, method = "Mace")
 res3 <- data.table(reso = gem_res,method = "Gem")
 
 reso <- rbind(res1,res2,res3)
-reso[ , method := factor(method, levels = c("dPeak","Apex","Gem"))]
+reso[ , method := factor(method, levels = c("dPeak","Mace","Gem"))]
 
-pdf(file = "figs/for_paper/algorithm_resolution.pdf")
-ggplot(reso[reso < 200], aes(method , reso))+geom_boxplot()+
+pdf(file = "figs/for_paper/algorithm_resolution.pdf",width = 5 , height = 5)
+ggplot(reso[reso < 200], aes(method , reso,fill = method))+geom_boxplot()+
   xlab("Algorithm")+ylab("Resolution")+coord_cartesian(ylim = c(0,150))+
-  theme_bw()+theme(plot.title = element_text(hjust = 0))
+  theme_bw()+
+  theme(plot.title = element_text(hjust = 0),legend.position = "none")+
+  scale_fill_brewer(palette = "Pastel1")
 dev.off()
 
 

### 2015-11-30 rmarkdown/prelims_paper.Rnw
second draft of slides

@@
 \url{https://www.encodeproject.org/data-standards/2012-quality-metrics/}),
-but it doesn't apply to ChIP-exo, if it did all the samples with
+but it does not apply to ChIP-exo, if it did all the samples with
 $\mbox{PBC} < 0.5$ should be reject due to be showing severe
@@
 ChIP-Seq dataset behaves as the successful case from \cite{encode_qc},
-but the ChIP-exo doesn't. Additionally, since the exonuclease enzyme
+but the ChIP-exo does not. Additionally, since the exonuclease enzyme
 digests the DNA fragments, then the shifts where the SCC may reach a
@@
 both strands. However, Figure \ref{fig:comp}B shows that this value
-doesn't represent the sample locally, hence the assumption doesn't
+does not represent the sample locally, hence the assumption does not
 hold well in practice.
@@
 tests to assess if the FSR's distribution was the same for the islands
-that overlap with a peak and the islands that didn't. The p.value for
+that overlap with a peak and the islands that did not. The p.value for
 the third replicate's test is \Sexpr{pval[[1]]}, for the first
@@
 Previously, ChIP-exo and SE ChIP-Seq have been compared at a fixed
-depth level in the literature, but this comparisons didn't included PE
+depth level in the literature, but this comparisons did not included PE
 ChIP-Seq. Hence, we sampled a fixed amount of reads for each of the
@@
 for ChIP-exo or SE ChIP-seq.
+
+% \subsection{Comparison with other algorithms}
+% \label{sec:comp_algo}
+
+% \begin{figure}[H]
+%   \centering
+%   \includegraphics[width = .5\textwidth]{../figs/for_paper/algorithm_resolution.pdf}
+%   \caption{Comparison of resolution of the algorithms for ChIP-exo
+%     data, using regulonDB annotations as a gold standard.}
+%   \label{fig:algor}
+% \end{figure}
 
@@
 between enriched samples and low complexity regions. It is shown that
-the ``peak-pair'' assumption doesn't hold locally in practice and we
+the ``peak-pair'' assumption does not hold locally in practice and we
 provide two out-of-the-box visualization capable to assess the strand
@@
 transcription factors under two conditions, one where rifampicin was
-applied by 20 minutes and another where it wasn't applied. As seen in
+applied by 20 minutes and another where it was not applied. As seen in
 the Figure, there is ChIP-exo data for the 3 factors and PE ChIP-Seq

### 2015-11-30 rmarkdown/prelims_paper.tex
second draft of slides

@@
 tests to assess if the FSR's distribution was the same for the islands
-that overlap with a peak and the islands that didn't. The p.value for
+that overlap with a peak and the islands that did not. The p.value for
 the third replicate's test is 0.1039, for the first

### 2015-11-30 rmarkdown/slides.Rnw
second draft of slides

@@
 
+\usepackage{wrapfig}
+
 \usepackage{multirow}
@@
 % Personal Data
+\def\Put(#1,#2)#3{\leavevmode\makebox(0,0){\put(#1,#2){#3}}}
 
@@
 \centering
-\includegraphics[width = \textwidth]{../figs/for_paper/coverage_diagram2_part3.pdf}
+\includegraphics[width = \textwidth]{../figs/for_paper/coverage_diagram2_part3P.pdf}
 \end{figure}
@@
  % \begin{itemize}
-  % \item $\mbox{ARC} = \frac{\text{Nr. of reads in the
-  %       region}}{\text{Width of the region}}$
-  % \item $\mbox{URCR} = \frac{\text{Nr. of reads mapped to only one
-  %       position in the region}}{\text{Nr. of reads in the region}}$
   % \item $\mbox{local-NSC}$
@@
 
-\section{Systematic Evaluation of ChIP-Seq and ChIP-exo}
-
-\begin{frame}[t,plain]
-  \frametitle{Library complexity and enrichment}
+\begin{frame}
+  \frametitle{ARC vs. URCR exploration}
 \only<1>{
+  \begin{itemize}
+  \item $\mbox{ARC} = \frac{\text{Nr. of reads in the
+        region}}{\text{Width of the region}}$
+  \item $\mbox{URCR} = \frac{\text{Nr. of reads mapped to only one
+        position in the region}}{\text{Nr. of reads in the region}}$
+  \end{itemize}
+}
+\only<2>{
+All regions
   \begin{figure}[H]
     \centering
-  \includegraphics[width = \textwidth]{../figs/Carroll_mice_for_paper/FoxA1_enrichment.pdf}
+    \includegraphics[width = .7\textwidth,page = 1]{../figs/for_paper/ARC_vs_URCR_example_npos.pdf}
+  \end{figure}
+}
+
+\only<3>{
+Mapped to more than 10 positions
+  \begin{figure}[H]
+    \centering
+    \includegraphics[width = .7\textwidth,page = 2]{../figs/for_paper/ARC_vs_URCR_example_npos.pdf}
+  \end{figure}
+}
+
+\only<4>{
+Mapped to more than 30 positions
+  \begin{figure}[H]
+    \centering
+    \includegraphics[width = .7\textwidth,page = 3]{../figs/for_paper/ARC_vs_URCR_example_npos.pdf}
+  \end{figure}
+}
+\only<5>{
+Mapped to more than 50 positions
+  \begin{figure}[H]
+    \centering
+    \includegraphics[width = .7\textwidth,page = 4]{../figs/for_paper/ARC_vs_URCR_example_npos.pdf}
+  \end{figure}
+}
+
+\end{frame}
+
+\begin{frame}
+  \frametitle{ARC vs. URCR exploration}
+
+\only<1-2>{
+  \begin{figure}[H]
+    \centering
+    \includegraphics[width = .7\textwidth,page =1]{../figs/for_paper/enrichment_example.pdf}
   \end{figure}
@@
 \only<2>{
-{\small
+\Put(120,200){\includegraphics[width = .6\textwidth,page = 1]{../figs/for_paper/coverage_example.pdf}}
+}
+
+\only<3-4>{
   \begin{figure}[H]
     \centering
-  \includegraphics[width = \textwidth]{../figs/for_paper/local_SCC_example.pdf}
+    \includegraphics[width = .7\textwidth,page =2]{../figs/for_paper/enrichment_example.pdf}
   \end{figure}
-}}
-\only<2>{
-\begin{align}
- y(\delta) = f(x_\delta) + \epsilon_\delta \quad 
-\mbox{local-NSC} = \frac{max_{x_\delta} \hat{f}(x_\delta)}{\hat{\sigma}_f}
-\nonumber
-\end{align}
 }
 
-\only<3>{
+\only<4>{
+\Put(120,200){\includegraphics[width = .6\textwidth,page = 2]{../figs/for_paper/coverage_example.pdf}}
+}
+
+\only<5-6>{
   \begin{figure}[H]
     \centering
-  \includegraphics[width = \textwidth,page =3]{../figs/Carroll_mice_for_paper/Local_SCC_indicator_by_strata.pdf}
+    \includegraphics[width = .7\textwidth,page =3]{../figs/for_paper/enrichment_example.pdf}
+  \end{figure}
+}
+
+\only<6>{
+\Put(120,200){\includegraphics[width = .6\textwidth,page = 3]{../figs/for_paper/coverage_example.pdf}}
+}
+
+\only<7-8>{
+  \begin{figure}[H]
+    \centering
+    \includegraphics[width = .7\textwidth,page = 4]{../figs/for_paper/enrichment_example.pdf}
+  \end{figure}
+}
+
+\only<8>{
+\Put(120,200){\includegraphics[width = .6\textwidth,page = 4]{../figs/for_paper/coverage_example.pdf}}
+}
+
+% \only<9-10>{
+%   \begin{figure}[H]
+%     \centering
+%     \includegraphics[width = .7\textwidth,page = 5]{../figs/for_paper/enrichment_example.pdf}
+%   \end{figure}
+% }
+
+% \only<10>{
+% \Put(120,200){\includegraphics[width = .6\textwidth,page = 5]{../figs/for_paper/coverage_example.pdf}}
+% }
+
+\only<9-10>{
+  \begin{figure}[H]
+    \centering
+    \includegraphics[width = .7\textwidth,page =6]{../figs/for_paper/enrichment_example.pdf}
+  \end{figure}
+}
+
+\only<10>{
+\Put(120,300){\includegraphics[width = .6\textwidth,page = 6]{../figs/for_paper/coverage_example.pdf}}
+}
+
+\only<11-12>{
+  \begin{figure}[H]
+    \centering
+    \includegraphics[width = .7\textwidth,page =7]{../figs/for_paper/enrichment_example.pdf}
+  \end{figure}
+}
+
+\only<12>{
+\Put(120,300){\includegraphics[width = .6\textwidth,page = 7]{../figs/for_paper/coverage_example.pdf}}
+}
+
+\only<13-14>{
+  \begin{figure}[H]
+    \centering
+    \includegraphics[width = .7\textwidth,page =8]{../figs/for_paper/enrichment_example.pdf}
+  \end{figure}
+}
+
+\only<14>{
+\Put(120,300){\includegraphics[width = .6\textwidth,page = 8]{../figs/for_paper/coverage_example.pdf}}
+}
+
+\end{frame}
+
+\begin{frame}
+  \frametitle{ARC vs. URCR exploration by replicate}
+\only<1>{
+\begin{figure}[H]
+  \centering
+  \includegraphics[width = \textwidth,page = 1]{../figs/Carroll_mice_for_paper/FoxA1_enrichment.pdf}
+\end{figure}
+}
+
+\only<2>{
+\begin{figure}[H]
+  \centering
+  \includegraphics[width = \textwidth,page = 3]{../figs/Carroll_mice_for_paper/FoxA1_enrichment.pdf}
+\end{figure}
+}
+
+\end{frame}
+
+\begin{frame}
+  \frametitle{ARC vs. URCR and local-NSC}
+
+\only<1>{
+\begin{figure}[H]
+  \centering
+  \includegraphics[width = \textwidth,page = 1]{../figs/for_paper/local_SCC_separated.pdf}
+\end{figure}
+}
+
+\only<2>{
+\begin{figure}[H]
+  \centering
+  \includegraphics[width = \textwidth,page = 2]{../figs/for_paper/local_SCC_separated.pdf}
 \end{figure}
@@
 \only<3>{
-{\small
+\begin{align}
+ y(\delta) = f(x_\delta) + \epsilon_\delta, \quad 
+\mbox{local-NSC} = \frac{max_{x_\delta} \hat{f}(x_\delta)}{\hat{\sigma}_f},
+\nonumber
+\end{align}
+\vspace*{\fill}
+}
+
+\only<3>{
+\begin{columns}
+
+\column{.4\textwidth}
 where:
+
 \begin{itemize}
-\item high - regions with nr. of unique positions $ > 100$
-\item med - regions with nr. of unique positions in $(50,100)$
-\item low - regions with nr. of unique positions in $(20,50)$
+\item $x_\delta$ is the shift.
+\item $y_\delta$ is the local - SCC at shift $x_\delta$
+\item $\epsilon \sim N(0, \sigma^2)$ i.i.d 
 \end{itemize}
-}}
 
+\column{.6\textwidth}
+{\scriptsize
+  \begin{table}[H]
+    \centering
+    \begin{tabularx}{\textwidth}{ @{} l | r | r |r @{}}
+      \toprule
+      Replicate & $\hat{\sigma}$ & $\max_{x_\delta} \hat{f}(x_\delta)$ & $\mbox{local-NSC}$ \\
+      \midrule
+Rep. 1 & 0.0533 & 0.2215 & 4.14 \\
+Rep. 2 & 0.0453 & 0.1606 & 3.54 \\
+Rep. 3 & 0.0475 & 0.0832 & 1.75 \\
+\bottomrule
+    \end{tabularx}
+  \end{table}
+}
+
+\end{columns}
+}
+
+\only<4>{
+  \begin{figure}[H]
+    \centering
+      \includegraphics[width = \textwidth,page =4]{../figs/Carroll_mice_for_paper/Local_SCC_indicator_by_strata.pdf}
+  \end{figure}
+
+  \begin{itemize}
+  \item {\color{RoyalBlue} high:} Number of unique positions $ > 100$
+  \item {\color{RoyalBlue} med:} Number of unique positions in $(50,100)$
+  \item {\color{RoyalBlue} low:} Number of unique positions in $(20,50)$
+  \end{itemize}
+}
 \end{frame}
@@
 
+{\scriptsize
+\begin{itemize}
+\item  
+\end{itemize}
+}
 \end{frame}
 
-\begin{frame}[t]
+\section{Systematic Evaluation of ChIP-Seq and ChIP-exo}
+
+\begin{frame}[plain]
   \frametitle{dPeak model for SET case}
@@
 \only<1>{
+
+      \begin{figure}[H]
+        \centering
+        \includegraphics[width =
+        \textwidth]{/p/keles/ChIPexo/volume3/ChIPexo/poster/dpeak.png}
+      \end{figure}
+}
+
+\only<2>{
+\begin{columns}[t]
+\column{.3\textwidth}
+  \begin{figure}[H]
+    \centering
+    \includegraphics[width =\textwidth]
+{/p/keles/ChIPexo/volume3/ChIPexo/poster/dpeak.png}
+  \end{figure}
+
+\column{.7\textwidth}
   We consider a region with $n$ reads and $m$ positions, for the
@@
       \item The reads belongs to the background: $R_i | Z_i = 0, D_i =
-        1 \sim \mbox{Unif}(1 - \beta + 1, m)$
+        1 \sim \mbox{Unif}(1, m)$
       \item The read belong to the $g$-th binding event: $R_i | Z_i =
@@
       \item The reads belongs to the background: $R_i | Z_i = 0, D_i =
-        0 \sim \mbox{Unif}(1 , m + \beta - 1)$
+        0 \sim \mbox{Unif}(1 , m )$
       \item The read belong to the $g$-th binding event: $R_i | Z_i =
@@
   \end{itemize}
+\end{columns}
 }
 
-\only<2>{
+\end{frame}
+
+\begin{frame}
+  \frametitle{Structure of the experiment}
+
   \begin{figure}[H]
     \centering
-        \includegraphics[width =
-        .8\textwidth]{/p/keles/ChIPexo/volume3/ChIPexo/poster/dpeak.png}
+
   \end{figure}
-}
+
 \end{frame}

### 2015-11-30 rmarkdown/slides.tex
second draft of slides

@@
 
-\begin{frame}[t]
+\section{Systematic Evaluation of ChIP-Seq and ChIP-exo}
+
+\begin{frame}[plain]
   \frametitle{dPeak model for SET case}

### 2015-11-30 rscripts/local_NSC_example.R

@@ -47,64 +47,79 @@ gr <- lapply(reads,function(x){
 })
 
 regs <- lapply(gr,create_regions,lower  = 1)
+regs <- lapply(regs,dt2gr)
 
-build_stats <- function(region,reads)
+
+
+build_stats <- function(region,reads,mc)
 {
   ## fix formats and stuff
-  
-  ov <- findOverlaps(region,reads)
-  reads <- gr2dt(reads)
-  w <- width(region)    
-  region <- gr2dt(region)
-  region[ , width := w]
-  region[, match := paste0(seqnames,":",start,"_",end)]
-  reads[  subjectHits(ov), match := region[queryHits(ov), (match)] ]  
-  reads[,strand := ifelse(strand == "+", "F","R")]
-  reads <- reads[!is.na(match)]
+  stat_by_chr <- function(reg,read)
+  {
+     ov <- findOverlaps(reg,read)  
+     read <- gr2dt(read)
+     w <- width(reg)    
+     region <- gr2dt(reg)
+     region[ , width := w]
+     region[, match := paste0(seqnames,":",start,"_",end)]
+     read[  subjectHits(ov), match := region[queryHits(ov), (match)] ]  
+     read[,strand := ifelse(strand == "+", "F","R")]
+     read <- read[!is.na(match)]
 
-  ## get base statistics
-  f <- reads[,sum(strand == "F"),by = match]
-  setnames(f,names(f),c("match","f"))
-  setkey(f,match)
-  r <- reads[,sum(strand == "R"),by = match]
-  setkey(r,match)
-  setnames(r,names(r),c("match","r"))
-  f_uniq <- reads[strand == "F",length(unique(start)),by = match]
-  setnames(f_uniq,names(f_uniq),c("match","f_pos"))
-  setkey(f_uniq,match)
-  r_uniq <- reads[strand == "R",length(unique(end)),by = match]
-  setnames(r_uniq,names(r_uniq),c("match","r_pos"))
-  setkey(r_uniq,match)
+     ## get base statistics
+     f <- read[,sum(strand == "F"),by = match]
+     setnames(f,names(f),c("match","f"))
+     setkey(f,match)
+     r <- read[,sum(strand == "R"),by = match]
+     setkey(r,match)
[…]

### 2015-11-30 rscripts/local_NSC_example_Rep1.R

@@ -1,0 +1,240 @@
+
+rm(list = ls())
+
+library(reshape2)
+library(ChIPUtils)
+library(ggplot2)
+library(data.table)
+library(GenomicAlignments)
+library(parallel)
+library(viridis)
+library(grid)
+library(gridExtra)
+
+dr <- "/p/keles/ChIPexo/volume3/CarrollData/mouse"
+files <- list.files(dr)
+
+files <- files[grep("bam",files)]
+files <- files[grep("bai",files,invert = TRUE)]
+
+files <- file.path(dr,files)
+
+reads <- mclapply(files,create_reads,mc.cores = 3)
+names(reads) <- files
+
+
+bamfiles <-c("ERR336942.bam","ERR336956.bam","ERR336935.bam")
+repl <- paste0("rep-",1:3)
+
+
+create_regions <- function(reads,lower,rangesOnly=TRUE)
+{
+  stopifnot(class(reads) == "GRanges")
+  stopifnot(lower > 0)
+  sreads <- split(reads,seqnames(reads))
+  sreads <- as.list(sreads)
+  cover <- mcmapply(function(x,y)coverage(x)[[y]],sreads,names(sreads),mc.cores = detectCores())
+  islands <- lapply(cover,slice,lower = lower, rangesOnly = rangesOnly)
+  islands <- mapply(function(x,y)GRanges(seqnames = y,ranges = x),islands,names(islands),SIMPLIFY =FALSE)
+  islands <- do.call(rbind,lapply(islands,gr2dt))
+  return(islands)
+}
+
+gr <- lapply(reads,function(x){
+  byChr <- mapply(rbind,readsF(x),readsR(x),SIMPLIFY =FALSE)
+  out <- do.call(rbind,byChr)
+  return(dt2gr(out))
+})
+
+regs <- lapply(gr,create_regions,lower  = 1)
+regs <- lapply(regs,dt2gr)
+
+build_stats <- function(region,reads,mc)
+{
+  ## fix formats and stuff
+  stat_by_chr <- function(reg,read)
+  {
+     ov <- findOverlaps(reg,read)  
+     read <- gr2dt(read)
+     w <- width(reg)    
+     region <- gr2dt(reg)
+     region[ , width := w]
+     region[, match := paste0(seqnames,":",start,"_",end)]
+     read[  subjectHits(ov), match := region[queryHits(ov), (match)] ]  
[…]

### 2015-11-30 rmarkdown/slides.Rnw
added diagrams to slides

@@
 ER & Human & - & 2 & 11,041,833 & 0.8024 & 21.422 \\
-ER & Human & - & 3 & 12,464,836 & 0.8203 & 19.699 \\
+R & Human & - & 3 & 12,464,836 & 0.8203 & 19.699 \\
 \hline
@@
 \only<1>{
+All regions
 \begin{figure}[H]
@@
 \only<2>{
+Regions formed by reads aligned to more than 10 unique positions
 \begin{figure}[H]
@@
   \includegraphics[width = \textwidth,page = 3]{../figs/Carroll_mice_for_paper/FoxA1_enrichment.pdf}
+\end{figure}
+}
+
+\only<3>{
+Regions formed by reads aligned to more than 30 unique positions
+\begin{figure}[H]
+  \centering
+  \includegraphics[width = \textwidth,page = 4]{../figs/Carroll_mice_for_paper/FoxA1_enrichment.pdf}
 \end{figure}
@@
 \only<1>{
+Calculate coverage for all regions
 \begin{figure}[H]
@@
 \only<2>{
+Calculate $\mbox{local-SCC}$ for regions
 \begin{figure}[H]
@@
 \only<3>{
+Fit local polynomial regression for the $\mbox{local-SCC}$ 
+\begin{figure}[H]
+  \centering
+  \includegraphics[width = \textwidth,page = 3]{../figs/for_paper/local_SCC_separated.pdf}
+\end{figure}
+}
+
+\only<4>{
 \begin{align}
@@
 
-\only<3>{
+\only<4>{
 \begin{columns}
@@
 
-\only<4>{
+\only<5>{
   \begin{figure}[H]
@@
 }
+
 \end{frame}
@@
 
-\begin{frame}
-  \frametitle{Structure of the experiment}
-
+\begin{frame}[plain]
+  \frametitle{Data structure}
+Data available for $\sig$:
   \begin{figure}[H]
     \centering
-
+    \includegraphics[width = \textwidth]{../figs/for_paper/chip_explanation.pdf}
   \end{figure}
@@
 \item $\delta$ measures the average distance of reads to their
-  respective binding sites
+  respective binding sites.
 \item $\sigma$ measures the dispersion of reads around their
-  respective binding sites
+  respective binding sites.
+\item Good news !! the model is reflecting the truth.
 \end{itemize}
@@
 }
-\only<1>{
-{\small
-  \begin{itemize}
-  \item ChIP-exo and PET ChIP-Seq are comparable and outperform SET
-    ChIP-Seq
-  \end{itemize}
-}}
 
@@
 }
+
+\only<2>{
+{\small
+  \begin{itemize}
+  \item ChIP-exo and PET ChIP-Seq are comparable and outperform SET
+    ChIP-Seq
+  \end{itemize}
+}}
 
@@
     outperforms PET and SET ChIP-Seq.
+  \item dPeak outperforms other algorithms in resolution.
   \end{itemize}
@@
   want to add a QC measure to the pipeline based on them.
-\item We want to assess if ChIP-Nexus library complexity is actually higher than ChIP-exo's by using the $\mbox{local-NSC}$.
+\item We want to assess if ChIP-Nexus library complexity is actually
+  higher than ChIP-exo's by using the $\mbox{local-NSC}$.
 \item We have been studying E. Coli's transcription initiation
-  complexes with PET ChIP-Seq, being able to label regions as open or
-  closed complexes. We want to improve this analysis by using ChIP-exo
-  data, and hopefully detecting intermediate step between this two
-  states.
+  complexes with PET ChIP-Seq. We want to improve this analysis by
+  using ChIP-exo data.
 \item Find a optimal strategy for labeling enhancer out of a

### 2015-11-30 rmarkdown/slides.tex
added diagrams to slides

@@
 \end{figure}
+
+{\scriptsize
+\begin{itemize}
+\item 
+\end{itemize}
+}
 

### 2015-12-01 rmarkdown/slides.Rnw
finished slides for prelim

@@
 
-\section{ChIP-exo Data Exploration and Qualiy Control}
+\section{ChIP-exo Data Exploration and Quality Control}
 
@@
 \item Number of reads. Self-explanatory, the higher the better
-}
+  (usually).}
   \item<2-> Number of reads.
@@
   {\color{RoyalBlue}EXACTLY} one reads maps and number of pos. to
-  which {\color{RoyalBlue}AT LEAST} one reads maps }
+  which {\color{RoyalBlue}AT LEAST} one reads maps.}
 \only<2>{
@@
   the shift is the fragment length and min. value of the SCC.
+}
+\item<6> Normalized Strand Cross-Correlation. 
+\only<6>{
+\begin{figure}[H]
+  \centering
+  \includegraphics[width = \textwidth]{../figs/for_paper/SCC.png}
+\caption{SCC as QC. Landt et al., 2012 \cite{encode_qc}}
+\end{figure}
 }
@@
 \begin{itemize}
-\item A shows that high density regions are similar between ChIP-Seq
+\item A) shows that high density regions are similar between ChIP-Seq
   and ChIP-exo but background regions are not.
@@
 \begin{itemize}
-\item There is a \emph{``phantom peak''} at read length.
-\item In ChIP-Seq SCC is maximized at the unobserved fragment length.
+\item There is a \emph{``phantom peak''} at read length for ChIP-Seq (SE).
+\item In ChIP-Seq, SCC is maximized at the unobserved fragment length.
 \item In ChIP-exo, the \emph{``phantom peak''} and the fragment length
@@
   \end{enumerate}
-\end{frame}
-
- % \begin{itemize}
-  % \item $\mbox{local-NSC}$
-  % \item $\mbox{FSR} = \frac{\text{Nr. of fwd. strand reads in region}}{\text{Nr. of reads in region}}$
-  % \end{itemize}
-
-\begin{frame}
-  \frametitle{ARC vs. URCR exploration}
-\only<1>{
-  \begin{itemize}
-  \item $\mbox{ARC} = \frac{\text{Nr. of reads in the
-        region}}{\text{Width of the region}}$
-  \item $\mbox{URCR} = \frac{\text{Nr. of reads mapped to only one
-        position in the region}}{\text{Nr. of reads in the region}}$
-  \end{itemize}
-}
-\only<2>{
-All regions
-  \begin{figure}[H]
-    \centering
-    \includegraphics[width = .7\textwidth,page = 1]{../figs/for_paper/ARC_vs_URCR_example_npos.pdf}
-  \end{figure}
-}
-
-\only<3>{
-Mapped to more than 10 positions
-  \begin{figure}[H]
-    \centering
-    \includegraphics[width = .7\textwidth,page = 2]{../figs/for_paper/ARC_vs_URCR_example_npos.pdf}
-  \end{figure}
-}
-
-\only<4>{
-Mapped to more than 30 positions
-  \begin{figure}[H]
-    \centering
-    \includegraphics[width = .7\textwidth,page = 3]{../figs/for_paper/ARC_vs_URCR_example_npos.pdf}
-  \end{figure}
-}
-\only<5>{
-Mapped to more than 50 positions
-  \begin{figure}[H]
-    \centering
-    \includegraphics[width = .7\textwidth,page = 4]{../figs/for_paper/ARC_vs_URCR_example_npos.pdf}
-  \end{figure}
-}
-
 \end{frame}
@@
 
-% \only<9-10>{
-%   \begin{figure}[H]
-%     \centering
-%     \includegraphics[width = .7\textwidth,page = 5]{../figs/for_paper/enrichment_example.pdf}
-%   \end{figure}
-% }
-
-% \only<10>{
-% \Put(120,200){\includegraphics[width = .6\textwidth,page = 5]{../figs/for_paper/coverage_example.pdf}}
-% }
-
 \only<9-10>{
@@
 \only<1>{
-All regions
+All regions:
 \begin{figure}[H]
@@
 \only<2>{
-Regions formed by reads aligned to more than 10 unique positions
+Regions formed by reads aligned to more than 10 unique positions:
 \begin{figure}[H]
@@
 \only<3>{
-Regions formed by reads aligned to more than 30 unique positions
+Regions formed by reads aligned to more than 30 unique positions:
 \begin{figure}[H]
@@
 \only<1>{
-Calculate coverage for all regions
+Calculate coverage for all regions:
 \begin{figure}[H]
@@
 \only<2>{
-Calculate $\mbox{local-SCC}$ for regions
+Calculate $\mbox{local-SCC}$ for regions:
 \begin{figure}[H]
@@
 \only<3>{
-Fit local polynomial regression for the $\mbox{local-SCC}$ 
+Fit local polynomial regression for the $\mbox{local-SCC}$ :
 \begin{figure}[H]
@@
 \item $x_\delta$ is the shift.
-\item $y_\delta$ is the local - SCC at shift $x_\delta$
-\item $\epsilon \sim N(0, \sigma^2)$ i.i.d 
+\item $y_\delta$ is the local - SCC at shift $x_\delta$.
+\item $\epsilon \sim N(0, \sigma^2)$ i.i.d .
 \end{itemize}
@@
   \begin{itemize}
-  \item {\color{RoyalBlue} high:} Number of unique positions $ > 100$
-  \item {\color{RoyalBlue} med:} Number of unique positions in $(50,100)$
-  \item {\color{RoyalBlue} low:} Number of unique positions in $(20,50)$
+  \item {\color{RoyalBlue} high:} Number of unique positions $ > 100$.
+  \item {\color{RoyalBlue} med:} Number of unique positions in $(50,100)$.
+  \item {\color{RoyalBlue} low:} Number of unique positions in $(20,50)$.
   \end{itemize}
@@
 \begin{itemize}
-\item 
+\only<1>{
+\item The quantiles of the first replicate, approximate the median
+  faster than the other replicate's quantiles.
+ }
+\only<2>{
+\item B) shows that the majority of the low depth islands are formed
+  exclusively by one strand.
+}
+
 \end{itemize}
@@
   \frametitle{Data structure}
-Data available for $\sig$:
+Available for $\sig$ (courtesy of Prof. Landick's lab):
   \begin{figure}[H]
@@
 
+\only<1>{
+{\small
+  \begin{itemize}
+  \item There are more candidate regions for both ChIP-Seq samples,
+    hence there are also more predicted events.
+  \end{itemize}
+}}
+
 \only<2>{
@@
   \centering
-  \includegraphics[width = .7\textwidth]{../figs/for_paper/algorithm_resolution.pdf}
+  \includegraphics[width = .6\textwidth]{../figs/for_paper/algorithm_resolution.pdf}
 \end{figure}
+
+{\scriptsize
+  \begin{itemize}
+  \item Resolution is defined as the min. absolute distance from the
+    regulonDB annotation to an estimated binding site.
+  \end{itemize}
+}
 
@@
     sample enrichment and library complexity.
-  \item We shown that the ``peak-pair'' assumption doesn't hold well
+  \item We shown that the ``peak-pair'' assumption does not hold well
     in practice, and implemented a visualization capable of detecting
@@
 
-\begin{frame}
+\begin{frame}[label=future]
 \frametitle{Future work}
@@
   ChIP-exo tag counts and both mappability and GC content scores. We
-  want to add a QC measure to the pipeline based on them.
+  want to add a QC measure to the pipeline based on
+  them. \hyperlink{gcmap}{\beamerbutton{GC \& Mapp}}
 \item We want to assess if ChIP-Nexus library complexity is actually
-  higher than ChIP-exo's by using the $\mbox{local-NSC}$.
+  higher than ChIP-exo's by using the
+  $\mbox{local-NSC}$. \hyperlink{nexus}{\beamerbutton{ChIP-Nexus}}
+
 \item We have been studying E. Coli's transcription initiation
   complexes with PET ChIP-Seq. We want to improve this analysis by
-  using ChIP-exo data.
+  using ChIP-exo data.  \hyperlink{tic}{\beamerbutton{TIC}}
 \item Find a optimal strategy for labeling enhancer out of a
@@
 
+\begin{frame}[label=gcmap]
+\frametitle{GC content and Mappability score bias}
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width = .46\textwidth,page = 1]{../figs/for_paper/eukaryotic_bias_CTCF.pdf}
+  \includegraphics[width = .46\textwidth,page = 2]{../figs/for_paper/eukaryotic_bias_CTCF.pdf}
+
+  \begin{itemize}
+  \item ChIP tag counts increase linearly as mappability scores
+    increases.
+  \item ChIP tag counts increase linearly as GC content score
+    increases when GC content is less than 0.6 and then ChIP tag
+    counts decrease as GC content
+    increases. \hyperlink{future}{\beamerbutton{back}}
+
+  \end{itemize}
+
+\end{figure}
+
+\end{frame}
+
+\begin{frame}[plain,label=nexus]
+\frametitle{ChIP-exo and ChIP- Nexus}
+
+\begin{columns}
+\column{.5\textwidth}
+  \begin{figure}[H]
+    \includegraphics[width = .9\textwidth]{../vignettes/exo-protocol.png}
+\caption{ChIP-exo explanation, Mendenhall and Bernstein, 2012 \cite{exo_gb}}
+  \end{figure}
+\column{.5\textwidth}
+
+\begin{itemize}
+\item ChIP-Nexus uses a random barcode for adapter $\#$ 1.
+\item Before step 5 uses self-circularization to attach the random
+  barcode to the 3' end, and uses it instead of adapter $\#$ 2.
+
+\end{itemize}
+  \begin{figure}[H]
+    \centering
+    \includegraphics[width = \textwidth]{../figs/for_paper/nexus.png}
+    \caption{ChIP-Nexus example, He et al., 2015, \cite{chipnexus}
+      \hyperlink{future}{\beamerbutton{back}}}
+  \end{figure}
+
+\end{columns}
+
+\end{frame}
+
+\begin{frame}[label=tic]
+\frametitle{Transcription Initiation Complexes}
+
+  \begin{figure}[H]
+  \centering
+ \includegraphics[width = \textwidth]{/p/keles/ChIPexo/volume5/OpenClosedComplexes/figs/filtered/heatmaps_centroid.png}
+ \caption{Centered and normalized heatmaps for PE ChIP-Seq peaks of
+   $\sig$ and $\beta'_f$ under rif0 and rif20. The regions are
+   organized by the centroids clustering of the $\beta'_f$-rif0
+   profiles. \hyperlink{future}{\beamerbutton{back}}}
+  \label{fig:hm1}
+\end{figure}
+
+\end{frame}
+
 \end{document}

### 2016-02-15 rmarkdown/paper.Rnw
New paper for ChIP-exo QC

% \VignetteEngine{knitr::knitr}
\documentclass{bmcart}

\usepackage{xcolor}
\usepackage{url}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows}
\usepackage{float}
\usepackage{verbatim}
\usepackage{hyperref}
\hypersetup{
     colorlinks   = true,
     citecolor    = black
}
\hypersetup{linkcolor=black}
% \hypersetup{linkcolor=black}
% \hypersetup{linkcolor=black}

% \usepackage{authblk}
\usepackage[utf8]{inputenc} %unicode support

\newcommand{\sig}{\sigma^{70}}

\begin{document}

%\maketitle

\begin{frontmatter}

\begin{fmbox}
\dochead{Draft}

\title{High Resolution Identification of Protein-DNA Binding
  Events and Quality Control for ChIP-exo data}

\author[
   addressref={aff6},                   % id's of addresses, e.g. {aff1,aff2}
   email={chungd@musc.edu}   % email address
]{\inits{DC}\fnm{Dongjun} \snm{Chung}}
\author[
   addressref={aff1},
   email={welch@stat.wisc.edu}
]{\inits{RW}\fnm{Rene} \snm{Welch}}
\author[
   addressref={aff3},
   email={ong@cs.wisc.edu}
]{\inits{IO}\fnm{Irene} \snm{Ong}}
\author[
   addressref={aff3,aff4},
   email={jagrass@wisc.edu}
]{\inits{JG}\fnm{Jeffrey} \snm{Grass}}
\author[
   addressref={aff3,aff4,aff5},
   email={landick@bact.wisc.edu}
]{\inits{RL}\fnm{Robert} \snm{Landick}}
\author[
   addressref={aff1,aff2},
   corref={aff1},
   email={keles@stat.wisc.edu}
]{\inits{SK}\fnm{S\"und\"uz} \snm{Kele\c{s}}}

\address[id=aff1]{%                           % unique id
  \orgname{Department of Statistics, University of Wisconsin Madison}, % university, etc
  \street{1300 University Avenue},                     %
  %\postcode{}                                % post or zip code
  \city{Madison},                              % city
  \cny{WI}                                    % country
}

\address[id=aff2]{%
  \orgname{Department of Biostatistics and Medical Informatics, University of Wisconsin Madison},
  \street{600 Highland Avenue},
%  \postcode{24105}
  \city{Madison},
  \cny{WI}
}
\address[id=aff3]{
  \orgname{Great Lakes Bioenergy Research Center, University of Wisconsin Madison},
  \street{1552 University Avenue},
  \city{Madison},
  \cny{WI}
}
\address[id=aff4]{
  \orgname{Department of Biochemistry, University of Wisconsin Madison},
  \street{433 Babcock Drive},
  \city{Madison},
  \cny{WI}
}
\address[id=aff5]{
  \orgname{Department of Bacteriology, University of Wisconsin Madison},
  \street{1550 Linden Drive},
  \city{Madison},
  \cny{WI}
}
\address[id=aff6]{
  \orgname{Department of Public Health Sciences, Medical University of South Carolina},
  \street{135 Cannon Street},
  \city{Charleston},
  \cny{SC}
}

% \begin{artnotes}
% %\note{Sample of title note}     % note to the article
% \note[id=n1]{Equal contributor} % note, connected to author
% \end{artnotes}

\end{fmbox}

\begin{abstractbox}

  \begin{abstract}

  ChIP-exo is a modification of the ChIP-Seq protocol for high
  resolution mapping of transcription factor binding sites. Although
  many aspect of the ChIP-exo data analysis are similar to those of
  ChIP-Seq, ChIP-exo presents a number of unique challenges. We
  present a quality control pipeline that analyzes ChIP-exo's strand
  imbalance and library complexity. Assessment of these biases and
  artifacts are facilitated through diagnostic plots and summary
  statistics calculated over regions of the genome with varying levels
  of coverage.

  We systematically evaluated diverse aspects of ChIP-exo and found
  the following characteristics: First, ChIP-exo's background is quite
  different from ChIP-Seq's. Second, although often assumed in
  ChIP-exo data analysis methods, the ``peak pair'' assumptions does
  not hold locally in actual ChIP-exo data. Third, we for the first
  time compared Paired End (PE) ChIP-Seq with ChIP-exo and found that
  both protocols are comparable in resolutions and sensitivity for
  closely located binding events, but as the distance between binding
  events increases ChIP-exo shows higher sensitivity that PE
  ChIP-Seq. Finally, at fixed sequencing depths, ChIP-exo provides
  higher sensitivity, specificity and spatial resolution than PE
  ChIP-Seq.

  \end{abstract}

\begin{keyword}
  \kwd{ChIP-exo}
  \kwd{QC}
  \kwd{TFBS}
  \kwd{BS identification on High-Res}
\end{keyword}

\end{abstractbox}

\end{frontmatter}

\newpage

\listoffigures

\newpage

\listoftables

\newpage

\section{Background}
\label{sec:intro}

ChIP-exo (Chromatin Immunoprecipitation followed by exonuclease
digestion and next generation sequencing) Rhee and Pugh, 2011
\cite{exo1} is the state-of-the-art experiment developed to attain
single base-pair resolution of protein binding site identification and
it is considered as a powerful alternative to popularly used ChIP-Seq
(Chromatin Immunoprecipitation coupled with next generation
sequencing) assay. ChIP-exo experiments first capture millions of DNA
fragments (150 -250 bp in length) that the protein under study
interacts with using random fragmentation of DNA and a
protein-specific antibody. Then, exonuclease is introduced to trim the
5' end of each DNA fragment to a fixed distance from the bound protein
compared to ChIP-Seq. This step is unique to ChIP-exo and could
potentially provide significantly higher spatial resolution compared
to ChIP-Seq. Finally, high throughput sequencing of a small region (25
to 100 bp) at the 5' end of each fragment generates millions of reads.

While the number of ChIP-exo data keeps increasing,characteristics of
ChIP-exo data are not fully investigated yet. First, DNA libraries
generated by the ChIP-exo protocol seem to be less complex than the
libraries generated by ChIP-Seq (Mahony et al., 2015,
\cite{exo_review}), i.e. the possible number of positions to which the
reads can be aligned has been reduced due to the exonuclease
digestion. Second, although there are roughly the same amount of read
in both strands, locally there may be more reads in one strand than in
the other. To address this challenges, we suggest a collection of
quality control visualizations to understand which of this biases are
present in a ChIP-exo experiment and globally asses the enrichment and
library complexity of a ChIP-exo sample. We gathered ChIP-exo data from
diverse organisms: CTCF factor in human \cite{exo1}; ER factor in
human and FoxA1 factor in mouse (Serandour et al., 2013
\cite{exoillumina}); and generated $\sig$ factor in Escherichia Coli
(\emph{E. Coli}) under aerobic ($ + O_2$) condition, and treated by
rifampicin by 0 and 20 minutes (courtesy of Professor Robert Landick's
lab). %% note to self, we may add more experiments here

Most of current ChIP-Seq quality control (QC) guidelines (Landt et
al., \cite{encode_qc}) may not be applicable on ChIP-exo, additionally
to our knowledge there are not established QC pipelines for
ChIP-exo. Previous ChIP-exo analyses used ChIP-Seq samples to compare
the resolution between experiments (\cite{exo1}, \cite{exo2},
\cite{exoillumina}); Carroll et al., 2014 \cite{carroll.qc} studied
the use of the Strand-Cross Correlation (SCC) (Kharchenko et al., 2008
\cite{strandcc}) and showed that by filtering blacklisted regions the
estimation of the SCC is improved. However, using the SCC may not be
helpful since the peaks that are attained at the read and fragment
lengths are confused in a typical ChIP-exo SCC curve. Additionally
this method requires to know blacklisted regions in advance which may
not be available. In our pipeline we propose an out-the-shelf analysis
that contrasts the enrichment of the experiment against its library
complexity.

In order to obtain the potential benefits of ChIP-exo on protein
binding site identification, it is critical to understand which are
the important characteristics of ChIP-exo data and to use algorithms
that could fully utilize information available in ChIP-exo data. Rhee
and Pugh, 2011 \cite{exo1} discussed that reads in the forward and
reverse strand might construct peak pairs around bound proteins, of
which heights were implicitly assumed to be symmetric. Hence, they
used the ``peak pair method'' that predicts the midpoint of two modes
of peak pairs as potential binding sites. Mace (Wang et al, 20114
\cite{mace}), CexoR (Madrigal, 2015 \cite{cexor}) and peakzilla
(Bardet et al., 2013 \cite{peakzilla}), recently developed ChIP-exo
data analysis methods are also based on this peak pair
assumption. However, appropriateness of such assumption was not fully
evaluated in the literature yet. Furthermore, it is still unknown
which factors could affect protein binding site identification using
ChIP-exo data. In order to address this problem, we investigated
various aspects of ChIP-exo data by contrasting them with their
respective ChIP-Seq experiments.

Currently, research on statistical methods for ChIP-exo data is still
in its very early stage. Although many methods have been proposed to
identify protein binding sites from ChIP-Seq data (reviewed by
Wilbanks and Facciotti, 2012 \cite{evaluation} and Pepke and Wold,
2009 \cite{computation}), such as MACS (Zhang et al., 2008
\cite{macs}), CisGenome (Ji et al., 2008 \cite{cisgenome}) and MOSAiCS
(Kuan et al., 2009 \cite{mosaics}), these approaches reveal protein
binding sites in lower resolution, i.e., at an interval of hundreds to
thousands of base pairs. Furthermore, they report only one ``mode'' or
``predicted binding location'' per peak. Hence, these methods are not
appropriate to evaluate the potential of ChIP-exo data for high
resolution identification of protein binding sites. More recently,
deconvolution algorithms such as Deconvolution (Lun et al., 2009
\cite{csdeconv}), GEM (Guo et al., 2012 \cite{gem}, an improved
version of Guo et al., 2010 \cite{gps} ) and PICS (Zhang et al., 2010
\cite{pics}) have been proposed to identify binding sites in higher
resolution using ChIP-Seq data. However, most of them are still not
tailored for ChIP-exo and PE and SE ChIP-Seq data in a unified
framework and as a result, currently available methods are not
appropriate for fair comparison between ChIP-exo and ChIP-Seq. To
address these limitations, we developed an improved of dPeak (Chung et
al., 2013 \cite{dpeak}), a high resolution binding site identification
(deconvolution) algorithm that we previously developed for PE and SE
ChIP-Seq data, so that it can also handle ChIP-exo data. The dPeak
algorithm implements a probabilistic model that accurately describes
the ChIP-exo and ChIP-Seq data generation process.

In this work we demonstrate that the ``peak pair'' assumption of Rhee
and Pugh, 2013 \cite{exo2} does not hold well in real ChIP-exo
data. Furthermore, we found that when we analyze ChIP-exo data from
eukaryotic genomes, it is important to consider sequence biases
inherent to ChIP-exo data, such as mappability and GC content in order
to improve sensitivity and specificity of binding site
identification. We evaluated several methods to identify binding events
and dPeak performs competitively respect to GEM and MACE when
analyzing ChIP-exo data. More importantly, when comparable number of
reads is used for both ChIP-exo and ChIP-Seq , dPeak coupled with
ChIP-exo data provides resolution comparable to PE ChIP-Seq and both
significantly improve the resolution of protein binding identification
compared to SE-based analysis with any of the available methods.

\section{Results and discussion}
\label{sec:results}

\subsection{Deeply sequence E. Coli $\sig$ ChIP-exo and ChIP-Seq data}

$\sig$ factor is a transcription initiation factor of housekeeping
genesin E. Coli. In this organism's genomes, many promoters contain
multiple transcription start sites (TSS) and these TSS are often
closely spaced (10 $\sim$ 150 bp). These closely spaced binding sites
are considered to be multiple ``switches'' that differentially
regulate gene expression under diverse growth conditions
\cite{regulondb}. Therefore, investigation of ChIP-exo's potential for
identification and differentiation of closely spaced binding sites are
invaluable for elucidating the transcriptional networks of prokaryotic
genomes.

\subsection{Current ChIP-Seq guidelines and quality metrics on ChIP-exo data}

We calculated the most commonly used quality indicators (list in table
\ref{tab:qc}) for ChIP-Seq data using all the ChIP-exo data we
gathered. The PCR Bottleneck Coefficient (PBC) and the Normalized
Strand Cross-Correlation (NSC) are calculated as in \cite{encode_qc}.
Additionally, we calculated the Forward Strand Ratio (FSR) to comped
the number of fragments sequenced from each strand. We can observe
that the PBC is low, hence it would be incorrectly suggested to repeat
the experiment.

\begin{table}[H]
  \centering
  \caption{Usual quality control indicators applied to the gathered ChIP-exo samples. 
    PBC stands for PCR Bottleneck Coefficient, NSC for Normalized Strand Cross-Correlation and
    FSR for Forward Strand Ratio. We ommited the Relative Strand Cross-Correlation (RSC) because
    a typical ChIP-exo experiment is not accompanied by an input file.}
  \label{tab:qc}
\end{table}

\bibliographystyle{bmc-mathphys} % Style BST file (bmc-mathphys, vancouver, spbasic).
\bibliography{chip_exo_paper}

% \nocite{exo_gb}
%  \nocite{maplot1}
% \nocite{maplot2}
% \nocite{chipbeyond}
% \nocite{esl}

\end{document}

### 2016-02-25 rmarkdown/paper.Rnw
most of the paper done, methods missing + conclusions

@@
 % \hypersetup{linkcolor=black}
+
+\usepackage{wrapfig}
 
@@
 \end{frontmatter}
-
-\newpage
-
-\listoffigures
-
-\newpage
-
-\listoftables
 
@@
 
-We calculated the most commonly used quality indicators (list in table
-\ref{tab:qc}) for ChIP-Seq data using all the ChIP-exo data we
-gathered. The PCR Bottleneck Coefficient (PBC) and the Normalized
+We started our exploration by investigating the suitability of the
+current state-of-the-art QC pipelines for ChIP-Seq for ChIP-exo. Table
+\ref{tab:qc} contains measures that are commonly calculated for
+ChIP-Seq samples: PCR Bottleneck Coefficient (PBC) and the Normalized
 Strand Cross-Correlation (NSC) are calculated as in \cite{encode_qc}.
 Additionally, we calculated the Forward Strand Ratio (FSR) to comped
-the number of fragments sequenced from each strand. We can observe
-that the PBC is low, hence it would be incorrectly suggested to repeat
-the experiment.
+the number of fragments sequenced from each strand.
 
-\begin{table}[H]
+%% We can observe that the PBC is low, hence it would be incorrectly
+%% suggested to repeat the experiment.
+
+\begin{table}[h!]
+%% need to fix format
   \centering
-  \caption{Usual quality control indicators applied to the gathered ChIP-exo samples. 
-    PBC stands for PCR Bottleneck Coefficient, NSC for Normalized Strand Cross-Correlation and
-    FSR for Forward Strand Ratio. We ommited the Relative Strand Cross-Correlation (RSC) because
-    a typical ChIP-exo experiment is not accompanied by an input file.}
-  \label{tab:qc}
+<<table1,include = TRUE,echo = FALSE,eval = TRUE>>=
+  load("../data/paper_tables/Landick_rif_summary.RData")
+  load("../data/paper_tables/Landick_aero_summary.RData")
+  load("../data/paper_tables/pugh_ctcf.RData")
+  load("../data/paper_tables/Carroll_mm9_FoxA1.RData")
+  load("../data/paper_tables/Meijsing_hg19_GR.RData")
+  load("../data/paper_tables/Carroll_hg19_ER.RData")
+  do.call(rbind,list(landick_aero,
+                     landick_rif,
+                     pugh_ctcf,
+                     carroll_mm9_FoxA1,
+                     meijing_hg19_GR,
+                     carroll_hg19_ER))
+@ 
+\caption{Usual quality control indicators applied to the gathered
+  ChIP-exo samples.  PBC stands for PCR Bottleneck Coefficient (0 -
+  0.5 is severe bottlenecking, 0.5 - 0.8 is moderate bottlenecking,
+  0.8 - 0.9 is mild bottlenecking, while 0.9 - 1 is no bottlenecking),
+  NSC for Normalized Strand Cross-Correlation and FSR for Forward
+  Strand Ratio. We omitted the Relative Strand Cross-Correlation (RSC)
+  because a typical ChIP-exo experiment is not accompanied by an input
+  file. $\sig$ samples were given by Dr. Landick's Lab, FoxA1 and ER
+  samples are from Serandour et al., 2013 \cite{exoillumina} and CTCF
+  sample is from Rhee and Pugh, 2011 \cite{exo1}}
+  \label{tab:qc} %% need to make this table
 \end{table}
+
+It is of special importance to notice that the values of PBC are quite
+low, hence it would be necessary to repeat these experiments by
+following ChIP-seq guidelines. On the other hand, in a good ChIP-Seq
+experiment the SCC is maximized at the fragment length, and a
+``phantom peak'' would be also present at the experiment's read length
+(\cite{encode_qc} and \cite{strandcc}). However, in ChIP-exo's case
+these two peaks are confounded. Hence an enrichment measure based in
+SCC such as the NSC is harder to interpret. In figure
+\ref{fig:scc_exo} we observe that both local maxima are hard to
+differentiate in the SCC curves for the $\sig$ ChIP-exo datasets used to
+calculate the NSC values in table \ref{tab:qc}.
+
+% % According to ENCODE's quality metrics, NSC values close
+% % to one corresponds to low enrichment samples or bad quality datasets
+
+\subsection{Comparison with ChIP-Seq data}
+
+<<comp_param , include = FALSE,echo = FALSE,eval = TRUE>>=
+
+  bin_size <- 150
+  frag_len <- 150
+
+@
+
+We first compared various factors that could affect binding site
+identification between ChIP-exo and ChIP-Seq data. In order to compare
+distribution of signal and background between ChIP-exo and ChIP-Seq
+data, we calculated ChIP tag counts across the genome by counting the
+number of reads mapping to each of \Sexpr{bin_size} non-overlapping
+window after extending reads by \Sexpr{frag_len} to their 3' end
+directions. ChIP tag counts in ChIP-exo data were linearly related to
+ChIP tag counts in ChIP-Seq data for the regions with high ChIP tag
+counts (Figure \ref{fig:comp}A). This implies that signals for
+potential binding sites are well reproducible between ChIP-exo and
+ChIP-Seq data. On the other hand, there was clear difference in the
+background distribution between them. In ChIP-Seq data reads were
+almost uniformly distributed over background (non-binding) regions and
+the ChIP tag counts in there regions were significantly larger than
+zero. In contrast, in ChIP-exo data, there was larger variation in
+ChIP tag counts among background regions and ChIP tag counts were much
+lower in these regions compared to ChIP-Seq data. There were also
+large proportion of regions without any read in ChIP-exo data. These
+results indicate that for ChIP-exo data a much smaller portion of the
+genome is expected to be background.
+
+We next evaluated the ``peak pair'' assumption from Rhee and Pugh,
+2011 \cite{exo1}, i.e. a peak of reads in the forward strand is
+usually paired with a peak of reads in the reverse strand that is
+located in the other site of the binding site. Wang et al., 2014
+\cite{mace}, Madrigal 2015 \cite{cexor} and Bardet et al., 2013
+\cite{peakzilla} proposed method rely in this assumption. In order to
+evaluate this assumption, we reviewed the proportion of reads in the
+forward strand in candidate regions (i.e. regions with at least one
+binding site) in $\sig$ ChIP-exo data. We found that strands of reads
+were much less balanced in ChIP-exo data than in ChIP-Seq data in
+these regions with potential binding sites (Fig. \ref{fig:comp}B) and
+this indicates that the peak pair assumption might not hold in real
+ChIP-exo data.
+
+We evaluated ChIP-exo data for CTCF factor from human genome
+\cite{exo1} to investigate issues specific to eukaryotic genomes for
+binding sites identification. Figures \ref{fig:comp}C and
+\ref{fig:comp}D display the bin-level average read counts against
+mappability and GC content. Each data point is obtained by averaging
+the read counts across bins with the same mappability of GC
+content. These results indicate that binding site identification in
+ChIP-exo sample might also benefit from the use of methods that take
+into account of apparent sequence biases such as mappability and GC
+content.
+
+\subsection{ChIP-exo Quality Control Pipeline}
+
+Figure \ref{fig:qcdiagram} shows a flowchart for the ChIP-exo QC
+pipeline. Which basically partitions the genome by keeping the
+non-digested ChIP-exo regions. Then, for each region calculates a
+series of summary statistics. Finally it creates several
+visualizations designed to assess the quality level of ChIP-exo
+sample.
+
+\subsubsection{Enrichment and library complexity in ChIP-exo data}
+
+In ChIP-exo experiments, the background is often digested by the
+exonuclease enzyme, therefore to determine the sample's quality is
+necessary to address the balance between the enrichment and library
+complexity of an experiment. To diagnose this, we considered the
+Average Read Coverage (ARC) and the Unique Read Coverage Rate (URCR)
+which are defined as follows:
+
+\begin{align*}
+  \text{ARC} &= \dfrac{\text{Nr. of reads in the region}}{\text{Width of the region}} \\
+  \text{URCR} &= \dfrac{\text{Nr. of reads in the region mapped to exactly one position}}{\text{Nr. of reads in the region}} 
+\end{align*}
+
+Using the FoxA1 in mouse liver cell lines from \cite{exoillumina} and
+these two quantities, the relationship between library complexity and
+experiment enrichment was explored. In figure \ref{fig:enrich}A we can
+observe both statistics for the regions that are common for the three
+replicates. This figure shows typical patterns for ChIP-exo
+experiment, there are two strong arms: The one on the left with low
+ARC and varying URCR corresponds to ChIP-exo's background, usually
+regions composed by scattered reads that were not digested during the
+exonuclease step; and the one on the right that corresponds to regions
+that are usually enriched, for these regions the URCR measures how the
+fragments are allocated into the possible positions in a region. Hence
+we would expect the first replicate to have have more enriched regions
+and the third replicate's library complexity to be lower than the
+other two replicates library complexities. To verify this statements,
+we extracted the sequences around high confidence binding events and
+look for the FoxA1 motif using FIMO \cite{fimo}. Figure
+\ref{fig:enrich}B shows the number of candidate regions, which shows
+that the first replicate is being allocated into more enriched regions
+than the other ones. Figure \ref{fig:enrich}C shows that for the first
+and third replicates, the FoxA1 motif is being detected in roughly the
+same proportion of sites, and finally in \ref{fig:enrich}D we observe
+that the first and third replicate can detect the FoxA1 motif with the
+same significance, while the second replicate does not.
+
+\subsubsection{Strand imbalance in ChIP-exo data}
+
+The strand imbalance assessment in based in the observation that the
+enriched regions usually are composed of a higher quantity of reads,
+therefore we examined the FSR (defined as the ratio of number of
+forward stranded reads divided by the total number of reads in a given
+region) as the regions with lower depth are being filtered out. This
+indicator is of particular importance, since several methods rely on
+the ``peak-pair'' assumption. In table \ref{tab:qc}, we calculated the
+global FSR and noticed that for all experiments is roughly $0.5$,
+which means there are roughly the same amount of reads in both
+strands. However figure \ref{fig:strand} shows that the global FSR
+does not represent the experiment's local strand imbalance, hence the
+``peak pair'' assumption may not hold well in real ChIP-exo data.
+
+In order to asses the strand imbalance we created the visualization
+shown in figure \ref{fig:strand}: Figure \ref{fig:strand}A presents
+the FSR's behavior as the lower depth regions are being filtered out,
+while B) shows which percentage of the regions are composed by reads
+in both strand or only one (forward or backward). In a good data set,
+it would be expected that all quantiles shown to be quickly converging
+towards the median (in panel A) or the regions composed of reads in
+one strand being made of few fragments (in panel B). For each
+replicate, we divided the partitioned regions by asking whether they
+overlap a set of high quality ChIP-exo peaks, and then we tested if
+the strand imbalance's distribution were the same for both
+classes. For regions composed by a higher amount of reads, is is
+harder to distinguish their peaks by considering only the strand
+imbalance, and in an experiment with lower quality it is harder to
+make this distinction.
+
+\subsection{Comparison with ChIP-Seq data using dPeak}
+
+<<comp_param,include = FALSE,echo = FALSE, eval = TRUE>>=
+
+  id_dist <- 20
+  ext <- 15
+
+@
+
+Figure \ref{fig:reso_all} shows different comparisons among ChIP-exo,
+PE ChIP-Seq and SE ChIP-Seq. A RegulonDB annotation (Salgado et al,
+2012 \cite{regulondb}) was considered identified if the distance
+between it and a dPeak binding site estimate was at most of
+$\Sexpr{id_dist}$ bp. That way, the sensitivity is defined as the
+proportion of RegulonDB annotations identified in a peak and the
+resolution is defined as the minimum distance between a RegulonDB
+annotation and the dPeak binding site estimates. The left panel of
+figure \ref{fig:reso_all} shows that the sensitivity increases as the
+mean distance between binding events increases. Despite the when the
+binding events in a peak are closer to each other, both ChIP-exo and
+PE ChIP-Seq are comparable, as the distance increases ChIP-exo
+identifies a higher proportion of the RegulonDB annotations;
+additionally SE ChIP-Seq is significantly less sensitive than both
+ChIP-exo and PE ChIP-Seq. The right panel shows that ChIP-exo and PE
+ChIP-Seq are comparable in resolution, while both protocols
+significantly outperform SE ChIP-Seq.
+
+\subsection{Systematic comparison of ChIP-Seq vs ChIP-exo under varying sequencing depth}
+
+Previously, ChIP-exo and SE ChIP-Seq have been compared at a fixed
+depth level in the literature, but this comparisons did not include PE
+ChIP-Seq. Hence, we sampled a fixed amount of reads for each of the
+ChIP-exo, PE ChIP-Seq and SE ChIP-Seq datasets of the $\sig$
+samples. For each sampled dataset we applied out lower-to-higher
+resolution pipeline by calling peaks with MOSAiCS \cite{mosaics} and
+then deconvolving the binding events by using dPeak \cite{dpeak}. For
+the ChIP-exo datasets we called peaks by using the GC-content and
+mappability models with MOSAiCS, and for the ChIP-Seq datasets we used
+their respective Input samples.
+
+Figures \ref{fig:design} shows the behavior of each data type in
+$\sig$ experiment under aerobic condition when their depth is
+fixed. It is remarkable that even when the number of candidate peaks
+or the number of predicted events is lower for ChIP-exo, it
+outperforms both PE and SE ChIP-Seq in number of identified targets
+and resolution.
+
+This may suggest that with ChIP-exo less positive peaks are being
+called and that when the targets are being identified, dPeak estimates
+binding locations closer to the true location. Additionally, we can
+see that as the read depth increases, all four indicator seem to
+stabilize and hit a plateau, which may indicate that with ChIP-exo a
+smaller amount of reads is necessary to identify a higher number of
+targets, but it may be also possible that this is an artifact occurring
+due to ChIP-exo's lower library complexity. Additionally, it is worth
+noting that for PE ChIP-Seq we sampled both ends of the fragment,
+hence for each sequencing depth we are sampling the half amount of
+pairs for PE ChIP-Seq than for ChIP-exo or SE ChIP-Seq.
+
+Figure \ref{fig:saturation_rif} shows an analogous analysis but using
+the $\sig$ replicates with and without rif treatment. The left, middle
+and right columns shows the fixed depth against the number of
+predicted events, identified targets and resolution being compared at
+a fixed depth level. The behavior of this quantities seems to be
+opposite to the one as in figure \ref{fig:design}, hence we used the
+ChIP-exo QC pipeline in the fixed depth ChIP-exo experiments. Figure
+\ref{fig:exoQC_sat_aero} shows hexbin of ARC vs URCR for several fixed
+sample sizes, as the fixed depth increases the two arm pattern becomes
+more distinctive while for lower depth, it seems that the majority of
+the sampled reads were aligned to enriched regions. On the other hand
+in figures \ref{fig:exoQC_sat1} to \ref{fig:exoQC_sat4}, we used the
+ChIP-exo QC pipeline on the samples that are outperformed by PE and SE
+ChIP-Seq. For a fixed low depth, we can see that the majority of the
+reads are being aligned to non-enriched regions since the vertical arm
+seems to be stronger for all 2 conditions and replicates; for higher
+depth we can see low URCR values being predominant, which indicates
+that the majority of the regions being formed by few positions with a
+higher read concentration. In low complexity regions, the reads are
+being aligned to fewer positions but there is no control over the
+amount of reads mapped. Hence, those regions are more likely to being
+strand-imbalance which in turn may bias the binding site estimate and
+therefore decrease the number of identified targets or increase an
+experiments resolution.
+
+% In supplement figures \ref{fig:exoQC_sat1} to \ref{fig:exoQC_sat4}, we
+% repeat the same analysis for both replicates under the two rif
+% conditions.
+
+\subsection{dPeak outperforms competing methods in discovering closely
+  spaced binding events from ChIP-exo and ChIP-Seq data}
+
+Figure \ref{fig:methods_comp} compares the resolution defined as the
+minimum distance between a RegulonDB annotation and a binding site
+predicted by either Peakzilla \cite{peakzilla}, MACE \cite{mace}, GEM
+\cite{gem} or dPeak \cite{dpeak}. In a good dataset such as both of
+the ChIP-exo experiments under aerobic (panel 933) or anaerobic (panel
+931) conditions, all the methods are comparable in resolution, and
+dPeak slightly outperforms the rest. On the other hand, for
+experiments with low library complexity the resolution calculated with
+dPeak's predictions is smaller than both Mace and Peakzilla, but it is
+greater than Gem. This may be due to the fact that Gem uses sequence
+information in addition to the 5' ends counts per coordinate that
+dPeak uses. \footnote{For here we may probably use only 933 as part of
+  the main article and keep the rest for the supplement}
+
+\section{Methods}
+\label{sec:methods}
 
@@
 \bibliography{chip_exo_paper}
+
+\newpage
+
+\section{Figures}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width = .8\textwidth]{/p/keles/ChIPexo/volume3/ChIPexo/figs/for_paper/EColi_strand_cross_corr.pdf}
+  \caption{SCC curves for $\sig$ samples. The ``phantom peak'' and the
+    summit that corresponds to the read and fragment length
+    respectively are confounded due to the exonuclease digestion.}
+  \label{fig:scc_exo}
+\end{figure}
+
+\newpage
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width = .46\textwidth,page = 3 ]{../figs/for_paper/ChIPseqPET_ChIPexo_tagCount_comparison.pdf}
+  \includegraphics[width = .46\textwidth]{../figs/for_paper/forward_strand_ratio_comp_old.pdf}
+\includegraphics[width = .46\textwidth,page = 1]{../figs/for_paper/eukaryotic_bias_CTCF.pdf}
+  \includegraphics[width = .46\textwidth,page = 2]{../figs/for_paper/eukaryotic_bias_CTCF.pdf}
+  \caption{ A) Hexbin plot of PE ChIP-Seq bin counts VS ChIP-exo bin
+    counts. B) Forward Strand Ratio densities for SE ChIP-Seq and
+    ChIP-exo peaks. C) ChIP tag counts increase linearly as
+    mappability scores increases. D) ChIP tag counts increase linearly
+    as GC content score increases when GC content is less than 0.6 and
+    then ChIP tag counts decrease as GC content increases. D)
+    Percentage of sites with at least p.value.}
+  \label{fig:comp}
+\end{figure}
+
+\newpage
+
+\begin{figure}[h!]
+  \centering
+  \centering
+  \includegraphics[width = \textwidth]{../figs/for_paper/coverage_diagram2.pdf}
+  \caption{Flowchart of the ChIP-exo quality control pipeline.}
+  \label{fig:qcdiagram}
+\end{figure}
+
+\newpage
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width = .9  \textwidth,page = 1]{../figs/Carroll_mice_for_paper/FoxA1_enrichment.pdf}
+  \newline
+  \includegraphics[width = .3 \textwidth,page = 1]{../figs/FOXA1_mm9_fimo/FOXA1_summaries_by_replicate.pdf}
+  \includegraphics[width = .3 \textwidth,page = 3]{../figs/FOXA1_mm9_fimo/FOXA1_summaries_by_replicate.pdf}
+  \includegraphics[width = .3 \textwidth,page = 1]{../figs/FOXA1_mm9_fimo/FOXA1_log10pval_cumulative_nmotifs.pdf}
+  \caption{Using the mouse-FoxA1 experiment from \cite{exoillumina}:
+    A) Hexbin plots of $\mbox{ARC}$ against $\mbox{URCR}$, in general
+    we can see a slight separation into two strong arms, one
+    corresponds to low $\mbox{ARC}$ and varying $\mbox{URCR}$, and for
+    the other $\mbox{URCR}$ decreases as $\mbox{ARC}$ increases. B)
+    Number of candidate sites for each replicate. C) Percentage of
+    candidate sites where the FoxA1 motif was detected. D) 
+ }
+  \label{fig:enrich}
+\end{figure}
+
+\newpage
+
+\begin{figure}[h!]
+  \centering  
+  \includegraphics[width = .9\textwidth,page = 1]{../figs/Carroll_mice_for_paper/Strand_imbalance.pdf} 
+  \newline
+  \includegraphics[width = .9\textwidth,page = 2]{../figs/Carroll_mice_for_paper/Strand_imbalance.pdf} 
+  \includegraphics[width = .4\textwidth]{../figs/for_paper/Carroll_FSR_depth_VS_pvalWilcoxon.pdf}
+  \caption{Strand imbalance QC plots for the same data as in Figure
+    \ref{fig:enrich}A. A) FSR distribution quantiles as the lower
+    depth regions are being filtered out, all quantiles approach to
+    the median as the lower bound increases. B) Stacked histogram with
+    the proportion of regions that are formed by two strands or only
+    one, in a good sample the single-stranded regions are going to be
+    filtered out quickly as in the middle row. C)
+    $-log_{10}(\text{p.value})$ of testing if the imbalance
+    distributions differs when ChIP-exo regions overlap their peaks.}
+  \label{fig:strand}
+\end{figure}
+
+\newpage
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width = .46\textwidth]{../figs/for_paper/sensitivity_exo_olda_data.pdf}
+  \includegraphics[width = .46\textwidth]{../figs/for_paper/resolution_by_dataset_old_data.pdf}
+   \caption{Comparison of (A) sensitivity and (B) resolution between
+     ChIP-exo and ChIP-Seq data. Sensitivity is defined as the
+     proportion of RegulonDB annotations identified using each
+     data. Resolution is defined as the distance between RegulonDB
+     annotation and its closest prediction.}
+  \label{fig:reso_all}
+\end{figure}
+
+\newpage
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width = .9\textwidth]{../figs/for_paper/saturation_analysis_old.pdf}
+  \caption{\textbf{ChIP-exo, PE ChIP-Seq and SE ChIP-Seq comparison at
+      varying sequencing depths.} Comparison of the number of candidate
+    regions (A), predicted events (B), identified targets (C) and
+    resolution (D) among ChIP-exo, PE ChIP-Seq and SE
+    ChIP-Seq. RegulonDB annotations are considered as a gold
+    standard. A gold standard binding events was deemed identified if
+    a binding event was estimated at a $\pm$ \Sexpr{ext} vicinity of
+    it.}
+  \label{fig:design}
+\end{figure}
+
+\newpage
+
+\begin{figure}[h!]
+  \centering
+\includegraphics[width = .9\textwidth]{../figs/saturation/K12_alignment/sig70_aerobic_enrichment.pdf}
+\caption{Hexbin plots of ARC vs URCR of the $\sig$ ChIP-exo experiment
+  under aerobic condition when 10K to 90K reads are being sampled.}
+  \label{fig:exoQC_sat_aero}
+\end{figure}
+
+\newpage
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width = .9\textwidth]{../figs/for_paper/sig70_methods_comparison_FDR5_topM300.pdf}
+  \caption{Comparison of the resolution between dPeak, Gem, Mace and
+    Peakzilla methods. Resolution is defined as the minimum distance
+    between a RegulonDB annotation an a predicted binding event.}
+  \label{fig:methods_comp}
+\end{figure}
+
+\newpage
+
+\section*{Supplement}
+\label{sec:supp}
+
+\begin{table}[h!]
+%% need to fix format
+  \centering
+<<table_chipseq,include = TRUE,echo = FALSE,eval = TRUE>>=
+  load("../data/paper_tables/Landick_rif_chipseq_summary.RData")
+  do.call(rbind,chipseq)
+@ 
+\caption{Same QC metrics as in table \ref{tab:qc} but applied to
+  Landick's chipseq data of the rif experiment}
+\end{table}
+
+\begin{figure}[h!]
+  \centering % note change to stat_density(geom = "line")
+   \includegraphics[width = .46\textwidth,page = 1]{../figs/for_paper/sigma_delta_old_densities.pdf}
+   \includegraphics[width = .46\textwidth,page = 2]{../figs/for_paper/sigma_delta_old_densities.pdf}   
+  \caption{ C) $\delta$ parameter in dPeak measures average distance
+    of the reads to their respective binding site. In ChIP-exo data,
+    reads were located much closer to the binding site than in SET
+    ChIP-Seq. D) $\sigma$ parameter measure the dispersion of reads
+    around each binding site. In ChIP-exo data, reads showed less
+    variation around the their respective binding sites compared to
+    SET ChIP-Seq.}
+
+\end{figure}
+
+\newpage
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width = .7\textwidth]{../figs/saturation/K12_alignment/PBC_saturation.pdf}
+  \caption{Average PBC (among all seeds) of the sampled ChIP-exo, PE
+    ChIP-Seq and SE ChIP-Seq experiments under the rif treatment
+    conditions used for saturation analysis.}
+  \label{fig:pbc_saturation}
+\end{figure}
+
+\newpage
+
+\begin{figure}[h!] %%% fix this plot as one ggplot, facet_grid(condition ~ variable,scales = "free_y")
+  \centering
+  \begin{itemize}
+  \item Rep-1 and rif-0min:
+  \end{itemize}
+  \includegraphics[width = .3\textwidth,page =2]{../figs/saturation/K12_alignment/Sig70_rep1_rif0_saturation_analysis.pdf}
+  \includegraphics[width = .3\textwidth,page =3]{../figs/saturation/K12_alignment/Sig70_rep1_rif0_saturation_analysis.pdf}
+  \includegraphics[width = .3\textwidth,page =4]{../figs/saturation/K12_alignment/Sig70_rep1_rif0_saturation_analysis.pdf}
+
+  \begin{itemize}
+  \item Rep-1 and rif-20min:
+  \end{itemize}
+
+  \includegraphics[width = .3\textwidth,page =2]{../figs/saturation/K12_alignment/Sig70_rep1_rif20_saturation_analysis.pdf}
+  \includegraphics[width = .3\textwidth,page =3]{../figs/saturation/K12_alignment/Sig70_rep1_rif20_saturation_analysis.pdf}
+  \includegraphics[width = .3\textwidth,page =4]{../figs/saturation/K12_alignment/Sig70_rep1_rif20_saturation_analysis.pdf}
+
+  \begin{itemize}
+  \item Rep-2 and rif-0min:
+  \end{itemize}
+
+  \includegraphics[width = .3\textwidth,page =2]{../figs/saturation/K12_alignment/Sig70_rep2_rif0_saturation_analysis.pdf}
+  \includegraphics[width = .3\textwidth,page =3]{../figs/saturation/K12_alignment/Sig70_rep2_rif0_saturation_analysis.pdf}
+  \includegraphics[width = .3\textwidth,page =4]{../figs/saturation/K12_alignment/Sig70_rep2_rif0_saturation_analysis.pdf}
+
+  \begin{itemize}
+  \item Rep-2 and rif-20min:
+  \end{itemize}
+
+  \includegraphics[width = .3\textwidth,page =2]{../figs/saturation/K12_alignment/Sig70_rep2_rif20_saturation_analysis.pdf}
+  \includegraphics[width = .3\textwidth,page =3]{../figs/saturation/K12_alignment/Sig70_rep2_rif20_saturation_analysis.pdf}
+  \includegraphics[width = .3\textwidth,page =4]{../figs/saturation/K12_alignment/Sig70_rep2_rif20_saturation_analysis.pdf}
+  \caption{\textbf{ChIP-exo, PE ChIP-Seq and SE ChIP-Seq comparison at
+      varying sequencing depth} Comparison of the number of predicted
+    events (left), identified targets (middle) and resolution (right)
+    among ChIP-exo, PE ChIP-Seq and SE ChIP-Seq. RegulonDB annotations
+    are considered as gold standard. A RegulonDB binding events was
+    deemed identified if a binding event was estimated at a $\pm$ \Sexpr{ext}
+    vicinity of it.}
+  \label{fig:saturation_rif}
+\end{figure}
+
+\newpage
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width = .9\textwidth,page = 1]{../figs/saturation/K12_alignment/sig70_rif_enrichment.pdf}
+  \caption{A) ARC vs URCR hexbin plots of Rep-1 and rif-0min from $\sig$ experiment when 100K to 900K reads are being
+    sampled for each panel. }
+  \label{fig:exoQC_sat1}
+\end{figure}
+
+\newpage
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width = .9\textwidth,page = 1]{../figs/saturation/K12_alignment/sig70_rif_enrichment.pdf}
+
+  \caption{ARC vs URCR hexbin plots of Rep-1 and rif-20min from $\sig$ experiment when 100K to 900K reads are being
+    sampled for each panel.}
+  \label{fig:exoQC_sat2}
+\end{figure}
+
+\newpage
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width = .9\textwidth,page = 1]{../figs/saturation/K12_alignment/sig70_rif_enrichment.pdf}
+  \caption{ARC vs URCR hexbin plots of Rep-2 and rif-0min from $\sig$ experiment when 100K to 900K reads are being
+    sampled for each panel.}
+  \label{fig:exoQC_sat3}
+\end{figure}
+
+\newpage
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width = .9\textwidth,page = 1]{../figs/saturation/K12_alignment/sig70_rif_enrichment.pdf}
+  \caption{ARC vs URCR hexbin plots of Rep-2 and rif-20min from $\sig$ experiment when 100K to 900K reads are being
+    sampled for each panel.}
+  \label{fig:exoQC_sat4}
+\end{figure}
+
+\newpage
 

### 2016-02-25 rmarkdown/paper.tex
New paper for ChIP-exo QC
most of the paper done, methods missing + conclusions

--- 2016-02-15 rmarkdown/paper.Rnw
@@
 % \VignetteEngine{knitr::knitr}
-\documentclass{bmcart}
+\documentclass{bmcart}\usepackage[]{graphicx}\usepackage[]{color}
+%% maxwidth is the original width if it is less than linewidth
+%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
+\makeatletter
+\def\maxwidth{ %
+  \ifdim\Gin@nat@width>\linewidth
+    \linewidth
+  \else
+    \Gin@nat@width
+  \fi
+}
+\makeatother
+
+\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
+\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
+\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
+\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
+\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
+\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
+\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
+\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
+\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
+\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
+
+\usepackage{framed}
+\makeatletter
+\newenvironment{kframe}{%
+ \def\at@end@of@kframe{}%
+ \ifinner\ifhmode%
+  \def\at@end@of@kframe{\end{minipage}}%
+  \begin{minipage}{\columnwidth}%
+ \fi\fi%
+ \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
+ \colorbox{shadecolor}{##1}\hskip-\fboxsep
+     % There is no \\@totalrightmargin, so:
+     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
+ \MakeFramed {\advance\hsize-\width
+   \@totalleftmargin\z@ \linewidth\hsize
+   \@setminipage}}%
+ {\par\unskip\endMakeFramed%
+ \at@end@of@kframe}
+\makeatother
+
+\definecolor{shadecolor}{rgb}{.97, .97, .97}
+\definecolor{messagecolor}{rgb}{0, 0, 0}
+\definecolor{warningcolor}{rgb}{1, 0, 1}
+\definecolor{errorcolor}{rgb}{1, 0, 0}
+\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX
+
+\usepackage{alltt}
 
@@
 
+\usepackage{wrapfig}
+
 % \usepackage{authblk}
@@
 
-We calculated the most commonly used quality indicators (list in table
-\ref{tab:qc}) for ChIP-Seq data using all the ChIP-exo data we
-gathered. The PCR Bottleneck Coefficient (PBC) and the Normalized
+We started our exploration by investigating the suitability of the
+current state-of-the-art QC pipelines for ChIP-Seq for ChIP-exo. Table
+\ref{tab:qc} contains measures that are commonly calculated for
+ChIP-Seq samples: PCR Bottleneck Coefficient (PBC) and the Normalized
 Strand Cross-Correlation (NSC) are calculated as in \cite{encode_qc}.
 Additionally, we calculated the Forward Strand Ratio (FSR) to comped
-the number of fragments sequenced from each strand. We can observe
-that the PBC is low, hence it would be incorrectly suggested to repeat
-the experiment.
+the number of fragments sequenced from each strand.
 
-\begin{table}[H]
+%% We can observe that the PBC is low, hence it would be incorrectly
+%% suggested to repeat the experiment.
+
+\begin{table}[h!]
+%% need to fix format
   \centering
-  \caption{Usual quality control indicators applied to the gathered ChIP-exo samples. 
-    PBC stands for PCR Bottleneck Coefficient, NSC for Normalized Strand Cross-Correlation and
-    FSR for Forward Strand Ratio. We ommited the Relative Strand Cross-Correlation (RSC) because
-    a typical ChIP-exo experiment is not accompanied by an input file.}
-  \label{tab:qc}
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{verbatim}
+##                files    nreads        pbc      nsc
+## 1      edsn931_Sig70  13961493 0.13994203 103.4239
+## 2      edsn933_Sig70  14810838 0.16339252 162.8002
+## 3      edsn935_Sig70  16108774 0.13535422 153.5088
+## 4      edsn937_Sig70  13636541 0.15322537 172.5815
+## 5     edsn1311_Sig70    902921 0.26895580 13.76908
+## 6     edsn1314_Sig70   1852124 0.25908459 17.91882
+## 7     edsn1317_Sig70   2104427 0.25841987  29.6083
+## 8     edsn1320_Sig70  11548572 0.15107930 13.08626
+## 9               CTCF  48478450 0.45796889 16.02482
+## 10         ERR336935  22421729 0.10683407 72.04243
+## 11         ERR336942  22210461 0.65622704 21.28197
+## 12         ERR336956  23307557 0.79962169 60.42191
+## 13 IMR90_GR_chip-exo  47443803 0.29785605 8.860176
+## 14  K562_GR_chip-exo 116518000 0.05047634 4.117883
+## 15  U2OS_GR_chip-exo   3255111 0.77142141 10.05883
+## 16         ERR336933   9289835 0.80825680 19.87523
+## 17         ERR336938  12464836 0.82035947 18.72806
+## 18         ERR336950  11041833 0.80241200 21.48509
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+\caption{Usual quality control indicators applied to the gathered
+  ChIP-exo samples.  PBC stands for PCR Bottleneck Coefficient (0 -
+  0.5 is severe bottlenecking, 0.5 - 0.8 is moderate bottlenecking,
+  0.8 - 0.9 is mild bottlenecking, while 0.9 - 1 is no bottlenecking),
+  NSC for Normalized Strand Cross-Correlation and FSR for Forward
+  Strand Ratio. We omitted the Relative Strand Cross-Correlation (RSC)
+  because a typical ChIP-exo experiment is not accompanied by an input
+  file. $\sig$ samples were given by Dr. Landick's Lab, FoxA1 and ER
+  samples are from Serandour et al., 2013 \cite{exoillumina} and CTCF
+  sample is from Rhee and Pugh, 2011 \cite{exo1}}
+  \label{tab:qc} %% need to make this table
 \end{table}
+
+It is of special importance to notice that the values of PBC are quite
+low, hence it would be necessary to repeat these experiments by
+following ChIP-seq guidelines. On the other hand, in a good ChIP-Seq
+experiment the SCC is maximized at the fragment length, and a
+``phantom peak'' would be also present at the experiment's read length
+(\cite{encode_qc} and \cite{strandcc}). However, in ChIP-exo's case
+these two peaks are confounded. Hence an enrichment measure based in
+SCC such as the NSC is harder to interpret. In figure
+\ref{fig:scc_exo} we observe that both local maxima are hard to
+differentiate in the SCC curves for the $\sig$ ChIP-exo datasets used to
+calculate the NSC values in table \ref{tab:qc}.
+
+% % According to ENCODE's quality metrics, NSC values close
+% % to one corresponds to low enrichment samples or bad quality datasets
+
+\subsection{Comparison with ChIP-Seq data}
+
+We first compared various factors that could affect binding site
+identification between ChIP-exo and ChIP-Seq data. In order to compare
+distribution of signal and background between ChIP-exo and ChIP-Seq
+data, we calculated ChIP tag counts across the genome by counting the
+number of reads mapping to each of 150 non-overlapping
+window after extending reads by 150 to their 3' end
+directions. ChIP tag counts in ChIP-exo data were linearly related to
+ChIP tag counts in ChIP-Seq data for the regions with high ChIP tag
+counts (Figure \ref{fig:comp}A). This implies that signals for
+potential binding sites are well reproducible between ChIP-exo and
+ChIP-Seq data. On the other hand, there was clear difference in the
+background distribution between them. In ChIP-Seq data reads were
+almost uniformly distributed over background (non-binding) regions and
+the ChIP tag counts in there regions were significantly larger than
+zero. In contrast, in ChIP-exo data, there was larger variation in
+ChIP tag counts among background regions and ChIP tag counts were much
+lower in these regions compared to ChIP-Seq data. There were also
+large proportion of regions without any read in ChIP-exo data. These
+results indicate that for ChIP-exo data a much smaller portion of the
+genome is expected to be background.
+
+We next evaluated the ``peak pair'' assumption from Rhee and Pugh,
+2011 \cite{exo1}, i.e. a peak of reads in the forward strand is
+usually paired with a peak of reads in the reverse strand that is
+located in the other site of the binding site. Wang et al., 2014
+\cite{mace}, Madrigal 2015 \cite{cexor} and Bardet et al., 2013
+\cite{peakzilla} proposed method rely in this assumption. In order to
+evaluate this assumption, we reviewed the proportion of reads in the
+forward strand in candidate regions (i.e. regions with at least one
+binding site) in $\sig$ ChIP-exo data. We found that strands of reads
+were much less balanced in ChIP-exo data than in ChIP-Seq data in
+these regions with potential binding sites (Fig. \ref{fig:comp}B) and
+this indicates that the peak pair assumption might not hold in real
+ChIP-exo data.
+
+We evaluated ChIP-exo data for CTCF factor from human genome
+\cite{exo1} to investigate issues specific to eukaryotic genomes for
+binding sites identification. Figures \ref{fig:comp}C and
+\ref{fig:comp}D display the bin-level average read counts against
+mappability and GC content. Each data point is obtained by averaging
+the read counts across bins with the same mappability of GC
+content. These results indicate that binding site identification in
+ChIP-exo sample might also benefit from the use of methods that take
+into account of apparent sequence biases such as mappability and GC
+content.
+
+\subsection{ChIP-exo Quality Control Pipeline}
+
+Figure \ref{fig:qcdiagram} shows a flowchart for the ChIP-exo QC
+pipeline. Which basically partitions the genome by keeping the
+non-digested ChIP-exo regions. Then, for each region calculates a
+series of summary statistics. Finally it creates several
+visualizations designed to assess the quality level of ChIP-exo
+sample.
+
+\subsubsection{Enrichment and library complexity in ChIP-exo data}
+
+In ChIP-exo experiments, the background is often digested by the
+exonuclease enzyme, therefore to determine the sample's quality is
+necessary to address the balance between the enrichment and library
+complexity of an experiment. To diagnose this, we considered the
+Average Read Coverage (ARC) and the Unique Read Coverage Rate (URCR)
+which are defined as follows:
+
+\begin{align*}
+  \text{ARC} &= \dfrac{\text{Nr. of reads in the region}}{\text{Width of the region}} \\
+  \text{URCR} &= \dfrac{\text{Nr. of reads in the region mapped to exactly one position}}{\text{Nr. of reads in the region}} 
+\end{align*}
+
+Using the FoxA1 in mouse liver cell lines from \cite{exoillumina} and
+these two quantities, the relationship between library complexity and
+experiment enrichment was explored. In figure \ref{fig:enrich}A we can
+observe both statistics for the regions that are common for the three
+replicates. This figure shows typical patterns for ChIP-exo
+experiment, there are two strong arms: The one on the left with low
+ARC and varying URCR corresponds to ChIP-exo's background, usually
+regions composed by scattered reads that were not digested during the
+exonuclease step; and the one on the right that corresponds to regions
+that are usually enriched, for these regions the URCR measures how the
+fragments are allocated into the possible positions in a region. Hence
+we would expect the first replicate to have have more enriched regions
+and the third replicate's library complexity to be lower than the
+other two replicates library complexities. To verify this statements,
+we extracted the sequences around high confidence binding events and
+look for the FoxA1 motif using FIMO \cite{fimo}. Figure
+\ref{fig:enrich}B shows the number of candidate regions, which shows
+that the first replicate is being allocated into more enriched regions
+than the other ones. Figure \ref{fig:enrich}C shows that for the first
+and third replicates, the FoxA1 motif is being detected in roughly the
+same proportion of sites, and finally in \ref{fig:enrich}D we observe
+that the first and third replicate can detect the FoxA1 motif with the
+same significance, while the second replicate does not.
+
+\subsubsection{Strand imbalance in ChIP-exo data}
+
+The strand imbalance assessment in based in the observation that the
+enriched regions usually are composed of a higher quantity of reads,
+therefore we examined the FSR (defined as the ratio of number of
+forward stranded reads divided by the total number of reads in a given
+region) as the regions with lower depth are being filtered out. This
+indicator is of particular importance, since several methods rely on
+the ``peak-pair'' assumption. In table \ref{tab:qc}, we calculated the
+global FSR and noticed that for all experiments is roughly $0.5$,
+which means there are roughly the same amount of reads in both
+strands. However figure \ref{fig:strand} shows that the global FSR
+does not represent the experiment's local strand imbalance, hence the
+``peak pair'' assumption may not hold well in real ChIP-exo data.
+
+In order to asses the strand imbalance we created the visualization
+shown in figure \ref{fig:strand}: Figure \ref{fig:strand}A presents
+the FSR's behavior as the lower depth regions are being filtered out,
+while B) shows which percentage of the regions are composed by reads
+in both strand or only one (forward or backward). In a good data set,
+it would be expected that all quantiles shown to be quickly converging
+towards the median (in panel A) or the regions composed of reads in
+one strand being made of few fragments (in panel B). For each
+replicate, we divided the partitioned regions by asking whether they
+overlap a set of high quality ChIP-exo peaks, and then we tested if
+the strand imbalance's distribution were the same for both
+classes. For regions composed by a higher amount of reads, is is
+harder to distinguish their peaks by considering only the strand
+imbalance, and in an experiment with lower quality it is harder to
+make this distinction.
+
+\subsection{Comparison with ChIP-Seq data using dPeak}
+
+Figure \ref{fig:reso_all} shows different comparisons among ChIP-exo,
+PE ChIP-Seq and SE ChIP-Seq. A RegulonDB annotation (Salgado et al,
+2012 \cite{regulondb}) was considered identified if the distance
+between it and a dPeak binding site estimate was at most of
+$20$ bp. That way, the sensitivity is defined as the
+proportion of RegulonDB annotations identified in a peak and the
+resolution is defined as the minimum distance between a RegulonDB
+annotation and the dPeak binding site estimates. The left panel of
+figure \ref{fig:reso_all} shows that the sensitivity increases as the
+mean distance between binding events increases. Despite the when the
+binding events in a peak are closer to each other, both ChIP-exo and
+PE ChIP-Seq are comparable, as the distance increases ChIP-exo
+identifies a higher proportion of the RegulonDB annotations;
+additionally SE ChIP-Seq is significantly less sensitive than both
+ChIP-exo and PE ChIP-Seq. The right panel shows that ChIP-exo and PE
+ChIP-Seq are comparable in resolution, while both protocols
+significantly outperform SE ChIP-Seq.
+
+\subsection{Systematic comparison of ChIP-Seq vs ChIP-exo under varying sequencing depth}
+
+Previously, ChIP-exo and SE ChIP-Seq have been compared at a fixed
+depth level in the literature, but this comparisons did not include PE
+ChIP-Seq. Hence, we sampled a fixed amount of reads for each of the
+ChIP-exo, PE ChIP-Seq and SE ChIP-Seq datasets of the $\sig$
+samples. For each sampled dataset we applied out lower-to-higher
+resolution pipeline by calling peaks with MOSAiCS \cite{mosaics} and
+then deconvolving the binding events by using dPeak \cite{dpeak}. For
+the ChIP-exo datasets we called peaks by using the GC-content and
+mappability models with MOSAiCS, and for the ChIP-Seq datasets we used
+their respective Input samples.
+
+Figures \ref{fig:design} shows the behavior of each data type in
+$\sig$ experiment under aerobic condition when their depth is
+fixed. It is remarkable that even when the number of candidate peaks
+or the number of predicted events is lower for ChIP-exo, it
+outperforms both PE and SE ChIP-Seq in number of identified targets
+and resolution.
+
+This may suggest that with ChIP-exo less positive peaks are being
+called and that when the targets are being identified, dPeak estimates
+binding locations closer to the true location. Additionally, we can
+see that as the read depth increases, all four indicator seem to
+stabilize and hit a plateau, which may indicate that with ChIP-exo a
+smaller amount of reads is necessary to identify a higher number of
+targets, but it may be also possible that this is an artifact occurring
+due to ChIP-exo's lower library complexity. Additionally, it is worth
+noting that for PE ChIP-Seq we sampled both ends of the fragment,
+hence for each sequencing depth we are sampling the half amount of
+pairs for PE ChIP-Seq than for ChIP-exo or SE ChIP-Seq.
+
+Figure \ref{fig:saturation_rif} shows an analogous analysis but using
+the $\sig$ replicates with and without rif treatment. The left, middle
+and right columns shows the fixed depth against the number of
+predicted events, identified targets and resolution being compared at
+a fixed depth level. The behavior of this quantities seems to be
+opposite to the one as in figure \ref{fig:design}, hence we used the
+ChIP-exo QC pipeline in the fixed depth ChIP-exo experiments. Figure
+\ref{fig:exoQC_sat_aero} shows hexbin of ARC vs URCR for several fixed
+sample sizes, as the fixed depth increases the two arm pattern becomes
+more distinctive while for lower depth, it seems that the majority of
+the sampled reads were aligned to enriched regions. On the other hand
+in figures \ref{fig:exoQC_sat1} to \ref{fig:exoQC_sat4}, we used the
+ChIP-exo QC pipeline on the samples that are outperformed by PE and SE
+ChIP-Seq. For a fixed low depth, we can see that the majority of the
+reads are being aligned to non-enriched regions since the vertical arm
+seems to be stronger for all 2 conditions and replicates; for higher
+depth we can see low URCR values being predominant, which indicates
+that the majority of the regions being formed by few positions with a
+higher read concentration. In low complexity regions, the reads are
+being aligned to fewer positions but there is no control over the
+amount of reads mapped. Hence, those regions are more likely to being
+strand-imbalance which in turn may bias the binding site estimate and
+therefore decrease the number of identified targets or increase an
+experiments resolution.
+
+% In supplement figures \ref{fig:exoQC_sat1} to \ref{fig:exoQC_sat4}, we
+% repeat the same analysis for both replicates under the two rif
+% conditions.
+
+\subsection{dPeak outperforms competing methods in discovering closely
+  spaced binding events from ChIP-exo and ChIP-Seq data}
+
+Figure \ref{fig:methods_comp} compares the resolution defined as the
+minimum distance between a RegulonDB annotation and a binding site
+predicted by either Peakzilla \cite{peakzilla}, MACE \cite{mace}, GEM
+\cite{gem} or dPeak \cite{dpeak}. In a good dataset such as both of
+the ChIP-exo experiments under aerobic (panel 933) or anaerobic (panel
+931) conditions, all the methods are comparable in resolution, and
+dPeak slightly outperforms the rest. On the other hand, for
+experiments with low library complexity the resolution calculated with
+dPeak's predictions is smaller than both Mace and Peakzilla, but it is
+greater than Gem. This may be due to the fact that Gem uses sequence
+information in addition to the 5' ends counts per coordinate that
+dPeak uses. \footnote{For here we may probably use only 933 as part of
+  the main article and keep the rest for the supplement}
+
+\section{Methods}
+\label{sec:methods}
 
@@
 \bibliography{chip_exo_paper}
+
+\newpage
+
+\section{Figures}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width = .8\textwidth]{/p/keles/ChIPexo/volume3/ChIPexo/figs/for_paper/EColi_strand_cross_corr.pdf}
+  \caption{SCC curves for $\sig$ samples. The ``phantom peak'' and the
+    summit that corresponds to the read and fragment length
+    respectively are confounded due to the exonuclease digestion.}
+  \label{fig:scc_exo}
+\end{figure}
+
+\newpage
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width = .46\textwidth,page = 3 ]{../figs/for_paper/ChIPseqPET_ChIPexo_tagCount_comparison.pdf}
+  \includegraphics[width = .46\textwidth]{../figs/for_paper/forward_strand_ratio_comp_old.pdf}
+\includegraphics[width = .46\textwidth,page = 1]{../figs/for_paper/eukaryotic_bias_CTCF.pdf}
+  \includegraphics[width = .46\textwidth,page = 2]{../figs/for_paper/eukaryotic_bias_CTCF.pdf}
+  \caption{ A) Hexbin plot of PE ChIP-Seq bin counts VS ChIP-exo bin
+    counts. B) Forward Strand Ratio densities for SE ChIP-Seq and
+    ChIP-exo peaks. C) ChIP tag counts increase linearly as
+    mappability scores increases. D) ChIP tag counts increase linearly
+    as GC content score increases when GC content is less than 0.6 and
+    then ChIP tag counts decrease as GC content increases. D)
+    Percentage of sites with at least p.value.}
+  \label{fig:comp}
+\end{figure}
+
+\newpage
+
+\begin{figure}[h!]
+  \centering
+  \centering
+  \includegraphics[width = \textwidth]{../figs/for_paper/coverage_diagram2.pdf}
+  \caption{Flowchart of the ChIP-exo quality control pipeline.}
+  \label{fig:qcdiagram}
+\end{figure}
+
+\newpage
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width = .9  \textwidth,page = 1]{../figs/Carroll_mice_for_paper/FoxA1_enrichment.pdf}
+  \newline
+  \includegraphics[width = .3 \textwidth,page = 1]{../figs/FOXA1_mm9_fimo/FOXA1_summaries_by_replicate.pdf}
+  \includegraphics[width = .3 \textwidth,page = 3]{../figs/FOXA1_mm9_fimo/FOXA1_summaries_by_replicate.pdf}
+  \includegraphics[width = .3 \textwidth,page = 1]{../figs/FOXA1_mm9_fimo/FOXA1_log10pval_cumulative_nmotifs.pdf}
+  \caption{Using the mouse-FoxA1 experiment from \cite{exoillumina}:
+    A) Hexbin plots of $\mbox{ARC}$ against $\mbox{URCR}$, in general
+    we can see a slight separation into two strong arms, one
+    corresponds to low $\mbox{ARC}$ and varying $\mbox{URCR}$, and for
+    the other $\mbox{URCR}$ decreases as $\mbox{ARC}$ increases. B)
+    Number of candidate sites for each replicate. C) Percentage of
+    candidate sites where the FoxA1 motif was detected. D) 
+ }
+  \label{fig:enrich}
+\end{figure}
+
+\newpage
+
+\begin{figure}[h!]
+  \centering  
+  \includegraphics[width = .9\textwidth,page = 1]{../figs/Carroll_mice_for_paper/Strand_imbalance.pdf} 
+  \newline
+  \includegraphics[width = .9\textwidth,page = 2]{../figs/Carroll_mice_for_paper/Strand_imbalance.pdf} 
+  \includegraphics[width = .4\textwidth]{../figs/for_paper/Carroll_FSR_depth_VS_pvalWilcoxon.pdf}
+  \caption{Strand imbalance QC plots for the same data as in Figure
+    \ref{fig:enrich}A. A) FSR distribution quantiles as the lower
+    depth regions are being filtered out, all quantiles approach to
+    the median as the lower bound increases. B) Stacked histogram with
+    the proportion of regions that are formed by two strands or only
+    one, in a good sample the single-stranded regions are going to be
+    filtered out quickly as in the middle row. C)
+    $-log_{10}(\text{p.value})$ of testing if the imbalance
+    distributions differs when ChIP-exo regions overlap their peaks.}
+  \label{fig:strand}
+\end{figure}
+
+\newpage
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width = .46\textwidth]{../figs/for_paper/sensitivity_exo_olda_data.pdf}
+  \includegraphics[width = .46\textwidth]{../figs/for_paper/resolution_by_dataset_old_data.pdf}
+   \caption{Comparison of (A) sensitivity and (B) resolution between
+     ChIP-exo and ChIP-Seq data. Sensitivity is defined as the
+     proportion of RegulonDB annotations identified using each
+     data. Resolution is defined as the distance between RegulonDB
+     annotation and its closest prediction.}
+  \label{fig:reso_all}
+\end{figure}
+
+\newpage
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width = .9\textwidth]{../figs/for_paper/saturation_analysis_old.pdf}
+  \caption{\textbf{ChIP-exo, PE ChIP-Seq and SE ChIP-Seq comparison at
+      varying sequencing depths.} Comparison of the number of candidate
+    regions (A), predicted events (B), identified targets (C) and
+    resolution (D) among ChIP-exo, PE ChIP-Seq and SE
+    ChIP-Seq. RegulonDB annotations are considered as a gold
+    standard. A gold standard binding events was deemed identified if
+    a binding event was estimated at a $\pm$ 15 vicinity of
+    it.}
+  \label{fig:design}
+\end{figure}
+
+\newpage
+
+\begin{figure}[h!]
+  \centering
+\includegraphics[width = .9\textwidth]{../figs/saturation/K12_alignment/sig70_aerobic_enrichment.pdf}
+\caption{Hexbin plots of ARC vs URCR of the $\sig$ ChIP-exo experiment
+  under aerobic condition when 10K to 90K reads are being sampled.}
+  \label{fig:exoQC_sat_aero}
+\end{figure}
+
+\newpage
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width = .9\textwidth]{../figs/for_paper/sig70_methods_comparison_FDR5_topM300.pdf}
+  \caption{Comparison of the resolution between dPeak, Gem, Mace and
+    Peakzilla methods. Resolution is defined as the minimum distance
+    between a RegulonDB annotation an a predicted binding event.}
+  \label{fig:methods_comp}
+\end{figure}
+
+\newpage
+
+\section*{Supplement}
+\label{sec:supp}
+
+\begin{table}[h!]
+%% need to fix format
+  \centering
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{verbatim}
+##                         files   nreads       pbc      nsc
+## pet.1 edsn1396_Sig70.sort.bam 13445022 0.9426179 8.865244
+## pet.2 edsn1398_Sig70.sort.bam 16538920 0.9378843 7.031836
+## pet.3 edsn1400_Sig70.sort.bam 11642722 0.8891744 10.77284
+## pet.4 edsn1402_Sig70.sort.bam 16854026 0.9407020 7.936239
+## set.1 edsn1396_Sig70.sort.bam  6722511 0.6632742  9.01779
+## set.2 edsn1398_Sig70.sort.bam  8269460 0.5594449 7.179539
+## set.3 edsn1400_Sig70.sort.bam  5821361 0.6472382 10.89898
+## set.4 edsn1402_Sig70.sort.bam  8427013 0.5895118 8.124717
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+\caption{Same QC metrics as in table \ref{tab:qc} but applied to
+  Landick's chipseq data of the rif experiment}
+\end{table}
+
+\begin{figure}[h!]
+  \centering % note change to stat_density(geom = "line")
+   \includegraphics[width = .46\textwidth,page = 1]{../figs/for_paper/sigma_delta_old_densities.pdf}
+   \includegraphics[width = .46\textwidth,page = 2]{../figs/for_paper/sigma_delta_old_densities.pdf}   
+  \caption{ C) $\delta$ parameter in dPeak measures average distance
+    of the reads to their respective binding site. In ChIP-exo data,
+    reads were located much closer to the binding site than in SET
+    ChIP-Seq. D) $\sigma$ parameter measure the dispersion of reads
+    around each binding site. In ChIP-exo data, reads showed less
+    variation around the their respective binding sites compared to
+    SET ChIP-Seq.}
+
+\end{figure}
+
+\newpage
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width = .7\textwidth]{../figs/saturation/K12_alignment/PBC_saturation.pdf}
+  \caption{Average PBC (among all seeds) of the sampled ChIP-exo, PE
+    ChIP-Seq and SE ChIP-Seq experiments under the rif treatment
+    conditions used for saturation analysis.}
+  \label{fig:pbc_saturation}
+\end{figure}
+
+\newpage
+
+\begin{figure}[h!] %%% fix this plot as one ggplot, facet_grid(condition ~ variable,scales = "free_y")
+  \centering
+  \begin{itemize}
+  \item Rep-1 and rif-0min:
+  \end{itemize}
+  \includegraphics[width = .3\textwidth,page =2]{../figs/saturation/K12_alignment/Sig70_rep1_rif0_saturation_analysis.pdf}
+  \includegraphics[width = .3\textwidth,page =3]{../figs/saturation/K12_alignment/Sig70_rep1_rif0_saturation_analysis.pdf}
+  \includegraphics[width = .3\textwidth,page =4]{../figs/saturation/K12_alignment/Sig70_rep1_rif0_saturation_analysis.pdf}
+
+  \begin{itemize}
+  \item Rep-1 and rif-20min:
+  \end{itemize}
+
+  \includegraphics[width = .3\textwidth,page =2]{../figs/saturation/K12_alignment/Sig70_rep1_rif20_saturation_analysis.pdf}
+  \includegraphics[width = .3\textwidth,page =3]{../figs/saturation/K12_alignment/Sig70_rep1_rif20_saturation_analysis.pdf}
+  \includegraphics[width = .3\textwidth,page =4]{../figs/saturation/K12_alignment/Sig70_rep1_rif20_saturation_analysis.pdf}
+
+  \begin{itemize}
+  \item Rep-2 and rif-0min:
+  \end{itemize}
+
+  \includegraphics[width = .3\textwidth,page =2]{../figs/saturation/K12_alignment/Sig70_rep2_rif0_saturation_analysis.pdf}
+  \includegraphics[width = .3\textwidth,page =3]{../figs/saturation/K12_alignment/Sig70_rep2_rif0_saturation_analysis.pdf}
+  \includegraphics[width = .3\textwidth,page =4]{../figs/saturation/K12_alignment/Sig70_rep2_rif0_saturation_analysis.pdf}
+
+  \begin{itemize}
+  \item Rep-2 and rif-20min:
+  \end{itemize}
+
+  \includegraphics[width = .3\textwidth,page =2]{../figs/saturation/K12_alignment/Sig70_rep2_rif20_saturation_analysis.pdf}
+  \includegraphics[width = .3\textwidth,page =3]{../figs/saturation/K12_alignment/Sig70_rep2_rif20_saturation_analysis.pdf}
+  \includegraphics[width = .3\textwidth,page =4]{../figs/saturation/K12_alignment/Sig70_rep2_rif20_saturation_analysis.pdf}
+  \caption{\textbf{ChIP-exo, PE ChIP-Seq and SE ChIP-Seq comparison at
+      varying sequencing depth} Comparison of the number of predicted
+    events (left), identified targets (middle) and resolution (right)
+    among ChIP-exo, PE ChIP-Seq and SE ChIP-Seq. RegulonDB annotations
+    are considered as gold standard. A RegulonDB binding events was
+    deemed identified if a binding event was estimated at a $\pm$ 15
+    vicinity of it.}
+  \label{fig:saturation_rif}
+\end{figure}
+
+\newpage
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width = .9\textwidth,page = 1]{../figs/saturation/K12_alignment/sig70_rif_enrichment.pdf}
+  \caption{A) ARC vs URCR hexbin plots of Rep-1 and rif-0min from $\sig$ experiment when 100K to 900K reads are being
+    sampled for each panel. }
+  \label{fig:exoQC_sat1}
+\end{figure}
+
+\newpage
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width = .9\textwidth,page = 1]{../figs/saturation/K12_alignment/sig70_rif_enrichment.pdf}
+
+  \caption{ARC vs URCR hexbin plots of Rep-1 and rif-20min from $\sig$ experiment when 100K to 900K reads are being
+    sampled for each panel.}
+  \label{fig:exoQC_sat2}
+\end{figure}
+
+\newpage
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width = .9\textwidth,page = 1]{../figs/saturation/K12_alignment/sig70_rif_enrichment.pdf}
+  \caption{ARC vs URCR hexbin plots of Rep-2 and rif-0min from $\sig$ experiment when 100K to 900K reads are being
+    sampled for each panel.}
+  \label{fig:exoQC_sat3}
+\end{figure}
+
+\newpage
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width = .9\textwidth,page = 1]{../figs/saturation/K12_alignment/sig70_rif_enrichment.pdf}
+  \caption{ARC vs URCR hexbin plots of Rep-2 and rif-20min from $\sig$ experiment when 100K to 900K reads are being
+    sampled for each panel.}
+  \label{fig:exoQC_sat4}
+\end{figure}
+
+\newpage
 

### 2016-02-28 rmarkdown/paper.Rnw
added conclusions to the draft

@@
 
-\title{High Resolution Identification of Protein-DNA Binding
-  Events and Quality Control for ChIP-exo data}
+\title{Data exploration, quality control and statistical analysis of
+  ChIP-exo experiments}
 
@@
 Strand Cross-Correlation (NSC) are calculated as in \cite{encode_qc}.
-Additionally, we calculated the Forward Strand Ratio (FSR) to comped
+Additionally, we calculated the Forward Strand Ratio (FSR) to compare
 the number of fragments sequenced from each strand.
@@
   \text{ARC} &= \dfrac{\text{Nr. of reads in the region}}{\text{Width of the region}} \\
-  \text{URCR} &= \dfrac{\text{Nr. of reads in the region mapped to exactly one position}}{\text{Nr. of reads in the region}} 
+  \text{URCR} &= \dfrac{\text{Nr. of reads in the region mapped to
+      exactly one position}}{\text{Nr. of reads in the region}}
 \end{align*}
@@
 
-\subsection{Systematic comparison of ChIP-Seq vs ChIP-exo under varying sequencing depth}
+\subsection{Systematic comparison of ChIP-Seq vs ChIP-exo under
+  varying sequencing depth}
 
@@
 dPeak's predictions is smaller than both Mace and Peakzilla, but it is
-greater than Gem. This may be due to the fact that Gem uses sequence
-information in addition to the 5' ends counts per coordinate that
-dPeak uses. \footnote{For here we may probably use only 933 as part of
-  the main article and keep the rest for the supplement}
+greater than Gem. This may be due that the fact that Gem uses sequence
+information in addition to the aligned 5' end counts that dPeak
+uses.\footnote{For here we may probably use only 933 as part of the
+  main article and keep the rest for the supplement}
+
+\section{Conclusions}
+\label{sec:conc}
+
+We made a systematic exploration of several ChIP-exo experiments and
+provide a list of factors that reflect the quality of a ChIP-exo
+experiment and we provide a QC pipeline which is capable of assessing
+the balance between the enrichment and the library complexity of a
+ChIP-exo experiment. Additionally, a set of diagnostics was established
+to assess the an experiment's quality. The QC pipeline only requires a
+set of aligned reads to give a global overview of a ChIP-exo
+experiment, this overview coincides with more elaborate analysis, such
+as motif detection in a set of high quality regions or resolution
+analysis given a set of annotations as gold-standard.
+
+To the extent of our knowledge, we made for the fist time a comparison
+between ChIP-exo and PE ChIP-Seq. Using a set of annotations as
+gold-standard, we showed that both protocols are comparable in
+resolution and that for regions with more than one binding site,
+ChIP-exo is more sensitive than both SE and PE ChIP-Seq. We made a
+systematic comparison between fixed depth ChIP-exo, PE ChIP-Seq and SE
+ChIP-Seq, and we probed that for sufficiently complex libraries,
+ChIP-exo experiments can outperform PE and SE ChIP-Seq in number of
+identified targets and resolution. Using the ChIP-exo QC pipeline, we
+show how to diagnose if the library complexity of ChIP-exo experiment
+is adequate.
 
@@
 \label{sec:methods}
+
+<<methods_param , include = FALSE,echo = FALSE,eval = TRUE>>=
+
+  fdr <- 0.01
+  topM <- 300
+
+@
+
+\subsection*{Growth conditions.}
+
+\subsection*{ChIP-exo experiments.}
+
+\subsection*{Construction of a SE ChIP-Seq experiment.}
+
+For the rif-treatment ChIP-Seq experiments, we sampled SE ChIP-Seq
+experiment from the PE ones by taking one of both ends randomly
+following a $\mbox{Ber}(0.5)$ model.
+
+\subsection*{Saturation analysis of ChIP-exo, PE ChIP-Seq and SE ChIP-Seq.}
+
+\subsection*{ChIP-Seq quality control metrics and Strand
+  Cross-Correlation.}
+
+The statistics in table \ref{tab:qc} and the SCC curves from figure
+\ref{fig:scc_exo} where calculated with the \textbf{ChIPUtils} package
+version 0.99.0, available in
+\url{https://github.com/welch16/ChIPUtils}.
+
+\subsection*{ChIP-exo quality control pipeline.}
+
+We used the R package \textbf{ChIPexoQual} to assess the quality of
+the ChIP-exo datasets by following the steps described in figure
+\ref{fig:qcdiagram}. We used version 1.0, and it is available in
+\url{https://github.com/welch16/ChIPexoqual}.
+
+\subsection*{dPeak analysis of $\sig$ ChIP-exo and ChIP-Seq data.}
+
+\vspace{2 cm}
+
+\subsection*{Method comparison for ChIP-exo.}
+
+We considered dPeak Chung et al., 2013 \cite{dpeak}, GEM Guo et al.,
+2012 \cite{gem}, MACE Wang et al., 2014 \cite{mace} and Peazilla
+Bardet et al., 2013 \cite{peakzilla} for the ChIP-exo data
+analysis. for the dPeak algorithm we used the R package \textbf{dPeak}
+version 2.0.1 which is available from
+\url{https://github.com/dongjunchung/dpeak}. For the GEM algorithm, we
+used it's Java implementation version 2.6 which is available from
+\url{http://groups.csail.mit.edu/cgs/gem/}. For the Mace algorithm, we
+used it Python implementation version 1.2, which is available from
+\url{http://dldcc-web.brc.bcm.edu/lilab/MACE/docs/html/}. For the
+Peakzilla algorithm, we used the version available in
+\url{https://github.com/steinmann/peakzilla}. Candidate regions for
+\textbf{dPeak} were identified for each replicate of ChIP-exo data
+using the \textbf{MOSAiCS} algorithm Kuan et al., 2011 \cite{mosaics}
+(one sample analysis using false discovery rate of \Sexpr{fdr}\%)
+implemented as an R package \textbf{mosaics} version 2.9.7 (available
+from \emph{bioconductor}). We further filtered out candidate regions
+by using the \Sexpr{topM} peaks with higher average ChIP tag count to
+avoid potential false positive based on the exploratory
+analysis. These regions were also explicitly provided to the GEM
+algorithm as candidate regions. Default tuning parameters were used
+during model fitting for all methods.
+
+%% CexoR
 

### 2016-02-28 rmarkdown/paper.tex
added conclusions to the draft

@@
 \label{sec:methods}
+
+\subsection*{Growth conditions.}
+
+\subsection*{ChIP-exo experiments.}
+
+\subsection*{Construction of a SE ChIP-Seq experiment.}
+
+For the rif-treatment ChIP-Seq experiments, we sampled SE ChIP-Seq
+experiment from the PE ones by taking one of both ends randomly
+following a $\mbox{Ber}(0.5)$ model.
+
+\subsection*{Saturation analysis of ChIP-exo, PE ChIP-Seq and SE ChIP-Seq.}
+
+\subsection*{ChIP-Seq quality control metrics and Strand
+  Cross-Correlation.}
+
+The statistics in table \ref{tab:qc} and the SCC curves from figure
+\ref{fig:scc_exo} where calculated with the \textbf{ChIPUtils} package
+version 0.99.0, available in
+\url{https://github.com/welch16/ChIPUtils}.
+
+\subsection*{ChIP-exo quality control pipeline.}
+
+We used the R package \textbf{ChIPexoQual} to assess the quality of
+the ChIP-exo datasets by following the steps described in figure
+\ref{fig:qcdiagram}. We used version 1.0, and it is available in
+\url{https://github.com/welch16/ChIPexoqual}.
+
+\subsection*{dPeak analysis of $\sig$ ChIP-exo and ChIP-Seq data.}
+
+\vspace{2 cm}
+
+\subsection*{Method comparison for ChIP-exo.}
+
+We considered dPeak Chung et al., 2013 \cite{dpeak}, GEM Guo et al.,
+2012 \cite{gem}, MACE Wang et al., 2014 \cite{mace} and Peazilla
+Bardet et al., 2013 \cite{peakzilla} for the ChIP-exo data
+analysis. for the dPeak algorithm we used the R package \textbf{dPeak}
+version 2.0.1 which is available from
+\url{https://github.com/dongjunchung/dpeak}. For the GEM algorithm, we
+used it's Java implementation version 2.6 which is available from
+\url{http://groups.csail.mit.edu/cgs/gem/}. For the Mace algorithm, we
+used it Python implementation version 1.2, which is available from
+\url{http://dldcc-web.brc.bcm.edu/lilab/MACE/docs/html/}. For the
+Peakzilla algorithm, we used the version available in
+\url{https://github.com/steinmann/peakzilla}. Candidate regions for
+\textbf{dPeak} were identified for each replicate of ChIP-exo data
+using the \textbf{MOSAiCS} algorithm Kuan et al., 2011 \cite{mosaics}
+(one sample analysis using false discovery rate of 0.01\%)
+implemented as an R package \textbf{mosaics} version 2.9.7 (available
+from \emph{bioconductor}). We further filtered out candidate regions
+by using the 300 peaks with higher average ChIP tag count to
+avoid potential false positive based on the exploratory
+analysis. These regions were also explicitly provided to the GEM
+algorithm as candidate regions. Default tuning parameters were used
+during model fitting for all methods.
+
+%% CexoR
 

### 2016-02-28 rmarkdown/paper.R

@@ -24,6 +24,12 @@
   ext <- 15
 
 
+## ----methods_param , include = FALSE,echo = FALSE,eval = TRUE------------
+
+  fdr <- 0.01
+  topM <- 300
+
+
 ## ----table_chipseq,include = TRUE,echo = FALSE,eval = TRUE---------------
   load("../data/paper_tables/Landick_rif_chipseq_summary.RData")
   do.call(rbind,chipseq)

### 2016-02-29 rmarkdown/paper.Rnw
finished paper , corrected table 1

@@
     ChIP-Seq, ChIP-exo presents a number of unique challenges. We
-  present a quality control pipeline that analyzes ChIP-exo's strand
-  imbalance and library complexity. Assessment of these biases and
-  artifacts are facilitated through diagnostic plots and summary
-  statistics calculated over regions of the genome with varying levels
-  of coverage.
+    present a quality control pipeline that analyzes a ChIP-exo
+    experiment's strand imbalance, enrichment and library
+    complexity. Assessment of these characteristics are facilitated
+    through diagnostic plots and summary statistics calculated over
+    regions of the genome with varying levels of coverage.
 
     We systematically evaluated diverse aspects of ChIP-exo and found
-  the following characteristics: First, ChIP-exo's background is quite
-  different from ChIP-Seq's. Second, although often assumed in
+    the following characteristics: First, ChIP-exo's background is
+    quite different from ChIP-Seq's. Second, although often assumed in
     ChIP-exo data analysis methods, the ``peak pair'' assumptions does
     not hold locally in actual ChIP-exo data. Third, we for the first
-  time compared Paired End (PE) ChIP-Seq with ChIP-exo and found that
-  both protocols are comparable in resolutions and sensitivity for
-  closely located binding events, but as the distance between binding
-  events increases ChIP-exo shows higher sensitivity that PE
+    time compared Paired End (PE) ChIP-Seq with ChIP-exo and found
+    that both protocols are comparable in resolutions and sensitivity
+    for closely located binding events, but as the distance between
+    binding events increases ChIP-exo shows higher sensitivity that PE
     ChIP-Seq. Finally, at fixed sequencing depths, ChIP-exo provides
@@
 present in a ChIP-exo experiment and globally asses the enrichment and
-library complexity of a ChIP-exo sample. We gathered ChIP-exo data from
-diverse organisms: CTCF factor in human \cite{exo1}; ER factor in
+library complexity of a ChIP-exo sample. We gathered ChIP-exo data
+from diverse organisms: CTCF factor in human \cite{exo1}; ER factor in
 human and FoxA1 factor in mouse (Serandour et al., 2013
-\cite{exoillumina}); and generated $\sig$ factor in Escherichia Coli
-(\emph{E. Coli}) under aerobic ($ + O_2$) condition, and treated by
-rifampicin by 0 and 20 minutes (courtesy of Professor Robert Landick's
-lab). %% note to self, we may add more experiments here
+\cite{exoillumina}); GR in IMR90, K562 and U20S cell lines (Starick et
+al., 2015 \cite{starick15}) and generated $\sig$ factor in Escherichia
+Coli (\emph{E. Coli}) under aerobic ($ + O_2$) condition, and treated
+by rifampicin by 0 and 20 minutes (courtesy of Professor Robert
+Landick's lab). %% note to self, we may add more experiments here
 
@@
 of peak pairs as potential binding sites. Mace (Wang et al, 20114
-\cite{mace}), CexoR (Madrigal, 2015 \cite{cexor}) and peakzilla
+\cite{mace}), CexoR (Madrigal, 2015 \cite{cexor}) and Peakzilla
 (Bardet et al., 2013 \cite{peakzilla}), recently developed ChIP-exo
@@
   \centering
-<<table1,include = TRUE,echo = FALSE,eval = TRUE>>=
+<<table1,include = FALSE,echo = FALSE,eval = TRUE>>=
+  library(data.table)
   load("../data/paper_tables/Landick_rif_summary.RData")
@@
   load("../data/paper_tables/Carroll_hg19_ER.RData")
-  do.call(rbind,list(landick_aero,
-                     landick_rif,
-                     pugh_ctcf,
-                     carroll_mm9_FoxA1,
-                     meijing_hg19_GR,
-                     carroll_hg19_ER))
+  landick_aero[,Organism := "E.Coli"]
+  landick_aero[,Replicate := c(1,2,1,2)]
+  landick_aero[,TF := "Sig70"]
+  landick_aero[ ,cond := c(rep("Aerobic",2),rep("Anaerobic", 2))]
+  landick_rif[,Organism := "E.Coli"]
+  landick_rif[,Replicate := c(1,1,2,2)]
+  landick_rif[,TF := "Sig70"]
+  landick_rif[ ,cond := rep(c("Rif-0min","Rif-20min"),each = 2)]
+  pugh_ctcf[,Organism := "Human"]
+  pugh_ctcf[,Replicate := 1]
+  pugh_ctcf[,TF := "CTCF"]
+  pugh_ctcf[ ,cond := "HeLA"]
+  carroll_mm9_FoxA1[,Organism := "Mouse"]
+  carroll_mm9_FoxA1[,Replicate := c(3,1,2)]
+  carroll_mm9_FoxA1[,TF := "Fox A1"]
+  carroll_mm9_FoxA1[ ,cond := "Mouse Liver"]
+  meijing_hg19_GR[,Organism := "Human"]
+  meijing_hg19_GR[,Replicate := 1]
+  meijing_hg19_GR[, TF := "GR"]
+  meijing_hg19_GR[,cond := c("IMR90","K562","U20S")]
+  carroll_hg19_ER[,Organism := "Human"]  
+  carroll_hg19_ER[,Replicate := c(1,3,2)]
+  carroll_hg19_ER[,TF := "ER"]
+  carroll_hg19_ER[ ,cond := "MCF-7"]
+  A <- do.call(rbind,list(landick_aero,
+                     landick_rif[order(Replicate)],
+                     pugh_ctcf[order(Replicate)],
+                     carroll_mm9_FoxA1[order(Replicate)],
+                     meijing_hg19_GR[order(Replicate)],
+                     carroll_hg19_ER[order(Replicate)]))
+  A[,nsc := unlist(nsc)]
+  setcolorder(A, c("Organism","TF","cond","Replicate","nreads","pbc","nsc","files"))
+  A[,files := NULL]
+  setnames(A,names(A),c("Organism","IP/TF","Condition/Cell","Rep.","Depth","PBC","NSC"))
+@ 
+<<table2,include = TRUE,echo = FALSE, eval = TRUE >>= 
+  A
 @ 
@@
 We made a systematic exploration of several ChIP-exo experiments and
-provide a list of factors that reflect the quality of a ChIP-exo
-experiment and we provide a QC pipeline which is capable of assessing
-the balance between the enrichment and the library complexity of a
-ChIP-exo experiment. Additionally, a set of diagnostics was established
-to assess the an experiment's quality. The QC pipeline only requires a
-set of aligned reads to give a global overview of a ChIP-exo
-experiment, this overview coincides with more elaborate analysis, such
-as motif detection in a set of high quality regions or resolution
-analysis given a set of annotations as gold-standard.
+provided a list of factors that reflect the quality of a ChIP-exo
+experiment and we developed a QC pipeline which is capable of
+assessing the balance between the enrichment and the library
+complexity of a ChIP-exo experiment. Additionally, a set of
+diagnostics was established to assess the quality of a ChIP-exo
+experiment. The QC pipeline only requires a set of aligned reads to
+give a global overview of a ChIP-exo experiment, this overview
+coincides with more elaborate analysis that is computationally more
+expensive to perform or requires additional inputs that may not be
+available, such as motif detection in a set of high quality regions or
+resolution analysis given a set of annotations as gold-standard.
 
@@
 ChIP-exo is more sensitive than both SE and PE ChIP-Seq. We made a
-systematic comparison between fixed depth ChIP-exo, PE ChIP-Seq and SE
+methodical comparison between fixed depth ChIP-exo, PE ChIP-Seq and SE
 ChIP-Seq, and we probed that for sufficiently complex libraries,
@@
 \subsection*{Saturation analysis of ChIP-exo, PE ChIP-Seq and SE ChIP-Seq.}
+
+To perform the saturation analysis, we sub-sampled $N$ fragments for
+both ChIP-exo and SE ChIP-Seq protocols. For PE ChIP-Seq we
+sub-sampled $N$ pairs or $N/2$ fragments. For all the experiments under a certain condition, we
 

### 2016-02-29 rmarkdown/paper.tex
finished paper , corrected table 1

@@
   \centering
+
 \begin{knitrout}
@@
 \begin{verbatim}
-##                files    nreads        pbc      nsc
-## 1      edsn931_Sig70  13961493 0.13994203 103.4239
-## 2      edsn933_Sig70  14810838 0.16339252 162.8002
-## 3      edsn935_Sig70  16108774 0.13535422 153.5088
-## 4      edsn937_Sig70  13636541 0.15322537 172.5815
-## 5     edsn1311_Sig70    902921 0.26895580 13.76908
-## 6     edsn1314_Sig70   1852124 0.25908459 17.91882
-## 7     edsn1317_Sig70   2104427 0.25841987  29.6083
-## 8     edsn1320_Sig70  11548572 0.15107930 13.08626
-## 9               CTCF  48478450 0.45796889 16.02482
-## 10         ERR336935  22421729 0.10683407 72.04243
-## 11         ERR336942  22210461 0.65622704 21.28197
-## 12         ERR336956  23307557 0.79962169 60.42191
-## 13 IMR90_GR_chip-exo  47443803 0.29785605 8.860176
-## 14  K562_GR_chip-exo 116518000 0.05047634 4.117883
-## 15  U2OS_GR_chip-exo   3255111 0.77142141 10.05883
-## 16         ERR336933   9289835 0.80825680 19.87523
-## 17         ERR336938  12464836 0.82035947 18.72806
-## 18         ERR336950  11041833 0.80241200 21.48509
+##     Organism  IP/TF Condition/Cell Rep.     Depth        PBC        NSC
+##  1:   E.Coli  Sig70        Aerobic    1  13961493 0.13994203 103.423876
+##  2:   E.Coli  Sig70        Aerobic    2  14810838 0.16339252 162.800188
+##  3:   E.Coli  Sig70      Anaerobic    1  16108774 0.13535422 153.508809
+##  4:   E.Coli  Sig70      Anaerobic    2  13636541 0.15322537 172.581549
+##  5:   E.Coli  Sig70       Rif-0min    1    902921 0.26895580  13.769077
+##  6:   E.Coli  Sig70       Rif-0min    1   1852124 0.25908459  17.918825
+##  7:   E.Coli  Sig70      Rif-20min    2   2104427 0.25841987  29.608305
+##  8:   E.Coli  Sig70      Rif-20min    2  11548572 0.15107930  13.086255
+##  9:    Human   CTCF           HeLA    1  48478450 0.45796889  16.024817
+## 10:    Mouse Fox A1    Mouse Liver    1  22210461 0.65622704  21.281972
+## 11:    Mouse Fox A1    Mouse Liver    2  23307557 0.79962169  60.421912
+## 12:    Mouse Fox A1    Mouse Liver    3  22421729 0.10683407  72.042428
+## 13:    Human     GR          IMR90    1  47443803 0.29785605   8.860176
+## 14:    Human     GR           K562    1 116518000 0.05047634   4.117883
+## 15:    Human     GR           U20S    1   3255111 0.77142141  10.058834
+## 16:    Human     ER          MCF-7    1   9289835 0.80825680  19.875232
+## 17:    Human     ER          MCF-7    2  11041833 0.80241200  21.485094
+## 18:    Human     ER          MCF-7    3  12464836 0.82035947  18.728062
 \end{verbatim}
@@
 ##                      files   nreads       pbc      nsc
-## pet.1 edsn1396_Sig70.sort.bam 13445022 0.9426179 8.865244
-## pet.2 edsn1398_Sig70.sort.bam 16538920 0.9378843 7.031836
-## pet.3 edsn1400_Sig70.sort.bam 11642722 0.8891744 10.77284
-## pet.4 edsn1402_Sig70.sort.bam 16854026 0.9407020 7.936239
-## set.1 edsn1396_Sig70.sort.bam  6722511 0.6632742  9.01779
-## set.2 edsn1398_Sig70.sort.bam  8269460 0.5594449 7.179539
-## set.3 edsn1400_Sig70.sort.bam  5821361 0.6472382 10.89898
-## set.4 edsn1402_Sig70.sort.bam  8427013 0.5895118 8.124717
+## 1: edsn1396_Sig70.sort.bam 13445022 0.9426179 8.865244
+## 2: edsn1398_Sig70.sort.bam 16538920 0.9378843 7.031836
+## 3: edsn1400_Sig70.sort.bam 11642722 0.8891744 10.77284
+## 4: edsn1402_Sig70.sort.bam 16854026 0.9407020 7.936239
+## 5: edsn1396_Sig70.sort.bam  6722511 0.6632742  9.01779
+## 6: edsn1398_Sig70.sort.bam  8269460 0.5594449 7.179539
+## 7: edsn1400_Sig70.sort.bam  5821361 0.6472382 10.89898
+## 8: edsn1402_Sig70.sort.bam  8427013 0.5895118 8.124717
 \end{verbatim}

### 2016-02-29 rmarkdown/paper.R

@@ -1,16 +1,48 @@
-## ----table1,include = TRUE,echo = FALSE,eval = TRUE----------------------
+## ----table1,include = FALSE,echo = FALSE,eval = TRUE---------------------
+  library(data.table)
   load("../data/paper_tables/Landick_rif_summary.RData")
   load("../data/paper_tables/Landick_aero_summary.RData")
   load("../data/paper_tables/pugh_ctcf.RData")
   load("../data/paper_tables/Carroll_mm9_FoxA1.RData")
   load("../data/paper_tables/Meijsing_hg19_GR.RData")
   load("../data/paper_tables/Carroll_hg19_ER.RData")
-  do.call(rbind,list(landick_aero,
-                     landick_rif,
-                     pugh_ctcf,
-                     carroll_mm9_FoxA1,
-                     meijing_hg19_GR,
-                     carroll_hg19_ER))
+  landick_aero[,Organism := "E.Coli"]
+  landick_aero[,Replicate := c(1,2,1,2)]
+  landick_aero[,TF := "Sig70"]
+  landick_aero[ ,cond := c(rep("Aerobic",2),rep("Anaerobic", 2))]
+  landick_rif[,Organism := "E.Coli"]
+  landick_rif[,Replicate := c(1,1,2,2)]
+  landick_rif[,TF := "Sig70"]
+  landick_rif[ ,cond := rep(c("Rif-0min","Rif-20min"),each = 2)]
+  pugh_ctcf[,Organism := "Human"]
+  pugh_ctcf[,Replicate := 1]
+  pugh_ctcf[,TF := "CTCF"]
+  pugh_ctcf[ ,cond := "HeLA"]
+  carroll_mm9_FoxA1[,Organism := "Mouse"]
+  carroll_mm9_FoxA1[,Replicate := c(3,1,2)]
+  carroll_mm9_FoxA1[,TF := "Fox A1"]
+  carroll_mm9_FoxA1[ ,cond := "Mouse Liver"]
+  meijing_hg19_GR[,Organism := "Human"]
+  meijing_hg19_GR[,Replicate := 1]
+  meijing_hg19_GR[, TF := "GR"]
+  meijing_hg19_GR[,cond := c("IMR90","K562","U20S")]
+  carroll_hg19_ER[,Organism := "Human"]  
+  carroll_hg19_ER[,Replicate := c(1,3,2)]
+  carroll_hg19_ER[,TF := "ER"]
[…]

### 2016-02-29 rmarkdown/paper.Rnw
draft finished, only need to review figure legends

@@
 these two peaks are confounded. Hence an enrichment measure based in
-SCC such as the NSC is harder to interpret. In figure
+SCC such as the NSC is harder to interpret. In Figure
 \ref{fig:scc_exo} we observe that both local maxima are hard to
 differentiate in the SCC curves for the $\sig$ ChIP-exo datasets used to
-calculate the NSC values in table \ref{tab:qc}.
+calculate the NSC values in Table \ref{tab:qc}.
 
@@
 series of summary statistics. Finally it creates several
-visualizations designed to assess the quality level of ChIP-exo
+visualizations designed to diagnose the quality level of ChIP-exo
 sample.
@@
 
-Using the FoxA1 in mouse liver cell lines from \cite{exoillumina} and
-these two quantities, the relationship between library complexity and
-experiment enrichment was explored. In figure \ref{fig:enrich}A we can
-observe both statistics for the regions that are common for the three
-replicates. This figure shows typical patterns for ChIP-exo
+Using the Fox A1 in mouse liver cell lines from \cite{exoillumina} and
+these two quantities, we explored the relationship between library
+complexity and experiment enrichment. In Figure \ref{fig:enrich}A we
+can observe both statistics for the regions that are common for the
+three replicates. This figure shows typical patterns for ChIP-exo
 experiment, there are two strong arms: The one on the left with low
@@
 
-The strand imbalance assessment in based in the observation that the
+The strand imbalance assessment is based in the observation that the
 enriched regions usually are composed of a higher quantity of reads,
@@
 indicator is of particular importance, since several methods rely on
-the ``peak-pair'' assumption. In table \ref{tab:qc}, we calculated the
-global FSR and noticed that for all experiments is roughly $0.5$,
-which means there are roughly the same amount of reads in both
-strands. However figure \ref{fig:strand} shows that the global FSR
-does not represent the experiment's local strand imbalance, hence the
-``peak pair'' assumption may not hold well in real ChIP-exo data.
+the ``peak-pair'' assumption. For every ChIP-exo experiment, we
+calculated the global FSR and noticed that for all experiments is
+roughly $0.5$, which means there are roughly the same amount of reads
+in both strands. However figure \ref{fig:strand} shows that the global
+FSR does not represent the experiment's local strand imbalance, hence
+the ``peak pair'' assumption may not hold well in real ChIP-exo data.
 
 In order to asses the strand imbalance we created the visualization
-shown in figure \ref{fig:strand}: Figure \ref{fig:strand}A presents
+shown in Figure \ref{fig:strand}: Figure \ref{fig:strand}A presents
 the FSR's behavior as the lower depth regions are being filtered out,
@@
 replicate, we divided the partitioned regions by asking whether they
-overlap a set of high quality ChIP-exo peaks, and then we tested if
-the strand imbalance's distribution were the same for both
-classes. For regions composed by a higher amount of reads, is is
-harder to distinguish their peaks by considering only the strand
-imbalance, and in an experiment with lower quality it is harder to
-make this distinction.
+overlap a set of high quality ChIP-exo peaks, and then we tested
+(using the Wilcoxon rank sum test over an imbalance index) if the
+strand imbalance's distribution is the same for both classes. For
+regions composed by a higher amount of reads, it is harder to
+distinguish their peaks by considering only the strand imbalance,
+hence in a better quality ChIP-exo experiment it is easier to
+distinguish enriched regions by the amount of reads in both
+strands. Additionally we may consider that the background of a
+ChIP-exo experiment it is more unbalanced than the enriched regions.
 
@@
 annotation and the dPeak binding site estimates. The left panel of
-figure \ref{fig:reso_all} shows that the sensitivity increases as the
+Figure \ref{fig:reso_all} shows that the sensitivity increases as the
 mean distance between binding events increases. Despite the when the
@@
 ChIP-Seq. Hence, we sampled a fixed amount of reads for each of the
-ChIP-exo, PE ChIP-Seq and SE ChIP-Seq datasets of the $\sig$
-samples. For each sampled dataset we applied out lower-to-higher
+ChIP-exo, PE ChIP-Seq and SE ChIP-Seq datasets of the $\sig$ samples
+($N$ reads for both ChIP-exo experiment and $N/2$ or $N$ pairs for PE
+ChIP-Seq). For each sampled dataset we applied out lower-to-higher
 resolution pipeline by calling peaks with MOSAiCS \cite{mosaics} and
@@
 \cite{gem} or dPeak \cite{dpeak}. In a good dataset such as both of
-the ChIP-exo experiments under aerobic (panel 933) or anaerobic (panel
-931) conditions, all the methods are comparable in resolution, and
-dPeak slightly outperforms the rest. On the other hand, for
-experiments with low library complexity the resolution calculated with
-dPeak's predictions is smaller than both Mace and Peakzilla, but it is
-greater than Gem. This may be due that the fact that Gem uses sequence
-information in addition to the aligned 5' end counts that dPeak
-uses.\footnote{For here we may probably use only 933 as part of the
-  main article and keep the rest for the supplement}
+the ChIP-exo experiments under aerobic (panel 931 and 933) conditions,
+all the methods are comparable in resolution, and dPeak slightly
+outperforms the rest. On the other hand, for experiments with low
+library complexity the resolution calculated with dPeak's predictions
+is smaller than both Mace and Peakzilla, but it is greater than
+Gem. This may be due that the fact that Gem uses sequence information
+in addition to the aligned 5' end counts that dPeak uses. \footnote{For
+  here we may probably use only 933 as part of the main article and
+  keep the rest for the supplement}
 
@@
 
+We studied the shared biases between ChIP-exo and ChIP-Seq data, and
+noticed that for eukaryotic genomes the relationship between ChIP-Seq
+data and either the mappability or the GC content scores are still
+present in ChIP-exo. We also examined ChIP-exo's background and
+noticed that is significantly different from the ChIP-Seq one, since
+it consists mostly on a small quantity of fragments that was not
+digested by the exonuclease enzyme. Additionally, we showed that the
+background regions are imbalanced respect the number of fragments in
+both strand, and that in a lower quality ChIP-exo experiment those
+regions are going to be harder to differentiate from the possibly
+enriched regions.
+
 To the extent of our knowledge, we made for the fist time a comparison
@@
 
-\subsection*{Construction of a SE ChIP-Seq experiment.}
-
-For the rif-treatment ChIP-Seq experiments, we sampled SE ChIP-Seq
-experiment from the PE ones by taking one of both ends randomly
-following a $\mbox{Ber}(0.5)$ model.
-
-\subsection*{Saturation analysis of ChIP-exo, PE ChIP-Seq and SE ChIP-Seq.}
-
-To perform the saturation analysis, we sub-sampled $N$ fragments for
-both ChIP-exo and SE ChIP-Seq protocols. For PE ChIP-Seq we
-sub-sampled $N$ pairs or $N/2$ fragments. For all the experiments under a certain condition, we
-
 \subsection*{ChIP-Seq quality control metrics and Strand
@@
 
-The statistics in table \ref{tab:qc} and the SCC curves from figure
+The statistics in Table \ref{tab:qc} and the SCC curves from Figure
 \ref{fig:scc_exo} where calculated with the \textbf{ChIPUtils} package
@@
 \url{https://github.com/welch16/ChIPexoqual}.
+
+\subsection*{Motif analysis of Fox A1 enriched regions}
+
+For all three replicates we called peaks using the GC + Mappability
+model from \textbf{MOSAiCS} using an FDR level of 5\%, filtering out
+the peaks with average ChIP counts below one hundred fragments and
+merging peaks gaped by at most 200 bp. Then, we fitted the
+\textbf{dPeak} model considering at most 5 binding events for each
+peak, and we searched for the Fox A1 motif over a 10 bp window around
+the estimated binding events. We used FIMO's command line on its 4.9.1
+version.
+
+\subsection*{Imbalance index}
+
+For every ChIP-exo experiment, we partitioned the experiment into
+regions using the QC pipeline. For each region we calculated the FSR
+defined as the ratio between the number of forward stranded reads and
+the total number of reads in a region. The imbalance index is defined
+as:
+
+\[
+\mbox{Imbalance index} = -\log_{10} (4 \times \mbox{FSR} \times (1 - \mbox{FSR}))
+\]
+
+\subsection*{Construction of a SE ChIP-Seq experiment.}
+
+For the rif-treatment ChIP-Seq experiments, we sampled SE ChIP-Seq
+experiment from the PE ones by taking one of both ends randomly
+following a $\mbox{Ber}(0.5)$ model.
+
+\subsection*{Saturation analysis of ChIP-exo, PE ChIP-Seq and SE ChIP-Seq.}
+
+To perform the saturation analysis, we sub-sampled $N$ fragments for
+both ChIP-exo and SE ChIP-Seq protocols. For PE ChIP-Seq we
+sub-sampled $N$ pairs or $N/2$ fragments. For all the experiments under a certain condition, we
 
@@
   \centering
-<<table_chipseq,include = TRUE,echo = FALSE,eval = TRUE>>=
+<<table_chipseq1,include = TRUE,echo = FALSE,eval = TRUE>>=
   load("../data/paper_tables/Landick_rif_chipseq_summary.RData")

### 2016-02-29 rmarkdown/paper.R

@@ -62,7 +62,7 @@
   topM <- 300
 
 
-## ----table_chipseq,include = TRUE,echo = FALSE,eval = TRUE---------------
+## ----table_chipseq1,include = TRUE,echo = FALSE,eval = TRUE--------------
   load("../data/paper_tables/Landick_rif_chipseq_summary.RData")
   do.call(rbind,chipseq)
 

### 2016-02-29 rmarkdown/paper.Rnw
finished methods part

@@
 
-For all three replicates we called peaks using the GC + Mappability
-model from \textbf{MOSAiCS} using an FDR level of 5\%, filtering out
-the peaks with average ChIP counts below one hundred fragments and
-merging peaks gaped by at most 200 bp. Then, we fitted the
-\textbf{dPeak} model considering at most 5 binding events for each
-peak, and we searched for the Fox A1 motif over a 10 bp window around
-the estimated binding events. We used FIMO's command line on its 4.9.1
-version.
+For all three replicates we called peaks using the MOSAiCS GC +
+Mappability model using an FDR level of 5\%, filtering out the peaks
+with average ChIP counts below one hundred fragments and merging peaks
+gaped by at most 200 bp. Then, we fitted the dPeak model considering
+at most 5 binding events for each peak, and we searched for the Fox A1
+motif over a 10 bp window around the estimated binding events. We used
+FIMO's command line 4.9.1 version.
 
@@
 
-\subsection*{Construction of a SE ChIP-Seq experiment.}
+\subsection*{Construction of a SE ChIP-Seq from a PE ChIP-Seq experiment.}
 
@@
 
+\subsection*{dPeak analysis of $\sig$ ChIP-exo and ChIP-Seq data.}
+
+For the resolution and sensitivity analysis, we used the MOSAiCS GC
+content + mappability model to call peaks for ChIP-exo experiments,
+while for both SE and PE ChIP-Seq experiments we used the MOSAiCS
+Input model. To avoid false positives we only considered ChIP-exo
+peaks with average ChIP counts greater than 3000 that overlapped both
+the SE and PE ChIP-Seq peaks. Then we estimated the binding site
+events using the dPeak model with a maximum number of 5 binding
+events. We considered RegulonDB annotation as gold-standard;
+Resolution is defined as the minimum distance from an annotation to an
+estimated binding events and Sensitivity is defined as the fraction of
+annotation in a are within 20 bp from an estimated binding event.
+
 \subsection*{Saturation analysis of ChIP-exo, PE ChIP-Seq and SE ChIP-Seq.}
@@
 both ChIP-exo and SE ChIP-Seq protocols. For PE ChIP-Seq we
-sub-sampled $N$ pairs or $N/2$ fragments. For all the experiments under a certain condition, we
-
-\subsection*{dPeak analysis of $\sig$ ChIP-exo and ChIP-Seq data.}
-
-\vspace{2 cm}
+sub-sampled $N$ pairs or $N/2$ fragments. For each seed, we called
+peaks using MOSAiCS \cite{mosaics} (GC content + mappability for
+ChIP-exo and Input for SE and PE ChIP-Seq) for the maximum sample size
+and to avoid false positives we considered only the top 500 peaks for
+each data protocol. We defined the number of candidate regions as the
+number of top sample peaks such that a binding events was estimated
+using the sampled reads and the dPeak's model; the number of predicted
+events is the total quantity of binding events estimated using the
+dPeak's model; the number of identified targets are number of
+gold-standard annotations within 20 bp from an estimated binding
+events; and the resolution is defined as the minimum distance from a
+gold-standard annotation to an estimated binding event. We repeated
+this analysis for ten seeds and reported the median between all those
+values.
 
@@
 algorithm as candidate regions. Default tuning parameters were used
-during model fitting for all methods.
-
-%% CexoR
+during model fitting for all methods. We were unable to use CexoR
+\cite{cexor} to estimate ChIP-exo binding sites.
 
@@
   \includegraphics[width = .9\textwidth]{../figs/for_paper/saturation_analysis_old.pdf}
-  \caption{\textbf{ChIP-exo, PE ChIP-Seq and SE ChIP-Seq comparison at
-      varying sequencing depths.} Comparison of the number of candidate
-    regions (A), predicted events (B), identified targets (C) and
-    resolution (D) among ChIP-exo, PE ChIP-Seq and SE
-    ChIP-Seq. RegulonDB annotations are considered as a gold
-    standard. A gold standard binding events was deemed identified if
-    a binding event was estimated at a $\pm$ \Sexpr{ext} vicinity of
-    it.}
+  \caption{Comparison of the number of candidate regions (A),
+    predicted events (B), identified targets (C) and resolution (D)
+    among ChIP-exo, PE ChIP-Seq and SE ChIP-Seq. RegulonDB annotations
+    are considered as a gold standard. A gold standard binding events
+    was deemed identified if a binding event was estimated at a $\pm$
+    \Sexpr{ext} vicinity of it.}
   \label{fig:design}
@@
 
+% \textbf{ChIP-exo, PE ChIP-Seq and SE ChIP-Seq comparison at varying
+%   sequencing depths.} 
  \newpage
@@
   \includegraphics[width = .3\textwidth,page =4]{../figs/saturation/K12_alignment/Sig70_rep2_rif20_saturation_analysis.pdf}
-  \caption{\textbf{ChIP-exo, PE ChIP-Seq and SE ChIP-Seq comparison at
-      varying sequencing depth} Comparison of the number of predicted
-    events (left), identified targets (middle) and resolution (right)
-    among ChIP-exo, PE ChIP-Seq and SE ChIP-Seq. RegulonDB annotations
-    are considered as gold standard. A RegulonDB binding events was
-    deemed identified if a binding event was estimated at a $\pm$ \Sexpr{ext}
-    vicinity of it.}
+  \caption{Comparison of the number of predicted events (left),
+    identified targets (middle) and resolution (right) among ChIP-exo,
+    PE ChIP-Seq and SE ChIP-Seq. RegulonDB annotations are considered
+    as gold standard. A RegulonDB binding events was deemed identified
+    if a binding event was estimated at a $\pm$ \Sexpr{ext} vicinity
+    of it.}
   \label{fig:saturation_rif}
 \end{figure}
+
+% \textbf{ChIP-exo, PE ChIP-Seq and SE ChIP-Seq comparison at varying
+%   sequencing depth}
 

### 2016-02-29 rmarkdown/paper.tex
finished methods part

@@
   \includegraphics[width = .9\textwidth]{../figs/for_paper/saturation_analysis_old.pdf}
-  \caption{\textbf{ChIP-exo, PE ChIP-Seq and SE ChIP-Seq comparison at
-      varying sequencing depths.} Comparison of the number of candidate
-    regions (A), predicted events (B), identified targets (C) and
-    resolution (D) among ChIP-exo, PE ChIP-Seq and SE
-    ChIP-Seq. RegulonDB annotations are considered as a gold
-    standard. A gold standard binding events was deemed identified if
-    a binding event was estimated at a $\pm$ 15 vicinity of
-    it.}
+  \caption{Comparison of the number of candidate regions (A),
+    predicted events (B), identified targets (C) and resolution (D)
+    among ChIP-exo, PE ChIP-Seq and SE ChIP-Seq. RegulonDB annotations
+    are considered as a gold standard. A gold standard binding events
+    was deemed identified if a binding event was estimated at a $\pm$
+    15 vicinity of it.}
   \label{fig:design}
@@
   \includegraphics[width = .3\textwidth,page =4]{../figs/saturation/K12_alignment/Sig70_rep2_rif20_saturation_analysis.pdf}
-  \caption{\textbf{ChIP-exo, PE ChIP-Seq and SE ChIP-Seq comparison at
-      varying sequencing depth} Comparison of the number of predicted
-    events (left), identified targets (middle) and resolution (right)
-    among ChIP-exo, PE ChIP-Seq and SE ChIP-Seq. RegulonDB annotations
-    are considered as gold standard. A RegulonDB binding events was
-    deemed identified if a binding event was estimated at a $\pm$ 15
-    vicinity of it.}
+  \caption{Comparison of the number of predicted events (left),
+    identified targets (middle) and resolution (right) among ChIP-exo,
+    PE ChIP-Seq and SE ChIP-Seq. RegulonDB annotations are considered
+    as gold standard. A RegulonDB binding events was deemed identified
+    if a binding event was estimated at a $\pm$ 15 vicinity
+    of it.}
   \label{fig:saturation_rif}
 \end{figure}
+
+% \textbf{ChIP-exo, PE ChIP-Seq and SE ChIP-Seq comparison at varying
+%   sequencing depth}
 

### 2016-03-01 rmarkdown/paper.tex
changed order fo first two authors

@@
 \author[
+   addressref={aff1},
+   email={welch@stat.wisc.edu}
+]{\inits{RW}\fnm{Rene} \snm{Welch}\thanks{These two authors contributed equally}\normalfont\textsuperscript{,}}
+\author[
    addressref={aff6},                   % id's of addresses, e.g. {aff1,aff2}
    email={chungd@musc.edu}   % email address
-]{\inits{DC}\fnm{Dongjun} \snm{Chung}}
-\author[
-   addressref={aff1},
-   email={welch@stat.wisc.edu}
-]{\inits{RW}\fnm{Rene} \snm{Welch}}
+]{\inits{DC}\fnm{Dongjun} \snm{Chung\normalfont\textsuperscript{[1],}}}
 \author[
@@
 the read counts across bins with the same mappability of GC
-content. These results indicate that binding site identification in
-ChIP-exo sample might also benefit from the use of methods that take
-into account of apparent sequence biases such as mappability and GC
-content.
+content. In Figure \ref{fig:comp}C it is shown that the ChIP-exo tag
+counts linearly increases with the mappability score and in Figure
+\ref{fig:comp}D it is shown that for GC - content below 0.6, the mean
+ChIP tag count increases and for GC - content greater than 0.6 it
+shows a decreasing trend.These results indicate that binding site
+identification in ChIP-exo sample might also benefit from the use of
+methods that take into account of apparent sequence biases such as
+mappability and GC content.
 
@@
 estimated binding events and Sensitivity is defined as the fraction of
-annotation in a are within 20 bp from an estimated binding event.
+annotation in a are within 15 bp from an estimated binding event.
 
@@
 dPeak's model; the number of identified targets are number of
-gold-standard annotations within 20 bp from an estimated binding
+gold-standard annotations within 15 bp from an estimated binding
 events; and the resolution is defined as the minimum distance from a
@@
 \bibliography{chip_exo_paper}
+\nocite{exo_gb}
+\nocite{maplot1}
+\nocite{maplot2}
+\nocite{chipbeyond}
+
+% \nocite{esl}
 
@@
   \includegraphics[width = .46\textwidth,page = 2]{../figs/for_paper/eukaryotic_bias_CTCF.pdf}
-  \caption{ A) Hexbin plot of PE ChIP-Seq bin counts VS ChIP-exo bin
+  \caption{ A) Hexbin plot of PE ChIP-Seq bin counts vs ChIP-exo bin
     counts. B) Forward Strand Ratio densities for SE ChIP-Seq and
-    ChIP-exo peaks. C) ChIP tag counts increase linearly as
-    mappability scores increases. D) ChIP tag counts increase linearly
-    as GC content score increases when GC content is less than 0.6 and
-    then ChIP tag counts decrease as GC content increases. D)
-    Percentage of sites with at least p.value.}
+    ChIP-exo peaks. C) Mappability score vs mean ChIP tag counts with
+    0.95 confidence bands. D) GC - content score vs mean ChIP tag
+    counts with 0.95 confidence bands.}
   \label{fig:comp}
@@
 
-\newpage
-
-\begin{figure}[h!]
-  \centering
+% ChIP tag counts increase linearly as mappability scores
+% increases. D) ChIP tag counts increase linearly as GC content score
+% increases when GC content is less than 0.6 and then ChIP tag counts
+% decrease as GC content increases. D) Percentage of sites with at
+% least p.value.
+
+\newpage
+
+\begin{figure}[h!]
   \centering
@@
   \includegraphics[width = .3 \textwidth,page = 1]{../figs/FOXA1_mm9_fimo/FOXA1_log10pval_cumulative_nmotifs.pdf}
-  \caption{Using the mouse-FoxA1 experiment from \cite{exoillumina}:
-    A) Hexbin plots of $\mbox{ARC}$ against $\mbox{URCR}$, in general
-    we can see a slight separation into two strong arms, one
-    corresponds to low $\mbox{ARC}$ and varying $\mbox{URCR}$, and for
-    the other $\mbox{URCR}$ decreases as $\mbox{ARC}$ increases. B)
-    Number of candidate sites for each replicate. C) Percentage of
-    candidate sites where the FoxA1 motif was detected. D) 
- }
+  \caption{Using the mouse FoxA1 experiment from \cite{exoillumina}:
+    A) Hexbin plots of $\mbox{ARC}$ against $\mbox{URCR}$, there is a
+    slight separation into two strong arms, one corresponds to low
+    $\mbox{ARC}$ and varying $\mbox{URCR}$, and for the other
+    $\mbox{URCR}$ decreases as $\mbox{ARC}$ increases. B) Number of
+    candidate sites for each replicate. C) Percentage of candidate
+    sites where the FoxA1 motif was detected. D) Cumulative
+    proportion of detected motifs by replicate.}
   \label{fig:enrich}
@@
     filtered out quickly as in the middle row. C)
-    $-log_{10}(\text{p.value})$ of testing if the imbalance
+    $-\log_{10}(\text{p.value})$ of testing if the imbalance
     distributions differs when ChIP-exo regions overlap their peaks.}
@@
 \label{sec:supp}
+
+\emph{Note: I am not sure wheter to include several figures on this supplement}
 
@@
 
-\newpage
-
-% \nocite{exo_gb}
-%  \nocite{maplot1}
-% \nocite{maplot2}
-% \nocite{chipbeyond}
-% \nocite{esl}
-
 \end{document}

### 2016-03-15 rmarkdown/paper.Rnw
first commits with reviews from DC

@@
 
+% \thanks{These two authors contributed equally.}
+
+\author[
+   addressref={aff1},
+   noteref={n1},
+   email={welch@stat.wisc.edu}   
+]{\inits{RW}\fnm{Rene} \snm{Welch} }
 \author[
    addressref={aff6},        % id's of addresses, e.g. {aff1,aff2}
+   noteref={n1},
    email={chungd@musc.edu}   % email address
-]{\inits{DC}\fnm{Dongjun} \snm{Chung}}
-\author[
-   addressref={aff1},
-   email={welch@stat.wisc.edu}
-]{\inits{RW}\fnm{Rene} \snm{Welch}}
+]{\inits{DC}\fnm{Dongjun} \snm{Chung} } 
 \author[
@@
 
-% \begin{artnotes}
-% %\note{Sample of title note}     % note to the article
-% \note[id=n1]{Equal contributor} % note, connected to author
-% \end{artnotes}
+\begin{artnotes}
+\note[id=n1]{These two authors contributed equally.} % note, connected to author
+\end{artnotes}
 
@@
   \kwd{ChIP-exo}
-  \kwd{QC}
-  \kwd{TFBS}
-  \kwd{BS identification on High-Res}
+  \kwd{Quality Control}
+  \kwd{ChIP-Seq}
+  \kwd{Spatial Resolution}
+  \kwd{Transcription Factor}  
+  \kwd{Binding Site Identification on High-Res}
+  \kwd{Deconvolution}
 \end{keyword}
@@
 sequencing) assay. ChIP-exo experiments first capture millions of DNA
-fragments (150 -250 bp in length) that the protein under study
-interacts with using random fragmentation of DNA and a
-protein-specific antibody. Then, exonuclease is introduced to trim the
-5' end of each DNA fragment to a fixed distance from the bound protein
-compared to ChIP-Seq. This step is unique to ChIP-exo and could
-potentially provide significantly higher spatial resolution compared
-to ChIP-Seq. Finally, high throughput sequencing of a small region (25
-to 100 bp) at the 5' end of each fragment generates millions of reads.
-
-While the number of ChIP-exo data keeps increasing,characteristics of
+fragments that the protein under study interacts with using random
+fragmentation of DNA and a protein-specific antibody. Then,
+exonuclease is introduced to trim the $5\prime$ end of each DNA
+fragment to a fixed distance from the bound protein compared to
+ChIP-Seq. This step is unique to ChIP-exo and could potentially
+provide significantly higher spatial resolution compared to
+ChIP-Seq. Finally, high throughput sequencing of a small region (36 to
+100 bp) at the $5\prime$ end of each fragment generates millions of
+reads.
+
+While the number of ChIP-exo data keeps increasing, characteristics of
 ChIP-exo data are not fully investigated yet. First, DNA libraries
@@
 in both strands, locally there may be more reads in one strand than in
-the other. To address this challenges, we suggest a collection of
-quality control visualizations to understand which of this biases are
-present in a ChIP-exo experiment and globally asses the enrichment and
-library complexity of a ChIP-exo sample. We gathered ChIP-exo data
-from diverse organisms: CTCF factor in human \cite{exo1}; ER factor in
+the other. To address these challenges, we suggest a collection of
+quality control visualizations to interrogate these biases in a
+ChIP-exo experiment and globally asses the enrichment and library
+complexity of a ChIP-exo sample. We gathered ChIP-exo data from
+diverse organisms: CTCF factor in human \cite{exo1}; ER factor in
 human and FoxA1 factor in mouse (Serandour et al., 2013
-\cite{exoillumina}); GR in IMR90, K562 and U20S cell lines (Starick et
-al., 2015 \cite{starick15}) and generated $\sig$ factor in Escherichia
-Coli (\emph{E. Coli}) under aerobic ($ + O_2$) condition, and treated
-by rifampicin by 0 and 20 minutes (courtesy of Professor Robert
-Landick's lab). %% note to self, we may add more experiments here
+\cite{exoillumina}); Glucocorticoid receptor (GR) in IMR90, K562 and
+U20S cell lines (Starick et al., 2015 \cite{starick15}).Furthermore,
+we also generated ChIP-exo and ChIP-Seq data for $\sig$ factor in
+\emph{Escherichia Coli} (\emph{E. Coli}) under aerobic ($ + O_2$)
+condition, and treated by rifampicin by 0 and 20 minutes (courtesy of
+Professor Robert Landick's lab).
+%% note to self, we may add more experiments here
 
@@
 Gem. This may be due that the fact that Gem uses sequence information
-in addition to the aligned 5' end counts that dPeak uses. \footnote{For
-  here we may probably use only 933 as part of the main article and
-  keep the rest for the supplement}
+in addition to the aligned $5\prime$ end counts that dPeak
+uses. \footnote{For here we may probably use only 933 as part of the
+  main article and keep the rest for the supplement}
 
@@
 estimated binding events and Sensitivity is defined as the fraction of
-annotation in a are within 20 bp from an estimated binding event.
+annotation in a are within \Sexpr{ext} bp from an estimated binding event.
 
@@
 dPeak's model; the number of identified targets are number of
-gold-standard annotations within 20 bp from an estimated binding
+gold-standard annotations within \Sexpr{ext} bp from an estimated binding
 events; and the resolution is defined as the minimum distance from a
@@
 \section{Figures}
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width =
+  .95 \textwidth]{/p/keles/ChIPexo/volume3/ChIPexo/figs/for_paper/chip_explanation_withEnzyme.pdf}
+  \caption{\textbf{Description of ChIP-exo, SE ChIP-Seq and PE
+      ChIP-Seq:} The DNA is sonicated, for ChIP-exo a exonuclease
+    enzyme (orange hexagon) trims the $5\prime$ ends of each DNA
+    fragment to a fixed distance from the bound protein, finally is
+    subjected to Immunoprecipitation and amplification. For both
+    ChIP-exo and SE ChIP-Seq an adapter is ligated (green triangles)
+    at the $5\prime$ ends, while for PE ChIP-Seq is ligated to both
+    ends.}
+  \label{fig:chip_diagram}
+\end{figure}
+\newpage
 
@@
 \label{sec:supp}
+
+\emph{Note: I am not sure whether to include several figures on this supplement}
 

### 2016-03-15 rmarkdown/paper.tex
first commits with reviews from DC

@@
 
+% \thanks{These two authors contributed equally.}
+
 \author[
    addressref={aff1},
+   noteref={n1},
    email={welch@stat.wisc.edu}   
-]{\inits{RW}\fnm{Rene} \snm{Welch}\thanks{These two authors contributed equally}\normalfont\textsuperscript{,}}
+]{\inits{RW}\fnm{Rene} \snm{Welch} }
 \author[
    addressref={aff6},        % id's of addresses, e.g. {aff1,aff2}
+   noteref={n1},
    email={chungd@musc.edu}   % email address
-]{\inits{DC}\fnm{Dongjun} \snm{Chung\normalfont\textsuperscript{[1],}}}
+]{\inits{DC}\fnm{Dongjun} \snm{Chung} } 
 \author[
@@
   \centering
+  \includegraphics[width =
+  .95 \textwidth]{/p/keles/ChIPexo/volume3/ChIPexo/figs/for_paper/chip_explanation_withEnzyme.pdf}
+  \caption{\textbf{Description of ChIP-exo, SE ChIP-Seq and PE
+      ChIP-Seq:} The DNA is sonicated, for ChIP-exo a exonuclease
+    enzyme (orange hexagon) trims the $5\prime$ ends of each DNA
+    fragment to a fixed distance from the bound protein, finally is
+    subjected to Immunoprecipitation and amplification. For both
+    ChIP-exo and SE ChIP-Seq an adapter is ligated (green triangles)
+    at the $5\prime$ ends, while for PE ChIP-Seq is ligated to both
+    ends.}
+  \label{fig:chip_diagram}
+\end{figure}
+\newpage
+
+\begin{figure}[h!]
+\centering
 \includegraphics[width = .8\textwidth]{/p/keles/ChIPexo/volume3/ChIPexo/figs/for_paper/EColi_strand_cross_corr.pdf}
@@
 
-\emph{Note: I am not sure wheter to include several figures on this supplement}
+\emph{Note: I am not sure whether to include several figures on this supplement}
 

### 2016-03-15 rmarkdown/paper.Rnw
new commit test

@@
   ChIP-exo experiments}
-
-% \thanks{These two authors contributed equally.}
 
@@
 100 bp) at the $5\prime$ end of each fragment generates millions of
-reads.
+reads. Figure \ref{fig:chip_diagram} illustrates the differences
+between ChIP-exo, Single End (SE) ChIP-Seq and Paired End (PE)
+ChIP-Seq: The $5\prime$ ends of a ChIP-exo experiment are located more
+tightly around the binding proteins than in a ChIP-Seq experiment; in
+a PE ChIP-Seq experiment both ends are observed while in a SE ChIP-Seq
+experiment only the $5\prime$ end.
+
+% Reffers to... First, DNA libraries... I think that these might be
+% too much details for the introduction. How about discussing them in
+% the results when we investigate the related
 
@@
 in both strands, locally there may be more reads in one strand than in
-the other. To address these challenges, we suggest a collection of
+the other.
+
+To address these challenges, we suggest a collection of
 quality control visualizations to interrogate these biases in a
@@
 ChIP-exo. Previous ChIP-exo analyses used ChIP-Seq samples to compare
-the resolution between experiments (\cite{exo1}, \cite{exo2},
-\cite{exoillumina}); Carroll et al., 2014 \cite{carroll.qc} studied
-the use of the Strand-Cross Correlation (SCC) (Kharchenko et al., 2008
-\cite{strandcc}) and showed that by filtering blacklisted regions the
-estimation of the SCC is improved. However, using the SCC may not be
-helpful since the peaks that are attained at the read and fragment
-lengths are confused in a typical ChIP-exo SCC curve. Additionally
-this method requires to know blacklisted regions in advance which may
-not be available. In our pipeline we propose an out-the-shelf analysis
-that contrasts the enrichment of the experiment against its library
-complexity.
+the resolution between experiments; Carroll et al., 2014
+\cite{carroll.qc} studied the use of the Strand-Cross Correlation
+(SCC) (Kharchenko et al., 2008 \cite{strandcc}) and showed that by
+filtering blacklisted regions the estimation of the SCC is
+improved. However, using the SCC may not be helpful since the peaks
+that are attained at the read and fragment lengths are confused in a
+typical ChIP-exo SCC curve. Additionally this method requires to know
+blacklisted regions in advance which may not be available. In our
+pipeline we propose an out-the-shelf analysis that contrasts the
+enrichment of the experiment against its library complexity.
 

### 2016-03-15 rmarkdown/paper.Rnw
fixed background section of paper

@@
 sequencing) assay. ChIP-exo experiments first capture millions of DNA
-fragments that the protein under study interacts with using random
-fragmentation of DNA and a protein-specific antibody. Then,
-exonuclease is introduced to trim the $5\prime$ end of each DNA
-fragment to a fixed distance from the bound protein compared to
-ChIP-Seq. This step is unique to ChIP-exo and could potentially
-provide significantly higher spatial resolution compared to
-ChIP-Seq. Finally, high throughput sequencing of a small region (36 to
-100 bp) at the $5\prime$ end of each fragment generates millions of
-reads. Figure \ref{fig:chip_diagram} illustrates the differences
-between ChIP-exo, Single End (SE) ChIP-Seq and Paired End (PE)
-ChIP-Seq: The $5\prime$ ends of a ChIP-exo experiment are located more
-tightly around the binding proteins than in a ChIP-Seq experiment; in
-a PE ChIP-Seq experiment both ends are observed while in a SE ChIP-Seq
-experiment only the $5\prime$ end.
+fragments (150 - 250 bp in length) that the protein under study
+interacts with using random fragmentation of DNA and a
+protein-specific antibody. Then, exonuclease is introduced to trim the
+$5\prime$ end of each DNA fragment to a fixed distance from the bound
+protein compared to ChIP-Seq. This step is unique to ChIP-exo and
+could potentially provide significantly higher spatial resolution
+compared to ChIP-Seq. Finally, high throughput sequencing of a small
+region (36 to 100 bp) at the $5\prime$ end of each fragment generates
+millions of reads. Figure \ref{fig:chip_diagram} illustrates the
+differences between ChIP-exo, Single End (SE) ChIP-Seq and Paired End
+(PE) ChIP-Seq: The $5\prime$ ends of a ChIP-exo experiment are located
+more tightly around the binding proteins than in a ChIP-Seq
+experiment; in a PE ChIP-Seq experiment both ends are observed while
+in a SE ChIP-Seq experiment only the $5\prime$ end.
 
@@
 generated by the ChIP-exo protocol seem to be less complex than the
-libraries generated by ChIP-Seq (Mahony et al., 2015,
-\cite{exo_review}), i.e. the possible number of positions to which the
-reads can be aligned has been reduced due to the exonuclease
-digestion. Second, although there are roughly the same amount of read
-in both strands, locally there may be more reads in one strand than in
-the other.
-
-To address these challenges, we suggest a collection of
-quality control visualizations to interrogate these biases in a
-ChIP-exo experiment and globally asses the enrichment and library
-complexity of a ChIP-exo sample. We gathered ChIP-exo data from
-diverse organisms: CTCF factor in human \cite{exo1}; ER factor in
-human and FoxA1 factor in mouse (Serandour et al., 2013
-\cite{exoillumina}); Glucocorticoid receptor (GR) in IMR90, K562 and
-U20S cell lines (Starick et al., 2015 \cite{starick15}).Furthermore,
-we also generated ChIP-exo and ChIP-Seq data for $\sig$ factor in
-\emph{Escherichia Coli} (\emph{E. Coli}) under aerobic ($ + O_2$)
-condition, and treated by rifampicin by 0 and 20 minutes (courtesy of
-Professor Robert Landick's lab).
-%% note to self, we may add more experiments here
-
-Most of current ChIP-Seq quality control (QC) guidelines (Landt et
-al., \cite{encode_qc}) may not be applicable on ChIP-exo, additionally
-to our knowledge there are not established QC pipelines for
-ChIP-exo. Previous ChIP-exo analyses used ChIP-Seq samples to compare
-the resolution between experiments; Carroll et al., 2014
-\cite{carroll.qc} studied the use of the Strand-Cross Correlation
-(SCC) (Kharchenko et al., 2008 \cite{strandcc}) and showed that by
-filtering blacklisted regions the estimation of the SCC is
-improved. However, using the SCC may not be helpful since the peaks
-that are attained at the read and fragment lengths are confused in a
-typical ChIP-exo SCC curve. Additionally this method requires to know
-blacklisted regions in advance which may not be available. In our
-pipeline we propose an out-the-shelf analysis that contrasts the
-enrichment of the experiment against its library complexity.
+libraries generated by ChIP-Seq (Mahony et al., 2015
+\cite{exo_review}). Second, although there are roughly the same amount
+of reads in both strands, locally there may be more reads in one
+strand than in the other. Finally, most of current ChIP-Seq quality
+control (QC) guidelines (Landt et al., \cite{encode_qc}) may not be
+applicable on ChIP-exo, while there are not established QC pipelines
+for ChIP-exo; previous ChIP-exo analyses used ChIP-Seq samples to
+compare the resolution between experiments (\cite{exo1},
+\cite{exoillumina}, \cite{exo2})). To address these challenges, we
+suggest a collection of quality control visualizations to interrogate
+these biases in a ChIP-exo experiment and globally asses the
+enrichment and library complexity of a ChIP-exo sample. We gathered
+ChIP-exo data from diverse organisms: CTCF factor in human
+\cite{exo1}; ER factor in human and FoxA1 factor in mouse (Serandour
+et al., 2013 \cite{exoillumina}); Glucocorticoid receptor (GR) in
+IMR90, K562 and U20S cell lines (Starick et al., 2015
+\cite{starick15}).Furthermore, we also generated ChIP-exo and ChIP-Seq
+data for $\sig$ factor in \emph{Escherichia Coli} (\emph{E. Coli})
+measured under aerobic ($ + O_2$) condition, and treated by rifampicin
+by 0 and 20 minutes (courtesy of Professor Robert Landick's lab).
 
 In order to obtain the potential benefits of ChIP-exo on protein
-binding site identification, it is critical to understand which are
-the important characteristics of ChIP-exo data and to use algorithms
-that could fully utilize information available in ChIP-exo data. Rhee
-and Pugh, 2011 \cite{exo1} discussed that reads in the forward and
-reverse strand might construct peak pairs around bound proteins, of
-which heights were implicitly assumed to be symmetric. Hence, they
-used the ``peak pair method'' that predicts the midpoint of two modes
-of peak pairs as potential binding sites. Mace (Wang et al, 20114
-\cite{mace}), CexoR (Madrigal, 2015 \cite{cexor}) and Peakzilla
-(Bardet et al., 2013 \cite{peakzilla}), recently developed ChIP-exo
-data analysis methods are also based on this peak pair
-assumption. However, appropriateness of such assumption was not fully
-evaluated in the literature yet. Furthermore, it is still unknown
-which factors could affect protein binding site identification using
-ChIP-exo data. In order to address this problem, we investigated
-various aspects of ChIP-exo data by contrasting them with their
-respective ChIP-Seq experiments.
+binding site identification, it is critical to use algorithms that
+could fully utilize information available in ChIP-exo data. Rhee and
+Pugh, 2011 \cite{exo1} discussed that reads in the forward and reverse
+strand might construct peak pairs around bound proteins, of which
+heights were implicitly assumed to be symmetric. Based on this
+rationale, they used the ``peak pair method'' that predicts the
+midpoint of two modes of peak pairs as potential binding
+sites. Recently developed ChIP-exo data analysis methods, such as Mace
+(Wang et al, 20114 \cite{mace}), CexoR (Madrigal, 2015 \cite{cexor})
+and Peakzilla (Bardet et al., 2013 \cite{peakzilla}), are also based
+on this peak pair assumption. However, appropriateness of such
+assumption was not fully evaluated in the literature yet. Furthermore,
+it is still unknown which factors could affect protein binding site
+identification using ChIP-exo data. In order to address this problem,
+we investigated various aspects of ChIP-exo data by contrasting them
+with their respective ChIP-Seq experiments.
 
@@
 \cite{macs}), CisGenome (Ji et al., 2008 \cite{cisgenome}) and MOSAiCS
-(Kuan et al., 2009 \cite{mosaics}), these approaches reveal protein
-binding sites in lower resolution, i.e., at an interval of hundreds to
-thousands of base pairs. Furthermore, they report only one ``mode'' or
-``predicted binding location'' per peak. Hence, these methods are not
-appropriate to evaluate the potential of ChIP-exo data for high
-resolution identification of protein binding sites. More recently,
+(Kuan et al., 2009 \cite{mosaics}), these approaches might not fully
+utilize potentials of ChIP-exo data for high resolution identification
+of protein binding sites. Specifically these approaches reveal protein
+binding sites only in lower resolution, i.e., at an interval of
+hundreds to thousands of base pairs. Furthermore, they implicitly
+assume that there is only one ``mode'' or ``predicted binding
+location'' per this wide genomic interval. More recently,
 deconvolution algorithms such as Deconvolution (Lun et al., 2009
@@
 appropriate for fair comparison between ChIP-exo and ChIP-Seq. To
-address these limitations, we developed an improved of dPeak (Chung et
-al., 2013 \cite{dpeak}), a high resolution binding site identification
-(deconvolution) algorithm that we previously developed for PE and SE
-ChIP-Seq data, so that it can also handle ChIP-exo data. The dPeak
-algorithm implements a probabilistic model that accurately describes
-the ChIP-exo and ChIP-Seq data generation process.
-
-In this work we demonstrate that the ``peak pair'' assumption of Rhee
-and Pugh, 2013 \cite{exo2} does not hold well in real ChIP-exo
-data. Furthermore, we found that when we analyze ChIP-exo data from
-eukaryotic genomes, it is important to consider sequence biases
-inherent to ChIP-exo data, such as mappability and GC content in order
-to improve sensitivity and specificity of binding site
-identification. We evaluated several methods to identify binding events
-and dPeak performs competitively respect to GEM and MACE when
-analyzing ChIP-exo data. More importantly, when comparable number of
-reads is used for both ChIP-exo and ChIP-Seq , dPeak coupled with
-ChIP-exo data provides resolution comparable to PE ChIP-Seq and both
-significantly improve the resolution of protein binding identification
-compared to SE-based analysis with any of the available methods.
+address these limitations, we developed and utilized an improved
+version of dPeak (Chung et al., 2013 \cite{dpeak}), a high resolution
+binding site identification (deconvolution) algorithm that we
+previously developed for PE and SE ChIP-Seq data, so that it can also
+handle ChIP-exo data. The dPeak algorithm implements a probabilistic
+model that accurately describes the ChIP-exo and ChIP-Seq data
+generation process.
+
+Some of the key findings in this work are as follows. First, we
+demonstrate that the ``peak pair'' assumption of Rhee and Pugh, 2013
+\cite{exo2} does not hold well in real ChIP-exo data. Second, we found
+that when we analyze ChIP-exo data from eukaryotic genomes, it is
+important to consider sequence biases inherent to ChIP-exo data, such
+as mappability and GC content, in order to improve sensitivity and
+specificity of binding site identification. Third, we evaluated
+several methods to identify binding events and dPeak performs
+competitively respect to GEM and MACE when analyzing ChIP-exo
+data. Finally, when comparable number of reads is used for both
+ChIP-exo and ChIP-Seq , dPeak coupled with ChIP-exo data provides
+resolution comparable to PE ChIP-Seq and both significantly improve
+the resolution of protein binding identification compared to SE-based
+analysis with any of the available methods.
 
@@
   \caption{\textbf{Description of ChIP-exo, SE ChIP-Seq and PE
-      ChIP-Seq:} The DNA is sonicated, for ChIP-exo a exonuclease
-    enzyme (orange hexagon) trims the $5\prime$ ends of each DNA
-    fragment to a fixed distance from the bound protein, finally is
-    subjected to Immunoprecipitation and amplification. For both
-    ChIP-exo and SE ChIP-Seq an adapter is ligated (green triangles)
-    at the $5\prime$ ends, while for PE ChIP-Seq is ligated to both
-    ends.}
+      ChIP-Seq:} A TF is bound to the forward (red) and backward
+    (blue) strand of the DNA. Then, is sonicated: For ChIP-exo a
+    exonuclease enzyme (orange hexagon) trims the $5\prime$ ends of
+    each DNA fragment to a fixed distance from the bound protein,
+    finally is subjected to Immunoprecipitation and amplification. For
+    both ChIP-exo and SE ChIP-Seq an adapter is ligated (green
+    triangles) at the $5\prime$ ends, while for PE ChIP-Seq is ligated
+    to both ends.}
   \label{fig:chip_diagram}
@@
 \caption{SCC curves for $\sig$ samples. The ``phantom peak'' and the
-    summit that corresponds to the read and fragment length
-    respectively are confounded due to the exonuclease digestion.}
+  summit that corresponds to the read and fragment length respectively
+  are confounded due to the exonuclease digestion.}
   \label{fig:scc_exo}

### 2016-03-16 rmarkdown/paper.Rnw
added qc definiions to methods and also updated the section on chip-seq qc metrics applied to chip-exo

@@
 
-\subsection{Deeply sequence E. Coli $\sig$ ChIP-exo and ChIP-Seq data}
+\subsection{Deeply sequenced E. Coli $\sig$ ChIP-exo and ChIP-Seq data}
 
 $\sig$ factor is a transcription initiation factor of housekeeping
-genesin E. Coli. In this organism's genomes, many promoters contain
+genes in E. Coli. In this organism's genomes, many promoters contain
 multiple transcription start sites (TSS) and these TSS are often
@@
 \cite{regulondb}. Therefore, investigation of ChIP-exo's potential for
-identification and differentiation of closely spaced binding sites are
+identification and differentiation of closely spaced binding sites is
 invaluable for elucidating the transcriptional networks of prokaryotic
@@
 
-\subsection{Current ChIP-Seq guidelines and quality metrics on ChIP-exo data}
-
-We started our exploration by investigating the suitability of the
-current state-of-the-art QC pipelines for ChIP-Seq for ChIP-exo. Table
-\ref{tab:qc} contains measures that are commonly calculated for
-ChIP-Seq samples: PCR Bottleneck Coefficient (PBC) and the Normalized
-Strand Cross-Correlation (NSC) are calculated as in \cite{encode_qc}.
-Additionally, we calculated the Forward Strand Ratio (FSR) to compare
-the number of fragments sequenced from each strand.
+\subsection{Application of current ChIP-Seq QC guidelines on ChIP-exo
+  data}
+
+We started our exploration by investigating whether the current
+state-of-the-art QC pipelines for ChIP-Seq are suitable for
+ChIP-exo. Table \ref{tab:qc} contains measures that are commonly
+calculated for ChIP-Seq samples: PCR Bottleneck Coefficient (PBC) and
+the Normalized Strand Cross-Correlation (NSC) are calculated as in
+\cite{encode_qc}. We omitted the Relative Strand Cross-Correlation
+(RSC) which is another commonly used QC measure because a typical
+ChIP-exo experiment is not accompanied by an input sample.
 
@@
   A[,files := NULL]
+  A[,nreads := prettyNum(nreads,big.mark = ",")]
+  A[,pbc := round(pbc,4)]
+  A[,nsc := round(nsc,4)]
   setnames(A,names(A),c("Organism","IP/TF","Condition/Cell","Rep.","Depth","PBC","NSC"))
 @
+
 <<table2,include = TRUE,echo = FALSE, eval = TRUE >>= 
@@
 @ 
-\caption{Usual quality control indicators applied to the gathered
-  ChIP-exo samples.  PBC stands for PCR Bottleneck Coefficient (0 -
-  0.5 is severe bottlenecking, 0.5 - 0.8 is moderate bottlenecking,
-  0.8 - 0.9 is mild bottlenecking, while 0.9 - 1 is no bottlenecking),
-  NSC for Normalized Strand Cross-Correlation and FSR for Forward
-  Strand Ratio. We omitted the Relative Strand Cross-Correlation (RSC)
-  because a typical ChIP-exo experiment is not accompanied by an input
-  file. $\sig$ samples were given by Dr. Landick's Lab, FoxA1 and ER
-  samples are from Serandour et al., 2013 \cite{exoillumina} and CTCF
-  sample is from Rhee and Pugh, 2011 \cite{exo1}}
+\caption{\textbf{Current QC metrics applied to ChIP-exo data.} PBC
+  stands for PCR Bottleneck Coefficient and NSC for Normalized Strand
+  Cross-Correlation.}
   \label{tab:qc} %% need to make this table
@@
 
-It is of special importance to notice that the values of PBC are quite
-low, hence it would be necessary to repeat these experiments by
-following ChIP-seq guidelines. On the other hand, in a good ChIP-Seq
-experiment the SCC is maximized at the fragment length, and a
-``phantom peak'' would be also present at the experiment's read length
-(\cite{encode_qc} and \cite{strandcc}). However, in ChIP-exo's case
-these two peaks are confounded. Hence an enrichment measure based in
-SCC such as the NSC is harder to interpret. In Figure
-\ref{fig:scc_exo} we observe that both local maxima are hard to
-differentiate in the SCC curves for the $\sig$ ChIP-exo datasets used to
-calculate the NSC values in Table \ref{tab:qc}.
-
-% % According to ENCODE's quality metrics, NSC values close
-% % to one corresponds to low enrichment samples or bad quality datasets
+DNA libraries generated by the ChIP-exo protocol seem to be less
+complex than the libraries generated by ChIP-Seq, since the possible
+number of positions to which the reads can be aligned is being reduced
+due to the exonuclease digestion, hence considerable amounts of reads
+are being mapped to specific positions. This affects the
+interpretation of the PBC, since for ChIP-Seq low PBC values indicate
+that the same read has been copied by the amplification process and
+aligned multiple times to the same position; while for ChIP-exo when
+several reads are aligned to the same position are not necessarily the
+same read amplified, but several reads that their $5\prime$ end was
+digested to the same position before the amplification step. It is of
+special importance to notice that for several ChIP-exo datasets, the
+PBC values are quite low. Therefore, by blindly following ChIP-Seq
+guidelines those experiment could have been considered as not useful
+and repeated.
+
+The Strand Cross-Correlation (SCC) introduced by Kharchenko et al.,
+2008 \cite{strandcc} is the most commonly used quality measure in
+ChIP-Seq. It is calculated as the correlation between both strand
+coverages, where each one is shifted $\delta / 2$ bp towards the
+$3\prime$ direction. In general it measures how well the reads mapped
+to each strand are clustered around the locations where the proteins
+are binding to the DNA, and usually it is expected to observed two
+local maxima, one when the profiles are shifted by the average read
+length and another when the profiles are shifted by the unobserved
+fragment length. In a good ChIP-Seq dataset the last one is also the
+SCC global maxima. However, in ChIP-exo's case these two peaks are
+confounded. Hence the Normalized Strand Cross-Correlation (NSC) which
+is a measure based on the SCC is harder to interpret. Figure
+\ref{fig:scc_exo} shows that both local maxima are hard to
+differentiate in the SCC curves for the $\sig$ ChIP-exo datasets used
+to calculate the NSC values in Table \ref{tab:qc}.
 
@@
 number of reads mapping to each of \Sexpr{bin_size} non-overlapping
-window after extending reads by \Sexpr{frag_len} to their 3' end
-directions. ChIP tag counts in ChIP-exo data were linearly related to
-ChIP tag counts in ChIP-Seq data for the regions with high ChIP tag
+window after extending reads by \Sexpr{frag_len} to their $3\prime$
+end directions. ChIP tag counts in ChIP-exo data were linearly related
+to ChIP tag counts in ChIP-Seq data for the regions with high ChIP tag
 counts (Figure \ref{fig:comp}A). This implies that signals for
@@
 
+\subsection*{Definition of current ChIP-Seq QC guidelines.}
+
+\subsubsection*{PCR Bottleneck Coefficient.}
+
+The PCR Bottleneck coefficient is a measure of library complexity in
+ChIP-Seq data:
+
+\begin{align}
+  \mbox{PBC} = \frac{\text{Nr. of positions to which exactly one
+      unique mapping read is aligned}}{\text{Nr. of positions to
+      which at least one unique mapping read is aligned}} \nonumber
+\end{align}
+
+For human and mouse genome, the ENCODE project states that a PBC value
+in the 0 - 0.5 range indicates severe bottlenecking, in the 0.5 - 0.8
+range moderate bottlenecking, in the 0.8 - 0.9 range indicates mild
+bottlenecking and in the 0.9 - 1 range indicates that there is no
+presence of bottlenecking.
+
+\subsubsection*{Strand Cross-Correlation.}
+
+The strand cross-correlation was proposed by Kharchenko et al., 2008
+\cite{strandcc} and it may be one of the most used of the ChIP-Seq QC
+metrics. The SCC curve is defined as:
+
+\begin{align}
+  y(\delta) = \sum_c w_c r\left[ n_c^+ \left(x + \frac{\delta}{2}
+    \right), n_c^- \left( x- \frac{\delta}{2} \right)\right],
+\label{scc}
+\end{align}
+
+where $y(\delta)$ is the SCC for a strand shift $\delta$, $r$ is the
+Pearson correlation, $w_c$ is the proportion of reads mapped to
+chromosome $c$ and $n_c^S$ is the read count vector for strand $S$ and
+chromosome $c$. Additionally, two QC metrics are defined:
+
+\begin{align}
+  \mbox{NSC} &= \frac{\max_\delta y(\delta)}{\min_\delta y(\delta)}, \label{nsc} \\
+  \mbox{RSC} &= \frac{\max_\delta y(\delta) -
+    y_{\text{bgd}}}{y_{\text{rl}} - y_{\text{bgd}}}. \label{rsc}
+\end{align}
+
+where $y_{\text{bgd}}$ is the background SCC level and $y_{\text{rl}}$
+is the SCC ``phantom peak'' value.
+
 \subsection*{ChIP-Seq quality control metrics and Strand
@@
     Peakzilla methods. Resolution is defined as the minimum distance
-    between a RegulonDB annotation an a predicted binding event.}
+    between a RegulonDB annotation and a predicted binding event.}
   \label{fig:methods_comp}

### 2016-03-16 rmarkdown/paper.tex
added qc definiions to methods and also updated the section on chip-seq qc metrics applied to chip-exo

@@
 ##     Organism  IP/TF Condition/Cell Rep.       Depth    PBC      NSC
-##  1:   E.Coli  Sig70        Aerobic    1  13961493 0.13994203 103.423876
-##  2:   E.Coli  Sig70        Aerobic    2  14810838 0.16339252 162.800188
-##  3:   E.Coli  Sig70      Anaerobic    1  16108774 0.13535422 153.508809
-##  4:   E.Coli  Sig70      Anaerobic    2  13636541 0.15322537 172.581549
-##  5:   E.Coli  Sig70       Rif-0min    1    902921 0.26895580  13.769077
-##  6:   E.Coli  Sig70       Rif-0min    1   1852124 0.25908459  17.918825
-##  7:   E.Coli  Sig70      Rif-20min    2   2104427 0.25841987  29.608305
-##  8:   E.Coli  Sig70      Rif-20min    2  11548572 0.15107930  13.086255
-##  9:    Human   CTCF           HeLA    1  48478450 0.45796889  16.024817
-## 10:    Mouse Fox A1    Mouse Liver    1  22210461 0.65622704  21.281972
-## 11:    Mouse Fox A1    Mouse Liver    2  23307557 0.79962169  60.421912
-## 12:    Mouse Fox A1    Mouse Liver    3  22421729 0.10683407  72.042428
-## 13:    Human     GR          IMR90    1  47443803 0.29785605   8.860176
-## 14:    Human     GR           K562    1 116518000 0.05047634   4.117883
-## 15:    Human     GR           U20S    1   3255111 0.77142141  10.058834
-## 16:    Human     ER          MCF-7    1   9289835 0.80825680  19.875232
-## 17:    Human     ER          MCF-7    2  11041833 0.80241200  21.485094
-## 18:    Human     ER          MCF-7    3  12464836 0.82035947  18.728062
+##  1:   E.Coli  Sig70        Aerobic    1  13,961,493 0.1399 103.4239
+##  2:   E.Coli  Sig70        Aerobic    2  14,810,838 0.1634 162.8002
+##  3:   E.Coli  Sig70      Anaerobic    1  16,108,774 0.1354 153.5088
+##  4:   E.Coli  Sig70      Anaerobic    2  13,636,541 0.1532 172.5815
+##  5:   E.Coli  Sig70       Rif-0min    1     902,921 0.2690  13.7691
+##  6:   E.Coli  Sig70       Rif-0min    1   1,852,124 0.2591  17.9188
+##  7:   E.Coli  Sig70      Rif-20min    2   2,104,427 0.2584  29.6083
+##  8:   E.Coli  Sig70      Rif-20min    2  11,548,572 0.1511  13.0863
+##  9:    Human   CTCF           HeLA    1  48,478,450 0.4580  16.0248
+## 10:    Mouse Fox A1    Mouse Liver    1  22,210,461 0.6562  21.2820
+## 11:    Mouse Fox A1    Mouse Liver    2  23,307,557 0.7996  60.4219
+## 12:    Mouse Fox A1    Mouse Liver    3  22,421,729 0.1068  72.0424
+## 13:    Human     GR          IMR90    1  47,443,803 0.2979   8.8602
+## 14:    Human     GR           K562    1 116,518,000 0.0505   4.1179
+## 15:    Human     GR           U20S    1   3,255,111 0.7714  10.0588
+## 16:    Human     ER          MCF-7    1   9,289,835 0.8083  19.8752
+## 17:    Human     ER          MCF-7    2  11,041,833 0.8024  21.4851
+## 18:    Human     ER          MCF-7    3  12,464,836 0.8204  18.7281
 \end{verbatim}
@@
 \end{knitrout}
-\caption{Usual quality control indicators applied to the gathered
-  ChIP-exo samples.  PBC stands for PCR Bottleneck Coefficient (0 -
-  0.5 is severe bottlenecking, 0.5 - 0.8 is moderate bottlenecking,
-  0.8 - 0.9 is mild bottlenecking, while 0.9 - 1 is no bottlenecking),
-  NSC for Normalized Strand Cross-Correlation and FSR for Forward
-  Strand Ratio. We omitted the Relative Strand Cross-Correlation (RSC)
-  because a typical ChIP-exo experiment is not accompanied by an input
-  file. $\sig$ samples were given by Dr. Landick's Lab, FoxA1 and ER
-  samples are from Serandour et al., 2013 \cite{exoillumina} and CTCF
-  sample is from Rhee and Pugh, 2011 \cite{exo1}}
+\caption{\textbf{Current QC metrics applied to ChIP-exo data.} PBC
+  stands for PCR Bottleneck Coefficient and NSC for Normalized Strand
+  Cross-Correlation.}
   \label{tab:qc} %% need to make this table
@@
 number of reads mapping to each of 150 non-overlapping
-window after extending reads by 150 to their 3' end
-directions. ChIP tag counts in ChIP-exo data were linearly related to
-ChIP tag counts in ChIP-Seq data for the regions with high ChIP tag
+window after extending reads by 150 to their $3\prime$
+end directions. ChIP tag counts in ChIP-exo data were linearly related
+to ChIP tag counts in ChIP-Seq data for the regions with high ChIP tag
 counts (Figure \ref{fig:comp}A). This implies that signals for

### 2016-03-16 rmarkdown/paper.R

@@ -39,6 +39,9 @@
   A[,nsc := unlist(nsc)]
   setcolorder(A, c("Organism","TF","cond","Replicate","nreads","pbc","nsc","files"))
   A[,files := NULL]
+  A[,nreads := prettyNum(nreads,big.mark = ",")]
+  A[,pbc := round(pbc,4)]
+  A[,nsc := round(nsc,4)]
   setnames(A,names(A),c("Organism","IP/TF","Condition/Cell","Rep.","Depth","PBC","NSC"))
 
 ## ----table2,include = TRUE,echo = FALSE, eval = TRUE---------------------

### 2016-03-17 rmarkdown/paper.Rnw
added mappability and gc content scores part to methods and added the corresponding bibliography entries

@@
 to ChIP tag counts in ChIP-Seq data for the regions with high ChIP tag
-counts (Figure \ref{fig:comp}A). This implies that signals for
-potential binding sites are well reproducible between ChIP-exo and
-ChIP-Seq data. On the other hand, there was clear difference in the
-background distribution between them. In ChIP-Seq data reads were
+counts (Upper part of Figure \ref{fig:comp}A). This implies that
+signals for potential binding sites are well reproducible between
+ChIP-exo and ChIP-Seq data. On the other hand, there was clear
+difference in the background distribution between them (lower part of
+Figure \ref{fig:comp}A). Specifically, in ChIP-Seq data reads were
 almost uniformly distributed over background (non-binding) regions and
@@
 usually paired with a peak of reads in the reverse strand that is
-located in the other site of the binding site. Wang et al., 2014
+located in the other site of the binding site. Note that currently
+available ChIP-exo data analysis methods, such as Wang et al., 2014
 \cite{mace}, Madrigal 2015 \cite{cexor} and Bardet et al., 2013
-\cite{peakzilla} proposed method rely in this assumption. In order to
-evaluate this assumption, we reviewed the proportion of reads in the
-forward strand in candidate regions (i.e. regions with at least one
-binding site) in $\sig$ ChIP-exo data. We found that strands of reads
-were much less balanced in ChIP-exo data than in ChIP-Seq data in
-these regions with potential binding sites (Fig. \ref{fig:comp}B) and
-this indicates that the peak pair assumption might not hold in real
+\cite{peakzilla} rely on this assumption. In order to evaluate this
+assumption, we reviewed the proportion of reads in the forward strand
+in ChIP-exo peaks such as at least one binding site is predicted in
+both ChIP-exo and ChIP-Seq data. We found that strands of reads were
+much less balanced in ChIP-exo data than in ChIP-Seq data in these
+regions with potential binding sites (Fig. \ref{fig:comp}B) and this
+indicates that the peak pair assumption might not hold in real
 ChIP-exo data.
@@
 ChIP tag count increases and for GC - content greater than 0.6 it
-shows a decreasing trend.These results indicate that binding site
-identification in ChIP-exo sample might also benefit from the use of
-methods that take into account of apparent sequence biases such as
-mappability and GC content.
+shows a decreasing trend.
+
+These results indicate that binding site identification in ChIP-exo
+sample might also benefit from the use of methods that take into
+account of apparent sequence biases such as mappability and GC
+content.
 
@@
 \begin{align*}
-  \text{ARC} &= \dfrac{\text{Nr. of reads in the region}}{\text{Width of the region}} \\
+  \text{ARC} &= \dfrac{\text{Nr. of reads in the region}}{\text{Width of the region}}, \\
   \text{URCR} &= \dfrac{\text{Nr. of reads in the region mapped to
-      exactly one position}}{\text{Nr. of reads in the region}}
+      exactly one position}}{\text{Nr. of reads in the region}}.
 \end{align*}
@@
 
+The statistics in Table \ref{tab:qc} and the SCC curves from Figure
+\ref{fig:scc_exo} were calculated with the \textbf{ChIPUtils} package
+version 0.99.0, available in
+\url{https://github.com/welch16/ChIPUtils}.
+
 \subsubsection*{PCR Bottleneck Coefficient.}
@@
 
-\subsection*{ChIP-Seq quality control metrics and Strand
-  Cross-Correlation.}
-
-The statistics in Table \ref{tab:qc} and the SCC curves from Figure
-\ref{fig:scc_exo} where calculated with the \textbf{ChIPUtils} package
-version 0.99.0, available in
-\url{https://github.com/welch16/ChIPUtils}.
+\subsection*{Mappability and GC content scores.}
+
+To define the mappability score we follow the definition from Rozowsky
+et al., 2009 \cite{peakseq}:
+
+\begin{align}
+  m_i &= \sum_{k = i - L +1}^{i + L - 1} \frac{\delta_k }{2L - 1}.
+\end{align}
+
+where $\delta_i$ is the indicator if the base at coordinate $i$ can be
+mapped uniquely by a 32 bp sequence at position $i$, and $L$ is the
+expected fragment length. GC - content score is defined analogously,
+where $\delta_i$ represents the ocurrence of a G or C at the $i$-th
+position in the genome.
+
+The mappability and GC - content scores for a bin are defined as the
+average of the scores across the nucleotides in the bin.
 
@@
 \includegraphics[width = .8\textwidth]{/p/keles/ChIPexo/volume3/ChIPexo/figs/for_paper/EColi_strand_cross_corr.pdf}
-\caption{SCC curves for $\sig$ samples. The ``phantom peak'' and the
-  summit that corresponds to the read and fragment length respectively
-  are confounded due to the exonuclease digestion.}
+\caption{\textbf{SCC curves for $\sig$ samples.} The ``phantom peak''
+  and the summit that corresponds to the read and fragment length
+  respectively are confounded due to the exonuclease digestion.}
   \label{fig:scc_exo}

### 2016-03-17 rmarkdown/paper.Rnw
updated GC content + mappability arguments

@@
 mappability and GC content. Each data point is obtained by averaging
-the read counts across bins with the same mappability of GC
+the read counts across bins with the same mappability of GC -
 content. In Figure \ref{fig:comp}C it is shown that the ChIP-exo tag
@@
 ChIP tag count increases and for GC - content greater than 0.6 it
-shows a decreasing trend.
-
-These results indicate that binding site identification in ChIP-exo
-sample might also benefit from the use of methods that take into
-account of apparent sequence biases such as mappability and GC
-content.
+shows a decreasing trend. Kuan et al., 2011 \cite{mosaics} studied the
+presence of the mappability and GC - content biases in ChIP-Seq's
+background. It is not surprising to see these biases also present in
+ChIP-exo data, since ChIP-Seq and ChIP-exo signal seems to be linearly
+correlated for enriched regions (Figure \ref{fig:comp}A). Rozowsky et
+al., 2009 \cite{peakseq} and Valouev et al., 2008 \cite{quest} provide
+in depth analysis of the mappability and GC - content biases for
+ChIP-Seq respectively. Finally, these results indicate that binding
+site identification for ChIP-exo data benefits from using methods that
+take into account of apparent sequence biases such as mappability and
+GC content.
 
@@
 expected fragment length. GC - content score is defined analogously,
-where $\delta_i$ represents the ocurrence of a G or C at the $i$-th
+where $\delta_i$ represents the occurrence of a G or C at the $i$-th
 position in the genome.

### 2016-03-21 rmarkdown/paper.Rnw
Qc pipeline section finished

@@
 Figure \ref{fig:qcdiagram} shows a flowchart for the ChIP-exo QC
-pipeline. Which basically partitions the genome by keeping the
-non-digested ChIP-exo regions. Then, for each region calculates a
-series of summary statistics. Finally it creates several
-visualizations designed to diagnose the quality level of ChIP-exo
-sample.
-
-\subsubsection{Enrichment and library complexity in ChIP-exo data}
-
-In ChIP-exo experiments, the background is often digested by the
-exonuclease enzyme, therefore to determine the sample's quality is
-necessary to address the balance between the enrichment and library
-complexity of an experiment. To diagnose this, we considered the
-Average Read Coverage (ARC) and the Unique Read Coverage Rate (URCR)
-which are defined as follows:
+pipeline. In the first step, we partition the genome by keeping the
+non-digested ChIP-exo regions. Then, for each region it counts the
+number of fragments that compose the region and the number of
+positions to which the reads are being mapped to in each strand. With
+these values it calculates the following summary statistics:
 
@@
   \text{URCR} &= \dfrac{\text{Nr. of reads in the region mapped to
-      exactly one position}}{\text{Nr. of reads in the region}}.
+      exactly one position}}{\text{Nr. of reads in the region}}, \\
+  \text{FSR} &= \dfrac{\text{Nr. of fwd. strand reads in the
+      region}}{\text{Number of reads in the region}}.
 \end{align*}
+
+Finally it creates several visualizations designed to diagnose the
+quality level of a ChIP-exo sample. Figure \ref{fig:qcdiagram}A shows
+the typical behavior of the ARC vs. URCR plot. In general, the plot
+depicts two strong arms: One on the left with low ARC values and
+varying URCR which corresponds to ChIP-exo's background, regions that
+are usually composed by scattered reads that were not digested during
+the exonuclease step; and another one where the URCR decreases as the
+ARC increases, which corresponds to regions that are usually enriched
+and as the URCR decreases the library complexity does it as well, on
+the other hand high URCR values correspond to regions composed by
+position with relatively few fragments aligned to. Figures
+\ref{fig:qcdiagram}B and \ref{fig:qcdiagram}C analyze the strand
+imbalance bias in a ChIP-exo experiment: The first one depicts how
+quickly the regions exclusively formed by fragments in one strand are
+being filtered out as regions with higher depth are observed; and the
+second one shows how quickly the FSR's distribution approach the
+median, since in a high quality sample it is expected for the median
+to be approximately 0.5 and the enriched regions are going to be
+composed by fragments sequenced from both strands.
+
+The ChIP-exo QC pipeline provides a model free framework to analyze
+the biases in a ChIP-exo experiment by taking advantage of the
+exonuclease enzyme that digests the non-enriched regions to partition
+the genome, calculates common ChIP-Seq QC metrics in ChIP-exo regions
+locally and allows the interpretation of these metrics by the use of
+diagnostic plots.
+
+\subsubsection{Enrichment and library complexity in ChIP-exo data}
+
+In ChIP-exo experiments, background fragments are often digested by
+the exonuclease enzyme, therefore the balance between the enrichment
+and library complexity of an experiment is a key factor determining
+the sample's data quality.
 
@@
 complexity and experiment enrichment. In Figure \ref{fig:enrich}A we
-can observe both statistics for the regions that are common for the
-three replicates. This figure shows typical patterns for ChIP-exo
-experiment, there are two strong arms: The one on the left with low
-ARC and varying URCR corresponds to ChIP-exo's background, usually
-regions composed by scattered reads that were not digested during the
-exonuclease step; and the one on the right that corresponds to regions
-that are usually enriched, for these regions the URCR measures how the
-fragments are allocated into the possible positions in a region. Hence
-we would expect the first replicate to have have more enriched regions
-and the third replicate's library complexity to be lower than the
-other two replicates library complexities. To verify this statements,
-we extracted the sequences around high confidence binding events and
-look for the FoxA1 motif using FIMO \cite{fimo}. Figure
-\ref{fig:enrich}B shows the number of candidate regions, which shows
-that the first replicate is being allocated into more enriched regions
-than the other ones. Figure \ref{fig:enrich}C shows that for the first
-and third replicates, the FoxA1 motif is being detected in roughly the
-same proportion of sites, and finally in \ref{fig:enrich}D we observe
-that the first and third replicate can detect the FoxA1 motif with the
-same significance, while the second replicate does not.
+present ARC vs. URCR plots for all three replicates. As a case of
+study, we compare the three plots to differentiate the quality of the
+three experiment. Hence, this might imply that the first replicate to
+have have more enriched regions and the third replicate's library
+complexity to be lower than the other two replicates library
+complexities. To verify this statement, we extracted the sequences
+around high confidence binding events and look for the FoxA1 motif
+using FIMO \cite{fimo}. Figure \ref{fig:enrich}B shows the number of
+candidate regions, which shows that the first replicate is being
+allocated into more enriched regions than the other ones. Figure
+\ref{fig:enrich}C shows that for the first and third replicates, the
+FoxA1 motif is being detected in roughly the same proportion of sites,
+and finally in \ref{fig:enrich}D we observe that the first and third
+replicate can detect the FoxA1 motif with the same significance, while
+the second replicate does not.
 
@@
 region) as the regions with lower depth are being filtered out. This
-indicator is of particular importance, since several methods rely on
-the ``peak-pair'' assumption. For every ChIP-exo experiment, we
-calculated the global FSR and noticed that for all experiments is
-roughly $0.5$, which means there are roughly the same amount of reads
-in both strands. However figure \ref{fig:strand} shows that the global
-FSR does not represent the experiment's local strand imbalance, hence
-the ``peak pair'' assumption may not hold well in real ChIP-exo data.
+indicator is of particular importance, as it evaluates the ``peak
+pair'' assumption that the original ChIP-exo paper suggested and
+multiple ChIP-exo data analysis methods rely on. For every ChIP-exo
+experiment, we calculated the global FSR and noticed that for all
+experiments is roughly $0.5$, which means there are roughly the same
+amount of reads in both strands.
 
@@
 the FSR's behavior as the lower depth regions are being filtered out,
-while B) shows which percentage of the regions are composed by reads
-in both strand or only one (forward or backward). In a good data set,
-it would be expected that all quantiles shown to be quickly converging
-towards the median (in panel A) or the regions composed of reads in
-one strand being made of few fragments (in panel B). For each
-replicate, we divided the partitioned regions by asking whether they
-overlap a set of high quality ChIP-exo peaks, and then we tested
-(using the Wilcoxon rank sum test over an imbalance index) if the
-strand imbalance's distribution is the same for both classes. For
-regions composed by a higher amount of reads, it is harder to
-distinguish their peaks by considering only the strand imbalance,
-hence in a better quality ChIP-exo experiment it is easier to
-distinguish enriched regions by the amount of reads in both
-strands. Additionally we may consider that the background of a
-ChIP-exo experiment it is more unbalanced than the enriched regions.
+while Figure \ref{fig:strand}B) shows which percentage of the regions
+are composed by reads in both strand or only one (forward or
+backward). In a good data set, it would be expected that all quantiles
+shown to be quickly converging towards the median (in panel A) or the
+regions composed of reads in one strand being made of few fragments
+(in panel B). For each replicate, we divided the partitioned regions
+by asking whether they overlap a set of high quality ChIP-exo peaks,
+and then we tested (using the Wilcoxon rank sum test over the
+imbalance index defined in the methods section) if the strand
+imbalance's distribution is the same for both classes. For regions
+composed by a higher amount of reads, it is harder to distinguish
+their peaks by considering only the strand imbalance, hence in a
+better quality ChIP-exo experiment it is easier to distinguish
+enriched regions by the amount of reads in both strands. Similarly, we
+may consider that the strands of reads for background of a ChIP-exo
+experiment is more unbalanced than those for the enriched regions. In
+conclusion, Figure \ref{fig:strand} shows that the global FSR does not
+represent the experiment's local strand imbalance, hence the ``peak
+pair'' assumption does not hold locally in ChIP-exo data.
 
@@
 
-Figure \ref{fig:reso_all} shows different comparisons among ChIP-exo,
-PE ChIP-Seq and SE ChIP-Seq. A RegulonDB annotation (Salgado et al,
-2012 \cite{regulondb}) was considered identified if the distance
-between it and a dPeak binding site estimate was at most of
-$\Sexpr{id_dist}$ bp. That way, the sensitivity is defined as the
-proportion of RegulonDB annotations identified in a peak and the
-resolution is defined as the minimum distance between a RegulonDB
-annotation and the dPeak binding site estimates. The left panel of
-Figure \ref{fig:reso_all} shows that the sensitivity increases as the
-mean distance between binding events increases. Despite the when the
-binding events in a peak are closer to each other, both ChIP-exo and
-PE ChIP-Seq are comparable, as the distance increases ChIP-exo
-identifies a higher proportion of the RegulonDB annotations;
-additionally SE ChIP-Seq is significantly less sensitive than both
-ChIP-exo and PE ChIP-Seq. The right panel shows that ChIP-exo and PE
-ChIP-Seq are comparable in resolution, while both protocols
-significantly outperform SE ChIP-Seq.
+Figure \ref{fig:reso_all} shows comparisons among ChIP-exo, PE
+ChIP-Seq and SE ChIP-Seq. We considered the RegulonDB data as ground
+truth, since those are the most recent annotation on \emph{Escherichia
+  Coli}. A RegulonDB annotation (Salgado et al, 2012 \cite{regulondb})
+was considered to be identified if the distance from the closest dPeak
+binding site estimate was less that or equal to $\Sexpr{id_dist}$
+bp. That way, the sensitivity is defined as the proportion of
+RegulonDB annotations identified in a peak and the resolution is
+defined as the minimum distance between a RegulonDB annotation and the
+closest dPeak binding site estimate. Figure \ref{fig:reso_all}A shows
+that the sensitivity increases as the mean distance between binding
+events increases. When the average disntace is greater than 200 bp,
+dPeak identifies more than the 75\% of the binding events in each
+peak, this is intuitive as the mean ChIP-Seq's fragment length is
+shorter than 200 bp, hence the read doesn't contribute to more than
+one binding event \cite{dpeak}. When the binding events in a peak are
+closer to each other, both ChIP-exo and PE ChIP-Seq are comparable, as
+the distance increases ChIP-exo identifies a higher proportion of the
+RegulonDB annotations; additionally SE ChIP-Seq is significantly less
+sensitive than both ChIP-exo and PE ChIP-Seq. Figure
+\ref{fig:reso_all}B shows that ChIP-exo and PE ChIP-Seq are comparable
+in resolution, while both protocols significantly outperform SE
+ChIP-Seq.
 
@@
 
-Previously, ChIP-exo and SE ChIP-Seq have been compared at a fixed
-depth level in the literature, but this comparisons did not include PE
-ChIP-Seq. Hence, we sampled a fixed amount of reads for each of the
+Previously, ChIP-exo and SE ChIP-Seq have been compared only at a
+fixed depth level in the literature, while they did not include PE
+ChIP-Seq as well either. In order to address this limitation in
+previous studies, we sampled a fixed amount of reads for each of the
 ChIP-exo, PE ChIP-Seq and SE ChIP-Seq datasets of the $\sig$ samples
 ($N$ reads for both ChIP-exo experiment and $N/2$ or $N$ pairs for PE
-ChIP-Seq). For each sampled dataset we applied out lower-to-higher
-resolution pipeline by calling peaks with MOSAiCS \cite{mosaics} and
-then deconvolving the binding events by using dPeak \cite{dpeak}. For
-the ChIP-exo datasets we called peaks by using the GC-content and
+ChIP-Seq to assume more realistic situation of a fixed cost). For
+each sampled dataset we applied out lower-to-higher resolution
+pipeline by calling peaks with MOSAiCS \cite{mosaics} and then
+deconvolving the binding events by using dPeak \cite{dpeak}. For the
+ChIP-exo datasets we called peaks by using the GC-content and
 mappability models with MOSAiCS, and for the ChIP-Seq datasets we used
@@
 Figures \ref{fig:design} shows the behavior of each data type in
-$\sig$ experiment under aerobic condition when their depth is
-fixed. It is remarkable that even when the number of candidate peaks
-or the number of predicted events is lower for ChIP-exo, it
-outperforms both PE and SE ChIP-Seq in number of identified targets
-and resolution.
-
-This may suggest that with ChIP-exo less positive peaks are being
-called and that when the targets are being identified, dPeak estimates
-binding locations closer to the true location. Additionally, we can
-see that as the read depth increases, all four indicator seem to
-stabilize and hit a plateau, which may indicate that with ChIP-exo a
-smaller amount of reads is necessary to identify a higher number of
-targets, but it may be also possible that this is an artifact occurring
-due to ChIP-exo's lower library complexity. Additionally, it is worth
-noting that for PE ChIP-Seq we sampled both ends of the fragment,
-hence for each sequencing depth we are sampling the half amount of
-pairs for PE ChIP-Seq than for ChIP-exo or SE ChIP-Seq.
+$\sig$ experiment under aerobic condition when comparable number of
+reads is used for all of ChIP-exo, SE ChIP-Seq and PE ChIP-Seq. It is
+remarkable that even when the number of candidate peaks or the number
+of predicted events is lower for ChIP-exo, it outperforms both PE and
+SE ChIP-Seq in number of identified targets and resolution.
+
+This may suggest that with ChIP-exo less false positive peaks are
+being called and that when the targets are being identified, dPeak
+estimates binding locations closer to the true location. Additionally,
+we can see that as the read depth increases, all four indicator seem
+to stabilize and hit a plateau earlier than the cases for ChIP-Seq,
+which may indicate that with ChIP-exo a smaller amount of reads is
+needed to identify the same number of targets than ChIP-Seq, but it
+may be also possible that this is an artifact occurring due to
+ChIP-exo's lower library complexity. Additionally, it is worth noting
+that for PE ChIP-Seq we sampled both ends of the fragment, hence for
+each sequencing depth we are sampling the half amount of pairs for PE
+ChIP-Seq than for ChIP-exo or SE ChIP-Seq.
 
@@
 the $\sig$ replicates with and without rif treatment. The left, middle
-and right columns shows the fixed depth against the number of
-predicted events, identified targets and resolution being compared at
-a fixed depth level. The behavior of this quantities seems to be
-opposite to the one as in figure \ref{fig:design}, hence we used the
-ChIP-exo QC pipeline in the fixed depth ChIP-exo experiments. Figure
+and right columns show the fixed depth against the number of predicted
+events, identified targets and resolution being compared at a fixed
+depth level. The behavior of this quantities seems to be opposite to
+the one as in Figure \ref{fig:design}, hence we used the ChIP-exo QC
+pipeline in the fixed depth ChIP-exo experiments. Figure
 \ref{fig:exoQC_sat_aero} shows hexbin of ARC vs URCR for several fixed
@@
 the sampled reads were aligned to enriched regions. On the other hand
-in figures \ref{fig:exoQC_sat1} to \ref{fig:exoQC_sat4}, we used the
+in Figures \ref{fig:exoQC_sat1} to \ref{fig:exoQC_sat4}, we used the
 ChIP-exo QC pipeline on the samples that are outperformed by PE and SE
@@
 version 0.99.0, available in
-\url{https://github.com/welch16/ChIPUtils}.
+\url{https://github.com/welch16/ChIPUtils}. This package provides an
+easy to use interface to calculate basic quality control metrics and
+diagnostic plots for ChIP-Seq data.
 
@@
 We used the R package \textbf{ChIPexoQual} to assess the quality of
-the ChIP-exo datasets by following the steps described in figure
+the ChIP-exo datasets by following the steps described in Figure
 \ref{fig:qcdiagram}. We used version 1.0, and it is available in
@@
 motif over a 10 bp window around the estimated binding events. We used
-FIMO's command line 4.9.1 version.
+FIMO's command line 4.9.1 version \cite{fimo}.
 
@@
   \centering
-  \includegraphics[width = \textwidth]{../figs/for_paper/coverage_diagram2.pdf}
-  \caption{Flowchart of the ChIP-exo quality control pipeline.}
+  \includegraphics[width = \textwidth]{../figs/for_paper/coverage_diagram3.pdf}
+  \caption{\textbf{Flowchart of the ChIP-exo quality control
+      pipeline.} A) ARC vs. URCR, this plot presents a global view of
+    the balance between library complexity and enrichment. There are
+    two arms: First, one with low ARC, which corresponds to regions
+    formed by few aligned positions; Second, where the URCR decreases
+    as the ARC increases. B) Min. depth vs Proportions of regions,
+    this plot shows the strand composition for all the regions formed
+    by a min. number of reads. Regions with low depth tend to be
+    formed by undigested reads from one unique strand, while regions
+    with higher signal are usually formed by reads in both strands. C)
+    Min depth vs. FSR, this plot depicts how quickly the FSR's
+    distribution approximate the median. In a high quality sample, the
+    median is around 0.5, and the other quantiles reach that value
+    quickly.}
   \label{fig:qcdiagram}

### 2016-03-21 rmarkdown/paper.tex
Qc pipeline section finished

@@
 
-Figure \ref{fig:reso_all} shows different comparisons among ChIP-exo,
-PE ChIP-Seq and SE ChIP-Seq. A RegulonDB annotation (Salgado et al,
-2012 \cite{regulondb}) was considered identified if the distance
-between it and a dPeak binding site estimate was at most of
-$20$ bp. That way, the sensitivity is defined as the
-proportion of RegulonDB annotations identified in a peak and the
-resolution is defined as the minimum distance between a RegulonDB
-annotation and the dPeak binding site estimates. The left panel of
-Figure \ref{fig:reso_all} shows that the sensitivity increases as the
-mean distance between binding events increases. Despite the when the
-binding events in a peak are closer to each other, both ChIP-exo and
-PE ChIP-Seq are comparable, as the distance increases ChIP-exo
-identifies a higher proportion of the RegulonDB annotations;
-additionally SE ChIP-Seq is significantly less sensitive than both
-ChIP-exo and PE ChIP-Seq. The right panel shows that ChIP-exo and PE
-ChIP-Seq are comparable in resolution, while both protocols
-significantly outperform SE ChIP-Seq.
+Figure \ref{fig:reso_all} shows comparisons among ChIP-exo, PE
+ChIP-Seq and SE ChIP-Seq. We considered the RegulonDB data as ground
+truth, since those are the most recent annotation on \emph{Escherichia
+  Coli}. A RegulonDB annotation (Salgado et al, 2012 \cite{regulondb})
+was considered to be identified if the distance from the closest dPeak
+binding site estimate was less that or equal to $20$
+bp. That way, the sensitivity is defined as the proportion of
+RegulonDB annotations identified in a peak and the resolution is
+defined as the minimum distance between a RegulonDB annotation and the
+closest dPeak binding site estimate. Figure \ref{fig:reso_all}A shows
+that the sensitivity increases as the mean distance between binding
+events increases. When the average disntace is greater than 200 bp,
+dPeak identifies more than the 75\% of the binding events in each
+peak, this is intuitive as the mean ChIP-Seq's fragment length is
+shorter than 200 bp, hence the read doesn't contribute to more than
+one binding event \cite{dpeak}. When the binding events in a peak are
+closer to each other, both ChIP-exo and PE ChIP-Seq are comparable, as
+the distance increases ChIP-exo identifies a higher proportion of the
+RegulonDB annotations; additionally SE ChIP-Seq is significantly less
+sensitive than both ChIP-exo and PE ChIP-Seq. Figure
+\ref{fig:reso_all}B shows that ChIP-exo and PE ChIP-Seq are comparable
+in resolution, while both protocols significantly outperform SE
+ChIP-Seq.
 

### 2016-03-21 rmarkdown/paper.Rnw
updated paper figures and fixed format

@@
 \usepackage{verbatim}
-\usepackage{hyperref}
-\hypersetup{
-     colorlinks   = true,
-     citecolor    = black
-}
-\hypersetup{linkcolor=black}
+% \usepackage{hyperref}
+% \hypersetup{
+%      colorlinks   = true,
+%      citecolor    = black
+% }
 % \hypersetup{linkcolor=black}
 % \hypersetup{linkcolor=black}
+% \hypersetup{linkcolor=black}
 
 \usepackage{wrapfig}
+
+\newcommand{\beginsupplement}{%
+        \setcounter{table}{0}
+        \renewcommand{\thetable}{S\arabic{table}}%
+        \setcounter{figure}{0}
+        \renewcommand{\thefigure}{S\arabic{figure}}%
+     }
 
@@
 that the sensitivity increases as the mean distance between binding
-events increases. When the average disntace is greater than 200 bp,
+events increases. When the average distance is greater than 200 bp,
 dPeak identifies more than the 75\% of the binding events in each
 peak, this is intuitive as the mean ChIP-Seq's fragment length is
-shorter than 200 bp, hence the read doesn't contribute to more than
+shorter than 200 bp, hence the read does not contribute to more than
 one binding event \cite{dpeak}. When the binding events in a peak are
@@
 ($N$ reads for both ChIP-exo experiment and $N/2$ or $N$ pairs for PE
-ChIP-Seq to assume more realistic situation of a fixed cost). For
-each sampled dataset we applied out lower-to-higher resolution
-pipeline by calling peaks with MOSAiCS \cite{mosaics} and then
-deconvolving the binding events by using dPeak \cite{dpeak}. For the
-ChIP-exo datasets we called peaks by using the GC-content and
-mappability models with MOSAiCS, and for the ChIP-Seq datasets we used
-their respective Input samples.
+ChIP-Seq to assume more realistic situation of a fixed cost). For each
+sampled dataset we applied our lower-to-higher resolution pipeline by
+calling peaks with MOSAiCS \cite{mosaics} and then deconvolving the
+binding events by using dPeak \cite{dpeak}. For the ChIP-exo datasets
+we called peaks by using the GC-content and mappability models with
+MOSAiCS, since it's background is usually composed of scattered reads
+across the genome; and for the ChIP-Seq datasets we used their
+respective Input samples. Additionally, it is worth noting that for PE
+ChIP-Seq we sampled both ends of the fragment, hence for each
+sequencing depth we are sampling the half amount of pairs for PE
+ChIP-Seq than for ChIP-exo or SE ChIP-Seq.
 
@@
 $\sig$ experiment under aerobic condition when comparable number of
-reads is used for all of ChIP-exo, SE ChIP-Seq and PE ChIP-Seq. It is
-remarkable that even when the number of candidate peaks or the number
-of predicted events is lower for ChIP-exo, it outperforms both PE and
-SE ChIP-Seq in number of identified targets and resolution.
+reads is used for all of ChIP-exo, SE ChIP-Seq and PE ChIP-Seq. In
+Figure \ref{fig:design}A we show the number of candidate regions
+defined as the number of regions where a binding event was identified
+in a collection of high quality peaks; in Figure \ref{fig:design}B we
+depict the number of binding events; in Figure \ref{fig:design}C we
+show the number of identified targets, where a RegulonDB was
+considered identified if a binding event was identified in a 15 bp
+vicinity of it; and finally Figure \ref{fig:design}D show the
+resolution defined as the distance from a RegulonDB annotation to the
+closest dPeak prediction. It is remarkable that even when the number
+of candidate peaks or the number of predicted events is lower for
+ChIP-exo, it outperforms both PE and SE ChIP-Seq in number of
+identified targets and resolution.
 
@@
 may be also possible that this is an artifact occurring due to
-ChIP-exo's lower library complexity. Additionally, it is worth noting
-that for PE ChIP-Seq we sampled both ends of the fragment, hence for
-each sequencing depth we are sampling the half amount of pairs for PE
-ChIP-Seq than for ChIP-exo or SE ChIP-Seq.
+ChIP-exo's lower library complexity.
 
@@
 pipeline in the fixed depth ChIP-exo experiments. Figure
-\ref{fig:exoQC_sat_aero} shows hexbin of ARC vs URCR for several fixed
-sample sizes, as the fixed depth increases the two arm pattern becomes
-more distinctive while for lower depth, it seems that the majority of
-the sampled reads were aligned to enriched regions. On the other hand
-in Figures \ref{fig:exoQC_sat1} to \ref{fig:exoQC_sat4}, we used the
-ChIP-exo QC pipeline on the samples that are outperformed by PE and SE
-ChIP-Seq. For a fixed low depth, we can see that the majority of the
-reads are being aligned to non-enriched regions since the vertical arm
-seems to be stronger for all 2 conditions and replicates; for higher
-depth we can see low URCR values being predominant, which indicates
-that the majority of the regions being formed by few positions with a
-higher read concentration. In low complexity regions, the reads are
-being aligned to fewer positions but there is no control over the
-amount of reads mapped. Hence, those regions are more likely to being
-strand-imbalance which in turn may bias the binding site estimate and
-therefore decrease the number of identified targets or increase an
-experiments resolution.
+\ref{fig:exoQC_sat_aero} shows scatter plots of ARC vs URCR for
+several fixed sample sizes, as the fixed depth increases the two arm
+pattern becomes more distinctive while for lower depth, it seems that
+the majority of the sampled reads were aligned to enriched regions. On
+the other hand in Figures \ref{fig:exoQC_sat1} to
+\ref{fig:exoQC_sat4}, we used the ChIP-exo QC pipeline on the samples
+that are outperformed by PE and SE ChIP-Seq. For a fixed low depth, we
+can see that the majority of the reads are being aligned to
+non-enriched regions since the vertical arm seems to be stronger for
+all 2 conditions and replicates; for higher depth we can see low URCR
+values being predominant, which indicates that the majority of the
+regions being formed by few positions with a higher read
+concentration. In low complexity regions, the reads are being aligned
+to fewer positions but there is no control over the amount of reads
+mapped. Hence, those regions are more likely to being strand-imbalance
+which in turn may bias the binding site estimate and therefore
+decrease the number of identified targets or increase an experiments
+resolution.
 
@@
 all the methods are comparable in resolution, and dPeak slightly
-outperforms the rest. On the other hand, for experiments with low
-library complexity the resolution calculated with dPeak's predictions
-is smaller than both Mace and Peakzilla, but it is greater than
-Gem. This may be due that the fact that Gem uses sequence information
-in addition to the aligned $5\prime$ end counts that dPeak
-uses. \footnote{For here we may probably use only 933 as part of the
-  main article and keep the rest for the supplement}
+outperforms the rest. On the other hand, for experiments dPeak
+identified binding sites with higher resolution than Mace and
+Peakzilla, but with low resolution than Gem. This may be due that the
+fact that Gem uses sequence information in addition to the aligned
+$5\prime$ end counts that dPeak uses. 
+ \footnote{For here we may
+  probably use only 933 as part of the main article and keep the rest
+  for the supplement}
+
+% are comparable if we use the init method with motif but not always
+%
 
@@
 
-We made a systematic exploration of several ChIP-exo experiments and
+We made a systematic exploration of several ChIP-exo experiments. We
 provided a list of factors that reflect the quality of a ChIP-exo
@@
 diagnostics was established to assess the quality of a ChIP-exo
-experiment. The QC pipeline only requires a set of aligned reads to
-give a global overview of a ChIP-exo experiment, this overview
+experiment. While the QC pipeline only requires a set of aligned reads
+to give a global overview of a ChIP-exo experiment, this overview
 coincides with more elaborate analysis that is computationally more
@@
 noticed that is significantly different from the ChIP-Seq one, since
-it consists mostly on a small quantity of fragments that was not
-digested by the exonuclease enzyme. Additionally, we showed that the
-background regions are imbalanced respect the number of fragments in
-both strand, and that in a lower quality ChIP-exo experiment those
-regions are going to be harder to differentiate from the possibly
-enriched regions.
+it consists of only a small quantity of fragments that was not
+digested by the exonuclease enzyme. Additionally, we showed that we
+have unbalanced number of reads in forward and reverse strands, and
+that in a lower quality ChIP-exo experiment those regions are going to
+be harder to differentiate from the possibly enriched regions.
 
@@
 ChIP-exo is more sensitive than both SE and PE ChIP-Seq. We made a
-methodical comparison between fixed depth ChIP-exo, PE ChIP-Seq and SE
+rigorous comparison between fixed depth ChIP-exo, PE ChIP-Seq and SE
 ChIP-Seq, and we probed that for sufficiently complex libraries,
 ChIP-exo experiments can outperform PE and SE ChIP-Seq in number of
-identified targets and resolution. Using the ChIP-exo QC pipeline, we
-show how to diagnose if the library complexity of ChIP-exo experiment
-is adequate.
+identified targets and resolution. The proposed ChIP-exo QC pipeline
+provides a rigorous, easily interpretable, computationally efficient
+framework to diagnose if the library complexity of a ChIP-exo
+experiment is adequate.
 
@@
 
-For every ChIP-exo experiment, we partitioned the experiment into
-regions using the QC pipeline. For each region we calculated the FSR
-defined as the ratio between the number of forward stranded reads and
-the total number of reads in a region. The imbalance index is defined
-as:
+For every ChIP-exo experiment, we partitioned the experiment into the
+non-digested regions using the QC pipeline. For each region we
+calculated the FSR defined as the ratio between the number of forward
+stranded reads and the total number of reads in a region.
+
+Then we
+transformed the FSR into an imbalance index is defined as:
 
@@
 experiment from the PE ones by taking one of both ends randomly
-following a $\mbox{Ber}(0.5)$ model.
+following with equal probability.
 
@@
 peaks with average ChIP counts greater than 3000 that overlapped both
-the SE and PE ChIP-Seq peaks. Then we estimated the binding site
-events using the dPeak model with a maximum number of 5 binding
-events. We considered RegulonDB annotation as gold-standard;
-Resolution is defined as the minimum distance from an annotation to an
-estimated binding events and Sensitivity is defined as the fraction of
-annotation in a are within \Sexpr{ext} bp from an estimated binding event.
+the SE and PE ChIP-Seq peaks, we considered other cutoff values but
+still obtained results similar to what we presented in this
+paper. Then we estimated the binding site events using the dPeak model
+with a maximum number of 5 binding events. We considered RegulonDB
+annotation as gold-standard; Resolution is defined as the minimum
+distance from an annotation to an estimated binding events and
+Sensitivity is defined as the fraction of annotation within
+\Sexpr{ext} bp from an estimated binding event.
 
@@
 
+\beginsupplement
+
 \section*{Supplement}

### 2016-03-21 rmarkdown/paper.tex
updated paper figures and fixed format

@@
 peaks with average ChIP counts greater than 3000 that overlapped both
-the SE and PE ChIP-Seq peaks. Then we estimated the binding site
-events using the dPeak model with a maximum number of 5 binding
-events. We considered RegulonDB annotation as gold-standard;
-Resolution is defined as the minimum distance from an annotation to an
-estimated binding events and Sensitivity is defined as the fraction of
-annotation in a are within 15 bp from an estimated binding event.
+the SE and PE ChIP-Seq peaks, we considered other cutoff values but
+still obtained results similar to what we presented in this
+paper. Then we estimated the binding site events using the dPeak model
+with a maximum number of 5 binding events. We considered RegulonDB
+annotation as gold-standard; Resolution is defined as the minimum
+distance from an annotation to an estimated binding events and
+Sensitivity is defined as the fraction of annotation within
+15 bp from an estimated binding event.
 

### 2016-03-21 rscripts/paper_analysis/carroll_foxA1_mm9_downstream_fimo_comparison.R

@@ -98,23 +98,43 @@ figs_dir <- "figs/FOXA1_mm9_fimo"
 dir.create(figs_dir,showWarnings = FALSE)
 
 pdf(file = file.path(figs_dir,"FOXA1_summaries_by_replicate.pdf"))
-ggplot(DTsummary,aes(rep,nMotif,fill = rep))+geom_bar(stat = "identity",colour = "black")+
-  theme_bw()+theme(legend.position = "none")+
-  xlab("Replicates")+ylab("Number of Motif occurrences")+scale_fill_brewer(palette = "Pastel1")
-ggplot(DTsummary,aes(rep,nSites,fill = rep))+geom_bar(stat = "identity",colour = "black")+
-  theme_bw()+theme(legend.position = "none")+
-  xlab("Replicates")+ylab("Number of FoxA1 binding events")+scale_fill_brewer(palette = "Pastel1")
-ggplot(DTsummary,aes(rep,Motif_perc,fill = rep))+geom_bar(stat = "identity",colour = "black")+
-  scale_y_continuous(labels = percent)+theme_bw()+theme(legend.position = "none")+
-  xlab("Replicates")+ylab("Percentage of sites with FoxA1 motif")+scale_fill_brewer(palette = "Pastel1")
+ggplot(DTsummary,aes(rep,nMotif,fill = rep))+
+  geom_bar(stat = "identity",colour = "black")+
+  theme_bw()+
+  theme(legend.position = "none",plot.title = element_text(hjust = 0,size = 32),
+        axis.text = element_text(size = 24),
+        axis.title = element_text(size = 26))+
+  xlab("Replicates")+ylab("Number of Motif occurrences")+
+  scale_fill_brewer(palette = "Pastel1")+ggtitle("B")
+ggplot(DTsummary,aes(rep,nSites,fill = rep))+
+  geom_bar(stat = "identity",colour = "black")+
+  theme_bw()+
+  theme(legend.position = "none",plot.title = element_text(hjust = 0,size = 32),
+        axis.text = element_text(size = 24),
+        axis.title = element_text(size = 26))+
+  xlab("Replicates")+ylab("Number of FoxA1 binding events")+
+  scale_fill_brewer(palette = "Pastel1")
+ggplot(DTsummary,aes(rep,Motif_perc,fill = rep))+
+  geom_bar(stat = "identity",colour = "black")+
+  scale_y_continuous(labels = percent)+
+  theme_bw()+
[…]

### 2016-03-21 rscripts/saturation_analysis_old.R

@@ -296,6 +296,10 @@ library(grid)
 library(gridExtra)
 library(RColorBrewer)
 
+    ## axis.text = element_text(size = 24),
+    ## axis.title = element_text(size = 26))+
+
+
 p1 <- ggplot(dt1,aes(nreads,candre, colour = Dataset))+
   geom_line(size = 1.2)+theme_bw()+
   scale_color_brewer(palette = "Set1",name="",guide = guide_legend(nrow = 2))+
@@ -304,20 +308,20 @@ p1 <- ggplot(dt1,aes(nreads,candre, colour = Dataset))+
   ylim(0,600)
 p2 <- ggplot(dt2,aes(nreads,predre, colour = Dataset))+
   geom_line(size = 1.2)+theme_bw()+
-  scale_color_brewer(palette = "Set1",name="",guide = guide_legend(nrow = 2))+    
   theme(legend.position = "top",plot.title = element_text(hjust = 0))+
+  scale_color_brewer(palette = "Set1",name="",guide = guide_legend(nrow = 2))+    
   xlab("Number of reads (million)")+ylab("Number of predicted events")+
   ylim(0,1500)
 p3 <- ggplot(dt3,aes(nreads,iden, colour = Dataset))+
   geom_line(size = 1.2)+theme_bw()+
-  scale_color_brewer(palette = "Set1",name="",guide = guide_legend(nrow = 2))+  
   theme(legend.position = "top",plot.title = element_text(hjust = 0))+
+  scale_color_brewer(palette = "Set1",name="",guide = guide_legend(nrow = 2))+  
   xlab("Number of reads (million)")+ylab("Number of identified targets")+
   ylim(0,250)
 p4 <- ggplot(dt4,aes(nreads,resol, colour = Dataset))+
   geom_line(size = 1.2)+theme_bw()+
-  scale_color_brewer(palette = "Set1",name="",guide = guide_legend(nrow = 2))+  
   theme(legend.position = "top",plot.title = element_text(hjust = 0))+
+  scale_color_brewer(palette = "Set1",name="",guide = guide_legend(nrow = 2))+  
   xlab("Number of reads (million)")+ylab("Resolution")+
   ylim(0,50)
 
@@ -330,7 +334,11 @@ print(p4)
 dev.off()
 
 pdf(file = "figs/for_paper/saturation_analysis_old.pdf")
[…]

### 2016-03-21 rscripts/peak_pair_assumption_old_data.R
fixed density figures for delta and sigma

@@ -168,7 +168,8 @@ dt[ , Dataset := plyr::mapvalues( Dataset ,
         from = c("ChIP-exo","ChIP-seq (PET)","ChIP-seq (SET)"),
         to = c("ChIP-exo","ChIP-seq (PE)","ChIP-seq (SE)"))]
 
-p1 = ggplot(dt,aes(delta,colour = Dataset))+geom_density(size=1)+facet_grid(Dataset~.)+
+p1 = ggplot(dt,aes(delta,colour = Dataset))+stat_density(size=1,geom = "line")+
+  facet_grid(Dataset~.)+
   theme_bw()+theme(legend.position = "none",plot.title = element_text(hjust = 0))+
   scale_color_manual( values = brewer.pal(3,"Set1")[c(1,3)])+
   ylab("")+xlab("delta estimate")
@@ -182,7 +183,8 @@ dt[ , Dataset := plyr::mapvalues( Dataset ,
         from = c("ChIP-exo","ChIP-seq (PET)","ChIP-seq (SET)"),
         to = c("ChIP-exo","ChIP-seq (PE)","ChIP-seq (SE)"))]
 
-p2 = ggplot(dt,aes(sigma,colour = Dataset))+geom_density(size=1)+facet_grid(Dataset~.)+
+p2 = ggplot(dt,aes(sigma,colour = Dataset))+stat_density(size=1,geom = "line")+
+  facet_grid(Dataset~.)+
   theme_bw()+theme(legend.position = "none",plot.title = element_text(hjust = 0))+
   scale_color_manual(values = brewer.pal(3,"Set1")[c(1,3)])+ylab("")+
   xlab("sigma estimate")

### 2016-04-07 rmarkdown/paper.Rnw
fixed density figures for delta and sigma
We updated the paper to add more datasets in table 1. We also, made several scripts to unify the pipeline runs.
We are working in adding a section in ChIPnexus

@@
 et al., 2013 \cite{exoillumina}); Glucocorticoid receptor (GR) in
-IMR90, K562 and U20S cell lines (Starick et al., 2015
-\cite{starick15}).Furthermore, we also generated ChIP-exo and ChIP-Seq
-data for $\sig$ factor in \emph{Escherichia Coli} (\emph{E. Coli})
-measured under aerobic ($ + O_2$) condition, and treated by rifampicin
-by 0 and 20 minutes (courtesy of Professor Robert Landick's lab).
+IMR90, K562 and U2OS cell lines (Starick et al., 2015
+\cite{starick15}); TATA-box Binding Protein (TBP) in K562 cell lines
+(Venters and Pugh, 2013 \cite{venters13}). Furthermore, we also
+generated ChIP-exo and ChIP-Seq data for $\sig$ factor in
+\emph{Escherichia Coli} (\emph{E. Coli}) measured under aerobic ($ +
+O_2$) condition, and treated by rifampicin by 0 and 20 minutes
+(courtesy of Professor Robert Landick's lab).
 
@@
   load("../data/paper_tables/Carroll_hg19_ER.RData")
+  load("../data/paper_tables/Venters_hg19_TBP.RData")
   landick_aero[,Organism := "E.Coli"]
@@
   carroll_hg19_ER[ ,cond := "MCF-7"]
+  venters_hg19_TBP[,Organism := "Human"]
+  venters_hg19_TBP[,cond := "K562"]
+  venters_hg19_TBP[,Replicate := 1:3]
+  venters_hg19_TBP[,TF := "TBP"]
+
+  ## this creates the table
   A <- do.call(rbind,list(landick_aero,
@@
                      meijing_hg19_GR[order(Replicate)],
-                     carroll_hg19_ER[order(Replicate)]))
+                     carroll_hg19_ER[order(Replicate)],
+                     venters_hg19_TBP[order(Replicate)]))
   A[,nsc := unlist(nsc)]
@@
 pair'' assumption does not hold locally in ChIP-exo data.
+
+\subsubsection{Analysis of ChIP-Nexus data with the QC pipeline}
+
+ChIP-nexus developed by He et al., 2015 \cite{chipnexus} is a
+modification of the ChIP-exo protocol, where both sequencing adaptors
+are ligated at the $3\prime$ end of the ChIP fragments. Then,
+exonuclease digestion and DNA self-circularization with circLigase,
+and restriction enzyme cutting between the two adaptors to create the
+final library that is amplified.
 
@@
   \includegraphics[width = .9\textwidth,page = 1]{../figs/saturation/K12_alignment/sig70_rif_enrichment.pdf}
-  \caption{ARC vs URCR hexbin plots of Rep-2 and rif-0min from $\sig$ experiment when 100K to 900K reads are being
-    sampled for each panel.}
+  \caption{ARC vs URCR hexbin plots of Rep-2 and rif-0min from $\sig$
+    experiment when 100K to 900K reads are being sampled for each
+    panel.}
   \label{fig:exoQC_sat3}

### 2016-04-07 rmarkdown/paper.tex
fixed density figures for delta and sigma
We updated the paper to add more datasets in table 1. We also, made several scripts to unify the pipeline runs.
We are working in adding a section in ChIPnexus

@@
 ## 18:    Human     ER          MCF-7    3  12,464,836 0.8204  18.7281
+## 19:    Human    TBP           K562    1  61,046,382 0.1232  12.0117
+## 20:    Human    TBP           K562    2  94,314,770 0.1682   7.9398
+## 21:    Human    TBP           K562    3 114,282,270 0.1464   9.2566
+##     Organism  IP/TF Condition/Cell Rep.       Depth    PBC      NSC
 \end{verbatim}

### 2016-04-07 rmarkdown/paper.R

@@ -6,6 +6,7 @@
   load("../data/paper_tables/Carroll_mm9_FoxA1.RData")
   load("../data/paper_tables/Meijsing_hg19_GR.RData")
   load("../data/paper_tables/Carroll_hg19_ER.RData")
+  load("../data/paper_tables/Venters_hg19_TBP.RData")
   landick_aero[,Organism := "E.Coli"]
   landick_aero[,Replicate := c(1,2,1,2)]
   landick_aero[,TF := "Sig70"]
@@ -30,12 +31,19 @@
   carroll_hg19_ER[,Replicate := c(1,3,2)]
   carroll_hg19_ER[,TF := "ER"]
   carroll_hg19_ER[ ,cond := "MCF-7"]
+  venters_hg19_TBP[,Organism := "Human"]
+  venters_hg19_TBP[,cond := "K562"]
+  venters_hg19_TBP[,Replicate := 1:3]
+  venters_hg19_TBP[,TF := "TBP"]
+
+  ## this creates the table
   A <- do.call(rbind,list(landick_aero,
                      landick_rif[order(Replicate)],
                      pugh_ctcf[order(Replicate)],
                      carroll_mm9_FoxA1[order(Replicate)],
                      meijing_hg19_GR[order(Replicate)],
-                     carroll_hg19_ER[order(Replicate)]))
+                     carroll_hg19_ER[order(Replicate)],
+                     venters_hg19_TBP[order(Replicate)]))
   A[,nsc := unlist(nsc)]
   setcolorder(A, c("Organism","TF","cond","Replicate","nreads","pbc","nsc","files"))
   A[,files := NULL]

### 2016-04-07 rscripts/ARC_vs_URCR_quantify.R

@@ -1,0 +1,93 @@
+
+rm(list = ls())
+
+library(devtools)
+library(viridis)
+library(hexbin)
+library(data.table)
+library(GenomicAlignments)
+library(parallel)
+library(scales)
+
+data_dir <- "/p/keles/ChIPexo/volume4/carroll_data/human/bamfiles"
+data_dir <- "/p/keles/ChIPexo/volume4/zeitlinger_data/bam/sortbam"
+
+files <- list.files(data_dir)
+files <- files[grep("sort",files)]
+files <- files[grep("bai",files,invert = TRUE)]
+ 
+files <- files[grep("embryo_dors",files)]
+load_all("~/Desktop/Docs/Code/ChIPexoQual")
+
+exo <- lapply(file.path(data_dir,files),
+  create_exo_experiment,parallel = TRUE,height = 1)
+
+stats <- lapply(exo,summary_stats)
+
+enrichment_plots <- function(stat)
+{
+  r <- viridis(100,option = "D")
+
+  plots <- list()
+  plots[[1]] <- ggplot(stat,aes(npos,depth))+stat_binhex(bins = 75)+
+    scale_fill_gradientn(colours = r,trans = "log10",
+      labels = trans_format("log10",math_format(10^.x)))+
+    theme_bw()+theme(legend.position = "top")+
+      xlim(0,500)+ylim(0,3000)
+  plots[[2]] <- ggplot(stat,aes(ave_reads,cover_rate))+
+    stat_binhex(bins = 75)+
+    scale_fill_gradientn(colours = r,trans = "log10",
+      labels = trans_format("log10",math_format(10^.x)))+
+    theme_bw()+theme(legend.position = "top")+
+      xlim(0,4)+ylim(0,1)
+  plots[[3]] <- ggplot(stat,aes(npos,width))+stat_binhex(bins = 75)+
+    scale_fill_gradientn(colours = r,trans = "log10",
+      labels = trans_format("log10",math_format(10^.x)))+
+    theme_bw()+theme(legend.position = "top")+
+      xlim(0,500)+ylim(0,1200)
+  plots[[4]] <- ggplot(stat,aes(depth,width))+stat_binhex(bins = 75)+
+    scale_fill_gradientn(colours = r,trans = "log10",
+      labels = trans_format("log10",math_format(10^.x)))+
[…]

### 2016-04-07 rscripts/ChIPexoQual_runs/Venters_hg19_runs.R

@@ -1,0 +1,155 @@
+rm(list = ls())
+library(viridis)
+library(data.table)
+library(GenomicAlignments)
+library(devtools)
+library(parallel)
+library(ggplot2)
+library(scales)
+library(RColorBrewer)
+library(hexbin)
+library(reshape2)
+
+load_all("~/Desktop/Docs/Code/ChIPexoQual")
+
+
+indir <- "/p/keles/ChIPexo/volume4/venters_data/sortbam"
+files <- list.files(indir)
+files <- files[grep("sort",files)]
+files <- files[grep("bai",files,invert= TRUE)]
+mc <- detectCores()
+
+
+files <- files[grep("TBP",files)]
+prefix <- "Venters_TBP_human"
+
+expts <- lapply(file.path(indir,files),create_exo_experiment,calc_summary = TRUE,height = 1,parallel = TRUE , mc = detectCores())
+
+names(expts) <- gsub(".sort.bam","",files)
+stats <- lapply(expts,summary_stats)
+
+
+fsr_DT <- function(stats,probs = c(0,.25,.5,.75,1),values = 1:750, mc = 8)
+{
+  stopifnot(is.numeric(values))
+  stopifnot(length(probs) <= 9)
+  stopifnot(length(values) > 1)
+  to_use <- stats[,.(depth, fsr)]
+  rows <- mclapply(values,function(i)to_use[depth > i,quantile(fsr, probs)],mc.cores = mc)
+  rows <- mapply(function(x,i)data.table(depth = i,quantiles = probs,fsr = x ),rows,values,SIMPLIFY = FALSE)
+  rows <- do.call(rbind,rows)
+  rows[,quantiles := factor(quantiles,levels = sort(probs))]
+  return(rows)
+}
+
+
+label_DT <- function(stats,values = 1:750,mc = 8,prop = FALSE)
+{
+  stopifnot(is.numeric(values))
+  stopifnot(length(values) > 1)
+
+  stats[,label := ifelse( f > 0 & r > 0 , "both",ifelse(f > 0,"fwd","bwd"))]
+
+  to_use <- stats[,.(depth,label)]
+  setkey(to_use,label)
+  fwd <- to_use["fwd"]
+  fwd <- do.call(c,mclapply(values,function(i)nrow(fwd[depth > i]),mc.cores = mc))
+  bwd <- to_use["bwd"]
[…]

### 2016-04-07 rscripts/paper_analysis/summary_arc_vs_urcr_regression.R

@@ -1,0 +1,66 @@
+
+rm(list = ls())
+
+library(broom)
+library(data.table)
+library(ggplot2)
+library(scales)
+library(RColorBrewer)
+
+in_dir <- "data/ARC_quantify"
+files <- list.files(in_dir)
+
+load_file <- function(x){load(x);coeff}
+
+coeffs <- lapply(file.path(in_dir,files),load_file)
+
+coeffs <- mapply(function(x,y){
+  x$id = y
+  x},coeffs,gsub(".RData","",files),SIMPLIFY = FALSE)
+
+
+coeffs <- data.table(do.call(rbind,coeffs))
+
+
+coeffs[ , Rep := "Rep-1"]
+
+coeffs[grep("Rep2",id), Rep := "Rep-2"]
+coeffs[grep("Rep3",id), Rep := "Rep-3"]
+
+coeffs[grep("1314",id), Rep := "Rep-2"]
+coeffs[grep("1320",id), Rep := "Rep-2"]
+coeffs[grep("933",id), Rep := "Rep-2"]
+coeffs[grep("937",id), Rep := "Rep-2"]
+
+coeffs[, expt := "a"]
+coeffs[ Rep == "Rep-1", expt := gsub("_Rep1","",id)]
+coeffs[ Rep == "Rep-2", expt := gsub("_Rep2","",id)]
+coeffs[ Rep == "Rep-3", expt := gsub("_Rep3","",id)]
+
+coeffs[grepl("landick_sig70_9",id), expt := "landick_aero"]
+coeffs[grepl("landick_sig70_13",id), expt := "landick_sig70_rif"]
+
+z <- qnorm(.95)
+
+figs_dir <- "figs/test_quantify"
+
+pdf(file = file.path(figs_dir,"arc_urcr_param_regression.pdf"),width = 12,height = 6)
+ggplot(coeffs[term == "b"],aes(Rep , estimate,fill = expt))+geom_bar(stat = "identity")+
+  facet_grid( . ~ expt,scales = "free",space = "free"  )+theme_bw()+
+  theme(legend.position = "none",
+        axis.text.x = element_text(angle = 90),
+        strip.text.x = element_text(angle = 70) )+
+  xlab("")+ylab("Intercept")
+ggplot(coeffs[term == "k"],aes(Rep ,estimate,fill = expt))+geom_bar(stat = "identity")+
+  facet_grid( . ~ expt,scales = "free",space = "free"  )+theme_bw()+
+  theme(legend.position = "none",
[…]

### 2016-04-07 rscripts/paper_analysis/summary_npos_depth_regression.R

@@ -1,0 +1,121 @@
+
+rm(list = ls())
+
+library(data.table)
+library(ggplot2)
+library(RColorBrewer)
+library(scales)
+
+data_dir <- "data/QC_quantify"
+files <- list.files(data_dir)
+
+load_file <- function(x){
+  load(x)
+  return(coeff)
+}
+
+coeffs <- lapply(file.path(data_dir,files),load_file)
+coeffs <- mapply(function(x,y)x[,id := y],coeffs,gsub(".RData","",files),SIMPLIFY = FALSE)
+DT <- do.call(rbind,coeffs)
+
+
+N <- 1e4
+K <- N
+
+mean_est <- sapply(coeffs,function(x)x[,mean(estimate[1:K])])
+var_est <- sapply(coeffs,function(x)x[,var(estimate[1:K])])
+
+z <- qnorm(.05,lower.tail = FALSE)
+
+dt <- data.table(id = sub(".RData","",files),
+                 m = mean_est,var = var_est)
+dt[,lb := m - z * sqrt(var) / N]
+dt[,ub := m + z * sqrt(var) / N]
+dt[,organism := "human"]
+dt[grepl("mouse",id),organism := "mouse"]
+dt[grepl("landick",id),organism := "e.coli"]
+dt[grepl("S2",id),organism := "drosophila"]
+dt[grepl("embryo",id),organism := "drosophila"]
+
+DT[,organism := "human"]
+DT[grepl("mouse",id),organism := "mouse"]
+DT[grepl("landick",id),organism := "e.coli"]
+DT[grepl("S2",id),organism := "drosophila"]
+DT[grepl("embryo",id),organism := "drosophila"]
+
+
+figs_dir <- "figs/test_quantify"
+pdf(file = file.path(figs_dir,"slope_npos_depth.pdf"),width = 12,height = 6)
+ggplot(dt,aes(id,m))+geom_bar(stat = "identity")+
+  theme(legend.position = "none",axis.text.x = element_text(angle = 90))+
+  facet_grid( ~ organism  ,scales = "free",space = "free")+xlab("")+
+  xlab("Experiment")+ylab("Depth to Nr. of Unique positions rate")
+ggplot(dt[!grepl("venter",id)],aes(id,m))+geom_bar(stat = "identity")+
+  theme(legend.position = "none",axis.text.x = element_text(angle = 90))+
+  facet_grid( ~ organism  ,scales = "free",space = "free")+xlab("")+
[…]

### 2016-04-07 rscripts/paper_tables/Venters_TBP_table.R

@@ -1,0 +1,32 @@
+rm(list = ls())
+library(ChIPUtils)
+library(data.table)
+library(GenomicAlignments)
+
+indir <- "/p/keles/ChIPexo/volume4/venters_data/sortbam"
+files <- list.files(indir)
+files <- files[grep(".sort.bam",files)]
+files <- files[grep("bai",files,invert = TRUE)]
+files <- files[grep("TBP",files)]
+
+reads <- lapply(file.path(indir,files),create_reads)#,mc.cores = 3)
+names(reads) <- gsub(".sort.bam","",files)
+nreads <- sapply(reads,nreads)
+pbc <- do.call(c,mclapply(reads,PBC,mc.cores = 3))
+
+fsr <- sapply(reads,function(x){
+  tab <- summary(x)
+  fwd <- sum(tab[,as.numeric(readsF)])
+  bwd <- sum(tab[,as.numeric(readsR)])
+  fwd / (fwd + bwd)
+})
+
+human.size <-system.file("extdata","chrom.sizes","hg19.chrom.sizes",package = "ChIPUtils")
+human.size <- data.table(read.table(human.size))
+scc <- lapply(reads,strand_cross_corr,shift = 1:300,chrom.size = human.size, parallel = TRUE)
+
+nsc <- sapply(scc,function(x)x[,max(cross.corr) / min(cross.corr)])
+
+venters_hg19_TBP <- data.table(files = names(reads),nreads,pbc,nsc)
+
+save(venters_hg19_TBP,file = "data/paper_tables/Venters_hg19_TBP.RData")

### 2016-04-07 rscripts/scripts/ChIPexoQual_run.R

@@ -1,0 +1,55 @@
+#! /usr/bin/env Rscript
+args <- commandArgs(TRUE)
+if("--help" %in% args){
+  cat("
+
+  Name:
+
+    ChIPexoQual_run.R - Applies the ChIPexoQual pipeline to partition
+       the aligned reads into regions and calculate a series of summary
+       statistics.
+
+  Arguments:
+
+
+   -- bamfile
+
+      File with bam format that contains the fragments of the ChIP-exo experiment.
+
+   -- outfile
+
+      RData file used to store the run and estimated parameters.
+
+   -- help
+
+      Show the help file.
+
+  Author:
+
+    Rene Wech, Department of Statistics, University of Wisconsin - Madison
+   
+");q()}
+
+stopifnot(length(args) == 2)
+
+library(GenomicAlignments)
+library(parallel)
+library(devtools)
+library(data.table)
+library(broom)
+
+load_all("~/Desktop/Docs/Code/ChIPexoQual")
+
+bamfile <- args[1]
+outfile <- args[2]
+
+mc <- detectCores()
+
+exo <- create_exo_experiment(bamfile,parallel = TRUE,mc.cores = mc)
+stats <- summary_stats(exo)
+
+ext_stats <- list()
+ext_stats[["stats"]] <- stats
+ext_stats[["nreads"]] <- length(reads(exo))
+
+save(ext_stats,file = outfile)

### 2016-04-07 rscripts/scripts/arc_vs_urcr_regression.R

@@ -1,0 +1,92 @@
+#! /usr/bin/env Rscript
+args <- commandArgs(TRUE)
+if("--help" %in% args){
+  cat("
+
+  Name:
+
+    arc_vs_urcr_regression.R - Performs a non-linear models between the ARC (average reads coverage) and
+      the URCR (unique read coverage rate)
+
+  Arguments:
+
+
+   -- bamfile
+
+      File with bam format that contains the fragments of the ChIP-exo experiment.
+
+   -- outfile
+
+      RData file used to store the run and estimated parameters.
+
+   -- minNpos
+
+      Min. number of unique position per region considered.
+
+   -- help
+
+      Show the help file.
+
+  Author:
+
+    Rene Wech, Department of Statistics, University of Wisconsin - Madison
+   
+");q()}
+
+stopifnot(length(args) == 3)
+
+library(GenomicAlignments)
+library(parallel)
+library(devtools)
+library(data.table)
+library(broom)
+
+load_all("~/Desktop/Docs/Code/ChIPexoQual")
+
+bamfile <- args[1]
+outfile <- args[2]
+minNpos <- as.numeric(args[3])
+
+mc <- detectCores()
+
+
+exo <- create_exo_experiment(bamfile,parallel = TRUE,mc.cores = mc)
+stats <- summary_stats(exo)
+
+rm(exo)
+
+stats <- stats[ f > 0 & r > 0 ]
+
+stats <- stats[npos > minNpos]
+
+model <- nls(cover_rate ~ b + k / ave_reads,data =stats,start = list(k =1,b=0))
+coeff <- tidy(model)
+coeff$n <- nrow(stats)
+
+save(coeff,file = outfile)
+
+
+## dr <- "/p/keles/ChIPexo/volume4/carroll_data/mouse/"
+
+## files <- list.files(dr)
+## files <- files[grep("sort",files)]
+## files <- files[grep("bai",files,invert = TRUE)]
+
+## bamfile <- file.path(dr,files[3])
+
+## samp <- function(stats,M)
+## {
+##   idx <- sample( nrow(stats),M)
+##   dt <- stats[idx,.(npos,depth)]
+##   model <- lm(depth ~ 0 + npos, data =dt)
+##   summ <- summary(model)
+##   out <- data.table(tidy(model))
+##   out[,term := NULL]
+##   out[,R2 := summ$r.squared]  
+##   return(out)
+
+## }
+
[…]

### 2016-04-07 scripts/ave_reads_vs_cover_rate_all_exo.sh

@@ -1,0 +1,75 @@
+#!/bin/sh
+
+outdir=data/ARC_quantify
+
+minNpos=10
+
+# carroll mouse
+dr=/p/keles/ChIPexo/volume4/carroll_data/mouse
+
+rscripts/scripts/arc_vs_urcr_regression.R $dr/ERR336942.sort.bam $outdir/carroll_mouse_Rep1.RData $minNpos
+rscripts/scripts/arc_vs_urcr_regression.R $dr/ERR336956.sort.bam $outdir/carroll_mouse_Rep2.RData $minNpos
+rscripts/scripts/arc_vs_urcr_regression.R $dr/ERR336935.sort.bam $outdir/carroll_mouse_Rep3.RData $minNpos
+
+# carroll human
+dr=/p/keles/ChIPexo/volume4/carroll_data/human/bamfiles
+
+rscripts/scripts/arc_vs_urcr_regression.R $dr/ERR336933.sort.bam $outdir/carroll_human_Rep1.RData $minNpos
+rscripts/scripts/arc_vs_urcr_regression.R $dr/ERR336950.sort.bam $outdir/carroll_human_Rep2.RData $minNpos
+rscripts/scripts/arc_vs_urcr_regression.R $dr/ERR336938.sort.bam $outdir/carroll_human_Rep3.RData $minNpos
+
+# pugh original ctcf
+dr=/p/keles/ChIPexo/volume4/pugh_data
+
+rscripts/scripts/arc_vs_urcr_regression.R $dr/CTCF.bam $outdir/pugh_CTCF.RData $minNpos
+
+# landick aerobic sig70
+dr=/p/keles/ChIPexo/volume7/Landick/K12/ChIPexo/aerobic_vs_anaerobic
+
+rscripts/scripts/arc_vs_urcr_regression.R $dr/edsn931_Sig70.sort.bam $outdir/landick_sig70_931.RData $minNpos
+rscripts/scripts/arc_vs_urcr_regression.R $dr/edsn933_Sig70.sort.bam $outdir/landick_sig70_933.RData $minNpos
+rscripts/scripts/arc_vs_urcr_regression.R $dr/edsn935_Sig70.sort.bam $outdir/landick_sig70_935.RData $minNpos
[…]

### 2016-04-07 scripts/npos_vs_depth_all_exo.sh

@@ -12,7 +12,7 @@ rscripts/scripts/npos_depth_regression.R $dr/ERR336956.sort.bam $outdir/carroll_
 rscripts/scripts/npos_depth_regression.R $dr/ERR336935.sort.bam $outdir/carroll_mouse_Rep3.RData $M $N
 
 # carroll human
-dr=/p/keles/ChIPexo/volume4/carroll_data/human
+dr=/p/keles/ChIPexo/volume4/carroll_data/human/bamfiles
 
 rscripts/scripts/npos_depth_regression.R $dr/ERR336933.sort.bam $outdir/carroll_human_Rep1.RData $M $N
 rscripts/scripts/npos_depth_regression.R $dr/ERR336950.sort.bam $outdir/carroll_human_Rep2.RData $M $N
@@ -63,3 +63,13 @@ dr=/p/keles/ChIPexo/volume4/venters_data/sortbam
 rscripts/scripts/npos_depth_regression.R $dr/TBP_K562_Rep1.sort.bam $outdir/venters_TBP_K562_Rep1.RData $M $N
 rscripts/scripts/npos_depth_regression.R $dr/TBP_K562_Rep2.sort.bam $outdir/venters_TBP_K562_Rep2.RData $M $N
 rscripts/scripts/npos_depth_regression.R $dr/TBP_K562_Rep3.sort.bam $outdir/venters_TBP_K562_Rep3.RData $M $N
+
+# meijsing GR
+dr=/p/keles/ChIPexo/volume4/meijsing_data/
+
+rscripts/scripts/npos_depth_regression.R $dr/IMR90_GR_chip-exo.sort.bam $outdir/meijsing_GR_IMR90_Rep1.RData $M $N
+rscripts/scripts/npos_depth_regression.R $dr/K562_GR_chip-exo.sort.bam $outdir/meijsing_GR_K562_Rep1.RData $M $N
+rscripts/scripts/npos_depth_regression.R $dr/U2OS_GR_chip-exo.sort.bam $outdir/meijsing_GR_U2OS_Rep1.RData $M $N
+
+
+

### 2016-04-12 rmarkdown/paper.Rnw
preliminary version of the paper, pre -change of sections

@@
     ChIP-Seq, ChIP-exo presents a number of unique challenges. We
-    present a quality control pipeline that analyzes a ChIP-exo
-    experiment's strand imbalance, enrichment and library
+    present a quality control pipeline that analyzes a ChIP-exo and
+    ChIP-nexus experiment's strand imbalance, enrichment and library
     complexity. Assessment of these characteristics are facilitated
@@
   \kwd{ChIP-exo}
+  \kwd{ChIP-nexus}
+  \kwd{ChIP-Seq}
   \kwd{Quality Control}
-  \kwd{ChIP-Seq}
   \kwd{Spatial Resolution}
@@
 \end{frontmatter}
+
+% \newpage
+
+% \tableofcontents
+% \listoffigures
 
@@
 
-% Reffers to... First, DNA libraries... I think that these might be
+% Refers to... First, DNA libraries... I think that these might be
 % too much details for the introduction. How about discussing them in
@@
   load("../data/paper_tables/Venters_hg19_TBP.RData")
+  load("../data/paper_tables/ChIPnexus_dm3.RData")
+  load("../data/paper_tables/ChIPnexus_hg19.RData")
   landick_aero[,Organism := "E.Coli"]
@@
   venters_hg19_TBP[,TF := "TBP"]
-
+  chipnexus_dm3[, Organism := "D. Melanogaster"]
+  chipnexus_dm3[, cond := c(rep("Embryo",4),rep("S2",4))]
+  chipnexus_dm3[, Replicate := rep(1:2,4)]
+  chipnexus_dm3[, TF := c(rep("Dorsal",2),rep("Twist",2),rep("Max",2),rep("MyC",2))]
+  chipnexus_K562[ , Organism := "Human"]
+  chipnexus_K562[ , cond  := "K562"]
+  chipnexus_K562[ , Replicate := 1:2]
+  chipnexus_K562[ , TF := "TBP"]
   ## this creates the table
@@
                      venters_hg19_TBP[order(Replicate)]))
+  B <- do.call(rbind,list(chipnexus_dm3,
+                     chipnexus_K562))
+  format_tab <- function(A){
     A[,nsc := unlist(nsc)]
@@
     setnames(A,names(A),c("Organism","IP/TF","Condition/Cell","Rep.","Depth","PBC","NSC"))
+    return(A)}
+
+  A <- format_tab(A)
+  B <- format_tab(B)
 @
@@
 
-\subsubsection{Analysis of ChIP-Nexus data with the QC pipeline}
+\subsubsection{Analysis of ChIP-nexus data with the QC pipeline}
 
 ChIP-nexus developed by He et al., 2015 \cite{chipnexus} is a
-modification of the ChIP-exo protocol, where both sequencing adaptors
-are ligated at the $3\prime$ end of the ChIP fragments. Then,
-exonuclease digestion and DNA self-circularization with circLigase,
-and restriction enzyme cutting between the two adaptors to create the
-final library that is amplified.
+modification to the ChIP-exo protocol, where both sequencing adaptors
+are ligated at the $3\prime$ end of the ChIP fragments. Then, after
+exonuclease digestion, DNA self-circularization with circLigase, and
+restriction enzyme cutting between the two adaptors, the final library
+is amplified. In Table \ref{tab:nexus} we show the usual ChIP-Seq QC
+metrics to ChIP-nexus data and found that the conclusions made for
+ChIP-exo data still holds, the NSC is high but when plotting the SCC
+the fragment and read length peaks are confounded and the PBC may
+indicate severe bottlenecking but it is caused by the exonuclease
+enzyme digestion which reduces the number of positions where the
+reads can be aligned.
+
+\begin{table}[h!]
+  \centering
+<<table3,include = TRUE,echo = FALSE, eval = TRUE >>= 
+  B
+@ 
+\caption{\textbf{Current QC metrics applied to ChIP-nexus data.} PBC
+  stands for PCR Bottleneck Coefficient and NSC for Normalized Strand
+  Cross-Correlation.}
+  \label{tab:nexus}
+\end{table}
+
+We applied the QC pipeline to the ChIP-nexus experiments and found
+that the overall pattern seen in the visualizations is similar to a
+high quality ChIP-exo experiment. Figure \ref{fig:nexus}A illustrates
+a typical pattern of the ARC vs. URCR plot with ChIP-nexus data: Both
+arms seems to be more narrow and therefore the separation between them
+seems to be more noticeable, this seems reasonable since libraries
+generated by ChIP-nexus usually yields higher complexity than
+libraries generated by ChIP-exo. Additionally, while in ChIP-exo
+experiments is common to observe regions formed by undigested
+fragments in only one strand, we can observe that in ChIP-nexus
+experiment the occurrence of this type of regions is reduced. Figure
+\ref{fig:nexus}B shows that in a ChIP-nexus experiment the FSR
+distribution approximates the median more quickly than in a ChIP-exo
+experiment and Figure \ref{fig:nexus}C we observe that most of these
+single-stranded regions are formed by few fragments, which means that
+good ChIP-nexus experiment are less likely to show very imbalanced
+regions.
+
+\subsection{Evaluating ChIP-exo and ChIP-nexus data}
+
+<<all_param,include = FALSE,echo = FALSE,eval = TRUE>>=
+
+  Ntimes <- prettyNum(1e4,big.mark = ",")
+  Msamp <- prettyNum(1e3,big.mark = ",")
+  thresh1 <- 10
+
+@
+
+We applied the QC pipeline to all ChIP-exo and ChIP-nexus datasets
+listed in Tables \ref{tab:qc} and \ref{tab:nexus}, that way we
+partitioned the reads of each experiment into the undigested regions
+by keeping the regions formed by at least one fragment. Figure
+\ref{fig:enrich_all}A shows compares the slope of the following model:
+
+\begin{align}
+  \mbox{depth} = \beta \times \text{(\# of unique positions)} +\epsilon  \nonumber
+\end{align}
+
+when sampling \Sexpr{Msamp} regions and repeating this process
+\Sexpr{Ntimes} times for every ChIP-exo and ChIP-nexus experiment. The
+slope can be interpreted as the change in depth as number of unique
+position varies. This parameter measures the library complexity on a
+ChIP-exo experiment, since when it is high it means that several
+fragments are being aligned to the same position in the genome.
+
+Figures \ref{fig:enrich_all}B and \ref{fig:enrich_all}C shows the
+parameters of the following model:
+
+\begin{align}
+  \mbox{URCR} = \frac{\kappa}{\mbox{ARC}} + \eta + \epsilon \nonumber
+\end{align}
+
+considering only the regions composed by 10 or more aligned reads,
+that way the model reflects the enriched arm of the ARC vs. URCR
+visualization. The intercept $\eta$ can be interpreted as the
+theoretical $\mbox{URCR}$ for relatively high depth regions and the
+slope $\kappa$ can be interpreted as the decay in $\mbox{URCR}$ as the
+$\mbox{ARC}$ increases. That way experiments with low complexity are
+showing low values for both parameters. We considered a linear model
+between the estimated parameters for each experiment and it's
+respective depth for which we were able to asses that there is not
+significant relationship between the estimated parameters and the
+interaction between depth and organism for each experiment. Hence,
+these parameters may be used as indicators of the overall quality of
+ChIP-exo experiment, but as with the SCC it is recommendable to also
+visualize the SCC curve.
 
@@
 
+\color{red}
+\emph{need to add this part}
+\color{black}
+
 \subsection*{ChIP-exo experiments.}
@@
 calculated the FSR defined as the ratio between the number of forward
-stranded reads and the total number of reads in a region.
-
-Then we
+stranded reads and the total number of reads in a region. Then we
 transformed the FSR into an imbalance index is defined as:
@@
 \]
+
+We used the imbalance index to asses if the regions considered by the
+QC pipeline are less imbalanced when they overlap peak regions. Since
+the strand imbalance distribution for both classes have the same
+median and differ in the tails, this transformation separates the
+tails from the median. Hence, when we reject the hypothesis of the
+distribution being the same for both classes, it means that the class
+of region that don't overlap peaks shows heavier tails than its
+complement. \footnote{not sure if I should add the FSR densities for
+  both classes in the supplement to prove this statement}
+
+\subsection*{Evaluation of QC visualizations}
+
+{\color{blue} \emph{need to describe figure \ref{fig:enrich_all}
+}
+\color{black}
 
@@
   \centering
+  \includegraphics[width = .8\textwidth]{../figs/ChIPnexus/ChIPnexus_S2_MyC_enrichment.pdf}
+\newline
+  \includegraphics[width = .4\textwidth]{../figs/ChIPnexus/ChIPnexus_S2_MyC_fsr_surv.pdf}  
+  \includegraphics[width = .4\textwidth]{../figs/ChIPnexus/ChIPnexus_S2_MyC_label_surv.pdf}  
+  \caption{\textbf{ChIP-exo QC pipeline on ChIP-nexus data.} A)
+    Enrichment vs. library complexity, B) Strand imbalance and C)
+    Single-strand region analysis for the MyC experiment in S2 cell
+    lines in \emph{D. Melanogaster} from He et al., 2015
+    \cite{chipnexus}}
+  \label{fig:nexus}
+\end{figure}
+
+\newpage
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width = .9  \textwidth,page = 1]{../figs/test_quantify/slope_npos_depth_boxplot.pdf}
+  \includegraphics[width = .9  \textwidth,page = 1]{../figs/test_quantify/arc_urcr_param_regression.pdf}
+  \includegraphics[width = .9  \textwidth,page = 2]{../figs/test_quantify/arc_urcr_param_regression.pdf}
+  \caption{A) Comparison of change in depth as the number of unique
+    position obtained by sampling \Sexpr{Msamp} regions and fitting
+    the model $\mbox{depth} = \beta \times \text{(\# unique
+      positions)}$, and repeating this process \Sexpr{Ntimes} times. B)
+    Intercept and C) Slope in the the $\mbox{URCR} = k / \mbox{ARC} +
+    b$ model for the regions formed by more than 10 aligned
+    positions.}
+  \label{fig:enrich_all}
+\end{figure}
+
+\newpage
+\begin{figure}[h!]
+  \centering
   \includegraphics[width = .46\textwidth]{../figs/for_paper/sensitivity_exo_olda_data.pdf}
@@
   load("../data/paper_tables/Landick_rif_chipseq_summary.RData")
-  do.call(rbind,chipseq)
+  chipseq$set <- chipseq$set[,Organism := "E.Coli"]
+  chipseq$set <- chipseq$set[,Replicate := c(1,1,2,2)]
+  chipseq$set <- chipseq$set[,TF := "Sig70"]
+  chipseq$set <- chipseq$set[ ,cond := rep(c("Rif-0min","Rif-20min"),each = 2)]
+  chipseq$pet <- chipseq$pet[,Organism := "E.Coli"]
+  chipseq$pet <- chipseq$pet[,Replicate := c(1,1,2,2)]
+  chipseq$pet <- chipseq$pet[,TF := "Sig70"]
+  chipseq$pet <- chipseq$pet[ ,cond := rep(c("Rif-0min","Rif-20min"),each = 2)]
+  chipseq <- lapply(chipseq,format_tab)
+  chipseq <- do.call(rbind,chipseq)
+  chipseq <- chipseq[, Protocol := rep(c("PE","SE"),each = 4)]
+  chipseq
 @ 
@@
 \end{document}
+
+% LocalWords:  Ntimes

### 2016-04-12 rmarkdown/paper.tex
preliminary version of the paper, pre -change of sections

@@
 
-\subsubsection{Analysis of ChIP-Nexus data with the QC pipeline}
+\subsubsection{Analysis of ChIP-nexus data with the QC pipeline}
 
 ChIP-nexus developed by He et al., 2015 \cite{chipnexus} is a
-modification of the ChIP-exo protocol, where both sequencing adaptors
-are ligated at the $3\prime$ end of the ChIP fragments. Then,
-exonuclease digestion and DNA self-circularization with circLigase,
-and restriction enzyme cutting between the two adaptors to create the
-final library that is amplified.
+modification to the ChIP-exo protocol, where both sequencing adaptors
+are ligated at the $3\prime$ end of the ChIP fragments. Then, after
+exonuclease digestion, DNA self-circularization with circLigase, and
+restriction enzyme cutting between the two adaptors, the final library
+is amplified. In Table \ref{tab:nexus} we show the usual ChIP-Seq QC
+metrics to ChIP-nexus data and found that the conclusions made for
+ChIP-exo data still holds, the NSC is high but when plotting the SCC
+the fragment and read length peaks are confounded and the PBC may
+indicate severe bottlenecking but it is caused by the exonuclease
+enzyme digestion which reduces the number of positions where the
+reads can be aligned.
+
+\begin{table}[h!]
+  \centering
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{verbatim}
+##            Organism  IP/TF Condition/Cell Rep.       Depth    PBC     NSC
+##  1: D. Melanogaster Dorsal         Embryo    1   8,863,170 0.6766  7.2771
+##  2: D. Melanogaster Dorsal         Embryo    2  10,003,562 0.5657  7.1958
+##  3: D. Melanogaster  Twist         Embryo    1  18,244,203 0.6592  5.8241
+##  4: D. Melanogaster  Twist         Embryo    2  52,546,982 0.4550  5.2775
+##  5: D. Melanogaster    Max             S2    1  18,320,743 0.5178  3.6057
+##  6: D. Melanogaster    Max             S2    2  24,965,642 0.2124  3.4701
+##  7: D. Melanogaster    MyC             S2    1   7,832,034 0.3935  5.9203
+##  8: D. Melanogaster    MyC             S2    2  22,824,467 0.1879  5.7645
+##  9:           Human    TBP           K562    1  33,708,245 0.3102 32.1617
+## 10:           Human    TBP           K562    2 129,675,001 0.0492 32.7037
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+\caption{\textbf{Current QC metrics applied to ChIP-nexus data.} PBC
+  stands for PCR Bottleneck Coefficient and NSC for Normalized Strand
+  Cross-Correlation.}
+  \label{tab:nexus}
+\end{table}
+
+We applied the QC pipeline to the ChIP-nexus experiments and found
+that the overall pattern seen in the visualizations is similar to a
+high quality ChIP-exo experiment. Figure \ref{fig:nexus}A illustrates
+a typical pattern of the ARC vs. URCR plot with ChIP-nexus data: Both
+arms seems to be more narrow and therefore the separation between them
+seems to be more noticeable, this seems reasonable since libraries
+generated by ChIP-nexus usually yields higher complexity than
+libraries generated by ChIP-exo. Additionally, while in ChIP-exo
+experiments is common to observe regions formed by undigested
+fragments in only one strand, we can observe that in ChIP-nexus
+experiment the occurrence of this type of regions is reduced. Figure
+\ref{fig:nexus}B shows that in a ChIP-nexus experiment the FSR
+distribution approximates the median more quickly than in a ChIP-exo
+experiment and Figure \ref{fig:nexus}C we observe that most of these
+single-stranded regions are formed by few fragments, which means that
+good ChIP-nexus experiment are less likely to show very imbalanced
+regions.
+
+\subsection{Evaluating ChIP-exo and ChIP-nexus data}
+
+We applied the QC pipeline to all ChIP-exo and ChIP-nexus datasets
+listed in Tables \ref{tab:qc} and \ref{tab:nexus}, that way we
+partitioned the reads of each experiment into the undigested regions
+by keeping the regions formed by at least one fragment. Figure
+\ref{fig:enrich_all}A shows compares the slope of the following model:
+
+\begin{align}
+  \mbox{depth} = \beta \times \text{(\# of unique positions)} +\epsilon  \nonumber
+\end{align}
+
+when sampling 1,000 regions and repeating this process
+10,000 times for every ChIP-exo and ChIP-nexus experiment. The
+slope can be interpreted as the change in depth as number of unique
+position varies. This parameter measures the library complexity on a
+ChIP-exo experiment, since when it is high it means that several
+fragments are being aligned to the same position in the genome.
+
+Figures \ref{fig:enrich_all}B and \ref{fig:enrich_all}C shows the
+parameters of the following model:
+
+\begin{align}
+  \mbox{URCR} = \frac{\kappa}{\mbox{ARC}} + \eta + \epsilon \nonumber
+\end{align}
+
+considering only the regions composed by 10 or more aligned reads,
+that way the model reflects the enriched arm of the ARC vs. URCR
+visualization. The intercept $\eta$ can be interpreted as the
+theoretical $\mbox{URCR}$ for relatively high depth regions and the
+slope $\kappa$ can be interpreted as the decay in $\mbox{URCR}$ as the
+$\mbox{ARC}$ increases. That way experiments with low complexity are
+showing low values for both parameters. We considered a linear model
+between the estimated parameters for each experiment and it's
+respective depth for which we were able to asses that there is not
+significant relationship between the estimated parameters and the
+interaction between depth and organism for each experiment. Hence,
+these parameters may be used as indicators of the overall quality of
+ChIP-exo experiment, but as with the SCC it is recommendable to also
+visualize the SCC curve.
 
@@
   \centering
+  \includegraphics[width = .8\textwidth]{../figs/ChIPnexus/ChIPnexus_S2_MyC_enrichment.pdf}
+\newline
+  \includegraphics[width = .4\textwidth]{../figs/ChIPnexus/ChIPnexus_S2_MyC_fsr_surv.pdf}  
+  \includegraphics[width = .4\textwidth]{../figs/ChIPnexus/ChIPnexus_S2_MyC_label_surv.pdf}  
+  \caption{\textbf{ChIP-exo QC pipeline on ChIP-nexus data.} A)
+    Enrichment vs. library complexity, B) Strand imbalance and C)
+    Single-strand region analysis for the MyC experiment in S2 cell
+    lines in \emph{D. Melanogaster} from He et al., 2015
+    \cite{chipnexus}}
+  \label{fig:nexus}
+\end{figure}
+
+\newpage
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width = .9  \textwidth,page = 1]{../figs/test_quantify/slope_npos_depth_boxplot.pdf}
+  \includegraphics[width = .9  \textwidth,page = 1]{../figs/test_quantify/arc_urcr_param_regression.pdf}
+  \includegraphics[width = .9  \textwidth,page = 2]{../figs/test_quantify/arc_urcr_param_regression.pdf}
+  \caption{A) Comparison of change in depth as the number of unique
+    position obtained by sampling 1,000 regions and fitting
+    the model $\mbox{depth} = \beta \times \text{(\# unique
+      positions)}$, and repeating this process 10,000 times. B)
+    Intercept and C) Slope in the the $\mbox{URCR} = k / \mbox{ARC} +
+    b$ model for the regions formed by more than 10 aligned
+    positions.}
+  \label{fig:enrich_all}
+\end{figure}
+
+\newpage
+\begin{figure}[h!]
+  \centering
   \includegraphics[width = .46\textwidth]{../figs/for_paper/sensitivity_exo_olda_data.pdf}
@@
 \begin{verbatim}
-##                      files   nreads       pbc      nsc
-## 1: edsn1396_Sig70.sort.bam 13445022 0.9426179 8.865244
-## 2: edsn1398_Sig70.sort.bam 16538920 0.9378843 7.031836
-## 3: edsn1400_Sig70.sort.bam 11642722 0.8891744 10.77284
-## 4: edsn1402_Sig70.sort.bam 16854026 0.9407020 7.936239
-## 5: edsn1396_Sig70.sort.bam  6722511 0.6632742  9.01779
-## 6: edsn1398_Sig70.sort.bam  8269460 0.5594449 7.179539
-## 7: edsn1400_Sig70.sort.bam  5821361 0.6472382 10.89898
-## 8: edsn1402_Sig70.sort.bam  8427013 0.5895118 8.124717
+##    Organism IP/TF Condition/Cell Rep.      Depth    PBC     NSC Protocol
+## 1:   E.Coli Sig70       Rif-0min    1 13,445,022 0.9426  8.8652       PE
+## 2:   E.Coli Sig70       Rif-0min    1 16,538,920 0.9379  7.0318       PE
+## 3:   E.Coli Sig70      Rif-20min    2 11,642,722 0.8892 10.7728       PE
+## 4:   E.Coli Sig70      Rif-20min    2 16,854,026 0.9407  7.9362       PE
+## 5:   E.Coli Sig70       Rif-0min    1  6,722,511 0.6633  9.0178       SE
+## 6:   E.Coli Sig70       Rif-0min    1  8,269,460 0.5594  7.1795       SE
+## 7:   E.Coli Sig70      Rif-20min    2  5,821,361 0.6472 10.8990       SE
+## 8:   E.Coli Sig70      Rif-20min    2  8,427,013 0.5895  8.1247       SE
 \end{verbatim}

### 2016-04-12 rmarkdown/paper.R

@@ -7,6 +7,8 @@
   load("../data/paper_tables/Meijsing_hg19_GR.RData")
   load("../data/paper_tables/Carroll_hg19_ER.RData")
   load("../data/paper_tables/Venters_hg19_TBP.RData")
+  load("../data/paper_tables/ChIPnexus_dm3.RData")
+  load("../data/paper_tables/ChIPnexus_hg19.RData")
   landick_aero[,Organism := "E.Coli"]
   landick_aero[,Replicate := c(1,2,1,2)]
   landick_aero[,TF := "Sig70"]
@@ -35,7 +37,14 @@
   venters_hg19_TBP[,cond := "K562"]
   venters_hg19_TBP[,Replicate := 1:3]
   venters_hg19_TBP[,TF := "TBP"]
-
+  chipnexus_dm3[, Organism := "D. Melanogaster"]
+  chipnexus_dm3[, cond := c(rep("Embryo",4),rep("S2",4))]
+  chipnexus_dm3[, Replicate := rep(1:2,4)]
+  chipnexus_dm3[, TF := c(rep("Dorsal",2),rep("Twist",2),rep("Max",2),rep("MyC",2))]
+  chipnexus_K562[ , Organism := "Human"]
+  chipnexus_K562[ , cond  := "K562"]
+  chipnexus_K562[ , Replicate := 1:2]
+  chipnexus_K562[ , TF := "TBP"]
   ## this creates the table
   A <- do.call(rbind,list(landick_aero,
                      landick_rif[order(Replicate)],
@@ -44,13 +53,20 @@
                      meijing_hg19_GR[order(Replicate)],
                      carroll_hg19_ER[order(Replicate)],
                      venters_hg19_TBP[order(Replicate)]))
-  A[,nsc := unlist(nsc)]
-  setcolorder(A, c("Organism","TF","cond","Replicate","nreads","pbc","nsc","files"))
-  A[,files := NULL]
-  A[,nreads := prettyNum(nreads,big.mark = ",")]
-  A[,pbc := round(pbc,4)]
-  A[,nsc := round(nsc,4)]
-  setnames(A,names(A),c("Organism","IP/TF","Condition/Cell","Rep.","Depth","PBC","NSC"))
+  B <- do.call(rbind,list(chipnexus_dm3,
+                     chipnexus_K562))
+  format_tab <- function(A){
+    A[,nsc := unlist(nsc)]
[…]

### 2016-04-28 rmarkdown/paper.Rnw
added outline for paper

@@
 \usepackage{verbatim}
+
 % \usepackage{hyperref}
@@
 
-\usepackage{wrapfig}
-
 \newcommand{\beginsupplement}{%
@@
 \begin{document}
-
-%\maketitle
 
@@
 
-% \newpage
-
-% \tableofcontents
-% \listoffigures
+\newpage
+
+outline
+
+\begin{enumerate}
+\item Available data, the motivation of this part is two fold:
+  \begin{itemize}
+  \item we want to list all data sets we obtained
+  \item we want to emphasize that for the sig70 datasets we have a
+    unique design that allows us to compare chip exo with SE and PE
+    chipseq , the later one for the first time
+  \end{itemize}
+\item comparison with chip seq ? with and without dpeak ?
+\item only in methods. list the align algorithm used
+\item chip seq qc metrics
+\item chip exo qc pipeline, description and justification of the steps
+\begin{itemize}
+\item need to justify the sampling step. for this we may use the depth
+  table and add the factor to normalize one depth to another. explain
+  the confounding of pcr amplification and the reduction of the number
+  of possible mappable positions.
+\item case study with FoxA1, analysis of enrichment vs library
+  complexity and analysis and the strand imbalance tools. in methods
+  we need to add which tests did we perform to show that the pipeline
+  works. how do we expect from the diagnostics and how to interpret
+  the plots.
+\item analysis of why the pipeline works. in methods need to specify
+  how and in results need to specify why it works, and what are the
+  conclusions from it
+\end{itemize}
+\item high resolution analysis with chip exo and comparison with PE
+  chip seq
+  \begin{itemize}
+  \item justify what do we used dpeak (in methods need to list the
+    other methods used). perhaps we can add gof plots in
+    supplement. we also want to compare with the same method,
+    therefore we can use the fact that dpeak have a unified model for
+    PE and SE dpeak
+  \item after comparing with other methods we can compare with PE and
+    SE ChIP-seq, this comparison gives also resolution and sensitivity
+  \item compare samples under same budget of reads
+  \end{itemize} 
+\end{enumerate}
 
@@
 O_2$) condition, and treated by rifampicin by 0 and 20 minutes
-(courtesy of Professor Robert Landick's lab).
+(courtesy of Professor Robert Landick's lab). Additionally we gathered
+the ChIP-nexus from TATA-box Binding Protein (TBP) in K562 cell lines;
+MyC and Max in S2 cell lines in \emph{Drosophila Melanogaster}
+(\emph{D. Melanogaster}); Twist and Dorsal in \emph{D. Melanogaster}
+embryo cell lines (He et al., 2015 \cite{chipnexus}).
 
@@
 
-\subsection{Application of current ChIP-Seq QC guidelines on ChIP-exo
+\subsection{Quality Control of ChIP-exo and ChIP-nexus data}
+
+\subsubsection{Application of current ChIP-Seq QC guidelines on ChIP-exo
   data}
@@
   \centering
+%% need to re-build this table
 <<table1,include = FALSE,echo = FALSE,eval = TRUE>>=
@@
 
-\subsection{Comparison with ChIP-Seq data}
+\subsubsection{Comparison with ChIP-Seq data}
 
@@
 
-\subsection{ChIP-exo Quality Control Pipeline}
+\subsubsection{ChIP-exo Quality Control Pipeline}
 
@@
 
-\subsubsection{Enrichment and library complexity in ChIP-exo data}
+\textbf{Enrichment and library complexity in ChIP-exo data}
 
@@
 
-\subsubsection{Strand imbalance in ChIP-exo data}
+\textbf{Strand imbalance in ChIP-exo data}
 
@@
 
-\subsubsection{Analysis of ChIP-nexus data with the QC pipeline}
+\textbf{Analysis of ChIP-nexus data with the QC pipeline}
 
@@
 
-\subsection{Evaluating ChIP-exo and ChIP-nexus data}
+\subsubsection{Evaluating ChIP-exo and ChIP-nexus data}
 
@@
 
-\subsection{Comparison with ChIP-Seq data using dPeak}
+\subsection{High Resolution Binding Site Identification with dPeak and
+  ChIP-exo}
+
+\subsubsection{Comparison with ChIP-Seq data using dPeak}
 
@@
 
-\subsection{Systematic comparison of ChIP-Seq vs ChIP-exo under
+\subsubsection{Systematic comparison of ChIP-Seq vs ChIP-exo under
   varying sequencing depth}
@@
 
-\subsection{dPeak outperforms competing methods in discovering closely
+\subsubsection{dPeak outperforms competing methods in discovering closely
   spaced binding events from ChIP-exo and ChIP-Seq data}
@@
 
-\subsection*{Growth conditions.}
+outline
+
+\begin{enumerate}
+\item Available data, the motivation of this part is two fold:
+  \begin{itemize}
+  \item we want to list all data sets we obtained
+  \item we want to emphasize that for the sig70 datasets we have a
+    unique design that allows us to compare chip exo with SE and PE
+    chipseq , the later one for the first time
+  \end{itemize}
+\item comparison with chip seq ? with and without dpeak ?
+\item only in methods. list the align algorithm used
+\item chip seq qc metrics
+\item chip exo qc pipeline, description and justification of the steps
+\begin{itemize}
+\item need to justify the sampling step. for this we may use the depth
+  table and add the factor to normalize one depth to another. explain
+  the confounding of pcr amplification and the reduction of the number
+  of possible mappable positions.
+\item case study with FoxA1, analysis of enrichment vs library
+  complexity and analysis and the strand imbalance tools. in methods
+  we need to add which tests did we perform to show that the pipeline
+  works. how do we expect from the diagnostics and how to interpret
+  the plots.
+\item analysis of why the pipeline works. in methods need to specify
+  how and in results need to specify why it works, and what are the
+  conclusions from it
+\end{itemize}
+\item high resolution analysis with chip exo and comparison with PE
+  chip seq
+  \begin{itemize}
+  \item justify what do we used dpeak (in methods need to list the
+    other methods used). perhaps we can add gof plots in
+    supplement. we also want to compare with the same method,
+    therefore we can use the fact that dpeak have a unified model for
+    PE and SE dpeak
+  \item after comparing with other methods we can compare with PE and
+    SE ChIP-seq, this comparison gives also resolution and sensitivity
+  \item compare samples under same budget of reads
+  \end{itemize} 
+\end{enumerate}
+
+\subsection*{Considered data.}
+
+\subsubsection*{Growth conditions of generated data.}
 
@@
 
-\subsection*{ChIP-exo experiments.}
+\subsubsection{ChIP-exo and ChIP-nexus experiments.}
+
+% While the number of ChIP-exo data keeps increasing, characteristics of
+% ChIP-exo data are not fully investigated yet. First, DNA libraries
+% generated by the ChIP-exo protocol seem to be less complex than the
+% libraries generated by ChIP-Seq (Mahony et al., 2015
+% \cite{exo_review}). Second, although there are roughly the same amount
+% of reads in both strands, locally there may be more reads in one
+% strand than in the other. Finally, most of current ChIP-Seq quality
+% control (QC) guidelines (Landt et al., \cite{encode_qc}) may not be
+% applicable on ChIP-exo, while there are not established QC pipelines
+% for ChIP-exo; previous ChIP-exo analyses used ChIP-Seq samples to
+% compare the resolution between experiments (\cite{exo1},
+% \cite{exoillumina}, \cite{exo2})). To address these challenges, we
+% suggest a collection of quality control visualizations to interrogate
+% these biases in a ChIP-exo experiment and globally asses the
+% enrichment and library complexity of a ChIP-exo sample. We gathered
+% ChIP-exo data from diverse organisms: CTCF factor in human
+% \cite{exo1}; ER factor in human and FoxA1 factor in mouse (Serandour
+% et al., 2013 \cite{exoillumina}); Glucocorticoid receptor (GR) in
+% IMR90, K562 and U2OS cell lines (Starick et al., 2015
+% \cite{starick15}); TATA-box Binding Protein (TBP) in K562 cell lines
+% (Venters and Pugh, 2013 \cite{venters13}). Furthermore, we also
+% generated ChIP-exo and ChIP-Seq data for $\sig$ factor in
+% \emph{Escherichia Coli} (\emph{E. Coli}) measured under aerobic ($ +
+% O_2$) condition, and treated by rifampicin by 0 and 20 minutes
+% (courtesy of Professor Robert Landick's lab). Additionally we gathered
+% the ChIP-nexus from TATA-box Binding Protein (TBP) in K562 cell lines;
+% MyC and Max in S2 cell lines in \emph{Drosophila Melanogaster}
+% (\emph{D. Melanogaster}); Twist and Dorsal in \emph{D. Melanogaster}
+% embryo cell lines (He et al., 2015 \cite{chipnexus}).
 
@@
 
-\subsection*{Evaluation of QC visualizations}
-
-{\color{blue} \emph{need to describe figure \ref{fig:enrich_all}
-}
-\color{black}
+% \subsection*{Evaluation of QC visualizations}
+
+% Each boxplot in Figure \ref{fig:enrich_all}A, was created by sampling
+% \Sexpr{Msamp} of the regions generated and fitting the model:
+
+% \begin{align*}
+%   \mbox{depth} = \beta \times \text{(\# unique
+%       positions)}
+% \end{align*}
+
+% and repeating the process \Sexpr{Ntimes} times.
+
+%     Intercept and C) Slope in the the $\mbox{URCR} = k / \mbox{ARC} +
+%     b$ model f
 
@@
 algorithm as candidate regions. Default tuning parameters were used
-during model fitting for all methods. We were unable to use CexoR
-\cite{cexor} to estimate ChIP-exo binding sites.
-
-\bibliographystyle{bmc-mathphys} % Style BST file (bmc-mathphys, vancouver, spbasic).
+during model fitting for all methods. We downloaded \textbf{CexoR}
+\cite{cexor} version on its 1.8 version from \emph{bioconductor} but
+were unable to use it for the $\sig$ experiments.
+
+ % Style BST file (bmc-mathphys, vancouver, spbasic).
+\bibliographystyle{vancouver}
 \bibliography{chip_exo_paper}
@@
   .95 \textwidth]{/p/keles/ChIPexo/volume3/ChIPexo/figs/for_paper/chip_explanation_withEnzyme.pdf}
-  \caption{\textbf{Description of ChIP-exo, SE ChIP-Seq and PE
-      ChIP-Seq:} A TF is bound to the forward (red) and backward
-    (blue) strand of the DNA. Then, is sonicated: For ChIP-exo a
-    exonuclease enzyme (orange hexagon) trims the $5\prime$ ends of
-    each DNA fragment to a fixed distance from the bound protein,
-    finally is subjected to Immunoprecipitation and amplification. For
-    both ChIP-exo and SE ChIP-Seq an adapter is ligated (green
-    triangles) at the $5\prime$ ends, while for PE ChIP-Seq is ligated
-    to both ends.}
+  % \caption{\textbf{Description of ChIP-exo, SE ChIP-Seq and PE
+  %     ChIP-Seq:} A TF is bound to the forward (red) and backward
+  %   (blue) strand of the DNA. Then, is sonicated: For ChIP-exo a
+  %   exonuclease enzyme (orange hexagon) trims the $5\prime$ ends of
+  %   each DNA fragment to a fixed distance from the bound protein,
+  %   finally is subjected to Immunoprecipitation and amplification. For
+  %   both ChIP-exo and SE ChIP-Seq an adapter is ligated (green
+  %   triangles) at the $5\prime$ ends, while for PE ChIP-Seq is ligated
+  %   to both ends.}
+  \caption{A TF is bound to the forward (red) and backward (blue)
+    strand of the DNA. Then, is sonicated: For ChIP-exo a exonuclease
+    enzyme (orange hexagon) trims the $5\prime$ ends of each DNA
+    fragment to a fixed distance from the bound protein, finally is
+    subjected to Immunoprecipitation and amplification. For both
+    ChIP-exo and SE ChIP-Seq an adapter is ligated (green triangles)
+    at the $5\prime$ ends, while for PE ChIP-Seq is ligated to both
+    ends.}
   \label{fig:chip_diagram}
@@
 \includegraphics[width = .8\textwidth]{/p/keles/ChIPexo/volume3/ChIPexo/figs/for_paper/EColi_strand_cross_corr.pdf}
-\caption{\textbf{SCC curves for $\sig$ samples.} The ``phantom peak''
-  and the summit that corresponds to the read and fragment length
-  respectively are confounded due to the exonuclease digestion.}
+% \caption{\textbf{SCC curves for $\sig$ samples.} The ``phantom peak''
+%   and the summit that corresponds to the read and fragment length
+%   respectively are confounded due to the exonuclease digestion.}
+\caption{SCC curves for $\sig$ samples. The ``phantom peak'' and the
+  summit that corresponds to the read and fragment length respectively
+  are confounded due to the exonuclease digestion.}
   \label{fig:scc_exo}
@@
   \includegraphics[width = \textwidth]{../figs/for_paper/coverage_diagram3.pdf}
-  \caption{\textbf{Flowchart of the ChIP-exo quality control
-      pipeline.} A) ARC vs. URCR, this plot presents a global view of
+  % \caption{\textbf{Flowchart of the ChIP-exo quality control
+  %     pipeline.} A) ARC vs. URCR, this plot presents a global view of
+  %   the balance between library complexity and enrichment. There are
+  %   two arms: First, one with low ARC, which corresponds to regions
+  %   formed by few aligned positions; Second, where the URCR decreases
+  %   as the ARC increases. B) Min. depth vs Proportions of regions,
+  %   this plot shows the strand composition for all the regions formed
+  %   by a min. number of reads. Regions with low depth tend to be
+  %   formed by undigested reads from one unique strand, while regions
+  %   with higher signal are usually formed by reads in both strands. C)
+  %   Min depth vs. FSR, this plot depicts how quickly the FSR's
+  %   distribution approximate the median. In a high quality sample, the
+  %   median is around 0.5, and the other quantiles reach that value
+  %   quickly.}
+  \caption{The ChIP-exo reads are partitioned by keeping the regions
+    in the genome formed by the undigested fragments. For each region,
+    we calculate a collection of summary statistics, which are
+    visualized as A) ARC vs. URCR, this plot presents a global view of
     the balance between library complexity and enrichment. There are
@@
   \includegraphics[width = .4\textwidth]{../figs/ChIPnexus/ChIPnexus_S2_MyC_label_surv.pdf}  
-  \caption{\textbf{ChIP-exo QC pipeline on ChIP-nexus data.} A)
-    Enrichment vs. library complexity, B) Strand imbalance and C)
-    Single-strand region analysis for the MyC experiment in S2 cell
+  % \caption{\textbf{ChIP-exo QC pipeline on ChIP-nexus data.} A)
+  %   Enrichment vs. library complexity, B) Strand imbalance and C)
+  %   Single-strand region analysis for the MyC experiment in S2 cell
+  %   lines in \emph{D. Melanogaster} from He et al., 2015
+  %   \cite{chipnexus}}
+  \caption{We applied the QC pipeline on ChIP-nexus data. A) ARC
+    vs. URCR hexbin plots, B) FSR distribution quantiles and C)
+    Stacked histogram with the proportion of regions composed by
+    fragments on a single strand for the MyC experiment in S2 cell
     lines in \emph{D. Melanogaster} from He et al., 2015
@@
   \includegraphics[width = .46\textwidth]{../figs/for_paper/resolution_by_dataset_old_data.pdf}
-   \caption{Comparison of (A) sensitivity and (B) resolution between
+  \caption{Comparison of A) sensitivity and B) resolution between
     ChIP-exo and ChIP-Seq data. Sensitivity is defined as the
@@
   \includegraphics[width = .9\textwidth]{../figs/for_paper/saturation_analysis_old.pdf}
-  \caption{Comparison of the number of candidate regions (A),
-    predicted events (B), identified targets (C) and resolution (D)
-    among ChIP-exo, PE ChIP-Seq and SE ChIP-Seq. RegulonDB annotations
-    are considered as a gold standard. A gold standard binding events
-    was deemed identified if a binding event was estimated at a $\pm$
+  \caption{Comparison of the number of A) candidate regions, B)
+    predicted events, C) identified targets and D) resolution among
+    ChIP-exo, PE ChIP-Seq and SE ChIP-Seq. RegulonDB annotations are
+    considered as a gold standard. A gold standard binding events was
+    deemed identified if a binding event was estimated at a $\pm$
     \Sexpr{ext} vicinity of it.}
@@
 @ 
-\caption{Same QC metrics as in table \ref{tab:qc} but applied to
-  Landick's chipseq data of the rif experiment}
+\caption{Usual ChIP-Seq QC metrics as in table \ref{tab:qc} but
+  applied to Landick's chipseq data of the rif experiment}
 \end{table}
@@
   \includegraphics[width = .9\textwidth,page = 1]{../figs/saturation/K12_alignment/sig70_rif_enrichment.pdf}
-  \caption{A) ARC vs URCR hexbin plots of Rep-1 and rif-0min from $\sig$ experiment when 100K to 900K reads are being
-    sampled for each panel. }
+  \caption{ARC vs URCR hexbin plots of Rep-1 and rif-0min from $\sig$
+    experiment when 100K to 900K reads are being sampled for each
+    panel. }
   \label{fig:exoQC_sat1}

### 2016-04-28 rmarkdown/paper.tex
added outline for paper

@@
   \centering
+%% need to re-build this table
 
@@
 
-\subsection*{Evaluation of QC visualizations}
-
-{\color{blue} \emph{need to describe figure \ref{fig:enrich_all}
-}
-\color{black}
+% \subsection*{Evaluation of QC visualizations}
+
+% Each boxplot in Figure \ref{fig:enrich_all}A, was created by sampling
+% Msamp of the regions generated and fitting the model:
+
+% \begin{align*}
+%   \mbox{depth} = \beta \times \text{(\# unique
+%       positions)}
+% \end{align*}
+
+% and repeating the process Ntimes times.
+
+%     Intercept and C) Slope in the the $\mbox{URCR} = k / \mbox{ARC} +
+%     b$ model f
 
@@
   \includegraphics[width = .9\textwidth]{../figs/for_paper/saturation_analysis_old.pdf}
-  \caption{Comparison of the number of candidate regions (A),
-    predicted events (B), identified targets (C) and resolution (D)
-    among ChIP-exo, PE ChIP-Seq and SE ChIP-Seq. RegulonDB annotations
-    are considered as a gold standard. A gold standard binding events
-    was deemed identified if a binding event was estimated at a $\pm$
+  \caption{Comparison of the number of A) candidate regions, B)
+    predicted events, C) identified targets and D) resolution among
+    ChIP-exo, PE ChIP-Seq and SE ChIP-Seq. RegulonDB annotations are
+    considered as a gold standard. A gold standard binding events was
+    deemed identified if a binding event was estimated at a $\pm$
     15 vicinity of it.}
@@
 \end{knitrout}
-\caption{Same QC metrics as in table \ref{tab:qc} but applied to
-  Landick's chipseq data of the rif experiment}
+\caption{Usual ChIP-Seq QC metrics as in table \ref{tab:qc} but
+  applied to Landick's chipseq data of the rif experiment}
 \end{table}

### 2016-05-02 rmarkdown/paper.Rnw
fixed methods para in paper

@@
 
-outline
-
-\begin{enumerate}
-\item Available data, the motivation of this part is two fold:
-  \begin{itemize}
-  \item we want to list all data sets we obtained
-  \item we want to emphasize that for the sig70 datasets we have a
-    unique design that allows us to compare chip exo with SE and PE
-    chipseq , the later one for the first time
-  \end{itemize}
-\item comparison with chip seq ? with and without dpeak ?
-\item only in methods. list the align algorithm used
-\item chip seq qc metrics
-\item chip exo qc pipeline, description and justification of the steps
-\begin{itemize}
-\item need to justify the sampling step. for this we may use the depth
-  table and add the factor to normalize one depth to another. explain
-  the confounding of pcr amplification and the reduction of the number
-  of possible mappable positions.
-\item case study with FoxA1, analysis of enrichment vs library
-  complexity and analysis and the strand imbalance tools. in methods
-  we need to add which tests did we perform to show that the pipeline
-  works. how do we expect from the diagnostics and how to interpret
-  the plots.
-\item analysis of why the pipeline works. in methods need to specify
-  how and in results need to specify why it works, and what are the
-  conclusions from it
-\end{itemize}
-\item high resolution analysis with chip exo and comparison with PE
-  chip seq
-  \begin{itemize}
-  \item justify what do we used dpeak (in methods need to list the
-    other methods used). perhaps we can add gof plots in
-    supplement. we also want to compare with the same method,
-    therefore we can use the fact that dpeak have a unified model for
-    PE and SE dpeak
-  \item after comparing with other methods we can compare with PE and
-    SE ChIP-seq, this comparison gives also resolution and sensitivity
-  \item compare samples under same budget of reads
-  \end{itemize} 
-\end{enumerate}
-
 \subsection*{Considered data.}
 
+% To asses the validity of the ChIP-exo QC pipeline several experiments
+% were considered.
+
 \subsubsection*{Growth conditions of generated data.}
 
-\color{red}
-\emph{need to add this part}
-\color{black}
-
-\subsubsection{ChIP-exo and ChIP-nexus experiments.}
-
-% While the number of ChIP-exo data keeps increasing, characteristics of
-% ChIP-exo data are not fully investigated yet. First, DNA libraries
-% generated by the ChIP-exo protocol seem to be less complex than the
-% libraries generated by ChIP-Seq (Mahony et al., 2015
-% \cite{exo_review}). Second, although there are roughly the same amount
-% of reads in both strands, locally there may be more reads in one
-% strand than in the other. Finally, most of current ChIP-Seq quality
-% control (QC) guidelines (Landt et al., \cite{encode_qc}) may not be
-% applicable on ChIP-exo, while there are not established QC pipelines
-% for ChIP-exo; previous ChIP-exo analyses used ChIP-Seq samples to
-% compare the resolution between experiments (\cite{exo1},
-% \cite{exoillumina}, \cite{exo2})). To address these challenges, we
-% suggest a collection of quality control visualizations to interrogate
-% these biases in a ChIP-exo experiment and globally asses the
-% enrichment and library complexity of a ChIP-exo sample. We gathered
-% ChIP-exo data from diverse organisms: CTCF factor in human
-% \cite{exo1}; ER factor in human and FoxA1 factor in mouse (Serandour
-% et al., 2013 \cite{exoillumina}); Glucocorticoid receptor (GR) in
-% IMR90, K562 and U2OS cell lines (Starick et al., 2015
-% \cite{starick15}); TATA-box Binding Protein (TBP) in K562 cell lines
-% (Venters and Pugh, 2013 \cite{venters13}). Furthermore, we also
-% generated ChIP-exo and ChIP-Seq data for $\sig$ factor in
-% \emph{Escherichia Coli} (\emph{E. Coli}) measured under aerobic ($ +
-% O_2$) condition, and treated by rifampicin by 0 and 20 minutes
-% (courtesy of Professor Robert Landick's lab). Additionally we gathered
-% the ChIP-nexus from TATA-box Binding Protein (TBP) in K562 cell lines;
-% MyC and Max in S2 cell lines in \emph{Drosophila Melanogaster}
-% (\emph{D. Melanogaster}); Twist and Dorsal in \emph{D. Melanogaster}
-% embryo cell lines (He et al., 2015 \cite{chipnexus}).
+We generated ChIP-exo and PE ChIP-Seq samples from $\sig$ in
+\emph{E.Coli}. For each PE ChIP-Seq experiment, a \emph{in silico} SE
+version was obtained by randomly sampling one of the two ends in each
+fragment.
+
+{\color{red}\emph{need to add the growth conditions}}
+
+\subsubsection*{ChIP-exo and ChIP-nexus experiments.}
+
+We gathered a collection of ChIP-exo experiments spanning different
+cell lines and transcription factors: CTCF in HeLA cells \cite{exo1},
+ER in MCF-7 cells \cite{exoillumina}, TBP in K562 cells
+\cite{venters13}, GR in IMR90, K562 and U2OS in K562 cells
+\cite{starick15} and FoxA1 in mouse liver cells
+\cite{exoillumina}. Additionally we used the ChIP-nexus experiments
+generated for He et al., 2015: TBP in human K562, MyC and Max factors
+in S2 cell lines in \emph{D. Melanogaster}; Twist and Dorsal factors
+in \emph{D. Melanogaster} embryo cell lines.
+
+The read files were aligned following the instructions provided in
+their respective sources when available. Otherwise we used
+\emph{bowtie} in its 1.1.2 version.
 
@@
 
-The statistics in Table \ref{tab:qc} and the SCC curves from Figure
-\ref{fig:scc_exo} were calculated with the \textbf{ChIPUtils} package
-version 0.99.0, available in
-\url{https://github.com/welch16/ChIPUtils}. This package provides an
-easy to use interface to calculate basic quality control metrics and
-diagnostic plots for ChIP-Seq data.
+% scc and related
+% change table to have pbc at the end
+
+We considered the definitions of ChIP-Seq QC guidelines provided in
+Landt et al., 2012 which are also listed in
+\url{https://genome.ucsc.edu/ENCODE/qualityMetrics.html}.
+
+\subsubsection*{Strand Cross-Correlation.}
+
+The strand cross-correlation was proposed by Kharchenko et al., 2008
+\cite{strandcc} and it may be one of the most used of the ChIP-Seq QC
+metrics. The SCC curve is defined as:
+
+\begin{align}
+  y(\delta) = \sum_c w_c r\left[ n_c^+ \left(x + \frac{\delta}{2}
+    \right), n_c^- \left( x- \frac{\delta}{2} \right)\right],
+\label{scc}
+\end{align}
+
+where $y(\delta)$ is the SCC for a strand shift $\delta$, $r$ is the
+Pearson correlation, $w_c$ is the proportion of reads mapped to
+chromosome $c$ and $n_c^S$ is the read count vector for strand $S$ and
+chromosome $c$. The following QC metrics are used to summarize the shape of the SCC curve:
+
+\begin{align}
+  \mbox{NSC} &= \frac{\max_\delta y(\delta)}{\min_\delta y(\delta)}, \label{nsc} \\
+  \mbox{RSC} &= \frac{\max_\delta y(\delta) -
+    y_{\text{bgd}}}{y_{\text{rl}} - y_{\text{bgd}}}. \label{rsc}
+\end{align}
+
+where $y_{\text{bgd}}$ is estimated as the background SCC level is
+defined as $y_{\text{bgd}} = \min_\delta y(\delta)$ and $y_{\text{rl}}$
+  is the value of the SCC curve when the shift equals the experiment's
+  read length.
 
@@
 
-\subsubsection*{Strand Cross-Correlation.}
-
-The strand cross-correlation was proposed by Kharchenko et al., 2008
-\cite{strandcc} and it may be one of the most used of the ChIP-Seq QC
-metrics. The SCC curve is defined as:
-
-\begin{align}
-  y(\delta) = \sum_c w_c r\left[ n_c^+ \left(x + \frac{\delta}{2}
-    \right), n_c^- \left( x- \frac{\delta}{2} \right)\right],
-\label{scc}
-\end{align}
-
-where $y(\delta)$ is the SCC for a strand shift $\delta$, $r$ is the
-Pearson correlation, $w_c$ is the proportion of reads mapped to
-chromosome $c$ and $n_c^S$ is the read count vector for strand $S$ and
-chromosome $c$. Additionally, two QC metrics are defined:
-
-\begin{align}
-  \mbox{NSC} &= \frac{\max_\delta y(\delta)}{\min_\delta y(\delta)}, \label{nsc} \\
-  \mbox{RSC} &= \frac{\max_\delta y(\delta) -
-    y_{\text{bgd}}}{y_{\text{rl}} - y_{\text{bgd}}}. \label{rsc}
-\end{align}
-
-where $y_{\text{bgd}}$ is the background SCC level and $y_{\text{rl}}$
-is the SCC ``phantom peak'' value.
-
-\subsection*{Mappability and GC content scores.}
+These QC guidelines were calculated with the \textbf{ChIPUtils}
+package in its 0.99.0 version (available in
+\url{https://github.com/welch16/ChIPUtils}). This package provides an
+easy to use interface to calculate basic quality control metrics and
+diagnostic plots for ChIP-Seq data.
+
+\subsection*{ChIP-exo quality control pipeline.}
+
+The ChIP-exo QC pipeline requires a set of aligned reads from a
+ChIP-exo (or ChIP-nexus) experiment. Then:
+
+\begin{enumerate}
+\item Partitions the experiment's coverage by keeping the regions
+  formed by one or more aligned reads.
+\item For each region, it counts the number of reads that are being
+  aligned to the region and the number of positions to which at least
+  one mapping read is aligned in each strand.
+\item For each region, it calculates the following statistics:
+  \begin{align*}
+    \mbox{ARC} &= \frac{\text{Total nr. of reads mapped to the region}}{\text{Width of the region}} \\
+    \mbox{URCR} &= \frac{\text{Total nr. of position that at least
+        one read is aligned in the region}}{\text{Total nr. of reads mapped 
+              to the region}} \\
+    \mbox{FSR} &= \frac{\text{Nr. of fwd. strand reads mapped to the region}}                 {\text{Total nr. of reads mapped to the region}} 
+  \end{align*}
+\item The ChIP-exo QC pipeline creates visualizations to summarize the
+  summary statistics and calculates the contribution of the number of
+  unique position and the width to the number of reads per region by
+  randomly sampling $M$ regions to fit the model:
+  \begin{align*}
+    \mbox{depth} = \beta_1 \mbox{npos} + \beta_2 \mbox{width} + \varepsilon
+  \end{align*}
+  where $\varepsilon \sim N(0,\sigma^2)$, $\beta_1$ and $\beta_2$ are
+  the contribution from the number of unique positions and width to
+  the total number of reads per regions respectively; this process is
+  repeated independently $N$ times. 
+\end{enumerate}
+
+We implemented this ChIP exo QC pipeline in the \emph{ChIPexoQual} R
+package, additionally it generates visualizations as seen in Figure
+\ref{fig:qcdiagram}. We used its 1.0 version which is available in
+\url{https://github.com/welch16/ChIPexoqual}.
+
+\subsubsection*{Motif analysis of FoxA1 enriched regions}
+
+For each replicate, we used the ChIP-exo QC pipeline to partition the
+mouse genome into a set of regions with their respective summary
+statistics; we filtered them into collections of high quality regions
+by:
+
+\begin{enumerate}
+\item Removing the regions formed by reads on only one strand.
+\item Removing regions with reads aligned to at most 15 unique
+  positions.
+\item Removing regions with less that 100 reads.
+\end{enumerate}
+
+Then we used \textbf{FIMO} on its 4.9.1 version \cite{fimo} to find
+the FoxA1 motif (with id MA0148.1 in the core JASPAR database).
+
+\subsubsection*{Imbalance test of FoxA1 replicates}
+
+For each replicate, we used the ChIP-exo QC pipeline to partition the
+mouse genome into a set of regions with their respective summary
+statistics. We filtered the regions with reads in only one strand and
+we transformed the FSR into an \emph{Imbalance index} that is zero
+when the region is perfectly balanced and infinity when it consists of
+reads in one strand exclusively:
+
+\begin{align*}
+  \mbox{Imbalance index} = -\log_{10} (4 \times \mbox{FSR} \times (1 -
+  \mbox{FSR}))
+\end{align*}
+
+We divided the ChIP-exo regions onto two classes by considering
+whether it overlaps a ChIP-exo peak of not. For each replicate we
+called peaks with the $\mbox{GC content} + \mbox{Mappability}$ model
+from MOSAiCS \cite{mosaics} (version 2.9.7) with bin size and fragment
+length of 200 bp. We called peaks with an FDR of $5\%$, a threshold of
+100 and a maximum gap size of 200 bp. We further filtered the peaks by
+keeping only the peaks with an average ChIP count of 200 extended
+reads.
+
+To show that the class that don't overlap with peaks exhibits heavier
+tails, we used a Wilcoxon test over the \emph{Imbalance index}.
+
+\subsubsection*{GC - content and Mappability}
 
 To define the mappability score we follow the definition from Rozowsky
-et al., 2009 \cite{peakseq}:
-
-\begin{align}
+et al., 2009:
+
+\begin{align*}
   m_i &= \sum_{k = i - L +1}^{i + L - 1} \frac{\delta_k }{2L - 1}.
-\end{align}
+\end{align*}
 
@@
 
-\subsection*{ChIP-exo quality control pipeline.}
-
-We used the R package \textbf{ChIPexoQual} to assess the quality of
-the ChIP-exo datasets by following the steps described in Figure
-\ref{fig:qcdiagram}. We used version 1.0, and it is available in
-\url{https://github.com/welch16/ChIPexoqual}.
-
-\subsection*{Motif analysis of Fox A1 enriched regions}
-
-For all three replicates we called peaks using the MOSAiCS GC +
-Mappability model using an FDR level of 5\%, filtering out the peaks
-with average ChIP counts below one hundred fragments and merging peaks
-gaped by at most 200 bp. Then, we fitted the dPeak model considering
-at most 5 binding events for each peak, and we searched for the Fox A1
-motif over a 10 bp window around the estimated binding events. We used
-FIMO's command line 4.9.1 version \cite{fimo}.
-
-\subsection*{Imbalance index}
-
-For every ChIP-exo experiment, we partitioned the experiment into the
-non-digested regions using the QC pipeline. For each region we
-calculated the FSR defined as the ratio between the number of forward
-stranded reads and the total number of reads in a region. Then we
-transformed the FSR into an imbalance index is defined as:
-
-\[
-\mbox{Imbalance index} = -\log_{10} (4 \times \mbox{FSR} \times (1 - \mbox{FSR}))
-\]
-
-We used the imbalance index to asses if the regions considered by the
-QC pipeline are less imbalanced when they overlap peak regions. Since
-the strand imbalance distribution for both classes have the same
-median and differ in the tails, this transformation separates the
-tails from the median. Hence, when we reject the hypothesis of the
-distribution being the same for both classes, it means that the class
-of region that don't overlap peaks shows heavier tails than its
-complement. \footnote{not sure if I should add the FSR densities for
-  both classes in the supplement to prove this statement}
-
-% \subsection*{Evaluation of QC visualizations}
-
-% Each boxplot in Figure \ref{fig:enrich_all}A, was created by sampling
-% \Sexpr{Msamp} of the regions generated and fitting the model:
-
-% \begin{align*}
-%   \mbox{depth} = \beta \times \text{(\# unique
-%       positions)}
-% \end{align*}
-
-% and repeating the process \Sexpr{Ntimes} times.
-
-%     Intercept and C) Slope in the the $\mbox{URCR} = k / \mbox{ARC} +
-%     b$ model f
-
-\subsection*{Construction of a SE ChIP-Seq from a PE ChIP-Seq experiment.}
-
-For the rif-treatment ChIP-Seq experiments, we sampled SE ChIP-Seq
-experiment from the PE ones by taking one of both ends randomly
-following with equal probability.
-
-\subsection*{dPeak analysis of $\sig$ ChIP-exo and ChIP-Seq data.}
-
-For the resolution and sensitivity analysis, we used the MOSAiCS GC
-content + mappability model to call peaks for ChIP-exo experiments,
-while for both SE and PE ChIP-Seq experiments we used the MOSAiCS
-Input model. To avoid false positives we only considered ChIP-exo
-peaks with average ChIP counts greater than 3000 that overlapped both
-the SE and PE ChIP-Seq peaks, we considered other cutoff values but
-still obtained results similar to what we presented in this
-paper. Then we estimated the binding site events using the dPeak model
-with a maximum number of 5 binding events. We considered RegulonDB
-annotation as gold-standard; Resolution is defined as the minimum
-distance from an annotation to an estimated binding events and
-Sensitivity is defined as the fraction of annotation within
-\Sexpr{ext} bp from an estimated binding event.
-
-\subsection*{Saturation analysis of ChIP-exo, PE ChIP-Seq and SE ChIP-Seq.}
-
-To perform the saturation analysis, we sub-sampled $N$ fragments for
-both ChIP-exo and SE ChIP-Seq protocols. For PE ChIP-Seq we
-sub-sampled $N$ pairs or $N/2$ fragments. For each seed, we called
-peaks using MOSAiCS \cite{mosaics} (GC content + mappability for
-ChIP-exo and Input for SE and PE ChIP-Seq) for the maximum sample size
-and to avoid false positives we considered only the top 500 peaks for
-each data protocol. We defined the number of candidate regions as the
-number of top sample peaks such that a binding events was estimated
-using the sampled reads and the dPeak's model; the number of predicted
-events is the total quantity of binding events estimated using the
-dPeak's model; the number of identified targets are number of
-gold-standard annotations within \Sexpr{ext} bp from an estimated binding
-events; and the resolution is defined as the minimum distance from a
-gold-standard annotation to an estimated binding event. We repeated
-this analysis for ten seeds and reported the median between all those
-values.
-
-\subsection*{Method comparison for ChIP-exo.}
-
-We considered dPeak Chung et al., 2013 \cite{dpeak}, GEM Guo et al.,
-2012 \cite{gem}, MACE Wang et al., 2014 \cite{mace} and Peazilla
-Bardet et al., 2013 \cite{peakzilla} for the ChIP-exo data
-analysis. for the dPeak algorithm we used the R package \textbf{dPeak}
-version 2.0.1 which is available from
+% Hence, when we reject the hypothesis of the
+% distribution being the same for both classes, it means that the class
+% of region that don't overlap peaks shows heavier tails than its
+% complement. \footnote{not sure if I should add the FSR densities for
+%   both classes in the supplement to prove this statement}
+
+% \subsection*{Construction of a SE ChIP-Seq from a PE ChIP-Seq experiment.}
+
+% For the rif-treatment ChIP-Seq experiments, we sampled SE ChIP-Seq
+% experiment from the PE ones by taking one of both ends randomly
+% following with equal probability.
+
+\subsection*{High resolution analysis with ChIP-exo}
+
+We considered RegulonDB \cite{regulondb} annotations as gold-standard
+and considered an annotation as being identified if the distance to a
+estimated binding events is less or equal than \Sexpr{id_dist} bp. We
+defined the resolution as the distance from an annotation to its
+closest predictions and  the sensitivity as the fraction of identified
+annotations in a genomic region.
+
+\subsubsection*{Method comparison for ChIP-exo.}
+
+We compared dPeak, Chung et al., 2013; GEM, Guo et al., 2012; MACE,
+Wang et al., 2014 and Peazilla, Bardet et al., 2013 for the ChIP-exo
+data analysis. For the dPeak algorithm we used the R package
+\textbf{dPeak} version 2.0.1 which is available from
 \url{https://github.com/dongjunchung/dpeak}. For the GEM algorithm, we
@@
 \url{https://github.com/steinmann/peakzilla}. Candidate regions for
-\textbf{dPeak} were identified for each replicate of ChIP-exo data
-using the \textbf{MOSAiCS} algorithm Kuan et al., 2011 \cite{mosaics}
-(one sample analysis using false discovery rate of \Sexpr{fdr}\%)
+\textbf{dPeak} and \textbf{GEM} were identified for each replicate of
+ChIP-exo data using the \textbf{MOSAiCS} algorithm \cite{mosaics} (one
+sample analysis using false discovery rate of \Sexpr{fdr}\%)
 implemented as an R package \textbf{mosaics} version 2.9.7 (available
@@
 during model fitting for all methods. We downloaded \textbf{CexoR}
-\cite{cexor} version on its 1.8 version from \emph{bioconductor} but
-were unable to use it for the $\sig$ experiments.
+\cite{cexor} on its 1.8 version from \emph{bioconductor} but were
+unable to use it for the $\sig$ experiments.
+
+\subsubsection*{dPeak analysis of $\sig$ ChIP-exo and ChIP-Seq data.}
+
+We compared the estimated binding events predicted by the MOSAiCS +
+dPeak pipeline using reads generated by ChIP-exo, PE and SE ChIP-Seq
+protocols. We called peaks at a $5\%$ FDR level using MOSAiCS (the
+$\mbox{GC content} + \mbox{Mappability}$ model for ChIP-exo and the
+$\mbox{Input only}$ model for PE and SE ChIP-Seq). Then, we
+deconvolved the peak into binding events with dPeak (version 2.0.1) by
+considering a maximum of 5 binding sites on each peak. To avoid false
+positives we only considered ChIP-exo peaks with average ChIP counts
+greater than 3000 that overlapped both the SE and PE ChIP-Seq peaks,
+we considered other cutoff values but still obtained results similar
+to what we presented in this paper.
+
+\subsubsection*{Saturation analysis of ChIP-exo, PE ChIP-Seq and SE ChIP-Seq.}
+
+We sub-sampled $N$ fragments for both ChIP-exo and SE ChIP-Seq
+protocols. For PE ChIP-Seq we sub-sampled $N$ pairs or $N/2$
+fragments. For each seed, we called peaks using MOSAiCS \cite{mosaics}
+($\mbox{GC content} + \mbox{Mappability}$ for ChIP-exo and
+$\mbox{Input only}$ for SE and PE ChIP-Seq) for the maximum sample
+size and to avoid false positives we considered only the top 500 peaks
+for each data protocol. We defined the number of candidate regions as
+the number of top sample peaks such that a binding events was
+estimated using the sampled reads and the dPeak's model; the number of
+predicted events is the total quantity of binding events estimated
+using the dPeak's model; the number of identified targets are number
+of gold-standard annotations within \Sexpr{ext} bp from an estimated
+binding events; and the resolution is defined as the minimum distance
+from a gold-standard annotation to an estimated binding event. We
+repeated this analysis for ten seeds and reported the median between
+all those values.
 

### 2016-05-02 rmarkdown/paper.tex
fixed methods para in paper

@@
 
-\subsection*{ChIP-exo quality control pipeline.}
-
-We used the R package \textbf{ChIPexoQual} to assess the quality of
-the ChIP-exo datasets by following the steps described in Figure
-\ref{fig:qcdiagram}. We used version 1.0, and it is available in
-\url{https://github.com/welch16/ChIPexoqual}.
-
-\subsection*{Motif analysis of Fox A1 enriched regions}
-
-For all three replicates we called peaks using the MOSAiCS GC +
-Mappability model using an FDR level of 5\%, filtering out the peaks
-with average ChIP counts below one hundred fragments and merging peaks
-gaped by at most 200 bp. Then, we fitted the dPeak model considering
-at most 5 binding events for each peak, and we searched for the Fox A1
-motif over a 10 bp window around the estimated binding events. We used
-FIMO's command line 4.9.1 version \cite{fimo}.
-
-\subsection*{Imbalance index}
-
-For every ChIP-exo experiment, we partitioned the experiment into the
-non-digested regions using the QC pipeline. For each region we
-calculated the FSR defined as the ratio between the number of forward
-stranded reads and the total number of reads in a region. Then we
-transformed the FSR into an imbalance index is defined as:
-
-\[
-\mbox{Imbalance index} = -\log_{10} (4 \times \mbox{FSR} \times (1 - \mbox{FSR}))
-\]
-
-We used the imbalance index to asses if the regions considered by the
-QC pipeline are less imbalanced when they overlap peak regions. Since
-the strand imbalance distribution for both classes have the same
-median and differ in the tails, this transformation separates the
-tails from the median. Hence, when we reject the hypothesis of the
-distribution being the same for both classes, it means that the class
-of region that don't overlap peaks shows heavier tails than its
-complement. \footnote{not sure if I should add the FSR densities for
-  both classes in the supplement to prove this statement}
-
-% \subsection*{Evaluation of QC visualizations}
-
-% Each boxplot in Figure \ref{fig:enrich_all}A, was created by sampling
-% Msamp of the regions generated and fitting the model:
-
-% \begin{align*}
-%   \mbox{depth} = \beta \times \text{(\# unique
-%       positions)}
-% \end{align*}
-
-% and repeating the process Ntimes times.
-
-%     Intercept and C) Slope in the the $\mbox{URCR} = k / \mbox{ARC} +
-%     b$ model f
-
-\subsection*{Construction of a SE ChIP-Seq from a PE ChIP-Seq experiment.}
-
-For the rif-treatment ChIP-Seq experiments, we sampled SE ChIP-Seq
-experiment from the PE ones by taking one of both ends randomly
-following with equal probability.
-
-\subsection*{dPeak analysis of $\sig$ ChIP-exo and ChIP-Seq data.}
-
-For the resolution and sensitivity analysis, we used the MOSAiCS GC
-content + mappability model to call peaks for ChIP-exo experiments,
-while for both SE and PE ChIP-Seq experiments we used the MOSAiCS
-Input model. To avoid false positives we only considered ChIP-exo
-peaks with average ChIP counts greater than 3000 that overlapped both
-the SE and PE ChIP-Seq peaks, we considered other cutoff values but
-still obtained results similar to what we presented in this
-paper. Then we estimated the binding site events using the dPeak model
-with a maximum number of 5 binding events. We considered RegulonDB
-annotation as gold-standard; Resolution is defined as the minimum
-distance from an annotation to an estimated binding events and
-Sensitivity is defined as the fraction of annotation within
-15 bp from an estimated binding event.
-
-\subsection*{Saturation analysis of ChIP-exo, PE ChIP-Seq and SE ChIP-Seq.}
-
-To perform the saturation analysis, we sub-sampled $N$ fragments for
-both ChIP-exo and SE ChIP-Seq protocols. For PE ChIP-Seq we
-sub-sampled $N$ pairs or $N/2$ fragments. For each seed, we called
-peaks using MOSAiCS \cite{mosaics} (GC content + mappability for
-ChIP-exo and Input for SE and PE ChIP-Seq) for the maximum sample size
-and to avoid false positives we considered only the top 500 peaks for
-each data protocol. We defined the number of candidate regions as the
-number of top sample peaks such that a binding events was estimated
-using the sampled reads and the dPeak's model; the number of predicted
-events is the total quantity of binding events estimated using the
-dPeak's model; the number of identified targets are number of
-gold-standard annotations within 15 bp from an estimated binding
-events; and the resolution is defined as the minimum distance from a
-gold-standard annotation to an estimated binding event. We repeated
-this analysis for ten seeds and reported the median between all those
-values.
-
-\subsection*{Method comparison for ChIP-exo.}
-
-We considered dPeak Chung et al., 2013 \cite{dpeak}, GEM Guo et al.,
-2012 \cite{gem}, MACE Wang et al., 2014 \cite{mace} and Peazilla
-Bardet et al., 2013 \cite{peakzilla} for the ChIP-exo data
-analysis. for the dPeak algorithm we used the R package \textbf{dPeak}
-version 2.0.1 which is available from
+% Hence, when we reject the hypothesis of the
+% distribution being the same for both classes, it means that the class
+% of region that don't overlap peaks shows heavier tails than its
+% complement. \footnote{not sure if I should add the FSR densities for
+%   both classes in the supplement to prove this statement}
+
+% \subsection*{Construction of a SE ChIP-Seq from a PE ChIP-Seq experiment.}
+
+% For the rif-treatment ChIP-Seq experiments, we sampled SE ChIP-Seq
+% experiment from the PE ones by taking one of both ends randomly
+% following with equal probability.
+
+\subsection*{High resolution analysis with ChIP-exo}
+
+We considered RegulonDB \cite{regulondb} annotations as gold-standard
+and considered an annotation as being identified if the distance to a
+estimated binding events is less or equal than 20 bp. We
+defined the resolution as the distance from an annotation to its
+closest predictions and  the sensitivity as the fraction of identified
+annotations in a genomic region.
+
+\subsubsection*{Method comparison for ChIP-exo.}
+
+We compared dPeak, Chung et al., 2013; GEM, Guo et al., 2012; MACE,
+Wang et al., 2014 and Peazilla, Bardet et al., 2013 for the ChIP-exo
+data analysis. For the dPeak algorithm we used the R package
+\textbf{dPeak} version 2.0.1 which is available from
 \url{https://github.com/dongjunchung/dpeak}. For the GEM algorithm, we
@@
 \url{https://github.com/steinmann/peakzilla}. Candidate regions for
-\textbf{dPeak} were identified for each replicate of ChIP-exo data
-using the \textbf{MOSAiCS} algorithm Kuan et al., 2011 \cite{mosaics}
-(one sample analysis using false discovery rate of 0.01\%)
+\textbf{dPeak} and \textbf{GEM} were identified for each replicate of
+ChIP-exo data using the \textbf{MOSAiCS} algorithm \cite{mosaics} (one
+sample analysis using false discovery rate of 0.01\%)
 implemented as an R package \textbf{mosaics} version 2.9.7 (available
@@
 during model fitting for all methods. We downloaded \textbf{CexoR}
-\cite{cexor} version on its 1.8 version from \emph{bioconductor} but
-were unable to use it for the $\sig$ experiments.
+\cite{cexor} on its 1.8 version from \emph{bioconductor} but were
+unable to use it for the $\sig$ experiments.
+
+\subsubsection*{dPeak analysis of $\sig$ ChIP-exo and ChIP-Seq data.}
+
+We compared the estimated binding events predicted by the MOSAiCS +
+dPeak pipeline using reads generated by ChIP-exo, PE and SE ChIP-Seq
+protocols. We called peaks at a $5\%$ FDR level using MOSAiCS (the
+$\mbox{GC content} + \mbox{Mappability}$ model for ChIP-exo and the
+$\mbox{Input only}$ model for PE and SE ChIP-Seq). Then, we
+deconvolved the peak into binding events with dPeak (version 2.0.1) by
+considering a maximum of 5 binding sites on each peak. To avoid false
+positives we only considered ChIP-exo peaks with average ChIP counts
+greater than 3000 that overlapped both the SE and PE ChIP-Seq peaks,
+we considered other cutoff values but still obtained results similar
+to what we presented in this paper.
+
+\subsubsection*{Saturation analysis of ChIP-exo, PE ChIP-Seq and SE ChIP-Seq.}
+
+We sub-sampled $N$ fragments for both ChIP-exo and SE ChIP-Seq
+protocols. For PE ChIP-Seq we sub-sampled $N$ pairs or $N/2$
+fragments. For each seed, we called peaks using MOSAiCS \cite{mosaics}
+($\mbox{GC content} + \mbox{Mappability}$ for ChIP-exo and
+$\mbox{Input only}$ for SE and PE ChIP-Seq) for the maximum sample
+size and to avoid false positives we considered only the top 500 peaks
+for each data protocol. We defined the number of candidate regions as
+the number of top sample peaks such that a binding events was
+estimated using the sampled reads and the dPeak's model; the number of
+predicted events is the total quantity of binding events estimated
+using the dPeak's model; the number of identified targets are number
+of gold-standard annotations within 15 bp from an estimated
+binding events; and the resolution is defined as the minimum distance
+from a gold-standard annotation to an estimated binding event. We
+repeated this analysis for ten seeds and reported the median between
+all those values.
 

### 2016-05-02 rmarkdown/paper.Rnw
abstract corrected

@@
     resolution mapping of transcription factor binding sites. Although
-    many aspect of the ChIP-exo data analysis are similar to those of
+    many aspects of the ChIP-exo data analysis are similar to those of
     ChIP-Seq, ChIP-exo presents a number of unique challenges. We
-    present a quality control pipeline that analyzes a ChIP-exo and
-    ChIP-nexus experiment's strand imbalance, enrichment and library
-    complexity. Assessment of these characteristics are facilitated
+    present a quality control pipeline to evaluate a number of key
+    issues including strand imbalance, library complexity and
+    enrichment. Assessment of these characteristics are facilitated
     through diagnostic plots and summary statistics calculated over
@@
 
-    We systematically evaluated diverse aspects of ChIP-exo and found
-    the following characteristics: First, ChIP-exo's background is
-    quite different from ChIP-Seq's. Second, although often assumed in
-    ChIP-exo data analysis methods, the ``peak pair'' assumptions does
-    not hold locally in actual ChIP-exo data. Third, we for the first
-    time compared Paired End (PE) ChIP-Seq with ChIP-exo and found
-    that both protocols are comparable in resolutions and sensitivity
-    for closely located binding events, but as the distance between
-    binding events increases ChIP-exo shows higher sensitivity that PE
-    ChIP-Seq. Finally, at fixed sequencing depths, ChIP-exo provides
-    higher sensitivity, specificity and spatial resolution than PE
-    ChIP-Seq.
-
+    The pipeline explores and quantifies these aspects by partitioning
+    the experiment reads into a collection of regions, calculating a
+    series of summary statistics for each region, providing
+    visualizations and calculating measures to globally asses the
+    quality of a ChIP-exo experiment. We provide guidelines to
+    distinguish libraries with low complexity from deeply sequenced
+    experiments by the use of the QC pipeline. We demonstrate that the
+    ChIP-exo QC pipeline it is also applicable to ChIP-nexus data,
+    showing that those experiments present higher quality than
+    ChIP-exo experiments under similar conditions.
+
+    We compared ChIP-exo with Paired End (PE) and Single End (SE)
+    ChIP-Seq and found the following characteristics: First, although
+    often assumed in ChIP-exo data analysis methods, the ``peak pair''
+    assumptions does not hold locally in actual ChIP-exo data. Second,
+    we for the first time compared PE ChIP-Seq with ChIP-exo and found
+    that at fixed sequencing depths, ChIP-exo provides higher
+    sensitivity, specificity and spatial resolution than PE
+    ChIP-Seq. Finally, we show that ChIP-exo and PE ChIP-Seq are
+    comparable in sensitivity for closely located binding events, but
+    as the average distance between binding events increases, ChIP-exo
+    show higher sensitivity than PE ChIP-Seq.
+
+    % both protocols are comparable in resolutions and sensitivity for
+    % closely located binding events, but as the distance between
+    % binding events increases ChIP-exo shows higher sensitivity that
+    % PE ChIP-Seq. Finally, at fixed sequencing depths, ChIP-exo
+    % provides higher sensitivity, specificity and spatial resolution
+    % than PE ChIP-Seq.
   \end{abstract}
@@
 \newpage
+
+\color{blue}
 
@@
 \end{enumerate}
+
+\color{black}
 

### 2016-05-02 rmarkdown/paper.tex
abstract corrected

@@
 
+\color{black}
+
 \newpage

### 2016-05-03 rmarkdown/paper.Rnw
fixed the qc tables and finished background

@@
 \usepackage{verbatim}
+\usepackage{tabu}
+\usepackage{multirow}
 
@@
 
+\section*{Background}
+\label{sec:intro}
+
+ChIP-exo (Chromatin Immunoprecipitation followed by exonuclease
+digestion and next generation sequencing) Rhee and Pugh, 2011
+\nocite{exo1} is the state-of-the-art experiment developed to attain
+single base-pair resolution of protein binding site identification and
+it is considered as a powerful alternative to popularly used ChIP-Seq
+(Chromatin Immunoprecipitation coupled with next generation
+sequencing) assay. ChIP-exo experiments first capture millions of DNA
+fragments (150 - 250 bp in length) that the protein under study
+interacts with using random fragmentation of DNA and a
+protein-specific antibody. Then, exonuclease is introduced to trim the
+$5\prime$ end of each DNA fragment to a fixed distance from the bound
+protein compared to ChIP-Seq. This step is unique to ChIP-exo and
+could potentially provide significantly higher spatial resolution
+compared to ChIP-Seq. Finally, high throughput sequencing of a small
+region (36 to 100 bp) at the $5\prime$ end of each fragment generates
+millions of reads. Figure \ref{fig:chip_diagram} illustrates the
+differences between ChIP-exo, Single End (SE) ChIP-Seq and Paired End
+(PE) ChIP-Seq: The $5\prime$ ends of a ChIP-exo experiment are located
+more tightly around the binding proteins than in a ChIP-Seq
+experiment; in a PE ChIP-Seq experiment both ends are observed while
+in a SE ChIP-Seq experiment only the $5\prime$ end.
+
+ChIP-nexus (Chromatin Immunoprecipitation followed by exonuclease
+digestion, unique barcode, single ligation and next generation
+ligation) He et al., 2015 \nocite{chipnexus} is a modification to the
+ChIP-exo protocol, where both sequencing adaptors are ligated at the
+end of the ChIP fragments. Then, after exonuclease digestion, DNA
+self-circularization with circLigase, and restriction enzyme cutting
+between the two adaptors, the final library is ammplified. ChIP-nexus
+arises as an alternative to ChIP-exo as it provides the possibility of
+attaining higher resolution analysis and yields higher complexity
+libraries.
+
+% Refers to... First, DNA libraries... I think that these might be
+% too much details for the introduction. How about discussing them in
+% the results when we investigate the related
+
+While the number of ChIP-exo data keeps increasing, characteristics of
+ChIP-exo data are not fully investigated yet. First, DNA libraries
+generated by the ChIP-exo protocol are expected to be less complex than the
+libraries generated by ChIP-Seq (Mahony et al.,
+2015\nocite{exo_review}). Second, although there are roughly the same
+amount of reads in both strands, locally there may be more reads in
+one strand than in the other. Finally, most of current ChIP-Seq
+quality control (QC) guidelines (Landt et al., 2012\nocite{encode_qc})
+may not be applicable on ChIP-exo, while there are not established QC
+pipelines for ChIP-exo; previous ChIP-exo analyses used ChIP-Seq
+samples to compare the resolution between experiments
+\cite{exo1,exoillumina,exo2}. To address these challenges, we suggest
+a collection of quality control visualizations to interrogate these
+biases in a ChIP-exo experiment and globally asses the enrichment and
+library complexity of a ChIP-exo sample and a procedure to distinguish
+low complexity libraries from deeply sequenced experiments. This
+aspect is unique to ChIP-exo, since the exonuclease enzyme it is
+expected to digested the reads that are not bound to a transcription
+factor, therefore the number of bases where a ChIP-exo fragment could
+be potentially aligned is reduced. That way, in a high quality
+ChIP-exo experiment is possible to observe a large amount of reads to
+be aligned to unique position due to genomic signal instead of PCR
+artifacts.
+
+In order to obtain the potential benefits of ChIP-exo on protein
+binding site identification, it is critical to use algorithms that
+could fully utilize information available in ChIP-exo data. Rhee and
+Pugh, 2011 \nocite{exo1} discussed that reads in the forward and reverse
+strand might construct peak pairs around bound proteins, of which
+heights were implicitly assumed to be symmetric. Based on this
+rationale, they used the ``peak pair method'' that predicts the
+midpoint of two modes of peak pairs as potential binding
+sites. Recently developed ChIP-exo data analysis methods, such as Mace
+(Wang et al, 20114\nocite{mace}), CexoR (Madrigal, 2015\nocite{cexor})
+and Peakzilla (Bardet et al., 2013\nocite{peakzilla}), are also based
+on this peak pair assumption. However, appropriateness of such
+assumption was not fully evaluated in the literature yet. Furthermore,
+it is still unknown which factors could affect protein binding site
+identification using ChIP-exo data. In order to address this problem,
+we investigated various aspects of ChIP-exo data by contrasting them
+with their respective ChIP-Seq experiments.
+
+Currently, research on statistical methods for ChIP-exo data is still
+in its very early stage. Although many methods have been proposed to
+identify protein binding sites from ChIP-Seq data (reviewed by
+Wilbanks and Facciotti, 2012\nocite{evaluation} and Pepke and Wold,
+2009 \nocite{computation}), such as MACS (Zhang et al.,
+2008\nocite{macs}), CisGenome (Ji et al., 2008\nocite{cisgenome}) and
+MOSAiCS (Kuan et al., 2009\nocite{mosaics}), these approaches might
+not fully utilize potentials of ChIP-exo data for high resolution
+identification of protein binding sites. Specifically these approaches
+reveal protein binding sites only in lower resolution, i.e., at an
+interval of hundreds to thousands of base pairs. Furthermore, they
+implicitly assume that there is only one ``mode'' or ``predicted
+binding location'' per this wide genomic interval. More recently,
+deconvolution algorithms such as Deconvolution (Lun et al.,
+2009\cite{csdeconv}), GEM (Guo et al., 2012\nocite{gem}, an improved
+version of Guo et al., 2010\nocite{gps} ) and PICS (Zhang et al.,
+2010\nocite{pics}) have been proposed to identify binding sites in
+higher resolution using ChIP-Seq data. However, most of them are still
+not tailored for ChIP-exo and PE and SE ChIP-Seq data in a unified
+framework and as a result, currently available methods are not
+appropriate for fair comparison between ChIP-exo and ChIP-Seq. To
+address these limitations, we developed and utilized an improved
+version of dPeak (Chung et al., 2013\nocite{dpeak}), a high resolution
+binding site identification (deconvolution) algorithm that we
+previously developed for PE and SE ChIP-Seq data, so that it can also
+handle ChIP-exo data. The dPeak algorithm implements a probabilistic
+model that accurately describes the ChIP-exo and ChIP-Seq data
+generation process.
+
+Some of the key findings in this work are as follows. First, we
+demonstrate that the ``peak pair'' assumption of Rhee and Pugh,
+2013\nocite{exo2} does not hold well in real ChIP-exo data. Second, we
+found that when analyzing ChIP-exo data and the Input control is not
+available, it is useful to adjust for GC content and mappability
+biases to improve peak calling and binding site identification. Third,
+we evaluated several methods to identify binding events and dPeak
+performs competitively respect to GEM and MACE when analyzing ChIP-exo
+data. Finally, when comparable number of reads is used for both
+ChIP-exo and ChIP-Seq , dPeak coupled with ChIP-exo data provides
+resolution comparable to PE ChIP-Seq and both significantly improve
+the resolution of protein binding identification compared to SE-based
+analysis with any of the available methods.
+
+\section*{Results and discussion}
+\label{sec:results}
+% CTCF factor in human \cite{exo1}; ER factor in
+% human and FoxA1 factor in mouse (Serandour et al., 2013
+% \cite{exoillumina}); Glucocorticoid receptor (GR) in IMR90, K562 and
+% U2OS cell lines (Starick et al., 2015 \cite{starick15}); TATA-box
+% Binding Protein (TBP) in K562 cell lines (Venters and Pugh, 2013
+% \cite{venters13}). Furthermore, we also generated ChIP-exo and
+% ChIP-Seq data for $\sig$ factor in \emph{Escherichia Coli}
+% (\emph{E. Coli}) measured under aerobic ($ + O_2$) condition, and
+% treated by rifampicin by 0 and 20 minutes (courtesy of Professor
+% Robert Landick's lab). Additionally we gathered the ChIP-nexus from
+% TATA-box Binding Protein (TBP) in K562 cell lines; MyC and Max in S2
+% cell lines in \emph{Drosophila Melanogaster} (\emph{D. Melanogaster});
+% Twist and Dorsal in \emph{D. Melanogaster} embryo cell lines (He et
+% al., 2015 \cite{chipnexus}).
+\subsection*{Publicly available and novel datasets}
+
+We generated $\sig$ factor ChIP-exo, PE and SE ChIP-Seq experiments in
+\emph{E. Coli} under aerobic ($+O_2$) and anaerobic ($-O_2$)
+conditions in glucose minimal media on the HiSeq2000 and Illumina GA
+IIx platforms. Similarly, we generated $\sig$ factor ChIP-exo, PE and
+SE (generated \emph{in silico}) ChIP-Seq experiments in \emph{E. Coli}
+under aerobic ($+O_2$) conditions where rifampicin was applied for 20
+minutes and a control sample (without rifampicin being applied). We
+used these experimental designs for comparisons of ChIP-exo and PE
+ChIP-Seq assays of high resolution analysis and binding site
+identification. This comparison benefits from using $\sig$ for this
+study since is a transcription initiation factor of housekeeping genes
+in \emph{E. Coli}. In this organism's genomes, many promoters contain
+multiple transcription start sites (TSS) and these TSS are often
+closely spaced (10 $\sim$ 150 bp). These closely spaced binding sites
+are considered to be multiple ``switches'' that differentially
+regulate gene expression under diverse growth conditions
+\cite{regulondb}.
+
+Additionally, we gathered ChIP-exo data from diverse organisms: CTCF
+factor in human HeLa cell lines \cite{exo1}; ER factor in human MCF-7
+cell lines \cite{exoillumina}; GR factor in IMR90, K562 and U2OS human
+cell lines \cite{starick15}; TBP factor in human K562 cell lines
+\cite{venters13}. Additionally, we also gathered the ChIP-nexus
+experiments provided by He et al., 2014: TBP in human K562 cell lines,
+MyC and Max in S2 \emph{D. Melanogaster} cell lines and, Twist and
+Dorsal in \emph{D. Melanogaster} embryo cell lines.
+
+% Therefore, investigation of ChIP-exo's potential for
+% identification and differentiation of closely spaced binding sites is
+% invaluable for elucidating the transcriptional networks of prokaryotic
+% genomes.
+
 \color{blue}
@@
 \item Available data, the motivation of this part is two fold:
-  \begin{itemize}
-  \item we want to list all data sets we obtained
-  \item we want to emphasize that for the sig70 datasets we have a
-    unique design that allows us to compare chip exo with SE and PE
-    chipseq , the later one for the first time
-  \end{itemize}
-\item comparison with chip seq ? with and without dpeak ?
-\item only in methods. list the align algorithm used
+%  \begin{itemize}
+%  \item we want to list all data sets we obtained
+%  \item we want to emphasize that for the sig70 datasets we have a
+%    unique design that allows us to compare chip exo with SE and PE
+%    chipseq , the later one for the first time
+%  \end{itemize}
+\item comparison with chip seq %? with and without dpeak ?
+%\item only in methods. list the align algorithm used
 \item chip seq qc metrics
@@
 
-\section{Background}
-\label{sec:intro}
-
-ChIP-exo (Chromatin Immunoprecipitation followed by exonuclease
-digestion and next generation sequencing) Rhee and Pugh, 2011
-\cite{exo1} is the state-of-the-art experiment developed to attain
-single base-pair resolution of protein binding site identification and
-it is considered as a powerful alternative to popularly used ChIP-Seq
-(Chromatin Immunoprecipitation coupled with next generation
-sequencing) assay. ChIP-exo experiments first capture millions of DNA
-fragments (150 - 250 bp in length) that the protein under study
-interacts with using random fragmentation of DNA and a
-protein-specific antibody. Then, exonuclease is introduced to trim the
-$5\prime$ end of each DNA fragment to a fixed distance from the bound
-protein compared to ChIP-Seq. This step is unique to ChIP-exo and
-could potentially provide significantly higher spatial resolution
-compared to ChIP-Seq. Finally, high throughput sequencing of a small
-region (36 to 100 bp) at the $5\prime$ end of each fragment generates
-millions of reads. Figure \ref{fig:chip_diagram} illustrates the
-differences between ChIP-exo, Single End (SE) ChIP-Seq and Paired End
-(PE) ChIP-Seq: The $5\prime$ ends of a ChIP-exo experiment are located
-more tightly around the binding proteins than in a ChIP-Seq
-experiment; in a PE ChIP-Seq experiment both ends are observed while
-in a SE ChIP-Seq experiment only the $5\prime$ end.
-
-% Refers to... First, DNA libraries... I think that these might be
-% too much details for the introduction. How about discussing them in
-% the results when we investigate the related
-
-While the number of ChIP-exo data keeps increasing, characteristics of
-ChIP-exo data are not fully investigated yet. First, DNA libraries
-generated by the ChIP-exo protocol seem to be less complex than the
-libraries generated by ChIP-Seq (Mahony et al., 2015
-\cite{exo_review}). Second, although there are roughly the same amount
-of reads in both strands, locally there may be more reads in one
-strand than in the other. Finally, most of current ChIP-Seq quality
-control (QC) guidelines (Landt et al., \cite{encode_qc}) may not be
-applicable on ChIP-exo, while there are not established QC pipelines
-for ChIP-exo; previous ChIP-exo analyses used ChIP-Seq samples to
-compare the resolution between experiments (\cite{exo1},
-\cite{exoillumina}, \cite{exo2})). To address these challenges, we
-suggest a collection of quality control visualizations to interrogate
-these biases in a ChIP-exo experiment and globally asses the
-enrichment and library complexity of a ChIP-exo sample. We gathered
-ChIP-exo data from diverse organisms: CTCF factor in human
-\cite{exo1}; ER factor in human and FoxA1 factor in mouse (Serandour
-et al., 2013 \cite{exoillumina}); Glucocorticoid receptor (GR) in
-IMR90, K562 and U2OS cell lines (Starick et al., 2015
-\cite{starick15}); TATA-box Binding Protein (TBP) in K562 cell lines
-(Venters and Pugh, 2013 \cite{venters13}). Furthermore, we also
-generated ChIP-exo and ChIP-Seq data for $\sig$ factor in
-\emph{Escherichia Coli} (\emph{E. Coli}) measured under aerobic ($ +
-O_2$) condition, and treated by rifampicin by 0 and 20 minutes
-(courtesy of Professor Robert Landick's lab). Additionally we gathered
-the ChIP-nexus from TATA-box Binding Protein (TBP) in K562 cell lines;
-MyC and Max in S2 cell lines in \emph{Drosophila Melanogaster}
-(\emph{D. Melanogaster}); Twist and Dorsal in \emph{D. Melanogaster}
-embryo cell lines (He et al., 2015 \cite{chipnexus}).
-
-In order to obtain the potential benefits of ChIP-exo on protein
-binding site identification, it is critical to use algorithms that
-could fully utilize information available in ChIP-exo data. Rhee and
-Pugh, 2011 \cite{exo1} discussed that reads in the forward and reverse
-strand might construct peak pairs around bound proteins, of which
-heights were implicitly assumed to be symmetric. Based on this
-rationale, they used the ``peak pair method'' that predicts the
-midpoint of two modes of peak pairs as potential binding
-sites. Recently developed ChIP-exo data analysis methods, such as Mace
-(Wang et al, 20114 \cite{mace}), CexoR (Madrigal, 2015 \cite{cexor})
-and Peakzilla (Bardet et al., 2013 \cite{peakzilla}), are also based
-on this peak pair assumption. However, appropriateness of such
-assumption was not fully evaluated in the literature yet. Furthermore,
-it is still unknown which factors could affect protein binding site
-identification using ChIP-exo data. In order to address this problem,
-we investigated various aspects of ChIP-exo data by contrasting them
-with their respective ChIP-Seq experiments.
-
-Currently, research on statistical methods for ChIP-exo data is still
-in its very early stage. Although many methods have been proposed to
-identify protein binding sites from ChIP-Seq data (reviewed by
-Wilbanks and Facciotti, 2012 \cite{evaluation} and Pepke and Wold,
-2009 \cite{computation}), such as MACS (Zhang et al., 2008
-\cite{macs}), CisGenome (Ji et al., 2008 \cite{cisgenome}) and MOSAiCS
-(Kuan et al., 2009 \cite{mosaics}), these approaches might not fully
-utilize potentials of ChIP-exo data for high resolution identification
-of protein binding sites. Specifically these approaches reveal protein
-binding sites only in lower resolution, i.e., at an interval of
-hundreds to thousands of base pairs. Furthermore, they implicitly
-assume that there is only one ``mode'' or ``predicted binding
-location'' per this wide genomic interval. More recently,
-deconvolution algorithms such as Deconvolution (Lun et al., 2009
-\cite{csdeconv}), GEM (Guo et al., 2012 \cite{gem}, an improved
-version of Guo et al., 2010 \cite{gps} ) and PICS (Zhang et al., 2010
-\cite{pics}) have been proposed to identify binding sites in higher
-resolution using ChIP-Seq data. However, most of them are still not
-tailored for ChIP-exo and PE and SE ChIP-Seq data in a unified
-framework and as a result, currently available methods are not
-appropriate for fair comparison between ChIP-exo and ChIP-Seq. To
-address these limitations, we developed and utilized an improved
-version of dPeak (Chung et al., 2013 \cite{dpeak}), a high resolution
-binding site identification (deconvolution) algorithm that we
-previously developed for PE and SE ChIP-Seq data, so that it can also
-handle ChIP-exo data. The dPeak algorithm implements a probabilistic
-model that accurately describes the ChIP-exo and ChIP-Seq data
-generation process.
-
-Some of the key findings in this work are as follows. First, we
-demonstrate that the ``peak pair'' assumption of Rhee and Pugh, 2013
-\cite{exo2} does not hold well in real ChIP-exo data. Second, we found
-that when we analyze ChIP-exo data from eukaryotic genomes, it is
-important to consider sequence biases inherent to ChIP-exo data, such
-as mappability and GC content, in order to improve sensitivity and
-specificity of binding site identification. Third, we evaluated
-several methods to identify binding events and dPeak performs
-competitively respect to GEM and MACE when analyzing ChIP-exo
-data. Finally, when comparable number of reads is used for both
-ChIP-exo and ChIP-Seq , dPeak coupled with ChIP-exo data provides
-resolution comparable to PE ChIP-Seq and both significantly improve
-the resolution of protein binding identification compared to SE-based
-analysis with any of the available methods.
-
-\section{Results and discussion}
-\label{sec:results}
-
-\subsection{Deeply sequenced E. Coli $\sig$ ChIP-exo and ChIP-Seq data}
-
-$\sig$ factor is a transcription initiation factor of housekeeping
-genes in E. Coli. In this organism's genomes, many promoters contain
-multiple transcription start sites (TSS) and these TSS are often
-closely spaced (10 $\sim$ 150 bp). These closely spaced binding sites
-are considered to be multiple ``switches'' that differentially
-regulate gene expression under diverse growth conditions
-\cite{regulondb}. Therefore, investigation of ChIP-exo's potential for
-identification and differentiation of closely spaced binding sites is
-invaluable for elucidating the transcriptional networks of prokaryotic
-genomes.
-
-\subsection{Quality Control of ChIP-exo and ChIP-nexus data}
-
-\subsubsection{Application of current ChIP-Seq QC guidelines on ChIP-exo
-  data}
+\subsection*{Comparison with ChIP-Seq data}
+
+<<comp_param , include = FALSE,echo = FALSE,eval = TRUE>>=
+
+  bin_size <- 150
+  frag_len <- 150
+
+@
+
+We first compared various factors that could affect binding site
+identification between ChIP-exo and ChIP-Seq data. In order to compare
+distribution of signal and background between ChIP-exo and ChIP-Seq
+data, we calculated ChIP tag counts across the genome by counting the
+number of reads mapping to each of \Sexpr{bin_size} non-overlapping
+window after extending reads by \Sexpr{frag_len} to their $3\prime$
+end directions. ChIP tag counts in ChIP-exo data were linearly related
+to ChIP tag counts in ChIP-Seq data for the regions with high ChIP tag
+counts (Upper part of Figure \ref{fig:comp}A). This implies that
+signals for potential binding sites are well reproducible between
+ChIP-exo and ChIP-Seq data. On the other hand, there was clear
+difference in the background distribution between them (lower part of
+Figure \ref{fig:comp}A). Specifically, in ChIP-Seq data reads were
+almost uniformly distributed over background (non-binding) regions and
+the ChIP tag counts in there regions were significantly larger than
+zero. In contrast, in ChIP-exo data, there was larger variation in
+ChIP tag counts among background regions and ChIP tag counts were much
+lower in these regions compared to ChIP-Seq data. There were also
+large proportion of regions without any read in ChIP-exo data. These
+results indicate that for ChIP-exo data a much smaller portion of the
+genome is expected to be background.
+
+We next evaluated the ``peak pair'' assumption from Rhee and Pugh,
+2011 \cite{exo1}, i.e. a peak of reads in the forward strand is
+usually paired with a peak of reads in the reverse strand that is
+located in the other site of the binding site. Note that currently
+available ChIP-exo data analysis methods, such as Wang et al., 2014
+\cite{mace}, Madrigal 2015 \cite{cexor} and Bardet et al., 2013
+\cite{peakzilla} rely on this assumption. In order to evaluate this
+assumption, we reviewed the proportion of reads in the forward strand
+in ChIP-exo peaks such as at least one binding site is predicted in
+both ChIP-exo and ChIP-Seq data. We found that strands of reads were
+much less balanced in ChIP-exo data than in ChIP-Seq data in these
+regions with potential binding sites (Fig. \ref{fig:comp}B) and this
+indicates that the peak pair assumption might not hold in real
+ChIP-exo data.
+
+We evaluated ChIP-exo data for CTCF factor from human genome
+\cite{exo1} to investigate issues specific to eukaryotic genomes for
+binding sites identification. Figures \ref{fig:comp}C and
+\ref{fig:comp}D display the bin-level average read counts against
+mappability and GC content. Each data point is obtained by averaging
+the read counts across bins with the same mappability of GC -
+content. In Figure \ref{fig:comp}C it is shown that the ChIP-exo tag
+counts linearly increases with the mappability score and in Figure
+\ref{fig:comp}D it is shown that for GC - content below 0.6, the mean
+ChIP tag count increases and for GC - content greater than 0.6 it
+shows a decreasing trend. Kuan et al., 2011 \cite{mosaics} studied the
+presence of the mappability and GC - content biases in ChIP-Seq's
+background. It is not surprising to see these biases also present in
+ChIP-exo data, since ChIP-Seq and ChIP-exo signal seems to be linearly
+correlated for enriched regions (Figure \ref{fig:comp}A). Rozowsky et
+al., 2009 \cite{peakseq} and Valouev et al., 2008 \cite{quest} provide
+in depth analysis of the mappability and GC - content biases for
+ChIP-Seq respectively. Finally, these results indicate that binding
+site identification for ChIP-exo data benefits from using methods that
+take into account of apparent sequence biases such as mappability and
+GC content.
+
+\subsection*{Application of ENCODE quality metrics to ChIP-exo and
+  ChIP-nexus data}
 
 We started our exploration by investigating whether the current
-state-of-the-art QC pipelines for ChIP-Seq are suitable for
-ChIP-exo. Table \ref{tab:qc} contains measures that are commonly
-calculated for ChIP-Seq samples: PCR Bottleneck Coefficient (PBC) and
-the Normalized Strand Cross-Correlation (NSC) are calculated as in
-\cite{encode_qc}. We omitted the Relative Strand Cross-Correlation
-(RSC) which is another commonly used QC measure because a typical
-ChIP-exo experiment is not accompanied by an input sample.
+state-of-the-art QC pipelines for ChIP-Seq are suitable for ChIP-exo
+and ChIP-neuxs. In Tables \ref{tab:qc_sig} and \ref{tab:qc} we
+calculated a collection of the commonly used ChIP-Seq QC metrics using
+the ChIP-exo and ChIP-nexus experiments instead: Normalized Strand
+Cross-Correlation (NSC), Relative Strand Cross-Correlation and PCR
+Bottleneck Coefficient (PBC) defined as in 
+\url{https://genome.ucsc.edu/ENCODE/qualityMetrics.html}\nocite{encode_qc}.
 
@@
 
+\arrayrulewidth=.1mm
+
 \begin{table}[h!]
-%% need to fix format
-  \centering
-%% need to re-build this table
-<<table1,include = FALSE,echo = FALSE,eval = TRUE>>=
+  \centering
+  \begin{tabu} to\linewidth{X[-1]|X[-1]|X[-1]|X[-1]|X[-1]|X[-1]|X[-1]}
+    \firsthline
+    \textbf{Condition} & \textbf{Treatment} & \textbf{Rep.} & \textbf{Depth} & \textbf{NSC} & \textbf{RSC} & 
+    \textbf{PBC}\\
+    \hline 
+    Aerobic & No Rif. & 1 & 13,961,493 & 103.15 & 2.0193 & 0.1399 \\
+    Aerobic & No Rif. & 2 & 14,810,838 & 162.70 & 1.7805 & 0.1633 \\
+    Anaerobic & No Rif. & 1 &  16,108,774 & 153.51 & 1.8035 & 0.1353 \\
+    Anaerobic & No Rif. & 2 &  13,636,541 & 172.59 & 2.014 & 0.1532 \\ 
+    Aerobic & No Rif. & 1 & 902,921 & 13.77 & 1.1270 & 0.2689\\
+    Aerobic & No Rif. & 1 & 1,852,124 & 17.91 & 1.5275 & 0.2590\\
+    Aerobic & Rif. 20 min & 2 & 2,104,427 & 29.60 & 1.2844 & 0.2584\\
+    Aerobic & Rif. 20 min & 2 & 11,548,572 & 13.08 & 1.5122 & 0.1510 \\
+    \lasthline
+    \end{tabu}
+    \caption{Current QC metrics applied to generated \emph{E. Coli} $\sig$ samples. NSC stands for Normalized 
+             Strand Cross-Correlation, RSC stands for Relative Strand Cross-Correlation and PBC stands for 
+             PCR Bottleneck Coefficient.}
+\label{tab:qc_sig}
+\end{table}
+
+\begin{table}[h!]
+  \centering
+  \begin{tabu} to\linewidth{X[1.3]|X[2.2]|X[-1]|X[-1]|X[-1]|X[1.8]|X[-1]|X[-1]|X[-1]}
+    \firsthline
+    \textbf{Protocol} & \textbf{Organism} & \textbf{TF} & \textbf{Cell line} & \textbf{Rep.} &
+    \textbf{Depth} & \textbf{NSC} & \textbf{RSC} &   \textbf{PBC} \\ 
+    \hline
+    \multirow{13}{4em}{ChIP-exo} & Human & CTCF & HeLa & & 48,478,450 & 16.02 & 1.1960 & 0.4579 \\
+    & \multirow{3}{4em}{Human} & \multirow{3}{4em}{ER} & \multirow{3}{4em}{MCF-7} & 1 & 9,289,835 & 
+    19.87 & 1.0127 & 0.8082 \\
+    &  &  & & 2 & 11,041,833 & 21.48& 1.0063 & 0.8024\\
+    &  &  & & 3 & 12,464,836 & 18.72& 1.0100 & 0.8203 \\    
+    &  \multirow{3}{4em}{Mouse} & \multirow{3}{4em}{FoxA1} & \multirow{3}{4em}{Liver} & 1 & 22,210,461 & 21.28 & 1.1104 &  0.6562 \\
+    &  & & & 2 & 23,307,557 & 60.42 & 1.1604 & 0.7996 \\ 
+    &  & & & 3 & 22,421,72  & 72.04 & 1.1975 & 0.1068 \\
+    & \multirow{3}{4em}{Human} & \multirow{3}{4em}{GR} & IM90 & & 47,443,803 & 8.86  & 1.3678 & 0.2978 \\
+    & &  & K562 & & 116,518,000   & 4.11 & 1.0441 &0.0504 \\
+    & &  & U2OS & & 3,255,111 &  10.05 & 1.0288 & 0.7714 \\
+    & \multirow{3}{4em}{Human} & \multirow{3}{4em}{TBP} & \multirow{3}{4em}{K562} &1 & 61,046,382 &  12.01  & 1.1119
+ & 0.1232 \\
+    & & & &2 & 94,314,770 & 7.93 & 1.0299 & 0.1681 \\
+    & & & & 3 & 114,282,270 & 9.25 & 1.1027 & 0.1464\\
+\hline
+\multirow{10}{4em}{ChIP-nexus} & \multirow{8}{4em}{D.Melanogaster} &
+       \multirow{2}{4em}{Dorsal} & \multirow{4}{4em}{embryo} & 1 &   8,863,170 &  7.27 & 1.0402 &  0.6766 \\
+ & & &  & 2 & 10,003,562 & 7.19 & 1.0672 & 0.5656\\
+ & &  \multirow{2}{4em}{Twist} & & 1 & 18,244,203 & 5.82 &  1.1637 &  0.6592 \\
+ & & &  & 2 & 52,546,982 & 5.27  &  1.1805  & 0.4549 \\
+ & & \multirow{2}{4em}{Max} & \multirow{4}{4em}{S2} & 1 & 18,320,743 & 3.60 & 1.3628 & 0.5178  \\
+ & & & & 2 & 24,965,642  & 3.47 & 1.0138 & 0.2124\\
+ & & \multirow{2}{4em}{MyC} & & 1 & 7,832,034 & 5.92 & 1.0115 &  0.3935  \\
+ & & & &  2 & 22,824,467 & 5.76 & 1.0045 & 0.1879\\
+ & \multirow{2}{4em}{Human} & \multirow{2}{4em}{TBP} & \multirow{2}{4em}{K562} & 1 &  
+              33,708,245 & 32.16 & 1.1712 & 0.3102 \\
+ & &  &  & 2 & 129,675,001 & 32.70 &  1.2455 & 0.0492 \\
+    \lasthline
+  \end{tabu}
+  \caption{Current QC metrics applied to gathered data. NSC stands for Normalized 
+    Strand Cross-Correlation, RSC stands for Relative Strand Cross-Correlation and PBC stands for 
+    PCR Bottleneck Coefficient.}  
+\label{tab:qc}
+\end{table}
+
+<<table1,include = FALSE,echo = FALSE,eval = FALSE>>=
   library(data.table)
@@
 
-<<table2,include = TRUE,echo = FALSE, eval = TRUE >>= 
-  A
-@ 
-\caption{\textbf{Current QC metrics applied to ChIP-exo data.} PBC
-  stands for PCR Bottleneck Coefficient and NSC for Normalized Strand
-  Cross-Correlation.}
-  \label{tab:qc} %% need to make this table
-\end{table}
-
-DNA libraries generated by the ChIP-exo protocol seem to be less
-complex than the libraries generated by ChIP-Seq, since the possible
-number of positions to which the reads can be aligned is being reduced
-due to the exonuclease digestion, hence considerable amounts of reads
-are being mapped to specific positions. This affects the
-interpretation of the PBC, since for ChIP-Seq low PBC values indicate
-that the same read has been copied by the amplification process and
-aligned multiple times to the same position; while for ChIP-exo when
-several reads are aligned to the same position are not necessarily the
-same read amplified, but several reads that their $5\prime$ end was
-digested to the same position before the amplification step. It is of
-special importance to notice that for several ChIP-exo datasets, the
-PBC values are quite low. Therefore, by blindly following ChIP-Seq
-guidelines those experiment could have been considered as not useful
-and repeated.
+DNA libraries generated by ChIP-exo and ChIP-nexus protocols are
+expected to be less complex than the libraries generated by ChIP-Seq,
+since the possible number of positions to which the reads can be
+aligned is being reduced due to the exonuclease digestion,
+considerable amounts of reads are being mapped to specific
+positions. This affects the interpretation of the PBC, since for
+ChIP-Seq low PBC values indicate that the same read has been copied by
+the amplification process and aligned multiple times to the same
+position; while for ChIP-exo when several reads are aligned to the
+same position are not necessarily the same read amplified, but several
+reads that their $5\prime$ end was digested to the same position
+before the amplification step. It is of special importance to notice
+that for deeply sequenced ChIP-exo and ChIP-nexus experiments, the PBC
+values are quite low, which by following the ChIP-Seq guidelines it
+would indicate that those experiment show severe bottlenecking
+problems, which may incorrectly suggest that there is not genomic
+signal int those experiments.
 
@@
 to calculate the NSC values in Table \ref{tab:qc}.
-
-\subsubsection{Comparison with ChIP-Seq data}
-
-<<comp_param , include = FALSE,echo = FALSE,eval = TRUE>>=
-
-  bin_size <- 150
-  frag_len <- 150
-
-@
-
-We first compared various factors that could affect binding site
-identification between ChIP-exo and ChIP-Seq data. In order to compare
-distribution of signal and background between ChIP-exo and ChIP-Seq
-data, we calculated ChIP tag counts across the genome by counting the
-number of reads mapping to each of \Sexpr{bin_size} non-overlapping
-window after extending reads by \Sexpr{frag_len} to their $3\prime$
-end directions. ChIP tag counts in ChIP-exo data were linearly related
-to ChIP tag counts in ChIP-Seq data for the regions with high ChIP tag
-counts (Upper part of Figure \ref{fig:comp}A). This implies that
-signals for potential binding sites are well reproducible between
-ChIP-exo and ChIP-Seq data. On the other hand, there was clear
-difference in the background distribution between them (lower part of
-Figure \ref{fig:comp}A). Specifically, in ChIP-Seq data reads were
-almost uniformly distributed over background (non-binding) regions and
-the ChIP tag counts in there regions were significantly larger than
-zero. In contrast, in ChIP-exo data, there was larger variation in
-ChIP tag counts among background regions and ChIP tag counts were much
-lower in these regions compared to ChIP-Seq data. There were also
-large proportion of regions without any read in ChIP-exo data. These
-results indicate that for ChIP-exo data a much smaller portion of the
-genome is expected to be background.
-
-We next evaluated the ``peak pair'' assumption from Rhee and Pugh,
-2011 \cite{exo1}, i.e. a peak of reads in the forward strand is
-usually paired with a peak of reads in the reverse strand that is
-located in the other site of the binding site. Note that currently
-available ChIP-exo data analysis methods, such as Wang et al., 2014
-\cite{mace}, Madrigal 2015 \cite{cexor} and Bardet et al., 2013
-\cite{peakzilla} rely on this assumption. In order to evaluate this
-assumption, we reviewed the proportion of reads in the forward strand
-in ChIP-exo peaks such as at least one binding site is predicted in
-both ChIP-exo and ChIP-Seq data. We found that strands of reads were
-much less balanced in ChIP-exo data than in ChIP-Seq data in these
-regions with potential binding sites (Fig. \ref{fig:comp}B) and this
-indicates that the peak pair assumption might not hold in real
-ChIP-exo data.
-
-We evaluated ChIP-exo data for CTCF factor from human genome
-\cite{exo1} to investigate issues specific to eukaryotic genomes for
-binding sites identification. Figures \ref{fig:comp}C and
-\ref{fig:comp}D display the bin-level average read counts against
-mappability and GC content. Each data point is obtained by averaging
-the read counts across bins with the same mappability of GC -
-content. In Figure \ref{fig:comp}C it is shown that the ChIP-exo tag
-counts linearly increases with the mappability score and in Figure
-\ref{fig:comp}D it is shown that for GC - content below 0.6, the mean
-ChIP tag count increases and for GC - content greater than 0.6 it
-shows a decreasing trend. Kuan et al., 2011 \cite{mosaics} studied the
-presence of the mappability and GC - content biases in ChIP-Seq's
-background. It is not surprising to see these biases also present in
-ChIP-exo data, since ChIP-Seq and ChIP-exo signal seems to be linearly
-correlated for enriched regions (Figure \ref{fig:comp}A). Rozowsky et
-al., 2009 \cite{peakseq} and Valouev et al., 2008 \cite{quest} provide
-in depth analysis of the mappability and GC - content biases for
-ChIP-Seq respectively. Finally, these results indicate that binding
-site identification for ChIP-exo data benefits from using methods that
-take into account of apparent sequence biases such as mappability and
-GC content.
 
@@
 \textbf{Analysis of ChIP-nexus data with the QC pipeline}
-
-ChIP-nexus developed by He et al., 2015 \cite{chipnexus} is a
-modification to the ChIP-exo protocol, where both sequencing adaptors
-are ligated at the $3\prime$ end of the ChIP fragments. Then, after
-exonuclease digestion, DNA self-circularization with circLigase, and
-restriction enzyme cutting between the two adaptors, the final library
-is amplified. In Table \ref{tab:nexus} we show the usual ChIP-Seq QC
-metrics to ChIP-nexus data and found that the conclusions made for
-ChIP-exo data still holds, the NSC is high but when plotting the SCC
-the fragment and read length peaks are confounded and the PBC may
-indicate severe bottlenecking but it is caused by the exonuclease
-enzyme digestion which reduces the number of positions where the
-reads can be aligned.
-
-\begin{table}[h!]
-  \centering
-<<table3,include = TRUE,echo = FALSE, eval = TRUE >>= 
-  B
-@ 
-\caption{\textbf{Current QC metrics applied to ChIP-nexus data.} PBC
-  stands for PCR Bottleneck Coefficient and NSC for Normalized Strand
-  Cross-Correlation.}
-  \label{tab:nexus}
-\end{table}
 
@@
 
-\section{Conclusions}
+\section*{Conclusions}
 \label{sec:conc}
@@
 
-\section{Methods}
+\section*{Methods}
 \label{sec:methods}
@@
 where $y_{\text{bgd}}$ is estimated as the background SCC level is
-defined as $y_{\text{bgd}} = \min_\delta y(\delta)$ and $y_{\text{rl}}$
-  is the value of the SCC curve when the shift equals the experiment's
-  read length.
+defined as $y_{\text{bgd}} = \min_\delta y(\delta)$ and
+$y_{\text{rl}}$ is the value of the SCC curve when the shift equals
+the experiment's read length. It is worth noticing that in Landt et
+al., 2012 the RSC is defined as:
+
+\begin{align*}
+  \mbox{RSC} &= \frac{\max_\delta y(\delta)}{y_{\text{rl}}}.  
+\end{align*}
+
+In ChIP-exo data, both definitions are equivalent since in a typical
+ChIP-exo experiment $y_{\text{bgd}}$ is approximately zero.
 
@@
 
-\begin{table}[h!]
-%% need to fix format
-  \centering
-<<table_chipseq1,include = TRUE,echo = FALSE,eval = TRUE>>=
+%% \begin{table}[h!]
+%%  %% need to fix format
+%%   \centering
+<<table_chipseq1,include = FALSE,echo = FALSE,eval = FALSE>>=
   load("../data/paper_tables/Landick_rif_chipseq_summary.RData")
@@
 @ 
-\caption{Usual ChIP-Seq QC metrics as in table \ref{tab:qc} but
-  applied to Landick's chipseq data of the rif experiment}
-\end{table}
+% \caption{Usual ChIP-Seq QC metrics as in table \ref{tab:qc} but
+%   applied to Landick's chipseq data of the rif experiment}
+% \end{table}
 

### 2016-05-03 rmarkdown/paper.tex
fixed the qc tables and finished background

@@
 
-\section{Background}
-\label{sec:intro}
-
-ChIP-exo (Chromatin Immunoprecipitation followed by exonuclease
-digestion and next generation sequencing) Rhee and Pugh, 2011
-\cite{exo1} is the state-of-the-art experiment developed to attain
-single base-pair resolution of protein binding site identification and
-it is considered as a powerful alternative to popularly used ChIP-Seq
-(Chromatin Immunoprecipitation coupled with next generation
-sequencing) assay. ChIP-exo experiments first capture millions of DNA
-fragments (150 - 250 bp in length) that the protein under study
-interacts with using random fragmentation of DNA and a
-protein-specific antibody. Then, exonuclease is introduced to trim the
-$5\prime$ end of each DNA fragment to a fixed distance from the bound
-protein compared to ChIP-Seq. This step is unique to ChIP-exo and
-could potentially provide significantly higher spatial resolution
-compared to ChIP-Seq. Finally, high throughput sequencing of a small
-region (36 to 100 bp) at the $5\prime$ end of each fragment generates
-millions of reads. Figure \ref{fig:chip_diagram} illustrates the
-differences between ChIP-exo, Single End (SE) ChIP-Seq and Paired End
-(PE) ChIP-Seq: The $5\prime$ ends of a ChIP-exo experiment are located
-more tightly around the binding proteins than in a ChIP-Seq
-experiment; in a PE ChIP-Seq experiment both ends are observed while
-in a SE ChIP-Seq experiment only the $5\prime$ end.
-
-% Refers to... First, DNA libraries... I think that these might be
-% too much details for the introduction. How about discussing them in
-% the results when we investigate the related
-
-While the number of ChIP-exo data keeps increasing, characteristics of
-ChIP-exo data are not fully investigated yet. First, DNA libraries
-generated by the ChIP-exo protocol seem to be less complex than the
-libraries generated by ChIP-Seq (Mahony et al., 2015
-\cite{exo_review}). Second, although there are roughly the same amount
-of reads in both strands, locally there may be more reads in one
-strand than in the other. Finally, most of current ChIP-Seq quality
-control (QC) guidelines (Landt et al., \cite{encode_qc}) may not be
-applicable on ChIP-exo, while there are not established QC pipelines
-for ChIP-exo; previous ChIP-exo analyses used ChIP-Seq samples to
-compare the resolution between experiments (\cite{exo1},
-\cite{exoillumina}, \cite{exo2})). To address these challenges, we
-suggest a collection of quality control visualizations to interrogate
-these biases in a ChIP-exo experiment and globally asses the
-enrichment and library complexity of a ChIP-exo sample. We gathered
-ChIP-exo data from diverse organisms: CTCF factor in human
-\cite{exo1}; ER factor in human and FoxA1 factor in mouse (Serandour
-et al., 2013 \cite{exoillumina}); Glucocorticoid receptor (GR) in
-IMR90, K562 and U2OS cell lines (Starick et al., 2015
-\cite{starick15}); TATA-box Binding Protein (TBP) in K562 cell lines
-(Venters and Pugh, 2013 \cite{venters13}). Furthermore, we also
-generated ChIP-exo and ChIP-Seq data for $\sig$ factor in
-\emph{Escherichia Coli} (\emph{E. Coli}) measured under aerobic ($ +
-O_2$) condition, and treated by rifampicin by 0 and 20 minutes
-(courtesy of Professor Robert Landick's lab). Additionally we gathered
-the ChIP-nexus from TATA-box Binding Protein (TBP) in K562 cell lines;
-MyC and Max in S2 cell lines in \emph{Drosophila Melanogaster}
-(\emph{D. Melanogaster}); Twist and Dorsal in \emph{D. Melanogaster}
-embryo cell lines (He et al., 2015 \cite{chipnexus}).
-
-In order to obtain the potential benefits of ChIP-exo on protein
-binding site identification, it is critical to use algorithms that
-could fully utilize information available in ChIP-exo data. Rhee and
-Pugh, 2011 \cite{exo1} discussed that reads in the forward and reverse
-strand might construct peak pairs around bound proteins, of which
-heights were implicitly assumed to be symmetric. Based on this
-rationale, they used the ``peak pair method'' that predicts the
-midpoint of two modes of peak pairs as potential binding
-sites. Recently developed ChIP-exo data analysis methods, such as Mace
-(Wang et al, 20114 \cite{mace}), CexoR (Madrigal, 2015 \cite{cexor})
-and Peakzilla (Bardet et al., 2013 \cite{peakzilla}), are also based
-on this peak pair assumption. However, appropriateness of such
-assumption was not fully evaluated in the literature yet. Furthermore,
-it is still unknown which factors could affect protein binding site
-identification using ChIP-exo data. In order to address this problem,
-we investigated various aspects of ChIP-exo data by contrasting them
-with their respective ChIP-Seq experiments.
-
-Currently, research on statistical methods for ChIP-exo data is still
-in its very early stage. Although many methods have been proposed to
-identify protein binding sites from ChIP-Seq data (reviewed by
-Wilbanks and Facciotti, 2012 \cite{evaluation} and Pepke and Wold,
-2009 \cite{computation}), such as MACS (Zhang et al., 2008
-\cite{macs}), CisGenome (Ji et al., 2008 \cite{cisgenome}) and MOSAiCS
-(Kuan et al., 2009 \cite{mosaics}), these approaches might not fully
-utilize potentials of ChIP-exo data for high resolution identification
-of protein binding sites. Specifically these approaches reveal protein
-binding sites only in lower resolution, i.e., at an interval of
-hundreds to thousands of base pairs. Furthermore, they implicitly
-assume that there is only one ``mode'' or ``predicted binding
-location'' per this wide genomic interval. More recently,
-deconvolution algorithms such as Deconvolution (Lun et al., 2009
-\cite{csdeconv}), GEM (Guo et al., 2012 \cite{gem}, an improved
-version of Guo et al., 2010 \cite{gps} ) and PICS (Zhang et al., 2010
-\cite{pics}) have been proposed to identify binding sites in higher
-resolution using ChIP-Seq data. However, most of them are still not
-tailored for ChIP-exo and PE and SE ChIP-Seq data in a unified
-framework and as a result, currently available methods are not
-appropriate for fair comparison between ChIP-exo and ChIP-Seq. To
-address these limitations, we developed and utilized an improved
-version of dPeak (Chung et al., 2013 \cite{dpeak}), a high resolution
-binding site identification (deconvolution) algorithm that we
-previously developed for PE and SE ChIP-Seq data, so that it can also
-handle ChIP-exo data. The dPeak algorithm implements a probabilistic
-model that accurately describes the ChIP-exo and ChIP-Seq data
-generation process.
-
-Some of the key findings in this work are as follows. First, we
-demonstrate that the ``peak pair'' assumption of Rhee and Pugh, 2013
-\cite{exo2} does not hold well in real ChIP-exo data. Second, we found
-that when we analyze ChIP-exo data from eukaryotic genomes, it is
-important to consider sequence biases inherent to ChIP-exo data, such
-as mappability and GC content, in order to improve sensitivity and
-specificity of binding site identification. Third, we evaluated
-several methods to identify binding events and dPeak performs
-competitively respect to GEM and MACE when analyzing ChIP-exo
-data. Finally, when comparable number of reads is used for both
-ChIP-exo and ChIP-Seq , dPeak coupled with ChIP-exo data provides
-resolution comparable to PE ChIP-Seq and both significantly improve
-the resolution of protein binding identification compared to SE-based
-analysis with any of the available methods.
-
-\section{Results and discussion}
-\label{sec:results}
-
-\subsection{Deeply sequenced E. Coli $\sig$ ChIP-exo and ChIP-Seq data}
-
-$\sig$ factor is a transcription initiation factor of housekeeping
-genes in E. Coli. In this organism's genomes, many promoters contain
-multiple transcription start sites (TSS) and these TSS are often
-closely spaced (10 $\sim$ 150 bp). These closely spaced binding sites
-are considered to be multiple ``switches'' that differentially
-regulate gene expression under diverse growth conditions
-\cite{regulondb}. Therefore, investigation of ChIP-exo's potential for
-identification and differentiation of closely spaced binding sites is
-invaluable for elucidating the transcriptional networks of prokaryotic
-genomes.
-
-\subsection{Quality Control of ChIP-exo and ChIP-nexus data}
-
-\subsubsection{Application of current ChIP-Seq QC guidelines on ChIP-exo
-  data}
-
-We started our exploration by investigating whether the current
-state-of-the-art QC pipelines for ChIP-Seq are suitable for
-ChIP-exo. Table \ref{tab:qc} contains measures that are commonly
-calculated for ChIP-Seq samples: PCR Bottleneck Coefficient (PBC) and
-the Normalized Strand Cross-Correlation (NSC) are calculated as in
-\cite{encode_qc}. We omitted the Relative Strand Cross-Correlation
-(RSC) which is another commonly used QC measure because a typical
-ChIP-exo experiment is not accompanied by an input sample.
-
-%% We can observe that the PBC is low, hence it would be incorrectly
-%% suggested to repeat the experiment.
-
-\begin{table}[h!]
-%% need to fix format
-  \centering
-%% need to re-build this table
-
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-##     Organism  IP/TF Condition/Cell Rep.       Depth    PBC      NSC
-##  1:   E.Coli  Sig70        Aerobic    1  13,961,493 0.1399 103.4239
-##  2:   E.Coli  Sig70        Aerobic    2  14,810,838 0.1634 162.8002
-##  3:   E.Coli  Sig70      Anaerobic    1  16,108,774 0.1354 153.5088
-##  4:   E.Coli  Sig70      Anaerobic    2  13,636,541 0.1532 172.5815
-##  5:   E.Coli  Sig70       Rif-0min    1     902,921 0.2690  13.7691
-##  6:   E.Coli  Sig70       Rif-0min    1   1,852,124 0.2591  17.9188
-##  7:   E.Coli  Sig70      Rif-20min    2   2,104,427 0.2584  29.6083
-##  8:   E.Coli  Sig70      Rif-20min    2  11,548,572 0.1511  13.0863
-##  9:    Human   CTCF           HeLA    1  48,478,450 0.4580  16.0248
-## 10:    Mouse Fox A1    Mouse Liver    1  22,210,461 0.6562  21.2820
-## 11:    Mouse Fox A1    Mouse Liver    2  23,307,557 0.7996  60.4219
-## 12:    Mouse Fox A1    Mouse Liver    3  22,421,729 0.1068  72.0424
-## 13:    Human     GR          IMR90    1  47,443,803 0.2979   8.8602
-## 14:    Human     GR           K562    1 116,518,000 0.0505   4.1179
-## 15:    Human     GR           U20S    1   3,255,111 0.7714  10.0588
-## 16:    Human     ER          MCF-7    1   9,289,835 0.8083  19.8752
-## 17:    Human     ER          MCF-7    2  11,041,833 0.8024  21.4851
-## 18:    Human     ER          MCF-7    3  12,464,836 0.8204  18.7281
-## 19:    Human    TBP           K562    1  61,046,382 0.1232  12.0117
-## 20:    Human    TBP           K562    2  94,314,770 0.1682   7.9398
-## 21:    Human    TBP           K562    3 114,282,270 0.1464   9.2566
-##     Organism  IP/TF Condition/Cell Rep.       Depth    PBC      NSC
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
-\caption{\textbf{Current QC metrics applied to ChIP-exo data.} PBC
-  stands for PCR Bottleneck Coefficient and NSC for Normalized Strand
-  Cross-Correlation.}
-  \label{tab:qc} %% need to make this table
-\end{table}
-
-DNA libraries generated by the ChIP-exo protocol seem to be less
-complex than the libraries generated by ChIP-Seq, since the possible
-number of positions to which the reads can be aligned is being reduced
-due to the exonuclease digestion, hence considerable amounts of reads
-are being mapped to specific positions. This affects the
-interpretation of the PBC, since for ChIP-Seq low PBC values indicate
-that the same read has been copied by the amplification process and
-aligned multiple times to the same position; while for ChIP-exo when
-several reads are aligned to the same position are not necessarily the
-same read amplified, but several reads that their $5\prime$ end was
-digested to the same position before the amplification step. It is of
-special importance to notice that for several ChIP-exo datasets, the
-PBC values are quite low. Therefore, by blindly following ChIP-Seq
-guidelines those experiment could have been considered as not useful
-and repeated.
-
-The Strand Cross-Correlation (SCC) introduced by Kharchenko et al.,
-2008 \cite{strandcc} is the most commonly used quality measure in
-ChIP-Seq. It is calculated as the correlation between both strand
-coverages, where each one is shifted $\delta / 2$ bp towards the
-$3\prime$ direction. In general it measures how well the reads mapped
-to each strand are clustered around the locations where the proteins
-are binding to the DNA, and usually it is expected to observed two
-local maxima, one when the profiles are shifted by the average read
-length and another when the profiles are shifted by the unobserved
-fragment length. In a good ChIP-Seq dataset the last one is also the
-SCC global maxima. However, in ChIP-exo's case these two peaks are
-confounded. Hence the Normalized Strand Cross-Correlation (NSC) which
-is a measure based on the SCC is harder to interpret. Figure
-\ref{fig:scc_exo} shows that both local maxima are hard to
-differentiate in the SCC curves for the $\sig$ ChIP-exo datasets used
-to calculate the NSC values in Table \ref{tab:qc}.
-
-\subsubsection{Comparison with ChIP-Seq data}
+\subsection*{Comparison with ChIP-Seq data}
 
@@
 GC content.
+
+\subsection*{Application of ENCODE quality metrics to ChIP-exo and
+  ChIP-nexus data}
+
+We started our exploration by investigating whether the current
+state-of-the-art QC pipelines for ChIP-Seq are suitable for ChIP-exo
+and ChIP-neuxs. In Tables \ref{tab:qc_sig} and \ref{tab:qc} we
+calculated a collection of the commonly used ChIP-Seq QC metrics using
+the ChIP-exo and ChIP-nexus experiments instead: Normalized Strand
+Cross-Correlation (NSC), Relative Strand Cross-Correlation and PCR
+Bottleneck Coefficient (PBC) defined as in 
+\url{https://genome.ucsc.edu/ENCODE/qualityMetrics.html}\nocite{encode_qc}.
+
+%% We can observe that the PBC is low, hence it would be incorrectly
+%% suggested to repeat the experiment.
+
+\arrayrulewidth=.1mm
+
+\begin{table}[h!]
+  \centering
+  \begin{tabu} to\linewidth{X[-1]|X[-1]|X[-1]|X[-1]|X[-1]|X[-1]|X[-1]}
+    \firsthline
+    \textbf{Condition} & \textbf{Treatment} & \textbf{Rep.} & \textbf{Depth} & \textbf{NSC} & \textbf{RSC} & 
+    \textbf{PBC}\\
+    \hline 
+    Aerobic & No Rif. & 1 & 13,961,493 & 103.15 & 2.0193 & 0.1399 \\
+    Aerobic & No Rif. & 2 & 14,810,838 & 162.70 & 1.7805 & 0.1633 \\
+    Anaerobic & No Rif. & 1 &  16,108,774 & 153.51 & 1.8035 & 0.1353 \\
+    Anaerobic & No Rif. & 2 &  13,636,541 & 172.59 & 2.014 & 0.1532 \\ 
+    Aerobic & No Rif. & 1 & 902,921 & 13.77 & 1.1270 & 0.2689\\
+    Aerobic & No Rif. & 1 & 1,852,124 & 17.91 & 1.5275 & 0.2590\\
+    Aerobic & Rif. 20 min & 2 & 2,104,427 & 29.60 & 1.2844 & 0.2584\\
+    Aerobic & Rif. 20 min & 2 & 11,548,572 & 13.08 & 1.5122 & 0.1510 \\
+    \lasthline
+    \end{tabu}
+    \caption{Current QC metrics applied to generated \emph{E. Coli} $\sig$ samples. NSC stands for Normalized 
+             Strand Cross-Correlation, RSC stands for Relative Strand Cross-Correlation and PBC stands for 
+             PCR Bottleneck Coefficient.}
+\label{tab:qc_sig}
+\end{table}
+
+\begin{table}[h!]
+  \centering
+  \begin{tabu} to\linewidth{X[1.3]|X[2.2]|X[-1]|X[-1]|X[-1]|X[1.8]|X[-1]|X[-1]|X[-1]}
+    \firsthline
+    \textbf{Protocol} & \textbf{Organism} & \textbf{TF} & \textbf{Cell line} & \textbf{Rep.} &
+    \textbf{Depth} & \textbf{NSC} & \textbf{RSC} &   \textbf{PBC} \\ 
+    \hline
+    \multirow{13}{4em}{ChIP-exo} & Human & CTCF & HeLa & & 48,478,450 & 16.02 & 1.1960 & 0.4579 \\
+    & \multirow{3}{4em}{Human} & \multirow{3}{4em}{ER} & \multirow{3}{4em}{MCF-7} & 1 & 9,289,835 & 
+    19.87 & 1.0127 & 0.8082 \\
+    &  &  & & 2 & 11,041,833 & 21.48& 1.0063 & 0.8024\\
+    &  &  & & 3 & 12,464,836 & 18.72& 1.0100 & 0.8203 \\    
+    &  \multirow{3}{4em}{Mouse} & \multirow{3}{4em}{FoxA1} & \multirow{3}{4em}{Liver} & 1 & 22,210,461 & 21.28 & 1.1104 &  0.6562 \\
+    &  & & & 2 & 23,307,557 & 60.42 & 1.1604 & 0.7996 \\ 
+    &  & & & 3 & 22,421,72  & 72.04 & 1.1975 & 0.1068 \\
+    & \multirow{3}{4em}{Human} & \multirow{3}{4em}{GR} & IM90 & & 47,443,803 & 8.86  & 1.3678 & 0.2978 \\
+    & &  & K562 & & 116,518,000   & 4.11 & 1.0441 &0.0504 \\
+    & &  & U2OS & & 3,255,111 &  10.05 & 1.0288 & 0.7714 \\
+    & \multirow{3}{4em}{Human} & \multirow{3}{4em}{TBP} & \multirow{3}{4em}{K562} &1 & 61,046,382 &  12.01  & 1.1119
+ & 0.1232 \\
+    & & & &2 & 94,314,770 & 7.93 & 1.0299 & 0.1681 \\
+    & & & & 3 & 114,282,270 & 9.25 & 1.1027 & 0.1464\\
+\hline
+\multirow{10}{4em}{ChIP-nexus} & \multirow{8}{4em}{D.Melanogaster} &
+       \multirow{2}{4em}{Dorsal} & \multirow{4}{4em}{embryo} & 1 &   8,863,170 &  7.27 & 1.0402 &  0.6766 \\
+ & & &  & 2 & 10,003,562 & 7.19 & 1.0672 & 0.5656\\
+ & &  \multirow{2}{4em}{Twist} & & 1 & 18,244,203 & 5.82 &  1.1637 &  0.6592 \\
+ & & &  & 2 & 52,546,982 & 5.27  &  1.1805  & 0.4549 \\
+ & & \multirow{2}{4em}{Max} & \multirow{4}{4em}{S2} & 1 & 18,320,743 & 3.60 & 1.3628 & 0.5178  \\
+ & & & & 2 & 24,965,642  & 3.47 & 1.0138 & 0.2124\\
+ & & \multirow{2}{4em}{MyC} & & 1 & 7,832,034 & 5.92 & 1.0115 &  0.3935  \\
+ & & & &  2 & 22,824,467 & 5.76 & 1.0045 & 0.1879\\
+ & \multirow{2}{4em}{Human} & \multirow{2}{4em}{TBP} & \multirow{2}{4em}{K562} & 1 &  
+              33,708,245 & 32.16 & 1.1712 & 0.3102 \\
+ & &  &  & 2 & 129,675,001 & 32.70 &  1.2455 & 0.0492 \\
+    \lasthline
+  \end{tabu}
+  \caption{Current QC metrics applied to gathered data. NSC stands for Normalized 
+    Strand Cross-Correlation, RSC stands for Relative Strand Cross-Correlation and PBC stands for 
+    PCR Bottleneck Coefficient.}  
+\label{tab:qc}
+\end{table}
+
+DNA libraries generated by ChIP-exo and ChIP-nexus protocols are
+expected to be less complex than the libraries generated by ChIP-Seq,
+since the possible number of positions to which the reads can be
+aligned is being reduced due to the exonuclease digestion,
+considerable amounts of reads are being mapped to specific
+positions. This affects the interpretation of the PBC, since for
+ChIP-Seq low PBC values indicate that the same read has been copied by
+the amplification process and aligned multiple times to the same
+position; while for ChIP-exo when several reads are aligned to the
+same position are not necessarily the same read amplified, but several
+reads that their $5\prime$ end was digested to the same position
+before the amplification step. It is of special importance to notice
+that for deeply sequenced ChIP-exo and ChIP-nexus experiments, the PBC
+values are quite low, which by following the ChIP-Seq guidelines it
+would indicate that those experiment show severe bottlenecking
+problems, which may incorrectly suggest that there is not genomic
+signal int those experiments.
+
+The Strand Cross-Correlation (SCC) introduced by Kharchenko et al.,
+2008 \cite{strandcc} is the most commonly used quality measure in
+ChIP-Seq. It is calculated as the correlation between both strand
+coverages, where each one is shifted $\delta / 2$ bp towards the
+$3\prime$ direction. In general it measures how well the reads mapped
+to each strand are clustered around the locations where the proteins
+are binding to the DNA, and usually it is expected to observed two
+local maxima, one when the profiles are shifted by the average read
+length and another when the profiles are shifted by the unobserved
+fragment length. In a good ChIP-Seq dataset the last one is also the
+SCC global maxima. However, in ChIP-exo's case these two peaks are
+confounded. Hence the Normalized Strand Cross-Correlation (NSC) which
+is a measure based on the SCC is harder to interpret. Figure
+\ref{fig:scc_exo} shows that both local maxima are hard to
+differentiate in the SCC curves for the $\sig$ ChIP-exo datasets used
+to calculate the NSC values in Table \ref{tab:qc}.
 
@@
 \textbf{Analysis of ChIP-nexus data with the QC pipeline}
-
-ChIP-nexus developed by He et al., 2015 \cite{chipnexus} is a
-modification to the ChIP-exo protocol, where both sequencing adaptors
-are ligated at the $3\prime$ end of the ChIP fragments. Then, after
-exonuclease digestion, DNA self-circularization with circLigase, and
-restriction enzyme cutting between the two adaptors, the final library
-is amplified. In Table \ref{tab:nexus} we show the usual ChIP-Seq QC
-metrics to ChIP-nexus data and found that the conclusions made for
-ChIP-exo data still holds, the NSC is high but when plotting the SCC
-the fragment and read length peaks are confounded and the PBC may
-indicate severe bottlenecking but it is caused by the exonuclease
-enzyme digestion which reduces the number of positions where the
-reads can be aligned.
-
-\begin{table}[h!]
-  \centering
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-##            Organism  IP/TF Condition/Cell Rep.       Depth    PBC     NSC
-##  1: D. Melanogaster Dorsal         Embryo    1   8,863,170 0.6766  7.2771
-##  2: D. Melanogaster Dorsal         Embryo    2  10,003,562 0.5657  7.1958
-##  3: D. Melanogaster  Twist         Embryo    1  18,244,203 0.6592  5.8241
-##  4: D. Melanogaster  Twist         Embryo    2  52,546,982 0.4550  5.2775
-##  5: D. Melanogaster    Max             S2    1  18,320,743 0.5178  3.6057
-##  6: D. Melanogaster    Max             S2    2  24,965,642 0.2124  3.4701
-##  7: D. Melanogaster    MyC             S2    1   7,832,034 0.3935  5.9203
-##  8: D. Melanogaster    MyC             S2    2  22,824,467 0.1879  5.7645
-##  9:           Human    TBP           K562    1  33,708,245 0.3102 32.1617
-## 10:           Human    TBP           K562    2 129,675,001 0.0492 32.7037
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
-\caption{\textbf{Current QC metrics applied to ChIP-nexus data.} PBC
-  stands for PCR Bottleneck Coefficient and NSC for Normalized Strand
-  Cross-Correlation.}
-  \label{tab:nexus}
-\end{table}
 
@@
 
-\begin{table}[h!]
-%% need to fix format
-  \centering
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-##    Organism IP/TF Condition/Cell Rep.      Depth    PBC     NSC Protocol
-## 1:   E.Coli Sig70       Rif-0min    1 13,445,022 0.9426  8.8652       PE
-## 2:   E.Coli Sig70       Rif-0min    1 16,538,920 0.9379  7.0318       PE
-## 3:   E.Coli Sig70      Rif-20min    2 11,642,722 0.8892 10.7728       PE
-## 4:   E.Coli Sig70      Rif-20min    2 16,854,026 0.9407  7.9362       PE
-## 5:   E.Coli Sig70       Rif-0min    1  6,722,511 0.6633  9.0178       SE
-## 6:   E.Coli Sig70       Rif-0min    1  8,269,460 0.5594  7.1795       SE
-## 7:   E.Coli Sig70      Rif-20min    2  5,821,361 0.6472 10.8990       SE
-## 8:   E.Coli Sig70      Rif-20min    2  8,427,013 0.5895  8.1247       SE
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
-\caption{Usual ChIP-Seq QC metrics as in table \ref{tab:qc} but
-  applied to Landick's chipseq data of the rif experiment}
-\end{table}
+%% \begin{table}[h!]
+%%  %% need to fix format
+%%   \centering
+
+% \caption{Usual ChIP-Seq QC metrics as in table \ref{tab:qc} but
+%   applied to Landick's chipseq data of the rif experiment}
+% \end{table}
 

### 2016-05-03 rmarkdown/paper.R

@@ -1,84 +1,78 @@
-## ----table1,include = FALSE,echo = FALSE,eval = TRUE---------------------
-  library(data.table)
-  load("../data/paper_tables/Landick_rif_summary.RData")
-  load("../data/paper_tables/Landick_aero_summary.RData")
-  load("../data/paper_tables/pugh_ctcf.RData")
-  load("../data/paper_tables/Carroll_mm9_FoxA1.RData")
-  load("../data/paper_tables/Meijsing_hg19_GR.RData")
-  load("../data/paper_tables/Carroll_hg19_ER.RData")
-  load("../data/paper_tables/Venters_hg19_TBP.RData")
-  load("../data/paper_tables/ChIPnexus_dm3.RData")
-  load("../data/paper_tables/ChIPnexus_hg19.RData")
-  landick_aero[,Organism := "E.Coli"]
-  landick_aero[,Replicate := c(1,2,1,2)]
-  landick_aero[,TF := "Sig70"]
-  landick_aero[ ,cond := c(rep("Aerobic",2),rep("Anaerobic", 2))]
-  landick_rif[,Organism := "E.Coli"]
-  landick_rif[,Replicate := c(1,1,2,2)]
-  landick_rif[,TF := "Sig70"]
-  landick_rif[ ,cond := rep(c("Rif-0min","Rif-20min"),each = 2)]
-  pugh_ctcf[,Organism := "Human"]
-  pugh_ctcf[,Replicate := 1]
-  pugh_ctcf[,TF := "CTCF"]
-  pugh_ctcf[ ,cond := "HeLA"]
-  carroll_mm9_FoxA1[,Organism := "Mouse"]
-  carroll_mm9_FoxA1[,Replicate := c(3,1,2)]
-  carroll_mm9_FoxA1[,TF := "Fox A1"]
-  carroll_mm9_FoxA1[ ,cond := "Mouse Liver"]
-  meijing_hg19_GR[,Organism := "Human"]
-  meijing_hg19_GR[,Replicate := 1]
-  meijing_hg19_GR[, TF := "GR"]
-  meijing_hg19_GR[,cond := c("IMR90","K562","U20S")]
-  carroll_hg19_ER[,Organism := "Human"]  
-  carroll_hg19_ER[,Replicate := c(1,3,2)]
-  carroll_hg19_ER[,TF := "ER"]
-  carroll_hg19_ER[ ,cond := "MCF-7"]
[…]

### 2016-05-03 rmarkdown/paper.Rnw
fixed exploratory analysis, comparison with chip seq and calculate of chip -seq qc metrics with chip-exo and chip-nexus data

@@
 \usepackage{multirow}
+
+\usepackage{soul}
 
@@
 self-circularization with circLigase, and restriction enzyme cutting
-between the two adaptors, the final library is ammplified. ChIP-nexus
+between the two adaptors, the final library is amplified. ChIP-nexus
 arises as an alternative to ChIP-exo as it provides the possibility of
@@
 
-\color{blue}
-
-outline
-
-\begin{enumerate}
-\item Available data, the motivation of this part is two fold:
-%  \begin{itemize}
-%  \item we want to list all data sets we obtained
-%  \item we want to emphasize that for the sig70 datasets we have a
-%    unique design that allows us to compare chip exo with SE and PE
-%    chipseq , the later one for the first time
-%  \end{itemize}
-\item comparison with chip seq %? with and without dpeak ?
-%\item only in methods. list the align algorithm used
-\item chip seq qc metrics
-\item chip exo qc pipeline, description and justification of the steps
-\begin{itemize}
-\item need to justify the sampling step. for this we may use the depth
-  table and add the factor to normalize one depth to another. explain
-  the confounding of pcr amplification and the reduction of the number
-  of possible mappable positions.
-\item case study with FoxA1, analysis of enrichment vs library
-  complexity and analysis and the strand imbalance tools. in methods
-  we need to add which tests did we perform to show that the pipeline
-  works. how do we expect from the diagnostics and how to interpret
-  the plots.
-\item analysis of why the pipeline works. in methods need to specify
-  how and in results need to specify why it works, and what are the
-  conclusions from it
-\end{itemize}
-\item high resolution analysis with chip exo and comparison with PE
-  chip seq
-  \begin{itemize}
-  \item justify what do we used dpeak (in methods need to list the
-    other methods used). perhaps we can add gof plots in
-    supplement. we also want to compare with the same method,
-    therefore we can use the fact that dpeak have a unified model for
-    PE and SE dpeak
-  \item after comparing with other methods we can compare with PE and
-    SE ChIP-seq, this comparison gives also resolution and sensitivity
-  \item compare samples under same budget of reads
-  \end{itemize} 
-\end{enumerate}
-
-\color{black}
-
-\newpage
-
 \subsection*{Comparison with ChIP-Seq data}
 
-<<comp_param , include = FALSE,echo = FALSE,eval = TRUE>>=
-
-  bin_size <- 150
-  frag_len <- 150
-
-@
-
 We first compared various factors that could affect binding site
-identification between ChIP-exo and ChIP-Seq data. In order to compare
-distribution of signal and background between ChIP-exo and ChIP-Seq
-data, we calculated ChIP tag counts across the genome by counting the
-number of reads mapping to each of \Sexpr{bin_size} non-overlapping
-window after extending reads by \Sexpr{frag_len} to their $3\prime$
-end directions. ChIP tag counts in ChIP-exo data were linearly related
-to ChIP tag counts in ChIP-Seq data for the regions with high ChIP tag
-counts (Upper part of Figure \ref{fig:comp}A). This implies that
-signals for potential binding sites are well reproducible between
-ChIP-exo and ChIP-Seq data. On the other hand, there was clear
-difference in the background distribution between them (lower part of
-Figure \ref{fig:comp}A). Specifically, in ChIP-Seq data reads were
-almost uniformly distributed over background (non-binding) regions and
-the ChIP tag counts in there regions were significantly larger than
+identification between ChIP-exo and ChIP-Seq data by using the
+ChIP-exo reads from the first biological sample and first replicate
+grown under aerobic condition (first line from Table \ref{tab:qc_sig})
+and a SE ChIP-Seq replicate grown under the same conditions. In order
+to compare distribution of signal and background between ChIP-exo and
+ChIP-Seq data, we counted the number of extended reads mapping to a
+partition of the genome into non-overlapping bins. ChIP tag counts in
+ChIP-exo data were linearly related to ChIP tag counts in ChIP-Seq
+data for the regions with high ChIP tag counts (Upper part of Figure
+\ref{fig:comp}A). This implies that signals for potential binding
+sites are well reproducible between ChIP-exo and ChIP-Seq data. On the
+other hand, there was a clear difference in the background
+distribution between them (lower part of Figure
+\ref{fig:comp}A). Specifically, in ChIP-Seq data reads were almost
+uniformly distributed over background (non-binding) regions and the
+ChIP tag counts in there regions were significantly larger than
 zero. In contrast, in ChIP-exo data, there was larger variation in
@@
 results indicate that for ChIP-exo data a much smaller portion of the
-genome is expected to be background.
-
-We next evaluated the ``peak pair'' assumption from Rhee and Pugh,
-2011 \cite{exo1}, i.e. a peak of reads in the forward strand is
-usually paired with a peak of reads in the reverse strand that is
-located in the other site of the binding site. Note that currently
-available ChIP-exo data analysis methods, such as Wang et al., 2014
-\cite{mace}, Madrigal 2015 \cite{cexor} and Bardet et al., 2013
-\cite{peakzilla} rely on this assumption. In order to evaluate this
-assumption, we reviewed the proportion of reads in the forward strand
-in ChIP-exo peaks such as at least one binding site is predicted in
-both ChIP-exo and ChIP-Seq data. We found that strands of reads were
-much less balanced in ChIP-exo data than in ChIP-Seq data in these
-regions with potential binding sites (Fig. \ref{fig:comp}B) and this
-indicates that the peak pair assumption might not hold in real
-ChIP-exo data.
+genome is expected to be background. Therefore, methods that
+specifically model the background distribution must take into account
+that ChIP-exo's background is going to be composed by a smaller
+proportion of the reads in the experiment.
+
+Using the same experiment, we next evaluated the ``peak pair''
+assumption from Rhee and Pugh, 2011\nocite{exo1}, i.e. a peak of reads
+in the forward strand is usually paired with a peak of reads in the
+reverse strand that is located in the other site of the binding
+site. Note that currently available ChIP-exo data analysis methods,
+such as Wang et al., 2014\nocite{mace}, Madrigal 2015\nocite{cexor}
+and Bardet et al., 2013\nocite{peakzilla} rely on this assumption. In
+order to evaluate this assumption, we reviewed the proportion of reads
+in the forward strand in high quality ChIP-exo peaks such as at least
+one binding site is predicted in both ChIP-exo and ChIP-Seq data. We
+found that strands of reads were much less balanced in ChIP-exo data
+than in ChIP-Seq data in these regions with potential binding sites
+(Fig. \ref{fig:comp}B) and this indicates that the peak pair
+assumption might not hold in ChIP-exo data. This is caused by either
+the enzyme digestion or the strength at which the protein binds to the
+DNA: Although it is expected for the exonuclease enzyme to digest the
+ChIP fragments starting by their $5\prime$ ends and stopping when
+finding a binding event, it may not be able to reach and digest every
+fragment in the ChIP sample. On the other hand, if the protein is not
+bound to both DNA strand, the enzyme may completely digest the
+fragment. Finally both effect are increased by the PCR amplification.
 
@@
 ChIP tag count increases and for GC - content greater than 0.6 it
-shows a decreasing trend. Kuan et al., 2011 \cite{mosaics} studied the
-presence of the mappability and GC - content biases in ChIP-Seq's
-background. It is not surprising to see these biases also present in
-ChIP-exo data, since ChIP-Seq and ChIP-exo signal seems to be linearly
-correlated for enriched regions (Figure \ref{fig:comp}A). Rozowsky et
-al., 2009 \cite{peakseq} and Valouev et al., 2008 \cite{quest} provide
-in depth analysis of the mappability and GC - content biases for
-ChIP-Seq respectively. Finally, these results indicate that binding
-site identification for ChIP-exo data benefits from using methods that
-take into account of apparent sequence biases such as mappability and
-GC content.
+shows a decreasing trend. Benjamini and Speed,
+2011\nocite{benjamini2011} and Kuan et al., 2011\nocite{mosaics}
+studied the presence of the mappability and GC - content biases in
+ChIP-Seq's background. It is not surprising to see these biases also
+present in ChIP-exo data, since ChIP-Seq and ChIP-exo signal seems to
+be linearly correlated for enriched regions (Figure
+\ref{fig:comp}A). Rozowsky et al., 2009\nocite{peakseq} and Valouev et
+al., 2008\nocite{quest} provide in depth analysis of the mappability
+and GC - content biases for ChIP-Seq respectively. Finally, these
+results indicate that binding site identification for ChIP-exo data
+benefits from using methods that take into account of apparent
+sequence biases such as mappability and GC content, mostly when an
+Input sample is not available.
 
@@
 
-We started our exploration by investigating whether the current
+We continued our exploration by investigating whether the current
 state-of-the-art QC pipelines for ChIP-Seq are suitable for ChIP-exo
-and ChIP-neuxs. In Tables \ref{tab:qc_sig} and \ref{tab:qc} we
+and ChIP-nexus. In Tables \ref{tab:qc_sig} and \ref{tab:qc} we
 calculated a collection of the commonly used ChIP-Seq QC metrics using
@@
   \centering
-  \begin{tabu} to\linewidth{X[-1]|X[-1]|X[-1]|X[-1]|X[-1]|X[-1]|X[-1]}
+  \begin{tabu} to\linewidth{X[-1,m,c]|X[-1]|X[-1]|X[-1,m,c]|X[-1]|X[-1]|X[-1]|X[-1]}
     \firsthline
-    \textbf{Condition} & \textbf{Treatment} & \textbf{Rep.} & \textbf{Depth} & \textbf{NSC} & \textbf{RSC} & 
+    \textbf{Bio Sample} & \textbf{Condition} & \textbf{Treatment} & \textbf{Rep.} & \textbf{Depth} & \textbf{NSC} & \textbf{RSC} & 
     \textbf{PBC}\\
     \hline 
-    Aerobic & No Rif. & 1 & 13,961,493 & 103.15 & 2.0193 & 0.1399 \\
-    Aerobic & No Rif. & 2 & 14,810,838 & 162.70 & 1.7805 & 0.1633 \\
-    Anaerobic & No Rif. & 1 &  16,108,774 & 153.51 & 1.8035 & 0.1353 \\
-    Anaerobic & No Rif. & 2 &  13,636,541 & 172.59 & 2.014 & 0.1532 \\ 
-    Aerobic & No Rif. & 1 & 902,921 & 13.77 & 1.1270 & 0.2689\\
-    Aerobic & No Rif. & 1 & 1,852,124 & 17.91 & 1.5275 & 0.2590\\
-    Aerobic & Rif. 20 min & 2 & 2,104,427 & 29.60 & 1.2844 & 0.2584\\
-    Aerobic & Rif. 20 min & 2 & 11,548,572 & 13.08 & 1.5122 & 0.1510 \\
+    \multirow{4}{1em}{1} & Aerobic & No Rif. & 1 & 13,961,493 & 103.15 & 2.0193 & 0.1399 \\
+    & Aerobic & No Rif. & 2 & 14,810,838 & 162.70 & 1.7805 & 0.1633 \\
+    & Anaerobic & No Rif. & 1 &  16,108,774 & 153.51 & 1.8035 & 0.1353 \\
+    & Anaerobic & No Rif. & 2 &  13,636,541 & 172.59 & 2.014 & 0.1532 \\ 
+    \hline
+    \multirow{4}{1em}{2} & Aerobic & No Rif. & 1 & 902,921 & 13.77 & 1.1270 & 0.2689\\
+    & Aerobic & No Rif. & 1 & 1,852,124 & 17.91 & 1.5275 & 0.2590\\
+    & Aerobic & Rif. 20 min & 2 & 2,104,427 & 29.60 & 1.2844 & 0.2584\\
+    & Aerobic & Rif. 20 min & 2 & 11,548,572 & 13.08 & 1.5122 & 0.1510 \\
     \lasthline
@@
   \centering
-  \begin{tabu} to\linewidth{X[1.3]|X[2.2]|X[-1]|X[-1]|X[-1]|X[1.8]|X[-1]|X[-1]|X[-1]}
+  \begin{tabu} to\linewidth{X[1.3]|X[2.2]|X[-1]|X[1.3]|X[-1,m,c]|X[1.8]|X[-1]|X[-1]|X[-1]}
     \firsthline
@@
     \hline
-    \multirow{13}{4em}{ChIP-exo} & Human & CTCF & HeLa & & 48,478,450 & 16.02 & 1.1960 & 0.4579 \\
-    & \multirow{3}{4em}{Human} & \multirow{3}{4em}{ER} & \multirow{3}{4em}{MCF-7} & 1 & 9,289,835 & 
+    \multirow{13}{1em}{ChIP-exo} & Human & CTCF & HeLa & & 48,478,450 & 16.02 & 1.1960 & 0.4579 \\
+    & \multirow{3}{1em}{Human} & \multirow{3}{1em}{ER} & \multirow{3}{3em}{MCF-7} & 1 & 9,289,835 & 
     19.87 & 1.0127 & 0.8082 \\
@@
     &  &  & & 3 & 12,464,836 & 18.72& 1.0100 & 0.8203 \\    
-    &  \multirow{3}{4em}{Mouse} & \multirow{3}{4em}{FoxA1} & \multirow{3}{4em}{Liver} & 1 & 22,210,461 & 21.28 & 1.1104 &  0.6562 \\
+    &  \multirow{3}{1em}{Mouse} & \multirow{3}{1em}{FoxA1} & \multirow{3}{1em}{Liver} & 1 & 22,210,461 & 21.28 & 1.1104 &  0.6562 \\
     &  & & & 2 & 23,307,557 & 60.42 & 1.1604 & 0.7996 \\ 
     &  & & & 3 & 22,421,72  & 72.04 & 1.1975 & 0.1068 \\
-    & \multirow{3}{4em}{Human} & \multirow{3}{4em}{GR} & IM90 & & 47,443,803 & 8.86  & 1.3678 & 0.2978 \\
+    & \multirow{3}{1em}{Human} & \multirow{3}{1em}{GR} & IMR90 & & 47,443,803 & 8.86  & 1.3678 & 0.2978 \\
     & &  & K562 & & 116,518,000   & 4.11 & 1.0441 &0.0504 \\
     & &  & U2OS & & 3,255,111 &  10.05 & 1.0288 & 0.7714 \\
-    & \multirow{3}{4em}{Human} & \multirow{3}{4em}{TBP} & \multirow{3}{4em}{K562} &1 & 61,046,382 &  12.01  & 1.1119
+    & \multirow{3}{1em}{Human} & \multirow{3}{1em}{TBP} & \multirow{3}{1em}{K562} &1 & 61,046,382 &  12.01  & 1.1119
  & 0.1232 \\
@@
 \hline
-\multirow{10}{4em}{ChIP-nexus} & \multirow{8}{4em}{D.Melanogaster} &
-       \multirow{2}{4em}{Dorsal} & \multirow{4}{4em}{embryo} & 1 &   8,863,170 &  7.27 & 1.0402 &  0.6766 \\
+\multirow{10}{1em}{ChIP-nexus} & \multirow{8}{1em}{D.Melanogaster} &
+       \multirow{2}{1em}{Dorsal} & \multirow{4}{1em}{embryo} & 1 &   8,863,170 &  7.27 & 1.0402 &  0.6766 \\
  & & &  & 2 & 10,003,562 & 7.19 & 1.0672 & 0.5656\\
- & &  \multirow{2}{4em}{Twist} & & 1 & 18,244,203 & 5.82 &  1.1637 &  0.6592 \\
+ & &  \multirow{2}{1em}{Twist} & & 1 & 18,244,203 & 5.82 &  1.1637 &  0.6592 \\
  & & &  & 2 & 52,546,982 & 5.27  &  1.1805  & 0.4549 \\
- & & \multirow{2}{4em}{Max} & \multirow{4}{4em}{S2} & 1 & 18,320,743 & 3.60 & 1.3628 & 0.5178  \\
+ & & \multirow{2}{1em}{Max} & \multirow{4}{1em}{S2} & 1 & 18,320,743 & 3.60 & 1.3628 & 0.5178  \\
  & & & & 2 & 24,965,642  & 3.47 & 1.0138 & 0.2124\\
- & & \multirow{2}{4em}{MyC} & & 1 & 7,832,034 & 5.92 & 1.0115 &  0.3935  \\
+ & & \multirow{2}{1em}{MyC} & & 1 & 7,832,034 & 5.92 & 1.0115 &  0.3935  \\
  & & & &  2 & 22,824,467 & 5.76 & 1.0045 & 0.1879\\
- & \multirow{2}{4em}{Human} & \multirow{2}{4em}{TBP} & \multirow{2}{4em}{K562} & 1 &  
+ & \multirow{2}{1em}{Human} & \multirow{2}{1em}{TBP} & \multirow{2}{1em}{K562} & 1 &  
               33,708,245 & 32.16 & 1.1712 & 0.3102 \\
@@
 the amplification process and aligned multiple times to the same
-position; while for ChIP-exo when several reads are aligned to the
-same position are not necessarily the same read amplified, but several
-reads that their $5\prime$ end was digested to the same position
-before the amplification step. It is of special importance to notice
-that for deeply sequenced ChIP-exo and ChIP-nexus experiments, the PBC
-values are quite low, which by following the ChIP-Seq guidelines it
-would indicate that those experiment show severe bottlenecking
-problems, which may incorrectly suggest that there is not genomic
-signal int those experiments.
+position; while for ChIP-exo and ChIP-nexus when several reads are
+aligned to the same position are not necessarily the same read
+amplified, but several reads that their $5\prime$ end was digested to
+the same position before the amplification step. It is of special
+importance to notice that for deeply sequenced ChIP-exo and ChIP-nexus
+experiments, the PBC values are quite low, which by following the
+ChIP-Seq guidelines it would indicate that those experiment show
+severe bottlenecking problems, which may incorrectly suggest that the
+positions with a large amount of aligned reads are being caused by PCR
+amplification rather than observed genomic signal.
 
 The Strand Cross-Correlation (SCC) introduced by Kharchenko et al.,
-2008 \cite{strandcc} is the most commonly used quality measure in
-ChIP-Seq. It is calculated as the correlation between both strand
-coverages, where each one is shifted $\delta / 2$ bp towards the
-$3\prime$ direction. In general it measures how well the reads mapped
-to each strand are clustered around the locations where the proteins
-are binding to the DNA, and usually it is expected to observed two
-local maxima, one when the profiles are shifted by the average read
-length and another when the profiles are shifted by the unobserved
-fragment length. In a good ChIP-Seq dataset the last one is also the
-SCC global maxima. However, in ChIP-exo's case these two peaks are
-confounded. Hence the Normalized Strand Cross-Correlation (NSC) which
-is a measure based on the SCC is harder to interpret. Figure
-\ref{fig:scc_exo} shows that both local maxima are hard to
-differentiate in the SCC curves for the $\sig$ ChIP-exo datasets used
-to calculate the NSC values in Table \ref{tab:qc}.
+2008\nocite{strandcc} is the most commonly used quality measure in
+ChIP-Seq. In general it measures how well the reads mapped to each
+strand are clustered around the locations where the proteins are
+binding to the DNA, and usually it is expected to observed two local
+maxima, one when the profiles are shifted by the average read length
+and another when the profiles are shifted by the unobserved fragment
+length. In a high quality ChIP-Seq dataset the last one is also the
+SCC global maxima. On the other hand, as a thought experiment in an
+idealized ChIP-exo experiment where the ChIP fragments are digested
+until they found the binding protein and this protein bound to the
+whole motif sequence, then the SCC would a flat function with a jump
+at the motif length; since usually this is not the case, we also
+expect to observe a local maxima at the read length. Figure
+\ref{fig:scc_exo} shows the SCC curves for the CTCF factor from human
+genome: The ChIP-exo curve shows a local maxima at the motif and read
+lengths, while the SE ChIP-Seq curves have a local maxima at the read
+length and a global maxima at the unobserved fragment
+length. Therefore, in ChIP-exo's cases both peaks are likely to be
+confounded. Hence measures based in the SCC such as the Normalized
+Strand Cross-Correlation (NSC) or the Relative Strand
+Cross-Correlation (RSC) are harder to interpret.
+
+% that both
+% local maxima are hard to differentiate in the SCC curves for the
+% $\sig$ ChIP-exo datasets used to calculate the NSC values in Table
+% \ref{tab:qc}.
+
+% It is calculated as the correlation between both strand
+% coverages, where each one is shifted $\delta / 2$ bp towards the
+% $3\prime$ direction
+\color{blue}
+
+outline
+
+\begin{enumerate}
+\item \st{Available data, the motivation of this part is two fold:}
+ \begin{itemize}
+ \item \st{we want to list all data sets we obtained}
+ \item \st{we want to emphasize that for the sig70 datasets we have a
+   unique design that allows us to compare chip exo with SE and PE
+   chipseq , the later one for the first time}
+ \end{itemize}
+\item \st{comparison with chip seq} %? with and without dpeak ?}
+\item \st{only in methods. list the align algorithm used}
+\item \st{chip seq qc metrics}
+\item chip exo qc pipeline, description and justification of the steps
+\begin{itemize}
+\item need to justify the sampling step. for this we may use the depth
+  table and add the factor to normalize one depth to another. explain
+  the confounding of pcr amplification and the reduction of the number
+  of possible mappable positions.
+\item case study with FoxA1, analysis of enrichment vs library
+  complexity and analysis and the strand imbalance tools. in methods
+  we need to add which tests did we perform to show that the pipeline
+  works. how do we expect from the diagnostics and how to interpret
+  the plots.
+\item analysis of why the pipeline works. in methods need to specify
+  how and in results need to specify why it works, and what are the
+  conclusions from it
+\end{itemize}
+\item high resolution analysis with chip exo and comparison with PE
+  chip seq
+  \begin{itemize}
+  \item justify what do we used dpeak (in methods need to list the
+    other methods used). perhaps we can add gof plots in
+    supplement. we also want to compare with the same method,
+    therefore we can use the fact that dpeak have a unified model for
+    PE and SE dpeak
+  \item after comparing with other methods we can compare with PE and
+    SE ChIP-seq, this comparison gives also resolution and sensitivity
+  \item compare samples under same budget of reads
+  \end{itemize} 
+\end{enumerate}
+
+\color{black}
 
@@
 \emph{bowtie} in its 1.1.2 version.
+
+\subsection*{Comparison with ChIP-Seq data}
+
+<<comp_param , include = FALSE,echo = FALSE,eval = TRUE>>=
+
+  bin_size <- 150
+  frag_len <- 150
+
+@
+
+Using the $\sig$ sample grown under aerobic condition for ChIP-exo and
+SE ChIP-Seq. We partitioned the \emph{E.Coli} genome into
+non-overlaping bins of length \Sexpr{bin_size}, and for each bin we
+counted how many ChIP-exo and SE ChIP-Seq extended reads overlapped
+each bin. The reads were extended by \Sexpr{frag_len} bp to their
+$3\prime$ directions.
+
+To model the strand imbalance, we called a set of high quality peaks
+for ChIP-exo and SE ChIP-Seq (with $\mbox{GC content + Mappability}$
+and $\mbox{Input only}$ models respectively), such that each peak
+contained at least one binding event (found by using \textbf{dPeak}
+with at most 5 binding events in each peak). For each peak, we
+calculated the FSR as:
+
+\begin{align*}
+    \mbox{FSR} &= \frac{\text{Nr. of fwd. strand reads mapped to the region}}                 {\text{Total nr. of reads mapped to the region}} 
+\end{align*}
+
+We estimated the $\mbox{FSR}$'s densities by using R's \emph{density}
+function using its default parameters, which computes gaussian kernel
+density estimates.
 
@@
   \centering
-\includegraphics[width = .8\textwidth]{/p/keles/ChIPexo/volume3/ChIPexo/figs/for_paper/EColi_strand_cross_corr.pdf}
-% \caption{\textbf{SCC curves for $\sig$ samples.} The ``phantom peak''
-%   and the summit that corresponds to the read and fragment length
-%   respectively are confounded due to the exonuclease digestion.}
-\caption{SCC curves for $\sig$ samples. The ``phantom peak'' and the
-  summit that corresponds to the read and fragment length respectively
-  are confounded due to the exonuclease digestion.}
-  \label{fig:scc_exo}
-\end{figure}
-
-\newpage
-
-\begin{figure}[h!]
-  \centering
   \includegraphics[width = .46\textwidth,page = 3 ]{../figs/for_paper/ChIPseqPET_ChIPexo_tagCount_comparison.pdf}
@@
 
-% ChIP tag counts increase linearly as mappability scores
-% increases. D) ChIP tag counts increase linearly as GC content score
-% increases when GC content is less than 0.6 and then ChIP tag counts
-% decrease as GC content increases. D) Percentage of sites with at
-% least p.value.
+\newpage
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width = .8 \textwidth]{/p/keles/ChIPexo/volume3/ChIPexo/figs/for_paper/scc_ctcf2.pdf}
+% \includegraphics[width = .8\textwidth]{/p/keles/ChIPexo/volume3/ChIPexo/figs/for_paper/EColi_strand_cross_corr.pdf}
+% \caption{\textbf{SCC curves for $\sig$ samples.} The ``phantom peak''
+%   and the summit that corresponds to the read and fragment length
+%   respectively are confounded due to the exonuclease digestion.}
+\caption{SCC curves for human CTCF on HeLa cell lines.  The SCC curve
+  for the ChIP-exo sample from \cite{exo1} is shown in the left panel,
+  and the SCC for ChIP-Seq samples from
+  \textbf{https://www.encodeproject.org/experiments/ENCSR000AOA/} are
+  shown in the right panel. The ChIP-exo curve shows local maxima at
+  the motif and read length. Both SE ChIP-Seq curves are maximized at
+  the fragments length and show a local maxima at its read length. }
+  \label{fig:scc_exo}
+\end{figure}
 
@@
 
-% LocalWords:  Ntimes
+% LocalWords:  Ntimes overlaping gaussian

### 2016-05-03 rmarkdown/paper.tex
fixed exploratory analysis, comparison with chip seq and calculate of chip -seq qc metrics with chip-exo and chip-nexus data

@@
 
+\subsection*{Comparison with ChIP-Seq data}
+
+We first compared various factors that could affect binding site
+identification between ChIP-exo and ChIP-Seq data by using the
+ChIP-exo reads from the first biological sample and first replicate
+grown under aerobic condition (first line from Table \ref{tab:qc_sig})
+and a SE ChIP-Seq replicate grown under the same conditions. In order
+to compare distribution of signal and background between ChIP-exo and
+ChIP-Seq data, we counted the number of extended reads mapping to a
+partition of the genome into non-overlapping bins. ChIP tag counts in
+ChIP-exo data were linearly related to ChIP tag counts in ChIP-Seq
+data for the regions with high ChIP tag counts (Upper part of Figure
+\ref{fig:comp}A). This implies that signals for potential binding
+sites are well reproducible between ChIP-exo and ChIP-Seq data. On the
+other hand, there was a clear difference in the background
+distribution between them (lower part of Figure
+\ref{fig:comp}A). Specifically, in ChIP-Seq data reads were almost
+uniformly distributed over background (non-binding) regions and the
+ChIP tag counts in there regions were significantly larger than
+zero. In contrast, in ChIP-exo data, there was larger variation in
+ChIP tag counts among background regions and ChIP tag counts were much
+lower in these regions compared to ChIP-Seq data. There were also
+large proportion of regions without any read in ChIP-exo data. These
+results indicate that for ChIP-exo data a much smaller portion of the
+genome is expected to be background. Therefore, methods that
+specifically model the background distribution must take into account
+that ChIP-exo's background is going to be composed by a smaller
+proportion of the reads in the experiment.
+
+Using the same experiment, we next evaluated the ``peak pair''
+assumption from Rhee and Pugh, 2011\nocite{exo1}, i.e. a peak of reads
+in the forward strand is usually paired with a peak of reads in the
+reverse strand that is located in the other site of the binding
+site. Note that currently available ChIP-exo data analysis methods,
+such as Wang et al., 2014\nocite{mace}, Madrigal 2015\nocite{cexor}
+and Bardet et al., 2013\nocite{peakzilla} rely on this assumption. In
+order to evaluate this assumption, we reviewed the proportion of reads
+in the forward strand in high quality ChIP-exo peaks such as at least
+one binding site is predicted in both ChIP-exo and ChIP-Seq data. We
+found that strands of reads were much less balanced in ChIP-exo data
+than in ChIP-Seq data in these regions with potential binding sites
+(Fig. \ref{fig:comp}B) and this indicates that the peak pair
+assumption might not hold in ChIP-exo data. This is caused by either
+the enzyme digestion or the strength at which the protein binds to the
+DNA: Although it is expected for the exonuclease enzyme to digest the
+ChIP fragments starting by their $5\prime$ ends and stopping when
+finding a binding event, it may not be able to reach and digest every
+fragment in the ChIP sample. On the other hand, if the protein is not
+bound to both DNA strand, the enzyme may completely digest the
+fragment. Finally both effect are increased by the PCR amplification.
+
+We evaluated ChIP-exo data for CTCF factor from human genome
+\cite{exo1} to investigate issues specific to eukaryotic genomes for
+binding sites identification. Figures \ref{fig:comp}C and
+\ref{fig:comp}D display the bin-level average read counts against
+mappability and GC content. Each data point is obtained by averaging
+the read counts across bins with the same mappability of GC -
+content. In Figure \ref{fig:comp}C it is shown that the ChIP-exo tag
+counts linearly increases with the mappability score and in Figure
+\ref{fig:comp}D it is shown that for GC - content below 0.6, the mean
+ChIP tag count increases and for GC - content greater than 0.6 it
+shows a decreasing trend. Benjamini and Speed,
+2011\nocite{benjamini2011} and Kuan et al., 2011\nocite{mosaics}
+studied the presence of the mappability and GC - content biases in
+ChIP-Seq's background. It is not surprising to see these biases also
+present in ChIP-exo data, since ChIP-Seq and ChIP-exo signal seems to
+be linearly correlated for enriched regions (Figure
+\ref{fig:comp}A). Rozowsky et al., 2009\nocite{peakseq} and Valouev et
+al., 2008\nocite{quest} provide in depth analysis of the mappability
+and GC - content biases for ChIP-Seq respectively. Finally, these
+results indicate that binding site identification for ChIP-exo data
+benefits from using methods that take into account of apparent
+sequence biases such as mappability and GC content, mostly when an
+Input sample is not available.
+
+\subsection*{Application of ENCODE quality metrics to ChIP-exo and
+  ChIP-nexus data}
+
+We continued our exploration by investigating whether the current
+state-of-the-art QC pipelines for ChIP-Seq are suitable for ChIP-exo
+and ChIP-nexus. In Tables \ref{tab:qc_sig} and \ref{tab:qc} we
+calculated a collection of the commonly used ChIP-Seq QC metrics using
+the ChIP-exo and ChIP-nexus experiments instead: Normalized Strand
+Cross-Correlation (NSC), Relative Strand Cross-Correlation and PCR
+Bottleneck Coefficient (PBC) defined as in
+\url{https://genome.ucsc.edu/ENCODE/qualityMetrics.html}\nocite{encode_qc}.
+
+%% We can observe that the PBC is low, hence it would be incorrectly
+%% suggested to repeat the experiment.
+
+\arrayrulewidth=.1mm
+
+\begin{table}[h!]
+  \centering
+  \begin{tabu} to\linewidth{X[-1,m,c]|X[-1]|X[-1]|X[-1,m,c]|X[-1]|X[-1]|X[-1]|X[-1]}
+    \firsthline
+    \textbf{Bio Sample} & \textbf{Condition} & \textbf{Treatment} & \textbf{Rep.} & \textbf{Depth} & \textbf{NSC} & \textbf{RSC} & 
+    \textbf{PBC}\\
+    \hline 
+    \multirow{4}{1em}{1} & Aerobic & No Rif. & 1 & 13,961,493 & 103.15 & 2.0193 & 0.1399 \\
+    & Aerobic & No Rif. & 2 & 14,810,838 & 162.70 & 1.7805 & 0.1633 \\
+    & Anaerobic & No Rif. & 1 &  16,108,774 & 153.51 & 1.8035 & 0.1353 \\
+    & Anaerobic & No Rif. & 2 &  13,636,541 & 172.59 & 2.014 & 0.1532 \\ 
+    \hline
+    \multirow{4}{1em}{2} & Aerobic & No Rif. & 1 & 902,921 & 13.77 & 1.1270 & 0.2689\\
+    & Aerobic & No Rif. & 1 & 1,852,124 & 17.91 & 1.5275 & 0.2590\\
+    & Aerobic & Rif. 20 min & 2 & 2,104,427 & 29.60 & 1.2844 & 0.2584\\
+    & Aerobic & Rif. 20 min & 2 & 11,548,572 & 13.08 & 1.5122 & 0.1510 \\
+    \lasthline
+    \end{tabu}
+    \caption{Current QC metrics applied to generated \emph{E. Coli} $\sig$ samples. NSC stands for Normalized 
+             Strand Cross-Correlation, RSC stands for Relative Strand Cross-Correlation and PBC stands for 
+             PCR Bottleneck Coefficient.}
+\label{tab:qc_sig}
+\end{table}
+
+\begin{table}[h!]
+  \centering
+  \begin{tabu} to\linewidth{X[1.3]|X[2.2]|X[-1]|X[1.3]|X[-1,m,c]|X[1.8]|X[-1]|X[-1]|X[-1]}
+    \firsthline
+    \textbf{Protocol} & \textbf{Organism} & \textbf{TF} & \textbf{Cell line} & \textbf{Rep.} &
+    \textbf{Depth} & \textbf{NSC} & \textbf{RSC} &   \textbf{PBC} \\ 
+    \hline
+    \multirow{13}{1em}{ChIP-exo} & Human & CTCF & HeLa & & 48,478,450 & 16.02 & 1.1960 & 0.4579 \\
+    & \multirow{3}{1em}{Human} & \multirow{3}{1em}{ER} & \multirow{3}{3em}{MCF-7} & 1 & 9,289,835 & 
+    19.87 & 1.0127 & 0.8082 \\
+    &  &  & & 2 & 11,041,833 & 21.48& 1.0063 & 0.8024\\
+    &  &  & & 3 & 12,464,836 & 18.72& 1.0100 & 0.8203 \\    
+    &  \multirow{3}{1em}{Mouse} & \multirow{3}{1em}{FoxA1} & \multirow{3}{1em}{Liver} & 1 & 22,210,461 & 21.28 & 1.1104 &  0.6562 \\
+    &  & & & 2 & 23,307,557 & 60.42 & 1.1604 & 0.7996 \\ 
+    &  & & & 3 & 22,421,72  & 72.04 & 1.1975 & 0.1068 \\
+    & \multirow{3}{1em}{Human} & \multirow{3}{1em}{GR} & IMR90 & & 47,443,803 & 8.86  & 1.3678 & 0.2978 \\
+    & &  & K562 & & 116,518,000   & 4.11 & 1.0441 &0.0504 \\
+    & &  & U2OS & & 3,255,111 &  10.05 & 1.0288 & 0.7714 \\
+    & \multirow{3}{1em}{Human} & \multirow{3}{1em}{TBP} & \multirow{3}{1em}{K562} &1 & 61,046,382 &  12.01  & 1.1119
+ & 0.1232 \\
+    & & & &2 & 94,314,770 & 7.93 & 1.0299 & 0.1681 \\
+    & & & & 3 & 114,282,270 & 9.25 & 1.1027 & 0.1464\\
+\hline
+\multirow{10}{1em}{ChIP-nexus} & \multirow{8}{1em}{D.Melanogaster} &
+       \multirow{2}{1em}{Dorsal} & \multirow{4}{1em}{embryo} & 1 &   8,863,170 &  7.27 & 1.0402 &  0.6766 \\
+ & & &  & 2 & 10,003,562 & 7.19 & 1.0672 & 0.5656\\
+ & &  \multirow{2}{1em}{Twist} & & 1 & 18,244,203 & 5.82 &  1.1637 &  0.6592 \\
+ & & &  & 2 & 52,546,982 & 5.27  &  1.1805  & 0.4549 \\
+ & & \multirow{2}{1em}{Max} & \multirow{4}{1em}{S2} & 1 & 18,320,743 & 3.60 & 1.3628 & 0.5178  \\
+ & & & & 2 & 24,965,642  & 3.47 & 1.0138 & 0.2124\\
+ & & \multirow{2}{1em}{MyC} & & 1 & 7,832,034 & 5.92 & 1.0115 &  0.3935  \\
+ & & & &  2 & 22,824,467 & 5.76 & 1.0045 & 0.1879\\
+ & \multirow{2}{1em}{Human} & \multirow{2}{1em}{TBP} & \multirow{2}{1em}{K562} & 1 &  
+              33,708,245 & 32.16 & 1.1712 & 0.3102 \\
+ & &  &  & 2 & 129,675,001 & 32.70 &  1.2455 & 0.0492 \\
+    \lasthline
+  \end{tabu}
+  \caption{Current QC metrics applied to gathered data. NSC stands for Normalized 
+    Strand Cross-Correlation, RSC stands for Relative Strand Cross-Correlation and PBC stands for 
+    PCR Bottleneck Coefficient.}  
+\label{tab:qc}
+\end{table}
+
+DNA libraries generated by ChIP-exo and ChIP-nexus protocols are
+expected to be less complex than the libraries generated by ChIP-Seq,
+since the possible number of positions to which the reads can be
+aligned is being reduced due to the exonuclease digestion,
+considerable amounts of reads are being mapped to specific
+positions. This affects the interpretation of the PBC, since for
+ChIP-Seq low PBC values indicate that the same read has been copied by
+the amplification process and aligned multiple times to the same
+position; while for ChIP-exo and ChIP-nexus when several reads are
+aligned to the same position are not necessarily the same read
+amplified, but several reads that their $5\prime$ end was digested to
+the same position before the amplification step. It is of special
+importance to notice that for deeply sequenced ChIP-exo and ChIP-nexus
+experiments, the PBC values are quite low, which by following the
+ChIP-Seq guidelines it would indicate that those experiment show
+severe bottlenecking problems, which may incorrectly suggest that the
+positions with a large amount of aligned reads are being caused by PCR
+amplification rather than observed genomic signal.
+
+The Strand Cross-Correlation (SCC) introduced by Kharchenko et al.,
+2008\nocite{strandcc} is the most commonly used quality measure in
+ChIP-Seq. In general it measures how well the reads mapped to each
+strand are clustered around the locations where the proteins are
+binding to the DNA, and usually it is expected to observed two local
+maxima, one when the profiles are shifted by the average read length
+and another when the profiles are shifted by the unobserved fragment
+length. In a high quality ChIP-Seq dataset the last one is also the
+SCC global maxima. On the other hand, as a thought experiment in an
+idealized ChIP-exo experiment where the ChIP fragments are digested
+until they found the binding protein and this protein bound to the
+whole motif sequence, then the SCC would a flat function with a jump
+at the motif length; since usually this is not the case, we also
+expect to observe a local maxima at the read length. Figure
+\ref{fig:scc_exo} shows the SCC curves for the CTCF factor from human
+genome: The ChIP-exo curve shows a local maxima at the motif and read
+lengths, while the SE ChIP-Seq curves have a local maxima at the read
+length and a global maxima at the unobserved fragment
+length. Therefore, in ChIP-exo's cases both peaks are likely to be
+confounded. Hence measures based in the SCC such as the Normalized
+Strand Cross-Correlation (NSC) or the Relative Strand
+Cross-Correlation (RSC) are harder to interpret.
+
+% that both
+% local maxima are hard to differentiate in the SCC curves for the
+% $\sig$ ChIP-exo datasets used to calculate the NSC values in Table
+% \ref{tab:qc}.
+
+% It is calculated as the correlation between both strand
+% coverages, where each one is shifted $\delta / 2$ bp towards the
+% $3\prime$ direction
 \color{blue}
@@
 \begin{enumerate}
-\item Available data, the motivation of this part is two fold:
-%  \begin{itemize}
-%  \item we want to list all data sets we obtained
-%  \item we want to emphasize that for the sig70 datasets we have a
-%    unique design that allows us to compare chip exo with SE and PE
-%    chipseq , the later one for the first time
-%  \end{itemize}
-\item comparison with chip seq %? with and without dpeak ?
-%\item only in methods. list the align algorithm used
-\item chip seq qc metrics
+\item \st{Available data, the motivation of this part is two fold:}
+ \begin{itemize}
+ \item \st{we want to list all data sets we obtained}
+ \item \st{we want to emphasize that for the sig70 datasets we have a
+   unique design that allows us to compare chip exo with SE and PE
+   chipseq , the later one for the first time}
+ \end{itemize}
+\item \st{comparison with chip seq} %? with and without dpeak ?}
+\item \st{only in methods. list the align algorithm used}
+\item \st{chip seq qc metrics}
 \item chip exo qc pipeline, description and justification of the steps
@@
 
-\newpage
-
-\subsection*{Comparison with ChIP-Seq data}
-
-We first compared various factors that could affect binding site
-identification between ChIP-exo and ChIP-Seq data. In order to compare
-distribution of signal and background between ChIP-exo and ChIP-Seq
-data, we calculated ChIP tag counts across the genome by counting the
-number of reads mapping to each of 150 non-overlapping
-window after extending reads by 150 to their $3\prime$
-end directions. ChIP tag counts in ChIP-exo data were linearly related
-to ChIP tag counts in ChIP-Seq data for the regions with high ChIP tag
-counts (Upper part of Figure \ref{fig:comp}A). This implies that
-signals for potential binding sites are well reproducible between
-ChIP-exo and ChIP-Seq data. On the other hand, there was clear
-difference in the background distribution between them (lower part of
-Figure \ref{fig:comp}A). Specifically, in ChIP-Seq data reads were
-almost uniformly distributed over background (non-binding) regions and
-the ChIP tag counts in there regions were significantly larger than
-zero. In contrast, in ChIP-exo data, there was larger variation in
-ChIP tag counts among background regions and ChIP tag counts were much
-lower in these regions compared to ChIP-Seq data. There were also
-large proportion of regions without any read in ChIP-exo data. These
-results indicate that for ChIP-exo data a much smaller portion of the
-genome is expected to be background.
-
-We next evaluated the ``peak pair'' assumption from Rhee and Pugh,
-2011 \cite{exo1}, i.e. a peak of reads in the forward strand is
-usually paired with a peak of reads in the reverse strand that is
-located in the other site of the binding site. Note that currently
-available ChIP-exo data analysis methods, such as Wang et al., 2014
-\cite{mace}, Madrigal 2015 \cite{cexor} and Bardet et al., 2013
-\cite{peakzilla} rely on this assumption. In order to evaluate this
-assumption, we reviewed the proportion of reads in the forward strand
-in ChIP-exo peaks such as at least one binding site is predicted in
-both ChIP-exo and ChIP-Seq data. We found that strands of reads were
-much less balanced in ChIP-exo data than in ChIP-Seq data in these
-regions with potential binding sites (Fig. \ref{fig:comp}B) and this
-indicates that the peak pair assumption might not hold in real
-ChIP-exo data.
-
-We evaluated ChIP-exo data for CTCF factor from human genome
-\cite{exo1} to investigate issues specific to eukaryotic genomes for
-binding sites identification. Figures \ref{fig:comp}C and
-\ref{fig:comp}D display the bin-level average read counts against
-mappability and GC content. Each data point is obtained by averaging
-the read counts across bins with the same mappability of GC -
-content. In Figure \ref{fig:comp}C it is shown that the ChIP-exo tag
-counts linearly increases with the mappability score and in Figure
-\ref{fig:comp}D it is shown that for GC - content below 0.6, the mean
-ChIP tag count increases and for GC - content greater than 0.6 it
-shows a decreasing trend. Kuan et al., 2011 \cite{mosaics} studied the
-presence of the mappability and GC - content biases in ChIP-Seq's
-background. It is not surprising to see these biases also present in
-ChIP-exo data, since ChIP-Seq and ChIP-exo signal seems to be linearly
-correlated for enriched regions (Figure \ref{fig:comp}A). Rozowsky et
-al., 2009 \cite{peakseq} and Valouev et al., 2008 \cite{quest} provide
-in depth analysis of the mappability and GC - content biases for
-ChIP-Seq respectively. Finally, these results indicate that binding
-site identification for ChIP-exo data benefits from using methods that
-take into account of apparent sequence biases such as mappability and
-GC content.
-
-\subsection*{Application of ENCODE quality metrics to ChIP-exo and
-  ChIP-nexus data}
-
-We started our exploration by investigating whether the current
-state-of-the-art QC pipelines for ChIP-Seq are suitable for ChIP-exo
-and ChIP-neuxs. In Tables \ref{tab:qc_sig} and \ref{tab:qc} we
-calculated a collection of the commonly used ChIP-Seq QC metrics using
-the ChIP-exo and ChIP-nexus experiments instead: Normalized Strand
-Cross-Correlation (NSC), Relative Strand Cross-Correlation and PCR
-Bottleneck Coefficient (PBC) defined as in 
-\url{https://genome.ucsc.edu/ENCODE/qualityMetrics.html}\nocite{encode_qc}.
-
-%% We can observe that the PBC is low, hence it would be incorrectly
-%% suggested to repeat the experiment.
-
-\arrayrulewidth=.1mm
-
-\begin{table}[h!]
-  \centering
-  \begin{tabu} to\linewidth{X[-1]|X[-1]|X[-1]|X[-1]|X[-1]|X[-1]|X[-1]}
-    \firsthline
-    \textbf{Condition} & \textbf{Treatment} & \textbf{Rep.} & \textbf{Depth} & \textbf{NSC} & \textbf{RSC} & 
-    \textbf{PBC}\\
-    \hline 
-    Aerobic & No Rif. & 1 & 13,961,493 & 103.15 & 2.0193 & 0.1399 \\
-    Aerobic & No Rif. & 2 & 14,810,838 & 162.70 & 1.7805 & 0.1633 \\
-    Anaerobic & No Rif. & 1 &  16,108,774 & 153.51 & 1.8035 & 0.1353 \\
-    Anaerobic & No Rif. & 2 &  13,636,541 & 172.59 & 2.014 & 0.1532 \\ 
-    Aerobic & No Rif. & 1 & 902,921 & 13.77 & 1.1270 & 0.2689\\
-    Aerobic & No Rif. & 1 & 1,852,124 & 17.91 & 1.5275 & 0.2590\\
-    Aerobic & Rif. 20 min & 2 & 2,104,427 & 29.60 & 1.2844 & 0.2584\\
-    Aerobic & Rif. 20 min & 2 & 11,548,572 & 13.08 & 1.5122 & 0.1510 \\
-    \lasthline
-    \end{tabu}
-    \caption{Current QC metrics applied to generated \emph{E. Coli} $\sig$ samples. NSC stands for Normalized 
-             Strand Cross-Correlation, RSC stands for Relative Strand Cross-Correlation and PBC stands for 
-             PCR Bottleneck Coefficient.}
-\label{tab:qc_sig}
-\end{table}
-
-\begin{table}[h!]
-  \centering
-  \begin{tabu} to\linewidth{X[1.3]|X[2.2]|X[-1]|X[-1]|X[-1]|X[1.8]|X[-1]|X[-1]|X[-1]}
-    \firsthline
-    \textbf{Protocol} & \textbf{Organism} & \textbf{TF} & \textbf{Cell line} & \textbf{Rep.} &
-    \textbf{Depth} & \textbf{NSC} & \textbf{RSC} &   \textbf{PBC} \\ 
-    \hline
-    \multirow{13}{4em}{ChIP-exo} & Human & CTCF & HeLa & & 48,478,450 & 16.02 & 1.1960 & 0.4579 \\
-    & \multirow{3}{4em}{Human} & \multirow{3}{4em}{ER} & \multirow{3}{4em}{MCF-7} & 1 & 9,289,835 & 
-    19.87 & 1.0127 & 0.8082 \\
-    &  &  & & 2 & 11,041,833 & 21.48& 1.0063 & 0.8024\\
-    &  &  & & 3 & 12,464,836 & 18.72& 1.0100 & 0.8203 \\    
-    &  \multirow{3}{4em}{Mouse} & \multirow{3}{4em}{FoxA1} & \multirow{3}{4em}{Liver} & 1 & 22,210,461 & 21.28 & 1.1104 &  0.6562 \\
-    &  & & & 2 & 23,307,557 & 60.42 & 1.1604 & 0.7996 \\ 
-    &  & & & 3 & 22,421,72  & 72.04 & 1.1975 & 0.1068 \\
-    & \multirow{3}{4em}{Human} & \multirow{3}{4em}{GR} & IM90 & & 47,443,803 & 8.86  & 1.3678 & 0.2978 \\
-    & &  & K562 & & 116,518,000   & 4.11 & 1.0441 &0.0504 \\
-    & &  & U2OS & & 3,255,111 &  10.05 & 1.0288 & 0.7714 \\
-    & \multirow{3}{4em}{Human} & \multirow{3}{4em}{TBP} & \multirow{3}{4em}{K562} &1 & 61,046,382 &  12.01  & 1.1119
- & 0.1232 \\
-    & & & &2 & 94,314,770 & 7.93 & 1.0299 & 0.1681 \\
-    & & & & 3 & 114,282,270 & 9.25 & 1.1027 & 0.1464\\
-\hline
-\multirow{10}{4em}{ChIP-nexus} & \multirow{8}{4em}{D.Melanogaster} &
-       \multirow{2}{4em}{Dorsal} & \multirow{4}{4em}{embryo} & 1 &   8,863,170 &  7.27 & 1.0402 &  0.6766 \\
- & & &  & 2 & 10,003,562 & 7.19 & 1.0672 & 0.5656\\
- & &  \multirow{2}{4em}{Twist} & & 1 & 18,244,203 & 5.82 &  1.1637 &  0.6592 \\
- & & &  & 2 & 52,546,982 & 5.27  &  1.1805  & 0.4549 \\
- & & \multirow{2}{4em}{Max} & \multirow{4}{4em}{S2} & 1 & 18,320,743 & 3.60 & 1.3628 & 0.5178  \\
- & & & & 2 & 24,965,642  & 3.47 & 1.0138 & 0.2124\\
- & & \multirow{2}{4em}{MyC} & & 1 & 7,832,034 & 5.92 & 1.0115 &  0.3935  \\
- & & & &  2 & 22,824,467 & 5.76 & 1.0045 & 0.1879\\
- & \multirow{2}{4em}{Human} & \multirow{2}{4em}{TBP} & \multirow{2}{4em}{K562} & 1 &  
-              33,708,245 & 32.16 & 1.1712 & 0.3102 \\
- & &  &  & 2 & 129,675,001 & 32.70 &  1.2455 & 0.0492 \\
-    \lasthline
-  \end{tabu}
-  \caption{Current QC metrics applied to gathered data. NSC stands for Normalized 
-    Strand Cross-Correlation, RSC stands for Relative Strand Cross-Correlation and PBC stands for 
-    PCR Bottleneck Coefficient.}  
-\label{tab:qc}
-\end{table}
-
-DNA libraries generated by ChIP-exo and ChIP-nexus protocols are
-expected to be less complex than the libraries generated by ChIP-Seq,
-since the possible number of positions to which the reads can be
-aligned is being reduced due to the exonuclease digestion,
-considerable amounts of reads are being mapped to specific
-positions. This affects the interpretation of the PBC, since for
-ChIP-Seq low PBC values indicate that the same read has been copied by
-the amplification process and aligned multiple times to the same
-position; while for ChIP-exo when several reads are aligned to the
-same position are not necessarily the same read amplified, but several
-reads that their $5\prime$ end was digested to the same position
-before the amplification step. It is of special importance to notice
-that for deeply sequenced ChIP-exo and ChIP-nexus experiments, the PBC
-values are quite low, which by following the ChIP-Seq guidelines it
-would indicate that those experiment show severe bottlenecking
-problems, which may incorrectly suggest that there is not genomic
-signal int those experiments.
-
-The Strand Cross-Correlation (SCC) introduced by Kharchenko et al.,
-2008 \cite{strandcc} is the most commonly used quality measure in
-ChIP-Seq. It is calculated as the correlation between both strand
-coverages, where each one is shifted $\delta / 2$ bp towards the
-$3\prime$ direction. In general it measures how well the reads mapped
-to each strand are clustered around the locations where the proteins
-are binding to the DNA, and usually it is expected to observed two
-local maxima, one when the profiles are shifted by the average read
-length and another when the profiles are shifted by the unobserved
-fragment length. In a good ChIP-Seq dataset the last one is also the
-SCC global maxima. However, in ChIP-exo's case these two peaks are
-confounded. Hence the Normalized Strand Cross-Correlation (NSC) which
-is a measure based on the SCC is harder to interpret. Figure
-\ref{fig:scc_exo} shows that both local maxima are hard to
-differentiate in the SCC curves for the $\sig$ ChIP-exo datasets used
-to calculate the NSC values in Table \ref{tab:qc}.
-
 \subsubsection{ChIP-exo Quality Control Pipeline}
@@
 \emph{bowtie} in its 1.1.2 version.
+
+\subsection*{Comparison with ChIP-Seq data}
+
+Using the $\sig$ sample grown under aerobic condition for ChIP-exo and
+SE ChIP-Seq. We partitioned the \emph{E.Coli} genome into
+non-overlaping bins of length 150, and for each bin we
+counted how many ChIP-exo and SE ChIP-Seq extended reads overlapped
+each bin. The reads were extended by 150 bp to their
+$3\prime$ directions.
+
+To model the strand imbalance, we called a set of high quality peaks
+for ChIP-exo and SE ChIP-Seq (with $\mbox{GC content + Mappability}$
+and $\mbox{Input only}$ models respectively), such that each peak
+contained at least one binding event (found by using \textbf{dPeak}
+with at most 5 binding events in each peak). For each peak, we
+calculated the FSR as:
+
+\begin{align*}
+    \mbox{FSR} &= \frac{\text{Nr. of fwd. strand reads mapped to the region}}                 {\text{Total nr. of reads mapped to the region}} 
+\end{align*}
+
+We estimated the $\mbox{FSR}$'s densities by using R's \emph{density}
+function using its default parameters, which computes gaussian kernel
+density estimates.
 

### 2016-05-03 rmarkdown/paper.R

@@ -1,9 +1,3 @@
-## ----comp_param , include = FALSE,echo = FALSE,eval = TRUE---------------
-
-  bin_size <- 150
-  frag_len <- 150
-
-
 ## ----table1,include = FALSE,echo = FALSE,eval = FALSE--------------------
 #    library(data.table)
 #    load("../data/paper_tables/Landick_rif_summary.RData")
@@ -91,6 +85,12 @@
 
   fdr <- 0.01
   topM <- 300
+
+
+## ----comp_param , include = FALSE,echo = FALSE,eval = TRUE---------------
+
+  bin_size <- 150
+  frag_len <- 150
 
 
 ## ----table_chipseq1,include = FALSE,echo = FALSE,eval = FALSE------------

### 2016-05-03 rmarkdown/paper.Rnw
added enrichment and library complexity section to paper, replaced respective figure with more comprehensive analysis

@@
 % $3\prime$ direction
+
+\subsubsection*{ChIP-exo Quality Control Pipeline}
+
+We first show the overall pipeline and consecutively discuss
+individual components with a case study using the FoxA1 factor in
+mouse liver cell lines generated by Serandour et al., 2013, and
+finally we show the evaluation of the ChIP-exo and ChIP-nexus samples
+with the QC pipeline. Figure \ref{fig:qcdiagram} shows a flowchart for
+the ChIP-exo QC pipeline. Given a sample of ChIP-exo or ChIP-nexus
+reads, in the first step, we partition the genome by keeping the
+non-digested ChIP-exo regions. Then, for each region it counts the
+number of fragments that compose the region and the number of
+positions to which the reads are being mapped to in each strand. With
+these values it calculates the following summary statistics:
+
+\begin{align*}
+  \text{ARC} &= \dfrac{\text{Nr. of reads in the region}}{\text{Width of the region}}, \\
+  \text{URCR} &= \dfrac{\text{Nr. of reads in the region mapped to
+      exactly one position}}{\text{Nr. of reads in the region}}, \\
+  \text{FSR} &= \dfrac{\text{Nr. of fwd. strand reads in the
+      region}}{\text{Number of reads in the region}}.
+\end{align*}
+
+Then it creates several visualizations designed to diagnose the
+quality level of a ChIP-exo sample. Figure \ref{fig:qcdiagram}A shows
+the typical behavior of the ARC vs. URCR plot. In general, the plot
+depicts two strong arms: One on the left with low ARC values and
+varying URCR which corresponds to ChIP-exo's background, regions that
+are usually composed by scattered reads that were not digested during
+the exonuclease step; and another one where the URCR decreases as the
+ARC increases, which corresponds to regions that are usually enriched
+and as the URCR decreases the library complexity does it as well, on
+the other hand high URCR values correspond to regions composed reads
+aligned to few positions. Finally we quantify we quantify the
+relationship between library complexity by the use of two indexes that
+represent the change in the number of unique positions per regions and
+the change in of the width of a regions as the depth changes. Figures
+\ref{fig:qcdiagram}B and \ref{fig:qcdiagram}C analyze the strand
+imbalance bias in a ChIP-exo experiment: The first one depicts how
+quickly the regions exclusively formed by fragments in one strand are
+being filtered out as regions with higher depth are observed, in a
+high quality sample the proportion of regions composed by only one
+strand is expected to decrease as higher depth regions are considered,
+while in a lower quality dataset this proportions remains
+approximately constant; and the second one shows how quickly the FSR's
+distribution approach the median, since in a high quality sample it is
+expected for the median to be approximately 0.5 and the enriched
+regions are going to be composed by fragments sequenced from both
+strands.
+
+% The ChIP-exo QC pipeline provides a model free framework to analyze
+% the biases in a ChIP-exo experiment by taking advantage of the
+% exonuclease enzyme that digests the non-enriched regions to partition
+% the genome, calculates common ChIP-Seq QC metrics in ChIP-exo regions
+% locally and allows the interpretation of these metrics by the use of
+% diagnostic plots.
+
+\textbf{Analysis of enrichment and library complexity in FoxA1
+  ChIP-exo data}
+
+In ChIP-exo experiments, background fragments are often digested by
+the exonuclease enzyme, therefore the balance between the enrichment
+and library complexity of an experiment is a key factor determining
+the sample's data quality.
+
+Using the Fox A1 in mouse liver cell lines generated by Serandour et
+al., 2013\nocite{exoillumina} and these two quantities, we explored
+the relationship between library complexity and experiment
+enrichment. In Figure \ref{fig:enrich}A we present ARC vs. URCR plots
+for all three replicates. The first and third replicate show a defined
+decreasing trend in the $\mbox{URCR}$ as the $\mbox{ARC}$ increases
+which indicates that this samples exhibit a higher enrichment than the
+second replicate. On the other hand, the overall $\mbox{URCR}$ level
+from the first two replicates is higher than the third replicate's
+level, which indicates that its libraries are more complex than the
+third replicate since three experiment are comparable in depth.
+
+Using the regions and summary statistics we obtained a set of
+candidate sites for each replicate, the total number of candidate
+sites is shown in Figure \ref{fig:enrich}B. For this sites, we
+extracted the sequence and searched for the FoxA1 motif using
+\textbf{FIMO} \cite{fimo}. Figure \ref{fig:enrich}C shows the fraction
+of candidate site where a motif was found. The first replicate
+outperforms the rest in both number of candidate sites and proportion
+of sites with motif; on the other hand we can see that the second
+replicate candidate sites are more likely to contain the motif
+sequence, but there more candidate sites in the third replicate than
+in the second. Finally, Figure \ref{fig:enrich}D shows the average
+profiles around the coordinate where the motif was found in the
+candidate sites: We can observe that the signal is more delimited for
+the first and third replicates than the second one, but overall the
+signal pattern is reproducible in the three ChIP-exo
+experiments. Figures \ref{fig:enrich}E, \ref{fig:enrich}F and
+\ref{fig:enrich}G show the purity of the motif found by visualizing
+the sequence to which the motif was matched. All these observations
+together shows the following findings: First, observing a defined
+decreasing trend in $\mbox{URCR}$ as $\mbox{ARC}$ increases implies
+than the sample is more enriched, therefore a higher amount of
+candidate sites. Second, for experiments that are not deeply
+sequenced, low $\mbox{URCR}$ values represent low complexity which
+decreases the amount of sites of motif found but it doesn't modify the
+read distribution around the motif. Finally, libraries with higher
+complexity returns sequences with a better motif match.
+
+% To verify this statement, we
+% extracted the sequences around high confidence binding events and look
+% for the FoxA1 motif using FIMO \cite{fimo}. Figure \ref{fig:enrich}B
+% shows the number of candidate regions, which shows that the first
+% replicate is being allocated into more enriched regions than the other
+% ones. Figure \ref{fig:enrich}C shows that for the first and third
+% replicates, the FoxA1 motif is being detected in roughly the same
+% proportion of sites, and finally in \ref{fig:enrich}D we observe that
+% the first and third replicate can detect the FoxA1 motif with the same
+% significance, while the second replicate does not.
 \color{blue}
@@
 \item \st{chip seq qc metrics}
-\item chip exo qc pipeline, description and justification of the steps
+\item \st{chip exo qc pipeline, description and justification of the steps}
 \begin{itemize}
@@
   of possible mappable positions.
-\item case study with FoxA1, analysis of enrichment vs library
+\item \st{case study with FoxA1, analysis of enrichment vs library
   complexity and analysis and the strand imbalance tools. in methods
@@
   works. how do we expect from the diagnostics and how to interpret
-  the plots.
+  the plots.}
 \item analysis of why the pipeline works. in methods need to specify
@@
 
-\subsubsection{ChIP-exo Quality Control Pipeline}
-
-Figure \ref{fig:qcdiagram} shows a flowchart for the ChIP-exo QC
-pipeline. In the first step, we partition the genome by keeping the
-non-digested ChIP-exo regions. Then, for each region it counts the
-number of fragments that compose the region and the number of
-positions to which the reads are being mapped to in each strand. With
-these values it calculates the following summary statistics:
-
-\begin{align*}
-  \text{ARC} &= \dfrac{\text{Nr. of reads in the region}}{\text{Width of the region}}, \\
-  \text{URCR} &= \dfrac{\text{Nr. of reads in the region mapped to
-      exactly one position}}{\text{Nr. of reads in the region}}, \\
-  \text{FSR} &= \dfrac{\text{Nr. of fwd. strand reads in the
-      region}}{\text{Number of reads in the region}}.
-\end{align*}
-
-Finally it creates several visualizations designed to diagnose the
-quality level of a ChIP-exo sample. Figure \ref{fig:qcdiagram}A shows
-the typical behavior of the ARC vs. URCR plot. In general, the plot
-depicts two strong arms: One on the left with low ARC values and
-varying URCR which corresponds to ChIP-exo's background, regions that
-are usually composed by scattered reads that were not digested during
-the exonuclease step; and another one where the URCR decreases as the
-ARC increases, which corresponds to regions that are usually enriched
-and as the URCR decreases the library complexity does it as well, on
-the other hand high URCR values correspond to regions composed by
-position with relatively few fragments aligned to. Figures
-\ref{fig:qcdiagram}B and \ref{fig:qcdiagram}C analyze the strand
-imbalance bias in a ChIP-exo experiment: The first one depicts how
-quickly the regions exclusively formed by fragments in one strand are
-being filtered out as regions with higher depth are observed; and the
-second one shows how quickly the FSR's distribution approach the
-median, since in a high quality sample it is expected for the median
-to be approximately 0.5 and the enriched regions are going to be
-composed by fragments sequenced from both strands.
-
-The ChIP-exo QC pipeline provides a model free framework to analyze
-the biases in a ChIP-exo experiment by taking advantage of the
-exonuclease enzyme that digests the non-enriched regions to partition
-the genome, calculates common ChIP-Seq QC metrics in ChIP-exo regions
-locally and allows the interpretation of these metrics by the use of
-diagnostic plots.
-
-\textbf{Enrichment and library complexity in ChIP-exo data}
-
-In ChIP-exo experiments, background fragments are often digested by
-the exonuclease enzyme, therefore the balance between the enrichment
-and library complexity of an experiment is a key factor determining
-the sample's data quality.
-
-Using the Fox A1 in mouse liver cell lines from \cite{exoillumina} and
-these two quantities, we explored the relationship between library
-complexity and experiment enrichment. In Figure \ref{fig:enrich}A we
-present ARC vs. URCR plots for all three replicates. As a case of
-study, we compare the three plots to differentiate the quality of the
-three experiment. Hence, this might imply that the first replicate to
-have have more enriched regions and the third replicate's library
-complexity to be lower than the other two replicates library
-complexities. To verify this statement, we extracted the sequences
-around high confidence binding events and look for the FoxA1 motif
-using FIMO \cite{fimo}. Figure \ref{fig:enrich}B shows the number of
-candidate regions, which shows that the first replicate is being
-allocated into more enriched regions than the other ones. Figure
-\ref{fig:enrich}C shows that for the first and third replicates, the
-FoxA1 motif is being detected in roughly the same proportion of sites,
-and finally in \ref{fig:enrich}D we observe that the first and third
-replicate can detect the FoxA1 motif with the same significance, while
-the second replicate does not.
-
-\textbf{Strand imbalance in ChIP-exo data}
+\textbf{Analysis of strand imbalance in FoxA1 ChIP-exo data}
 
@@
   \newline
-  \includegraphics[width = .3 \textwidth,page = 1]{../figs/FOXA1_mm9_fimo/FOXA1_summaries_by_replicate.pdf}
-  \includegraphics[width = .3 \textwidth,page = 3]{../figs/FOXA1_mm9_fimo/FOXA1_summaries_by_replicate.pdf}
-  \includegraphics[width = .3 \textwidth,page = 1]{../figs/FOXA1_mm9_fimo/FOXA1_log10pval_cumulative_nmotifs.pdf}
+  \includegraphics[width = .23 \textwidth, page =
+  1]{../figs/for_paper/FoxA1_barplots.pdf}
+  \includegraphics[width = .23 \textwidth, page =
+  3]{../figs/for_paper/FoxA1_barplots.pdf}
+  \includegraphics[width =
+  .4\textwidth]{../figs/for_paper/FoxA1_profiles_around_motif.pdf}
+  \includegraphics[width = .3\textwidth,page =
+  2]{../figs/for_paper/FoxA1_matched_motif_sequence.pdf}
+  \includegraphics[width = .3\textwidth,page =
+  3]{../figs/for_paper/FoxA1_matched_motif_sequence.pdf}
+  \includegraphics[width = .3\textwidth,page =
+  1]{../figs/for_paper/FoxA1_matched_motif_sequence.pdf}
+  % \includegraphics[width = .3 \textwidth,page = 1]{../figs/FOXA1_mm9_fimo/FOXA1_summaries_by_replicate.pdf}
+  % \includegraphics[width = .3 \textwidth,page = 3]{../figs/FOXA1_mm9_fimo/FOXA1_summaries_by_replicate.pdf}
+  % \includegraphics[width = .3 \textwidth,page = 1]{../figs/FOXA1_mm9_fimo/FOXA1_log10pval_cumulative_nmotifs.pdf}
+  % \caption{Using the mouse FoxA1 experiment from \cite{exoillumina}:
+  %   A) Hexbin plots of $\mbox{ARC}$ against $\mbox{URCR}$, there is a
+  %   slight separation into two strong arms, one corresponds to low
+  %   $\mbox{ARC}$ and varying $\mbox{URCR}$, and for the other
+  %   $\mbox{URCR}$ decreases as $\mbox{ARC}$ increases. B) Number of
+  %   candidate sites for each replicate. C) Percentage of candidate
+  %   sites where the FoxA1 motif was detected. D) Cumulative
+  %   proportion of detected motifs by replicate.}
   \caption{Using the mouse FoxA1 experiment from \cite{exoillumina}:
@@
     candidate sites for each replicate. C) Percentage of candidate
-    sites where the FoxA1 motif was detected. D) Cumulative
-    proportion of detected motifs by replicate.}
+    sites where the FoxA1 motif was detected. D) Average coverage
+    around FoxA1 motif. Base distribution for matched sequence for
+    Rep-1 (E), Rep-2 (F) and Rep-3 (G).}
   \label{fig:enrich}
 \end{figure}
+
+{\color{red} \emph{need to add labels for each figure and think about
+    order, in the supplement we have ecdf from fimo score or pval, and
+    overlaps with peaks called by mosaics}}
 

### 2016-05-09 rmarkdown/paper.Rnw
fixed version of paper

@@
 
-Using the regions and summary statistics we obtained a set of
+Considering the regions and summary statistics we obtained a set of
 candidate sites for each replicate, the total number of candidate
@@
 % significance, while the second replicate does not.
-\color{blue}
-
-outline
-
-\begin{enumerate}
-\item \st{Available data, the motivation of this part is two fold:}
- \begin{itemize}
- \item \st{we want to list all data sets we obtained}
- \item \st{we want to emphasize that for the sig70 datasets we have a
-   unique design that allows us to compare chip exo with SE and PE
-   chipseq , the later one for the first time}
- \end{itemize}
-\item \st{comparison with chip seq} %? with and without dpeak ?}
-\item \st{only in methods. list the align algorithm used}
-\item \st{chip seq qc metrics}
-\item \st{chip exo qc pipeline, description and justification of the steps}
-\begin{itemize}
-\item need to justify the sampling step. for this we may use the depth
-  table and add the factor to normalize one depth to another. explain
-  the confounding of pcr amplification and the reduction of the number
-  of possible mappable positions.
-\item \st{case study with FoxA1, analysis of enrichment vs library
-  complexity and analysis and the strand imbalance tools. in methods
-  we need to add which tests did we perform to show that the pipeline
-  works. how do we expect from the diagnostics and how to interpret
-  the plots.}
-\item analysis of why the pipeline works. in methods need to specify
-  how and in results need to specify why it works, and what are the
-  conclusions from it
-\end{itemize}
-\item high resolution analysis with chip exo and comparison with PE
-  chip seq
-  \begin{itemize}
-  \item justify what do we used dpeak (in methods need to list the
-    other methods used). perhaps we can add gof plots in
-    supplement. we also want to compare with the same method,
-    therefore we can use the fact that dpeak have a unified model for
-    PE and SE dpeak
-  \item after comparing with other methods we can compare with PE and
-    SE ChIP-seq, this comparison gives also resolution and sensitivity
-  \item compare samples under same budget of reads
-  \end{itemize} 
-\end{enumerate}
-
-\color{black}
 
 \textbf{Analysis of strand imbalance in FoxA1 ChIP-exo data}
+
+% Using the Fox A1 in mouse liver cell lines generated by Serandour et
+% al., 2013\nocite{exoillumina}
 
@@
 enriched regions usually are composed of a higher quantity of reads,
-therefore we examined the FSR (defined as the ratio of number of
-forward stranded reads divided by the total number of reads in a given
-region) as the regions with lower depth are being filtered out. This
-indicator is of particular importance, as it evaluates the ``peak
-pair'' assumption that the original ChIP-exo paper suggested and
-multiple ChIP-exo data analysis methods rely on. For every ChIP-exo
-experiment, we calculated the global FSR and noticed that for all
-experiments is roughly $0.5$, which means there are roughly the same
-amount of reads in both strands.
-
-In order to asses the strand imbalance we created the visualization
-shown in Figure \ref{fig:strand}: Figure \ref{fig:strand}A presents
-the FSR's behavior as the lower depth regions are being filtered out,
-while Figure \ref{fig:strand}B) shows which percentage of the regions
-are composed by reads in both strand or only one (forward or
-backward). In a good data set, it would be expected that all quantiles
-shown to be quickly converging towards the median (in panel A) or the
-regions composed of reads in one strand being made of few fragments
-(in panel B). For each replicate, we divided the partitioned regions
-by asking whether they overlap a set of high quality ChIP-exo peaks,
-and then we tested (using the Wilcoxon rank sum test over the
-imbalance index defined in the methods section) if the strand
-imbalance's distribution is the same for both classes. For regions
-composed by a higher amount of reads, it is harder to distinguish
-their peaks by considering only the strand imbalance, hence in a
-better quality ChIP-exo experiment it is easier to distinguish
-enriched regions by the amount of reads in both strands. Similarly, we
-may consider that the strands of reads for background of a ChIP-exo
-experiment is more unbalanced than those for the enriched regions. In
-conclusion, Figure \ref{fig:strand} shows that the global FSR does not
-represent the experiment's local strand imbalance, hence the ``peak
-pair'' assumption does not hold locally in ChIP-exo data.
-
-\textbf{Analysis of ChIP-nexus data with the QC pipeline}
-
-We applied the QC pipeline to the ChIP-nexus experiments and found
-that the overall pattern seen in the visualizations is similar to a
-high quality ChIP-exo experiment. Figure \ref{fig:nexus}A illustrates
-a typical pattern of the ARC vs. URCR plot with ChIP-nexus data: Both
-arms seems to be more narrow and therefore the separation between them
-seems to be more noticeable, this seems reasonable since libraries
-generated by ChIP-nexus usually yields higher complexity than
-libraries generated by ChIP-exo. Additionally, while in ChIP-exo
-experiments is common to observe regions formed by undigested
-fragments in only one strand, we can observe that in ChIP-nexus
-experiment the occurrence of this type of regions is reduced. Figure
-\ref{fig:nexus}B shows that in a ChIP-nexus experiment the FSR
-distribution approximates the median more quickly than in a ChIP-exo
-experiment and Figure \ref{fig:nexus}C we observe that most of these
-single-stranded regions are formed by few fragments, which means that
-good ChIP-nexus experiment are less likely to show very imbalanced
-regions.
-
-\subsubsection{Evaluating ChIP-exo and ChIP-nexus data}
-
-<<all_param,include = FALSE,echo = FALSE,eval = TRUE>>=
-
-  Ntimes <- prettyNum(1e4,big.mark = ",")
-  Msamp <- prettyNum(1e3,big.mark = ",")
-  thresh1 <- 10
-
-@
-
-We applied the QC pipeline to all ChIP-exo and ChIP-nexus datasets
-listed in Tables \ref{tab:qc} and \ref{tab:nexus}, that way we
-partitioned the reads of each experiment into the undigested regions
-by keeping the regions formed by at least one fragment. Figure
-\ref{fig:enrich_all}A shows compares the slope of the following model:
-
-\begin{align}
-  \mbox{depth} = \beta \times \text{(\# of unique positions)} +\epsilon  \nonumber
-\end{align}
-
-when sampling \Sexpr{Msamp} regions and repeating this process
-\Sexpr{Ntimes} times for every ChIP-exo and ChIP-nexus experiment. The
-slope can be interpreted as the change in depth as number of unique
-position varies. This parameter measures the library complexity on a
-ChIP-exo experiment, since when it is high it means that several
-fragments are being aligned to the same position in the genome.
-
-Figures \ref{fig:enrich_all}B and \ref{fig:enrich_all}C shows the
-parameters of the following model:
-
-\begin{align}
-  \mbox{URCR} = \frac{\kappa}{\mbox{ARC}} + \eta + \epsilon \nonumber
-\end{align}
-
-considering only the regions composed by 10 or more aligned reads,
-that way the model reflects the enriched arm of the ARC vs. URCR
-visualization. The intercept $\eta$ can be interpreted as the
-theoretical $\mbox{URCR}$ for relatively high depth regions and the
-slope $\kappa$ can be interpreted as the decay in $\mbox{URCR}$ as the
-$\mbox{ARC}$ increases. That way experiments with low complexity are
-showing low values for both parameters. We considered a linear model
-between the estimated parameters for each experiment and it's
-respective depth for which we were able to asses that there is not
-significant relationship between the estimated parameters and the
-interaction between depth and organism for each experiment. Hence,
-these parameters may be used as indicators of the overall quality of
-ChIP-exo experiment, but as with the SCC it is recommendable to also
-visualize the SCC curve.
-
-\subsection{High Resolution Binding Site Identification with dPeak and
+therefore we examined the FSR as the regions with lower depth are
+being filtered out. This indicator is of particular importance, as it
+evaluates the ``peak pair'' assumption that the original ChIP-exo
+paper suggested and multiple ChIP-exo data analysis methods rely
+on. For every ChIP-exo experiment, we calculated the global FSR and
+noticed that for all experiments is roughly $0.5$, which means there
+are approximately the same amount of reads in both strands.
+
+In order to study the strand imbalance of the FoxA1 factor ChIP-exo
+experiments in mouse liver generated by Serandour et al.,
+2013\nocite{exoillumina} we considered the visualizations shown in
+Figure \ref{fig:strand}: Figure \ref{fig:strand}A presents the FSR's
+behavior as the lower depth regions are being filtered out, while
+Figure \ref{fig:strand}B) shows the proportion of regions that are
+composed by reads from both strands against the regions formed by
+reads in exclusively one strand. In a high quality dataset, it is
+expected for all quantiles to quickly converging towards the median
+(in panel A) or the regions formed by reads in one strand (either
+forward or backward) to be formed by few fragments (in panel
+B).
+
+Additionally, it is worth mentioning that the background in a ChIP-exo
+experiment is expected to be formed by the undigested ChIP fragments
+that are not bound to a protein, hence it is for the reads in both
+strand to be imbalanced in a background region. Figure
+\ref{fig:strand}C compares the strand imbalance when ChIP-exo regions
+overlap with high quality peaks. It is noticeable that for regions
+composed by a large amount of reads, it is harder to distinguish their
+peaks by considering only the strand imbalance, hence in a high
+quality ChIP-exo experiment the background is expected to show a
+higher imbalance that the enriched regions. In conclusion, Figure
+\ref{fig:strand} shows that the global FSR does not accurately
+represent a ChIP-exo experiment's strand imbalance locally, hence the
+``peak pair'' assumption does not completely hold in every ChIP-exo
+enriched region.
+
+% \textbf{Analysis of ChIP-nexus data with the QC pipeline}
+
+% We applied the QC pipeline to the ChIP-nexus experiments and found
+% that the overall pattern seen in the visualizations is similar to a
+% high quality ChIP-exo experiment. Figure \ref{fig:nexus}A illustrates
+% a typical pattern of the ARC vs. URCR plot with ChIP-nexus data: Both
+% arms seems to be more narrow and therefore the separation between them
+% seems to be more noticeable, this seems reasonable since libraries
+% generated by ChIP-nexus usually yields higher complexity than
+% libraries generated by ChIP-exo. Additionally, while in ChIP-exo
+% experiments is common to observe regions formed by undigested
+% fragments in only one strand, we can observe that in ChIP-nexus
+% experiment the occurrence of this type of regions is reduced. Figure
+% \ref{fig:nexus}B shows that in a ChIP-nexus experiment the FSR
+% distribution approximates the median more quickly than in a ChIP-exo
+% experiment and Figure \ref{fig:nexus}C we observe that most of these
+% single-stranded regions are formed by few fragments, which means that
+% good ChIP-nexus experiment are less likely to show very imbalanced
+% regions.
+
+\subsubsection*{Evaluation of ChIP-exo and ChIP-nexus data with QC
+  pipeline}
+
+We evaluated every experiment listed in Tables \ref{tab:qc_sig} and
+\ref{tab:qc} with the QC pipeline. For each experiment, we explored
+the relationship between depth, width and the number of unique
+positions to summarize the shape obtained in the $\mbox{ARC}$
+vs. $\mbox{URCR}$ visualization. Figure \ref{fig:eval}A shows the
+overall change in depth as the number of unique positions varies for
+the gathered experiment in eukaryotic genomes. This parameter can be
+interpreted as the inverse of $\gamma$ in equation (\ref{mod}), which
+can is also the experiment's large depth $\mbox{URCR}$. In a high
+quality ChIP-exo experiment, it is expected in the $\mbox{ARC}$
+vs. $\mbox{URCR}$ plot to observe two arms, one with low $\mbox{ARC}$
+and varying $\mbox{URCR}$ and another that with a decreasing
+$\mbox{URCR}$ as $\mbox{ARC}$ increases that stabilizes in
+$\gamma$. When the ChIP-exo experiment is not deeply sequenced, high
+values in Figure \ref{fig:eval} indicate that the library complexity
+from the experiment is low. On the other hand, low values corresponds
+to higher quality ChIP-exo experiment.
+
+Figure \ref{fig:eval}B shows the average read coverage bias when the
+experiment's sequencing depth is large/ Under perfect digestion
+setting, most of the reads aligned to binding regions are going to be
+accumulated around a binding event, which suggest to be unlikely to
+observe as the sequencing depth increases to observe a reads being
+aligned to another position that was not previously covered. Low
+quality ChIP-exo experiment exhibit more scattered reads in both
+strands across the genome; therefore it is more likely to observe
+consider reads that align to position in the genome that were not
+previously covered. We can compare the second replicate from the human
+ChIP-nexus experiment against both second and third replicates from
+the TBP factor experiment in K562 cell lines. The sequencing depth of
+this three experiment is comparable but the large depth $\mbox{ARC}$'s
+bias is considerably higher for both ChIP-exo experiments.
+
+The interaction between these two parameters represents the quality of
+a ChIP-exo and ChIP-nexus experiment. When either the adjusted
+$\mbox{ARC}$ or the $\mbox{ARC}$'s bias is large, then it is suggested
+to sub-sample reads from the experiment and then evaluate the
+sub-sampled experiment with the QC pipeline. Considering the TBP
+factor ChIP-exo \cite{venters13} and ChIP-nexus \cite{chipnexus}
+experiment in human K562 cell lines, we sub-sampled 20M to 50M reads
+to cover the sequencing depth's range from all the human experiments
+that were considered in this study and evaluated them with QC
+pipeline. Figure \ref{fig:eval}C exhibits an increasing trend in the
+high-depth adjusted $\mbox{ARC}$ in every panel, where the ChIP-nexus
+trend being considerable lower than both ChIP-exo replicates. On the
+other hand, in Figure \ref{fig:eval}D the bias remains approximately
+constant in ChIP-nexus sub-samples, while it increases for both
+ChIP-exo replicates. This suggest that the ChIP-exo replicates have
+low library complexity and overall lower quality than the ChIP-nexus
+experiment, regardless to the fact that all three experiment are
+deeply sequenced with more than 100 million reads each.
+
+\subsection*{High Resolution Binding Site Identification with dPeak and
   ChIP-exo}
 
-\subsubsection{Comparison with ChIP-Seq data using dPeak}
-
-<<comp_param,include = FALSE,echo = FALSE, eval = TRUE>>=
-
-  id_dist <- 20
-  ext <- 15
-
-@
-
-Figure \ref{fig:reso_all} shows comparisons among ChIP-exo, PE
-ChIP-Seq and SE ChIP-Seq. We considered the RegulonDB data as ground
-truth, since those are the most recent annotation on \emph{Escherichia
-  Coli}. A RegulonDB annotation (Salgado et al, 2012 \cite{regulondb})
-was considered to be identified if the distance from the closest dPeak
-binding site estimate was less that or equal to $\Sexpr{id_dist}$
-bp. That way, the sensitivity is defined as the proportion of
-RegulonDB annotations identified in a peak and the resolution is
-defined as the minimum distance between a RegulonDB annotation and the
-closest dPeak binding site estimate. Figure \ref{fig:reso_all}A shows
-that the sensitivity increases as the mean distance between binding
-events increases. When the average distance is greater than 200 bp,
-dPeak identifies more than the 75\% of the binding events in each
-peak, this is intuitive as the mean ChIP-Seq's fragment length is
-shorter than 200 bp, hence the read does not contribute to more than
-one binding event \cite{dpeak}. When the binding events in a peak are
-closer to each other, both ChIP-exo and PE ChIP-Seq are comparable, as
-the distance increases ChIP-exo identifies a higher proportion of the
-RegulonDB annotations; additionally SE ChIP-Seq is significantly less
-sensitive than both ChIP-exo and PE ChIP-Seq. Figure
-\ref{fig:reso_all}B shows that ChIP-exo and PE ChIP-Seq are comparable
-in resolution, while both protocols significantly outperform SE
-ChIP-Seq.
-
-\subsubsection{Systematic comparison of ChIP-Seq vs ChIP-exo under
+\subsubsection*{dPeak outperforms competing methods in discovering closely
+  spaced binding events from ChIP-exo and ChIP-Seq data}
+
+After we determined which biases could affect a ChIP-exo experiment's
+quality we decided to proceed and examine binding sites in high
+resolution on high quality ChIP-exo experiments. We examined first
+both aerobic replicates from the \emph{E.Coli} first biosample.
+
+Figures \ref{fig:methods_comp}A and \ref{fig:methods_comp}B compares
+binding sites calculated with Peakzilla \cite{peakzilla}, MACE
+\cite{mace}, GEM \cite{gem} and dPeak \cite{dpeak} in resolution,
+where the resolution is defined as the distance from a RegulonDB
+annotation to its closest prediction. For the first replicate (Figure
+\ref{fig:methods_comp}A), the resolution calculated with all methods
+is comparable, while dPeak outperforms the rest in resolution when
+using the second replicate as seen in Figure \ref{fig:methods_comp}B.
+
+Additionally, Figures \ref{fig:methods_comp}C and
+\ref{fig:methods_comp}D compares the densities of the parameters
+estimated by dPeak for ChIP-exo and SE ChIP-Seq data. As expected, the
+$\delta$ parameter which measures the average distance from the
+$5\prime$ ends to their respective binding events and the $\sigma$
+parameter which is a measure from the reads dispersion around their
+respective binding events are lower for ChIP-exo than for SE
+ChIP-Seq. Therefore, we can asses that dPeak is accurately
+representing that the ChIP-exo fragments are allocated more tightly
+around the transcription factor binding sites.
+
+% \color{blue}
+% outline
+
+% \begin{enumerate}
+% \item \st{Available data, the motivation of this part is two fold:}
+%  \begin{itemize}
+%  \item \st{we want to list all data sets we obtained}
+%  \item \st{we want to emphasize that for the sig70 datasets we have a
+%    unique design that allows us to compare chip exo with SE and PE
+%    chipseq , the later one for the first time}
+%  \end{itemize}
+% \item \st{comparison with chip seq} %? with and without dpeak ?}
+% \item \st{only in methods. list the align algorithm used}
+% \item \st{chip seq qc metrics}
+% \item \st{chip exo qc pipeline, description and justification of the steps}
+% \begin{itemize}
+% \item \st{need to justify the sampling step. for this we may use the depth
+%   table and add the factor to normalize one depth to another. explain
+%   the confounding of pcr amplification and the reduction of the number
+%   of possible mappable positions.}
+% \item \st{case study with FoxA1, analysis of enrichment vs library
+%   complexity and analysis and the strand imbalance tools. in methods
+%   we need to add which tests did we perform to show that the pipeline
+%   works. how do we expect from the diagnostics and how to interpret
+%   the plots.}
+% \item \st{analysis of why the pipeline works. in methods need to specify
+%   how and in results need to specify why it works, and what are the
+%   conclusions from it}
+% \end{itemize}
+% \item high resolution analysis with chip exo and comparison with PE
+%   chip seq
+%   \begin{itemize}
+%   \item justify what do we used dpeak (in methods need to list the
+%     other methods used). perhaps we can add gof plots in
+%     supplement. we also want to compare with the same method,
+%     therefore we can use the fact that dpeak have a unified model for
+%     PE and SE dpeak
+%   \item after comparing with other methods we can compare with PE and
+%     SE ChIP-seq, this comparison gives also resolution and sensitivity
+%   \item compare samples under same budget of reads
+%   \end{itemize} 
+% \end{enumerate}
+
+% \color{black}
+% Figure \ref{fig:methods_comp} compares the resolution defined as the
+% minimum distance between a RegulonDB annotation and a binding site
+% predicted by either Peakzilla \cite{peakzilla}, MACE \cite{mace}, GEM
+% \cite{gem} or dPeak \cite{dpeak}. In a good dataset such as both of
+% the ChIP-exo experiments under aerobic (panel 931 and 933) conditions,
+% all the methods are comparable in resolution, and dPeak slightly
+% outperforms the rest. On the other hand, for experiments dPeak
+% identified binding sites with higher resolution than Mace and
+% Peakzilla, but with low resolution than Gem. This may be due that the
+% fact that Gem uses sequence information in addition to the aligned
+% $5\prime$ end counts that dPeak uses.
+
+\subsubsection*{Systematic comparison of ChIP-Seq vs ChIP-exo under
   varying sequencing depth}
@@
 ($N$ reads for both ChIP-exo experiment and $N/2$ or $N$ pairs for PE
-ChIP-Seq to assume more realistic situation of a fixed cost). For each
-sampled dataset we applied our lower-to-higher resolution pipeline by
-calling peaks with MOSAiCS \cite{mosaics} and then deconvolving the
-binding events by using dPeak \cite{dpeak}. For the ChIP-exo datasets
-we called peaks by using the GC-content and mappability models with
-MOSAiCS, since it's background is usually composed of scattered reads
-across the genome; and for the ChIP-Seq datasets we used their
-respective Input samples. Additionally, it is worth noting that for PE
-ChIP-Seq we sampled both ends of the fragment, hence for each
-sequencing depth we are sampling the half amount of pairs for PE
-ChIP-Seq than for ChIP-exo or SE ChIP-Seq.
-
-Figures \ref{fig:design} shows the behavior of each data type in
+ChIP-Seq to assume more realistic situation of a fixed
+cost). Additionally, it is worth noticing that for PE ChIP-Seq we
+sampled both ends of the fragment, hence for each sequencing depth we
+are sampling the half amount of pairs for PE ChIP-Seq than for
+ChIP-exo or SE ChIP-Seq.
+
+% For each sampled dataset we applied our lower-to-higher resolution
+% pipeline by calling peaks with MOSAiCS \cite{mosaics} and then
+% deconvolving the binding events by using dPeak \cite{dpeak}. For the
+% ChIP-exo experiments we called peaks by using the GC-content and
+% mappability models with MOSAiCS, since it's background is usually
+% composed of scattered reads across the genome; and for the ChIP-Seq
+% datasets we used their respective Input samples.
+
+Figure \ref{fig:design} illustrates the behavior of each data type in
 $\sig$ experiment under aerobic condition when comparable number of
@@
 
-Figure \ref{fig:saturation_rif} shows an analogous analysis but using
-the $\sig$ replicates with and without rif treatment. The left, middle
-and right columns show the fixed depth against the number of predicted
-events, identified targets and resolution being compared at a fixed
-depth level. The behavior of this quantities seems to be opposite to
-the one as in Figure \ref{fig:design}, hence we used the ChIP-exo QC
-pipeline in the fixed depth ChIP-exo experiments. Figure
-\ref{fig:exoQC_sat_aero} shows scatter plots of ARC vs URCR for
-several fixed sample sizes, as the fixed depth increases the two arm
-pattern becomes more distinctive while for lower depth, it seems that
-the majority of the sampled reads were aligned to enriched regions. On
-the other hand in Figures \ref{fig:exoQC_sat1} to
-\ref{fig:exoQC_sat4}, we used the ChIP-exo QC pipeline on the samples
-that are outperformed by PE and SE ChIP-Seq. For a fixed low depth, we
-can see that the majority of the reads are being aligned to
-non-enriched regions since the vertical arm seems to be stronger for
-all 2 conditions and replicates; for higher depth we can see low URCR
-values being predominant, which indicates that the majority of the
-regions being formed by few positions with a higher read
-concentration. In low complexity regions, the reads are being aligned
-to fewer positions but there is no control over the amount of reads
-mapped. Hence, those regions are more likely to being strand-imbalance
-which in turn may bias the binding site estimate and therefore
-decrease the number of identified targets or increase an experiments
-resolution.
-
+% Figure \ref{fig:saturation_rif} shows an analogous analysis but using
+% the $\sig$ replicates with and without rif treatment. The left, middle
+% and right columns show the fixed depth against the number of predicted
+% events, identified targets and resolution being compared at a fixed
+% depth level. The behavior of this quantities seems to be opposite to
+% the one as in Figure \ref{fig:design}, hence we used the ChIP-exo QC
+% pipeline in the fixed depth ChIP-exo experiments. Figure
+% \ref{fig:exoQC_sat_aero} shows scatter plots of ARC vs URCR for
+% several fixed sample sizes, as the fixed depth increases the two arm
+% pattern becomes more distinctive while for lower depth, it seems that
+% the majority of the sampled reads were aligned to enriched regions. On
+% the other hand in Figures \ref{fig:exoQC_sat1} to
+% \ref{fig:exoQC_sat4}, we used the ChIP-exo QC pipeline on the samples
+% that are outperformed by PE and SE ChIP-Seq. For a fixed low depth, we
+% can see that the majority of the reads are being aligned to
+% non-enriched regions since the vertical arm seems to be stronger for
+% all 2 conditions and replicates; for higher depth we can see low URCR
+% values being predominant, which indicates that the majority of the
+% regions being formed by few positions with a higher read
+% concentration. In low complexity regions, the reads are being aligned
+% to fewer positions but there is no control over the amount of reads
+% mapped. Hence, those regions are more likely to being strand-imbalance
+% which in turn may bias the binding site estimate and therefore
+% decrease the number of identified targets or increase an experiments
+% resolution.
+
+\subsubsection*{Comparison with ChIP-Seq data using dPeak}
+
+<<comp_param,include = FALSE,echo = FALSE, eval = TRUE>>=
+
+  id_dist <- 20
+  ext <- 15
+
+@
+
+Figure \ref{fig:reso_all} shows comparisons among ChIP-exo, PE
+ChIP-Seq and SE ChIP-Seq. We considered the RegulonDB data as ground
+truth, since those are the most recent annotation on \emph{E. Coli}. A
+RegulonDB annotation (Salgado et al, 2012\nocite{regulondb}) was
+considered to be identified if the distance from the closest dPeak
+binding site estimate was less that or equal to $\Sexpr{id_dist}$
+bp. That way, the sensitivity is defined as the proportion of
+RegulonDB annotations identified in a peak and the resolution is
+defined as the minimum distance between a RegulonDB annotation and the
+closest dPeak binding site estimate. Figure \ref{fig:reso_all}A shows
+that the sensitivity increases as the mean distance between binding
+events increases. When the binding events in a peak are closer to each
+other, both ChIP-exo and PE ChIP-Seq are comparable, as the distance
+increases ChIP-exo identifies a higher proportion of the RegulonDB
+annotations; additionally SE ChIP-Seq is significantly less sensitive
+than both ChIP-exo and PE ChIP-Seq. Figure \ref{fig:reso_all}B shows
+that ChIP-exo and PE ChIP-Seq are comparable in resolution, while both
+protocols significantly outperform SE ChIP-Seq.
+
+% When the average distance is greater than 200 bp,
+% dPeak identifies more than the 75\% of the binding events in each
+% peak, this is intuitive as the mean ChIP-Seq's fragment length is
+% shorter than 200 bp, hence the read does not contribute to more than
+% one binding event \cite{dpeak}.
 % In supplement figures \ref{fig:exoQC_sat1} to \ref{fig:exoQC_sat4}, we
@@
 % conditions.
-
-\subsubsection{dPeak outperforms competing methods in discovering closely
-  spaced binding events from ChIP-exo and ChIP-Seq data}
-
-Figure \ref{fig:methods_comp} compares the resolution defined as the
-minimum distance between a RegulonDB annotation and a binding site
-predicted by either Peakzilla \cite{peakzilla}, MACE \cite{mace}, GEM
-\cite{gem} or dPeak \cite{dpeak}. In a good dataset such as both of
-the ChIP-exo experiments under aerobic (panel 931 and 933) conditions,
-all the methods are comparable in resolution, and dPeak slightly
-outperforms the rest. On the other hand, for experiments dPeak
-identified binding sites with higher resolution than Mace and
-Peakzilla, but with low resolution than Gem. This may be due that the
-fact that Gem uses sequence information in addition to the aligned
-$5\prime$ end counts that dPeak uses. 
- \footnote{For here we may
-  probably use only 933 as part of the main article and keep the rest
-  for the supplement}
 
@@
   topM <- 300
+  Ntimes <- prettyNum(1e4,big.mark = ",")
+  Msamp <- prettyNum(1e3,big.mark = ",")
+  thresh1 <- 10
 
@@
 \url{https://github.com/welch16/ChIPexoqual}.
+
+\textbf{Interpretation of the linear model in the QC pipeline}
+
+The linear model:
+
+\begin{align*}
+  \mbox{depth} = \beta_1 \mbox{npos} + \beta_2 \mbox{width}
+\end{align*}
+
+is a re-parametrization of the following relationship:
+
+\begin{align}
+  \mbox{URCR} = \frac{\kappa}{\mbox{ARC}} + \gamma
+\label{mod}
+\end{align}
+
+with $\beta_1 = 1 / \gamma$ and $\beta_2 = - \kappa / \gamma$. In this
+setting, $\gamma$ can be considered as the large - depth
+$\mbox{URCR}$, i.e. the ratio between the number of unique positions
+and depth as the depth limits to infinity.
+
+On the other hand, to interpret $\beta_2$, if we assume that the
+$\gamma$ value is already given, then by expressing $\kappa$ as a
+function of $\mbox{ARC}$ and $\mbox{URCR}$ we can observe the
+following identities:
+
+\begin{align*}
+  \kappa &= \frac{\mbox{npos}}{\mbox{width}} - \gamma \mbox{ARC} \\
+  \frac{\kappa}{\gamma} &= \frac{1}{\gamma} \frac{\mbox{npos}}{\mbox{width}} - \mbox{ARC} 
+\end{align*}
+
+where $\mbox{npos}$ and $\mbox{width}$ denote the region's number of
+unique position where the reads are aligned to and the region's width
+respectively. This is particularly important, $\gamma$ approximates
+the true $\mbox{URCR}$ as the sequencing depth increases, which
+implies that $- \kappa / \gamma$ can be interpreted as the large
+sequencing depth bias of the $\mbox{ARC}$ since as the depth
+increases, the first term of $\kappa / \gamma$ is going to approximate
+the average read coverage:
+
+\begin{align*}
+  \frac{\mbox{npos}}{\mbox{width}} \times \lim_{d \rightarrow
+    \infty} \frac{d }{\mbox{npos}} \sim \mbox{ARC}
+\end{align*}
+
+Hence, we define the $\mbox{ARC}$ bias as $- \kappa / \gamma$.
 
@@
   1]{../figs/for_paper/FoxA1_matched_motif_sequence.pdf}
-  % \includegraphics[width = .3 \textwidth,page = 1]{../figs/FOXA1_mm9_fimo/FOXA1_summaries_by_replicate.pdf}
-  % \includegraphics[width = .3 \textwidth,page = 3]{../figs/FOXA1_mm9_fimo/FOXA1_summaries_by_replicate.pdf}
-  % \includegraphics[width = .3 \textwidth,page = 1]{../figs/FOXA1_mm9_fimo/FOXA1_log10pval_cumulative_nmotifs.pdf}
-  % \caption{Using the mouse FoxA1 experiment from \cite{exoillumina}:
-  %   A) Hexbin plots of $\mbox{ARC}$ against $\mbox{URCR}$, there is a
-  %   slight separation into two strong arms, one corresponds to low
-  %   $\mbox{ARC}$ and varying $\mbox{URCR}$, and for the other
-  %   $\mbox{URCR}$ decreases as $\mbox{ARC}$ increases. B) Number of
-  %   candidate sites for each replicate. C) Percentage of candidate
-  %   sites where the FoxA1 motif was detected. D) Cumulative
-  %   proportion of detected motifs by replicate.}
   \caption{Using the mouse FoxA1 experiment from \cite{exoillumina}:
@@
     order, in the supplement we have ecdf from fimo score or pval, and
-    overlaps with peaks called by mosaics}}
+    overlaps with peaks called by mosaics.
+\newline
+Also, need to change labels from repk to Rep-k
+}}
 
@@
   \includegraphics[width = .4\textwidth]{../figs/for_paper/Carroll_FSR_depth_VS_pvalWilcoxon.pdf}
-  \caption{Strand imbalance QC plots for the same data as in Figure
-    \ref{fig:enrich}A. A) FSR distribution quantiles as the lower
+  \caption{Strand imbalance using the mouse FoxA1 experiment from
+    \cite{exoillumina}: A) FSR distribution quantiles as the lower
     depth regions are being filtered out, all quantiles approach to
@@
   \centering
-  \includegraphics[width = .8\textwidth]{../figs/ChIPnexus/ChIPnexus_S2_MyC_enrichment.pdf}
+  \includegraphics[width = .9  \textwidth,page = 1]{../figs/for_paper/QC_pipeline_eval_boxplot.pdf}
 \newline
-  \includegraphics[width = .4\textwidth]{../figs/ChIPnexus/ChIPnexus_S2_MyC_fsr_surv.pdf}  
-  \includegraphics[width = .4\textwidth]{../figs/ChIPnexus/ChIPnexus_S2_MyC_label_surv.pdf}  
-  % \caption{\textbf{ChIP-exo QC pipeline on ChIP-nexus data.} A)
-  %   Enrichment vs. library complexity, B) Strand imbalance and C)
-  %   Single-strand region analysis for the MyC experiment in S2 cell
-  %   lines in \emph{D. Melanogaster} from He et al., 2015
-  %   \cite{chipnexus}}
-  \caption{We applied the QC pipeline on ChIP-nexus data. A) ARC
-    vs. URCR hexbin plots, B) FSR distribution quantiles and C)
-    Stacked histogram with the proportion of regions composed by
-    fragments on a single strand for the MyC experiment in S2 cell
-    lines in \emph{D. Melanogaster} from He et al., 2015
-    \cite{chipnexus}}
-  \label{fig:nexus}
+  \includegraphics[width = .9  \textwidth,page = 2]{../figs/for_paper/QC_pipeline_eval_boxplot.pdf}
+\newline
+  \includegraphics[width = .45  \textwidth,page = 1]{../figs/for_paper/K562_TBP_sample.pdf}
+  \includegraphics[width = .45  \textwidth,page = 2]{../figs/for_paper/K562_TBP_sample.pdf}
+  \caption{Comparison of adjusted Average Read Coverage and
+    $\mbox{ARC}$ bias for eukaryotic genome ChIP-exo and ChIP-nexus
+    experiment. A) Boxplot of adjusted $\mbox{ARC}$. B) Boxplot of
+    large sequencing depth $\mbox{ARC}$'s bias. Comparison of
+    sub-sampled ChIP-exo and ChIP-nexus experiments of TBP factor in
+    human K562 cell lines: C) Adjusted $\mbox{ARC}$ and D) $\mbox{ARC}$'s
+    bias.}
+  \label{fig:eval}
 \end{figure}
+
+  % \includegraphics[width = .9  \textwidth,page = 1]{../figs/test_quantify/arc_urcr_param_regression.pdf}
+  % \includegraphics[width = .9  \textwidth,page = 2]{../figs/test_quantify/arc_urcr_param_regression.pdf}
+  % \caption{A) Comparison of the contribution of the number of unique
+  %   positions to depth among all ChIP-exo and ChIP-nexus datasets
+  %   separated by organism. Comparison of the contributions of number
+  %   of unique positions (B) and width (C) to depth for sub-samples of
+  %   20M to 50M reads of TBP factor in human K562 cell lines ChIP-exo
+  %   and ChIP-nexus experiments.}
 
@@
   \centering
-  \includegraphics[width = .9  \textwidth,page = 1]{../figs/test_quantify/slope_npos_depth_boxplot.pdf}
-  \includegraphics[width = .9  \textwidth,page = 1]{../figs/test_quantify/arc_urcr_param_regression.pdf}
-  \includegraphics[width = .9  \textwidth,page = 2]{../figs/test_quantify/arc_urcr_param_regression.pdf}
-  \caption{A) Comparison of change in depth as the number of unique
-    position obtained by sampling \Sexpr{Msamp} regions and fitting
-    the model $\mbox{depth} = \beta \times \text{(\# unique
-      positions)}$, and repeating this process \Sexpr{Ntimes} times. B)
-    Intercept and C) Slope in the the $\mbox{URCR} = k / \mbox{ARC} +
-    b$ model for the regions formed by more than 10 aligned
-    positions.}
-  \label{fig:enrich_all}
+  \includegraphics[width = .45\textwidth,page = 1]{../figs/for_paper/methods_comp_FDR5_topM400.pdf}
+  \includegraphics[width = .45\textwidth,page = 2]{../figs/for_paper/methods_comp_FDR5_topM400.pdf}
+   \includegraphics[width = .45\textwidth,page = 1]{../figs/for_paper/sigma_delta_old_densities.pdf}
+   \includegraphics[width = .45\textwidth,page = 2]{../figs/for_paper/sigma_delta_old_densities.pdf}   
+  \caption{Comparison of the resolution between dPeak, Gem, Mace and
+    Peakzilla methods with \emph{E.Coli} aerobic experiment from the
+    first biosample A) Replicate 1 and B) Replicate 2. Resolution is
+    defined as the minimum distance between a RegulonDB annotation and
+    a predicted binding event. C) $\delta$ parameter in dPeak measures
+    average distance of the reads to their respective binding site. In
+    ChIP-exo data, reads were located much closer to the binding site
+    than in SET ChIP-Seq. D) $\sigma$ parameter measure the dispersion
+    of reads around each binding site. In ChIP-exo data, reads showed
+    less variation around the their respective binding sites compared
+    to SET ChIP-Seq.}
+  \label{fig:methods_comp}
 \end{figure}
+
+\newpage
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width = .9\textwidth]{../figs/for_paper/saturation_analysis_old.pdf}
+  \caption{Comparison of the number of A) candidate regions, B)
+    predicted events, C) identified targets and D) resolution among
+    ChIP-exo, PE ChIP-Seq and SE ChIP-Seq. RegulonDB annotations are
+    considered as a gold standard. A gold standard binding events was
+    deemed identified if a binding event was estimated at a $\pm$
+    \Sexpr{ext} vicinity of it.}
+  \label{fig:design}
+\end{figure}
+
+% \textbf{ChIP-exo, PE ChIP-Seq and SE ChIP-Seq comparison at varying
+%   sequencing depths.} 
+%  \newpage
+
+% \begin{figure}[h!]
+%   \centering
+% \includegraphics[width = .9\textwidth]{../figs/saturation/K12_alignment/sig70_aerobic_enrichment.pdf}
+% \caption{Hexbin plots of ARC vs URCR of the $\sig$ ChIP-exo experiment
+%   under aerobic condition when 10K to 90K reads are being sampled.}
+%   \label{fig:exoQC_sat_aero}
+% \end{figure}
 
@@
 
-\newpage
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width = .9\textwidth]{../figs/for_paper/saturation_analysis_old.pdf}
-  \caption{Comparison of the number of A) candidate regions, B)
-    predicted events, C) identified targets and D) resolution among
-    ChIP-exo, PE ChIP-Seq and SE ChIP-Seq. RegulonDB annotations are
-    considered as a gold standard. A gold standard binding events was
-    deemed identified if a binding event was estimated at a $\pm$
-    \Sexpr{ext} vicinity of it.}
-  \label{fig:design}
-\end{figure}
-
-% \textbf{ChIP-exo, PE ChIP-Seq and SE ChIP-Seq comparison at varying
-%   sequencing depths.} 
- \newpage
-
-\begin{figure}[h!]
-  \centering
-\includegraphics[width = .9\textwidth]{../figs/saturation/K12_alignment/sig70_aerobic_enrichment.pdf}
-\caption{Hexbin plots of ARC vs URCR of the $\sig$ ChIP-exo experiment
-  under aerobic condition when 10K to 90K reads are being sampled.}
-  \label{fig:exoQC_sat_aero}
-\end{figure}
-
-\newpage
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = .9\textwidth]{../figs/for_paper/sig70_methods_comparison_FDR5_topM300.pdf}
-  \caption{Comparison of the resolution between dPeak, Gem, Mace and
-    Peakzilla methods. Resolution is defined as the minimum distance
-    between a RegulonDB annotation and a predicted binding event.}
-  \label{fig:methods_comp}
-\end{figure}
-
-\newpage
-
-\beginsupplement
-
-\section*{Supplement}
-\label{sec:supp}
-
-\emph{Note: I am not sure whether to include several figures on this supplement}
+% \newpage
+
+% \beginsupplement
+
+% \section*{Supplement}
+% \label{sec:supp}
+
+% \emph{Note: I am not sure whether to include several figures on this supplement}
 
@@
 
-\begin{figure}[h!]
-  \centering % note change to stat_density(geom = "line")
-   \includegraphics[width = .46\textwidth,page = 1]{../figs/for_paper/sigma_delta_old_densities.pdf}
-   \includegraphics[width = .46\textwidth,page = 2]{../figs/for_paper/sigma_delta_old_densities.pdf}   
-  \caption{ C) $\delta$ parameter in dPeak measures average distance
-    of the reads to their respective binding site. In ChIP-exo data,
-    reads were located much closer to the binding site than in SET
-    ChIP-Seq. D) $\sigma$ parameter measure the dispersion of reads
-    around each binding site. In ChIP-exo data, reads showed less
-    variation around the their respective binding sites compared to
-    SET ChIP-Seq.}
-
-\end{figure}
-
-\newpage
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = .7\textwidth]{../figs/saturation/K12_alignment/PBC_saturation.pdf}
-  \caption{Average PBC (among all seeds) of the sampled ChIP-exo, PE
-    ChIP-Seq and SE ChIP-Seq experiments under the rif treatment
-    conditions used for saturation analysis.}
-  \label{fig:pbc_saturation}
-\end{figure}
-
-\newpage
-
-\begin{figure}[h!] %%% fix this plot as one ggplot, facet_grid(condition ~ variable,scales = "free_y")
-  \centering
-  \begin{itemize}
-  \item Rep-1 and rif-0min:
-  \end{itemize}
-  \includegraphics[width = .3\textwidth,page =2]{../figs/saturation/K12_alignment/Sig70_rep1_rif0_saturation_analysis.pdf}
-  \includegraphics[width = .3\textwidth,page =3]{../figs/saturation/K12_alignment/Sig70_rep1_rif0_saturation_analysis.pdf}
-  \includegraphics[width = .3\textwidth,page =4]{../figs/saturation/K12_alignment/Sig70_rep1_rif0_saturation_analysis.pdf}
-
-  \begin{itemize}
-  \item Rep-1 and rif-20min:
-  \end{itemize}
-
-  \includegraphics[width = .3\textwidth,page =2]{../figs/saturation/K12_alignment/Sig70_rep1_rif20_saturation_analysis.pdf}
-  \includegraphics[width = .3\textwidth,page =3]{../figs/saturation/K12_alignment/Sig70_rep1_rif20_saturation_analysis.pdf}
-  \includegraphics[width = .3\textwidth,page =4]{../figs/saturation/K12_alignment/Sig70_rep1_rif20_saturation_analysis.pdf}
-
-  \begin{itemize}
-  \item Rep-2 and rif-0min:
-  \end{itemize}
-
-  \includegraphics[width = .3\textwidth,page =2]{../figs/saturation/K12_alignment/Sig70_rep2_rif0_saturation_analysis.pdf}
-  \includegraphics[width = .3\textwidth,page =3]{../figs/saturation/K12_alignment/Sig70_rep2_rif0_saturation_analysis.pdf}
-  \includegraphics[width = .3\textwidth,page =4]{../figs/saturation/K12_alignment/Sig70_rep2_rif0_saturation_analysis.pdf}
-
-  \begin{itemize}
-  \item Rep-2 and rif-20min:
-  \end{itemize}
-
-  \includegraphics[width = .3\textwidth,page =2]{../figs/saturation/K12_alignment/Sig70_rep2_rif20_saturation_analysis.pdf}
-  \includegraphics[width = .3\textwidth,page =3]{../figs/saturation/K12_alignment/Sig70_rep2_rif20_saturation_analysis.pdf}
-  \includegraphics[width = .3\textwidth,page =4]{../figs/saturation/K12_alignment/Sig70_rep2_rif20_saturation_analysis.pdf}
-  \caption{Comparison of the number of predicted events (left),
-    identified targets (middle) and resolution (right) among ChIP-exo,
-    PE ChIP-Seq and SE ChIP-Seq. RegulonDB annotations are considered
-    as gold standard. A RegulonDB binding events was deemed identified
-    if a binding event was estimated at a $\pm$ \Sexpr{ext} vicinity
-    of it.}
-  \label{fig:saturation_rif}
-\end{figure}
-
-% \textbf{ChIP-exo, PE ChIP-Seq and SE ChIP-Seq comparison at varying
-%   sequencing depth}
-
-\newpage
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = .9\textwidth,page = 1]{../figs/saturation/K12_alignment/sig70_rif_enrichment.pdf}
-  \caption{ARC vs URCR hexbin plots of Rep-1 and rif-0min from $\sig$
-    experiment when 100K to 900K reads are being sampled for each
-    panel. }
-  \label{fig:exoQC_sat1}
-\end{figure}
-
-\newpage
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = .9\textwidth,page = 1]{../figs/saturation/K12_alignment/sig70_rif_enrichment.pdf}
-
-  \caption{ARC vs URCR hexbin plots of Rep-1 and rif-20min from $\sig$ experiment when 100K to 900K reads are being
-    sampled for each panel.}
-  \label{fig:exoQC_sat2}
-\end{figure}
-
-\newpage
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = .9\textwidth,page = 1]{../figs/saturation/K12_alignment/sig70_rif_enrichment.pdf}
-  \caption{ARC vs URCR hexbin plots of Rep-2 and rif-0min from $\sig$
-    experiment when 100K to 900K reads are being sampled for each
-    panel.}
-  \label{fig:exoQC_sat3}
-\end{figure}
-
-\newpage
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = .9\textwidth,page = 1]{../figs/saturation/K12_alignment/sig70_rif_enrichment.pdf}
-  \caption{ARC vs URCR hexbin plots of Rep-2 and rif-20min from $\sig$ experiment when 100K to 900K reads are being
-    sampled for each panel.}
-  \label{fig:exoQC_sat4}
-\end{figure}
+% \begin{figure}[h!]
+%   \centering
+%   \includegraphics[width = .7\textwidth]{../figs/saturation/K12_alignment/PBC_saturation.pdf}
+%   \caption{Average PBC (among all seeds) of the sampled ChIP-exo, PE
+%     ChIP-Seq and SE ChIP-Seq experiments under the rif treatment
+%     conditions used for saturation analysis.}
+%   \label{fig:pbc_saturation}
+% \end{figure}
+
+% \newpage
+
+% \begin{figure}[h!] %%% fix this plot as one ggplot, facet_grid(condition ~ variable,scales = "free_y")
+%   \centering
+%   \begin{itemize}
+%   \item Rep-1 and rif-0min:
+%   \end{itemize}
+%   \includegraphics[width = .3\textwidth,page =2]{../figs/saturation/K12_alignment/Sig70_rep1_rif0_saturation_analysis.pdf}
+%   \includegraphics[width = .3\textwidth,page =3]{../figs/saturation/K12_alignment/Sig70_rep1_rif0_saturation_analysis.pdf}
+%   \includegraphics[width = .3\textwidth,page =4]{../figs/saturation/K12_alignment/Sig70_rep1_rif0_saturation_analysis.pdf}
+
+%   \begin{itemize}
+%   \item Rep-1 and rif-20min:
+%   \end{itemize}
+
+%   \includegraphics[width = .3\textwidth,page =2]{../figs/saturation/K12_alignment/Sig70_rep1_rif20_saturation_analysis.pdf}
+%   \includegraphics[width = .3\textwidth,page =3]{../figs/saturation/K12_alignment/Sig70_rep1_rif20_saturation_analysis.pdf}
+%   \includegraphics[width = .3\textwidth,page =4]{../figs/saturation/K12_alignment/Sig70_rep1_rif20_saturation_analysis.pdf}
+
+%   \begin{itemize}
+%   \item Rep-2 and rif-0min:
+%   \end{itemize}
+
+%   \includegraphics[width = .3\textwidth,page =2]{../figs/saturation/K12_alignment/Sig70_rep2_rif0_saturation_analysis.pdf}
+%   \includegraphics[width = .3\textwidth,page =3]{../figs/saturation/K12_alignment/Sig70_rep2_rif0_saturation_analysis.pdf}
+%   \includegraphics[width = .3\textwidth,page =4]{../figs/saturation/K12_alignment/Sig70_rep2_rif0_saturation_analysis.pdf}
+
+%   \begin{itemize}
+%   \item Rep-2 and rif-20min:
+%   \end{itemize}
+
+%   \includegraphics[width = .3\textwidth,page =2]{../figs/saturation/K12_alignment/Sig70_rep2_rif20_saturation_analysis.pdf}
+%   \includegraphics[width = .3\textwidth,page =3]{../figs/saturation/K12_alignment/Sig70_rep2_rif20_saturation_analysis.pdf}
+%   \includegraphics[width = .3\textwidth,page =4]{../figs/saturation/K12_alignment/Sig70_rep2_rif20_saturation_analysis.pdf}
+%   \caption{Comparison of the number of predicted events (left),
+%     identified targets (middle) and resolution (right) among ChIP-exo,
+%     PE ChIP-Seq and SE ChIP-Seq. RegulonDB annotations are considered
+%     as gold standard. A RegulonDB binding events was deemed identified
+%     if a binding event was estimated at a $\pm$ \Sexpr{ext} vicinity
+%     of it.}
+%   \label{fig:saturation_rif}
+% \end{figure}
+
+% % \textbf{ChIP-exo, PE ChIP-Seq and SE ChIP-Seq comparison at varying
+% %   sequencing depth}
+
+% \newpage
+
+% \begin{figure}[h!]
+%   \centering
+%   \includegraphics[width = .9\textwidth,page = 1]{../figs/saturation/K12_alignment/sig70_rif_enrichment.pdf}
+%   \caption{ARC vs URCR hexbin plots of Rep-1 and rif-0min from $\sig$
+%     experiment when 100K to 900K reads are being sampled for each
+%     panel. }
+%   \label{fig:exoQC_sat1}
+% \end{figure}
+
+% \newpage
+
+% \begin{figure}[h!]
+%   \centering
+%   \includegraphics[width = .9\textwidth,page = 1]{../figs/saturation/K12_alignment/sig70_rif_enrichment.pdf}
+
+%   \caption{ARC vs URCR hexbin plots of Rep-1 and rif-20min from $\sig$ experiment when 100K to 900K reads are being
+%     sampled for each panel.}
+%   \label{fig:exoQC_sat2}
+% \end{figure}
+
+% \newpage
+
+% \begin{figure}[h!]
+%   \centering
+%   \includegraphics[width = .9\textwidth,page = 1]{../figs/saturation/K12_alignment/sig70_rif_enrichment.pdf}
+%   \caption{ARC vs URCR hexbin plots of Rep-2 and rif-0min from $\sig$
+%     experiment when 100K to 900K reads are being sampled for each
+%     panel.}
+%   \label{fig:exoQC_sat3}
+% \end{figure}
+
+% \newpage
+
+% \begin{figure}[h!]
+%   \centering
+%   \includegraphics[width = .9\textwidth,page = 1]{../figs/saturation/K12_alignment/sig70_rif_enrichment.pdf}
+%   \caption{ARC vs URCR hexbin plots of Rep-2 and rif-20min from $\sig$ experiment when 100K to 900K reads are being
+%     sampled for each panel.}
+%   \label{fig:exoQC_sat4}
+% \end{figure}
 
@@
 
-% LocalWords:  Ntimes overlaping gaussian
+% LocalWords:  Ntimes overlaping gaussian repk

### 2016-05-09 rmarkdown/paper.tex
fixed version of paper

@@
 enriched regions usually are composed of a higher quantity of reads,
-therefore we examined the FSR (defined as the ratio of number of
-forward stranded reads divided by the total number of reads in a given
-region) as the regions with lower depth are being filtered out. This
-indicator is of particular importance, as it evaluates the ``peak
-pair'' assumption that the original ChIP-exo paper suggested and
-multiple ChIP-exo data analysis methods rely on. For every ChIP-exo
-experiment, we calculated the global FSR and noticed that for all
-experiments is roughly $0.5$, which means there are roughly the same
-amount of reads in both strands.
-
-In order to asses the strand imbalance we created the visualization
-shown in Figure \ref{fig:strand}: Figure \ref{fig:strand}A presents
-the FSR's behavior as the lower depth regions are being filtered out,
-while Figure \ref{fig:strand}B) shows which percentage of the regions
-are composed by reads in both strand or only one (forward or
-backward). In a good data set, it would be expected that all quantiles
-shown to be quickly converging towards the median (in panel A) or the
-regions composed of reads in one strand being made of few fragments
-(in panel B). For each replicate, we divided the partitioned regions
-by asking whether they overlap a set of high quality ChIP-exo peaks,
-and then we tested (using the Wilcoxon rank sum test over the
-imbalance index defined in the methods section) if the strand
-imbalance's distribution is the same for both classes. For regions
-composed by a higher amount of reads, it is harder to distinguish
-their peaks by considering only the strand imbalance, hence in a
-better quality ChIP-exo experiment it is easier to distinguish
-enriched regions by the amount of reads in both strands. Similarly, we
-may consider that the strands of reads for background of a ChIP-exo
-experiment is more unbalanced than those for the enriched regions. In
-conclusion, Figure \ref{fig:strand} shows that the global FSR does not
-represent the experiment's local strand imbalance, hence the ``peak
-pair'' assumption does not hold locally in ChIP-exo data.
-
-\textbf{Analysis of ChIP-nexus data with the QC pipeline}
-
-We applied the QC pipeline to the ChIP-nexus experiments and found
-that the overall pattern seen in the visualizations is similar to a
-high quality ChIP-exo experiment. Figure \ref{fig:nexus}A illustrates
-a typical pattern of the ARC vs. URCR plot with ChIP-nexus data: Both
-arms seems to be more narrow and therefore the separation between them
-seems to be more noticeable, this seems reasonable since libraries
-generated by ChIP-nexus usually yields higher complexity than
-libraries generated by ChIP-exo. Additionally, while in ChIP-exo
-experiments is common to observe regions formed by undigested
-fragments in only one strand, we can observe that in ChIP-nexus
-experiment the occurrence of this type of regions is reduced. Figure
-\ref{fig:nexus}B shows that in a ChIP-nexus experiment the FSR
-distribution approximates the median more quickly than in a ChIP-exo
-experiment and Figure \ref{fig:nexus}C we observe that most of these
-single-stranded regions are formed by few fragments, which means that
-good ChIP-nexus experiment are less likely to show very imbalanced
-regions.
-
-\subsubsection{Evaluating ChIP-exo and ChIP-nexus data}
-
-We applied the QC pipeline to all ChIP-exo and ChIP-nexus datasets
-listed in Tables \ref{tab:qc} and \ref{tab:nexus}, that way we
-partitioned the reads of each experiment into the undigested regions
-by keeping the regions formed by at least one fragment. Figure
-\ref{fig:enrich_all}A shows compares the slope of the following model:
-
-\begin{align}
-  \mbox{depth} = \beta \times \text{(\# of unique positions)} +\epsilon  \nonumber
-\end{align}
-
-when sampling 1,000 regions and repeating this process
-10,000 times for every ChIP-exo and ChIP-nexus experiment. The
-slope can be interpreted as the change in depth as number of unique
-position varies. This parameter measures the library complexity on a
-ChIP-exo experiment, since when it is high it means that several
-fragments are being aligned to the same position in the genome.
-
-Figures \ref{fig:enrich_all}B and \ref{fig:enrich_all}C shows the
-parameters of the following model:
-
-\begin{align}
-  \mbox{URCR} = \frac{\kappa}{\mbox{ARC}} + \eta + \epsilon \nonumber
-\end{align}
-
-considering only the regions composed by 10 or more aligned reads,
-that way the model reflects the enriched arm of the ARC vs. URCR
-visualization. The intercept $\eta$ can be interpreted as the
-theoretical $\mbox{URCR}$ for relatively high depth regions and the
-slope $\kappa$ can be interpreted as the decay in $\mbox{URCR}$ as the
-$\mbox{ARC}$ increases. That way experiments with low complexity are
-showing low values for both parameters. We considered a linear model
-between the estimated parameters for each experiment and it's
-respective depth for which we were able to asses that there is not
-significant relationship between the estimated parameters and the
-interaction between depth and organism for each experiment. Hence,
-these parameters may be used as indicators of the overall quality of
-ChIP-exo experiment, but as with the SCC it is recommendable to also
-visualize the SCC curve.
-
-\subsection{High Resolution Binding Site Identification with dPeak and
+therefore we examined the FSR as the regions with lower depth are
+being filtered out. This indicator is of particular importance, as it
+evaluates the ``peak pair'' assumption that the original ChIP-exo
+paper suggested and multiple ChIP-exo data analysis methods rely
+on. For every ChIP-exo experiment, we calculated the global FSR and
+noticed that for all experiments is roughly $0.5$, which means there
+are approximately the same amount of reads in both strands.
+
+In order to study the strand imbalance of the FoxA1 factor ChIP-exo
+experiments in mouse liver generated by Serandour et al.,
+2013\nocite{exoillumina} we considered the visualizations shown in
+Figure \ref{fig:strand}: Figure \ref{fig:strand}A presents the FSR's
+behavior as the lower depth regions are being filtered out, while
+Figure \ref{fig:strand}B) shows the proportion of regions that are
+composed by reads from both strands against the regions formed by
+reads in exclusively one strand. In a high quality dataset, it is
+expected for all quantiles to quickly converging towards the median
+(in panel A) or the regions formed by reads in one strand (either
+forward or backward) to be formed by few fragments (in panel
+B).
+
+Additionally, it is worth mentioning that the background in a ChIP-exo
+experiment is expected to be formed by the undigested ChIP fragments
+that are not bound to a protein, hence it is for the reads in both
+strand to be imbalanced in a background region. Figure
+\ref{fig:strand}C compares the strand imbalance when ChIP-exo regions
+overlap with high quality peaks. It is noticeable that for regions
+composed by a large amount of reads, it is harder to distinguish their
+peaks by considering only the strand imbalance, hence in a high
+quality ChIP-exo experiment the background is expected to show a
+higher imbalance that the enriched regions. In conclusion, Figure
+\ref{fig:strand} shows that the global FSR does not accurately
+represent a ChIP-exo experiment's strand imbalance locally, hence the
+``peak pair'' assumption does not completely hold in every ChIP-exo
+enriched region.
+
+% \textbf{Analysis of ChIP-nexus data with the QC pipeline}
+
+% We applied the QC pipeline to the ChIP-nexus experiments and found
+% that the overall pattern seen in the visualizations is similar to a
+% high quality ChIP-exo experiment. Figure \ref{fig:nexus}A illustrates
+% a typical pattern of the ARC vs. URCR plot with ChIP-nexus data: Both
+% arms seems to be more narrow and therefore the separation between them
+% seems to be more noticeable, this seems reasonable since libraries
+% generated by ChIP-nexus usually yields higher complexity than
+% libraries generated by ChIP-exo. Additionally, while in ChIP-exo
+% experiments is common to observe regions formed by undigested
+% fragments in only one strand, we can observe that in ChIP-nexus
+% experiment the occurrence of this type of regions is reduced. Figure
+% \ref{fig:nexus}B shows that in a ChIP-nexus experiment the FSR
+% distribution approximates the median more quickly than in a ChIP-exo
+% experiment and Figure \ref{fig:nexus}C we observe that most of these
+% single-stranded regions are formed by few fragments, which means that
+% good ChIP-nexus experiment are less likely to show very imbalanced
+% regions.
+
+\subsubsection*{Evaluation of ChIP-exo and ChIP-nexus data with QC
+  pipeline}
+
+We evaluated every experiment listed in Tables \ref{tab:qc_sig} and
+\ref{tab:qc} with the QC pipeline. For each experiment, we explored
+the relationship between depth, width and the number of unique
+positions to summarize the shape obtained in the $\mbox{ARC}$
+vs. $\mbox{URCR}$ visualization. Figure \ref{fig:eval}A shows the
+overall change in depth as the number of unique positions varies for
+the gathered experiment in eukaryotic genomes. This parameter can be
+interpreted as the inverse of $\gamma$ in equation (\ref{mod}), which
+can is also the experiment's large depth $\mbox{URCR}$. In a high
+quality ChIP-exo experiment, it is expected in the $\mbox{ARC}$
+vs. $\mbox{URCR}$ plot to observe two arms, one with low $\mbox{ARC}$
+and varying $\mbox{URCR}$ and another that with a decreasing
+$\mbox{URCR}$ as $\mbox{ARC}$ increases that stabilizes in
+$\gamma$. When the ChIP-exo experiment is not deeply sequenced, high
+values in Figure \ref{fig:eval} indicate that the library complexity
+from the experiment is low. On the other hand, low values corresponds
+to higher quality ChIP-exo experiment.
+
+Figure \ref{fig:eval}B shows the average read coverage bias when the
+experiment's sequencing depth is large/ Under perfect digestion
+setting, most of the reads aligned to binding regions are going to be
+accumulated around a binding event, which suggest to be unlikely to
+observe as the sequencing depth increases to observe a reads being
+aligned to another position that was not previously covered. Low
+quality ChIP-exo experiment exhibit more scattered reads in both
+strands across the genome; therefore it is more likely to observe
+consider reads that align to position in the genome that were not
+previously covered. We can compare the second replicate from the human
+ChIP-nexus experiment against both second and third replicates from
+the TBP factor experiment in K562 cell lines. The sequencing depth of
+this three experiment is comparable but the large depth $\mbox{ARC}$'s
+bias is considerably higher for both ChIP-exo experiments.
+
+The interaction between these two parameters represents the quality of
+a ChIP-exo and ChIP-nexus experiment. When either the adjusted
+$\mbox{ARC}$ or the $\mbox{ARC}$'s bias is large, then it is suggested
+to sub-sample reads from the experiment and then evaluate the
+sub-sampled experiment with the QC pipeline. Considering the TBP
+factor ChIP-exo \cite{venters13} and ChIP-nexus \cite{chipnexus}
+experiment in human K562 cell lines, we sub-sampled 20M to 50M reads
+to cover the sequencing depth's range from all the human experiments
+that were considered in this study and evaluated them with QC
+pipeline. Figure \ref{fig:eval}C exhibits an increasing trend in the
+high-depth adjusted $\mbox{ARC}$ in every panel, where the ChIP-nexus
+trend being considerable lower than both ChIP-exo replicates. On the
+other hand, in Figure \ref{fig:eval}D the bias remains approximately
+constant in ChIP-nexus sub-samples, while it increases for both
+ChIP-exo replicates. This suggest that the ChIP-exo replicates have
+low library complexity and overall lower quality than the ChIP-nexus
+experiment, regardless to the fact that all three experiment are
+deeply sequenced with more than 100 million reads each.
+
+\subsection*{High Resolution Binding Site Identification with dPeak and
   ChIP-exo}
 
-\subsubsection{Comparison with ChIP-Seq data using dPeak}
-
-Figure \ref{fig:reso_all} shows comparisons among ChIP-exo, PE
-ChIP-Seq and SE ChIP-Seq. We considered the RegulonDB data as ground
-truth, since those are the most recent annotation on \emph{Escherichia
-  Coli}. A RegulonDB annotation (Salgado et al, 2012 \cite{regulondb})
-was considered to be identified if the distance from the closest dPeak
-binding site estimate was less that or equal to $20$
-bp. That way, the sensitivity is defined as the proportion of
-RegulonDB annotations identified in a peak and the resolution is
-defined as the minimum distance between a RegulonDB annotation and the
-closest dPeak binding site estimate. Figure \ref{fig:reso_all}A shows
-that the sensitivity increases as the mean distance between binding
-events increases. When the average distance is greater than 200 bp,
-dPeak identifies more than the 75\% of the binding events in each
-peak, this is intuitive as the mean ChIP-Seq's fragment length is
-shorter than 200 bp, hence the read does not contribute to more than
-one binding event \cite{dpeak}. When the binding events in a peak are
-closer to each other, both ChIP-exo and PE ChIP-Seq are comparable, as
-the distance increases ChIP-exo identifies a higher proportion of the
-RegulonDB annotations; additionally SE ChIP-Seq is significantly less
-sensitive than both ChIP-exo and PE ChIP-Seq. Figure
-\ref{fig:reso_all}B shows that ChIP-exo and PE ChIP-Seq are comparable
-in resolution, while both protocols significantly outperform SE
-ChIP-Seq.
-
-\subsubsection{Systematic comparison of ChIP-Seq vs ChIP-exo under
+\subsubsection*{dPeak outperforms competing methods in discovering closely
+  spaced binding events from ChIP-exo and ChIP-Seq data}
+
+After we determined which biases could affect a ChIP-exo experiment's
+quality we decided to proceed and examine binding sites in high
+resolution on high quality ChIP-exo experiments. We examined first
+both aerobic replicates from the \emph{E.Coli} first biosample.
+
+Figures \ref{fig:methods_comp}A and \ref{fig:methods_comp}B compares
+binding sites calculated with Peakzilla \cite{peakzilla}, MACE
+\cite{mace}, GEM \cite{gem} and dPeak \cite{dpeak} in resolution,
+where the resolution is defined as the distance from a RegulonDB
+annotation to its closest prediction. For the first replicate (Figure
+\ref{fig:methods_comp}A), the resolution calculated with all methods
+is comparable, while dPeak outperforms the rest in resolution when
+using the second replicate as seen in Figure \ref{fig:methods_comp}B.
+
+Additionally, Figures \ref{fig:methods_comp}C and
+\ref{fig:methods_comp}D compares the densities of the parameters
+estimated by dPeak for ChIP-exo and SE ChIP-Seq data. As expected, the
+$\delta$ parameter which measures the average distance from the
+$5\prime$ ends to their respective binding events and the $\sigma$
+parameter which is a measure from the reads dispersion around their
+respective binding events are lower for ChIP-exo than for SE
+ChIP-Seq. Therefore, we can asses that dPeak is accurately
+representing that the ChIP-exo fragments are allocated more tightly
+around the transcription factor binding sites.
+
+% \color{blue}
+% outline
+
+% \begin{enumerate}
+% \item \st{Available data, the motivation of this part is two fold:}
+%  \begin{itemize}
+%  \item \st{we want to list all data sets we obtained}
+%  \item \st{we want to emphasize that for the sig70 datasets we have a
+%    unique design that allows us to compare chip exo with SE and PE
+%    chipseq , the later one for the first time}
+%  \end{itemize}
+% \item \st{comparison with chip seq} %? with and without dpeak ?}
+% \item \st{only in methods. list the align algorithm used}
+% \item \st{chip seq qc metrics}
+% \item \st{chip exo qc pipeline, description and justification of the steps}
+% \begin{itemize}
+% \item \st{need to justify the sampling step. for this we may use the depth
+%   table and add the factor to normalize one depth to another. explain
+%   the confounding of pcr amplification and the reduction of the number
+%   of possible mappable positions.}
+% \item \st{case study with FoxA1, analysis of enrichment vs library
+%   complexity and analysis and the strand imbalance tools. in methods
+%   we need to add which tests did we perform to show that the pipeline
+%   works. how do we expect from the diagnostics and how to interpret
+%   the plots.}
+% \item \st{analysis of why the pipeline works. in methods need to specify
+%   how and in results need to specify why it works, and what are the
+%   conclusions from it}
+% \end{itemize}
+% \item high resolution analysis with chip exo and comparison with PE
+%   chip seq
+%   \begin{itemize}
+%   \item justify what do we used dpeak (in methods need to list the
+%     other methods used). perhaps we can add gof plots in
+%     supplement. we also want to compare with the same method,
+%     therefore we can use the fact that dpeak have a unified model for
+%     PE and SE dpeak
+%   \item after comparing with other methods we can compare with PE and
+%     SE ChIP-seq, this comparison gives also resolution and sensitivity
+%   \item compare samples under same budget of reads
+%   \end{itemize} 
+% \end{enumerate}
+
+% \color{black}
+% Figure \ref{fig:methods_comp} compares the resolution defined as the
+% minimum distance between a RegulonDB annotation and a binding site
+% predicted by either Peakzilla \cite{peakzilla}, MACE \cite{mace}, GEM
+% \cite{gem} or dPeak \cite{dpeak}. In a good dataset such as both of
+% the ChIP-exo experiments under aerobic (panel 931 and 933) conditions,
+% all the methods are comparable in resolution, and dPeak slightly
+% outperforms the rest. On the other hand, for experiments dPeak
+% identified binding sites with higher resolution than Mace and
+% Peakzilla, but with low resolution than Gem. This may be due that the
+% fact that Gem uses sequence information in addition to the aligned
+% $5\prime$ end counts that dPeak uses.
+
+\subsubsection*{Systematic comparison of ChIP-Seq vs ChIP-exo under
   varying sequencing depth}
@@
 
-Figure \ref{fig:saturation_rif} shows an analogous analysis but using
-the $\sig$ replicates with and without rif treatment. The left, middle
-and right columns show the fixed depth against the number of predicted
-events, identified targets and resolution being compared at a fixed
-depth level. The behavior of this quantities seems to be opposite to
-the one as in Figure \ref{fig:design}, hence we used the ChIP-exo QC
-pipeline in the fixed depth ChIP-exo experiments. Figure
-\ref{fig:exoQC_sat_aero} shows scatter plots of ARC vs URCR for
-several fixed sample sizes, as the fixed depth increases the two arm
-pattern becomes more distinctive while for lower depth, it seems that
-the majority of the sampled reads were aligned to enriched regions. On
-the other hand in Figures \ref{fig:exoQC_sat1} to
-\ref{fig:exoQC_sat4}, we used the ChIP-exo QC pipeline on the samples
-that are outperformed by PE and SE ChIP-Seq. For a fixed low depth, we
-can see that the majority of the reads are being aligned to
-non-enriched regions since the vertical arm seems to be stronger for
-all 2 conditions and replicates; for higher depth we can see low URCR
-values being predominant, which indicates that the majority of the
-regions being formed by few positions with a higher read
-concentration. In low complexity regions, the reads are being aligned
-to fewer positions but there is no control over the amount of reads
-mapped. Hence, those regions are more likely to being strand-imbalance
-which in turn may bias the binding site estimate and therefore
-decrease the number of identified targets or increase an experiments
-resolution.
-
+% Figure \ref{fig:saturation_rif} shows an analogous analysis but using
+% the $\sig$ replicates with and without rif treatment. The left, middle
+% and right columns show the fixed depth against the number of predicted
+% events, identified targets and resolution being compared at a fixed
+% depth level. The behavior of this quantities seems to be opposite to
+% the one as in Figure \ref{fig:design}, hence we used the ChIP-exo QC
+% pipeline in the fixed depth ChIP-exo experiments. Figure
+% \ref{fig:exoQC_sat_aero} shows scatter plots of ARC vs URCR for
+% several fixed sample sizes, as the fixed depth increases the two arm
+% pattern becomes more distinctive while for lower depth, it seems that
+% the majority of the sampled reads were aligned to enriched regions. On
+% the other hand in Figures \ref{fig:exoQC_sat1} to
+% \ref{fig:exoQC_sat4}, we used the ChIP-exo QC pipeline on the samples
+% that are outperformed by PE and SE ChIP-Seq. For a fixed low depth, we
+% can see that the majority of the reads are being aligned to
+% non-enriched regions since the vertical arm seems to be stronger for
+% all 2 conditions and replicates; for higher depth we can see low URCR
+% values being predominant, which indicates that the majority of the
+% regions being formed by few positions with a higher read
+% concentration. In low complexity regions, the reads are being aligned
+% to fewer positions but there is no control over the amount of reads
+% mapped. Hence, those regions are more likely to being strand-imbalance
+% which in turn may bias the binding site estimate and therefore
+% decrease the number of identified targets or increase an experiments
+% resolution.
+
+\subsubsection*{Comparison with ChIP-Seq data using dPeak}
+
+Figure \ref{fig:reso_all} shows comparisons among ChIP-exo, PE
+ChIP-Seq and SE ChIP-Seq. We considered the RegulonDB data as ground
+truth, since those are the most recent annotation on \emph{E. Coli}. A
+RegulonDB annotation (Salgado et al, 2012\nocite{regulondb}) was
+considered to be identified if the distance from the closest dPeak
+binding site estimate was less that or equal to $20$
+bp. That way, the sensitivity is defined as the proportion of
+RegulonDB annotations identified in a peak and the resolution is
+defined as the minimum distance between a RegulonDB annotation and the
+closest dPeak binding site estimate. Figure \ref{fig:reso_all}A shows
+that the sensitivity increases as the mean distance between binding
+events increases. When the binding events in a peak are closer to each
+other, both ChIP-exo and PE ChIP-Seq are comparable, as the distance
+increases ChIP-exo identifies a higher proportion of the RegulonDB
+annotations; additionally SE ChIP-Seq is significantly less sensitive
+than both ChIP-exo and PE ChIP-Seq. Figure \ref{fig:reso_all}B shows
+that ChIP-exo and PE ChIP-Seq are comparable in resolution, while both
+protocols significantly outperform SE ChIP-Seq.
+
+% When the average distance is greater than 200 bp,
+% dPeak identifies more than the 75\% of the binding events in each
+% peak, this is intuitive as the mean ChIP-Seq's fragment length is
+% shorter than 200 bp, hence the read does not contribute to more than
+% one binding event \cite{dpeak}.
 % In supplement figures \ref{fig:exoQC_sat1} to \ref{fig:exoQC_sat4}, we
@@
   \centering
-  \includegraphics[width = .9  \textwidth,page = 1]{../figs/test_quantify/slope_npos_depth_boxplot.pdf}
-  \includegraphics[width = .9  \textwidth,page = 1]{../figs/test_quantify/arc_urcr_param_regression.pdf}
-  \includegraphics[width = .9  \textwidth,page = 2]{../figs/test_quantify/arc_urcr_param_regression.pdf}
-  \caption{A) Comparison of change in depth as the number of unique
-    position obtained by sampling 1,000 regions and fitting
-    the model $\mbox{depth} = \beta \times \text{(\# unique
-      positions)}$, and repeating this process 10,000 times. B)
-    Intercept and C) Slope in the the $\mbox{URCR} = k / \mbox{ARC} +
-    b$ model for the regions formed by more than 10 aligned
-    positions.}
-  \label{fig:enrich_all}
+  \includegraphics[width = .45\textwidth,page = 1]{../figs/for_paper/methods_comp_FDR5_topM400.pdf}
+  \includegraphics[width = .45\textwidth,page = 2]{../figs/for_paper/methods_comp_FDR5_topM400.pdf}
+   \includegraphics[width = .45\textwidth,page = 1]{../figs/for_paper/sigma_delta_old_densities.pdf}
+   \includegraphics[width = .45\textwidth,page = 2]{../figs/for_paper/sigma_delta_old_densities.pdf}   
+  \caption{Comparison of the resolution between dPeak, Gem, Mace and
+    Peakzilla methods with \emph{E.Coli} aerobic experiment from the
+    first biosample A) Replicate 1 and B) Replicate 2. Resolution is
+    defined as the minimum distance between a RegulonDB annotation and
+    a predicted binding event. C) $\delta$ parameter in dPeak measures
+    average distance of the reads to their respective binding site. In
+    ChIP-exo data, reads were located much closer to the binding site
+    than in SET ChIP-Seq. D) $\sigma$ parameter measure the dispersion
+    of reads around each binding site. In ChIP-exo data, reads showed
+    less variation around the their respective binding sites compared
+    to SET ChIP-Seq.}
+  \label{fig:methods_comp}
 \end{figure}
+
+\newpage
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width = .9\textwidth]{../figs/for_paper/saturation_analysis_old.pdf}
+  \caption{Comparison of the number of A) candidate regions, B)
+    predicted events, C) identified targets and D) resolution among
+    ChIP-exo, PE ChIP-Seq and SE ChIP-Seq. RegulonDB annotations are
+    considered as a gold standard. A gold standard binding events was
+    deemed identified if a binding event was estimated at a $\pm$
+    15 vicinity of it.}
+  \label{fig:design}
+\end{figure}
+
+% \textbf{ChIP-exo, PE ChIP-Seq and SE ChIP-Seq comparison at varying
+%   sequencing depths.} 
+%  \newpage
+
+% \begin{figure}[h!]
+%   \centering
+% \includegraphics[width = .9\textwidth]{../figs/saturation/K12_alignment/sig70_aerobic_enrichment.pdf}
+% \caption{Hexbin plots of ARC vs URCR of the $\sig$ ChIP-exo experiment
+%   under aerobic condition when 10K to 90K reads are being sampled.}
+%   \label{fig:exoQC_sat_aero}
+% \end{figure}
 
@@
 
-\newpage
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width = .9\textwidth]{../figs/for_paper/saturation_analysis_old.pdf}
-  \caption{Comparison of the number of A) candidate regions, B)
-    predicted events, C) identified targets and D) resolution among
-    ChIP-exo, PE ChIP-Seq and SE ChIP-Seq. RegulonDB annotations are
-    considered as a gold standard. A gold standard binding events was
-    deemed identified if a binding event was estimated at a $\pm$
-    15 vicinity of it.}
-  \label{fig:design}
-\end{figure}
-
-% \textbf{ChIP-exo, PE ChIP-Seq and SE ChIP-Seq comparison at varying
-%   sequencing depths.} 
- \newpage
-
-\begin{figure}[h!]
-  \centering
-\includegraphics[width = .9\textwidth]{../figs/saturation/K12_alignment/sig70_aerobic_enrichment.pdf}
-\caption{Hexbin plots of ARC vs URCR of the $\sig$ ChIP-exo experiment
-  under aerobic condition when 10K to 90K reads are being sampled.}
-  \label{fig:exoQC_sat_aero}
-\end{figure}
-
-\newpage
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = .9\textwidth]{../figs/for_paper/sig70_methods_comparison_FDR5_topM300.pdf}
-  \caption{Comparison of the resolution between dPeak, Gem, Mace and
-    Peakzilla methods. Resolution is defined as the minimum distance
-    between a RegulonDB annotation and a predicted binding event.}
-  \label{fig:methods_comp}
-\end{figure}
-
-\newpage
-
-\beginsupplement
-
-\section*{Supplement}
-\label{sec:supp}
-
-\emph{Note: I am not sure whether to include several figures on this supplement}
+% \newpage
+
+% \beginsupplement
+
+% \section*{Supplement}
+% \label{sec:supp}
+
+% \emph{Note: I am not sure whether to include several figures on this supplement}
 
@@
 
-\begin{figure}[h!]
-  \centering % note change to stat_density(geom = "line")
-   \includegraphics[width = .46\textwidth,page = 1]{../figs/for_paper/sigma_delta_old_densities.pdf}
-   \includegraphics[width = .46\textwidth,page = 2]{../figs/for_paper/sigma_delta_old_densities.pdf}   
-  \caption{ C) $\delta$ parameter in dPeak measures average distance
-    of the reads to their respective binding site. In ChIP-exo data,
-    reads were located much closer to the binding site than in SET
-    ChIP-Seq. D) $\sigma$ parameter measure the dispersion of reads
-    around each binding site. In ChIP-exo data, reads showed less
-    variation around the their respective binding sites compared to
-    SET ChIP-Seq.}
-
-\end{figure}
-
-\newpage
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = .7\textwidth]{../figs/saturation/K12_alignment/PBC_saturation.pdf}
-  \caption{Average PBC (among all seeds) of the sampled ChIP-exo, PE
-    ChIP-Seq and SE ChIP-Seq experiments under the rif treatment
-    conditions used for saturation analysis.}
-  \label{fig:pbc_saturation}
-\end{figure}
-
-\newpage
-
-\begin{figure}[h!] %%% fix this plot as one ggplot, facet_grid(condition ~ variable,scales = "free_y")
-  \centering
-  \begin{itemize}
-  \item Rep-1 and rif-0min:
-  \end{itemize}
-  \includegraphics[width = .3\textwidth,page =2]{../figs/saturation/K12_alignment/Sig70_rep1_rif0_saturation_analysis.pdf}
-  \includegraphics[width = .3\textwidth,page =3]{../figs/saturation/K12_alignment/Sig70_rep1_rif0_saturation_analysis.pdf}
-  \includegraphics[width = .3\textwidth,page =4]{../figs/saturation/K12_alignment/Sig70_rep1_rif0_saturation_analysis.pdf}
-
-  \begin{itemize}
-  \item Rep-1 and rif-20min:
-  \end{itemize}
-
-  \includegraphics[width = .3\textwidth,page =2]{../figs/saturation/K12_alignment/Sig70_rep1_rif20_saturation_analysis.pdf}
-  \includegraphics[width = .3\textwidth,page =3]{../figs/saturation/K12_alignment/Sig70_rep1_rif20_saturation_analysis.pdf}
-  \includegraphics[width = .3\textwidth,page =4]{../figs/saturation/K12_alignment/Sig70_rep1_rif20_saturation_analysis.pdf}
-
-  \begin{itemize}
-  \item Rep-2 and rif-0min:
-  \end{itemize}
-
-  \includegraphics[width = .3\textwidth,page =2]{../figs/saturation/K12_alignment/Sig70_rep2_rif0_saturation_analysis.pdf}
-  \includegraphics[width = .3\textwidth,page =3]{../figs/saturation/K12_alignment/Sig70_rep2_rif0_saturation_analysis.pdf}
-  \includegraphics[width = .3\textwidth,page =4]{../figs/saturation/K12_alignment/Sig70_rep2_rif0_saturation_analysis.pdf}
-
-  \begin{itemize}
-  \item Rep-2 and rif-20min:
-  \end{itemize}
-
-  \includegraphics[width = .3\textwidth,page =2]{../figs/saturation/K12_alignment/Sig70_rep2_rif20_saturation_analysis.pdf}
-  \includegraphics[width = .3\textwidth,page =3]{../figs/saturation/K12_alignment/Sig70_rep2_rif20_saturation_analysis.pdf}
-  \includegraphics[width = .3\textwidth,page =4]{../figs/saturation/K12_alignment/Sig70_rep2_rif20_saturation_analysis.pdf}
-  \caption{Comparison of the number of predicted events (left),
-    identified targets (middle) and resolution (right) among ChIP-exo,
-    PE ChIP-Seq and SE ChIP-Seq. RegulonDB annotations are considered
-    as gold standard. A RegulonDB binding events was deemed identified
-    if a binding event was estimated at a $\pm$ 15 vicinity
-    of it.}
-  \label{fig:saturation_rif}
-\end{figure}
-
-% \textbf{ChIP-exo, PE ChIP-Seq and SE ChIP-Seq comparison at varying
-%   sequencing depth}
-
-\newpage
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = .9\textwidth,page = 1]{../figs/saturation/K12_alignment/sig70_rif_enrichment.pdf}
-  \caption{ARC vs URCR hexbin plots of Rep-1 and rif-0min from $\sig$
-    experiment when 100K to 900K reads are being sampled for each
-    panel. }
-  \label{fig:exoQC_sat1}
-\end{figure}
-
-\newpage
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = .9\textwidth,page = 1]{../figs/saturation/K12_alignment/sig70_rif_enrichment.pdf}
-
-  \caption{ARC vs URCR hexbin plots of Rep-1 and rif-20min from $\sig$ experiment when 100K to 900K reads are being
-    sampled for each panel.}
-  \label{fig:exoQC_sat2}
-\end{figure}
-
-\newpage
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = .9\textwidth,page = 1]{../figs/saturation/K12_alignment/sig70_rif_enrichment.pdf}
-  \caption{ARC vs URCR hexbin plots of Rep-2 and rif-0min from $\sig$
-    experiment when 100K to 900K reads are being sampled for each
-    panel.}
-  \label{fig:exoQC_sat3}
-\end{figure}
-
-\newpage
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = .9\textwidth,page = 1]{../figs/saturation/K12_alignment/sig70_rif_enrichment.pdf}
-  \caption{ARC vs URCR hexbin plots of Rep-2 and rif-20min from $\sig$ experiment when 100K to 900K reads are being
-    sampled for each panel.}
-  \label{fig:exoQC_sat4}
-\end{figure}
+% \begin{figure}[h!]
+%   \centering
+%   \includegraphics[width = .7\textwidth]{../figs/saturation/K12_alignment/PBC_saturation.pdf}
+%   \caption{Average PBC (among all seeds) of the sampled ChIP-exo, PE
+%     ChIP-Seq and SE ChIP-Seq experiments under the rif treatment
+%     conditions used for saturation analysis.}
+%   \label{fig:pbc_saturation}
+% \end{figure}
+
+% \newpage
+
+% \begin{figure}[h!] %%% fix this plot as one ggplot, facet_grid(condition ~ variable,scales = "free_y")
+%   \centering
+%   \begin{itemize}
+%   \item Rep-1 and rif-0min:
+%   \end{itemize}
+%   \includegraphics[width = .3\textwidth,page =2]{../figs/saturation/K12_alignment/Sig70_rep1_rif0_saturation_analysis.pdf}
+%   \includegraphics[width = .3\textwidth,page =3]{../figs/saturation/K12_alignment/Sig70_rep1_rif0_saturation_analysis.pdf}
+%   \includegraphics[width = .3\textwidth,page =4]{../figs/saturation/K12_alignment/Sig70_rep1_rif0_saturation_analysis.pdf}
+
+%   \begin{itemize}
+%   \item Rep-1 and rif-20min:
+%   \end{itemize}
+
+%   \includegraphics[width = .3\textwidth,page =2]{../figs/saturation/K12_alignment/Sig70_rep1_rif20_saturation_analysis.pdf}
+%   \includegraphics[width = .3\textwidth,page =3]{../figs/saturation/K12_alignment/Sig70_rep1_rif20_saturation_analysis.pdf}
+%   \includegraphics[width = .3\textwidth,page =4]{../figs/saturation/K12_alignment/Sig70_rep1_rif20_saturation_analysis.pdf}
+
+%   \begin{itemize}
+%   \item Rep-2 and rif-0min:
+%   \end{itemize}
+
+%   \includegraphics[width = .3\textwidth,page =2]{../figs/saturation/K12_alignment/Sig70_rep2_rif0_saturation_analysis.pdf}
+%   \includegraphics[width = .3\textwidth,page =3]{../figs/saturation/K12_alignment/Sig70_rep2_rif0_saturation_analysis.pdf}
+%   \includegraphics[width = .3\textwidth,page =4]{../figs/saturation/K12_alignment/Sig70_rep2_rif0_saturation_analysis.pdf}
+
+%   \begin{itemize}
+%   \item Rep-2 and rif-20min:
+%   \end{itemize}
+
+%   \includegraphics[width = .3\textwidth,page =2]{../figs/saturation/K12_alignment/Sig70_rep2_rif20_saturation_analysis.pdf}
+%   \includegraphics[width = .3\textwidth,page =3]{../figs/saturation/K12_alignment/Sig70_rep2_rif20_saturation_analysis.pdf}
+%   \includegraphics[width = .3\textwidth,page =4]{../figs/saturation/K12_alignment/Sig70_rep2_rif20_saturation_analysis.pdf}
+%   \caption{Comparison of the number of predicted events (left),
+%     identified targets (middle) and resolution (right) among ChIP-exo,
+%     PE ChIP-Seq and SE ChIP-Seq. RegulonDB annotations are considered
+%     as gold standard. A RegulonDB binding events was deemed identified
+%     if a binding event was estimated at a $\pm$ ext vicinity
+%     of it.}
+%   \label{fig:saturation_rif}
+% \end{figure}
+
+% % \textbf{ChIP-exo, PE ChIP-Seq and SE ChIP-Seq comparison at varying
+% %   sequencing depth}
+
+% \newpage
+
+% \begin{figure}[h!]
+%   \centering
+%   \includegraphics[width = .9\textwidth,page = 1]{../figs/saturation/K12_alignment/sig70_rif_enrichment.pdf}
+%   \caption{ARC vs URCR hexbin plots of Rep-1 and rif-0min from $\sig$
+%     experiment when 100K to 900K reads are being sampled for each
+%     panel. }
+%   \label{fig:exoQC_sat1}
+% \end{figure}
+
+% \newpage
+
+% \begin{figure}[h!]
+%   \centering
+%   \includegraphics[width = .9\textwidth,page = 1]{../figs/saturation/K12_alignment/sig70_rif_enrichment.pdf}
+
+%   \caption{ARC vs URCR hexbin plots of Rep-1 and rif-20min from $\sig$ experiment when 100K to 900K reads are being
+%     sampled for each panel.}
+%   \label{fig:exoQC_sat2}
+% \end{figure}
+
+% \newpage
+
+% \begin{figure}[h!]
+%   \centering
+%   \includegraphics[width = .9\textwidth,page = 1]{../figs/saturation/K12_alignment/sig70_rif_enrichment.pdf}
+%   \caption{ARC vs URCR hexbin plots of Rep-2 and rif-0min from $\sig$
+%     experiment when 100K to 900K reads are being sampled for each
+%     panel.}
+%   \label{fig:exoQC_sat3}
+% \end{figure}
+
+% \newpage
+
+% \begin{figure}[h!]
+%   \centering
+%   \includegraphics[width = .9\textwidth,page = 1]{../figs/saturation/K12_alignment/sig70_rif_enrichment.pdf}
+%   \caption{ARC vs URCR hexbin plots of Rep-2 and rif-20min from $\sig$ experiment when 100K to 900K reads are being
+%     sampled for each panel.}
+%   \label{fig:exoQC_sat4}
+% \end{figure}
 

### 2016-05-09 rmarkdown/paper.R

@@ -68,13 +68,6 @@
 #    A <- format_tab(A)
 #    B <- format_tab(B)
 
-## ----all_param,include = FALSE,echo = FALSE,eval = TRUE------------------
-
-  Ntimes <- prettyNum(1e4,big.mark = ",")
-  Msamp <- prettyNum(1e3,big.mark = ",")
-  thresh1 <- 10
-
-
 ## ----comp_param,include = FALSE,echo = FALSE, eval = TRUE----------------
 
   id_dist <- 20
@@ -85,6 +78,9 @@
 
   fdr <- 0.01
   topM <- 300
+  Ntimes <- prettyNum(1e4,big.mark = ",")
+  Msamp <- prettyNum(1e3,big.mark = ",")
+  thresh1 <- 10
 
 
 ## ----comp_param , include = FALSE,echo = FALSE,eval = TRUE---------------

### 2016-05-12 rmarkdown/paper.Rnw
new paper version

@@
 
-The strand cross-correlation was proposed by Kharchenko et al., 2008
-\cite{strandcc} and it may be one of the most used of the ChIP-Seq QC
-metrics. The SCC curve is defined as:
+The strand cross-correlation was proposed by Kharchenko et al.,
+2008\nocite{strandcc} and it may be one of the most used of the
+ChIP-Seq QC metrics. The SCC curve is defined as:
 
@@
 
-is a re-parametrization of the following relationship:
+where $\mbox{npos}$ and $\mbox{width}$ denote the region's number of
+unique position where the reads are aligned to and the region's width
+respectively. Is a re-parametrization of the following relationship:
 
@@
 $\mbox{URCR}$, i.e. the ratio between the number of unique positions
-and depth as the depth limits to infinity.
-
-On the other hand, to interpret $\beta_2$, if we assume that the
-$\gamma$ value is already given, then by expressing $\kappa$ as a
-function of $\mbox{ARC}$ and $\mbox{URCR}$ we can observe the
-following identities:
+and depth as the depth limits to infinity. On the other hand, to
+interpret $\beta_2 = - \kappa / \gamma $, by expressing $\kappa$ as a
+function of $\mbox{ARC}$ and $\mbox{URCR}$ and assuming that $\gamma$
+is already estimated, we can observe the following identities:
 
@@
 
-where $\mbox{npos}$ and $\mbox{width}$ denote the region's number of
-unique position where the reads are aligned to and the region's width
-respectively. This is particularly important, $\gamma$ approximates
-the true $\mbox{URCR}$ as the sequencing depth increases, which
-implies that $- \kappa / \gamma$ can be interpreted as the large
-sequencing depth bias of the $\mbox{ARC}$ since as the depth
-increases, the first term of $\kappa / \gamma$ is going to approximate
-the average read coverage:
+This is important: $\gamma$ approximates the $\mbox{URCR}$ as the
+sequencing depth increases, which implies that $- \kappa / \gamma$ can
+be interpreted as the large sequencing depth bias of the $\mbox{ARC}$
+since as the depth increases, the first term of $\kappa / \gamma$ is
+going to approximate the average read coverage:
 
@@
 
-Hence, we define the $\mbox{ARC}$ bias as $- \kappa / \gamma$.
+Therefore, $\beta_2 = - \kappa / \gamma$ is interpreted as the
+$\mbox{ARC}$ bias as the sequencing depth increases.
 
@@
 
-We compared the estimated binding events predicted by the MOSAiCS +
-dPeak pipeline using reads generated by ChIP-exo, PE and SE ChIP-Seq
-protocols. We called peaks at a $5\%$ FDR level using MOSAiCS (the
-$\mbox{GC content} + \mbox{Mappability}$ model for ChIP-exo and the
-$\mbox{Input only}$ model for PE and SE ChIP-Seq). Then, we
-deconvolved the peak into binding events with dPeak (version 2.0.1) by
-considering a maximum of 5 binding sites on each peak. To avoid false
-positives we only considered ChIP-exo peaks with average ChIP counts
-greater than 3000 that overlapped both the SE and PE ChIP-Seq peaks,
-we considered other cutoff values but still obtained results similar
-to what we presented in this paper.
+We compared the estimated binding events predicted by the
+\textbf{MOSAiCS} + \textbf{dPeak} pipeline using reads generated by
+ChIP-exo, PE and SE ChIP-Seq protocols. We called peaks at a $5\%$ FDR
+level using \textbf{MOSAiCS} (the $\mbox{GC content} +
+\mbox{Mappability}$ model for ChIP-exo and the $\mbox{Input only}$
+model for PE and SE ChIP-Seq). Then, we deconvolved the peak into
+binding events with \textbf{dPeak} (version 2.0.1) by considering a
+maximum of 5 binding sites on each peak. To avoid false positives we
+only considered ChIP-exo peaks with average ChIP counts greater than
+3000 that overlapped both the SE and PE ChIP-Seq peaks, we considered
+other cutoff values but still obtained results similar to what we
+presented in this paper.
 
@@
 \nocite{chipbeyond}
+\nocite{meme}
 

### 2016-07-05 paper/aa.tex
previous version of the paper

Chromatin Immunoprecipitation followed by exonuclease digestion and
next generation sequencing (ChIP-exo) is currently one of the
state-of-the-art high throughput assays for profiling protein-DNA
interactions at or close to single base-pair resolution
\cite{exo1}. It presents a powerful alternative to popularly used
ChIP-seq (Chromatin immunoprecipitation coupled with next generation
sequencing) assay. ChIP-exo experiments first capture millions of DNA
fragments ($150$ - $250$ bps in length) that the protein under study
interacts with using a protein-specific antibody and random
fragmentation of DNA. Then, $\lambda$-exonuclease ($\lambda$-exo) is
deployed to trim the $5^{\prime}$ end of each DNA fragment to each
protein-DNA interaction boundary. This step is unique to ChIP-exo and
aims to achieve significantly higher spatial resolution compared to
ChIP-seq. Finally, high throughput sequencing of a small region ($36$
to $100$ bps) at the $5^{\prime}$ end of each fragment generates
millions of reads.  Similarly, ChIP-nexus (Chromatin
Immunoprecipitation followed by exonuclease digestion, unique barcode,
single ligation and next generation ligation) \cite{chipnexus} is a
further modification on the ChIP-exo protocol. ChIP-nexus aims to
overcome limitations of ChIP-exo by by yielding high complexity
libraries with numbers of cells comparable to that of ChIP-seq
experiments. This is achieved by reducing the numbers of ligations in
the standard ChIP-exo protocol from two to one, and adding unique,
randomized barcodes to adaptors to enable monitoring of
overamplification.
%where both sequencing adaptors are ligated at the
%end of the ChIP fragments. Then, after exonuclease digestion, DNA self-circularization with circLigase, and restriction enzyme cutting between the two adaptors, the final library is amplified. Compared to ChIP-exo, ChIP-nexus aims to   attain  higher resolution analysis while yielding higher complexity libraries.
Figure~\ref{fig:chip_diagram} illustrates the differences between
different ChIP-based protocols: ChIP-exo, single-end (SE) ChIP-seq,
paired-end (PE) ChIP-seq, ChIP-nexus. The $5^{\prime}$ ends of a
ChIP-exo/nexus experiment are clustered more tightly around the
binding sites of the protein than in a ChIP-seq experiment. In a PE
ChIP-seq experiment, both ends are sequenced as opposed to only the
$5^{\prime}$ end in a SE ChIP-seq. \SK{Note to Rene: ChIP-Nexus paper
  has a good description of what can go wrong with ChIP-exo; It seems
  like most of the read imbalance could also be due to ligation
  inefficiency. Although this probably does not explain why we have
  reads only from one strand in some regions. This is more likely
  explained by over digestion in one strand or single stranded TF-DNA
  interaction. Think a bit more on these issues and understand how the
  ligation could give rise to strand imbalance, we can ask Bob to have
  a closer look at the section where we have this discussion.}

Although ChIP-exo/nexus protocols are being adopted by research
community, features of ChIP-exo data, especially those pertaining to
data quality, have not been investigated much. The key features of
ChIP-exo/nexus that separate them from ChIP-seq are broadly as
follows. First, DNA libraries generated by the ChIP-exo protocol are
expected to be less complex than the libraries generated by ChIP-seq
\cite{exo_review}) because digestion by $\lambda$-exo is expected to
restrict the space of genomic positions that sequencing reads can map
to, to small local regions around the actual binding sites. Therefore,
in high quality and especially deeply sequenced ChIP-exo datasets, it
is possible to observe large numbers of reads accumulating at a small
number of bases due to actual signal rather than overamplification
bias as commonly observed in ChIP-seq experiments.  Second, although
we expect approximately the same numbers of reads from both DNA
strands at a given binding site, there may be locally more reads in
one strand than in the other, owing to $\lambda$-exo efficiency,
ligation efficiency, or other factors.  This is a key point with
implications on the statistical analysis of ChIP-exo
data. Specifically, currently available ChIP-exo specific statistical
analysis methods (e.g., Mace \cite{mace}, CexoR \cite{cexor}, and
Peakzilla \cite{peakzilla}) rely on the existence of peak-pairs formed
by forward and reverse strand reads at the binding site.  Finally,
most of current widely used ChIP-seq quality control (QC) guidelines
\cite{encode_qc} may not be directly applicable to ChIP-exo data.

To address these challenges, we develop a suite of diagnostic plots
and summary statistics and implement them in a versatile \texttt{R
  package} named \pname{}.  We apply this pipeline on a large
collection of public and newly generated ChIP-exo/nexus data. We
validate implications of the QC pipeline by evaluation of the samples
for features that capture high signal to noise such as occurrences of
motifs recognized by the profiled DNA interacting protein. Our
analyses of this large collection of data revealed that the so-called
ChIP-exo peak-pair assumption is subject to violations. To further
address this and provide a platform where ChIP-exo and ChIP-seq
experiments can be evaluated with comparable methods, we assess
performances of recently developed methods suitable for ChIP-exo
analysis, including dpeak \cite{dpeak} and GEM \cite{gem}. We observe
that dPeak performs as good or better than the available ChIP-exo
methods and provides a platform where PE and SE ChIP-seq can be
compared with their ChIP-exo counterpart. Our comparisons of PE
ChIP-seq with ChIP-exo interestingly highlights that while ChIP-exo
outperforms PE ChIP-seq in terms of resolution and detection power,
both are significantly better than SE ChIP-seq.

### 2016-07-05 paper/chipexo_QC_vs1_SK.tex
previous version of the paper

\documentclass{bmcart}

\usepackage{xcolor}
\usepackage{url}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows}
\usepackage{float}
\usepackage{verbatim}
\usepackage{tabu}
\usepackage{multirow}
\usepackage{framed}

\usepackage{soul}

\newcommand{\beginsupplement}{%
        \setcounter{table}{0}
        \renewcommand{\thetable}{S\arabic{table}}%
        \setcounter{figure}{0}
        \renewcommand{\thefigure}{S\arabic{figure}}%
     }

\newcommand{\pname}[1]{\texttt{ChIPexoQual}}
\newcommand{\SK}[1]{\textcolor{red}{SK: #1}}

% \usepackage{authblk}
\usepackage[utf8]{inputenc} %unicode support

\newcommand{\sig}{\sigma^{70}}

\begin{document}

\begin{frontmatter}

\begin{fmbox}
\dochead{Draft}

\title{Data Exploration, Quality Control, and Statistical Analysis of
  ChIP-exo/nexus Experiments}

\author[
   addressref={aff1},
   noteref={n1},
   email={welch@stat.wisc.edu}      
]{\inits{RW}\fnm{Rene} \snm{Welch} }
\author[
   addressref={aff6},        % id's of addresses, e.g. {aff1,aff2}
   noteref={n1},
   email={chungd@musc.edu}   % email address
]{\inits{DC}\fnm{Dongjun} \snm{Chung} } 
\author[
   addressref={aff3},
   email={ong@cs.wisc.edu}
]{\inits{IO}\fnm{Irene} \snm{Ong}}
\author[
   addressref={aff3,aff4},
   email={jagrass@wisc.edu}
]{\inits{JG}\fnm{Jeffrey} \snm{Grass}}
\author[
   addressref={aff3,aff4,aff5},
   email={landick@bact.wisc.edu}
]{\inits{RL}\fnm{Robert} \snm{Landick}}
\author[
   addressref={aff1,aff2},
   corref={aff1},
   email={keles@stat.wisc.edu}
]{\inits{SK}\fnm{S\"und\"uz} \snm{Kele\c{s}}}

\address[id=aff1]{%                           % unique id
  \orgname{Department of Statistics, University of Wisconsin Madison}, % university, etc
  \street{1300 University Avenue},                     %
  %\postcode{}                                % post or zip code
  \city{Madison},                              % city
  \cny{WI}                                    % country
}

\address[id=aff2]{%
  \orgname{Department of Biostatistics and Medical Informatics, University of Wisconsin Madison},
  \street{600 Highland Avenue},
%  \postcode{24105}
  \city{Madison},
  \cny{WI}
}
\address[id=aff3]{
  \orgname{Great Lakes Bioenergy Research Center, University of Wisconsin Madison},
  \street{1552 University Avenue},
  \city{Madison},
  \cny{WI}
}
\address[id=aff4]{
  \orgname{Department of Biochemistry, University of Wisconsin Madison},
  \street{433 Babcock Drive},
  \city{Madison},
  \cny{WI}
}
\address[id=aff5]{
  \orgname{Department of Bacteriology, University of Wisconsin Madison},
  \street{1550 Linden Drive},
  \city{Madison},
  \cny{WI}
}
\address[id=aff6]{
  \orgname{Department of Public Health Sciences, Medical University of South Carolina},
  \street{135 Cannon Street},
  \city{Charleston},
  \cny{SC}
}

\begin{artnotes}
\note[id=n1]{These two authors contributed equally.} % note, connected to author
\end{artnotes}

\end{fmbox}

\begin{abstractbox}

  \begin{abstract}
ChIP-exo/nexus experiments present  modifications on the commonly used ChIP-seq protocol
    for high
    resolution mapping of transcription factor binding sites. Although
    many aspects of the ChIP-exo data analysis are similar to those of
    ChIP-seq, these high throughput experiments present a number of unique quality control and analysis challenges. We
    develop a statistical quality control pipeline and accompanying \texttt{R} package, \pname{}, 
    to enable exploration and analysis of ChIP-exo and related experiments. \pname{}  evaluates a number of key
    issues including strand imbalance, library complexity, and
    signal enrichment of data from ChIP-exo/nexus experiments. Assessment of these properties  are facilitated
    through diagnostic plots and summary statistics calculated over
    regions of the genome with varying levels of coverage.
    %\pname pipeline explores and quantifies these aspects by partitioning the experiment reads into a collection of regions, calculating a series of summary statistics for each region, providing visualizations and calculating measures to globally asses t quality of a ChIP-exo experiment.

    We evaluated our QC pipeline with both  large collections of public ChIP-exo/nexus data and 
    multiple, new ChIP-exo datasets from \textit{E. Coli}.  \pname{} analysis of these datasets resulted in  guidelines  for
   using these QC metrics across a wide range of sequencing depths 
   and further insights for modeling ChIP-exo data. %In addition, we demonstrate that the
    %ChIP-exo QC pipeline it is also applicable to ChIP-nexus data,
    %showing that those experiments present higher quality than
    %ChIP-exo experiments under similar conditions.
    Finally, although ChIP-exo experiments have been compared to ChIP-seq experiments with single-end (SE) sequencing, we provide, for the first time, comparisons with paired-end (PE)  ChIP-seq experiments.
    %We compared ChIP-exo with Paired End (PE) and Single End (SE)
    %ChIP-Seq and found the following characteristics: First, although
    %often assumed in ChIP-exo data analysis methods, the ``peak pair''
    %assumptions does not hold locally in actual ChIP-exo data. Second,
    %we for the first time compared PE ChIP-Seq with ChIP-exo and found
%    that 
    We illustrate that, at fixed sequencing depths, ChIP-exo provides higher
    sensitivity, specificity, and spatial resolution than PE
    ChIP-seq and both significantly outperform their SE ChIP-seq counterpart. Furthermore, we show that ChIP-exo and PE ChIP-seq are
    comparable in sensitivity for closely located binding events, but
    as the average distance between binding events increases, ChIP-exo
    exhibits higher sensitivity than PE ChIP-Seq. \SK{The last sentence does not really make much sense? }

  \end{abstract}

\begin{keyword}
  \kwd{ChIP-exo}
  \kwd{ChIP-nexus}
  \kwd{ChIP-seq}
  \kwd{Statistical Quality Control}
  \kwd{Spatial Resolution}
  \kwd{Transcription Factor}  
  \kwd{Binding Site Identification with High-Resolution}
  \kwd{Deconvolution}
\end{keyword}

\end{abstractbox}

\end{frontmatter}

\newpage

\section*{Background}
\label{sec:intro}

Chromatin Immunoprecipitation followed by exonuclease
digestion and next generation sequencing (ChIP-exo) 
is currently one of the state-of-the-art high throughput assays for profiling protein-DNA interactions at or close to
single base-pair resolution \cite{exo1}. It presents a powerful alternative to popularly used ChIP-seq
(Chromatin immunoprecipitation coupled with next generation
sequencing) assay. ChIP-exo experiments first capture millions of DNA
fragments ($150$ - $250$ bps in length) that the protein under study
interacts with using  a
protein-specific antibody and random fragmentation of DNA. Then, $\lambda$-exonuclease ($\lambda$-exo) is deployed to trim the
$5^{\prime}$ end of each DNA fragment to each protein-DNA interaction boundary. This step is unique to ChIP-exo and
aims to achieve  significantly higher spatial resolution
compared to ChIP-seq. Finally, high throughput sequencing of a small
region ($36$ to $100$ bps) at the $5^{\prime}$ end of each fragment generates
millions of reads. 
 Similarly, ChIP-nexus (Chromatin Immunoprecipitation followed by exonuclease
digestion, unique barcode, single ligation and next generation
ligation)  \cite{chipnexus} is a further modification on the
ChIP-exo protocol. ChIP-nexus aims to overcome limitations of ChIP-exo by 
by yielding high complexity libraries with numbers of cells comparable to that of ChIP-seq experiments. This is achieved by reducing the numbers of ligations in the standard ChIP-exo protocol from two to one, and adding unique, randomized barcodes to adoptors to enable monitoring of overamplification. 
%where both sequencing adaptors are ligated at the
%end of the ChIP fragments. Then, after exonuclease digestion, DNA self-circularization with circLigase, and restriction enzyme cutting between the two adaptors, the final library is amplified. Compared to ChIP-exo, ChIP-nexus aims to   attain  higher resolution analysis while yielding higher complexity libraries.
Figure~\ref{fig:chip_diagram} illustrates the
differences between different ChIP-based protocols: ChIP-exo, single-end (SE) ChIP-seq, paired-end
(PE) ChIP-seq, ChIP-nexus. The $5^{\prime}$ ends of a ChIP-exo/nexus experiment are clustered
more tightly around the binding sites of the protein than in a ChIP-seq
experiment. In a PE ChIP-seq experiment, both ends are sequenced as opposed to only the $5^{\prime}$ end
in a SE ChIP-seq. \SK{Note to Rene: ChIP-Nexus paper has a good description of what can go wrong with ChIP-exo; It seems like most of the read imbalance could also be due to ligation inefficiency. Although this probably does not explain why we have reads only from one strand in some regions. This is more likely explained by over digestion in one strand or single stranded TF-DNA interaction. Think a bit more on these issues and understand how the ligation could give rise to strand imbalance, we can ask Bob to have a closer look at the section where we have this discussion.}

Although ChIP-exo/nexus protocols are being adopted by  research community, features of 
ChIP-exo data, especially those pertaining to data quality,  have not been investigated much. The key features of ChIP-exo/nexus that separate them from ChIP-seq are broadly as follows. First, 
DNA libraries
generated by the ChIP-exo protocol are expected to be less complex than the
libraries generated by ChIP-seq \cite{exo_review})  because digestion by $\lambda$-exo   is expected to restrict the space of genomic positions that sequencing reads can map to, to small local regions around the actual binding sites. Therefore, in  high quality and especially deeply sequenced ChIP-exo datasets, it is possible to observe large numbers of reads accumulating at a small number of bases due to actual signal rather than overamplification bias as commonly observed in ChIP-seq experiments.
Second, although we expect approximately the same
numbers  of reads from  both DNA strands at a given binding site,  there may be locally more reads in
one strand than in the other, owing to $\lambda$-exo efficiency, ligation efficiency, or other factors. 
This is a key point with implications on the statistical analysis of ChIP-exo data. Specifically, 
currently available ChIP-exo specific statistical analysis methods   (e.g., 
 Mace \cite{mace}, CexoR \cite{cexor}, and   Peakzilla \cite{peakzilla}) rely on the existence of 
peak-pairs formed by forward and reverse strand reads at the binding site. 
Finally, most of current widely used ChIP-seq
quality control (QC) guidelines  \cite{encode_qc}
may not be directly applicable to ChIP-exo data.

To address these challenges, we develop 
a suite of diagnostic plots and summary statistics and implement them in a versatile \texttt{R package} named \pname{}.
We apply this pipeline on  a large collection of public and newly generated ChIP-exo/nexus data. We validate implications of the QC pipeline  by  evaluation of the samples for features that capture high signal to noise such as occurrences of   motifs recognized by the profiled DNA interacting protein. Our analyses of this large collection of data revealed that  the so-called ChIP-exo peak-pair assumption is subject to violations. To further address this and provide a platform where ChIP-exo and  ChIP-seq experiments can be evaluated with comparable methods, we assess performances of recently developed methods suitable for ChIP-exo analysis, including dpeak \cite{dpeak} and GEM \cite{gem}. We observe that dPeak performs as good or better than the available ChIP-exo methods and provides a platform where PE and SE ChIP-seq can be compared with their ChIP-exo counterpart. Our comparisons of PE ChIP-seq with ChIP-exo interestingly highlights that while ChIP-exo outperforms PE ChIP-seq in terms of resolution and detection power, both are significantly better than SE ChIP-seq.

% while there are not established QC pipelines for ChIP-exo; previous ChIP-exo analyses used ChIP-Seq samples to compare the resolution between experiments \cite{exo1,exoillumina,exo2}.

%to interrogate these
%biases in a ChIP-exo experiment and globally asses the enrichment and
%library complexity of a ChIP-exo sample and a procedure to distinguish
%low complexity libraries from deeply sequenced experiments. 
%This
%aspect is unique to ChIP-exo, since the exonuclease enzyme it is
%expected to digested the reads that are not bound to a transcription
%factor, therefore the number of bases where a ChIP-exo fragment could
%be potentially aligned is reduced. That way, in a high quality
%ChIP-exo experiment is possible to observe a large amount of reads to
%be aligned to unique position due to genomic signal instead of PCR
%artifacts.

%In order to obtain the potential benefits of ChIP-exo on protein
%binding site identification, it is critical to use algorithms that
%could fully utilize information available in ChIP-exo data. Rhee and
%Pugh, 2011 \nocite{exo1} discussed that reads in the forward and reverse
%strand might construct peak pairs around bound proteins, of which
%heights were implicitly assumed to be symmetric. Based on this
%rationale, they used the ``peak pair method'' that predicts the
%midpoint of two modes of peak pairs as potential binding
%sites. Recently developed ChIP-exo data analysis methods, such as Mace
%(Wang et al, 20114\nocite{mace}), CexoR (Madrigal, 2015\nocite{cexor})
%and Peakzilla (Bardet et al., 2013\nocite{peakzilla}), are also based
%on this peak pair assumption. However, appropriateness of such
%assumption was not fully evaluated in the literature yet. Furthermore,
%it is still unknown which factors could affect protein binding site
%identification using ChIP-exo data. In order to address this problem,
%we investigated various aspects of ChIP-exo data by contrasting them
%with their respective ChIP-Seq experiments.

%Currently, research on statistical methods for ChIP-exo data is still
%in its very early stage. Although many methods have been proposed to
%identify protein binding sites from ChIP-Seq data (reviewed by
%Wilbanks and Facciotti, 2012\nocite{evaluation} and Pepke and Wold,
%2009 \nocite{computation}), such as MACS (Zhang et al.,
%2008\nocite{macs}), CisGenome (Ji et al., 2008\nocite{cisgenome}) and
%MOSAiCS (Kuan et al., 2009\nocite{mosaics}), these approaches might
%not fully utilize potentials of ChIP-exo data for high resolution
%identification of protein binding sites. Specifically these approaches
%reveal protein binding sites only in lower resolution, i.e., at an
%interval of hundreds to thousands of base pairs. Furthermore, they
%implicitly assume that there is only one ``mode'' or ``predicted
%binding location'' per this wide genomic interval. More recently,
%deconvolution algorithms such as Deconvolution (Lun et al.,
%2009\cite{csdeconv}), GEM (Guo et al., 2012\nocite{gem}, an improved
%version of Guo et al., 2010\nocite{gps} ) and PICS (Zhang et al.,
%2010\nocite{pics}) have been proposed to identify binding sites in
%higher resolution using ChIP-Seq data. However, most of them are still
%not tailored for ChIP-exo and PE and SE ChIP-Seq data in a unified
%framework and as a result, currently available methods are not
%appropriate for fair comparison between ChIP-exo and ChIP-Seq. To
%address these limitations, we developed and utilized an improved
%version of dPeak (Chung et al., 2013\nocite{dpeak}), a high resolution
%binding site identification (deconvolution) algorithm that we
%previously developed for PE and SE ChIP-Seq data, so that it can also
%handle ChIP-exo data. The dPeak algorithm implements a probabilistic
%model that accurately describes the ChIP-exo and ChIP-Seq data
%generation process.

%Some of the key findings in this work are as follows. First, we
%demonstrate that the ``peak pair'' assumption of Rhee and Pugh,
%2013\nocite{exo2} does not hold well in real ChIP-exo data. Second, we
%found that when analyzing ChIP-exo data and the Input control is not
%available, it is useful to adjust for GC content and mappability
%biases to improve peak calling and binding site identification. Third,
%we evaluated several methods to identify binding events and dPeak
%performs competitively respect to GEM and MACE when analyzing ChIP-exo
%data. Finally, when comparable number of reads is used for both
%ChIP-exo and ChIP-Seq , dPeak coupled with ChIP-exo data provides
%resolution comparable to PE ChIP-Seq and both significantly improve
%the resolution of protein binding identification compared to SE-based
%analysis with any of the available methods.

\section*{Results and discussion}
\label{sec:results}

\subsection*{Publicly available ChIP-exo/nexus and novel \textit{E. coli} ChIP-seq/exo datasets}
We utilized a rich collection of publicly available ChIP-exo/nexus data from multiple organisms to build and evaluate our quality control pipeline (Table~\ref{tab:qc}). These include:   CTCF factor
in human HeLa cells \cite{exo1}; ER factor in human MCF-7
cells \cite{exoillumina}; GR factor in IMR90, K562, and U2OS human
cells \cite{starick15}; TBP factor in human K562 cells
\cite{venters13}. 
ChIP-nexus data included experiments from \cite{chipnexus} profiling
TBP in human K562 cells, MyC and Max in \emph{D. Melanogaster} S2 cells, and Twist and
Dorsal in \emph{D. Melanogaster} embryo.

In order to have a setting where we can compare SE and PE ChIP-seq with their ChIP-exo counterpart, we profiled 
$\sig$ under a variety of  conditions in \textit{E. coli} with both ChIP-exo (Table~\ref{tab:qc_sig}) and PE ChIP-seq (\SK{Table SXX}). Collectively, we 
 generated $\sig$ factor ChIP-exo, PE and SE ChIP-seq experiments  under aerobic ($+O_2$) and anaerobic ($-O_2$)
conditions in glucose minimal media. Similarly, we generated $\sig$ factor ChIP-exo, PE and
SE (generated \emph{in silico}) ChIP-seq experiments in \emph{E. Coli}
under aerobic ($+O_2$) conditions with  and without  rifampicin  treatment.
We refer to these latter set of experiments  as time point 20 minutes and time point 0 minutes \SK{Check for consistency with other parts of the paper and correct, ok to use different naming convention}.
We specifically used these experimental data  for comparisons of ChIP-exo and PE
ChIP-seq assays in identifying closely spaced binding events and their resolution, i.e., physical proximity of the predicted events to the actual binding sites.  This comparison benefits from using $\sig$  which is a transcription initiation factor of housekeeping genes
in \emph{E. Coli}. Many \textit{E. coli} promoters contain
multiple transcription start sites (TSS). These TSSs are often
closely spaced, i.e., within 10 $\sim$ 150 bps of each other, and are considered  to be multiple ``switches'' that differentially
regulate gene expression under diverse growth conditions
\cite{regulondb}.

\subsection*{ChIP-exo versus ChIP-seq: general features}

\textit{Read distributions within signal and background regions.}
We first compared ChIP-seq and ChIP-exo in terms of data features that are well studied in ChIP-seq studies.
Our $\sig$  ChIP-seq and ChIP-exo samples from \textit{E. coli} are especially well suited for this task since they are all deeply sequenced compared to the genome size of \textit{E. coli}.   Figure~\ref{fig:comp} summarizes this comparison for one biological replicate  of ChIP-exo and ChIP-seq experiments from the same biological conditions (samples x and y from Table~\ref{tab:qc_sig} \SK{Maybe add a column as sample number to this table to make referencing easy}). Comparisons with other paired \textit{E. coli} ChIP-seq and ChIP-exo  samples led to similar conclusions (Supplementary Figure \SK{SXXX}).
%compared various factors that could affect binding site
%identification between ChIP-exo and ChIP-Seq data by using the
%ChIP-exo reads from the first biological sample and first replicate
%grown under aerobic condition (first line from Table \ref{tab:qc_sig})
%and a SE ChIP-Seq replicate grown under the same conditions. 
We summarized the extended raw read counts within $150$ bps non-overlapping intervals, i.e., bins, interrogating the genome.
%In order
%to compare distribution of signal and background between ChIP-exo and
%ChIP-Seq data, we counted the number of extended reads mapping to a
%partition of the genome into non-overlapping bins. 
Figure~\ref{fig:comp}(A) depicts that, as expected,  ChIP read counts from ChIP-exo and ChIP-seq are linearly 
correlated especially at high read counts. This indicates that signals for potential binding
sites are well reproducible between ChIP-exo and ChIP-seq data. 
In contrast, there is a clear difference
among the two data types for bins with low read counts, highlighting expected differences in the  background read distributions 
of the two data types. 
\SK{***After seeing these types of plots from other samples (e..g., ER in MCF7, it is unlikely that what we are seeing from $\sig$ samples regarding background reads (the next few sentences) is typical. For CTCF and $\sig$, the ChIP-exo samples have a wider dynamic range, but this could be due to differences in sequencing depth. *** Specifically, background reads from ChIP-seq are  almost
uniformly distributed over background (non-enriched) regions. In comparison, background regions in ChIP-exo show larger variation but have overall lower read counts. Furthermore, there are considerably large numbers of bins with zero read counts in ChIP-exo and non-zero read counts in ChIP-seq.
%and the
%ChIP tag counts in there regions were significantly larger than
%zero. In contrast, in ChIP-exo data, there was larger variation in
%ChIP tag counts among background regions and ChIP tag counts were much
%lower in these regions compared to ChIP-Seq data. There were also
%large proportion of regions without any read in ChIP-exo data. 
Overall, these observations indicate that  a much smaller portion of the
genome is expected to be background in ChIP-exo compared to ChIP-seq and 
 methods that
specifically model the background read distribution might benefit from acknowledging this.}

\textit{Peak-pair assumption.} We next evaluated the peak-pair
assumption, i.e., a cluster of reads
in the forward strand is usually paired with a cluster of reads in the
reverse strand that is located on the right-hand-side of  the binding
site, that is commonly utilized in designing statistical analysis methods for ChIP-exo data \cite{mace, cexor, peakzilla}.
We considered the set of peaks identified in both the ChIP-seq and ChIP-exo samples as high quality peaks (Materials and Methods) and calculated the proportion of forward strand reads in these regions (Figure~\ref{fig:comp}(B)). This plot reveals a higher level of strand imbalance for ChIP-exo compared to ChIP-seq. 
%In order to evaluate this assumption, we reviewed the proportion of reads in the forward strand in high quality ChIP-exo peaks such as at least one binding site is predicted in both ChIP-exo and ChIP-Seq data.
%We
%found that strands of reads were much less balanced in ChIP-exo data
%than in ChIP-Seq data in these regions with potential binding sites
%(Fig. \ref{fig:comp}B) and this indicates that the peak pair
%assumption might not hold in ChIP-exo data. 
Potential reasons for this observation include ligation efficiency, efficiency of $\lambda$-exo digestion, 
and single-stranded protein-DNA interactions. Overall, such an imbalance is prevalent in  \SK{X}\% of the ChIP-exo samples used in this paper.
%the enzyme digestion or the strength at which the protein binds to the
%DNA: Although it is expected for the exonuclease enzyme to digest the
%ChIP fragments starting by their $5\prime$ ends and stopping when
%finding a binding event, it may not be able to reach and digest every
%fragment in the ChIP sample. On the other hand, if the protein is not
%bound to both DNA strand, the enzyme may completely digest the
%fragment. Finally both effect are increased by the PCR amplification.

\textit{Mappability and GC-content bias.} We next evaluated ChIP-exo data of CTCF in HeLa cells
\cite{exo1} to investigate biases inherent to next generation sequencing experiments with eukaryotic genomes. 
Figures \ref{fig:comp}(C) and
\ref{fig:comp}(D) display the bin-level average read counts against
mappability and GC-content. Each data point is obtained by averaging
the read counts across bins with the same mappability of GC-content. 
These biases, increasing linear trend with mappability and non-linear trend with GC-content, are similar to those observed in ChIP-seq datasets \cite{benjamini2011, mosaics,quest}.
%In Figure \ref{fig:comp}C it is shown that the ChIP-exo tag
%counts linearly increases with the mappability score and in Figure
%\ref{fig:comp}D it is shown that for GC - content below 0.6, the mean
%ChIP tag count increases and for GC - content greater than 0.6 it
%shows a decreasing trend. Benjamini and Speed,
%2011\nocite{benjamini2011} and Kuan et al., 2011\nocite{mosaics}
%studied the presence of the mappability and GC - content biases in
%ChIP-Seq's background. It is not surprising to see these biases also
%present in ChIP-exo data, since ChIP-Seq and ChIP-exo signal seems to
%be linearly correlated for enriched regions (Figure
%\ref{fig:comp}A). Rozowsky et al., 2009\nocite{peakseq} and Valouev et
%al., 2008\nocite{quest} provide in depth analysis of the mappability
%and GC - content biases for ChIP-Seq respectively. 
This observation  indicates that analysis of  ChIP-exo data
should benefit from methods that take into account  apparent
sequencing biases such as mappability and GC content, mostly when an
input  control sample is not available.

\subsection*{Application of ENCODE ChIP-seq quality metrics   to ChIP-exo and
  ChIP-nexus data}

ENCODE consortium established empirical and widely used QC metrics on ChIP-seq data
\cite{encode_qc}. Currently, these constitute the state-of-the-art QC pipelines for these high throughput experiments.
We evaluated how these metrics, namely Normalized Strand
Cross-Correlation (NSC), Relative Strand Cross-Correlation (RSC), and PCR
Bottleneck Coefficient (PBC) defined at 
\url{https://genome.ucsc.edu/ENCODE/qualityMetrics.html} \cite{encode_qc}, behave
 on ChIP-exo/nexus data in  Tables~\ref{tab:qc} and \ref{tab:qc_sig}.

%We continued our exploration by investigating whether the current
%state-of-the-art QC pipelines for ChIP-Seq are suitable for ChIP-exo
%and ChIP-nexus. In Tables \ref{tab:qc_sig} and \ref{tab:qc} we
%calculated a collection of the commonly used ChIP-Seq QC metrics using
%the ChIP-exo and ChIP-nexus experiments instead: Normalized Strand
%Cross-Correlation (NSC), Relative Strand Cross-Correlation and PCR
%Bottleneck Coefficient (PBC) defined as in
%\url{https://genome.ucsc.edu/ENCODE/qualityMetrics.html}\nocite{encode_qc}.

\arrayrulewidth=.1mm

\begin{table}[h!]
  \centering
  \begin{tabu} to\linewidth{X[1.3]|X[2.2]|X[-1]|X[1.3]|X[-1,m,c]|X[1.8]|X[-1]|X[-1]|X[-1]}
    \firsthline
    \textbf{Protocol} & \textbf{Organism} & \textbf{TF} & \textbf{Cell type} & \textbf{Rep.} &
    \textbf{Depth} & \textbf{NSC} & \textbf{RSC} &   \textbf{PBC} \\ 
    \hline
    \multirow{13}{1em}{ChIP-exo} & Human & CTCF & HeLa & & 48,478,450 & 16.02 & 1.1960 & 0.4579 \\
    & \multirow{3}{1em}{Human} & \multirow{3}{1em}{ER} & \multirow{3}{3em}{MCF-7} & 1 & 9,289,835 & 
    19.87 & 1.0127 & 0.8082 \\
    &  &  & & 2 & 11,041,833 & 21.48& 1.0063 & 0.8024\\
    &  &  & & 3 & 12,464,836 & 18.72& 1.0100 & 0.8203 \\    
    &  \multirow{3}{1em}{Mouse} & \multirow{3}{1em}{FoxA1} & \multirow{3}{1em}{Liver} & 1 & 22,210,461 & 21.28 & 1.1104 &  0.6562 \\
    &  & & & 2 & 23,307,557 & 60.42 & 1.1604 & 0.7996 \\ 
    &  & & & 3 & 22,421,72  & 72.04 & 1.1975 & 0.1068 \\
    & \multirow{3}{1em}{Human} & \multirow{3}{1em}{GR} & IMR90 & & 47,443,803 & 8.86  & 1.3678 & 0.2978 \\
    & &  & K562 & & 116,518,000   & 4.11 & 1.0441 &0.0504 \\
    & &  & U2OS & & 3,255,111 &  10.05 & 1.0288 & 0.7714 \\
    & \multirow{3}{1em}{Human} & \multirow{3}{1em}{TBP} & \multirow{3}{1em}{K562} &1 & 61,046,382 &  12.01  & 1.1119
 & 0.1232 \\
    & & & &2 & 94,314,770 & 7.93 & 1.0299 & 0.1681 \\
    & & & & 3 & 114,282,270 & 9.25 & 1.1027 & 0.1464\\
\hline
\multirow{10}{1em}{ChIP-nexus} & \multirow{8}{1em}{D.Melanogaster} &
       \multirow{2}{1em}{Dorsal} & \multirow{4}{1em}{embryo} & 1 &   8,863,170 &  7.27 & 1.0402 &  0.6766 \\
 & & &  & 2 & 10,003,562 & 7.19 & 1.0672 & 0.5656\\
 & &  \multirow{2}{1em}{Twist} & & 1 & 18,244,203 & 5.82 &  1.1637 &  0.6592 \\
 & & &  & 2 & 52,546,982 & 5.27  &  1.1805  & 0.4549 \\
 & & \multirow{2}{1em}{Max} & \multirow{4}{1em}{S2} & 1 & 18,320,743 & 3.60 & 1.3628 & 0.5178  \\
 & & & & 2 & 24,965,642  & 3.47 & 1.0138 & 0.2124\\
 & & \multirow{2}{1em}{MyC} & & 1 & 7,832,034 & 5.92 & 1.0115 &  0.3935  \\
 & & & &  2 & 22,824,467 & 5.76 & 1.0045 & 0.1879\\
 & \multirow{2}{1em}{Human} & \multirow{2}{1em}{TBP} & \multirow{2}{1em}{K562} & 1 &  
              33,708,245 & 32.16 & 1.1712 & 0.3102 \\
 & &  &  & 2 & 129,675,001 & 32.70 &  1.2455 & 0.0492 \\
    \lasthline
  \end{tabu}
  \caption{Summary of publicly available data used for development and evaluation of \pname{}. The last three columns depict ENCODE QC metrics on these data:  NSC:  Normalized 
    Strand Cross-Correlation; RSC: Relative Strand Cross-Correlation;  PBC:  
    PCR Bottleneck Coefficient.}  
\label{tab:qc}
\end{table}

\begin{table}[h!]
  \centering
  \begin{tabu} to\linewidth{X[-1,m,c]|X[-1]|X[-1]|X[-1,m,c]|X[-1]|X[-1]|X[-1]|X[-1]}
    \firsthline
    \textbf{Bio Sample} & \textbf{Condition} & \textbf{Treatment} & \textbf{Rep.} & \textbf{Depth} & \textbf{NSC} & \textbf{RSC} & 
    \textbf{PBC}\\
    \hline 
    \multirow{4}{1em}{1} & Aerobic & No Rif. & 1 & 13,961,493 & 103.15 & 2.0193 & 0.1399 \\
    & Aerobic & No Rif. & 2 & 14,810,838 & 162.70 & 1.7805 & 0.1633 \\
    & Anaerobic & No Rif. & 1 &  16,108,774 & 153.51 & 1.8035 & 0.1353 \\
    & Anaerobic & No Rif. & 2 &  13,636,541 & 172.59 & 2.014 & 0.1532 \\ 
    \hline
    \multirow{4}{1em}{2} & Aerobic & No Rif. & 1 & 902,921 & 13.77 & 1.1270 & 0.2689\\
    & Aerobic & No Rif. & 1 & 1,852,124 & 17.91 & 1.5275 & 0.2590\\
    & Aerobic & Rif. 20 min & 2 & 2,104,427 & 29.60 & 1.2844 & 0.2584\\
    & Aerobic & Rif. 20 min & 2 & 11,548,572 & 13.08 & 1.5122 & 0.1510 \\
    \lasthline
    \end{tabu}
    \caption{ Summary of the \emph{E. Coli} $\sig$ ChIP-exo and ChIP-seq samples. 
The last three columns depict ENCODE QC metrics on these data:  NSC:  Normalized 
    Strand Cross-Correlation; RSC: Relative Strand Cross-Correlation;  PBC:  
    PCR Bottleneck Coefficient.    }
\label{tab:qc_sig}
\end{table}

DNA libraries generated by ChIP-exo and ChIP-nexus protocols are
expected to be less complex than the libraries generated by ChIP-seq because
 the numbers of positions to which the reads can 
align to are reduced  due to the exonuclease digestion. 
%considerable amounts of reads are being mapped to specific positions.
This affects the interpretation of the PBC, which is defined as the ratio of the 
number of genomic positions to which exactly one read maps to   the number of genomic positions to which at least one  read maps.
For ChIP-seq samples,  low PBC values (e.g., $\le 0.5$) indicate high levels of PCR amplification bias, i.e., PC bottleneck, 
 unless the sequencing depth is high enough to sature all targets of the factor profiled.
In contrast,  for ChIP-exo/nexus, exonuclease digestion will lead to reads with same exact 
$5^{\prime}$ end even before the PCR amplification step.
%because of the exonuclease digestion,
% several reads that their $5\prime$ end was digested to
%the same position before the amplification step.
We note  that  the PBC values are especially low
for deeply sequenced ChIP-exo and ChIP-nexus
samples; however, this does not automatically indicate severe 
bottlenecking as suggested by standard ChIP-seq guidelines.
%, which by following the
%ChIP-Seq guidelines it would indicate that those experiment show
%severe bottlenecking problems, which may incorrectly suggest that the
%positions with a large amount of aligned reads are being caused by PCR
%amplification rather than observed genomic signal.

The Strand Cross-Correlation (SCC), introduced by \cite{strandcc}, is the most commonly used quality metric in assessing 
ChIP-seq enrichment quality. It aims to quantify  how well the reads mapped to each
strand are clustered around the locations of the protein-DNA interaction sites by calculating the Pearson correlation among forward and backward strands reads by shifting them across a range that covers both the read length of the experiment and the expected average fragment length.
Typical SCC profiles exhibit two local maxima: at the average fragment length and the read length. In high quality experiments with clear ChIP enrichment, the average fragment length maximum coincides with the global maximum.
%, and usually it is expected to observed two local
%maxima, one when the profiles are shifted by the average read length
%and another when the profiles are shifted by the unobserved fragment
%length. 
%In a high quality ChIP-Seq dataset the last one is also the
%SCC global maxima. 
In an
idealized ChIP-exo experiment where the DNA fragments are digested
 to  the boundaries of the protein-DNA interaction sites,  we would expect the SCC profile to maximize at the motif length indicating clustering of the forward and reverse strand reads around the  binding site.
 Figure~\ref{fig:scc_exo} displays the SCC curves for the CTCF  HeLa samples where the ChIP-exo curve shows  local maxima at the motif and read
lengths, while the SE ChIP-seq curves have a local maxima at the read
length and a global maxima at the average fragment
length. SCC profiles for other samples  are available in Supplementary Figures \SK{XXX}. The read length and motif length maxima are often in close proximity of each other; as a result, this renders
QC metrics such as the Normalized
Strand Cross-Correlation (NSC) or the Relative Strand
Cross-Correlation (RSC) harder to interpret; however, the profile itself seems informative about the enrichment signal in ChIP-exo/nexus experiments \SK{this seems true at least based on my initial look at the plots, but we should be cautious in reporting this}.

\subsection*{ChIP-exo quality control pipeline}

We first present the overall pipeline and then discuss
individual components with a case study using ChIP-exo data of FoxA1  
from \cite{exoillumina} and ChIP-nexus data from \cite{chipnexus}.
Figure~\ref{fig:qcdiagram} summarizes  the 4-step pipeline. 
Given aligned reads from a  ChIP-exo/nexus sample, the first step partitions the reference genome into islands by keeping the
non-digested ChIP-exo regions.  In step 2,   the total
number of extended reads \SK{check: extended or not? What do we extend to in ChIP-exo? } overlapping each island ($D_i$) and the number of
unique island positions with at least one aligned read re recorded ($(U_i)$. Then,  three summary statistics $\text{ARC}_i$, $\text{URCR}_i$, and $\text{FSR}_i$ are computed for each region $i$.  $\text{ARC}_i$ denotes the \textit{average read coverage per base pair} and is defined as the ratio of the
\#  of reads in island $i$ ($D_i$) to the width of the island $i$ ($W_i$);  $\text{URCR}_i$, \textit{unique read coverage ratio}, \SK{Is unique read coverage ratio a good name for this?} quantifies inverse of the effective coverage and is defined as the ratio of the \# of genomic positions with at least one aligned read within island $i$ ($U_i$) to the \# of reads in island $i$ ($D_i$);  and  $\text{FSR}_i$ denotes the proportion of forward strand reads.
%\begin{align*}
%  \text{ARC}_i &= \dfrac{\text{\#  of reads in island $i$}}{\text{Width of the island $i$}}, \\
%  \text{URCR}_i & = \dfrac{\text{\# of genomic positions with at least one aligned read within island $i$}}{\text{\# of reads in island %$i$}}, \\
%  \text{FSR}_i &= \dfrac{\text{\# of forward strand reads in island $i$}}{\text{\# of reads in island $i$}}.
%\end{align*}
Step 3 of the pipeline generates several diagnostic plots aimed at quantifying ChIP enrichment and strand imbalance and step 4 generates quantitative summaries of these diagnostic plots \SK{Aren't the linear model fit etc part of the pipeline?}

Figure~\ref{fig:qcdiagram}A presents
the typical behavior of the URCR vs. ARC plot for a high quality ChIP-exo sample. In general, the plot
depicts two strong arms. 
High URCR values correspond to regions with reads
concentrated on a small number of positions. 
The  left arm, with low ARC  and
varying URCR values,  corresponds to  background islands, regions that
are usually composed of scattered reads that were not digested during
the exonuclease step. The right arm  where the URCR decreases as the
ARC increases  corresponds to regions that are usually ChIP enriched.
%and as the URCR decreases the library complexity does it as well, on
%the other hand 
\SK{Not sure what this next sentence is referring to: Finally we quantify we quantify the
relationship between library complexity by the use of two indexes that
represent the change in the number of unique positions per regions and
the change in of the width of a regions as the depth changes. }
Figures~\ref{fig:qcdiagram}B and \ref{fig:qcdiagram}C aim to quantify the strand imbalance as part of the QC pipeline. The former  depicts how
quickly the islands exclusively formed by reads SK{***I changed fragments to reads, but should this stay as fragments***} from a single strand are  filtered out as islands with higher depths are observed. In a
high quality sample,  the proportion of islands with reads from only one strand 
is expected to decrease rapidly  as we consider higher depth regions. In contrast, 
 this proportion remains
approximately constant in  lower quality samples. The latter plot illustrates how quickly the quantiles of 
the  FSR approaches to 0.5, the expected  FSR value in high quality samples.
%median, since in a high quality sample it is
%expected for the median to be approximately 0.5 and the enriched
%regions are going to be composed by fragments sequenced from both
%strands.

\subsubsection*{Application and validation of \pname{} with the FoxA1
  ChIP-exo dataset}
\SK{We should present the whole QC results and end with validation -  panels from the two figures should be switched.}
We next use Fox A1 ChIP-exo datasets, with three biological replicates at comparable sequencing depths from mouse liver cells, 
to illustrate the QC pipeline. 
Figure~\ref{fig:enrich}A presents  URCR vs. ARC plots
for all three replicates. The first and third replicates exhibit a defined
decreasing trend in  $\mbox{URCR}$ as the $\mbox{ARC}$ increases. This indicates that
these samples exhibit a higher ChIP enrichment than the
second replicate. On the other hand, the overall $\mbox{URCR}$ level
from the first two replicates is higher than that of the third replicate
level,  elucidating  that the  libraries for the first two replicates are more complex than that of  the
third replicate. 
%As we discussed above,  background fragments are often digested by
%the exonuclease enzyme in ChIP-exo experiments; therefore, the balance between the enrichment
%and library complexity of an experiment is one of the key factors determining
%the sample's data quality.
%Using the Fox A1 in mouse liver cell lines generated by Serandour et
%al., 2013\nocite{exoillumina} and these two quantities, we explored
%the relationship between library complexity and experiment
%enrichment.

%\textbf{Analysis of strand imbalance in FoxA1 ChIP-exo data}

%The strand imbalance assessment is based in the observation that the
%enriched regions usually are composed of a higher quantity of reads,
%therefore we examined the FSR as the regions with lower depth are
%being filtered out. This indicator is of particular importance, as it
%evaluates the ``peak pair'' assumption that the original ChIP-exo
%paper suggested and multiple ChIP-exo data analysis methods rely
%on. For every ChIP-exo experiment, we calculated the global FSR and
%noticed that for all experiments is roughly $0.5$, which means there
%are approximately the same amount of reads in both strands.

Figure~\ref{fig:strand}A and B display the strand-imbalance diagnostic plots and highlight specific problems with replicates 2 and 3. In particular, \SK{I have a hard time interpreting B and C specifically, need to include something about their implications for replicates 2 and 3.}
%In order to study the strand imbalance of the FoxA1 factor ChIP-exo
%experiments in mouse liver generated by Serandour et al.,
%2013\nocite{exoillumina} we considered the visualizations shown in
%Figure \ref{fig:strand}: Figure \ref{fig:strand}A presents the FSR's
%behavior as the lower depth regions are being filtered out, while
%Figure \ref{fig:strand}B) shows the proportion of regions that are
%composed by reads from both strands against the regions formed by
%reads in exclusively one strand. In a high quality dataset, it is
%expected for all quantiles to quickly converging towards the median
%(in panel A) or the regions formed by reads in one strand (either
%forward or backward) to be formed by few fragments (in panel
%B). 
%Additionally, it is worth mentioning that the background in a ChIP-exo
%experiment is expected to be formed by the undigested ChIP fragments
%that are not bound to a protein, hence it is for the reads in both
%strand to be imbalanced in a background region. 
\SK{I also don't follow the discussion for panel C: Figure
\ref{fig:strand}C compares the strand imbalance when ChIP-exo islands 
overlap with high quality peaks. It is noticeable that for regions
composed by a large amount of reads, it is harder to distinguish their
peaks by considering only the strand imbalance, hence in a high
quality ChIP-exo experiment the background is expected to show a
higher imbalance that the enriched regions. In conclusion, Figure
\ref{fig:strand} shows that the global FSR does not accurately
represent a ChIP-exo experiment's strand imbalance locally, hence the
``peak pair'' assumption does not completely hold in every ChIP-exo
enriched region.}

Overall, we conclude that replicate 1 is higher quality than both  of replicates 2 and 3.
We validate this observation with a motif analysis on the candidate binding regions identified from these replicates \SK{Could we say binding events? or is the peak analysis on the whole binding region, i.e., peak? If so, we may want o change "candidate regions" to peaks?} 
Figure~\ref{fig:enrich}B summarizes the total numbers of regions identified from each replicate. The lower number of enriched regions from replicate 2 is consistent with the lower ChIP enrichment pattern in the $UCRC$ vs. $ARC$ diagnostic plot.  Scanning of these regions for the occurrence of FoxA1 sequence motif with the FIMO tool \cite{fimo} indicates that the first replicate outperforms the other two in terms of percentage of candidate regions with the FoxA1 motif (Figure~\ref{fig:enrich}C).
%Considering the regions and summary statistics we obtained a set of
%candidate sites for each replicate, the total number of candidate
%sites is shown in Figure \ref{fig:enrich}B.
%For this sites, we
%extracted the sequence and searched for the FoxA1 motif using
%\textbf{FIMO} \cite{fimo}. Figure \ref{fig:enrich}C shows the fraction
%of candidate site where a motif was found. The first replicate
%outperforms the rest in both number of candidate sites and proportion
%of sites with motif; on the other hand we can see that the second
%replicate candidate sites are more likely to contain the motif
%sequence, but there more candidate sites in the third replicate than
%in the second. 
Furthermore, Figure~\ref{fig:enrich}D displays the average normalized
read coverage around the actual motif locations in the candidate binding regions.
These coverage plots reveal that the ChIP signal is more defined for
the first and third replicates than the second one, indicating overall strength of the ChIP enrichment in these samples compared to the second replicate.
% however,  overall, the signal pattern is reproducible in all of the three biological replicates.
Figures \ref{fig:enrich}E, \ref{fig:enrich}F, and
\ref{fig:enrich}G further highlight the overall quality of the identified motif sequences for each replicate and suggest  that libraries with high library complexity (replicates 1 and 3) capture binding sites with better motif matches.
\SK{Worth thinking a bit more about the interpretation.}

%In summary,  the diagnostics plots of the QC pipeline highlights that  observing a defined
%decreasing trend in $\mbox{URCR}$ as $\mbox{ARC}$ increases implies
%ChIP enrichment and multiple samples can be compared with respect to the degree of this trend. 
%Second, for experiments that are not deeply
%sequenced, low $\mbox{URCR}$ values represent low complexity. 
%decreases the amount of sites of motif found but it doesn't modify the
%read distribution around the motif. Finally, libraries with higher
%complexity returns sequences with a better motif match. \SK{We do not have enough analysis with motif validation to conclude these.}

\subsubsection*{High sequencing depth may confound low-complexity library issues}
We next evaluated every sample listed in Tables~\ref{tab:qc} and
\ref{tab:qc_sig} with the \pname{} QC pipeline (Supplementary Figures \SK{XXX}). A key, albeit not suprising, observation from large scale analysis is that  the URCR vs. ARC plots  typically display the three patterns captured in the FoxA1 study. We will refer to these as pattern I (FoxA1 replicate 1), II (FoxA1 replicate 2), and III (FoxA1 replicate 3), respectively. 
Pattern III where the two arms along $\mbox{ARC}$ are not distinguishable can arise due to either low-complexity library or high sequencing depth.  For example, all three replicates of the TBP ChIP-exo from K562, with sequencing depths between $\sim$ 60M to 115M reads, and replicate two of TBP ChIP-nexus in K562, with sequencing depth of $\sim$ 130M reads, exhibit this pattern. A simple but effective strategy to distinguish the two plausible scenarios from Pattern III is to apply the QC pipeline to sub-samples randomly generated from the full dataset at varying sequencing depths.  We applied this strategy by  sub-sampling 20M to 50M reads, a range that represents the 
 sequencing depths of the human samples we are using in the paper, from the TBP samples. 
$\mbox{URCR}$ vs. $\mbox{ARC}$ diagnostics of these sub-samples  (Supplementary Figures \SK{XXX}) indicate that of the four TBP samples with this pattern, replicates two and three of K562 ChIP-exo suffer from low-complexity library issues, whereas the other samples exhibit the pattern specific to high quality samples. \SK{Add: Validation with motif on TBP ChIP-exo replicates.}

\subsubsection*{Evaluation of a large collection of ChIP-exo and ChIP-nexus data with \pname{}}

We next performed an overall analysis of the \pname{} 
QC pipeline results for the samples in Tables~\ref{tab:qc} and
\ref{tab:qc_sig}.  We  quantified the relationship between $\mbox{ARC}$ and $\mbox{URCR}$ by fitting a reparametrized regression model of $\mbox{URCR}$ as a function of  $\mbox{ARC}$.   Specifically, we considered $D_i = \beta_1 U_i +\beta_2 W_i +\varepsilon_i$, where 
$\varepsilon_i$ represents the random error term.  As we discuss in Materials and methods, this parametrization has a direct connection to $\mbox{URCR}_i = \frac{\kappa}{\mbox{ARC}_i} + \gamma+ \epsilon_i$,   which aims to recapitulate the relationship in the  $\mbox{URCR}$ vs. $\mbox{ARC}$ plots.
%between depth, width, and the number of unique
%positions to summarize the shape obtained in the $\mbox{ARC}$
%vs. $\mbox{URCR}$ visualization. 
Figure~\ref{fig:eval}A displays estimated
overall change in depth ($\hat{\beta}_1$) as the number of positions with at least one aligned read varies across a large collection of ChIP-exo samples from eukaryotic genomes. The $\beta_1$ parameter can be interpreted as the  limiting (i.e., large depth) 
\SK{need a better name for:} average read depth intensity $\mbox{URCR}$ of a sample.
%This parameter can be
%interpreted as the inverse of $\gamma$ in equation (\ref{mod}), which
%can is also the experiment's large depth $\mbox{URCR}$. 
As discussed earlier, high
quality ChIP-exo samples are expected to have two arms in the $\mbox{URCR}$
vs. $\mbox{ARC}$ plots: one with low $\mbox{ARC}$
and varying $\mbox{URCR}$ and another with a decreasing
$\mbox{URCR}$as  $\mbox{ARC}$ increases and stabilizes
$\beta_1$. When the ChIP-exo sample is not deeply sequenced, high
values of $\hat{\beta}_1$ in  Figure~\ref{fig:eval}A indicate that the library complexity
is low. On the other hand, lower values correspond
to higher quality ChIP-exo experiments. Taking into account the depths of these samples and visualizing all the diagnostic plots (\SK{Supplementary Figures XXX}), we conclude that samples with estimated $\beta_1$ values less than $10$ seem to be high quality samples.

We interpret the $\beta_2$ parameter above as the bias for the average read coverage  and display its estimates across all the eukaryotic samples in Figure~\ref{fig:eval}B.
% shows the average read coverage bias estimates when the
%experiment's sequencing depth is large/
 Under perfect digestion by $\lambda$-exo, most of the reads aligned to binding regions are expected to accumulate
 around a binding event.  
 \SK{I don't follow the next argument: We should say something about the expected behavior of $\beta_2$
 which suggest to be unlikely to
observe as the sequencing depth increases to observe a reads being
aligned to another position that was not previously covered. Low
quality ChIP-exo experiment exhibit more scattered reads in both
strands across the genome; therefore it is more likely to observe
consider reads that align to position in the genome that were not
previously covered. }
Although the 
third replicate of the TBP ChIP-exo experiment has comparable sequencing depth to the 
second replicate of the TBP ChIP-nexus experiment, the  
(Figure~\ref{fig:eval}(B)), $\mbox{ARC}$ bias is considerably higher for the ChIP-exo experiment.
This potentially indicates that additional sequencing reads in comparison to replicates 1 and 2 are scattered around new positions instead of accumulating on the existing binding sites.
%We can compare the second replicate from the human
%ChIP-nexus experiment against both second and third replicates from
%the TBP factor experiment in K562 cell lines. The sequencing depths of
%these three experiments  are comparable but the large depth $\mbox{ARC}$'s
%bias is considerably higher for both ChIP-exo experiments.

The interaction between these two parameters  has implications regarding the quality of 
a ChIP-exo and ChIP-nexus sample. When either the adjusted
$\mbox{ARC}$ or the $\mbox{ARC}$ bias is large \SK{Not clear what large for $\beta_2$ means if we are only looking at a single sample} owing to potentially the high sequencing depth of the sample, we suggest
randomly sub-sampling reads to form samples of lower depth and evaluating the 
sub-samples with the QC pipeline. 
As an illustration, we revisit this strategy for  the three replicates of TBP ChIP-exo in K562 \cite{venters13} and 
second replicate from the K562  ChIP-nexus experiments \cite{chipnexus}.
Figure~\ref{fig:eval}C exhibits an increasing trend in the
estimated $\beta_1$ across varying sequencing depths for replicates 2 and 3 which we deem as lower quality than replicate 1. Furthermore, the estimates with lower depths are still higher than that of the replicate 1 and the overall trend of the ChIP-nexus sub-samples. 
%, where the ChIP-nexus
%trend being considerable lower than both ChIP-exo replicates. 
Figure \ref{fig:eval}D illustrates that the $ARC$ bias remains approximately
constant in ChIP-nexus sub-samples and sub-samples of first replicate of ChIP-exo, while it increases for the second and third
ChIP-exo replicates. This suggests that these two ChIP-exo replicates have
low library complexity and overall lower quality than the ChIP-nexus
samples, regardless of the fact that all three experiments are
deeply sequenced with more than 90M  reads each. Furthermore, the \pname{}  diagnostic plots for each sub-sample (\SK{Supplementary Figures XXX}) illustrate that the two arms of the ARC versus UCRC plots are clearly visible in moderate depth sub-samples of TBP ChIP-nexus data.  \SK{Some more work/reorganization not to have too much repetition with this last paragraph and the section that follows right after FoxA1.}

\subsection*{High-resolution binding event identification with statistical analysis  of
  ChIP-exo data}

Our QC pipeline \pname{} operates on aligned read files and does not require any statistical analysis of the data such as identification of potential binding regions/events. This enables its broad and easy usability before any statistical analysis  for identifying binding events from ChIP-exo/nexus data.  We next evaluated recent analytical approaches for ChIP-exo data analysis and further compared ChIP-exo with PE ChIP-seq data on our \textit{E. coli} samples.

\subsubsection*{Evaluation of available methods in discovering closely
  spaced binding events from \textit{E. coli} ChIP-exo  data}

We specifically considered recently developed Peakzilla \cite{peakzilla}, MACE
\cite{mace}, GEM \cite{gem}, and dPeak \cite{dpeak} in our evaluations with 
high quality \emph{E.Coli} samples (samples \SK{x and y} from the aerobic condition). Among these methods, Peakzilla and MACE are specifically developed for ChIP-exo analysis, whereas both dPeak and GEM can identify high resolution binding events from ChIP experiments, rendering them suitable for ChIP-exo analysis as well. 
%After we determined which biases could affect a ChIP-exo experiment's
%quality we decided to proceed and examine binding sites in high
%resolution on high quality ChIP-exo experiments. We examined first
%both aerobic replicates from the \emph{E.Coli} first biosample.

Figures~\ref{fig:methods_comp}A and \ref{fig:methods_comp}B compare
binding events identified with all the four methods  in terms of resolution,
where the resolution is defined as the distance from a RegulonDB \cite{regulondb}
annotation to its closest prediction. The resolutions of  all the methods
are comparable for the first replicate (Figure~\ref{fig:methods_comp}(A)) and dPeak, on average, has slightly better resolution than the rest in the second replicate (Figure~\ref{fig:methods_comp}(B)).

dPeak model  has two parameters that further elucidate characteristics of ChIP-exo data.
The $\delta$ parameter of dPeak quantifies the average distance from the
$5^{\prime}$ ends of the reads to the binding event they are profiling and the $\sigma$
parameter is a measure of dispersion of $5^{\prime}$ ends around the binding events.
Figures~\ref{fig:methods_comp}C and
\ref{fig:methods_comp}D compare the densities of these parameters
estimated by dPeak for \textit{E. coli}  ChIP-exo and SE ChIP-seq data. As expected, 
both   are lower for ChIP-exo than for SE
ChIP-Seq, indicating that dPeak, although originally developed for high-resolution binding event identification from ChIP-seq, can learn this information.
%Therefore, we can asses that dPeak is accurately
%representing that the ChIP-exo fragments are allocated more tightly
%around the transcription factor binding sites.

\subsubsection*{Saturation analysis with ChIP-exo and ChIP-seq: Systematic comparison of ChIP-seq vs ChIP-exo under
  varying sequencing depths}
Establishing competitive performance of dPeak compared to other ChIP-exo analysis methods enable comparison of ChIP-exo and ChIP-seq with a unified analysis framework using dPeak. Previous comparisons of ChIP-exo and SE ChIP-seq
were all performed without controlling for the sequencing depths of the samples. More importantly, although it is well established that PE ChIP-seq leads to better resolution that SE ChIP-seq in terms of binding site identification \SK{Cite Qi's BMC Bioinformatics paper}, previous work does not have an in depth comparison of ChIP-exo and PE ChIP-seq. 
%Previously, ChIP-exo and SE ChIP-Seq have been compared only at a
%fixed depth level in the literature, while they did not include PE
%ChIP-Seq as well either. 
In order to address these limitations of the
previous studies, we performed a sub-sampling experiment by sampling fixed numbers of reads from each of the
$\sig$ ChIP-exo, PE ChIP-seq, and SE ChIP-seq datasets. Specifically, for every $N$ reads in ChIP-exo and SE ChIP-seq, $N/2$ read pairs 
were sampled for PE ChIP-seq to operate under fixed sequencing costs.
%Additionally, it is worth noticing that for PE ChIP-Seq we
%sampled both ends of the fragment, hence for each sequencing depth we
%are sampling the half amount of pairs for PE ChIP-Seq than for
%ChIP-exo or SE ChIP-Seq.

Figure~\ref{fig:design} summarizes the comparisons of the three data types for $\sig$ under aerobic condition as a function of sequencing depth. For these computational experiments, we used $\sig$ RegulonDB \cite{regulondb} binding events as gold standard.
Figure~\ref{fig:design}A displays the number of candidate regions (i.e., peaks)
 where at least one  binding event was identified whereas Figure~\ref{fig:design}B depicts the number of identified binding events, i.e., each candidate region can harbor multiple binding events.
In terms of the number of  binding events, ChIP-exo ranks on the conservative side.  
In Figure~\ref{fig:design}C,  we
display the number of correctly identified binding events, where we consider a  RegulonDB event as
correctly identified if an estimated binding event is identified within 15 bps
of it. Finally, Figure~\ref{fig:design}D presents the
resolution defined as the distance from a RegulonDB binding event to the
closest dPeak prediction. These comparisons indicate that although PE ChIP-seq performs much better than SE ChIP-seq in terms of identifying closely spaced binding events, ChIP-exo outperforms PE ChIP-seq at comparable depths.
%It is remarkable that even when the number
%of candidate peaks or the number of predicted events is lower for
%ChIP-exo, it outperforms both PE and SE ChIP-Seq in number of
%identified targets and resolution.

%This may suggest that with ChIP-exo less false positive peaks are
%being called and that when the targets are being identified, dPeak
%estimates binding locations closer to the true location. 
%\SK{The lines do not cross each other, hard to support the following statement:}
%Additionally, we can see that as the read depth increases, all four indicator seem to stabilize and hit a plateau earlier than the cases for ChIP-Seq, which may indicate that with ChIP-exo a smaller amount of reads is needed to identify the same number of targets than ChIP-Seq, but it may be also possible that this is an artifact occurring due to ChIP-exo's lower library complexity.

%\subsubsection*{Comparison with ChIP-Seq data using dPeak}

Figure~\ref{fig:reso_all} shows comparisons among dPeak analysis of full ChIP-exo, PE
ChIP-seq and SE ChIP-seq $\sig$ datasets \SK{Again, make sure these ChIP-seq samples are introduced before and numbering system will help to refer each sample}. \SK{Are the sequencing depths comparable? We are making a big deal about matching depths and then jumping into this analysis. The only way these results would work is that either the depts are comparable or all the samples have depths sufficient enough for saturation}. Utilizing  RegulonDB binding events as ground truth, 
%We considered the RegulonDB data as ground
%truth, since those are the most recent annotation on \emph{E. Coli}. A
%RegulonDB annotation (Salgado et al, 2012\nocite{regulondb}) was
%considered to be identified if the distance from the closest dPeak
%binding site estimate was less that or equal to 20 bp. 
we computed sensitivity  as the proportion of RegulonDB events 
identified by dPeak in each analysis and the resolution as the minimum
distance between a RegulonDB event and the closest dPeak binding
event prediction. Figure \ref{fig:reso_all}A illustrates that as the mean distance between binding events increases, sensitivity
for all data types increase. Consistent with the sub-sampling experiments, both PE ChIP-seq and ChIP-exo significantly outperform SE ChIP-seq and ChIP-exo exhibits more power than PE ChIP-seq in deconvolving binding events.
%shows that the sensitivity
%increases as the mean distance between binding events increases. When
%the binding events in a peak are closer to each other, both ChIP-exo
%and PE ChIP-Seq are comparable, as the distance increases ChIP-exo
%identifies a higher proportion of the RegulonDB annotations;
%additionally SE ChIP-Seq is significantly less sensitive than both
%ChIP-exo and PE ChIP-Seq. 
Figure~\ref{fig:reso_all}B highlights that 
ChIP-exo and PE ChIP-seq are comparable in resolution with these deeply sequenced samples, while both
protocols significantly outperform SE ChIP-seq.

\section*{Conclusions}
\label{sec:conc}

We presented a systematic exploration of several ChIP-exo/nexus experiments. We
provided a list of factors that reflect the quality of a ChIP-exo
experiment and developed a QC pipeline, named \pname{}. \pname{} takes as input aligned reads and automatically generates several diagnostic plots and summary measures that enable assessing enrichment and library complexity.
%Additionally, a set of
%diagnostics was established to assess the quality of a ChIP-exo
%experiment. 
Our analysis of several datasets indicated that
while the QC pipeline only requires a set of aligned reads
to give a global overview of the quality of a given ChIP-exo dataset, 
implications of the diagnostic plots and the summary measures align well
with more elaborate analysis that is computationally more
expensive to perform and/or requires additional inputs that may not be
available, such as motif occurrences in a set of high quality regions or
resolution analysis based on a gold-standard.

%We studied the shared biases between ChIP-exo and ChIP-Seq data, and
%noticed that for eukaryotic genomes the relationship between ChIP-Seq
%data and either the mappability or the GC content scores are still
%present in ChIP-exo. We also examined ChIP-exo's background and
%noticed that is significantly different from the ChIP-Seq one, since
%it consists of only a small quantity of fragments that was not
%digested by the exonuclease enzyme. Additionally, we showed that we
%have unbalanced number of reads in forward and reverse strands, and
%that in a lower quality ChIP-exo experiment those regions are going to
%be harder to differentiate from the possibly enriched regions.

To the best of our knowledge, we also provide the first systematic comparison 
between ChIP-exo and PE ChIP-seq datasets using our \textit{E. coli} $\sig$ samples. 
This comparison revealed that PE ChIP-seq compares much more competitively with ChIP-exo compared to SE ChIP-seq. However,
overall, ChIP-exo provides the best performance in terms of deconvolving closely spaced binding events and resolution.
The pname{} is available at \url{https://github.com/keleslab/ChIPexoQual}.
%Using a set of annotations as
%gold-standard, we showed that both protocols are comparable in
%resolution and that for regions with more than one binding site,
%ChIP-exo is more sensitive than both SE and PE ChIP-Seq. We made a
%rigorous comparison between fixed depth ChIP-exo, PE ChIP-Seq and SE
%ChIP-Seq, and we probed that for sufficiently complex libraries,
%ChIP-exo experiments can outperform PE and SE ChIP-Seq in number of
%identified targets and resolution. The proposed ChIP-exo QC pipeline
%provides a rigorous, easily interpretable, computationally efficient
%framework to diagnose if the library complexity of a ChIP-exo
%experiment is adequate.

\section*{Materials and methods}
\label{sec:methods}

\subsection*{ChIP-seq/exo/nexus datasets}
% To asses the validity of the ChIP-exo QC pipeline several experiments
% were considered.

\subsubsection*{\textit{E. coli} ChIP-exo and ChIp-seq samples}

We generated ChIP-exo and PE ChIP-seq samples from $\sig$ in
\emph{E.Coli}. For each PE ChIP-seq experiment, an \emph{in silico} SE
version was obtained by randomly sampling one of the two ends in a read pair.

{\color{red}\emph{need to add the growth conditions. \SK{Add something based on dpeak paper, we can ask Jeff to check.}}}

\subsubsection*{Processing of the ChIP-exo and ChIP-nexus samples}
%We gathered a collection of ChIP-exo experiments spanning different
%cell lines and transcription factors: CTCF in HeLA cells \cite{exo1},
%ER in MCF-7 cells \cite{exoillumina}, TBP in K562 cells
%\cite{venters13}, GR in IMR90, K562 and U2OS in K562 cells
%\cite{starick15} and FoxA1 in mouse liver cells
%\cite{exoillumina}. Additionally we used the ChIP-nexus experiments
%generated for He et al., 2015: TBP in human K562, MyC and Max factors
%in S2 cell lines in \emph{D. Melanogaster}; Twist and Dorsal factors
%in \emph{D. Melanogaster} embryo cell lines.
We aligned the read files of the samples listed in Table~\ref{tab:qc_sig}
either following the directions in their original publications when available or with \texttt{bowtie} (version 1.1.2) \cite{bowtie}. 
\SK{include bowtie comment to show how parameters were set.}
%The read files were aligned following the instructions provided in
%their respective sources when available. Otherwise we used
%\emph{bowtie} in its 1.1.2 version.

\subsection*{Generation of a set of high signal regions from \textit{E. coli} samples to assess strand imbalance}

%We used the $\sig$ sample grown under aerobic condition for ChIP-exo and
%SE ChIP-Seq.

We partitioned the \emph{E.Coli} genome into
non-overlaping intervals, i.e., bins, of length 150 bps and counted the number of reads overlapping each bin. As is usually the practice with ChIP-seq analysis, each read was extended to the average fragment length of $150$ bps towards the $3^{\prime}$ direction.
%for each bin we counted how
%many ChIP-exo and SE ChIP-Seq extended reads overlapped each bin. The
%reads were extended by 150 bp to their $3\prime$ directions.
To evaluate the strand imbalance, we identified a set of high quality peaks
for ChIP-exo and SE ChIP-seq by analyzing both with \texttt{mosaics} \cite{mosaics}
under the  $\mbox{GC content + Mappability}$ and $\mbox{Input only}$ modes for background estimation, respectively.
The subset of these peaks for which \texttt{dPeak} analysis  identified one or more binding events were used in FSR assessments (Figure~\SK{Fig num}).  \SK{Do the plots change if we consider only peaks with one binding event? One could argue that when there are multiple events, there might be more background reads, which could skew the FSR distribution.}
%(with $\mbox{GC content + Mappability}$
%and $\mbox{Input only}$ models respectively), such that each peak
%contained at least one binding event (found by using \textbf{dPeak}
%with at most 5 binding events in each peak).

%For each peak, we
%calculated the FSR as:

%\begin{align*}
%    \mbox{FSR} &= \frac{\text{Nr. of fwd. strand reads mapped to the region}}                 {\text{Total nr. of reads mapped to the region}} 
%\end{align*}

%We estimated the $\mbox{FSR}$'s densities by using R's \emph{density}
%function using its default parameters, which computes gaussian kernel
%density estimates.

\subsection*{ENCODE  ChIP-seq QC metric guidelines}

% scc and related
% change table to have pbc at the end

We used the ChIP-seq QC metric definitions  established by \cite{encode_qc} and described in detail at
%Landt et al., 2012 which are also listed in
\url{https://genome.ucsc.edu/ENCODE/qualityMetrics.html}.
These QC metrics were calculated with the \textbf{ChIPUtils}
package (version 0.99.0 from 
\url{https://github.com/welch16/ChIPUtils}). 
%This package provides an
%easy to use interface to calculate basic quality control metrics and
%diagnostic plots for ChIP-Seq data.
Empirical data from the ENCODE project suggests the following guidelines for interpretation of the QC metrics for human and mouse genomes: a PBC value
between $0$ to $0.5$  indicates severe bottlenecking,  $0.5$ to $0.8$
moderate bottlenecking,  $0.8$ to $0.9$  mild
bottlenecking, and  $0.9$ - $1$ no bottlenecking.

\SK{I am in favor of skipping these definitions of ENCODE metrics  for the paper - you might want to keep it for your dissertation-}

%\subsubsection*{Strand Cross-Correlation.}

%The strand cross-correlation was proposed by Kharchenko et al.,
%2008\nocite{strandcc} and it may be one of the most used of the
%ChIP-Seq QC metrics. The SCC curve is defined as:

%\begin{align}
%  y(\delta) = \sum_c w_c r\left[ n_c^+ \left(x + \frac{\delta}{2}
%    \right), n_c^- \left( x- \frac{\delta}{2} \right)\right],
%\label{scc}
%\end{align}

%where $y(\delta)$ is the SCC for a strand shift $\delta$, $r$ is the
%Pearson correlation, $w_c$ is the proportion of reads mapped to
%chromosome $c$ and $n_c^S$ is the read count vector for strand $S$ and
%chromosome $c$. The following QC metrics are used to summarize the shape of the SCC curve:

%\begin{align}
%  \mbox{NSC} &= \frac{\max_\delta y(\delta)}{\min_\delta y(\delta)}, \label{nsc} \\
%  \mbox{RSC} &= \frac{\max_\delta y(\delta) -
%    y_{\text{bgd}}}{y_{\text{rl}} - y_{\text{bgd}}}. \label{rsc}
%\end{align}

%where $y_{\text{bgd}}$ is estimated as the background SCC level is
%defined as $y_{\text{bgd}} = \min_\delta y(\delta)$ and
%$y_{\text{rl}}$ is the value of the SCC curve when the shift equals
%the experiment's read length. It is worth noticing that in Landt et
%al., 2012 the RSC is defined as:

%\begin{align*}
%  \mbox{RSC} &= \frac{\max_\delta y(\delta)}{y_{\text{rl}}}.  
%\end{align*}

%In ChIP-exo data, both definitions are equivalent since in a typical
%ChIP-exo experiment $y_{\text{bgd}}$ is approximately zero.

%\subsubsection*{PCR Bottleneck Coefficient.}

%The PCR Bottleneck coefficient is a measure of library complexity in
%ChIP-Seq data:

%\begin{align}
%  \mbox{PBC} = \frac{\text{Nr. of positions to which exactly one
%      unique mapping read is aligned}}{\text{Nr. of positions to
%      which at least one unique mapping read is aligned}} \nonumber
%\end{align}

\subsection*{ChIP-exo quality control with \texttt{R} package \texttt{ChIPexoQual}}

We implemented our proposed QC pipeline with an \texttt{R} package named \texttt{ChIPexoQual}, available at 
\url{https://github.com/welch16/ChIPexoqual}. The analysis in this paper used version 1.0 of the \pname{} package.

\begin{framed}
\noindent \texttt{ChIPexoQual}: 
The package takes in as input a set of aligned reads from a
ChIP-exo (or ChIP-nexus) experiment and performs the following steps.

\begin{enumerate}
\item Identify read islands, i.e., overlapping clusters of reads separated by gaps, from  read coverage.

\item Compute $D_i$,  number of reads in the  island $i$ , and $U_i$, number of island $i$ positions with at least one aligning read, $i=1, \cdots, I$.

\item For each island $i$, $i=1, \cdots, I$, compute island statistics:
  \begin{align*}
    \mbox{ARC}_i &= \frac{D_i}{W_i}, \quad \mbox{URCR}_i = \frac{U_i}{D_i},  \\
    %\mbox{URCR}_i &= \frac{U_i}{D_i}, \\
    \mbox{FSR}_i &= (\text{\# of forward strand reads aligning to island $i$})/D_i, 
  \end{align*}
  where $W_i$ denotes the width of island $i$.
\item Generate diagnostic plots (i) URCR vs. ARC plot; (ii) \SK{Name this FSR plot}; (iii) \SK{Name this FSR plot} as in Figure~\ref{fig:qcdiagram}.
\item Randomly sample $M$ (at least \SK{?}) islands and fit, 
  \begin{align*}
    D_i = \beta_1 U_i + \beta_2 W_i + \varepsilon,
  \end{align*}
  where $\varepsilon$ denotes the independent error term.  Repeat this process $B$ times and generate  box plots of estimated $\beta_1$ and $\beta_2$. 
\end{enumerate}

\end{framed}

%We implemented this ChIP exo QC pipeline in the \emph{ChIPexoQual} R
%package, additionally it generates visualizations as seen in Figure
%\ref{fig:qcdiagram}. We used its 1.0 version which is available in
%\url{https://github.com/welch16/ChIPexoqual}.

\textit{Interpretation of the linear model in the QC pipeline}
The linear model
\begin{align*}
  D_i = \beta_1 U_i + \beta_2 W_i+\varepsilon_i
\end{align*}
 re-parametrization of the following relationship from $UCRC$ vs. $ARC$ diagnostic plot:
\begin{align}
  \mbox{URCR}_i = \frac{\kappa}{\mbox{ARC}_i} + \gamma + \epsilon_i
\label{mod}
\end{align}
with $\beta_1 = 1 / \gamma$ and $\beta_2 = - \kappa / \gamma$. In this
setting, $\gamma$ can be considered as the large-depth
$\mbox{URCR}_i$, i.e., the limiting ratio between the number of positions with at least one mapping read
and depth as the depth tends to infinity.

\SK{***$\beta_2$ needs more clarification. }
On the other hand, to
interpret $\beta_2 = - \kappa / \gamma $, by expressing $\kappa$ as a
function of $\mbox{ARC}$ and $\mbox{URCR}$ and assuming that $\gamma$
is already estimated, we can observe the following identities:

\begin{align*}
  \kappa &= \frac{\mbox{npos}}{\mbox{width}} - \gamma \mbox{ARC} \\
  \frac{\kappa}{\gamma} &= \frac{1}{\gamma} \frac{\mbox{npos}}{\mbox{width}} - \mbox{ARC} 
\end{align*}

This is important: $\gamma$ approximates the $\mbox{URCR}$ as the
sequencing depth increases, which implies that $- \kappa / \gamma$ can
be interpreted as the large sequencing depth bias of the $\mbox{ARC}$
since as the depth increases, the first term of $\kappa / \gamma$ is
going to approximate the average read coverage:

\begin{align*}
  \frac{\mbox{npos}}{\mbox{width}} \times \lim_{d \rightarrow
    \infty} \frac{d }{\mbox{npos}} \sim \mbox{ARC} 
\end{align*}

Therefore, $\beta_2 = - \kappa / \gamma$ is interpreted as the
$\mbox{ARC}$ bias as the sequencing depth increases.
\SK{***}

\subsection*{Motif analysis of FoxA1 enriched regions}
For each FoxA1 ChIP-exo replicate, we used the ChIP-exo QC pipeline to partition the
mouse genome into a set of islands  with their respective summary
statistics. We  then filtered them into collections of high quality regions
by (i) removing the islands with reads residing only on one strand; (ii) removing the islands with $U_i \le 15$; (iii) removing islands with $D_i < 100$. These thresholds were empirically selected and the overall conclusions were robust to their variation.
%\begin{enumerate}
%\item Removing the regions formed by reads on only one strand.
%\item Removing regions with reads aligned to at most 15 unique
%  positions.
%\item Removing regions with less that 100 reads.
%\end{enumerate}
We used \texttt{FIMO} (version 4.9.1, SK{include command line to show parameters}) \cite{fimo} to identify
the FoxA1 motif within each enriched region using FoxA1 position weight matrix MA0148.1 from the JASPAR database \cite{jaspar}.

\subsection*{Testing for strand imbalance in FoxA1 ChIP-exo replicates}
\SK{ This paragraph could benefit from a bit more polishing.}

We used the ChIP-exo QC pipeline to partition the
mouse genome into a set of islands with their respective summary
statistics. We filtered the islands with reads in only one strand and
 transformed the FSR into an \emph{imbalance index} that is zero
when the resulting region is perfectly balanced in terms of forward and reverse strands reads and infinity when it consists of
reads in one strand exclusively:
\begin{align*}
  \mbox{Imbalance index} = -\log_{10} (4 \times \mbox{FSR} \times (1 -
  \mbox{FSR}))
\end{align*}
We divided the ChIP-exo regions into two classes depending on 
whether or not they overlap  ChIP-exo peaks, where the we identified the peaks with \texttt{mosaics}
using the  $\mbox{GC content} + \mbox{Mappability}$ background model \cite{mosaics} (version 2.9.7) and at a false discovery rate (FDR) level of $0.05$ and minimum read threshold of $100$. 
The  bin size and fragment length for MOSAiCS runs were set to $200$ bps. 
%We called peaks with an FDR of $5\%$, a threshold of
%100 and a maximum gap size of 200 bp. 
We further filtered the resulting peaks by
keeping only the peaks with an average extended ChIP read count of 200 and  used a Wilcoxon test to assess the difference in imbalance index between the two groups of ChIP-exo regions.

%\SK{Is this showing the imbalance index is different between peak and non-peak regions?
%To show that the class that don't overlap with peaks exhibits heavier
%tails, we used a Wilcoxon test over the \emph{Imbalance index}.}

%\subsubsection*{GC - content and Mappability}

%To define the mappability score we follow the definition from Rozowsky
%et al., 2009:

%\begin{align*}
%  m_i &= \sum_{k = i - L +1}^{i + L - 1} \frac{\delta_k }{2L - 1}.
%\end{align*}

%where $\delta_i$ is the indicator if the base at coordinate $i$ can be
%mapped uniquely by a 32 bp sequence at position $i$, and $L$ is the
%expected fragment length. GC - content score is defined analogously,
%where $\delta_i$ represents the occurrence of a G or C at the $i$-th
%position in the genome.

%The mappability and GC - content scores for a bin are defined as the
%average of the scores across the nucleotides in the bin.

% Hence, when we reject the hypothesis of the
% distribution being the same for both classes, it means that the class
% of region that don't overlap peaks shows heavier tails than its
% complement. \footnote{not sure if I should add the FSR densities for
%   both classes in the supplement to prove this statement}

% \subsection*{Construction of a SE ChIP-Seq from a PE ChIP-Seq experiment.}

% For the rif-treatment ChIP-Seq experiments, we sampled SE ChIP-Seq
% experiment from the PE ones by taking one of both ends randomly
% following with equal probability.

\subsection*{High resolution analysis with ChIP-exo}

In all the evaluations using \textit{E. coli} samples, 
we considered RegulonDB \cite{regulondb} $\sig$ sites as gold-standard. A $sig$ site was deemed 
as identified  if there existed a binding event within its  20 bps proximity. We defined the
resolution as the distance from an $\sig$ site to its closest
predicted binding event and the sensitivity as the fraction of correctly identified
$\sig$ sites in a genomic region.

\subsubsection*{Statistical methods for ChIP-exo}

We compared dPeak \cite{dpeak} (\url{https://github.com/dongjunchung/dpeak}, version 2.0.1), GEM \cite{gem} (\url{http://groups.csail.mit.edu/cgs/gem/}, version 2.6), MACE \cite{mace} (\url{http://dldcc-web.brc.bcm.edu/lilab/MACE/docs/html/}, version 1.2), peakzilla \cite{peakzilla} (\url{https://github.com/steinmann/peakzilla}).
%, Chung et al., 2013; GEM, Guo et al., 2012; MACE,
%Wang et al., 2014 and Peazilla, Bardet et al., 2013 for the ChIP-exo
%data analysis.

%For the dPeak algorithm we used the R package
%\textbf{dPeak} version 2.0.1 which is available from
%
%\url{https://github.com/dongjunchung/dpeak}. For the GEM algorithm, we
%used it's Java implementation version 2.6 which is available from
%\url{http://groups.csail.mit.edu/cgs/gem/}. For the Mace algorithm, we
%used it Python implementation version 1.2, which is available from
%\url{http://dldcc-web.brc.bcm.edu/lilab/MACE/docs/html/}. For the
%Peakzilla algorithm, we used the version available in
%\url{https://github.com/steinmann/peakzilla}.

\SK{Too many repeats of mosaics version etc, try to minimize.}

Candidate regions for
\texttt{dPeak} and \texttt{GEM} were identified for each replicate of
ChIP-exo data using the \textbf{MOSAiCS} algorithm \cite{mosaics} (one
sample analysis using false discovery rate of 1\%) implemented as an \texttt{R}
package \textbf{mosaics} (version 2.9.7 from
Bioconductor). We further filtered out candidate regions by
using the top 400 peaks with highest average ChIP read counts to avoid
potential false positives based on the exploratory analysis. These
regions were also explicitly provided to the GEM algorithm as
candidate regions. Default tuning parameters were used during model
fitting for all methods. Although we were able to download \texttt{CexoR} \cite{cexor} (version 1.8  from \emph{bioconductor}), we were unable to use it for
the $\sig$ experiments because of the \SK{*** issues}.

\subsubsection*{dPeak analysis of $\sig$ ChIP-exo and ChIP-seq data}

We compared the estimated binding events predicted by the
\textbf{MOSAiCS} + \textbf{dPeak} pipeline using reads generated by
ChIP-exo, PE and SE ChIP-seq protocols. We called peaks at a $5\%$ FDR
level using \textbf{MOSAiCS} (the $\mbox{GC content} +
\mbox{Mappability}$ background model for ChIP-exo and the $\mbox{Input only}$
model for PE and SE ChIP-seq). Then, we deconvolved the peaks into
binding events with \textbf{dPeak} (version 2.0.1) by considering a
maximum of 5 binding events within each peak. To avoid false positives, we
only considered ChIP-exo peaks with average ChIP read count greater than
3,000 that overlapped both the SE and PE ChIP-seq peaks. Repeating the analysis with other cutoff values led to similar conclusions.

\subsubsection*{Saturation analysis of ChIP-exo, PE and SE ChIP-seq}
\SK{Next paragraph could use more editing.}
We sub-sampled $N$ reads for both ChIP-exo and SE ChIP-seq
protocols. For PE ChIP-seq, we sub-sampled $N/2$ pairs equaling to a total of $N$ reads. 
For each sub-sample, we identified  peaks using MOSAiCS \cite{mosaics}
($\mbox{GC content} + \mbox{Mappability}$ background model for ChIP-exo and
$\mbox{Input only}$ for SE and PE ChIP-seq) \SK{not sure: for the maximum sample
size}. To avoid potential false positives, we considered only the top 500 peaks
for each data protocol. We defined the number of candidate peaks as
the number of top sample peaks with at least one predicted dPeak binding event; the number of
predicted events is the total number of dPeak predicted binding events; the number of identified $\sig$ sites as the number
of gold-standard $\sig$ sites within 15 bps from an estimated binding
event; and the resolution as the   minimum distance from a
gold-standard $\sig$ site to an estimated binding event. We repeated
this analysis for ten random sub-sampling experiments and reported the median across these experiments.

\SK{Yeah, Materials and Methods has too many repeats, needs to be organized a bit and cleaned up. Figure captions need to be modified and made more easy follow. }
 % Style BST file (bmc-mathphys, vancouver, spbasic).
\bibliographystyle{vancouver}
 \bibliography{chip_exo_paper} 
\nocite{exo_gb}
\nocite{maplot1}
\nocite{maplot2}
\nocite{chipbeyond}
\nocite{meme}

% \nocite{esl}

\newpage

\section{Figures}

\begin{figure}[h!]
  \centering
  \includegraphics[width =
  .95 \textwidth]{figures/fig1/chip_explanation_withEnzyme.pdf}
  % \caption{\textbf{Description of ChIP-exo, SE ChIP-Seq and PE
  %     ChIP-Seq:} A TF is bound to the forward (red) and backward
  %   (blue) strand of the DNA. Then, is sonicated: For ChIP-exo a
  %   exonuclease enzyme (orange hexagon) trims the $5\prime$ ends of
  %   each DNA fragment to a fixed distance from the bound protein,
  %   finally is subjected to Immunoprecipitation and amplification. For
  %   both ChIP-exo and SE ChIP-Seq an adapter is ligated (green
  %   triangles) at the $5\prime$ ends, while for PE ChIP-Seq is ligated
  %   to both ends.}
  \caption{A TF is bound to the forward (red) and backward (blue)
    strand of the DNA. Then, is sonicated: For ChIP-exo a exonuclease
    enzyme (orange hexagon) trims the $5\prime$ ends of each DNA
    fragment to a fixed distance from the bound protein, finally is
    subjected to Immunoprecipitation and amplification. For both
    ChIP-exo and SE ChIP-Seq an adapter is ligated (green triangles)
    at the $5\prime$ ends, while for PE ChIP-Seq is ligated to both
    ends.}
  \label{fig:chip_diagram}
\end{figure}
\newpage

\begin{figure}[h!]
  \centering
  \includegraphics[width = .46\textwidth,page = 3]{figures/fig2/ChIPseqPET_ChIPexo_tagCount_comparison.pdf}
  \includegraphics[width =.46\textwidth]{figures/fig2/forward_strand_ratio_comp_old.pdf}
  \includegraphics[width = .46\textwidth,page = 1]{figures/fig2/eukaryotic_bias_CTCF.pdf}
  \includegraphics[width = .46\textwidth,page = 2]{figures/fig2/eukaryotic_bias_CTCF.pdf}
  \caption{ A) Hexbin plot of PE ChIP-Seq bin counts vs ChIP-exo bin
    counts. B) Forward Strand Ratio densities for SE ChIP-Seq and
    ChIP-exo peaks. C) Mappability score vs mean ChIP tag counts with
    0.95 confidence bands. D) GC - content score vs mean ChIP tag
    counts with 0.95 confidence bands.}
  \label{fig:comp}
\end{figure}

\newpage

\begin{figure}[h!]
\centering
\includegraphics[width = .8 \textwidth]{figures/fig3/scc_ctcf2.pdf}
% \includegraphics[width = .8\textwidth]{/p/keles/ChIPexo/volume3/ChIPexo/figs/for_paper/EColi_strand_cross_corr.pdf}
% \caption{\textbf{SCC curves for $\sig$ samples.} The ``phantom peak''
%   and the summit that corresponds to the read and fragment length
%   respectively are confounded due to the exonuclease digestion.}
\caption{SCC curves for human CTCF on HeLa cell lines.  The SCC curve
  for the ChIP-exo sample from \cite{exo1} is shown in the left panel,
  and the SCC for ChIP-Seq samples from
  \textbf{https://www.encodeproject.org/experiments/ENCSR000AOA/} are
  shown in the right panel. The ChIP-exo curve shows local maxima at
  the motif and read length. Both SE ChIP-Seq curves are maximized at
  the fragments length and show a local maxima at its read length. }
  \label{fig:scc_exo}
\end{figure}

\newpage

\begin{figure}[h!]
  \centering
  \includegraphics[width = \textwidth]{figures/fig4/coverage_diagram3.pdf}
  % \caption{\textbf{Flowchart of the ChIP-exo quality control
  %     pipeline.} A) ARC vs. URCR, this plot presents a global view of
  %   the balance between library complexity and enrichment. There are
  %   two arms: First, one with low ARC, which corresponds to regions
  %   formed by few aligned positions; Second, where the URCR decreases
  %   as the ARC increases. B) Min. depth vs Proportions of regions,
  %   this plot shows the strand composition for all the regions formed
  %   by a min. number of reads. Regions with low depth tend to be
  %   formed by undigested reads from one unique strand, while regions
  %   with higher signal are usually formed by reads in both strands. C)
  %   Min depth vs. FSR, this plot depicts how quickly the FSR's
  %   distribution approximate the median. In a high quality sample, the
  %   median is around 0.5, and the other quantiles reach that value
  %   quickly.}
  \caption{The ChIP-exo reads are partitioned by keeping the regions
    in the genome formed by the undigested fragments. For each region,
    we calculate a collection of summary statistics, which are
    visualized as A) ARC vs. URCR, this plot presents a global view of
    the balance between library complexity and enrichment. There are
    two arms: First, one with low ARC, which corresponds to regions
    formed by few aligned positions; Second, where the URCR decreases
    as the ARC increases. B) Min. depth vs Proportions of regions,
    this plot shows the strand composition for all the regions formed
    by a min. number of reads. Regions with low depth tend to be
    formed by undigested reads from one unique strand, while regions
    with higher signal are usually formed by reads in both strands. C)
    Min depth vs. FSR, this plot depicts how quickly the FSR's
    distribution approximate the median. In a high quality sample, the
    median is around 0.5, and the other quantiles reach that value
    quickly.}
  \label{fig:qcdiagram}
\end{figure}

\newpage

\begin{figure}[h!]
  \centering
  \includegraphics[width = .9  \textwidth,page = 1]{figures/fig5/FoxA1_enrichment.pdf}
  \newline
  \includegraphics[width = .23 \textwidth, page = 1]{figures/fig5/FoxA1_barplots.pdf}
  \includegraphics[width = .23 \textwidth, page = 3]{figures/fig5/FoxA1_barplots.pdf}
  \includegraphics[width =.4\textwidth]{figures/fig5/FoxA1_profiles_around_motif.pdf}
  \includegraphics[width = .3\textwidth,page = 2]{figures/fig5/FoxA1_matched_motif_sequence.pdf}
  \includegraphics[width = .3\textwidth,page = 3]{figures/fig5/FoxA1_matched_motif_sequence.pdf}
  \includegraphics[width = .3\textwidth,page = 1]{figures/fig5/FoxA1_matched_motif_sequence.pdf}
  \caption{Using the mouse FoxA1 experiment from \cite{exoillumina}:
    A) Hexbin plots of $\mbox{ARC}$ against $\mbox{URCR}$, there is a
    slight separation into two strong arms, one corresponds to low
    $\mbox{ARC}$ and varying $\mbox{URCR}$, and for the other
    $\mbox{URCR}$ decreases as $\mbox{ARC}$ increases. B) Number of
    candidate sites for each replicate. C) Percentage of candidate
    sites where the FoxA1 motif was detected. D) Average coverage
    around FoxA1 motif. Base distribution for matched sequence for
    Rep-1 (E), Rep-2 (F) and Rep-3 (G).}
  \label{fig:enrich}
\end{figure}

{\color{red} \emph{need to add labels for each figure and think about
    order, in the supplement we have ecdf from fimo score or pval, and
    overlaps with peaks called by mosaics.
\newline
Also, need to change labels from repk to Rep-k
}}

\newpage

\begin{figure}[h!]
  \centering  
  \includegraphics[width = .9\textwidth,page = 1]{figures/fig6/Strand_imbalance.pdf} 
  \newline
  \includegraphics[width = .9\textwidth,page = 2]{figures/fig6/Strand_imbalance.pdf} 
  \includegraphics[width = .4\textwidth]{figures/fig6/Carroll_FSR_depth_VS_pvalWilcoxon.pdf}
  \caption{Strand imbalance using the mouse FoxA1 experiment from
    \cite{exoillumina}: A) FSR distribution quantiles as the lower
    depth regions are being filtered out, all quantiles approach to
    the median as the lower bound increases. B) Stacked histogram with
    the proportion of regions that are formed by two strands or only
    one, in a good sample the single-stranded regions are going to be
    filtered out quickly as in the middle row. C)
    $-\log_{10}(\text{p.value})$ of testing if the imbalance
    distributions differs when ChIP-exo regions overlap their peaks.}
  \label{fig:strand}
\end{figure}

\newpage

\begin{figure}[h!]
  \centering
  \includegraphics[width = .9  \textwidth,page = 1]{figures/fig7/QC_pipeline_eval_boxplot.pdf}
\newline
  \includegraphics[width = .9  \textwidth,page = 2]{figures/fig7/QC_pipeline_eval_boxplot.pdf}
\newline
  \includegraphics[width = .45  \textwidth,page = 1]{figures/fig7/K562_TBP_sample.pdf}
  \includegraphics[width = .45  \textwidth,page = 2]{figures/fig7/K562_TBP_sample.pdf}
  \caption{Comparison of adjusted Average Read Coverage and
    $\mbox{ARC}$ bias for eukaryotic genome ChIP-exo and ChIP-nexus
    experiment. A) Boxplot of adjusted $\mbox{ARC}$. B) Boxplot of
    large sequencing depth $\mbox{ARC}$'s bias. Comparison of
    sub-sampled ChIP-exo and ChIP-nexus experiments of TBP factor in
    human K562 cell lines: C) Adjusted $\mbox{ARC}$ and D) $\mbox{ARC}$'s
    bias.}
  \label{fig:eval}
\end{figure}

\newpage

\begin{figure}[h!]
  \centering
  \includegraphics[width = .45\textwidth,page = 1]{figures/fig8/methods_comp_FDR5_topM400.pdf}
  \includegraphics[width = .45\textwidth,page = 2]{figures/fig8/methods_comp_FDR5_topM400.pdf}
   \includegraphics[width = .45\textwidth,page = 1]{figures/fig8/sigma_delta_old_densities.pdf}
   \includegraphics[width = .45\textwidth,page = 2]{figures/fig8/sigma_delta_old_densities.pdf}   
  \caption{Comparison of the resolution between dPeak, Gem, Mace and
    Peakzilla methods with \emph{E.Coli} aerobic experiment from the
    first biosample A) Replicate 1 and B) Replicate 2. Resolution is
    defined as the minimum distance between a RegulonDB annotation and
    a predicted binding event. C) $\delta$ parameter in dPeak measures
    average distance of the reads to their respective binding site. In
    ChIP-exo data, reads were located much closer to the binding site
    than in SET ChIP-Seq. D) $\sigma$ parameter measure the dispersion
    of reads around each binding site. In ChIP-exo data, reads showed
    less variation around the their respective binding sites compared
    to SET ChIP-Seq.}
  \label{fig:methods_comp}
\end{figure}

\newpage

\begin{figure}[h]
  \centering
  \includegraphics[width = .9\textwidth]{figures/fig9/saturation_analysis_old.pdf}
  \caption{Comparison of the number of A) candidate regions, B)
    predicted events, C) identified targets and D) resolution among
    ChIP-exo, PE ChIP-Seq and SE ChIP-Seq. RegulonDB annotations are
    considered as a gold standard. A gold standard binding events was
    deemed identified if a binding event was estimated at a $\pm$ 15
    vicinity of it.}
  \label{fig:design}
\end{figure}

\newpage
\begin{figure}[h!]
  \centering
  \includegraphics[width = .46\textwidth]{figures/fig10/sensitivity_exo_olda_data.pdf}
  \includegraphics[width = .46\textwidth]{figures/fig10/resolution_by_dataset_old_data.pdf}
  \caption{Comparison of A) sensitivity and B) resolution between
    ChIP-exo and ChIP-Seq data. Sensitivity is defined as the
    proportion of RegulonDB annotations identified using each
    data. Resolution is defined as the distance between RegulonDB
    annotation and its closest prediction.}
  \label{fig:reso_all}
\end{figure}

\end{document}

% LocalWords:  Ntimes overlaping gaussian repk

### 2016-07-05 paper/chipexo_QC_vs2.tex
previous version of the paper

--- 2016-07-05 paper/chipexo_QC_vs1_SK.tex
@@
 \usepackage{framed}
+\usepackage{xr}
+\externaldocument{supplement}
+
+\usepackage{verbatim}
+\usepackage{gensymb}
+\usepackage{listings}
 
@@
 \newcommand{\SK}[1]{\textcolor{red}{SK: #1}}
+\newcommand{\RW}[1]{\textcolor{blue}{RW: #1}}
 
@@
   \begin{abstract}
-ChIP-exo/nexus experiments present  modifications on the commonly used ChIP-seq protocol
-    for high
-    resolution mapping of transcription factor binding sites. Although
-    many aspects of the ChIP-exo data analysis are similar to those of
-    ChIP-seq, these high throughput experiments present a number of unique quality control and analysis challenges. We
-    develop a statistical quality control pipeline and accompanying \texttt{R} package, \pname{}, 
-    to enable exploration and analysis of ChIP-exo and related experiments. \pname{}  evaluates a number of key
-    issues including strand imbalance, library complexity, and
-    signal enrichment of data from ChIP-exo/nexus experiments. Assessment of these properties  are facilitated
-    through diagnostic plots and summary statistics calculated over
-    regions of the genome with varying levels of coverage.
+    ChIP-exo/nexus experiments present modifications on the commonly
+    used ChIP-seq protocol for high resolution mapping of
+    transcription factor binding sites. Although many aspects of the
+    ChIP-exo data analysis are similar to those of ChIP-seq, these
+    high throughput experiments present a number of unique quality
+    control and analysis challenges. We develop a statistical quality
+    control pipeline and accompanying \texttt{R} package, \pname{}, to
+    enable exploration and analysis of ChIP-exo and related
+    experiments. \pname{} evaluates a number of key issues including
+    strand imbalance, library complexity, and signal enrichment of
+    data from ChIP-exo/nexus experiments. Assessment of these
+    properties are facilitated through diagnostic plots and summary
+    statistics calculated over regions of the genome with varying
+    levels of coverage.
     %\pname pipeline explores and quantifies these aspects by partitioning the experiment reads into a collection of regions, calculating a series of summary statistics for each region, providing visualizations and calculating measures to globally asses t quality of a ChIP-exo experiment.
 
-    We evaluated our QC pipeline with both  large collections of public ChIP-exo/nexus data and 
-    multiple, new ChIP-exo datasets from \textit{E. Coli}.  \pname{} analysis of these datasets resulted in  guidelines  for
-   using these QC metrics across a wide range of sequencing depths 
-   and further insights for modeling ChIP-exo data. %In addition, we demonstrate that the
+    We evaluated our QC pipeline with both large collections of public
+    ChIP-exo/nexus data and multiple, new ChIP-exo datasets from
+    \textit{E. Coli}.  \pname{} analysis of these datasets resulted in
+    guidelines for using these QC metrics across a wide range of
+    sequencing depths and further insights for modeling ChIP-exo
+    data. %In addition, we demonstrate that the
     %ChIP-exo QC pipeline it is also applicable to ChIP-nexus data,
@@
     %ChIP-exo experiments under similar conditions.
-    Finally, although ChIP-exo experiments have been compared to ChIP-seq experiments with single-end (SE) sequencing, we provide, for the first time, comparisons with paired-end (PE)  ChIP-seq experiments.
+    Finally, although ChIP-exo experiments have been compared to
+    ChIP-seq experiments with single-end (SE) sequencing, we provide,
+    for the first time, comparisons with paired-end (PE) ChIP-seq
+    experiments.
     %We compared ChIP-exo with Paired End (PE) and Single End (SE)
@@
 %    that 
-    We illustrate that, at fixed sequencing depths, ChIP-exo provides higher
-    sensitivity, specificity, and spatial resolution than PE
-    ChIP-seq and both significantly outperform their SE ChIP-seq counterpart. Furthermore, we show that ChIP-exo and PE ChIP-seq are
-    comparable in sensitivity for closely located binding events, but
-    as the average distance between binding events increases, ChIP-exo
-    exhibits higher sensitivity than PE ChIP-Seq. \SK{The last sentence does not really make much sense? }
+    We illustrate that, at fixed sequencing depths, ChIP-exo provides
+    higher sensitivity, specificity, and spatial resolution than PE
+    ChIP-seq and both significantly outperform their SE ChIP-seq
+    counterpart. Furthermore, we show that ChIP-exo and PE ChIP-seq
+    are comparable in sensitivity for closely located binding events,
+    but as the average distance between binding events increases,
+    ChIP-exo exhibits higher sensitivity than PE ChIP-Seq.
+
+    \SK{The last sentence does not really make much sense? }, \RW{I
+      meant that when we have a peak composed by more than one binding
+      event, as the mean distance between binding events increases
+      then ChIP-exo is more likely to identify all the binding events,
+      I used sensitivity because we defined it as the proportion of
+      identified binding events in a given peak. I think it depends if
+      we decide to add Fig8 in the main article or in the
+      supplement.}
 
@@
 
-Chromatin Immunoprecipitation followed by exonuclease
-digestion and next generation sequencing (ChIP-exo) 
-is currently one of the state-of-the-art high throughput assays for profiling protein-DNA interactions at or close to
-single base-pair resolution \cite{exo1}. It presents a powerful alternative to popularly used ChIP-seq
-(Chromatin immunoprecipitation coupled with next generation
+Chromatin Immunoprecipitation followed by exonuclease digestion and
+next generation sequencing (ChIP-exo) is currently one of the
+state-of-the-art high throughput assays for profiling protein-DNA
+interactions at or close to single base-pair resolution
+\cite{exo1}. It presents a powerful alternative to popularly used
+ChIP-seq (Chromatin immunoprecipitation coupled with next generation
 sequencing) assay. ChIP-exo experiments first capture millions of DNA
 fragments ($150$ - $250$ bps in length) that the protein under study
-interacts with using  a
-protein-specific antibody and random fragmentation of DNA. Then, $\lambda$-exonuclease ($\lambda$-exo) is deployed to trim the
-$5^{\prime}$ end of each DNA fragment to each protein-DNA interaction boundary. This step is unique to ChIP-exo and
-aims to achieve  significantly higher spatial resolution
-compared to ChIP-seq. Finally, high throughput sequencing of a small
-region ($36$ to $100$ bps) at the $5^{\prime}$ end of each fragment generates
-millions of reads. 
- Similarly, ChIP-nexus (Chromatin Immunoprecipitation followed by exonuclease
-digestion, unique barcode, single ligation and next generation
-ligation)  \cite{chipnexus} is a further modification on the
-ChIP-exo protocol. ChIP-nexus aims to overcome limitations of ChIP-exo by 
-by yielding high complexity libraries with numbers of cells comparable to that of ChIP-seq experiments.
-This is achieved by reducing the numbers of ligations in the standard ChIP-exo protocol from two to one, and adding unique, randomized barcodes to adoptors to enable monitoring of overamplification.
+interacts with using a protein-specific antibody and random
+fragmentation of DNA. Then, $\lambda$-exonuclease ($\lambda$-exo) is
+deployed to trim the $5^{\prime}$ end of each DNA fragment to each
+protein-DNA interaction boundary. This step is unique to ChIP-exo and
+aims to achieve significantly higher spatial resolution compared to
+ChIP-seq. Finally, high throughput sequencing of a small region ($36$
+to $100$ bps) at the $5^{\prime}$ end of each fragment generates
+millions of reads.  Similarly, ChIP-nexus (Chromatin
+Immunoprecipitation followed by exonuclease digestion, unique barcode,
+single ligation and next generation ligation) \cite{chipnexus} is a
+further modification on the ChIP-exo protocol. ChIP-nexus aims to
+overcome limitations of ChIP-exo by yielding high complexity libraries
+with numbers of cells comparable to that of ChIP-seq experiments. This
+is achieved by reducing the numbers of ligations in the standard
+ChIP-exo protocol from two to one, and adding unique, randomized
+barcodes to adaptors to enable monitoring of
+overamplification. Figure~\ref{fig:1}A illustrates the differences between
+different ChIP-based protocols: ChIP-exo, single-end (SE) ChIP-seq,
+paired-end (PE) ChIP-seq, ChIP-nexus. The $5^{\prime}$ ends of a
+ChIP-exo/nexus experiment are clustered more tightly around the
+binding sites of the protein than in a ChIP-seq experiment. In a PE
+ChIP-seq experiment, both ends are sequenced as opposed to only the
+$5^{\prime}$ end in a SE ChIP-seq.
+
+% \RW{This modification makes a library generated
+%   using the ChIP-nexus protocol to be larger than one generated with
+%   ChIP-exo, which }
 %where both sequencing adaptors are ligated at the
@@
 Compared to ChIP-exo, ChIP-nexus aims to   attain  higher resolution analysis while yielding higher complexity libraries.
-Figure~\ref{fig:chip_diagram} illustrates the
-differences between different ChIP-based protocols: ChIP-exo, single-end (SE) ChIP-seq, paired-end
-(PE) ChIP-seq, ChIP-nexus. The $5^{\prime}$ ends of a ChIP-exo/nexus experiment are clustered
-more tightly around the binding sites of the protein than in a ChIP-seq
-experiment. In a PE ChIP-seq experiment, both ends are sequenced as opposed to only the $5^{\prime}$ end
-in a SE ChIP-seq.
-\SK{Note to Rene: ChIP-Nexus paper has a good description of what can go wrong with ChIP-exo; It seems like most of the read imbalance could also be due to ligation inefficiency.
-Although this probably does not explain why we have reads only from one strand in some regions.
-This is more likely explained by over digestion in one strand or single stranded TF-DNA interaction.
-Think a bit more on these issues and understand how the ligation could give rise to strand imbalance, we can ask Bob to have a closer look at the section where we have this discussion.}
-
-Although ChIP-exo/nexus protocols are being adopted by  research community, features of 
-ChIP-exo data, especially those pertaining to data quality,  have not been investigated much. The key features of ChIP-exo/nexus that separate them from ChIP-seq are broadly as follows. First, 
-DNA libraries
-generated by the ChIP-exo protocol are expected to be less complex than the
-libraries generated by ChIP-seq \cite{exo_review})  because digestion by $\lambda$-exo   is expected to restrict the space of genomic positions that sequencing reads can map to, to small local regions around the actual binding sites.
-Therefore, in  high quality and especially deeply sequenced ChIP-exo datasets, it is possible to observe large numbers of reads accumulating at a small number of bases due to actual signal rather than overamplification bias as commonly observed in ChIP-seq experiments.
-Second, although we expect approximately the same
-numbers  of reads from  both DNA strands at a given binding site,  there may be locally more reads in
-one strand than in the other, owing to $\lambda$-exo efficiency, ligation efficiency, or other factors. 
-This is a key point with implications on the statistical analysis of ChIP-exo data. Specifically, 
-currently available ChIP-exo specific statistical analysis methods   (e.g., 
- Mace \cite{mace}, CexoR \cite{cexor}, and   Peakzilla \cite{peakzilla}) rely on the existence of 
-peak-pairs formed by forward and reverse strand reads at the binding site. 
-Finally, most of current widely used ChIP-seq
-quality control (QC) guidelines  \cite{encode_qc}
-may not be directly applicable to ChIP-exo data.
-
-To address these challenges, we develop 
-a suite of diagnostic plots and summary statistics and implement them in a versatile \texttt{R package} named \pname{}.
-We apply this pipeline on  a large collection of public and newly generated ChIP-exo/nexus data.
-We validate implications of the QC pipeline  by  evaluation of the samples for features that capture high signal to noise such as occurrences of   motifs recognized by the profiled DNA interacting protein.
-Our analyses of this large collection of data revealed that  the so-called ChIP-exo peak-pair assumption is subject to violations.
-To further address this and provide a platform where ChIP-exo and  ChIP-seq experiments can be evaluated with comparable methods, we assess performances of recently developed methods suitable for ChIP-exo analysis, including dpeak \cite{dpeak} and GEM \cite{gem}.
-We observe that dPeak performs as good or better than the available ChIP-exo methods and provides a platform where PE and SE ChIP-seq can be compared with their ChIP-exo counterpart.
-Our comparisons of PE ChIP-seq with ChIP-exo interestingly highlights that while ChIP-exo outperforms PE ChIP-seq in terms of resolution and detection power, both are significantly better than SE ChIP-seq.
-
-% while there are not established QC pipelines for ChIP-exo; previous ChIP-exo analyses used ChIP-Seq samples to compare the resolution between experiments \cite{exo1,exoillumina,exo2}.
+
+\SK{Note to Rene: ChIP-nexus paper has a good description of what can
+  go wrong with ChIP-exo; It seems like most of the read imbalance
+  could also be due to ligation inefficiency. Although this probably
+  does not explain why we have reads only from one strand in some
+  regions. This is more likely explained by over digestion in one
+  strand or single stranded TF-DNA interaction. Think a bit more on
+  these issues and understand how the ligation could give rise to
+  strand imbalance, we can ask Bob to have a closer look at the
+  section where we have this discussion.}
+
+% \RW{1 - I think it may come from the starting material in each side. I
+%   think that at least the number of fragments on each side must
+%   preserve that if there is more starting reads in one side, then this
+%   must be the case after the amplification.}
+
+% \RW{2 - There is not control about how the enzyme digests the
+%   fragments, as far as we know it starts on the $5\prime$ side and
+%   stops until it finds the TF or it runs out of energy to keep
+%   digesting. I think Bob mentioned that it may be possible for the
+%   enzyme to pass though the TF if it is left digesting for a longer
+%   time, but if that's the case then the question may be why it
+%   preffers one strand over the other, which may lead to the GC-bias
+%   but as far as I can tell is very hard to observe that because that
+%   bias appear also in ChIP-seq (and it seeems that also appear in the
+%   circularization step from ChIP-nexus).}
+
+% \RW{3 - I think that the regions formed by fragments in one strand,
+%   may be copies of few undigested fragments, in a good experiment the
+%   fragments are expected to be clustered around the binding sites,
+%   therefore it is less likely to observe this fragments scattered in
+%   the genome. In a bad experiment is the opposite case, so the
+%   fragments are not clustered so if the enzyme doesn't have a strong
+%   preference to how is attracted to the fragments then the remaining
+%   ones can be seen as random positions around the genome, this
+%   positions are copied several times during the amplification and that
+%   may be the reason why we observe something as scattered. Perhaps if
+%   the enzyme was left for longer time, then the starting material that
+%   later was amplified into this regions would have been digested and
+%   some of this regions wouldn't appear (I am thinking about the deeply
+%   sequenced TBP experiments.)}
+
+Although ChIP-exo/nexus protocols are being adopted by research
+community, features of ChIP-exo data, especially those pertaining to
+data quality, have not been investigated much. The key features of
+ChIP-exo/nexus that separate them from ChIP-seq are broadly as
+follows. First, DNA libraries generated by the ChIP-exo protocol are
+expected to be less complex than the libraries generated by ChIP-seq
+\cite{exo_review}) because digestion by $\lambda$-exo is expected to
+restrict the space of genomic positions that sequencing reads can map
+to, to small local regions around the actual binding sites. Therefore,
+in high quality and especially deeply sequenced ChIP-exo datasets, it
+is possible to observe large numbers of reads accumulating at a small
+number of bases due to actual signal rather than overamplification
+bias as commonly observed in ChIP-seq experiments.  Second, although
+we expect approximately the same numbers of reads from both DNA
+strands at a given binding site, there may be locally more reads in
+one strand than in the other, owing to $\lambda$-exo efficiency,
+ligation efficiency, or other factors.  This is a key point with
+implications on the statistical analysis of ChIP-exo
+data. Specifically, currently available ChIP-exo specific statistical
+analysis methods (e.g., Mace \cite{mace}, CexoR \cite{cexor}, and
+Peakzilla \cite{peakzilla}) rely on the existence of peak-pairs formed
+by forward and reverse strand reads at the binding site.  Finally,
+most of current widely used ChIP-seq quality control (QC) guidelines
+\cite{encode_qc} may not be directly applicable to ChIP-exo data.
+
+To address these challenges, we develop a suite of diagnostic plots
+and summary statistics and implement them in a versatile \texttt{R
+  package} named \pname{}.  We apply this pipeline on a large
+collection of public and newly generated ChIP-exo/nexus data. We
+validate implications of the QC pipeline by evaluation of the samples
+for features that capture high signal to noise such as occurrences of
+motifs recognized by the profiled DNA interacting protein. Our
+analyses of this large collection of data revealed that the so-called
+ChIP-exo peak-pair assumption is subject to violations. To further
+address this and provide a platform where ChIP-exo and ChIP-seq
+experiments can be evaluated with comparable methods, we assess
+performances of recently developed methods suitable for ChIP-exo
+analysis, including dpeak \cite{dpeak} and GEM \cite{gem}. We observe
+that dPeak performs as good or better than the available ChIP-exo
+methods and provides a platform where PE and SE ChIP-seq can be
+compared with their ChIP-exo counterpart. Our comparisons of PE
+ChIP-seq with ChIP-exo interestingly highlights that while ChIP-exo
+outperforms PE ChIP-seq in terms of resolution and detection power,
+both are significantly better than SE ChIP-seq.
+
+% while there are not established QC pipelines for ChIP-exo; previous
+% ChIP-exo analyses used ChIP-Seq samples to compare the resolution
+% between experiments \cite{exo1,exoillumina,exo2}.
 
@@
 \subsection*{Publicly available ChIP-exo/nexus and novel \textit{E. coli} ChIP-seq/exo datasets}
-We utilized a rich collection of publicly available ChIP-exo/nexus data from multiple organisms to build and evaluate our quality control pipeline (Table~\ref{tab:qc}). These include:   CTCF factor
-in human HeLa cells \cite{exo1}; ER factor in human MCF-7
-cells \cite{exoillumina}; GR factor in IMR90, K562, and U2OS human
-cells \cite{starick15}; TBP factor in human K562 cells
-\cite{venters13}. 
+We utilized a rich collection of publicly available ChIP-exo/nexus
+data from multiple organisms to build and evaluate our quality control
+pipeline (Table~\ref{tab:qc}). These include: CTCF factor in human
+HeLa cells \cite{exo1}; ER factor in human MCF-7 cells
+\cite{exoillumina}; GR factor in IMR90, K562, and U2OS human cells
+\cite{starick15}; TBP factor in human K562 cells \cite{venters13}.
 ChIP-nexus data included experiments from \cite{chipnexus} profiling
-TBP in human K562 cells, MyC and Max in \emph{D. Melanogaster} S2 cells, and Twist and
-Dorsal in \emph{D. Melanogaster} embryo.
-
-In order to have a setting where we can compare SE and PE ChIP-seq with their ChIP-exo counterpart, we profiled 
-$\sig$ under a variety of  conditions in \textit{E. coli} with both ChIP-exo (Table~\ref{tab:qc_sig}) and PE ChIP-seq (\SK{Table SXX}). Collectively, we 
- generated $\sig$ factor ChIP-exo, PE and SE ChIP-seq experiments  under aerobic ($+O_2$) and anaerobic ($-O_2$)
-conditions in glucose minimal media. Similarly, we generated $\sig$ factor ChIP-exo, PE and
-SE (generated \emph{in silico}) ChIP-seq experiments in \emph{E. Coli}
-under aerobic ($+O_2$) conditions with  and without  rifampicin  treatment.
-We refer to these latter set of experiments  as time point 20 minutes and time point 0 minutes \SK{Check for consistency with other parts of the paper and correct, ok to use different naming convention}.
-We specifically used these experimental data  for comparisons of ChIP-exo and PE
-ChIP-seq assays in identifying closely spaced binding events and their resolution, i.e., physical proximity of the predicted events to the actual binding sites.
-This comparison benefits from using $\sig$  which is a transcription initiation factor of housekeeping genes
-in \emph{E. Coli}. Many \textit{E. coli} promoters contain
-multiple transcription start sites (TSS). These TSSs are often
-closely spaced, i.e., within 10 $\sim$ 150 bps of each other, and are considered  to be multiple ``switches'' that differentially
-regulate gene expression under diverse growth conditions
-\cite{regulondb}.
+TBP in human K562 cells, MyC and Max in \textit{D. melanogaster} S2
+cells, and Twist and Dorsal in \textit{D. melanogaster} embryo.
+
+In order to have a setting where we can compare SE and PE ChIP-seq
+with their ChIP-exo counterpart, we profiled $\sig$ under a variety of
+conditions in \textit{E. coli} with both ChIP-exo and PE ChIP-seq
+(Table~\ref{tab:qc_sig}). Collectively, we generated $\sig$ factor
+ChIP-exo, PE and SE ChIP-seq experiments under aerobic ($+O_2$) and
+anaerobic ($-O_2$) conditions in glucose minimal media, for simplicity
+we named these experiments as E1, P1 and S1 respectively. Similarly,
+we generated $\sig$ factor ChIP-exo, PE and SE (generated \textit{in
+  silico}) ChIP-seq experiments in \textit{E. coli} under aerobic
+($+O_2$) conditions with and without rifampicin treatment.  We refer
+to these latter set of experiments as time point 20 minutes and time
+point 0 minutes, we also named these experiments E2, P2 and S2 for
+ChIP-exo, PE ChIP-seq and SE ChIP-seq respectively.
+% \SK{Check for
+%   consistency with other parts of the paper and correct, ok to use
+%   different naming convention}.
+We specifically used these experimental data for comparisons of
+ChIP-exo and PE ChIP-seq assays in identifying closely spaced binding
+events and their resolution, i.e., physical proximity of the predicted
+events to the actual binding sites.  This comparison benefits from
+using $\sig$ which is a transcription initiation factor of
+housekeeping genes in \textit{E. coli}. Many \textit{E. coli}
+promoters contain multiple transcription start sites (TSS). These TSSs
+are often closely spaced, i.e., within 10 $\sim$ 150 bps of each
+other, and are considered to be multiple ``switches'' that
+differentially regulate gene expression under diverse growth
+conditions \cite{regulondb}.
 
@@
 
-\textit{Read distributions within signal and background regions.}
-We first compared ChIP-seq and ChIP-exo in terms of data features that are well studied in ChIP-seq studies.
-Our $\sig$  ChIP-seq and ChIP-exo samples from \textit{E.
-coli} are especially well suited for this task since they are all deeply sequenced compared to the genome size of \textit{E.
-coli}.
-Figure~\ref{fig:comp} summarizes this comparison for one biological replicate  of ChIP-exo and ChIP-seq experiments from the same biological conditions (samples x and y from Table~\ref{tab:qc_sig} \SK{Maybe add a column as sample number to this table to make referencing easy}).
-Comparisons with other paired \textit{E.
-coli} ChIP-seq and ChIP-exo  samples led to similar conclusions (Supplementary Figure \SK{SXXX}).
+ % \SK{Maybe add a column as sample
+ %  number to this table to make referencing easy}
+\textit{Read distributions within signal and background regions.} We
+first compared ChIP-seq and ChIP-exo in terms of data features that
+are well studied in ChIP-seq studies.  Our $\sig$ ChIP-seq and
+ChIP-exo samples from \textit{E. coli} are especially well suited for
+this task since they are all deeply sequenced compared to the genome
+size of \textit{E. coli}.  Figure~\ref{fig:1} summarizes this
+comparison for one biological replicate of ChIP-exo and ChIP-seq
+experiments from the same biological conditions (samples EB1-1 and
+P1-1 from Table~\ref{tab:qc_sig}). Comparisons with other paired
+\textit{E. coli} ChIP-seq and ChIP-exo samples led to similar
+conclusions (Supplementary Figure~\ref{sfig:comp1}).
 %compared various factors that could affect binding site
@@
 %and a SE ChIP-Seq replicate grown under the same conditions. 
-We summarized the extended raw read counts within $150$ bps non-overlapping intervals, i.e., bins, interrogating the genome.
+We summarized the extended raw read counts within $150$ bps
+non-overlapping intervals, i.e., bins, interrogating the genome.
 %In order
@@
 %partition of the genome into non-overlapping bins. 
-Figure~\ref{fig:comp}(A) depicts that, as expected,  ChIP read counts from ChIP-exo and ChIP-seq are linearly 
-correlated especially at high read counts. This indicates that signals for potential binding
-sites are well reproducible between ChIP-exo and ChIP-seq data. 
-In contrast, there is a clear difference
-among the two data types for bins with low read counts, highlighting expected differences in the  background read distributions 
-of the two data types. 
-\SK{***After seeing these types of plots from other samples (e..g., ER in MCF7, it is unlikely that what we are seeing from $\sig$ samples regarding background reads (the next few sentences) is typical.
-For CTCF and $\sig$, the ChIP-exo samples have a wider dynamic range, but this could be due to differences in sequencing depth.
-*** Specifically, background reads from ChIP-seq are  almost
-uniformly distributed over background (non-enriched) regions.
-In comparison, background regions in ChIP-exo show larger variation but have overall lower read counts.
-Furthermore, there are considerably large numbers of bins with zero read counts in ChIP-exo and non-zero read counts in ChIP-seq.
+Figure~\ref{fig:1}B depicts that, as expected, ChIP read counts
+from ChIP-exo and ChIP-seq are linearly correlated especially at high
+read counts. This indicates that signals for potential binding sites
+are well reproducible between ChIP-exo and ChIP-seq data.  In
+contrast, there is a clear difference among the two data types for
+bins with low read counts, highlighting expected differences in the
+background read distributions of the two data types.  \SK{***After
+  seeing these types of plots from other samples (e..g., ER in MCF7,
+  it is unlikely that what we are seeing from $\sig$ samples regarding
+  background reads (the next few sentences) is typical. For CTCF and
+  $\sig$, the ChIP-exo samples have a wider dynamic range, but this
+  could be due to differences in sequencing depth. *** Specifically,
+  background reads from ChIP-seq are almost uniformly distributed over
+  background (non-enriched) regions. In comparison, background regions
+  in ChIP-exo show larger variation but have overall lower read
+  counts. Furthermore, there are considerably large numbers of bins
+  with zero read counts in ChIP-exo and non-zero read counts in
+  ChIP-seq.
 %and the
@@
 %large proportion of regions without any read in ChIP-exo data. 
-Overall, these observations indicate that  a much smaller portion of the
-genome is expected to be background in ChIP-exo compared to ChIP-seq and 
- methods that
-specifically model the background read distribution might benefit from acknowledging this.}
+  Overall, these observations indicate that a much smaller portion of
+  the genome is expected to be background in ChIP-exo compared to
+  ChIP-seq and methods that specifically model the background read
+  distribution might benefit from acknowledging this.} \RW{Figure 2B
+  from the chip-nexus paper makes a similar comparison but specific to
+  peaks, i.e. our argument about both ChIP-seq and ChIP-exo is
+  correlated in high signal region still holds.}
 
 \textit{Peak-pair assumption.} We next evaluated the peak-pair
-assumption, i.e., a cluster of reads
-in the forward strand is usually paired with a cluster of reads in the
-reverse strand that is located on the right-hand-side of  the binding
-site, that is commonly utilized in designing statistical analysis methods for ChIP-exo data \cite{mace, cexor, peakzilla}.
-We considered the set of peaks identified in both the ChIP-seq and ChIP-exo samples as high quality peaks (Materials and Methods) and calculated the proportion of forward strand reads in these regions (Figure~\ref{fig:comp}(B)).
-This plot reveals a higher level of strand imbalance for ChIP-exo compared to ChIP-seq.
-%In order to evaluate this assumption, we reviewed the proportion of reads in the forward strand in high quality ChIP-exo peaks such as at least one binding site is predicted in both ChIP-exo and ChIP-Seq data.
-%We
-%found that strands of reads were much less balanced in ChIP-exo data
-%than in ChIP-Seq data in these regions with potential binding sites
-%(Fig. \ref{fig:comp}B) and this indicates that the peak pair
-%assumption might not hold in ChIP-exo data. 
-Potential reasons for this observation include ligation efficiency, efficiency of $\lambda$-exo digestion, 
-and single-stranded protein-DNA interactions. Overall, such an imbalance is prevalent in  \SK{X}\% of the ChIP-exo samples used in this paper.
+assumption, i.e., a cluster of reads in the forward strand is usually
+paired with a cluster of reads in the reverse strand that is located
+on the right-hand-side of the binding site, that is commonly utilized
+in designing statistical analysis methods for ChIP-exo data
+\cite{mace, cexor, peakzilla}.  We considered the set of peaks
+identified in both the ChIP-seq and ChIP-exo samples as high quality
+peaks (Materials and Methods) and calculated the proportion of forward
+strand reads in these regions (Figure~\ref{fig:1}C). This plot
+reveals a higher level of strand imbalance for ChIP-exo compared to
+ChIP-seq.
+% In order to evaluate this assumption, we reviewed the proportion of
+% reads in the forward strand in high quality ChIP-exo peaks such as
+% at least one binding site is predicted in both ChIP-exo and ChIP-Seq
+% data.  We found that strands of reads were much less balanced in
+% ChIP-exo data than in ChIP-Seq data in these regions with potential
+% binding sites (Fig. \ref{fig:comp}B) and this indicates that the
+% peak pair assumption might not hold in ChIP-exo data.
+Potential reasons for this observation include ligation efficiency,
+efficiency of $\lambda$-exo digestion, and single-stranded protein-DNA
+interactions. Overall, such an imbalance is prevalent in 85\% of the
+ChIP-exo samples used in this paper \RW{(Supplement
+  Figures~\ref{sfig:comp1} to \ref{sfig:comp3})}. \RW{Note: The only ChIP-exo
+  experiment were it seems that ChIP-exo is not more imbalanced than
+  ChIP-seq is Carroll's ER dataset.}
 %the enzyme digestion or the strength at which the protein binds to the
@@
 
-\textit{Mappability and GC-content bias.} We next evaluated ChIP-exo data of CTCF in HeLa cells
-\cite{exo1} to investigate biases inherent to next generation sequencing experiments with eukaryotic genomes. 
-Figures \ref{fig:comp}(C) and
-\ref{fig:comp}(D) display the bin-level average read counts against
-mappability and GC-content. Each data point is obtained by averaging
-the read counts across bins with the same mappability of GC-content. 
-These biases, increasing linear trend with mappability and non-linear trend with GC-content, are similar to those observed in ChIP-seq datasets \cite{benjamini2011, mosaics,quest}.
+\textit{Mappability and GC-content bias.} We next evaluated ChIP-exo
+data of CTCF in HeLa cells \cite{exo1} to investigate biases inherent
+to next generation sequencing experiments with eukaryotic genomes.
+Figures \ref{fig:1}E and \ref{fig:1}F display the bin-level
+average read counts against mappability and GC-content. Each data
+point is obtained by averaging the read counts across bins with the
+same mappability of GC-content.  These biases, increasing linear trend
+with mappability and non-linear trend with GC-content, are similar to
+those observed in ChIP-seq datasets \cite{benjamini2011,
+  mosaics,quest}.
 %In Figure \ref{fig:comp}C it is shown that the ChIP-exo tag
@@
 
-\subsection*{Application of ENCODE ChIP-seq quality metrics   to ChIP-exo and
-  ChIP-nexus data}
-
-ENCODE consortium established empirical and widely used QC metrics on ChIP-seq data
-\cite{encode_qc}. Currently, these constitute the state-of-the-art QC pipelines for these high throughput experiments.
+\subsection*{Application of ENCODE ChIP-seq quality metrics to
+  ChIP-exo and ChIP-nexus data}
+
+ENCODE consortium established empirical and widely used QC metrics on
+ChIP-seq data \cite{encode_qc}. Currently, these constitute the
+state-of-the-art QC pipelines for these high throughput experiments.
 We evaluated how these metrics, namely Normalized Strand
-Cross-Correlation (NSC), Relative Strand Cross-Correlation (RSC), and PCR
-Bottleneck Coefficient (PBC) defined at 
-\url{https://genome.ucsc.edu/ENCODE/qualityMetrics.html} \cite{encode_qc}, behave
- on ChIP-exo/nexus data in  Tables~\ref{tab:qc} and \ref{tab:qc_sig}.
+Cross-Correlation (NSC), Relative Strand Cross-Correlation (RSC), and
+PCR Bottleneck Coefficient (PBC) defined at
+\url{https://genome.ucsc.edu/ENCODE/qualityMetrics.html}
+\cite{encode_qc}, behave on ChIP-exo/nexus data in Tables~\ref{tab:qc}
+and \ref{tab:qc_sig}.
 
@@
   \centering
-  \begin{tabu} to\linewidth{X[1.3]|X[2.2]|X[-1]|X[1.3]|X[-1,m,c]|X[1.8]|X[-1]|X[-1]|X[-1]}
+  \begin{tabu} to\linewidth{X[1.3]|X[2.3]|X[-1]|X[1.3]|X[-1,m,c]|X[1.8]|X[-1]|X[-1]|X[-1]}
     \firsthline
@@
   \end{tabu}
-  \caption{Summary of publicly available data used for development and evaluation of \pname{}. The last three columns depict ENCODE QC metrics on these data:  NSC:  Normalized 
-    Strand Cross-Correlation; RSC: Relative Strand Cross-Correlation;  PBC:  
-    PCR Bottleneck Coefficient.}  
+  \caption{Summary of publicly available data used for development and evaluation of \pname{}.
+The last three columns depict ENCODE QC metrics on these data:  NSC:  Normalized Strand Cross-Correlation; RSC: Relative Strand Cross-Correlation; PBC: PCR Bottleneck Coefficient.}
 \label{tab:qc}
@@
   \centering
-  \begin{tabu} to\linewidth{X[-1,m,c]|X[-1]|X[-1]|X[-1,m,c]|X[-1]|X[-1]|X[-1]|X[-1]}
+  \begin{tabu} to\linewidth{X[1.4,m,c]|X[1.2]|X[1.3,m,c]|X[.4,m,c]|X[.6,m,c]|X[1.2]|X[.6]|X[.6]|X[.5]}
     \firsthline
-    \textbf{Bio Sample} & \textbf{Condition} & \textbf{Treatment} & \textbf{Rep.} & \textbf{Depth} & \textbf{NSC} & \textbf{RSC} & 
-    \textbf{PBC}\\
+    \textbf{Group} & \textbf{Growth} & \textbf{Treatment} & \textbf{Rep.} & \textbf{Id.} &
+    \textbf{Depth} & \textbf{NSC} & \textbf{RSC} & 
+    \textbf{PBC}  \\
     \hline 
-    \multirow{4}{1em}{1} & Aerobic & No Rif. & 1 & 13,961,493 & 103.15 & 2.0193 & 0.1399 \\
-    & Aerobic & No Rif. & 2 & 14,810,838 & 162.70 & 1.7805 & 0.1633 \\
-    & Anaerobic & No Rif. & 1 &  16,108,774 & 153.51 & 1.8035 & 0.1353 \\
-    & Anaerobic & No Rif. & 2 &  13,636,541 & 172.59 & 2.014 & 0.1532 \\ 
+    &  Exp. $+O_2$ & No Rif. & 1 & 1 & 13,961,493 & 103.15 & 2.0193 & 0.1399 \\
+    ChIP-exo  & Exp. $+O_2$ & No Rif. & 2 & 2 &14,810,838 & 162.70 &
+    1.7805 & 0.1633 \\
+    (E1)   & Stat. $+O_2$ & No Rif. & 1 & 3& 16,108,774 & 153.51 &
+    1.8035 & 0.1353 \\
+    & Stat. $+O_2$ & No Rif. & 2 & 4 & 13,636,541 & 172.59 & 2.014 & 0.1532 \\ 
     \hline
-    \multirow{4}{1em}{2} & Aerobic & No Rif. & 1 & 902,921 & 13.77 & 1.1270 & 0.2689\\
-    & Aerobic & No Rif. & 1 & 1,852,124 & 17.91 & 1.5275 & 0.2590\\
-    & Aerobic & Rif. 20 min & 2 & 2,104,427 & 29.60 & 1.2844 & 0.2584\\
-    & Aerobic & Rif. 20 min & 2 & 11,548,572 & 13.08 & 1.5122 & 0.1510 \\
+    ChIP-seq PE & Exp. $+O_2$ & No Rif. &  & 1 & 27,665,432 & 4.01 & 
+    3.9582 & 0.3869 \\
+    (P1) & Exp. $-O_2$ & No Rif. &  & 2 & 44,707,340 & 3.56 & 
+    3.3045 & 0.3134 \\
+    \hline
+    ChIP-seq SE & Exp. $+O_2$ & No Rif. &  & 1 & 7,456,068 & 3.27 & 
+    2.3863 & 0.5629 \\
+    (S1) & Exp. $-O_2$ & No Rif. &  & 2 & 11,467,086 & 2.91 & 
+    2.1362 & 0.5452 \\
+    \hline\hline
+    &  Exp. $+O_2$  & No Rif. & 1 & 1 &902,921 & 13.77 & 1.1270 & 0.2689\\
+    ChIP-exo   & Exp. $+O_2$ & Rif. 20 min & 1 & 2 & 1,852,124 & 17.91 &
+    1.5275 & 0.2590\\
+    (E2)  & Exp. $+O_2$ & No Rif. &  2 & 3 & 2,104,427 & 29.60 & 1.2844 & 
+    0.2584\\
+    & Exp. $+O_2$ & Rif. 20 min & 2 & 4 & 11,548,572 & 13.08 & 1.5122 & 0.1510 \\
+    \hline
+    & Exp. $+O_2$ & No Rif. & 1 & 1 & 13,445,022 & 8.86 & 1.0541 & 0.9426 \\
+    ChIP-seq PE & Exp. $+O_2$ & Rif. 20 min & 1 & 2 & 16,538,920 & 7.03 & 
+    1.0157 & 0.9378 \\
+    (P2) & Exp. $+O_2$ & No Rif. & 2  & 3 & 11,642,722 & 10.77 & 1.0145 &
+    0.8891 \\
+    & Exp. $+O_2$ & Rif. 20 min & 2 & 4 & 16,854,026 & 7.93 &1.0048& 0.9407 \\
+    \hline
+    & Exp. $+O_2$ & No Rif. & 1 & 1 & 6,722,511 & 9.01 & 2.8461 & 0.6632 \\
+    ChIP-seq SE  & Exp. $+O_2$ & Rif. 20 min & 1 & 2 & 8,269,460 & 7.17 &
+    2.5168 & 0.5594 \\
+    (S2)  & Exp. $+O_2$ & No Rif. & 2  & 3 & 5,821,361 & 10.89 & 3.1291 &
+    0.6472 \\
+    & Exp. $+O_2$ & Rif. 20 min & 2 & 4 & 8,427,013 & 8.12 & 2.6908 & 0.5895 \\
     \lasthline
     \end{tabu}
-    \caption{ Summary of the \emph{E. Coli} $\sig$ ChIP-exo and ChIP-seq samples. 
-The last three columns depict ENCODE QC metrics on these data:  NSC:  Normalized 
-    Strand Cross-Correlation; RSC: Relative Strand Cross-Correlation;  PBC:  
-    PCR Bottleneck Coefficient.    }
+    \caption{Summary of the \textit{E. coli} $\sig$ ChIP-exo and ChIP-seq 
+      samples. Exp. stands for exponential and Stat. for stationary growth 
+      conditions. The last three columns depict ENCODE QC metrics on these 
+      data:  NSC:  Normalized Strand Cross-Correlation; RSC: Relative Strand 
+      Cross-Correlation;  PBC: PCR Bottleneck Coefficient.}
 \label{tab:qc_sig}
@@
 DNA libraries generated by ChIP-exo and ChIP-nexus protocols are
-expected to be less complex than the libraries generated by ChIP-seq because
- the numbers of positions to which the reads can 
-align to are reduced  due to the exonuclease digestion. 
+expected to be less complex than the libraries generated by ChIP-seq
+because the numbers of positions to which the reads can align to are
+reduced due to the exonuclease digestion.
 %considerable amounts of reads are being mapped to specific positions.
-This affects the interpretation of the PBC, which is defined as the ratio of the 
-number of genomic positions to which exactly one read maps to   the number of genomic positions to which at least one  read maps.
-For ChIP-seq samples,  low PBC values (e.g., $\le 0.5$) indicate high levels of PCR amplification bias, i.e., PC bottleneck, 
- unless the sequencing depth is high enough to sature all targets of the factor profiled.
-In contrast,  for ChIP-exo/nexus, exonuclease digestion will lead to reads with same exact 
-$5^{\prime}$ end even before the PCR amplification step.
+This affects the interpretation of the PBC, which is defined as the
+ratio of the number of genomic positions to which exactly one read
+maps to the number of genomic positions to which at least one read
+maps.  For ChIP-seq samples, low PBC values (e.g., $\le 0.5$) indicate
+high levels of PCR amplification bias, i.e., PC bottleneck, unless the
+sequencing depth is high enough to sature all targets of the factor
+profiled.  In contrast, for ChIP-exo/nexus, exonuclease digestion will
+lead to reads with same exact $5^{\prime}$ end even before the PCR
+amplification step.
 %because of the exonuclease digestion,
@@
 %SCC global maxima. 
-In an
-idealized ChIP-exo experiment where the DNA fragments are digested
- to  the boundaries of the protein-DNA interaction sites,  we would expect the SCC profile to maximize at the motif length indicating clustering of the forward and reverse strand reads around the  binding site.
- Figure~\ref{fig:scc_exo} displays the SCC curves for the CTCF  HeLa samples where the ChIP-exo curve shows  local maxima at the motif and read
-lengths, while the SE ChIP-seq curves have a local maxima at the read
-length and a global maxima at the average fragment
-length. SCC profiles for other samples  are available in Supplementary Figures \SK{XXX}. The read length and motif length maxima are often in close proximity of each other; as a result, this renders
-QC metrics such as the Normalized
-Strand Cross-Correlation (NSC) or the Relative Strand
-Cross-Correlation (RSC) harder to interpret; however, the profile itself seems informative about the enrichment signal in ChIP-exo/nexus experiments \SK{this seems true at least based on my initial look at the plots, but we should be cautious in reporting this}.
+In an idealized ChIP-exo experiment where the DNA fragments are
+digested to the boundaries of the protein-DNA interaction sites, we
+would expect the SCC profile to maximize at the motif length
+indicating clustering of the forward and reverse strand reads around
+the binding site. Figure~\ref{fig:1}D displays the SCC curves for the
+CTCF HeLa samples where the ChIP-exo curve shows local maxima at the
+motif and read lengths, while the SE ChIP-seq curves have a local
+maxima at the read length and a global maxima at the average fragment
+length. SCC profiles for other samples are available in Supplementary
+Figures \ref{sfig:scc1} to \ref{sfig:scc8}. The read length and motif
+length maxima are often in close proximity of each other; as a result,
+this renders QC metrics such as the Normalized Strand
+Cross-Correlation (NSC) or the Relative Strand Cross-Correlation (RSC)
+harder to interpret; however, the profile itself seems informative
+about the enrichment signal in ChIP-exo/nexus experiments \SK{this
+  seems true at least based on my initial look at the plots, but we
+  should be cautious in reporting this}. \RW{i agree, and additionally
+  Carroll et al, 2014 kinda agrees with us that the SCC-related
+  measure may not be useful since the fragment length coincides with
+  read length}
+
+ % %http://journal.frontiersin.org/article/10.3389/fgene.2014.00075/full#h5
+ %  the paragraph is a little bit above conclusions
 
@@
 
-We first present the overall pipeline and then discuss
-individual components with a case study using ChIP-exo data of FoxA1  
-from \cite{exoillumina} and ChIP-nexus data from \cite{chipnexus}.
-Figure~\ref{fig:qcdiagram} summarizes  the 4-step pipeline. 
-Given aligned reads from a  ChIP-exo/nexus sample, the first step partitions the reference genome into islands by keeping the
-non-digested ChIP-exo regions.  In step 2,   the total
-number of extended reads \SK{check: extended or not? What do we extend to in ChIP-exo? } overlapping each island ($D_i$) and the number of
-unique island positions with at least one aligned read re recorded ($(U_i)$.
-Then,  three summary statistics $\text{ARC}_i$, $\text{URCR}_i$, and $\text{FSR}_i$ are computed for each region $i$.
-$\text{ARC}_i$ denotes the \textit{average read coverage per base pair} and is defined as the ratio of the
-\#  of reads in island $i$ ($D_i$) to the width of the island $i$ ($W_i$);  $\text{URCR}_i$, \textit{unique read coverage ratio}, \SK{Is unique read coverage ratio a good name for this?} quantifies inverse of the effective coverage and is defined as the ratio of the \# of genomic positions with at least one aligned read within island $i$ ($U_i$) to the \# of reads in island $i$ ($D_i$);  and  $\text{FSR}_i$ denotes the proportion of forward strand reads.
+We first present the overall pipeline and then discuss individual
+components with a case study using ChIP-exo data of FoxA1 from
+\cite{exoillumina} and ChIP-nexus data from \cite{chipnexus}.
+Figure~\ref{fig:2} summarizes the 4-step pipeline.  Given
+aligned reads from a ChIP-exo/nexus sample, the first step partitions
+the reference genome into islands by keeping the non-digested ChIP-exo
+regions.  In step 2, the total number of extended reads (by the
+experiment's read length) \SK{check: extended or not? What do we
+  extend to in ChIP-exo?} \RW{right now, we don't extend. we use the
+  coverage plot with the experiment's read length to partition the
+  coverage into the regions. In the package I left it as a parameter
+  but I don't is going to change a lot. For example, the ChIP-exo
+  datasets read length is ~45 and it's regions seems to be composed by
+  scattered reads in both sides. So, I think that using a higher value
+  will only made the regions wider but will not increase the number of
+  unique position or the depth by a lot} overlapping each island
+($D_i$) and the number of unique island positions with at least one
+aligned read re recorded ($(U_i)$. Then, three summary statistics
+$\text{ARC}_i$, $\text{URC}_i$, and $\text{FSR}_i$ are computed for
+each region $i$.  $\text{ARC}_i$ denotes the \textit{average read
+  coefficient} and is defined as the ratio of the \# of reads in
+island $i$ ($D_i$) to the width of the island $i$ ($W_i$);
+$\text{URC}_i$, \textit{unique read coefficient}, quantifies the
+inverse of the effective coverage and is defined as the ratio of the
+\# of genomic positions with at least one aligned read within island
+$i$ ($U_i$) to the \# of reads in island $i$ ($D_i$); and
+$\text{FSR}_i$ denotes the proportion of forward strand reads.
 %\begin{align*}
 %  \text{ARC}_i &= \dfrac{\text{\#  of reads in island $i$}}{\text{Width of the island $i$}}, \\
-%  \text{URCR}_i & = \dfrac{\text{\# of genomic positions with at least one aligned read within island $i$}}{\text{\# of reads in island %$i$}}, \\
-%  \text{FSR}_i &= \dfrac{\text{\# of forward strand reads in island $i$}}{\text{\# of reads in island $i$}}.
+%  \text{URC}_i & = \dfrac{\text{\# of genomic positions with at least
+%      one aligned read within island $i$}}{\text{\# of reads in
+%      island %$i$}}, \\
+%      \text{FSR}_i &= \dfrac{\text{\# of forward strand reads in
+%          island $i$}}{\text{\# of reads in island $i$}}.
 %\end{align*}
-Step 3 of the pipeline generates several diagnostic plots aimed at quantifying ChIP enrichment and strand imbalance and step 4 generates quantitative summaries of these diagnostic plots \SK{Aren't the linear model fit etc part of the pipeline?}
-
-Figure~\ref{fig:qcdiagram}A presents
-the typical behavior of the URCR vs. ARC plot for a high quality ChIP-exo sample. In general, the plot
-depicts two strong arms. 
-High URCR values correspond to regions with reads
-concentrated on a small number of positions. 
-The  left arm, with low ARC  and
-varying URCR values,  corresponds to  background islands, regions that
-are usually composed of scattered reads that were not digested during
-the exonuclease step. The right arm  where the URCR decreases as the
-ARC increases  corresponds to regions that are usually ChIP enriched.
+Step 3 of the pipeline generates several diagnostic plots aimed at
+quantifying ChIP enrichment and strand imbalance and step 4 generates
+quantitative summaries of these diagnostic plots.
+
+Figure~\ref{fig:2}A presents the typical behavior of the URC vs. ARC
+plot for a high quality ChIP-exo sample. In general, the plot depicts
+two strong arms.  High URC values correspond to regions with reads
+concentrated on a small number of positions.  The left arm, with low
+ARC and varying URC values, corresponds to background islands, regions
+that are usually composed of scattered reads that were not digested
+during the exonuclease step. The right arm where the URC decreases as
+the ARC increases corresponds to regions that are usually ChIP
+enriched. \RW{Finally we quantify the shape of the ARC vs. URC plot by
+  the use of two estimated parameters: $\beta_1$ which
+  represents the average number of reads aligned to the unique
+  positions in large depth regions and $\beta_2$ which
+  represents the overall change in depth as the width varies across a
+  large set of regions.}  \RW{Figures~\ref{fig:2}B and \ref{fig:2}C
+  presents the typical behavior of the Region Composition and Forward
+  Strand Ratio distribution plots respectively, both quantify the
+  strand imbalance as part of the QC pipeline. The former depicts how
+  quickly the islands exclusively composed by fragments on a single
+  strand are filtered out as islands with higher depths are
+  observed. In a high quality sample, the proportion of islands with
+  reads from only one strand is expected to decrease rapidly as we
+  consider higher depth regions. In contrast, this proportion remains
+  approximately constant in lower quality samples. The latter plot
+  illustrates how quickly the quantiles of the FSR approaches to 0.5,
+  the expected FSR value in high quality samples. Even though not
+  every region in a ChIP-exo experiment is perfectly balanced, the
+  most enriched regions are expected to be composed by approximately
+  the same quantity of reads in both strands.}
+
 %and as the URCR decreases the library complexity does it as well, on
 %the other hand 
-\SK{Not sure what this next sentence is referring to: Finally we quantify we quantify the
-relationship between library complexity by the use of two indexes that
-represent the change in the number of unique positions per regions and
-the change in of the width of a regions as the depth changes. }
-Figures~\ref{fig:qcdiagram}B and \ref{fig:qcdiagram}C aim to quantify the strand imbalance as part of the QC pipeline. The former  depicts how
-quickly the islands exclusively formed by reads SK{***I changed fragments to reads, but should this stay as fragments***} from a single strand are  filtered out as islands with higher depths are observed.
-In a
-high quality sample,  the proportion of islands with reads from only one strand 
-is expected to decrease rapidly  as we consider higher depth regions. In contrast, 
- this proportion remains
-approximately constant in  lower quality samples. The latter plot illustrates how quickly the quantiles of 
-the  FSR approaches to 0.5, the expected  FSR value in high quality samples.
+% \SK{Not sure what this next sentence is referring to: Finally we
+%   quantify the relationship between library complexity by the use of
+%   two indexes that represent the change in the number of unique
+%   positions per regions and the change in of the width of a regions as
+%   the depth changes.}
+
 %median, since in a high quality sample it is
@@
   ChIP-exo dataset}
-\SK{We should present the whole QC results and end with validation -  panels from the two figures should be switched.}
-We next use Fox A1 ChIP-exo datasets, with three biological replicates at comparable sequencing depths from mouse liver cells, 
-to illustrate the QC pipeline. 
-Figure~\ref{fig:enrich}A presents  URCR vs. ARC plots
-for all three replicates. The first and third replicates exhibit a defined
-decreasing trend in  $\mbox{URCR}$ as the $\mbox{ARC}$ increases. This indicates that
-these samples exhibit a higher ChIP enrichment than the
-second replicate. On the other hand, the overall $\mbox{URCR}$ level
-from the first two replicates is higher than that of the third replicate
-level,  elucidating  that the  libraries for the first two replicates are more complex than that of  the
-third replicate. 
+% \SK{We should present the whole QC results and end with validation -  panels from the two figures should be switched.}
+
+We next use Fox A1 ChIP-exo datasets, with three biological replicates
+at comparable sequencing depths from mouse liver cells, to illustrate
+the QC pipeline.  Figure~\ref{fig:3}A presents URC vs. ARC plots
+for all three replicates. The first and third replicates exhibit a
+defined decreasing trend in URC as the ARC increases. This indicates
+that these samples exhibit a higher ChIP enrichment than the second
+replicate. On the other hand, the overall URC level from the first two
+replicates is higher than that of the third replicate level,
+elucidating that the libraries for the first two replicates are more
+complex than that of the third replicate.
+
 %As we discussed above,  background fragments are often digested by
@@
 
-Figure~\ref{fig:strand}A and B display the strand-imbalance diagnostic plots and highlight specific problems with replicates 2 and 3.
-In particular, \SK{I have a hard time interpreting B and C specifically, need to include something about their implications for replicates 2 and 3.}
+\SK{I have a hard time interpreting B and C specifically, need to
+  include something about their implications for replicates 2 and 3.}
+
+Figure~\ref{fig:3}B and \ref{fig:3}C display the Read Composition and
+FSR distribution plots, which highlight specific problems with
+replicates 2 and 3. \RW{In Figure~\ref{fig:3}B, we can observe
+  decreasing trends in the proportions of regions formed by fragments
+  in one exclusive strand. In a high quality experiment its is
+  expected to observe an exponential decay in the proportion of single
+  stranded regions, while in lower quality experiments the trend may
+  be linear of even constant (Supplement
+  Figure~\ref{sfig:qcsamp3}). In the FSR distribution plot, both
+  replicate 2 and 3 FSR's distributions are more spread around their
+  respective medians. The approximation of the 0.1 and 0..9 quantiles
+  to the median indicated the aforementioned lower enrichment in the
+  second replicate and the low complexity in the third one.
+ }
+ % In particular, \RW{both replicate 2 and replicate 3 FSR's
+ %  distributions are more spread around the median. The approximation
+ %  of the 0.1 and 0.9 quantiles to the median is indicative of the
+ %  aforementioned lower enrichment of the second replicate, since there
+ %  are less regions with a higher concentration of reads, this makes
+ %  the FSR's distribution to be dominated by non-enriched regions. On
+ %  the other hand, the slow approximation of the quantiles to the FSR
+ %  median in the third replicate is indicative of low library
+ %  complexity. In Figure~\ref{fig:3}C, we can observe an decreasing
+ %  trend in the proportion of regions formed by fragments in one
+ %  exclusive strand. In a high quality experiment it is expected to
+ %  observe an exponential trend, while in lower quality experiments the
+ %  trend may be linear or even constant (Supplement
+ %  Figure~\ref{sfig:qcsamp3}).}
+
 %In order to study the strand imbalance of the FoxA1 factor ChIP-exo
@@
 %strand to be imbalanced in a background region.
-\SK{I also don't follow the discussion for panel C: Figure
-\ref{fig:strand}C compares the strand imbalance when ChIP-exo islands 
-overlap with high quality peaks. It is noticeable that for regions
-composed by a large amount of reads, it is harder to distinguish their
-peaks by considering only the strand imbalance, hence in a high
-quality ChIP-exo experiment the background is expected to show a
-higher imbalance that the enriched regions. In conclusion, Figure
-\ref{fig:strand} shows that the global FSR does not accurately
-represent a ChIP-exo experiment's strand imbalance locally, hence the
-``peak pair'' assumption does not completely hold in every ChIP-exo
-enriched region.}
-
-Overall, we conclude that replicate 1 is higher quality than both  of replicates 2 and 3.
-We validate this observation with a motif analysis on the candidate binding regions identified from these replicates \SK{Could we say binding events?
-or is the peak analysis on the whole binding region, i.e., peak?
-If so, we may want o change "candidate regions" to peaks?}
-Figure~\ref{fig:enrich}B summarizes the total numbers of regions identified from each replicate.
-The lower number of enriched regions from replicate 2 is consistent with the lower ChIP enrichment pattern in the $UCRC$ vs.
-$ARC$ diagnostic plot.
-Scanning of these regions for the occurrence of FoxA1 sequence motif with the FIMO tool \cite{fimo} indicates that the first replicate outperforms the other two in terms of percentage of candidate regions with the FoxA1 motif (Figure~\ref{fig:enrich}C).
+
+ Overall, we conclude that replicate 1 is higher quality than both of
+ replicates 2 and 3.  We validate this observation with a motif
+ analysis on the candidate binding regions identified from these
+ replicates \SK{Could we say binding events? or is the peak analysis
+   on the whole binding region, i.e., peak? If so, we may want o
+   change "candidate regions" to peaks?} \RW{I used a set of high
+   quality regions, i.e. using the ChIP-exo pipeline I filtered them
+   by npos, FSR and depth. Figure~\ref{sfig:venn} show the overlap
+   between these regions and the peaks (roughly speaking these regions
+   are almost a proper subset of the peaks)}. Figure~\ref{fig:4}A
+ summarizes the total numbers of regions identified from each
+ replicate. The lower number of enriched regions from replicate 2 is
+ consistent with the lower ChIP enrichment pattern in the $UCR$
+ vs. $ARC$ diagnostic plot.  Scanning of these regions for the
+ occurrence of FoxA1 sequence motif with the FIMO tool \cite{fimo}
+ indicates that the first replicate outperforms the other two in terms
+ of percentage of candidate regions with the FoxA1 motif
+ (Figure~\ref{fig:4}B).
+
 %Considering the regions and summary statistics we obtained a set of
@@
 %in the second. 
-Furthermore, Figure~\ref{fig:enrich}D displays the average normalized
-read coverage around the actual motif locations in the candidate binding regions.
-These coverage plots reveal that the ChIP signal is more defined for
-the first and third replicates than the second one, indicating overall strength of the ChIP enrichment in these samples compared to the second replicate.
+Furthermore, Figure~\ref{fig:4}C displays the average normalized
+read coverage around the actual motif locations in the candidate
+binding regions.  These coverage plots reveal that the ChIP signal is
+more defined for the first and third replicates than the second one,
+indicating overall strength of the ChIP enrichment in these samples
+compared to the second replicate.
 % however,  overall, the signal pattern is reproducible in all of the three biological replicates.
-Figures \ref{fig:enrich}E, \ref{fig:enrich}F, and
-\ref{fig:enrich}G further highlight the overall quality of the identified motif sequences for each replicate and suggest  that libraries with high library complexity (replicates 1 and 3) capture binding sites with better motif matches.
-\SK{Worth thinking a bit more about the interpretation.}
+Figure~\ref{fig:4}D further highlights the overall quality of the
+identified motif sequences for each replicate and suggest that
+libraries with high library complexity (replicates 1 and \RW{2})
+capture binding sites with better motif
+matches. \RW{Figure~\ref{fig:4}E compares the top FIMO scores among
+  the three replicates, not-surprisingly confirming that the first
+  explicate exhibits the highest quality. Finally, Figure~\ref{fig:4}F
+  compares the Strand Imbalance by considering the regions overlapping
+  with their respective sets of ChIP-exo peaks. Not surprisingly, for
+  the replicate 2 we can see that the Imbalance distribution differ
+  among both classes, which is caused by the aforementioned lower
+  enrichment of that sample; for the remaining replicates we can see
+  that the Imbalance is more similar when regions composed by more
+  reads are considered. This suggests that enriched regions are more
+  likely to be balanced, therefore this suggest that the ``peak-pair''
+  assumption may not hold in non-enriched regions.}
+
+ % Hence, in a low enriched sample the
+ %  ``peak-pair'' assumption does not hold globally.}
+% \SK{I also don't follow the discussion for panel C: Figure
+% \ref{fig:4}C compares the strand imbalance when ChIP-exo islands 
+% overlap with high quality peaks. It is noticeable that for regions
+% composed by a large amount of reads, it is harder to distinguish their
+% peaks by considering only the strand imbalance, hence in a high
+% quality ChIP-exo experiment the background is expected to show a
+% higher imbalance that the enriched regions. In conclusion, Figure
+% \ref{fig:4} shows that the global FSR does not accurately
+% represent a ChIP-exo experiment's strand imbalance locally, hence the
+% ``peak pair'' assumption does not completely hold in every ChIP-exo
+% enriched region.}
+% \SK{Worth thinking a bit more about the interpretation.}
 
@@
 
-\subsubsection*{High sequencing depth may confound low-complexity library issues}
+\subsubsection*{High sequencing depth may confound low-complexity
+  library issues}
 We next evaluated every sample listed in Tables~\ref{tab:qc} and
-\ref{tab:qc_sig} with the \pname{} QC pipeline (Supplementary Figures \SK{XXX}).
-A key, albeit not suprising, observation from large scale analysis is that  the URCR vs.
-ARC plots  typically display the three patterns captured in the FoxA1 study.
-We will refer to these as pattern I (FoxA1 replicate 1), II (FoxA1 replicate 2), and III (FoxA1 replicate 3), respectively.
-Pattern III where the two arms along $\mbox{ARC}$ are not distinguishable can arise due to either low-complexity library or high sequencing depth.
-For example, all three replicates of the TBP ChIP-exo from K562, with sequencing depths between $\sim$ 60M to 115M reads, and replicate two of TBP ChIP-nexus in K562, with sequencing depth of $\sim$ 130M reads, exhibit this pattern.
-A simple but effective strategy to distinguish the two plausible scenarios from Pattern III is to apply the QC pipeline to sub-samples randomly generated from the full dataset at varying sequencing depths.
-We applied this strategy by  sub-sampling 20M to 50M reads, a range that represents the
- sequencing depths of the human samples we are using in the paper, from the TBP samples. 
-$\mbox{URCR}$ vs.
-$\mbox{ARC}$ diagnostics of these sub-samples  (Supplementary Figures \SK{XXX}) indicate that of the four TBP samples with this pattern, replicates two and three of K562 ChIP-exo suffer from low-complexity library issues, whereas the other samples exhibit the pattern specific to high quality samples.
-\SK{Add: Validation with motif on TBP ChIP-exo replicates.}
-
-\subsubsection*{Evaluation of a large collection of ChIP-exo and ChIP-nexus data with \pname{}}
-
-We next performed an overall analysis of the \pname{} 
-QC pipeline results for the samples in Tables~\ref{tab:qc} and
-\ref{tab:qc_sig}.
-We  quantified the relationship between $\mbox{ARC}$ and $\mbox{URCR}$ by fitting a reparametrized regression model of $\mbox{URCR}$ as a function of  $\mbox{ARC}$.
-Specifically, we considered $D_i = \beta_1 U_i +\beta_2 W_i +\varepsilon_i$, where
-$\varepsilon_i$ represents the random error term.
-As we discuss in Materials and methods, this parametrization has a direct connection to $\mbox{URCR}_i = \frac{\kappa}{\mbox{ARC}_i} + \gamma+ \epsilon_i$,   which aims to recapitulate the relationship in the  $\mbox{URCR}$ vs.
-$\mbox{ARC}$ plots.
+\ref{tab:qc_sig} with the \pname{} QC pipeline (Supplementary
+Figures~\ref{sfig:qc1} to \ref{sfig:qc11}). A key, albeit not
+surprising, observation from large scale analysis is that the URC
+vs. ARC plots typically display the three patterns captured in the
+FoxA1 study. We will refer to these as pattern I (FoxA1 replicate 1),
+II (FoxA1 replicate 2), and III (FoxA1 replicate 3), respectively.
+Pattern III where the two arms along ARC are not distinguishable can
+arise due to either low-complexity library or high sequencing depth.
+For example, all three replicates of the TBP ChIP-exo from K562, with
+sequencing depths between $\sim$ 60M to 115M reads, and replicate two
+of TBP ChIP-nexus in K562, with sequencing depth of $\sim$ 130M reads,
+exhibit this pattern. A simple but effective strategy to distinguish
+the two plausible scenarios from Pattern III is to apply the QC
+pipeline to sub-samples randomly generated from the full dataset at
+varying sequencing depths. We applied this strategy by sub-sampling
+20M to 50M reads, a range that represents the sequencing depths of the
+human samples we are using in the paper, from the TBP samples.  URC
+vs. ARC diagnostics of these sub-samples (Supplementary Figures S23 to
+\ref{sfig:qcsamp3}) indicate that of the four TBP samples with this
+pattern, replicates two and three of K562 ChIP-exo suffer from
+low-complexity library issues, whereas the other samples exhibit the
+pattern specific to high quality samples. \RW{To confirm this, we
+  compared the top FIMO scores \cite{fimo} of the TBP motif calculated
+  with the ChIP-exo and ChIP-nexus replicates. Figure~\ref{fig:5}E
+  illustrates that the scores of the ChIP-nexus replicates and the
+  first ChIP-exo replicate are higher than the other two samples when
+  considering the top scores. As expected the ChIP-nexus scores are
+  comparable among and show better performance than the three ChIP-exo
+  experiment and among those three replicates, replicate 1 outperforms
+  the other two.}
+
+%  \RW{Figure~\ref{fig:4}}
+% exhibits the \texttt{FIMO}
+
+\subsubsection*{Evaluation of a large collection of ChIP-exo and
+  ChIP-nexus data with \pname{}}
+
+We next performed an overall analysis of the \pname{} QC pipeline
+results for the samples in Tables~\ref{tab:qc} and \ref{tab:qc_sig}.
+We quantified the relationship between ARC and URC by fitting a
+reparametrized regression model of URC as a function of ARC.
+Specifically, we considered $D_i = \beta_1 U_i +\beta_2 W_i
++\varepsilon_i$, where $\varepsilon_i$ represents the random error
+term.  As we discuss in Materials and methods, this parametrization
+has a direct connection to $\mbox{URC}_i = \frac{\kappa}{\mbox{ARC}_i}
++ \gamma+ \epsilon_i$, which aims to recapitulate the relationship in
+the URC vs. ARC plots.
 %between depth, width, and the number of unique
@@
 %vs. $\mbox{URCR}$ visualization. 
-Figure~\ref{fig:eval}A displays estimated
-overall change in depth ($\hat{\beta}_1$) as the number of positions with at least one aligned read varies across a large collection of ChIP-exo samples from eukaryotic genomes.
-The $\beta_1$ parameter can be interpreted as the  limiting (i.e., large depth)
-\SK{need a better name for:} average read depth intensity $\mbox{URCR}$ of a sample.
+Figure~\ref{fig:5}A displays estimated overall change in depth
+($\hat{\beta}_1$) as the number of positions with at least one aligned
+read varies across a large collection of ChIP-exo samples from
+eukaryotic genomes. The $\beta_1$ parameter can be interpreted as the
+limiting (i.e., large depth) for the URC of a sample.
 %This parameter can be
@@
 %can is also the experiment's large depth $\mbox{URCR}$. 
-As discussed earlier, high
-quality ChIP-exo samples are expected to have two arms in the $\mbox{URCR}$
-vs. $\mbox{ARC}$ plots: one with low $\mbox{ARC}$
-and varying $\mbox{URCR}$ and another with a decreasing
-$\mbox{URCR}$as  $\mbox{ARC}$ increases and stabilizes
+As discussed earlier, high quality ChIP-exo samples are expected to
+have two arms in the URC vs. ARC plots: one with low ARC and varying
+URC and another with a decreasing URC as ARC increases and stabilizes
 $\beta_1$. When the ChIP-exo sample is not deeply sequenced, high
-values of $\hat{\beta}_1$ in  Figure~\ref{fig:eval}A indicate that the library complexity
-is low. On the other hand, lower values correspond
-to higher quality ChIP-exo experiments.
-Taking into account the depths of these samples and visualizing all the diagnostic plots (\SK{Supplementary Figures XXX}), we conclude that samples with estimated $\beta_1$ values less than $10$ seem to be high quality samples.
-
-We interpret the $\beta_2$ parameter above as the bias for the average read coverage  and display its estimates across all the eukaryotic samples in Figure~\ref{fig:eval}B.
+values of $\hat{\beta}_1$ in Figure~\ref{fig:5}A indicate that the
+library complexity is low. On the other hand, lower values correspond
+to higher quality ChIP-exo experiments. Taking into account the depths
+of these samples and visualizing all the diagnostic plots
+(Supplementary Figures~\ref{sfig:qc1} to \ref{sfig:qc11}), we conclude
+that samples with estimated $\beta_1$ values less than $10$ seem to be
+high quality samples.
+
+\RW{We interpret the $\beta_2$ as the overall change in depth as the
+  width varies and display its estimates across all the eukaryotic
+  samples in Figure~\ref{fig:5}B.}
+% parameter above as the bias for the average
+% read coverage and display its estimates across all the eukaryotic
+% samples in Figure~\ref{fig:5}B.
 % shows the average read coverage bias estimates when the
 %experiment's sequencing depth is large/
- Under perfect digestion by $\lambda$-exo, most of the reads aligned to binding regions are expected to accumulate
- around a binding event.  
- \SK{I don't follow the next argument: We should say something about the expected behavior of $\beta_2$
- which suggest to be unlikely to
-observe as the sequencing depth increases to observe a reads being
-aligned to another position that was not previously covered. Low
-quality ChIP-exo experiment exhibit more scattered reads in both
-strands across the genome; therefore it is more likely to observe
-consider reads that align to position in the genome that were not
-previously covered. }
-Although the 
-third replicate of the TBP ChIP-exo experiment has comparable sequencing depth to the 
-second replicate of the TBP ChIP-nexus experiment, the  
-(Figure~\ref{fig:eval}(B)), $\mbox{ARC}$ bias is considerably higher for the ChIP-exo experiment.
-This potentially indicates that additional sequencing reads in comparison to replicates 1 and 2 are scattered around new positions instead of accumulating on the existing binding sites.
+Under perfect digestion by $\lambda$-exo, most of the reads aligned to
+binding regions are expected to accumulate around a binding
+event. \RW{In a high quality experiment the overall variation in depth
+  is going to be small as the overall width of the regions changes, as
+  this change is not going to involve the majority of the reads as
+  those are located tightly around the binding sites. On the other
+  hand, in low quality experiments regions are usually composed of a
+  fixed proportion of reads aligned to a small number of unique
+  positions; hence the overall change in depth as the width varies is
+  proportional to this fixed proportion.} \SK{I don't follow the next
+  argument: We should say something about the expected behavior of
+  $\beta_2$} \RW{Informally, this is what I tried to say: If we have
+  an enriched region and the width is being reduced, then the depth of
+  the region is not going to change a lot since the majority of the
+  reads are localized around the binding sites and not the
+  extremes. On the other hand, in a non-enriched region, by reducing
+  its width then several reads may have been lost (in this case I am
+  thinking about a regions of width $w_i$, with $k$ reads aligned to
+  any of the $u_i$ positions, in this case, since the region is not
+  enriched we observe $u_i << w_i$ and $d_i = k u_i$. Finally,
+  $\frac{\Delta d_i}{\Delta w_i} \approx k /h $, where $h = \Delta w_i
+  \leq rl$.)}
+
+ % which suggest to be unlikely to
+ %  observe as the sequencing depth increases to observe a reads being
+ %  aligned to another position that was not previously covered. Low
+ %  quality ChIP-exo experiment exhibit more scattered reads in both
+ %  strands across the genome; therefore it is more likely to observe
+ %  consider reads that align to position in the genome that were not
+ %  previously covered. }
+Although the third replicate of the TBP ChIP-exo experiment has
+comparable sequencing depth to the second replicate of the TBP
+ChIP-nexus experiment, the (Figure~\ref{fig:5}B), $\hat{\beta_2}$ is
+considerably higher for the ChIP-exo experiment.  This potentially
+indicates that additional sequencing reads in comparison to replicates
+1 and 2 are scattered around new positions instead of accumulating on
+the existing binding sites.
 %We can compare the second replicate from the human
@@
 
-The interaction between these two parameters  has implications regarding the quality of 
-a ChIP-exo and ChIP-nexus sample. When either the adjusted
-$\mbox{ARC}$ or the $\mbox{ARC}$ bias is large \SK{Not clear what large for $\beta_2$ means if we are only looking at a single sample} owing to potentially the high sequencing depth of the sample, we suggest
-randomly sub-sampling reads to form samples of lower depth and evaluating the 
-sub-samples with the QC pipeline. 
-As an illustration, we revisit this strategy for  the three replicates of TBP ChIP-exo in K562 \cite{venters13} and 
-second replicate from the K562  ChIP-nexus experiments \cite{chipnexus}.
-Figure~\ref{fig:eval}C exhibits an increasing trend in the
-estimated $\beta_1$ across varying sequencing depths for replicates 2 and 3 which we deem as lower quality than replicate 1.
-Furthermore, the estimates with lower depths are still higher than that of the replicate 1 and the overall trend of the ChIP-nexus sub-samples.
+The interaction between these two parameters has implications
+regarding the quality of a ChIP-exo and ChIP-nexus sample. When either
+the $\hat{\beta_1}$ is large or $\hat{\beta_2}$ is different
+from zero 
+% \SK{Not clear what large for $\beta_2$ means if we are only looking
+%   at a single sample}
+owing to potentially the high sequencing depth of the sample, we
+suggest randomly sub-sampling reads to form samples of lower depth and
+evaluating the sub-samples with the QC pipeline.  As an illustration,
+we revisit this strategy for the three replicates of TBP ChIP-exo in
+K562 \cite{venters13} and second replicate from the K562 ChIP-nexus
+experiments \cite{chipnexus}. Figure~\ref{fig:5}C exhibits an
+increasing trend in the estimated $\beta_1$ across varying sequencing
+depths for replicates 2 and 3 which we deem as lower quality than
+replicate 1. Furthermore, the estimates with lower depths are still
+higher than that of the replicate 1 and the overall trend of the
+ChIP-nexus sub-samples.
 %, where the ChIP-nexus
 %trend being considerable lower than both ChIP-exo replicates. 
-Figure \ref{fig:eval}D illustrates that the $ARC$ bias remains approximately
-constant in ChIP-nexus sub-samples and sub-samples of first replicate of ChIP-exo, while it increases for the second and third
-ChIP-exo replicates. This suggests that these two ChIP-exo replicates have
-low library complexity and overall lower quality than the ChIP-nexus
-samples, regardless of the fact that all three experiments are
-deeply sequenced with more than 90M  reads each.
-Furthermore, the \pname{}  diagnostic plots for each sub-sample (\SK{Supplementary Figures XXX}) illustrate that the two arms of the ARC versus UCRC plots are clearly visible in moderate depth sub-samples of TBP ChIP-nexus data.
-\SK{Some more work/reorganization not to have too much repetition with this last paragraph and the section that follows right after FoxA1.}
-
-\subsection*{High-resolution binding event identification with statistical analysis  of
-  ChIP-exo data}
-
-Our QC pipeline \pname{} operates on aligned read files and does not require any statistical analysis of the data such as identification of potential binding regions/events.
-This enables its broad and easy usability before any statistical analysis  for identifying binding events from ChIP-exo/nexus data.
-We next evaluated recent analytical approaches for ChIP-exo data analysis and further compared ChIP-exo with PE ChIP-seq data on our \textit{E.
-coli} samples.
+Figure~\ref{fig:5}D illustrates that the $\beta_2$ estimates remains
+approximately constant in ChIP-nexus sub-samples and sub-samples of
+first replicate of ChIP-exo, while it increases for the second and
+third ChIP-exo replicates. This suggests that these two ChIP-exo
+replicates have low library complexity and overall lower quality than
+the ChIP-nexus samples, regardless of the fact that all three
+experiments are deeply sequenced with more than 90M reads
+each. Furthermore, the \pname{} diagnostic plots for each sub-sample
+(Supplementary Figures S23 to \ref{sfig:qcsamp3}) illustrate that the
+two arms of the ARC versus URC plots are clearly visible in moderate
+depth sub-samples of TBP ChIP-nexus data.  \SK{Some more
+  work/reorganization not to have too much repetition with this last
+  paragraph and the section that follows right after FoxA1.}
+
+\subsection*{High-resolution binding event identification with
+  statistical analysis of ChIP-exo data}
+
+Our QC pipeline \pname{} operates on aligned read files and does not
+require any statistical analysis of the data such as identification of
+potential binding regions/events. This enables its broad and easy
+usability before any statistical analysis for identifying binding
+events from ChIP-exo/nexus data.  We next evaluated recent analytical
+approaches for ChIP-exo data analysis and further compared ChIP-exo
+with PE ChIP-seq data on our \textit{E. coli} samples.
 
@@
 
-We specifically considered recently developed Peakzilla \cite{peakzilla}, MACE
-\cite{mace}, GEM \cite{gem}, and dPeak \cite{dpeak} in our evaluations with 
-high quality \emph{E.Coli} samples (samples \SK{x and y} from the aerobic condition).
-Among these methods, Peakzilla and MACE are specifically developed for ChIP-exo analysis, whereas both dPeak and GEM can identify high resolution binding events from ChIP experiments, rendering them suitable for ChIP-exo analysis as well.
+We specifically considered recently developed Peakzilla
+\cite{peakzilla}, MACE \cite{mace}, GEM \cite{gem}, and dPeak
+\cite{dpeak} in our evaluations with high quality \textit{E. coli}
+samples (samples E1-1 and E1-2). Among these methods, Peakzilla and
+MACE are specifically developed for ChIP-exo analysis, whereas both
+dPeak and GEM can identify high resolution binding events from ChIP
+experiments, rendering them suitable for ChIP-exo analysis as well.
 %After we determined which biases could affect a ChIP-exo experiment's
@@
 
-Figures~\ref{fig:methods_comp}A and \ref{fig:methods_comp}B compare
-binding events identified with all the four methods  in terms of resolution,
-where the resolution is defined as the distance from a RegulonDB \cite{regulondb}
-annotation to its closest prediction. The resolutions of  all the methods
-are comparable for the first replicate (Figure~\ref{fig:methods_comp}(A)) and dPeak, on average, has slightly better resolution than the rest in the second replicate (Figure~\ref{fig:methods_comp}(B)).
-
-dPeak model  has two parameters that further elucidate characteristics of ChIP-exo data.
-The $\delta$ parameter of dPeak quantifies the average distance from the
-$5^{\prime}$ ends of the reads to the binding event they are profiling and the $\sigma$
-parameter is a measure of dispersion of $5^{\prime}$ ends around the binding events.
-Figures~\ref{fig:methods_comp}C and
-\ref{fig:methods_comp}D compare the densities of these parameters
-estimated by dPeak for \textit{E. coli}  ChIP-exo and SE ChIP-seq data. As expected, 
-both   are lower for ChIP-exo than for SE
-ChIP-Seq, indicating that dPeak, although originally developed for high-resolution binding event identification from ChIP-seq, can learn this information.
+Figures~\ref{fig:6}A and \ref{fig:6}B compare binding events
+identified with all the four methods in terms of resolution, where the
+resolution is defined as the distance from a RegulonDB
+\cite{regulondb} annotation to its closest prediction. The resolutions
+of all the methods are comparable for the first replicate
+(Figure~\ref{fig:6}A) and dPeak, on average, has slightly better
+resolution than the rest in the second replicate
+(Figure~\ref{fig:6}B).
+
+dPeak model has two parameters that further elucidate characteristics
+of ChIP-exo data.  The $\delta$ parameter of dPeak quantifies the
+average distance from the $5^{\prime}$ ends of the reads to the
+binding event they are profiling and the $\sigma$ parameter is a
+measure of dispersion of $5^{\prime}$ ends around the binding events.
+Figures~\ref{fig:6}C and \ref{fig:6}D compare the densities of these
+parameters estimated by dPeak for \textit{E. coli} ChIP-exo and SE
+ChIP-seq data (sample E1-1 and S1-1, respectively). As expected, both
+are lower for ChIP-exo than for SE ChIP-Seq, indicating that dPeak,
+although originally developed for high-resolution binding event
+identification from ChIP-seq, can learn this information.
 %Therefore, we can asses that dPeak is accurately
@@
   varying sequencing depths}
-Establishing competitive performance of dPeak compared to other ChIP-exo analysis methods enable comparison of ChIP-exo and ChIP-seq with a unified analysis framework using dPeak.
-Previous comparisons of ChIP-exo and SE ChIP-seq
-were all performed without controlling for the sequencing depths of the samples.
-More importantly, although it is well established that PE ChIP-seq leads to better resolution that SE ChIP-seq in terms of binding site identification \SK{Cite Qi's BMC Bioinformatics paper}, previous work does not have an in depth comparison of ChIP-exo and PE ChIP-seq.
+Establishing competitive performance of dPeak compared to other
+ChIP-exo analysis methods enable comparison of ChIP-exo and ChIP-seq
+with a unified analysis framework using dPeak. Previous comparisons of
+ChIP-exo and SE ChIP-seq were all performed without controlling for
+the sequencing depths of the samples. More importantly, although it is
+well established that PE ChIP-seq leads to better resolution that SE
+ChIP-seq in terms of binding site identification \cite{zhang16},
+previous work does not have an in depth comparison of ChIP-exo and PE
+ChIP-seq.
 %Previously, ChIP-exo and SE ChIP-Seq have been compared only at a
@@
 %ChIP-Seq as well either. 
-In order to address these limitations of the
-previous studies, we performed a sub-sampling experiment by sampling fixed numbers of reads from each of the
-$\sig$ ChIP-exo, PE ChIP-seq, and SE ChIP-seq datasets. Specifically, for every $N$ reads in ChIP-exo and SE ChIP-seq, $N/2$ read pairs 
-were sampled for PE ChIP-seq to operate under fixed sequencing costs.
+In order to address these limitations of the previous studies, we
+performed a sub-sampling experiment by sampling fixed numbers of reads
+from each of the $\sig$ ChIP-exo, PE ChIP-seq, and SE ChIP-seq
+datasets. Specifically, for every $N$ reads in ChIP-exo and SE
+ChIP-seq, $N/2$ read pairs were sampled for PE ChIP-seq to operate
+under fixed sequencing costs.
 %Additionally, it is worth noticing that for PE ChIP-Seq we
@@
 
-Figure~\ref{fig:design} summarizes the comparisons of the three data types for $\sig$ under aerobic condition as a function of sequencing depth.
-For these computational experiments, we used $\sig$ RegulonDB \cite{regulondb} binding events as gold standard.
-Figure~\ref{fig:design}A displays the number of candidate regions (i.e., peaks)
- where at least one  binding event was identified whereas Figure~\ref{fig:design}B depicts the number of identified binding events, i.e., each candidate region can harbor multiple binding events.
-In terms of the number of  binding events, ChIP-exo ranks on the conservative side.  
-In Figure~\ref{fig:design}C,  we
-display the number of correctly identified binding events, where we consider a  RegulonDB event as
-correctly identified if an estimated binding event is identified within 15 bps
-of it. Finally, Figure~\ref{fig:design}D presents the
-resolution defined as the distance from a RegulonDB binding event to the
-closest dPeak prediction.
-These comparisons indicate that although PE ChIP-seq performs much better than SE ChIP-seq in terms of identifying closely spaced binding events, ChIP-exo outperforms PE ChIP-seq at comparable depths.
+Figure~\ref{fig:7} summarizes the comparisons of the three data
+types for $\sig$ under aerobic condition (samples E1-1, P1-1 and S1-1)
+as a function of sequencing depth. For these computational
+experiments, we used $\sig$ RegulonDB \cite{regulondb} binding events
+as gold standard.  Figure~\ref{fig:7}A displays the number of
+candidate regions (i.e., peaks) where at least one binding event was
+identified whereas Figure~\ref{fig:7}B depicts the number of
+identified binding events, i.e., each candidate region can harbor
+multiple binding events.  In terms of the number of binding events,
+ChIP-exo ranks on the conservative side.  In Figure~\ref{fig:7}C,
+we display the number of correctly identified binding events, where we
+consider a RegulonDB event as correctly identified if an estimated
+binding event is identified within 15 bps of it. Finally,
+Figure~\ref{fig:7}D presents the resolution defined as the
+distance from a RegulonDB binding event to the closest dPeak
+prediction. These comparisons indicate that although PE ChIP-seq
+performs much better than SE ChIP-seq in terms of identifying closely
+spaced binding events, ChIP-exo outperforms PE ChIP-seq at comparable
+depths.
 %It is remarkable that even when the number
@@
 
-Figure~\ref{fig:reso_all} shows comparisons among dPeak analysis of full ChIP-exo, PE
-ChIP-seq and SE ChIP-seq $\sig$ datasets \SK{Again, make sure these ChIP-seq samples are introduced before and numbering system will help to refer each sample}.
-\SK{Are the sequencing depths comparable?
-We are making a big deal about matching depths and then jumping into this analysis.
-The only way these results would work is that either the depts are comparable or all the samples have depths sufficient enough for saturation}.
-Utilizing  RegulonDB binding events as ground truth,
+Figure~\ref{fig:8} shows comparisons among dPeak analysis of
+full ChIP-exo, PE ChIP-seq and SE ChIP-seq $\sig$ datasets \SK{Again,
+  make sure these ChIP-seq samples are introduced before and numbering
+  system will help to refer each sample}. \SK{Are the sequencing
+  depths comparable? We are making a big deal about matching depths
+  and then jumping into this analysis. The only way these results
+  would work is that either the depths are comparable or all the
+  samples have depths sufficient enough for saturation} \RW{The
+  samples have sufficient depth for saturation: ChIP-exo is ~14M, PE
+  ChIP-seq is ~27M and SE ChIP-seq is ~7.5M. Since the main point is
+  to compare ChIP-exo and PE-ChIP-seq I think this plots are good as
+  PE ChIP-seq has twice as many reads as ChIP-exo. About the
+  comparison with SE ChIP-seq I don't it works very well, as the \# of
+  ChIP-exo fragments is almost double the amount of SE ChIP-seq. So, I
+  think we have the following options: A) Keep it with the disclaimer
+  that the comparison against SE ChIP-seq is not as important as the
+  other, B) Remove the SE ChIP-seq part but keep ChIP-exo vs. PE
+  ChIP-seq or C) Send the whole figure / section to the
+  supplement.}. Utilizing RegulonDB binding events as ground truth,
 %We considered the RegulonDB data as ground
@@
 distance between a RegulonDB event and the closest dPeak binding
-event prediction. Figure \ref{fig:reso_all}A illustrates that as the mean distance between binding events increases, sensitivity
+event prediction. Figure \ref{fig:8}A illustrates that as the mean distance between binding events increases, sensitivity
 for all data types increase.
@@
 %ChIP-exo and PE ChIP-Seq. 
-Figure~\ref{fig:reso_all}B highlights that 
+Figure~\ref{fig:8}B highlights that 
 ChIP-exo and PE ChIP-seq are comparable in resolution with these deeply sequenced samples, while both
@@
 
-To the best of our knowledge, we also provide the first systematic comparison 
-between ChIP-exo and PE ChIP-seq datasets using our \textit{E. coli} $\sig$ samples. 
-This comparison revealed that PE ChIP-seq compares much more competitively with ChIP-exo compared to SE ChIP-seq. However,
-overall, ChIP-exo provides the best performance in terms of deconvolving closely spaced binding events and resolution.
-The pname{} is available at \url{https://github.com/keleslab/ChIPexoQual}.
+To the best of our knowledge, we also provide the first systematic
+comparison between ChIP-exo and PE ChIP-seq datasets using our
+\textit{E. coli} $\sig$ samples.  This comparison revealed that PE
+ChIP-seq compares much more competitively with ChIP-exo compared to SE
+ChIP-seq. However, overall, ChIP-exo provides the best performance in
+terms of deconvolving closely spaced binding events and resolution.
+The \pname{} package is available at
+\url{https://github.com/keleslab/ChIPexoQual}.
 %Using a set of annotations as
@@
 
-\subsubsection*{\textit{E. coli} ChIP-exo and ChIp-seq samples}
-
-We generated ChIP-exo and PE ChIP-seq samples from $\sig$ in
-\emph{E.Coli}. For each PE ChIP-seq experiment, an \emph{in silico} SE
-version was obtained by randomly sampling one of the two ends in a read pair.
-
-{\color{red}\emph{need to add the growth conditions. \SK{Add something based on dpeak paper, we can ask Jeff to check.}}}
-
-\subsubsection*{Processing of the ChIP-exo and ChIP-nexus samples}
+\subsubsection*{\textit{E. coli} ChIP-exo and ChIP-seq samples}
+
+% We generated ChIP-exo, SE ChIP-seq and PE ChIP-seq samples from $\sig$
+% in \textit{E. coli}.
+\RW{
+\textit{Growth conditions.} All strains were grown in MOPS minimal
+medium supplemented with $0.2\%$ glucose \cite{neidhardt74} at
+$37\degree C$. E1 samples were sparged with a gas mix of $1\%$ $CO_2$,
+$30\%$ $O_2$ and $69\%$ $N_2$; P1 and S1 aerobic samples with a gas
+mix of $95\%$ $N_2$ and $5\%$ $CO_2$ (anaerobic) or $70\%$ $N_2$,
+$5\%$ $CO_2$o and $25\%$ $O_2$ (aerobic); E2 and P2 samples with a gas
+mix of $5\%$ $CO_2$, $25\%$ $O_2$ and $70\%$ $N_2$, additionally
+samples E2-2, E2-4, P2-2 and P2-4 were treated with
+\textit{Rifampicin} by 20 minutes. P1-2 and S1-2 samples were
+harvested during a stationary growth ($OD_{600} \geq 1.0$ for 1
+hour). The remaining samples were harvested during a mid growth
+($OD_{600} = 0.3$ using a Perkin Elmer Lambda 25 $UV/Vis$
+Spectrophotometer). WT \textit{E. coli} K-12 MG1655 was used for the
+E1, P1 and S1 experiments and WT \textit{E. coli} K-12 RL3000 was used
+for the E2 and P2 experiments.}
+
+\RW{ \textit{Library preparation and sequencing.} For ChIP-Seq
+  experiments, $10$ ng of immunoprecipitated and purified DNA
+  fragments from the aerobic and anaerobic $\sig$ samples (one
+  biological sample for both aerobic and anaerobic growth conditions),
+  along with $10$ ng of input control (two biological replicates for
+  anaerobic Input and one biological sample for aerobic Input), were
+  submitted to the University of Wisconsin-Madison DNA Sequencing
+  Facility for ChIP-Seq library preparation. Samples were sheared to
+  $200$ - $500$ nt during the IP process to facilitate library
+  preparation. All libraries were generated using reagents from the
+  Illumina Paired End Sample Preparation Kit (Illumina) and the
+  Illumina protocol \textit{``Preparing Samples for ChIP Sequencing of
+    DNA''} (Illumina part \# 11257047 RevA) as per the manufacturer's
+  instructions, except products of the ligation reaction were purified
+  by gel electrophoresis using $2\%$ SizeSelect agarose gels
+  (Invitrogen) targeting $275$ bp fragments. \RW{need part about
+    enzyme digestion} After library construction and amplification,
+  quality and quantity were assessed using an Agilent DNA 1000 series
+  chip assay (Agilent) and QuantIT PicoGreen dsDNA Kit (Invitrogen),
+  respectively, and libraries were standardized to $10 \mu$M. \RW{For
+    Chip-exo?}For PE ChIP-seq data, cluster generation was performed
+  using an Illumina cBot Paired End Cluster Generation Kit
+  (v3). Paired reads, $36$ bp run was performed for each end, using
+  $200$ bp v3 SBS reagents and CASAVA (the Illumina pipeline) v 1.8.2,
+  on the HiSeq2000. For SE ChIP-seq data, cluster generation was
+  performed using an Illumina cBot Single Read Cluster Generation Kit
+  (v4) and placed on the Illumina cBot. A single read, $32$ bp run was
+  performed, using standard $36$ bp SBS kits (v4) and SCS 2.6 on an
+  Illumina Genome Analyzer IIx. Base calling was performed using the
+  standard Illumina Pipeline version 1.6.}
+
+\RW{The S2 group samples were built \textit{in silico}, by randomly
+  sampling one of the two ends in a read pair from the respective PE
+  ChIP-seq experiment in the P2 group with the same conditions.}
+
+% For each PE ChIP-seq experiment, an \emph{in silico}
+% SE version was obtained by randomly sampling one of the two ends in a
+% read pair.
+% {\color{red}\emph{need to add the growth conditions. \SK{Add something based on dpeak paper, we can ask Jeff to check.}}}
+
 %We gathered a collection of ChIP-exo experiments spanning different
@@
 %in \emph{D. Melanogaster} embryo cell lines.
-We aligned the read files of the samples listed in Table~\ref{tab:qc_sig}
-either following the directions in their original publications when available or with \texttt{bowtie} (version 1.1.2) \cite{bowtie}. 
-\SK{include bowtie comment to show how parameters were set.}
+
+\subsubsection*{Processing of the ChIP-exo and ChIP-nexus samples}
+
+We aligned the read files of the samples listed in
+Table~\ref{tab:qc_sig} either following the directions in their
+original publications when available or with \texttt{bowtie} (version
+1.1.2) \cite{bowtie}. The E1 samples were aligned by using
+\texttt{bowtie -q -m 1 -l 55 -k 1 -5 3 -3 40 --best -S} and the E2
+samples were aligned by using \texttt{bowtie -q -m 1 -v 2 --best}.
+
+%\SK{include bowtie comment to show how parameters were set.}
 %The read files were aligned following the instructions provided in
@@
 
-\subsection*{Generation of a set of high signal regions from \textit{E. coli} samples to assess strand imbalance}
+\subsection*{ChIP-exo and ChIP-seq peak calling with MOSAiCS}
+
+\RW{ \texttt{MOSAiCS} \cite{mosaics} is a model-based approach for the
+  analysis of ChIP-exo and ChIP-seq data. We used \texttt{MOSAiCS} to
+  identify sets of high quality peaks for ChIP-exo and ChIP-seq under
+  the $\mbox{GC + Mappability}$ and $\mbox{Input Only}$ modes for
+  background estimation, respectively. Subsequently, we called peaks
+  with a 5\% FDR and a threshold of at least 100 extended fragments.}
+
+\subsection*{Generation of a set of high signal regions from
+  \textit{E. coli} samples to assess strand imbalance}
 
@@
 
-We partitioned the \emph{E.Coli} genome into
-non-overlaping intervals, i.e., bins, of length 150 bps and counted the number of reads overlapping each bin.
-As is usually the practice with ChIP-seq analysis, each read was extended to the average fragment length of $150$ bps towards the $3^{\prime}$ direction.
-%for each bin we counted how
-%many ChIP-exo and SE ChIP-Seq extended reads overlapped each bin. The
-%reads were extended by 150 bp to their $3\prime$ directions.
-To evaluate the strand imbalance, we identified a set of high quality peaks
-for ChIP-exo and SE ChIP-seq by analyzing both with \texttt{mosaics} \cite{mosaics}
-under the  $\mbox{GC content + Mappability}$ and $\mbox{Input only}$ modes for background estimation, respectively.
-The subset of these peaks for which \texttt{dPeak} analysis  identified one or more binding events were used in FSR assessments (Figure~\SK{Fig num}).
-\SK{Do the plots change if we consider only peaks with one binding event?
-One could argue that when there are multiple events, there might be more background reads, which could skew the FSR distribution.}
+We partitioned the \textit{E. coli} genome into non-overlaping
+intervals, i.e., bins, of length 150 bps and counted the number of
+reads overlapping each bin. As is usually the practice with ChIP-seq
+analysis, each read was extended to the average fragment length of
+$150$ bps towards the $3^{\prime}$ direction. To evaluate the strand
+imbalance, we identified a set of high quality peaks for ChIP-exo and
+SE ChIP-seq. The subset of these peaks for which \texttt{dPeak}
+analysis identified one or more binding events were used in FSR
+assessments (Figure~\ref{fig:1}C and Figure~\ref{sfig:comp1}).
+
+% We partitioned the \textit{E. coli} genome into non-overlaping
+% intervals, i.e., bins, of length 150 bps and counted the number of
+% reads overlapping each bin. As is usually the practice with ChIP-seq
+% analysis, each read was extended to the average fragment length of
+% $150$ bps towards the $3^{\prime}$ direction. To evaluate the strand
+% imbalance, we identified a set of high quality peaks for ChIP-exo and
+% SE ChIP-seq by analyzing both with \texttt{mosaics} \cite{mosaics}
+% under the $\mbox{GC content + Mappability}$ and $\mbox{Input only}$
+% modes for background estimation, respectively.  The subset of these
+% peaks for which \texttt{dPeak} analysis identified one or more binding
+% events were used in FSR assessments (Figure~\ref{fig:comp} and
+% Figure~\ref{sfig:comp1}).
+
+\SK{Do the plots change if we consider only peaks with one binding
+  event? One could argue that when there are multiple events, there
+  might be more background reads, which could skew the FSR
+  distribution.}
+
+\RW{I did a quick check and the plot looks more like fig. S1: The
+  ChIP-seq density seems to be more localized around 0.5 (expected),
+  and the ChIP-exo density seems to become more flat. I think, this
+  happens because when we remove peaks with more than 1 TFBS, we are
+  also removing peaks where at least two binding sites are located
+  close to each other, which makes harder for the $\lambda$ enzyme to
+  reach those regions, so as the distance between the binding events
+  get shorter we are observing regions that behave more like ChIP-seq
+  in than ChIP-exo. in fig10A we can see that the mean distance
+  between binding sites is below 100 for the majority of the peaks.}
+
 %(with $\mbox{GC content + Mappability}$
@@
 %\begin{align*}
-%    \mbox{FSR} &= \frac{\text{Nr. of fwd. strand reads mapped to the region}}                 {\text{Total nr. of reads mapped to the region}} 
+%    \mbox{FSR} &= \frac{\text{Nr. of fwd. strand reads mapped to the region}}{\text{Total nr. of reads mapped to the region}} 
 %\end{align*}
@@
 
-We used the ChIP-seq QC metric definitions  established by \cite{encode_qc} and described in detail at
+We used the ChIP-seq QC metric definitions established by
+\cite{encode_qc} and described in detail at
 %Landt et al., 2012 which are also listed in
-\url{https://genome.ucsc.edu/ENCODE/qualityMetrics.html}.
-These QC metrics were calculated with the \textbf{ChIPUtils}
-package (version 0.99.0 from 
-\url{https://github.com/welch16/ChIPUtils}). 
+\url{https://genome.ucsc.edu/ENCODE/qualityMetrics.html}.  These QC
+metrics were calculated with the \textbf{ChIPUtils} package (version
+0.99.0 from \url{https://github.com/welch16/ChIPUtils}).
 %This package provides an
@@
 %diagnostic plots for ChIP-Seq data.
-Empirical data from the ENCODE project suggests the following guidelines for interpretation of the QC metrics for human and mouse genomes: a PBC value
-between $0$ to $0.5$  indicates severe bottlenecking,  $0.5$ to $0.8$
-moderate bottlenecking,  $0.8$ to $0.9$  mild
-bottlenecking, and  $0.9$ - $1$ no bottlenecking.
-
-\SK{I am in favor of skipping these definitions of ENCODE metrics  for the paper - you might want to keep it for your dissertation-}
-
-%\subsubsection*{Strand Cross-Correlation.}
-
-%The strand cross-correlation was proposed by Kharchenko et al.,
-%2008\nocite{strandcc} and it may be one of the most used of the
-%ChIP-Seq QC metrics. The SCC curve is defined as:
-
-%\begin{align}
-%  y(\delta) = \sum_c w_c r\left[ n_c^+ \left(x + \frac{\delta}{2}
-%    \right), n_c^- \left( x- \frac{\delta}{2} \right)\right],
-%\label{scc}
-%\end{align}
-
-%where $y(\delta)$ is the SCC for a strand shift $\delta$, $r$ is the
-%Pearson correlation, $w_c$ is the proportion of reads mapped to
-%chromosome $c$ and $n_c^S$ is the read count vector for strand $S$ and
-%chromosome $c$. The following QC metrics are used to summarize the shape of the SCC curve:
-
-%\begin{align}
-%  \mbox{NSC} &= \frac{\max_\delta y(\delta)}{\min_\delta y(\delta)}, \label{nsc} \\
-%  \mbox{RSC} &= \frac{\max_\delta y(\delta) -
-%    y_{\text{bgd}}}{y_{\text{rl}} - y_{\text{bgd}}}. \label{rsc}
-%\end{align}
-
-%where $y_{\text{bgd}}$ is estimated as the background SCC level is
-%defined as $y_{\text{bgd}} = \min_\delta y(\delta)$ and
-%$y_{\text{rl}}$ is the value of the SCC curve when the shift equals
-%the experiment's read length. It is worth noticing that in Landt et
-%al., 2012 the RSC is defined as:
-
-%\begin{align*}
-%  \mbox{RSC} &= \frac{\max_\delta y(\delta)}{y_{\text{rl}}}.  
-%\end{align*}
-
-%In ChIP-exo data, both definitions are equivalent since in a typical
-%ChIP-exo experiment $y_{\text{bgd}}$ is approximately zero.
-
-%\subsubsection*{PCR Bottleneck Coefficient.}
-
-%The PCR Bottleneck coefficient is a measure of library complexity in
-%ChIP-Seq data:
-
-%\begin{align}
-%  \mbox{PBC} = \frac{\text{Nr. of positions to which exactly one
-%      unique mapping read is aligned}}{\text{Nr. of positions to
-%      which at least one unique mapping read is aligned}} \nonumber
-%\end{align}
-
-\subsection*{ChIP-exo quality control with \texttt{R} package \texttt{ChIPexoQual}}
-
-We implemented our proposed QC pipeline with an \texttt{R} package named \texttt{ChIPexoQual}, available at 
-\url{https://github.com/welch16/ChIPexoqual}. The analysis in this paper used version 1.0 of the \pname{} package.
+Empirical data from the ENCODE project suggests the following
+guidelines for interpretation of the QC metrics for human and mouse
+genomes: a PBC value between $0$ to $0.5$ indicates severe
+bottlenecking, $0.5$ to $0.8$ moderate bottlenecking, $0.8$ to $0.9$
+mild bottlenecking, and $0.9$ - $1$ no bottlenecking.
+
+\subsection*{ChIP-exo quality control with \texttt{R} package
+  \texttt{ChIPexoQual}}
+
+We implemented our proposed QC pipeline with an \texttt{R} package
+named \texttt{ChIPexoQual}, available at
+\url{https://github.com/welch16/ChIPexoqual}. The analysis in this
+paper used version 1.0 of the \pname{} package.
 
 \begin{framed}
-\noindent \texttt{ChIPexoQual}: 
-The package takes in as input a set of aligned reads from a
-ChIP-exo (or ChIP-nexus) experiment and performs the following steps.
+  \noindent \texttt{ChIPexoQual}: The package takes in as input a set
+  of aligned reads from a ChIP-exo (or ChIP-nexus) experiment and
+  performs the following steps.
 
   \begin{enumerate}
-\item Identify read islands, i.e., overlapping clusters of reads separated by gaps, from  read coverage.
-
-\item Compute $D_i$,  number of reads in the  island $i$ , and $U_i$, number of island $i$ positions with at least one aligning read, $i=1, \cdots, I$.
-
-\item For each island $i$, $i=1, \cdots, I$, compute island statistics:
+  \item Identify read islands, i.e., overlapping clusters of reads
+    separated by gaps, from read coverage.
+
+  \item Compute $D_i$, number of reads in the island $i$ , and $U_i$,
+    number of island $i$ positions with at least one aligning read,
+    $i=1, \cdots, I$.
+
+  \item For each island $i$, $i=1, \cdots, I$, compute island
+    statistics:
   \begin{align*}
-    \mbox{ARC}_i &= \frac{D_i}{W_i}, \quad \mbox{URCR}_i = \frac{U_i}{D_i},  \\
-    %\mbox{URCR}_i &= \frac{U_i}{D_i}, \\
-    \mbox{FSR}_i &= (\text{\# of forward strand reads aligning to island $i$})/D_i, 
+    \mbox{ARC}_i &= \frac{D_i}{W_i}, \quad \mbox{URC}_i = \frac{U_i}{D_i},  \\
+    %\mbox{URC}_i &= \frac{U_i}{D_i}, \\
+    \mbox{FSR}_i &= (\text{\# of forward strand reads aligning to
+      island $i$})/D_i,
   \end{align*}
   where $W_i$ denotes the width of island $i$.
-\item Generate diagnostic plots (i) URCR vs. ARC plot; (ii) \SK{Name this FSR plot}; (iii) \SK{Name this FSR plot} as in Figure~\ref{fig:qcdiagram}.
-\item Randomly sample $M$ (at least \SK{?}) islands and fit, 
+\item Generate diagnostic plots (i) URC vs. ARC plot; (ii) Region
+  Composition plot;
+  % \SK{Name this FSR plot}
+  (iii) FSR distribution plot.
+  % \SK{Name this FSR plot as in Figure~\ref{fig:2}.}
+\item Randomly sample $M$ (at least 1000) islands and fit,
   \begin{align*}
@@
   \end{align*}
-  where $\varepsilon$ denotes the independent error term.  Repeat this process $B$ times and generate  box plots of estimated $\beta_1$ and $\beta_2$. 
+  where $\varepsilon$ denotes the independent error term.  Repeat this
+  process $B$ times and generate box plots of estimated $\beta_1$ and
+  $\beta_2$.
 \end{enumerate}
@@
 \begin{align}
-  \mbox{URCR}_i = \frac{\kappa}{\mbox{ARC}_i} + \gamma + \epsilon_i
+  \mbox{URC}_i = \frac{\kappa}{\mbox{ARC}_i} + \gamma + \epsilon_i
 \label{mod}
@@
 with $\beta_1 = 1 / \gamma$ and $\beta_2 = - \kappa / \gamma$. In this
-setting, $\gamma$ can be considered as the large-depth
-$\mbox{URCR}_i$, i.e., the limiting ratio between the number of positions with at least one mapping read
-and depth as the depth tends to infinity.
-
-\SK{***$\beta_2$ needs more clarification. }
-On the other hand, to
+setting, $\gamma$ can be considered as the large-depth $\mbox{URC}_i$,
+i.e., the limiting ratio between the number of positions with at least
+one mapping read and depth as the depth tends to infinity.
+
+\SK{***$\beta_2$ needs more clarification. } On the other hand, to
 interpret $\beta_2 = - \kappa / \gamma $, by expressing $\kappa$ as a
-function of $\mbox{ARC}$ and $\mbox{URCR}$ and assuming that $\gamma$
+function of $\mbox{ARC}$ and $\mbox{URC}$ and assuming that $\gamma$
 is already estimated, we can observe the following identities:
@@
 \begin{align*}
-  \kappa &= \frac{\mbox{npos}}{\mbox{width}} - \gamma \mbox{ARC} \\
-  \frac{\kappa}{\gamma} &= \frac{1}{\gamma} \frac{\mbox{npos}}{\mbox{width}} - \mbox{ARC} 
+  \kappa &= \frac{U}{W} - \gamma \mbox{ARC} \\
+  \frac{\kappa}{\gamma} &= \frac{1}{\gamma} \frac{U}{W}   - \mbox{ARC} 
 \end{align*}
 
-This is important: $\gamma$ approximates the $\mbox{URCR}$ as the
-sequencing depth increases, which implies that $- \kappa / \gamma$ can
-be interpreted as the large sequencing depth bias of the $\mbox{ARC}$
-since as the depth increases, the first term of $\kappa / \gamma$ is
-going to approximate the average read coverage:
-
-\begin{align*}
-  \frac{\mbox{npos}}{\mbox{width}} \times \lim_{d \rightarrow
-    \infty} \frac{d }{\mbox{npos}} \sim \mbox{ARC} 
-\end{align*}
-
-Therefore, $\beta_2 = - \kappa / \gamma$ is interpreted as the
-$\mbox{ARC}$ bias as the sequencing depth increases.
+This is important: $\gamma$ approximates the URC as the sequencing
+depth increases, which implies that $- \kappa / \gamma$ can be
+interpreted as the large sequencing depth bias of the ARC since as the
+depth increases, the first term of $\kappa / \gamma$ is going to
+approximate the average read coverage:
+
+\begin{align}
+  \frac{U}{W}\times \frac{1}{\hat{\gamma}}= \frac{U}{W} \times \lim_{D
+    \rightarrow \infty} \frac{D}{U(D)} \sim \mbox{ARC}
+\label{eqlim}
+\end{align}
+
+Therefore, $\beta_2 = - \kappa / \gamma$ is interpreted as the ARC
+bias as the sequencing depth increases. \RW{In a high quality
+  experiment $\gamma$ is estimated by a positive value, additionally
+  for a given region in the aforementioned high quality experiment
+  reads align to the majority of the possible mappable positions,
+  therefore we can consider that the first term in (\ref{eqlim})
+  converges to the observed ARC of that region. On the other hand, if
+  we consider a non-enriched region in low quality experiment, then
+  the reads are aligned to few unique positions, showing that $\kappa
+  / \gamma$ theoretically diverges.}
+
+% \RW{relative difference between the observed and
+%   large depth limit $\mbox{URC}$, weighted by the $\mbox{ARC}$. In a
+%   high quality experiment this parameter is expected to be zero, since
+%   $\gamma$ is expected to approximate the $\mbox{URC}$ from deeply
+%   sequenced regions in the genome. On the other hand, on lower quality
+%   experiments this parameter may deviate from zero specially in
+%   regions with PCR artifacts.}
+
  \SK{***}
 
-\subsection*{Motif analysis of FoxA1 enriched regions}
-For each FoxA1 ChIP-exo replicate, we used the ChIP-exo QC pipeline to partition the
-mouse genome into a set of islands  with their respective summary
-statistics. We  then filtered them into collections of high quality regions
-by (i) removing the islands with reads residing only on one strand; (ii) removing the islands with $U_i \le 15$; (iii) removing islands with $D_i < 100$.
-These thresholds were empirically selected and the overall conclusions were robust to their variation.
+% large sequencing depth bias of the ARC since as the
+% depth increases, the first term of $\kappa / \gamma$ is going to
+% approximate the average read coverage:
+
+% \begin{align*}
+%   \frac{\mbox{npos}}{\mbox{width}} \times \lim_{d \rightarrow
+%     \infty} \frac{d }{\mbox{npos}} \sim \mbox{ARC} 
+% \end{align*}
+
+% Therefore, $\beta_2 = - \kappa / \gamma$ is interpreted as the
+% $\mbox{ARC}$ bias as the sequencing depth increases.
+
+\subsection*{Motif analysis of FoxA1 and TBP enriched regions}
+
+\color{blue}
+
+For each ChIP-exo/nexus sample, we used the ChIP-exo QC pipeline to
+partition its respective genome into a set of islands with their
+respective summary statistics. We then filtered them into collections
+of high quality regions by:
+
+\begin{itemize}
+\item FoxA1 experiments: (i) removing the islands with reads residing
+  only on one strand; (ii) removing the islands with $U_i \leq 15$;
+  (iii) removing islands with $D_i < 100$.
+\item TBP experiments: (i) removing the islands with reads residing
+  only on one strand; (ii) removing islands with $W_i < 50$ or $W_i
+  \geq 2000$ bp; (iii) removing islands with the $D_i <=
+  \mbox{median}_j D_j$.
+\end{itemize}
+
+These thresholds were empirically selected and the overall conclusions
+were robust to their variation.
+
+We used \texttt{FIMO} (version 4.9.1) \cite{fimo} to identify the
+FoxA1 and TBP motifs within each enriched regions using the FoxA1
+MA0148.1 and TBP MA0108.1 position weight matrices from the
+\texttt{JASPAR} database \cite{jaspar} respectively. For the FoxA1
+experiments we used the default parameters and for the TBP experiments
+we considered all motifs found with p.value less than 0.01.
+
+\color{black}
+
+% We used \texttt{FIMO} (version 4.9.1, \SK{include command line to show parameters}) \cite{fimo} to identify
+% the FoxA1 motif within each enriched region using FoxA1 position weight matrix MA0148.1 from the JASPAR database \cite{jaspar}.
+
+% For each FoxA1 ChIP-exo replicate, we used the ChIP-exo QC pipeline to
+% partition the mouse genome into a set of islands with their respective
+% summary statistics. We then filtered them into collections of high
+% quality regions by (i) removing the islands with reads residing only
+% on one strand; (ii) removing the islands with $U_i \le 15$; (iii)
+% removing islands with $D_i < 100$. These thresholds were empirically
+% selected and the overall conclusions were robust to their variation.
 %\begin{enumerate}
@@
 %\end{enumerate}
-We used \texttt{FIMO} (version 4.9.1, SK{include command line to show parameters}) \cite{fimo} to identify
-the FoxA1 motif within each enriched region using FoxA1 position weight matrix MA0148.1 from the JASPAR database \cite{jaspar}.
+% We used \texttt{FIMO} (version 4.9.1, \SK{include command line to show parameters}) \cite{fimo} to identify
+% the FoxA1 motif within each enriched region using FoxA1 position weight matrix MA0148.1 from the JASPAR database \cite{jaspar}.
 
 \subsection*{Testing for strand imbalance in FoxA1 ChIP-exo replicates}
-\SK{ This paragraph could benefit from a bit more polishing.}
-
-We used the ChIP-exo QC pipeline to partition the
-mouse genome into a set of islands with their respective summary
-statistics. We filtered the islands with reads in only one strand and
- transformed the FSR into an \emph{imbalance index} that is zero
-when the resulting region is perfectly balanced in terms of forward and reverse strands reads and infinity when it consists of
-reads in one strand exclusively:
+%% \SK{ This paragraph could benefit from a bit more polishing.}
+
+  % \item FoxA1 experiments: (i) removing the islands with reads
+  %   residing only on one strand; (ii) removing the islands with $U_i
+  %   \leq 15$; (iii) removing islands with $D_i < 100$.
+
+\color{blue}
+
+We partitioned the mouse genome into a set of islands by using the
+ChIP-exo QC pipeline. Subsequently, we filtered the islands with reads
+in only one strand. We define an \texttt{Imbalance}, that is
+zero when the region is composed by the same amount of reads in both
+strands and infinity when consists of reads in one strand exclusively,
+as:
+
 \begin{align*}
-  \mbox{Imbalance index} = -\log_{10} (4 \times \mbox{FSR} \times (1 -
-  \mbox{FSR}))
+  \text{\texttt{Imbalance}} = -\log_{10} (4 \times \mbox{FSR}
+  \times (1 - \mbox{FSR}))
 \end{align*}
-We divided the ChIP-exo regions into two classes depending on 
-whether or not they overlap  ChIP-exo peaks, where the we identified the peaks with \texttt{mosaics}
-using the  $\mbox{GC content} + \mbox{Mappability}$ background model \cite{mosaics} (version 2.9.7) and at a false discovery rate (FDR) level of $0.05$ and minimum read threshold of $100$. 
-The  bin size and fragment length for MOSAiCS runs were set to $200$ bps. 
+
+We divided the ChIP-exo regions into two classes depending on whether
+they overlap the respective replicate set of high quality ChIP-exo
+peaks and compared the two classes \texttt{Imbalance}'s distribution.
+
+% . We used a Wilcoxon test to asses that the \texttt{Imbalance}'s
+% median differs between the two classes.
+
+% The  bin size and fragment length for MOSAiCS runs were set to $200$ bps. 
 %We called peaks with an FDR of $5\%$, a threshold of
 %100 and a maximum gap size of 200 bp.
-We further filtered the resulting peaks by
-keeping only the peaks with an average extended ChIP read count of 200 and  used a Wilcoxon test to assess the difference in imbalance index between the two groups of ChIP-exo regions.
-
-%\SK{Is this showing the imbalance index is different between peak and non-peak regions?
-%To show that the class that don't overlap with peaks exhibits heavier
-%tails, we used a Wilcoxon test over the \emph{Imbalance index}.}
+
+% We further filtered the resulting peaks by keeping only the peaks with
+% an average extended ChIP read count of 200 and used a Wilcoxon test to
+% assess the difference in imbalance index between the two groups of
+% ChIP-exo regions.
+\SK{Is this showing the imbalance index is different between peak and
+  non-peak regions?  To show that the class that don't overlap with
+  peaks exhibits heavier tails, we used a Wilcoxon test over the
+  \emph{Imbalance index}.}
+
+\color{black}
+
+% We used the ChIP-exo QC pipeline to partition the mouse genome into a
+% set of islands with their respective summary statistics. We filtered
+% the islands with reads in only one strand and transformed the FSR into
+% an \emph{imbalance index} that is zero when the resulting region is
+% perfectly balanced in terms of forward and reverse strands reads and
+% infinity when it consists of reads in one strand exclusively:
+% \begin{align*}
+%   \mbox{Imbalance index} = -\log_{10} (4 \times \mbox{FSR} \times (1 -
+%   \mbox{FSR}))
+% \end{align*}
+
+% We divided the ChIP-exo regions into two classes depending on 
+% whether or not they overlap  ChIP-exo peaks, where the we identified the peaks with \texttt{mosaics}
+% using the  $\mbox{GC content} + \mbox{Mappability}$ background model \cite{mosaics} (version 2.9.7) and at a false discovery rate (FDR) level of $0.05$ and minimum read threshold of $100$.
 
@@
 
-In all the evaluations using \textit{E. coli} samples, 
-we considered RegulonDB \cite{regulondb} $\sig$ sites as gold-standard. A $sig$ site was deemed 
-as identified  if there existed a binding event within its  20 bps proximity. We defined the
-resolution as the distance from an $\sig$ site to its closest
-predicted binding event and the sensitivity as the fraction of correctly identified
-$\sig$ sites in a genomic region.
+In all the evaluations using \textit{E. coli} samples, we considered
+RegulonDB \cite{regulondb} $\sig$ sites as gold-standard. A $\sig$
+site was deemed as identified if there exists a binding event within
+its 20 bps proximity. We defined the resolution as the distance from
+an $\sig$ site to its closest predicted binding event and the
+sensitivity as the fraction of correctly identified $\sig$ sites in a
+genomic region.
 
@@
 
-We compared dPeak \cite{dpeak} (\url{https://github.com/dongjunchung/dpeak}, version 2.0.1), GEM \cite{gem} (\url{http://groups.csail.mit.edu/cgs/gem/}, version 2.6), MACE \cite{mace} (\url{http://dldcc-web.brc.bcm.edu/lilab/MACE/docs/html/}, version 1.2), peakzilla \cite{peakzilla} (\url{https://github.com/steinmann/peakzilla}).
+We compared dPeak \cite{dpeak}
+(\url{https://github.com/dongjunchung/dpeak}, version 2.0.1), GEM
+\cite{gem} (\url{http://groups.csail.mit.edu/cgs/gem/}, version 2.6),
+MACE \cite{mace}
+(\url{http://dldcc-web.brc.bcm.edu/lilab/MACE/docs/html/}, version
+1.2), peakzilla \cite{peakzilla}
+(\url{https://github.com/steinmann/peakzilla}). \RW{For each method,
+  we calculated the resolution by considering only the annotations
+  overlapping any of the top 400 peaks called by
+  \texttt{MOSAiCS}. Default tuning parameters were used during model
+  fitting for all methods, except for GEM where the peaks were
+  explicitly provided as candidate regions. Although we were able to
+  download \texttt{CexoR} \cite{cexor} (version 1.8 from
+  \emph{bioconductor}), we were unable to use it for the $\sig$
+  experiments because it requires more than one replicate to call
+  ChIP-exo peaks.}
+
 %, Chung et al., 2013; GEM, Guo et al., 2012; MACE,
@@
 
-\SK{Too many repeats of mosaics version etc, try to minimize.}
-
-Candidate regions for
-\texttt{dPeak} and \texttt{GEM} were identified for each replicate of
-ChIP-exo data using the \textbf{MOSAiCS} algorithm \cite{mosaics} (one
-sample analysis using false discovery rate of 1\%) implemented as an \texttt{R}
-package \textbf{mosaics} (version 2.9.7 from
-Bioconductor). We further filtered out candidate regions by
-using the top 400 peaks with highest average ChIP read counts to avoid
-potential false positives based on the exploratory analysis. These
-regions were also explicitly provided to the GEM algorithm as
-candidate regions. Default tuning parameters were used during model
-fitting for all methods. Although we were able to download \texttt{CexoR} \cite{cexor} (version 1.8  from \emph{bioconductor}), we were unable to use it for
-the $\sig$ experiments because of the \SK{*** issues}.
-
-\subsubsection*{dPeak analysis of $\sig$ ChIP-exo and ChIP-seq data}
-
-We compared the estimated binding events predicted by the
-\textbf{MOSAiCS} + \textbf{dPeak} pipeline using reads generated by
-ChIP-exo, PE and SE ChIP-seq protocols. We called peaks at a $5\%$ FDR
-level using \textbf{MOSAiCS} (the $\mbox{GC content} +
-\mbox{Mappability}$ background model for ChIP-exo and the $\mbox{Input only}$
-model for PE and SE ChIP-seq). Then, we deconvolved the peaks into
-binding events with \textbf{dPeak} (version 2.0.1) by considering a
-maximum of 5 binding events within each peak. To avoid false positives, we
-only considered ChIP-exo peaks with average ChIP read count greater than
-3,000 that overlapped both the SE and PE ChIP-seq peaks. Repeating the analysis with other cutoff values led to similar conclusions.
+% \SK{Too many repeats of mosaics version etc, try to minimize.}
+
+% Candidate regions for \texttt{dPeak} and \texttt{GEM} were identified
+% for each replicate of ChIP-exo data using the \textbf{MOSAiCS}
+% algorithm \cite{mosaics} (one sample analysis using false discovery
+% rate of 1\%) implemented as an \texttt{R} package \textbf{mosaics}
+% (version 2.9.7 from Bioconductor). We further filtered out candidate
+% regions by using the top 400 peaks with highest average ChIP read
+% counts to avoid potential false positives based on the exploratory
+% analysis.
+
+% These regions were also explicitly provided to the GEM
+% algorithm as candidate regions. Default tuning parameters were used
+% during model fitting for all methods. Although we were able to
+% download \texttt{CexoR} \cite{cexor} (version 1.8 from
+% \emph{bioconductor}), we were unable to use it for the $\sig$
+% experiments because of the \SK{*** issues}.
 
@@
 \SK{Next paragraph could use more editing.}
-We sub-sampled $N$ reads for both ChIP-exo and SE ChIP-seq
-protocols. For PE ChIP-seq, we sub-sampled $N/2$ pairs equaling to a total of $N$ reads. 
-For each sub-sample, we identified  peaks using MOSAiCS \cite{mosaics}
-($\mbox{GC content} + \mbox{Mappability}$ background model for ChIP-exo and
-$\mbox{Input only}$ for SE and PE ChIP-seq) \SK{not sure: for the maximum sample
-size}. To avoid potential false positives, we considered only the top 500 peaks
-for each data protocol. We defined the number of candidate peaks as
-the number of top sample peaks with at least one predicted dPeak binding event; the number of
-predicted events is the total number of dPeak predicted binding events; the number of identified $\sig$ sites as the number
-of gold-standard $\sig$ sites within 15 bps from an estimated binding
-event; and the resolution as the   minimum distance from a
-gold-standard $\sig$ site to an estimated binding event. We repeated
-this analysis for ten random sub-sampling experiments and reported the median across these experiments.
-
-\SK{Yeah, Materials and Methods has too many repeats, needs to be organized a bit and cleaned up. Figure captions need to be modified and made more easy follow. }
+
+\RW{
+We sub-sampled 10K to 100K (in 10K increments) reads from each of the
+$\sig$ ChIP-exo, PE and SE ChIP-seq datasets. Specifically, for every
+$N$ reads in ChIP-exo and SE ChIP-seq, $N/2$ read pairs were sampled
+from PE ChIP-seq to operate under fixed sequencing costs. For each
+dataset, we called peaks using \texttt{MOSAiCS} \cite{mosaics} and
+estimated TFBS using \texttt{dPeak} \cite{dpeak} by considering at
+most 5 binding sites for each peak. To avoid potential false
+positives, we considered only the top 500 peaks for each data
+protocol. We defined the number of candidate peaks as the number of
+top sample peaks with at least one predicted dPeak binding event; the
+number of predicted events is the total number of dPeak predicted
+binding events; the number of identified $\sig$ sites as the number of
+gold-standard $\sig$ sites within 15 bps from an estimated binding
+event; and the resolution as the minimum distance from a gold-standard
+$\sig$ site to an estimated binding event. We repeated this analysis
+for ten different seeds and reported the median across there
+experiments.
+}
+
+\nocite{robinson14}
+
+% We sub-sampled $N$ reads for both ChIP-exo and SE ChIP-seq
+% protocols. For PE ChIP-seq, we sub-sampled $N/2$ pairs equaling to a
+% total of $N$ reads.  For each sub-sample, we identified peaks using
+% MOSAiCS \cite{mosaics} ($\mbox{GC content} + \mbox{Mappability}$
+% background model for ChIP-exo and $\mbox{Input only}$ for SE and PE
+% ChIP-seq) \SK{not sure: for the maximum sample size}.
+
+% To avoid
+% potential false positives, we considered only the top 500 peaks for
+% each data protocol. We defined the number of candidate peaks as the
+% number of top sample peaks with at least one predicted dPeak binding
+% event; the number of predicted events is the total number of dPeak
+% predicted binding events; the number of identified $\sig$ sites as the
+% number of gold-standard $\sig$ sites within 15 bps from an estimated
+% binding event; and the resolution as the minimum distance from a
+% gold-standard $\sig$ site to an estimated binding event. We repeated
+% this analysis for ten random sub-sampling experiments and reported the
+% median across these experiments.
+
+\subsubsection*{dPeak analysis of $\sig$ ChIP-exo and ChIP-seq data}
+
+\RW{ We compared the estimated binding events predicted by the
+  \textbf{MOSAiCS} + \textbf{dPeak} pipeline using reads generated by
+  ChIP-exo, PE and SE ChIP-seq protocols. We deconvolved the peaks
+  into binding events with \texttt{dPeak} by considering a maximum of
+  5 binding events within each peak. To avoid false positives, we only
+  considered ChIP-exo peaks with average ChIP read count greater than
+  3,000 that overlapped both the SE and PE ChIP-seq peaks. Repeating
+  the analysis with other cutoff values led to similar conclusions.  }
+
+% We compared the estimated binding events predicted by the
+% \textbf{MOSAiCS} + \textbf{dPeak} pipeline using reads generated by
+% ChIP-exo, PE and SE ChIP-seq protocols. We called peaks at a $5\%$ FDR
+% level using \textbf{MOSAiCS} (the $\mbox{GC content} +
+% \mbox{Mappability}$ background model for ChIP-exo and the $\mbox{Input
+%   only}$ model for PE and SE ChIP-seq). Then, we deconvolved the peaks
+% into binding events with \textbf{dPeak} (version 2.0.1) by considering
+% a maximum of 5 binding events within each peak. To avoid false
+% positives, we only considered ChIP-exo peaks with average ChIP read
+% count greater than 3,000 that overlapped both the SE and PE ChIP-seq
+% peaks. Repeating the analysis with other cutoff values led to similar
+% conclusions.
+
+\SK{Yeah, Materials and Methods has too many repeats, needs to be
+  organized a bit and cleaned up. Figure captions need to be modified
+  and made more easy follow. } 
  % Style BST file (bmc-mathphys, vancouver, spbasic).
@@
 
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width =
-  .95 \textwidth]{figures/fig1/chip_explanation_withEnzyme.pdf}
+% figure 1
+
+\RW{The enumeration of some figures is a little bit off, I am going to
+  fix it after we finish the text}
   % \caption{\textbf{Description of ChIP-exo, SE ChIP-Seq and PE
@@
   %   to both ends.}
-  \caption{A TF is bound to the forward (red) and backward (blue)
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width =
+  .45 \textwidth]{figures/fig1/chip_explanation_withEnzyme_nexus.pdf}
+  \includegraphics[width = .45\textwidth,page = 3]{figures/fig1/ChIPseqPET_ChIPexo_tagCount_comparison.pdf}
+  \includegraphics[width =.3\textwidth]{figures/fig1/forward_strand_ratio_comp_old.pdf}
+\includegraphics[width = .6 \textwidth]{figures/fig1/scc_ctcf2.pdf}
+  \includegraphics[width = .4\textwidth,page = 1]{figures/fig1/eukaryotic_bias_CTCF.pdf}
+  \includegraphics[width = .4\textwidth,page = 2]{figures/fig1/eukaryotic_bias_CTCF.pdf}
+  \caption{A) A TF is bound to the forward (red) and backward (blue)
     strand of the DNA. Then, is sonicated: For ChIP-exo a exonuclease
@@
     fragment to a fixed distance from the bound protein, finally is
-    subjected to Immunoprecipitation and amplification. For both
-    ChIP-exo and SE ChIP-Seq an adapter is ligated (green triangles)
-    at the $5\prime$ ends, while for PE ChIP-Seq is ligated to both
-    ends.}
-  \label{fig:chip_diagram}
+    subjected to Immunoprecipitation and amplification. For ChIP-nexus
+    a random barcode is added on the $3\prime$ end, and transferred to
+    the $5\prime$ stopping base by self-circularization, after the
+    library is amplified and the reads aligned, reads with identical
+    barcode and start position are removed. For both ChIP-exo and SE
+    ChIP-seq an adapter is ligated (green triangles) at the $5\prime$
+    ends, while for PE ChIP-seq is ligated to both ends.  B) Hexbin
+    plot of PE ChIP-seq bin counts vs ChIP-exo bin counts. C) Forward
+    Strand Ratio densities for SE ChIP-seq and ChIP-exo peaks. D) SCC
+    curves for human CTCF on HeLa cell lines.  The SCC curve for the
+    ChIP-exo sample from \cite{exo1} is shown in the left panel, and
+    the SCC for ChIP-seq samples from \cite{encode1} are shown in the
+    right panel. The ChIP-exo curve shows local maxima at the motif
+    and read length. Both SE ChIP-seq curves are maximized at the
+    fragments length and show a local maxima at its read length. E)
+    Mappability score vs mean ChIP tag counts with 0.95 confidence
+    bands. F) GC - content score vs mean ChIP tag counts with 0.95
+    confidence bands.}
+  \label{fig:1}
 \end{figure}
+
+    % \textbf{https://www.encodeproject.org/experiments/ENCSR000AOA/}
+
 \newpage
+
+% figure 2
 
@@
   \centering
-  \includegraphics[width = .46\textwidth,page = 3]{figures/fig2/ChIPseqPET_ChIPexo_tagCount_comparison.pdf}
-  \includegraphics[width =.46\textwidth]{figures/fig2/forward_strand_ratio_comp_old.pdf}
-  \includegraphics[width = .46\textwidth,page = 1]{figures/fig2/eukaryotic_bias_CTCF.pdf}
-  \includegraphics[width = .46\textwidth,page = 2]{figures/fig2/eukaryotic_bias_CTCF.pdf}
-  \caption{ A) Hexbin plot of PE ChIP-Seq bin counts vs ChIP-exo bin
-    counts. B) Forward Strand Ratio densities for SE ChIP-Seq and
-    ChIP-exo peaks. C) Mappability score vs mean ChIP tag counts with
-    0.95 confidence bands. D) GC - content score vs mean ChIP tag
-    counts with 0.95 confidence bands.}
-  \label{fig:comp}
+  \includegraphics[width = \textwidth]{figures/fig2/coverage_diagram4.pdf}
+  \caption{\RW{The ChIP-exo reads are partitioned by keeping the
+      regions in the genome composed by the undigested fragments. For
+      each region the following summary statistics are calculated:
+      Average Read Coefficient (ARC), Unique Read Coefficient (URC)
+      and Forward Strand Ratio (FSR). These statistics are visualized
+      as: A) ARC vs. URC plots, which presents the overall balance
+      between library complexity and enrichment; there are two arms,
+      one with low ARC, that corresponds to regions formed by few
+      aligned positions and one where the URC decreases as the ARC
+      increases, that corresponds to potentially enriched regions. B)
+      Region Composition plot, which shows the strand composition for
+      all the regions formed by a min. number of reads; low depth
+      regions are commonly formed by undigested reads in one unique
+      strand, while high depth regions are usually formed by reads in
+      both. C) FSR distribution plot, it illustrates the FSR's
+      distribution as as higher depth regions are being considered: In
+      a high quality experiment the FSR's distribution is tightly
+      centered around the FSR's median.}}
+
+  % There are two arms: First, one with low ARC, which corresponds to
+  % regions formed by few aligned positions; Second, where the URC
+  % decreases as the ARC increases.
+
+  % Regions with low depth tend to be formed by undigested reads from
+  % one unique strand, while regions with higher signal are usually
+  % formed by reads in both strands.
+
+  % In a high quality sample, the median is around 0.5, and the other
+  % quantiles reach that value quickly.
+  \label{fig:2}
 \end{figure}
@@
 \newpage
+
+% figure 3
 
@@
   \centering
-\includegraphics[width = .8 \textwidth]{figures/fig3/scc_ctcf2.pdf}
-% \includegraphics[width = .8\textwidth]{/p/keles/ChIPexo/volume3/ChIPexo/figs/for_paper/EColi_strand_cross_corr.pdf}
-% \caption{\textbf{SCC curves for $\sig$ samples.} The ``phantom peak''
-%   and the summit that corresponds to the read and fragment length
-%   respectively are confounded due to the exonuclease digestion.}
-\caption{SCC curves for human CTCF on HeLa cell lines.  The SCC curve
-  for the ChIP-exo sample from \cite{exo1} is shown in the left panel,
-  and the SCC for ChIP-Seq samples from
-  \textbf{https://www.encodeproject.org/experiments/ENCSR000AOA/} are
-  shown in the right panel. The ChIP-exo curve shows local maxima at
-  the motif and read length. Both SE ChIP-Seq curves are maximized at
-  the fragments length and show a local maxima at its read length. }
-  \label{fig:scc_exo}
+  \includegraphics[width = .8\textwidth,page = 1]{figures/fig3/Carroll_FoxA1_mouse_enrichment.pdf}
+  \includegraphics[width = .8\textwidth,page = 3]{figures/fig3/Carroll_FoxA1_mouse_strand_imbalance.pdf} 
+  \includegraphics[width = .8\textwidth,page = 1]{figures/fig3/Carroll_FoxA1_mouse_strand_imbalance.pdf} 
+  \caption{\RW{Diagnostic plots generated by \pname{}. Comparison of
+      A) ARC vs. URC plot, B) Region Composition plot and C) FSR
+      distribution plot between the three FoxA1 replicates in mouse
+      liver tissue from \cite{exoillumina}.}}
+  \label{fig:3}
 \end{figure}
@@
 \newpage
+
+% , there is a slight separation into two strong arms, one corresponds
+% to low $\mbox{ARC}$ and varying $\mbox{URCR}$, and for the other
+% $\mbox{URCR}$ decreases as $\mbox{ARC}$ increases
+
+%     , all quantiles approach to the median as the lower bound
+%     increases
+
+%     , in a good sample the single-stranded regions are going to be
+%     filtered out quickly as in the middle row.
+
+\newpage
+
+% figure 4
 
@@
   \centering  
-  \includegraphics[width = \textwidth]{figures/fig4/coverage_diagram3.pdf}
-  % \caption{\textbf{Flowchart of the ChIP-exo quality control
-  %     pipeline.} A) ARC vs. URCR, this plot presents a global view of
-  %   the balance between library complexity and enrichment. There are
-  %   two arms: First, one with low ARC, which corresponds to regions
-  %   formed by few aligned positions; Second, where the URCR decreases
-  %   as the ARC increases. B) Min. depth vs Proportions of regions,
-  %   this plot shows the strand composition for all the regions formed
-  %   by a min. number of reads. Regions with low depth tend to be
-  %   formed by undigested reads from one unique strand, while regions
-  %   with higher signal are usually formed by reads in both strands. C)
-  %   Min depth vs. FSR, this plot depicts how quickly the FSR's
-  %   distribution approximate the median. In a high quality sample, the
-  %   median is around 0.5, and the other quantiles reach that value
-  %   quickly.}
-  \caption{The ChIP-exo reads are partitioned by keeping the regions
-    in the genome formed by the undigested fragments. For each region,
-    we calculate a collection of summary statistics, which are
-    visualized as A) ARC vs. URCR, this plot presents a global view of
-    the balance between library complexity and enrichment. There are
-    two arms: First, one with low ARC, which corresponds to regions
-    formed by few aligned positions; Second, where the URCR decreases
-    as the ARC increases. B) Min. depth vs Proportions of regions,
-    this plot shows the strand composition for all the regions formed
-    by a min. number of reads. Regions with low depth tend to be
-    formed by undigested reads from one unique strand, while regions
-    with higher signal are usually formed by reads in both strands. C)
-    Min depth vs. FSR, this plot depicts how quickly the FSR's
-    distribution approximate the median. In a high quality sample, the
-    median is around 0.5, and the other quantiles reach that value
-    quickly.}
-  \label{fig:qcdiagram}
+  \includegraphics[width = .23 \textwidth, page = 1]{figures/fig4/FoxA1_barplots.pdf}
+  \includegraphics[width = .23 \textwidth, page = 3]{figures/fig4/FoxA1_barplots.pdf}
+  \includegraphics[width =.4\textwidth]{figures/fig4/FoxA1_profiles_around_motif.pdf}
+  \includegraphics[width = .25\textwidth,page = 2]{figures/fig4/FoxA1_matched_motif_sequence.pdf}
+  \includegraphics[width = .25\textwidth,page = 3]{figures/fig4/FoxA1_matched_motif_sequence.pdf}
+  \includegraphics[width = .25\textwidth,page = 1]{figures/fig4/FoxA1_matched_motif_sequence.pdf}  
+  \includegraphics[width = .8\textwidth,page = 1]{figures/fig4/FoxA1_fimo_score.pdf}
+  \includegraphics[width = .8\textwidth,page = 2]{figures/fig4/strand_imbalance_FoxA1.pdf}
+  \caption{Using the mouse FoxA1 experiment from \cite{exoillumina}
+    and the FoxA1 motif with MA0148.1 id in the \texttt{JASPAR}
+    database: \RW{ A) Number of high quality regions where the FoxA1
+      motif was searched. B) Proportion of candidate regions with
+      motif. C) FoxA1 Average Coverage plots centered around motif
+      start positions separated by replicate and strand. D) Sequence
+      composition of sites with motif for each replicate. E)
+      Comparison of the top 50, 100, 250, 500, 1000 and 2000 FIMO
+      scores for each replicate. F) Strand Imbalance comparison of
+      regions with at least 50, 100, 150 and 250 reads overlapping
+      with ChIP-exo peaks for each replicate.}}
+% Strand imbalance using the mouse FoxA1 experiment from
+%     \cite{exoillumina}:  C)
+%     $-\log_{10}(\text{p.value})$ of testing if the imbalance
+%     distributions differs when ChIP-exo regions overlap their peaks.}
+  \label{fig:4}
 \end{figure}
 
+ % B) Number of
+ %    candidate sites for each replicate. C) Percentage of candidate
+ %    sites where the FoxA1 motif was detected. D) Average coverage
+ %    around FoxA1 motif. Base distribution for matched sequence for
+ %    Rep-1 (E), Rep-2 (F) and Rep-3 (G).
+
 \newpage
+
+% figure 5
 
@@
   \centering
-  \includegraphics[width = .9  \textwidth,page = 1]{figures/fig5/FoxA1_enrichment.pdf}
+  \includegraphics[width = .9  \textwidth,page = 1]{figures/fig5/QC_pipeline_eval_boxplot.pdf}
 \newline
-  \includegraphics[width = .23 \textwidth, page = 1]{figures/fig5/FoxA1_barplots.pdf}
-  \includegraphics[width = .23 \textwidth, page = 3]{figures/fig5/FoxA1_barplots.pdf}
-  \includegraphics[width =.4\textwidth]{figures/fig5/FoxA1_profiles_around_motif.pdf}
-  \includegraphics[width = .3\textwidth,page = 2]{figures/fig5/FoxA1_matched_motif_sequence.pdf}
-  \includegraphics[width = .3\textwidth,page = 3]{figures/fig5/FoxA1_matched_motif_sequence.pdf}
-  \includegraphics[width = .3\textwidth,page = 1]{figures/fig5/FoxA1_matched_motif_sequence.pdf}
-  \caption{Using the mouse FoxA1 experiment from \cite{exoillumina}:
-    A) Hexbin plots of $\mbox{ARC}$ against $\mbox{URCR}$, there is a
-    slight separation into two strong arms, one corresponds to low
-    $\mbox{ARC}$ and varying $\mbox{URCR}$, and for the other
-    $\mbox{URCR}$ decreases as $\mbox{ARC}$ increases. B) Number of
-    candidate sites for each replicate. C) Percentage of candidate
-    sites where the FoxA1 motif was detected. D) Average coverage
-    around FoxA1 motif. Base distribution for matched sequence for
-    Rep-1 (E), Rep-2 (F) and Rep-3 (G).}
-  \label{fig:enrich}
+  \includegraphics[width = .9  \textwidth,page = 2]{figures/fig5/QC_pipeline_eval_boxplot.pdf}
+\newline
+\includegraphics[width = .45\textwidth,page = 1]{figures/fig5/TBP_param_depth_trend.pdf}
+  % \includegraphics[width = .45  \textwidth,page = 1]{figures/fig5/K562_TBP_sample.pdf}
+  % \includegraphics[width = .45  \textwidth,page = 2]{figures/fig5/K562_TBP_sample.pdf}
+\includegraphics[width = .5\textwidth]{figures/fig5/TBP_scores_comparison.pdf}
+\caption{\RW{Comparison of all estimated A) $\beta_1$ and B) $\beta_2$
+    for all eukaryotic ChIP-exo/nexus samples. Comparison of the
+    average estimated A) $\beta_1$ and B) $\beta_2$ for the
+    ChIP-exo/nexus TBP samples in K562 cell lines when sub-sampling
+    20M to 50M reads. E) Comparison of the top 50, 100, 250, 500,
+    1000, 2000, 4000 and 8000 FIMO scores for each TBP experiment. The
+    scores were calculated with the TBP MA0108.1 motif in the
+    \texttt{JASPAR} database.}}
+% \caption{Comparison of adjusted Average Read Coverage and $\mbox{ARC}$
+%   bias for eukaryotic genome ChIP-exo and ChIP-nexus experiment. A)
+%   Boxplot of adjusted $\mbox{ARC}$. B) Boxplot of large sequencing
+%   depth $\mbox{ARC}$'s bias. Comparison of sub-sampled ChIP-exo and
+%   ChIP-nexus experiments of TBP factor in human K562 cell lines: C)
+%   Adjusted $\mbox{ARC}$ and D) $\mbox{ARC}$'s bias. E) FIMO scores of
+%   TBP ChIP-exo and ChIP-nexus samples for best motif scores.}
+  \label{fig:5}
 \end{figure}
 
-{\color{red} \emph{need to add labels for each figure and think about
-    order, in the supplement we have ecdf from fimo score or pval, and
-    overlaps with peaks called by mosaics.
-\newline
-Also, need to change labels from repk to Rep-k
-}}
-
 \newpage
+
+% figure 6
 
@@
   \centering
-  \includegraphics[width = .9\textwidth,page = 1]{figures/fig6/Strand_imbalance.pdf} 
-  \newline
-  \includegraphics[width = .9\textwidth,page = 2]{figures/fig6/Strand_imbalance.pdf} 
-  \includegraphics[width = .4\textwidth]{figures/fig6/Carroll_FSR_depth_VS_pvalWilcoxon.pdf}
-  \caption{Strand imbalance using the mouse FoxA1 experiment from
-    \cite{exoillumina}: A) FSR distribution quantiles as the lower
-    depth regions are being filtered out, all quantiles approach to
-    the median as the lower bound increases. B) Stacked histogram with
-    the proportion of regions that are formed by two strands or only
-    one, in a good sample the single-stranded regions are going to be
-    filtered out quickly as in the middle row. C)
-    $-\log_{10}(\text{p.value})$ of testing if the imbalance
-    distributions differs when ChIP-exo regions overlap their peaks.}
-  \label{fig:strand}
+  \includegraphics[width = .45\textwidth,page = 1]{figures/fig6/methods_comp_FDR5_topM400.pdf}
+  \includegraphics[width = .45\textwidth,page = 2]{figures/fig6/methods_comp_FDR5_topM400.pdf}
+   \includegraphics[width = .45\textwidth,page = 1]{figures/fig6/sigma_delta_old_densities.pdf}
+   \includegraphics[width = .45\textwidth,page = 2]{figures/fig6/sigma_delta_old_densities.pdf}   
+   \caption{\RW{Resolution comparison between dPeak, Gem, Mace and
+       Peakzilla using the A) E1-1 and B) E1-2 \textit{E.coli}
+       ChIP-exo experiments. Estimated C) $\delta$ and D) $\sigma$
+       parameter densities for ChIP-exo (E1-2 sample) and SE ChIP-seq
+       (S1-1 sample).}}
+% Resolution is the defined as the distance
+%        between a RegulonDB annotation and the closest predicted
+%        binding event. 
+  % \caption{Comparison of the resolution between dPeak, Gem, Mace and
+  %   Peakzilla methods with \emph{E.Coli} aerobic experiment from the
+  %   first biosample A) Replicate 1 and B) Replicate 2. Resolution is
+  %   defined as the minimum distance between a RegulonDB annotation and
+  %   a predicted binding event. C) $\delta$ parameter in dPeak measures
+  %   average distance of the reads to their respective binding site. In
+  %   ChIP-exo data, reads were located much closer to the binding site
+  %   than in SET ChIP-Seq. D) $\sigma$ parameter measure the dispersion
+  %   of reads around each binding site. In ChIP-exo data, reads showed
+  %   less variation around the their respective binding sites compared
+  %   to SET ChIP-Seq.}
+  \label{fig:6}
 \end{figure}
@@
 \newpage
+
+% figure 7
+% \RW{For now, I kept the next two figures separated. If we decide to
+%   include the analysis of Fig8 in the main text, then I consider we
+%   should merge both into fig 7. ChIP-exo depths is rougly the half of
+%   PE ChIP-seq which may work into our advantage since we are obtaining
+%   comparable resolution and better sensitivity with less reads. On the
+%   other hand, the SE ChIP-seq depth is roughly the half from the
+%   ChIP-exo making that comparison a little bit unfair (which may not
+%   be a big deal since the other comparison is the most important
+%   one).}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width = .9\textwidth]{figures/fig7/saturation_analysis_old.pdf}
+  \caption{Comparison of the number of A) candidate regions, B)
+    predicted events, C) identified targets and D) resolution among
+    ChIP-exo, PE ChIP-seq and SE ChIP-seq.}
+ % A RegulonDB binding event
+ %    was deemed identified if a binding event was estimated at a $\pm$
+ %    15 vicinity of it.
+  \label{fig:7}
+\end{figure}
+
+% figure 8
 
@@
   \centering
-  \includegraphics[width = .9  \textwidth,page = 1]{figures/fig7/QC_pipeline_eval_boxplot.pdf}
-\newline
-  \includegraphics[width = .9  \textwidth,page = 2]{figures/fig7/QC_pipeline_eval_boxplot.pdf}
-\newline
-  \includegraphics[width = .45  \textwidth,page = 1]{figures/fig7/K562_TBP_sample.pdf}
-  \includegraphics[width = .45  \textwidth,page = 2]{figures/fig7/K562_TBP_sample.pdf}
-  \caption{Comparison of adjusted Average Read Coverage and
-    $\mbox{ARC}$ bias for eukaryotic genome ChIP-exo and ChIP-nexus
-    experiment. A) Boxplot of adjusted $\mbox{ARC}$. B) Boxplot of
-    large sequencing depth $\mbox{ARC}$'s bias. Comparison of
-    sub-sampled ChIP-exo and ChIP-nexus experiments of TBP factor in
-    human K562 cell lines: C) Adjusted $\mbox{ARC}$ and D) $\mbox{ARC}$'s
-    bias.}
-  \label{fig:eval}
+  \includegraphics[width = .46\textwidth]{figures/fig8/sensitivity_exo_olda_data.pdf}
+  \includegraphics[width = .46\textwidth]{figures/fig8/resolution_by_dataset_old_data.pdf}
+  \caption{Comparison of A) sensitivity and B) resolution between
+    ChIP-exo and ChIP-seq data.}
+  % Sensitivity is defined as the proportion of RegulonDB annotations
+  % identified using each data. Resolution is defined as the distance
+  % between RegulonDB annotation and its closest prediction.
+  \label{fig:8}
 \end{figure}
 
-\newpage
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = .45\textwidth,page = 1]{figures/fig8/methods_comp_FDR5_topM400.pdf}
-  \includegraphics[width = .45\textwidth,page = 2]{figures/fig8/methods_comp_FDR5_topM400.pdf}
-   \includegraphics[width = .45\textwidth,page = 1]{figures/fig8/sigma_delta_old_densities.pdf}
-   \includegraphics[width = .45\textwidth,page = 2]{figures/fig8/sigma_delta_old_densities.pdf}   
-  \caption{Comparison of the resolution between dPeak, Gem, Mace and
-    Peakzilla methods with \emph{E.Coli} aerobic experiment from the
-    first biosample A) Replicate 1 and B) Replicate 2. Resolution is
-    defined as the minimum distance between a RegulonDB annotation and
-    a predicted binding event. C) $\delta$ parameter in dPeak measures
-    average distance of the reads to their respective binding site. In
-    ChIP-exo data, reads were located much closer to the binding site
-    than in SET ChIP-Seq. D) $\sigma$ parameter measure the dispersion
-    of reads around each binding site. In ChIP-exo data, reads showed
-    less variation around the their respective binding sites compared
-    to SET ChIP-Seq.}
-  \label{fig:methods_comp}
-\end{figure}
-
-\newpage
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width = .9\textwidth]{figures/fig9/saturation_analysis_old.pdf}
-  \caption{Comparison of the number of A) candidate regions, B)
-    predicted events, C) identified targets and D) resolution among
-    ChIP-exo, PE ChIP-Seq and SE ChIP-Seq. RegulonDB annotations are
-    considered as a gold standard. A gold standard binding events was
-    deemed identified if a binding event was estimated at a $\pm$ 15
-    vicinity of it.}
-  \label{fig:design}
-\end{figure}
-
-\newpage
-\begin{figure}[h!]
-  \centering
-  \includegraphics[width = .46\textwidth]{figures/fig10/sensitivity_exo_olda_data.pdf}
-  \includegraphics[width = .46\textwidth]{figures/fig10/resolution_by_dataset_old_data.pdf}
-  \caption{Comparison of A) sensitivity and B) resolution between
-    ChIP-exo and ChIP-Seq data. Sensitivity is defined as the
-    proportion of RegulonDB annotations identified using each
-    data. Resolution is defined as the distance between RegulonDB
-    annotation and its closest prediction.}
-  \label{fig:reso_all}
-\end{figure}
-
 \end{document}
 
-% LocalWords:  Ntimes overlaping gaussian repk
+% LocalWords:  Ntimes overlaping gaussian repk depts

### 2016-07-05 paper/paper.tex
previous version of the paper

--- 2016-05-12 rmarkdown/paper.tex
@@
-% \VignetteEngine{knitr::knitr}
-\documentclass{bmcart}\usepackage[]{graphicx}\usepackage[]{color}
-%% maxwidth is the original width if it is less than linewidth
-%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
-\makeatletter
-\def\maxwidth{ %
-  \ifdim\Gin@nat@width>\linewidth
-    \linewidth
-  \else
-    \Gin@nat@width
-  \fi
-}
-\makeatother
-
-\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
-\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
-\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
-\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
-\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
-\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
-\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
-\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
-\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
-\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
-
-\usepackage{framed}
-\makeatletter
-\newenvironment{kframe}{%
- \def\at@end@of@kframe{}%
- \ifinner\ifhmode%
-  \def\at@end@of@kframe{\end{minipage}}%
-  \begin{minipage}{\columnwidth}%
- \fi\fi%
- \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
- \colorbox{shadecolor}{##1}\hskip-\fboxsep
-     % There is no \\@totalrightmargin, so:
-     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
- \MakeFramed {\advance\hsize-\width
-   \@totalleftmargin\z@ \linewidth\hsize
-   \@setminipage}}%
- {\par\unskip\endMakeFramed%
- \at@end@of@kframe}
-\makeatother
-
-\definecolor{shadecolor}{rgb}{.97, .97, .97}
-\definecolor{messagecolor}{rgb}{0, 0, 0}
-\definecolor{warningcolor}{rgb}{1, 0, 1}
-\definecolor{errorcolor}{rgb}{1, 0, 0}
-\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX
-
-\usepackage{alltt}
+\documentclass{bmcart}
 
@@
 
-% \usepackage{hyperref}
-% \hypersetup{
-%      colorlinks   = true,
-%      citecolor    = black
-% }
-% \hypersetup{linkcolor=black}
-% \hypersetup{linkcolor=black}
-% \hypersetup{linkcolor=black}
-
 \newcommand{\beginsupplement}{%
@@
 \newcommand{\sig}{\sigma^{70}}
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+
 \begin{document}
@@
 
-    % both protocols are comparable in resolutions and sensitivity for
-    % closely located binding events, but as the distance between
-    % binding events increases ChIP-exo shows higher sensitivity that
-    % PE ChIP-Seq. Finally, at fixed sequencing depths, ChIP-exo
-    % provides higher sensitivity, specificity and spatial resolution
-    % than PE ChIP-Seq.
   \end{abstract}
@@
 libraries.
-
-% Refers to... First, DNA libraries... I think that these might be
-% too much details for the introduction. How about discussing them in
-% the results when we investigate the related
 
@@
 \label{sec:results}
-% CTCF factor in human \cite{exo1}; ER factor in
-% human and FoxA1 factor in mouse (Serandour et al., 2013
-% \cite{exoillumina}); Glucocorticoid receptor (GR) in IMR90, K562 and
-% U2OS cell lines (Starick et al., 2015 \cite{starick15}); TATA-box
-% Binding Protein (TBP) in K562 cell lines (Venters and Pugh, 2013
-% \cite{venters13}). Furthermore, we also generated ChIP-exo and
-% ChIP-Seq data for $\sig$ factor in \emph{Escherichia Coli}
-% (\emph{E. Coli}) measured under aerobic ($ + O_2$) condition, and
-% treated by rifampicin by 0 and 20 minutes (courtesy of Professor
-% Robert Landick's lab). Additionally we gathered the ChIP-nexus from
-% TATA-box Binding Protein (TBP) in K562 cell lines; MyC and Max in S2
-% cell lines in \emph{Drosophila Melanogaster} (\emph{D. Melanogaster});
-% Twist and Dorsal in \emph{D. Melanogaster} embryo cell lines (He et
-% al., 2015 \cite{chipnexus}).
+
 \subsection*{Publicly available and novel datasets}
@@
 Dorsal in \emph{D. Melanogaster} embryo cell lines.
-
-% Therefore, investigation of ChIP-exo's potential for
-% identification and differentiation of closely spaced binding sites is
-% invaluable for elucidating the transcriptional networks of prokaryotic
-% genomes.
 
@@
 
-%% We can observe that the PBC is low, hence it would be incorrectly
-%% suggested to repeat the experiment.
-
 \arrayrulewidth=.1mm
@@
     \multirow{4}{1em}{2} & Aerobic & No Rif. & 1 & 902,921 & 13.77 & 1.1270 & 0.2689\\
-    & Aerobic & No Rif. & 1 & 1,852,124 & 17.91 & 1.5275 & 0.2590\\
-    & Aerobic & Rif. 20 min & 2 & 2,104,427 & 29.60 & 1.2844 & 0.2584\\
+    & Aerobic & Rif. 20 min & 1 & 1,852,124 & 17.91 & 1.5275 & 0.2590\\
+    & Aerobic & No Rif. & 2 & 2,104,427 & 29.60 & 1.2844 & 0.2584\\
     & Aerobic & Rif. 20 min & 2 & 11,548,572 & 13.08 & 1.5122 & 0.1510 \\
@@
 
-% that both
-% local maxima are hard to differentiate in the SCC curves for the
-% $\sig$ ChIP-exo datasets used to calculate the NSC values in Table
-% \ref{tab:qc}.
-
-% It is calculated as the correlation between both strand
-% coverages, where each one is shifted $\delta / 2$ bp towards the
-% $3\prime$ direction
-
 \subsubsection*{ChIP-exo Quality Control Pipeline}
@@
 
-% The ChIP-exo QC pipeline provides a model free framework to analyze
-% the biases in a ChIP-exo experiment by taking advantage of the
-% exonuclease enzyme that digests the non-enriched regions to partition
-% the genome, calculates common ChIP-Seq QC metrics in ChIP-exo regions
-% locally and allows the interpretation of these metrics by the use of
-% diagnostic plots.
-
 \textbf{Analysis of enrichment and library complexity in FoxA1
@@
 
-% To verify this statement, we
-% extracted the sequences around high confidence binding events and look
-% for the FoxA1 motif using FIMO \cite{fimo}. Figure \ref{fig:enrich}B
-% shows the number of candidate regions, which shows that the first
-% replicate is being allocated into more enriched regions than the other
-% ones. Figure \ref{fig:enrich}C shows that for the first and third
-% replicates, the FoxA1 motif is being detected in roughly the same
-% proportion of sites, and finally in \ref{fig:enrich}D we observe that
-% the first and third replicate can detect the FoxA1 motif with the same
-% significance, while the second replicate does not.
-
 \textbf{Analysis of strand imbalance in FoxA1 ChIP-exo data}
-
-% Using the Fox A1 in mouse liver cell lines generated by Serandour et
-% al., 2013\nocite{exoillumina}
 
@@
 enriched region.
-
-% \textbf{Analysis of ChIP-nexus data with the QC pipeline}
-
-% We applied the QC pipeline to the ChIP-nexus experiments and found
-% that the overall pattern seen in the visualizations is similar to a
-% high quality ChIP-exo experiment. Figure \ref{fig:nexus}A illustrates
-% a typical pattern of the ARC vs. URCR plot with ChIP-nexus data: Both
-% arms seems to be more narrow and therefore the separation between them
-% seems to be more noticeable, this seems reasonable since libraries
-% generated by ChIP-nexus usually yields higher complexity than
-% libraries generated by ChIP-exo. Additionally, while in ChIP-exo
-% experiments is common to observe regions formed by undigested
-% fragments in only one strand, we can observe that in ChIP-nexus
-% experiment the occurrence of this type of regions is reduced. Figure
-% \ref{fig:nexus}B shows that in a ChIP-nexus experiment the FSR
-% distribution approximates the median more quickly than in a ChIP-exo
-% experiment and Figure \ref{fig:nexus}C we observe that most of these
-% single-stranded regions are formed by few fragments, which means that
-% good ChIP-nexus experiment are less likely to show very imbalanced
-% regions.
 
@@
 
-% \color{blue}
-% outline
-
-% \begin{enumerate}
-% \item \st{Available data, the motivation of this part is two fold:}
-%  \begin{itemize}
-%  \item \st{we want to list all data sets we obtained}
-%  \item \st{we want to emphasize that for the sig70 datasets we have a
-%    unique design that allows us to compare chip exo with SE and PE
-%    chipseq , the later one for the first time}
-%  \end{itemize}
-% \item \st{comparison with chip seq} %? with and without dpeak ?}
-% \item \st{only in methods. list the align algorithm used}
-% \item \st{chip seq qc metrics}
-% \item \st{chip exo qc pipeline, description and justification of the steps}
-% \begin{itemize}
-% \item \st{need to justify the sampling step. for this we may use the depth
-%   table and add the factor to normalize one depth to another. explain
-%   the confounding of pcr amplification and the reduction of the number
-%   of possible mappable positions.}
-% \item \st{case study with FoxA1, analysis of enrichment vs library
-%   complexity and analysis and the strand imbalance tools. in methods
-%   we need to add which tests did we perform to show that the pipeline
-%   works. how do we expect from the diagnostics and how to interpret
-%   the plots.}
-% \item \st{analysis of why the pipeline works. in methods need to specify
-%   how and in results need to specify why it works, and what are the
-%   conclusions from it}
-% \end{itemize}
-% \item high resolution analysis with chip exo and comparison with PE
-%   chip seq
-%   \begin{itemize}
-%   \item justify what do we used dpeak (in methods need to list the
-%     other methods used). perhaps we can add gof plots in
-%     supplement. we also want to compare with the same method,
-%     therefore we can use the fact that dpeak have a unified model for
-%     PE and SE dpeak
-%   \item after comparing with other methods we can compare with PE and
-%     SE ChIP-seq, this comparison gives also resolution and sensitivity
-%   \item compare samples under same budget of reads
-%   \end{itemize} 
-% \end{enumerate}
-
-% \color{black}
-% Figure \ref{fig:methods_comp} compares the resolution defined as the
-% minimum distance between a RegulonDB annotation and a binding site
-% predicted by either Peakzilla \cite{peakzilla}, MACE \cite{mace}, GEM
-% \cite{gem} or dPeak \cite{dpeak}. In a good dataset such as both of
-% the ChIP-exo experiments under aerobic (panel 931 and 933) conditions,
-% all the methods are comparable in resolution, and dPeak slightly
-% outperforms the rest. On the other hand, for experiments dPeak
-% identified binding sites with higher resolution than Mace and
-% Peakzilla, but with low resolution than Gem. This may be due that the
-% fact that Gem uses sequence information in addition to the aligned
-% $5\prime$ end counts that dPeak uses.
-
 \subsubsection*{Systematic comparison of ChIP-Seq vs ChIP-exo under
@@
 ChIP-exo or SE ChIP-Seq.
-
-% For each sampled dataset we applied our lower-to-higher resolution
-% pipeline by calling peaks with MOSAiCS \cite{mosaics} and then
-% deconvolving the binding events by using dPeak \cite{dpeak}. For the
-% ChIP-exo experiments we called peaks by using the GC-content and
-% mappability models with MOSAiCS, since it's background is usually
-% composed of scattered reads across the genome; and for the ChIP-Seq
-% datasets we used their respective Input samples.
 
@@
 
-% Figure \ref{fig:saturation_rif} shows an analogous analysis but using
-% the $\sig$ replicates with and without rif treatment. The left, middle
-% and right columns show the fixed depth against the number of predicted
-% events, identified targets and resolution being compared at a fixed
-% depth level. The behavior of this quantities seems to be opposite to
-% the one as in Figure \ref{fig:design}, hence we used the ChIP-exo QC
-% pipeline in the fixed depth ChIP-exo experiments. Figure
-% \ref{fig:exoQC_sat_aero} shows scatter plots of ARC vs URCR for
-% several fixed sample sizes, as the fixed depth increases the two arm
-% pattern becomes more distinctive while for lower depth, it seems that
-% the majority of the sampled reads were aligned to enriched regions. On
-% the other hand in Figures \ref{fig:exoQC_sat1} to
-% \ref{fig:exoQC_sat4}, we used the ChIP-exo QC pipeline on the samples
-% that are outperformed by PE and SE ChIP-Seq. For a fixed low depth, we
-% can see that the majority of the reads are being aligned to
-% non-enriched regions since the vertical arm seems to be stronger for
-% all 2 conditions and replicates; for higher depth we can see low URCR
-% values being predominant, which indicates that the majority of the
-% regions being formed by few positions with a higher read
-% concentration. In low complexity regions, the reads are being aligned
-% to fewer positions but there is no control over the amount of reads
-% mapped. Hence, those regions are more likely to being strand-imbalance
-% which in turn may bias the binding site estimate and therefore
-% decrease the number of identified targets or increase an experiments
-% resolution.
-
 \subsubsection*{Comparison with ChIP-Seq data using dPeak}
@@
 considered to be identified if the distance from the closest dPeak
-binding site estimate was less that or equal to $20$
-bp. That way, the sensitivity is defined as the proportion of
-RegulonDB annotations identified in a peak and the resolution is
-defined as the minimum distance between a RegulonDB annotation and the
-closest dPeak binding site estimate. Figure \ref{fig:reso_all}A shows
-that the sensitivity increases as the mean distance between binding
-events increases. When the binding events in a peak are closer to each
-other, both ChIP-exo and PE ChIP-Seq are comparable, as the distance
-increases ChIP-exo identifies a higher proportion of the RegulonDB
-annotations; additionally SE ChIP-Seq is significantly less sensitive
-than both ChIP-exo and PE ChIP-Seq. Figure \ref{fig:reso_all}B shows
-that ChIP-exo and PE ChIP-Seq are comparable in resolution, while both
+binding site estimate was less that or equal to 20 bp. That way, the
+sensitivity is defined as the proportion of RegulonDB annotations
+identified in a peak and the resolution is defined as the minimum
+distance between a RegulonDB annotation and the closest dPeak binding
+site estimate. Figure \ref{fig:reso_all}A shows that the sensitivity
+increases as the mean distance between binding events increases. When
+the binding events in a peak are closer to each other, both ChIP-exo
+and PE ChIP-Seq are comparable, as the distance increases ChIP-exo
+identifies a higher proportion of the RegulonDB annotations;
+additionally SE ChIP-Seq is significantly less sensitive than both
+ChIP-exo and PE ChIP-Seq. Figure \ref{fig:reso_all}B shows that
+ChIP-exo and PE ChIP-Seq are comparable in resolution, while both
 protocols significantly outperform SE ChIP-Seq.
-
-% When the average distance is greater than 200 bp,
-% dPeak identifies more than the 75\% of the binding events in each
-% peak, this is intuitive as the mean ChIP-Seq's fragment length is
-% shorter than 200 bp, hence the read does not contribute to more than
-% one binding event \cite{dpeak}.
-% In supplement figures \ref{fig:exoQC_sat1} to \ref{fig:exoQC_sat4}, we
-% repeat the same analysis for both replicates under the two rif
-% conditions.
-
-% are comparable if we use the init method with motif but not always
-%
 
@@
 SE ChIP-Seq. We partitioned the \emph{E.Coli} genome into
-non-overlaping bins of length 150, and for each bin we
-counted how many ChIP-exo and SE ChIP-Seq extended reads overlapped
-each bin. The reads were extended by 150 bp to their
-$3\prime$ directions.
+non-overlaping bins of length 150, and for each bin we counted how
+many ChIP-exo and SE ChIP-Seq extended reads overlapped each bin. The
+reads were extended by 150 bp to their $3\prime$ directions.
 
@@
 and considered an annotation as being identified if the distance to a
-estimated binding events is less or equal than 20 bp. We
-defined the resolution as the distance from an annotation to its
-closest predictions and the sensitivity as the fraction of identified
+estimated binding events is less or equal than 20 bp. We defined the
+resolution as the distance from an annotation to its closest
+predictions and the sensitivity as the fraction of identified
 annotations in a genomic region.
@@
 ChIP-exo data using the \textbf{MOSAiCS} algorithm \cite{mosaics} (one
-sample analysis using false discovery rate of 0.01\%)
-implemented as an R package \textbf{mosaics} version 2.9.7 (available
-from \emph{bioconductor}). We further filtered out candidate regions
-by using the 300 peaks with higher average ChIP tag count to
-avoid potential false positive based on the exploratory
-analysis. These regions were also explicitly provided to the GEM
-algorithm as candidate regions. Default tuning parameters were used
-during model fitting for all methods. We downloaded \textbf{CexoR}
-\cite{cexor} on its 1.8 version from \emph{bioconductor} but were
-unable to use it for the $\sig$ experiments.
+sample analysis using false discovery rate of 1\%) implemented as an R
+package \textbf{mosaics} version 2.9.7 (available from
+\emph{bioconductor}). We further filtered out candidate regions by
+using the 400 peaks with higher average ChIP tag count to avoid
+potential false positive based on the exploratory analysis. These
+regions were also explicitly provided to the GEM algorithm as
+candidate regions. Default tuning parameters were used during model
+fitting for all methods. We downloaded \textbf{CexoR} \cite{cexor} on
+its 1.8 version from \emph{bioconductor} but were unable to use it for
+the $\sig$ experiments.
 
@@
 using the dPeak's model; the number of identified targets are number
-of gold-standard annotations within 15 bp from an estimated
-binding events; and the resolution is defined as the minimum distance
-from a gold-standard annotation to an estimated binding event. We
-repeated this analysis for ten seeds and reported the median between
-all those values.
+of gold-standard annotations within 15 bp from an estimated binding
+events; and the resolution is defined as the minimum distance from a
+gold-standard annotation to an estimated binding event. We repeated
+this analysis for ten seeds and reported the median between all those
+values.
 
@@
   \includegraphics[width =
-  .95 \textwidth]{/p/keles/ChIPexo/volume3/ChIPexo/figs/for_paper/chip_explanation_withEnzyme.pdf}
+  .95 \textwidth]{figures/fig1/chip_explanation_withEnzyme.pdf}
   % \caption{\textbf{Description of ChIP-exo, SE ChIP-Seq and PE
@@
   \centering
-  \includegraphics[width = .46\textwidth,page = 3 ]{../figs/for_paper/ChIPseqPET_ChIPexo_tagCount_comparison.pdf}
-  \includegraphics[width = .46\textwidth]{../figs/for_paper/forward_strand_ratio_comp_old.pdf}
-\includegraphics[width = .46\textwidth,page = 1]{../figs/for_paper/eukaryotic_bias_CTCF.pdf}
-  \includegraphics[width = .46\textwidth,page = 2]{../figs/for_paper/eukaryotic_bias_CTCF.pdf}
+  \includegraphics[width = .46\textwidth,page = 3]{figures/fig2/ChIPseqPET_ChIPexo_tagCount_comparison.pdf}
+  \includegraphics[width =.46\textwidth]{figures/fig2/forward_strand_ratio_comp_old.pdf}
+  \includegraphics[width = .46\textwidth,page = 1]{figures/fig2/eukaryotic_bias_CTCF.pdf}
+  \includegraphics[width = .46\textwidth,page = 2]{figures/fig2/eukaryotic_bias_CTCF.pdf}
   \caption{ A) Hexbin plot of PE ChIP-Seq bin counts vs ChIP-exo bin
@@
 \centering
-\includegraphics[width = .8 \textwidth]{/p/keles/ChIPexo/volume3/ChIPexo/figs/for_paper/scc_ctcf2.pdf}
+\includegraphics[width = .8 \textwidth]{figures/fig3/scc_ctcf2.pdf}
 % \includegraphics[width = .8\textwidth]{/p/keles/ChIPexo/volume3/ChIPexo/figs/for_paper/EColi_strand_cross_corr.pdf}
@@
   \centering
-  \includegraphics[width = \textwidth]{../figs/for_paper/coverage_diagram3.pdf}
+  \includegraphics[width = \textwidth]{figures/fig4/coverage_diagram3.pdf}
   % \caption{\textbf{Flowchart of the ChIP-exo quality control
@@
   \centering
-  \includegraphics[width = .9  \textwidth,page = 1]{../figs/Carroll_mice_for_paper/FoxA1_enrichment.pdf}
+  \includegraphics[width = .9  \textwidth,page = 1]{figures/fig5/FoxA1_enrichment.pdf}
   \newline
-  \includegraphics[width = .23 \textwidth, page =
-  1]{../figs/for_paper/FoxA1_barplots.pdf}
-  \includegraphics[width = .23 \textwidth, page =
-  3]{../figs/for_paper/FoxA1_barplots.pdf}
-  \includegraphics[width =
-  .4\textwidth]{../figs/for_paper/FoxA1_profiles_around_motif.pdf}
-  \includegraphics[width = .3\textwidth,page =
-  2]{../figs/for_paper/FoxA1_matched_motif_sequence.pdf}
-  \includegraphics[width = .3\textwidth,page =
-  3]{../figs/for_paper/FoxA1_matched_motif_sequence.pdf}
-  \includegraphics[width = .3\textwidth,page =
-  1]{../figs/for_paper/FoxA1_matched_motif_sequence.pdf}
+  \includegraphics[width = .23 \textwidth, page = 1]{figures/fig5/FoxA1_barplots.pdf}
+  \includegraphics[width = .23 \textwidth, page = 3]{figures/fig5/FoxA1_barplots.pdf}
+  \includegraphics[width =.4\textwidth]{figures/fig5/FoxA1_profiles_around_motif.pdf}
+  \includegraphics[width = .3\textwidth,page = 2]{figures/fig5/FoxA1_matched_motif_sequence.pdf}
+  \includegraphics[width = .3\textwidth,page = 3]{figures/fig5/FoxA1_matched_motif_sequence.pdf}
+  \includegraphics[width = .3\textwidth,page = 1]{figures/fig5/FoxA1_matched_motif_sequence.pdf}
   \caption{Using the mouse FoxA1 experiment from \cite{exoillumina}:
@@
   \centering  
-  \includegraphics[width = .9\textwidth,page = 1]{../figs/Carroll_mice_for_paper/Strand_imbalance.pdf} 
+  \includegraphics[width = .9\textwidth,page = 1]{figures/fig6/Strand_imbalance.pdf} 
   \newline
-  \includegraphics[width = .9\textwidth,page = 2]{../figs/Carroll_mice_for_paper/Strand_imbalance.pdf} 
-  \includegraphics[width = .4\textwidth]{../figs/for_paper/Carroll_FSR_depth_VS_pvalWilcoxon.pdf}
+  \includegraphics[width = .9\textwidth,page = 2]{figures/fig6/Strand_imbalance.pdf} 
+  \includegraphics[width = .4\textwidth]{figures/fig6/Carroll_FSR_depth_VS_pvalWilcoxon.pdf}
   \caption{Strand imbalance using the mouse FoxA1 experiment from
@@
   \centering
-  \includegraphics[width = .9  \textwidth,page = 1]{../figs/for_paper/QC_pipeline_eval_boxplot.pdf}
+  \includegraphics[width = .9  \textwidth,page = 1]{figures/fig7/QC_pipeline_eval_boxplot.pdf}
 \newline
-  \includegraphics[width = .9  \textwidth,page = 2]{../figs/for_paper/QC_pipeline_eval_boxplot.pdf}
+  \includegraphics[width = .9  \textwidth,page = 2]{figures/fig7/QC_pipeline_eval_boxplot.pdf}
 \newline
-  \includegraphics[width = .45  \textwidth,page = 1]{../figs/for_paper/K562_TBP_sample.pdf}
-  \includegraphics[width = .45  \textwidth,page = 2]{../figs/for_paper/K562_TBP_sample.pdf}
+  \includegraphics[width = .45  \textwidth,page = 1]{figures/fig7/K562_TBP_sample.pdf}
+  \includegraphics[width = .45  \textwidth,page = 2]{figures/fig7/K562_TBP_sample.pdf}
   \caption{Comparison of adjusted Average Read Coverage and
@@
 
-  % \includegraphics[width = .9  \textwidth,page = 1]{../figs/test_quantify/arc_urcr_param_regression.pdf}
-  % \includegraphics[width = .9  \textwidth,page = 2]{../figs/test_quantify/arc_urcr_param_regression.pdf}
-  % \caption{A) Comparison of the contribution of the number of unique
-  %   positions to depth among all ChIP-exo and ChIP-nexus datasets
-  %   separated by organism. Comparison of the contributions of number
-  %   of unique positions (B) and width (C) to depth for sub-samples of
-  %   20M to 50M reads of TBP factor in human K562 cell lines ChIP-exo
-  %   and ChIP-nexus experiments.}
-
 \newpage
@@
   \centering
-  \includegraphics[width = .45\textwidth,page = 1]{../figs/for_paper/methods_comp_FDR5_topM400.pdf}
-  \includegraphics[width = .45\textwidth,page = 2]{../figs/for_paper/methods_comp_FDR5_topM400.pdf}
-   \includegraphics[width = .45\textwidth,page = 1]{../figs/for_paper/sigma_delta_old_densities.pdf}
-   \includegraphics[width = .45\textwidth,page = 2]{../figs/for_paper/sigma_delta_old_densities.pdf}   
+  \includegraphics[width = .45\textwidth,page = 1]{figures/fig8/methods_comp_FDR5_topM400.pdf}
+  \includegraphics[width = .45\textwidth,page = 2]{figures/fig8/methods_comp_FDR5_topM400.pdf}
+   \includegraphics[width = .45\textwidth,page = 1]{figures/fig8/sigma_delta_old_densities.pdf}
+   \includegraphics[width = .45\textwidth,page = 2]{figures/fig8/sigma_delta_old_densities.pdf}   
   \caption{Comparison of the resolution between dPeak, Gem, Mace and
@@
   \centering
-  \includegraphics[width = .9\textwidth]{../figs/for_paper/saturation_analysis_old.pdf}
+  \includegraphics[width = .9\textwidth]{figures/fig9/saturation_analysis_old.pdf}
   \caption{Comparison of the number of A) candidate regions, B)
@@
     considered as a gold standard. A gold standard binding events was
-    deemed identified if a binding event was estimated at a $\pm$
-    15 vicinity of it.}
+    deemed identified if a binding event was estimated at a $\pm$ 15
+    vicinity of it.}
   \label{fig:design}
 \end{figure}
-
-% \textbf{ChIP-exo, PE ChIP-Seq and SE ChIP-Seq comparison at varying
-%   sequencing depths.} 
-%  \newpage
-
-% \begin{figure}[h!]
-%   \centering
-% \includegraphics[width = .9\textwidth]{../figs/saturation/K12_alignment/sig70_aerobic_enrichment.pdf}
-% \caption{Hexbin plots of ARC vs URCR of the $\sig$ ChIP-exo experiment
-%   under aerobic condition when 10K to 90K reads are being sampled.}
-%   \label{fig:exoQC_sat_aero}
-% \end{figure}
 
@@
   \centering
-  \includegraphics[width = .46\textwidth]{../figs/for_paper/sensitivity_exo_olda_data.pdf}
-  \includegraphics[width = .46\textwidth]{../figs/for_paper/resolution_by_dataset_old_data.pdf}
+  \includegraphics[width = .46\textwidth]{figures/fig10/sensitivity_exo_olda_data.pdf}
+  \includegraphics[width = .46\textwidth]{figures/fig10/resolution_by_dataset_old_data.pdf}
   \caption{Comparison of A) sensitivity and B) resolution between
@@
 
-% \newpage
-
-% \beginsupplement
-
-% \section*{Supplement}
-% \label{sec:supp}
-
-% \emph{Note: I am not sure whether to include several figures on this supplement}
-
-%% \begin{table}[h!]
-%%  %% need to fix format
-%%   \centering
-
-% \caption{Usual ChIP-Seq QC metrics as in table \ref{tab:qc} but
-%   applied to Landick's chipseq data of the rif experiment}
-% \end{table}
-
-% \begin{figure}[h!]
-%   \centering
-%   \includegraphics[width = .7\textwidth]{../figs/saturation/K12_alignment/PBC_saturation.pdf}
-%   \caption{Average PBC (among all seeds) of the sampled ChIP-exo, PE
-%     ChIP-Seq and SE ChIP-Seq experiments under the rif treatment
-%     conditions used for saturation analysis.}
-%   \label{fig:pbc_saturation}
-% \end{figure}
-
-% \newpage
-
-% \begin{figure}[h!] %%% fix this plot as one ggplot, facet_grid(condition ~ variable,scales = "free_y")
-%   \centering
-%   \begin{itemize}
-%   \item Rep-1 and rif-0min:
-%   \end{itemize}
-%   \includegraphics[width = .3\textwidth,page =2]{../figs/saturation/K12_alignment/Sig70_rep1_rif0_saturation_analysis.pdf}
-%   \includegraphics[width = .3\textwidth,page =3]{../figs/saturation/K12_alignment/Sig70_rep1_rif0_saturation_analysis.pdf}
-%   \includegraphics[width = .3\textwidth,page =4]{../figs/saturation/K12_alignment/Sig70_rep1_rif0_saturation_analysis.pdf}
-
-%   \begin{itemize}
-%   \item Rep-1 and rif-20min:
-%   \end{itemize}
-
-%   \includegraphics[width = .3\textwidth,page =2]{../figs/saturation/K12_alignment/Sig70_rep1_rif20_saturation_analysis.pdf}
-%   \includegraphics[width = .3\textwidth,page =3]{../figs/saturation/K12_alignment/Sig70_rep1_rif20_saturation_analysis.pdf}
-%   \includegraphics[width = .3\textwidth,page =4]{../figs/saturation/K12_alignment/Sig70_rep1_rif20_saturation_analysis.pdf}
-
-%   \begin{itemize}
-%   \item Rep-2 and rif-0min:
-%   \end{itemize}
-
-%   \includegraphics[width = .3\textwidth,page =2]{../figs/saturation/K12_alignment/Sig70_rep2_rif0_saturation_analysis.pdf}
-%   \includegraphics[width = .3\textwidth,page =3]{../figs/saturation/K12_alignment/Sig70_rep2_rif0_saturation_analysis.pdf}
-%   \includegraphics[width = .3\textwidth,page =4]{../figs/saturation/K12_alignment/Sig70_rep2_rif0_saturation_analysis.pdf}
-
-%   \begin{itemize}
-%   \item Rep-2 and rif-20min:
-%   \end{itemize}
-
-%   \includegraphics[width = .3\textwidth,page =2]{../figs/saturation/K12_alignment/Sig70_rep2_rif20_saturation_analysis.pdf}
-%   \includegraphics[width = .3\textwidth,page =3]{../figs/saturation/K12_alignment/Sig70_rep2_rif20_saturation_analysis.pdf}
-%   \includegraphics[width = .3\textwidth,page =4]{../figs/saturation/K12_alignment/Sig70_rep2_rif20_saturation_analysis.pdf}
-%   \caption{Comparison of the number of predicted events (left),
-%     identified targets (middle) and resolution (right) among ChIP-exo,
-%     PE ChIP-Seq and SE ChIP-Seq. RegulonDB annotations are considered
-%     as gold standard. A RegulonDB binding events was deemed identified
-%     if a binding event was estimated at a $\pm$ ext vicinity
-%     of it.}
-%   \label{fig:saturation_rif}
-% \end{figure}
-
-% % \textbf{ChIP-exo, PE ChIP-Seq and SE ChIP-Seq comparison at varying
-% %   sequencing depth}
-
-% \newpage
-
-% \begin{figure}[h!]
-%   \centering
-%   \includegraphics[width = .9\textwidth,page = 1]{../figs/saturation/K12_alignment/sig70_rif_enrichment.pdf}
-%   \caption{ARC vs URCR hexbin plots of Rep-1 and rif-0min from $\sig$
-%     experiment when 100K to 900K reads are being sampled for each
-%     panel. }
-%   \label{fig:exoQC_sat1}
-% \end{figure}
-
-% \newpage
-
-% \begin{figure}[h!]
-%   \centering
-%   \includegraphics[width = .9\textwidth,page = 1]{../figs/saturation/K12_alignment/sig70_rif_enrichment.pdf}
-
-%   \caption{ARC vs URCR hexbin plots of Rep-1 and rif-20min from $\sig$ experiment when 100K to 900K reads are being
-%     sampled for each panel.}
-%   \label{fig:exoQC_sat2}
-% \end{figure}
-
-% \newpage
-
-% \begin{figure}[h!]
-%   \centering
-%   \includegraphics[width = .9\textwidth,page = 1]{../figs/saturation/K12_alignment/sig70_rif_enrichment.pdf}
-%   \caption{ARC vs URCR hexbin plots of Rep-2 and rif-0min from $\sig$
-%     experiment when 100K to 900K reads are being sampled for each
-%     panel.}
-%   \label{fig:exoQC_sat3}
-% \end{figure}
-
-% \newpage
-
-% \begin{figure}[h!]
-%   \centering
-%   \includegraphics[width = .9\textwidth,page = 1]{../figs/saturation/K12_alignment/sig70_rif_enrichment.pdf}
-%   \caption{ARC vs URCR hexbin plots of Rep-2 and rif-20min from $\sig$ experiment when 100K to 900K reads are being
-%     sampled for each panel.}
-%   \label{fig:exoQC_sat4}
-% \end{figure}
-
 \end{document}

### 2016-07-05 paper/supplement.tex
previous version of the paper

\documentclass{article}

\usepackage{authblk}
\usepackage{graphicx}
\usepackage{float}
\usepackage{tabu}
\usepackage{multirow}
\usepackage{xcolor}

\title{Supplement: Data exploration, quality control and statistical
  analysis of ChIP-exo experiments}
\date{}

\newcommand{\beginsupplement}{%
        \setcounter{table}{0}
        \renewcommand{\thetable}{S\arabic{table}}%
        \setcounter{figure}{0}
        \renewcommand{\thefigure}{S\arabic{figure}}%
     }

\author[1]{Rene Welch}
\author[6]{Dongjun Chung}
\author[3]{Irene Ong}
\author[3,4]{Jeffrey Grass}
\author[3,4,5]{Robert Landick}
\author[1,2]{S\"und\"uz Kele\c{s}}

\affil[1]{Department of Statistics, University of Wisconsin Madison}
\affil[2]{Department of Biostatistics and Medical Informatics,
  University of Wisconsin Madison}
\affil[3]{Great Lakes Bioenergy Research Center, University of Wisconsin Madison}
\affil[4]{Department of Biochemistry, University of Wisconsin Madison}
\affil[5]{Department of Bacteriology, University of Wisconsin Madison}
\affil[6]{Department of Public Health Sciences, Medical University of South Carolina}

\newcommand{\sig}{\sigma^{70}}

\newcommand{\pname}[1]{\texttt{ChIPexoQual}}
\newcommand{\SK}[1]{\textcolor{red}{SK: #1}}
\newcommand{\RW}[1]{\textcolor{blue}{RW: #1}}

\begin{document}

\maketitle

\beginsupplement

\tableofcontents

\newpage

\section{Comparison of ChIP-exo with ChIP-Seq data.}

\begin{figure}[H]
  \centering
  \includegraphics[width =
  .35\textwidth,page = 1]{figures/supplement/ChIPSeq_comp/sig70_rif_count_comparison.pdf}
  \includegraphics[width =
  .35\textwidth,page = 2]{figures/supplement/ChIPSeq_comp/sig70_rif_count_comparison.pdf}
  \includegraphics[width =
  .35\textwidth,page = 3]{figures/supplement/ChIPSeq_comp/sig70_rif_count_comparison.pdf}
  \includegraphics[width =
  .35\textwidth,page = 4]{figures/supplement/ChIPSeq_comp/sig70_rif_count_comparison.pdf}
  \includegraphics[width =
  .9\textwidth]{figures/supplement/ChIPSeq_comp/Sig70_strand_imbalance_rif.pdf}
  \caption{Hexbin plots of PE ChIP-Seq bin counts vs. ChIP-exo bin
    counts for $\sigma^{70}$ second biosample: A) E2-1, B) E2-2 C)
    E2-3 and D) E2-4 samples. E) Forward Strand Ratio densities for SE
    ChIP-Seq and ChIP-exo peaks for $\sigma^{70}$, S2 and E2 groups
    respectively.}
  \label{sfig:comp1}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width = .35\textwidth,page =
  1]{figures/supplement/ChIPSeq_comp/pugh_CTCF_human_bin_comp.pdf}
  \includegraphics[width = .35\textwidth,page =
  1]{figures/supplement/ChIPSeq_comp/CTCF_imbalance.pdf}
  \includegraphics[width = .35\textwidth,page =
  2]{figures/supplement/ChIPSeq_comp/pugh_CTCF_human_bin_comp.pdf}
  \includegraphics[width = .35\textwidth,page =
  2]{figures/supplement/ChIPSeq_comp/CTCF_imbalance.pdf}
  \caption{Hexbin plot comparing A) Rep-1 and B) Rep-2 ENCODE SE
    ChIP-seq bin counts vs. Pugh ChIP-exo bin counts for CTCF in HeLa
    cell. C) Rep-1 and D) Rep-2 ChIP-seq compared against ChIP-exo
    Forward Strand Ratio densities.}
  \label{sfig:comp2}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width =.35\textwidth,page =
  1]{figures/supplement/ChIPSeq_comp/carroll_ER_human_bin_comp.pdf}
  \includegraphics[width = .35\textwidth,page =
  1]{figures/supplement/ChIPSeq_comp/ER_imbalance.pdf}
  \includegraphics[width =.35\textwidth,page =
  2]{figures/supplement/ChIPSeq_comp/carroll_ER_human_bin_comp.pdf}
  \includegraphics[width = .35\textwidth,page =
  2]{figures/supplement/ChIPSeq_comp/ER_imbalance.pdf}
  \includegraphics[width =.35\textwidth,page =
  3]{figures/supplement/ChIPSeq_comp/carroll_ER_human_bin_comp.pdf}
  \includegraphics[width = .35\textwidth,page =
  3]{figures/supplement/ChIPSeq_comp/ER_imbalance.pdf}
  \caption{Hexbin plots comparing SE ChIP-Seq bin counts vs. ChIP-exo
    bin counts for ER in human MCF-7 cell lines. }
  \label{sfig:comp3}
\end{figure}

% \RW{ we also have data to add comparison of ChIP-exo vs ChIP-Seq TBP
%   in K562 cell lines, I am not eager to include those because 1) the
%   chip-exo data is of low quality, 2)}

\newpage

\section{SCC curves for ChIP-exo and ChIP-nexus data.}

\begin{figure}[H]
  \centering
  \includegraphics[width =
  .7\textwidth]{figures/supplement/SCC/Carroll_human_SCC.pdf}
  \caption{Comparison of ChIP-exo vs. ChIP-Seq SCC curves for ER factor in
    human MCF-7 cell lines. }
\label{sfig:scc1}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width =
  .5\textwidth]{figures/supplement/SCC/Carroll_mouse_SCC.pdf}
  \caption{SCC curves for FoxA1 factor in mouse liver cell lines.}
\label{sfig:scc2}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width =
  .5\textwidth]{figures/supplement/SCC/Meijsing_GR_SCC.pdf}
  \caption{SCC curves for GR factor in IMR90, K562 and U2OS human cell
    lines.}
  \label{sfig:scc3}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width =
  .75\textwidth]{figures/supplement/SCC/TBP_K562_SCC.pdf}
  \caption{Comparison of ChIP-exo vs. ChIP-nexus SCC curves for TBP
    factor in human K562 cell lines.}
  \label{sfig:scc4}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width =        
  .5\textwidth]{figures/supplement/SCC/Nexus_embryo_dorsal_SCC.pdf}
  \caption{SCC curves for dorsal factor in embryo
    \emph{D. Melanogaster} cell lines.}
  \label{sfig:scc5}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width =
  .5\textwidth]{figures/supplement/SCC/Nexus_embryo_twist_SCC.pdf}
  \caption{SCC curves for twist factor in embryo
    \emph{D. Melanogaster} cell lines.}
  \label{sfig:scc6}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width =
  .5\textwidth]{figures/supplement/SCC/Nexus_S2_Max_SCC.pdf}
  \caption{SCC curves for Max factor in S2 \emph{D. Melanogaster} cell
    lines.}
  \label{sfig:scc7}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width =
  .5\textwidth]{figures/supplement/SCC/Nexus_S2_MyC_SCC.pdf}
  \caption{SCC curves for MyC factor in S2 \emph{D. Melanogaster} cell
    lines.}
  \label{sfig:scc8}
\end{figure}

\RW{need to generate this curves for sig70 samples, compared against
  PE and SE ChIP-seq}

\newpage

\section{QC pipeline applied to ChIP-exo and ChIP-nexus data.}

\begin{figure}[H]
  \centering
  \includegraphics[width = .5\textwidth,page =
1]{figures/supplement/QC/Sig70_bios1_enrichment.pdf}\\
\includegraphics[width = .45\textwidth,page =
1]{figures/supplement/QC/Sig70_bios1_strand_imbalance.pdf}
\includegraphics[width = .45\textwidth,page =
3]{figures/supplement/QC/Sig70_bios1_strand_imbalance.pdf}
\caption{ChIP-exo QC pipeline diagnostics for $\sigma^{70}$ E1
  biosample in \emph{E. Coli}.}
  \label{sfig:qc1}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width = .5\textwidth,page =
1]{figures/supplement/QC/Sig70_bios2_enrichment.pdf}\\
\includegraphics[width = .45\textwidth,page =
1]{figures/supplement/QC/Sig70_bios2_strand_imbalance.pdf}
\includegraphics[width = .45\textwidth,page =
3]{figures/supplement/QC/Sig70_bios2_strand_imbalance.pdf}
\caption{ChIP-exo QC pipeline diagnostics for $\sigma^{70}$ E2
  biosample in \emph{E. Coli}.}
  \label{sfig:qc2}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width = .35\textwidth,page =
1]{figures/supplement/QC/Pugh_CTCF_hela_enrichment.pdf}\\
\includegraphics[width = .65\textwidth,page =
1]{figures/supplement/QC/Pugh_CTCF_hela_strand_imbalance.pdf}
\includegraphics[width = .65\textwidth,page =
3]{figures/supplement/QC/Pugh_CTCF_hela_strand_imbalance.pdf}
\caption{ChIP-exo QC pipeline diagnostics for CTCF factor in human
  HeLa cell lines.}
  \label{sfig:qc3}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width = .75\textwidth,page =
1]{figures/supplement/QC/Carroll_ER_MCF7_enrichment.pdf}\\
\includegraphics[width = .65\textwidth,page =
1]{figures/supplement/QC/Carroll_ER_MCF7_strand_imbalance.pdf}
\includegraphics[width = .65\textwidth,page =
3]{figures/supplement/QC/Carroll_ER_MCF7_strand_imbalance.pdf}
  \caption{ChIP-exo QC pipeline diagnostics for ER factor in human MCF-7 cell lines.}
  \label{sfig:qc4}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width = .75\textwidth,page =
1]{figures/supplement/QC/Venters_K562_TBP_enrichment.pdf}\\
\includegraphics[width = .65\textwidth,page =
1]{figures/supplement/QC/Venters_K562_TBP_strand_imbalance.pdf}
\includegraphics[width = .65\textwidth,page =
3]{figures/supplement/QC/Venters_K562_TBP_strand_imbalance.pdf}
\caption{ChIP-exo QC pipeline diagnostics for TBP factor in human K562
  cell lines.}
  \label{sfig:qc5}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width = .75\textwidth,page =
1]{figures/supplement/QC/Meijsing_GR_enrichment.pdf}\\
\includegraphics[width = .65\textwidth,page =
1]{figures/supplement/QC/Meijsing_GR_strand_imbalance.pdf}
\includegraphics[width = .65\textwidth,page =
3]{figures/supplement/QC/Meijsing_GR_strand_imbalance.pdf}
\caption{ChIP-exo QC pipeline diagnostics for GR factor in human
  IMR90, K562 and U2OS cell lines respectively.}
  \label{sfig:qc6}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width = .65\textwidth,page =
1]{figures/supplement/QC/ChIPnexus_embryo_dorsal_enrichment.pdf}\\
\includegraphics[width = .65\textwidth,page =
1]{figures/supplement/QC/ChIPnexus_embryo_dorsal_strand_imbalance.pdf}
\includegraphics[width = .65\textwidth,page =
3]{figures/supplement/QC/ChIPnexus_embryo_dorsal_strand_imbalance.pdf}
\caption{ChIP-exo QC pipeline diagnostics for ChIP-nexus experiment of
  dorsal factor in embryo \emph{D. Melanogaster} cell lines.}
  \label{sfig:qc7}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width = .65\textwidth,page =
1]{figures/supplement/QC/ChIPnexus_embryo_twist_enrichment.pdf}\\
\includegraphics[width = .65\textwidth,page =
1]{figures/supplement/QC/ChIPnexus_embryo_twist_strand_imbalance.pdf}
\includegraphics[width = .65\textwidth,page =
3]{figures/supplement/QC/ChIPnexus_embryo_twist_strand_imbalance.pdf}
\caption{ChIP-exo QC pipeline diagnostics for ChIP-nexus experiment of
  twist factor in embryo \emph{D. Melanogaster} cell lines.}
  \label{sfig:qc8}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width = .65\textwidth,page =
1]{figures/supplement/QC/ChIPnexus_S2_max_enrichment.pdf}\\
\includegraphics[width = .65\textwidth,page =
1]{figures/supplement/QC/ChIPnexus_S2_max_strand_imbalance.pdf}
\includegraphics[width = .65\textwidth,page =
3]{figures/supplement/QC/ChIPnexus_S2_max_strand_imbalance.pdf}
\caption{ChIP-exo QC pipeline diagnostics for ChIP-nexus experiment of
  Max factor in S2 \emph{D. Melanogaster} cell lines.}
  \label{sfig:qc9}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width = .65\textwidth,page =
1]{figures/supplement/QC/ChIPnexus_S2_myc_enrichment.pdf}\\
\includegraphics[width = .65\textwidth,page =
1]{figures/supplement/QC/ChIPnexus_S2_myc_strand_imbalance.pdf}
\includegraphics[width = .65\textwidth,page =
3]{figures/supplement/QC/ChIPnexus_S2_myc_strand_imbalance.pdf}
\caption{ChIP-exo QC pipeline diagnostics for ChIP-nexus experiment of
  MyC factor in S2 \emph{D. Melanogaster} cell lines.}
  \label{sfig:qc10}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width = .65\textwidth,page =
1]{figures/supplement/QC/ChIPnexus_K562_TBP_enrichment.pdf}\\
\includegraphics[width = .65\textwidth,page =
1]{figures/supplement/QC/ChIPnexus_K562_TBP_strand_imbalance.pdf}
\includegraphics[width = .65\textwidth,page =
3]{figures/supplement/QC/ChIPnexus_K562_TBP_strand_imbalance.pdf}
\caption{ChIP-exo QC pipeline diagnostics for ChIP-nexus experiment of
  TBP factor in K562 human cell lines.}
  \label{sfig:qc11}
\end{figure}

\newpage

\subsection*{Comparison of QC pipeline for different sub-sampled
  sequencing depths.}

\begin{figure}[H]
  \centering
  \includegraphics[width =.95\textheight,angle = 90,page =
1]{figures/supplement/QC_samp/Sig70_sample_bios1_enrichment.pdf} 
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width =.9\textheight,angle = 90,page =
1]{figures/supplement/QC_samp/Sig70_sample_bios1_strand_imbalance.pdf} 
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width =.9\textheight,angle = 90,page =
3]{figures/supplement/QC_samp/Sig70_sample_bios1_strand_imbalance.pdf} 
  \label{sfig:qcsamp1}
  \caption{ChIP-exo QC pipeline diagnostics for $\sigma^{70}$ E1
    biosample on \emph{E. Coli} when sampling 10K to 100K reads from
    each experiment}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width =.95\textheight,angle = 90,page =
1]{figures/supplement/QC_samp/Sig70_sample_bios2_enrichment.pdf} 
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width =.9\textheight,angle = 90,page =
1]{figures/supplement/QC_samp/Sig70_sample_bios2_strand_imbalance.pdf} 
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width =.9\textheight,angle = 90,page =
3]{figures/supplement/QC_samp/Sig70_sample_bios2_strand_imbalance.pdf} 
  \caption{ChIP-exo QC pipeline diagnostics for $\sigma^{70}$ E2
    biosample on \emph{E. Coli} when sampling 10K to 100K reads from
    each experiment}
  \label{sfig:qcsamp2}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[height = .4\textheight,page =
  1]{figures/supplement/QC_samp/TBP_sample_enrichment.pdf}
  \includegraphics[height = .25\textheight,page =
1]{figures/supplement/QC_samp/TBP_sample_strand_imbalance.pdf} 
  \includegraphics[height = .25\textheight,page =
3]{figures/supplement/QC_samp/TBP_sample_strand_imbalance.pdf} 
\caption{ChIP-exo QC pipeline diagnostics applied to ChIP-nexus and
  ChIP-exo experiments of TBP factor in K562 human cell lines when
  sampling 20M to 50M reads from each experiment.}
  \label{sfig:qcsamp3}
\end{figure}

\newpage

\subsection{FoxA1 peak overlaps with high quality regions.}

\begin{figure}[H]
\begin{center}
  \includegraphics[width=.45\textwidth,page =
  1]{figures/supplement/peak_overlaps/FoxA1_VennDiagram_QC_region_w_motif_and_peaks.pdf} \\
\includegraphics[width=.45\textwidth,page =
2]{figures/supplement/peak_overlaps/FoxA1_VennDiagram_QC_region_w_motif_and_peaks.pdf}\\
\includegraphics[width=.45\textwidth,page =
3]{figures/supplement/peak_overlaps/FoxA1_VennDiagram_QC_region_w_motif_and_peaks.pdf}\\
\end{center}
\caption{Venn diagrams of high quality regions that overlap peaks for
  FoxA1 factor in mouse liver cell lines. Top) Rep-1, Middle) Rep-2
  and Bottom) Rep-3.}
\label{sfig:venn}
\end{figure}

\newpage

% \subsection{Evaluation of sub-sampled experiment with QC pipeline.}

% \newpage

% \section{Comparison with other methods rif samples.}

% \newpage

% \section{Saturation analysis rif samples.} %% need to re -do

% \newpage

% \section{Resolution and sensitivity.}

\end{document}

### 2016-07-05 rscripts/FoxA1_profiles_around_motif.R
modified scripts

@@ -380,6 +380,33 @@ ggplot(venn2,aes(rep,prop,fill = rep))+geom_bar(stat = "identity")+
 dev.off()
 
 
+fimo2 <- fimo
+names(fimo2) <- c("Rep-3","Rep-1","Rep-2")
 
 
+topK <- c(50,250,100,500,1000,2000)
 
+filter_topK <- function(fimo,K){
+  fimo[order(-score)][1:K]
+}
+
+fimo2 <- mapply(function(x,y)x[,repl := y],fimo2,
+                names(fimo2),SIMPLIFY = FALSE)
+
+dt_list <- lapply(topK,
+  function(k,fimo2){
+    out <- lapply(fimo2,filter_topK,k)
+    out <- do.call(rbind,out)
+    out[,K := k]
+    return(out)},fimo2)
+dt <- do.call(rbind,dt_list)
+
+pdf(file = "figs/FoxA1_fimo_score.pdf",width =12 ,height = 5)
+ggplot(dt,aes(repl,score,fill = repl))+
+  geom_boxplot(outlier.size = NA)+
+  facet_grid( . ~ K)+theme_bw()+
+  theme(legend.position = "none",
+        axis.text.x = element_text(angle = 75,hjust = 1))+
+  scale_fill_brewer(palette = "Pastel1")+
+  xlab("Replicate")+ylab("FIMO score")
+dev.off()

### 2016-07-05 rscripts/Strand_cross_corr_comp_CTCF.R

@@ -12,9 +12,9 @@ hg19 <- system.file("extdata","chrom.sizes","hg19.chrom.sizes",
 sizes <- data.table(read.table(hg19))
 shift <- 1:200
 
-exo_file <- "/p/keles/ChIPexo/volume4/pugh_data/CTCF.bam"
-set_file1 <- "/p/keles/ChIPexo/volume4/pugh_data/encode-Uw-Helas3-Ctcf-rep1.bam"
-set_file2 <- "/p/keles/ChIPexo/volume4/pugh_data/encode-Uw-Helas3-Ctcf-rep2.bam"
+exo_file <- "/p/keles/ChIPexo/volume4/pugh_data/raw_data/CTCF_hg18.sort.bam"
+set_file1 <- "/p/keles/ChIPexo/volume4/pugh_data/encode/encode-Uw-Helas3-Ctcf-rep1.bam"
+set_file2 <- "/p/keles/ChIPexo/volume4/pugh_data/encode/encode-Uw-Helas3-Ctcf-rep2.bam"
 
 exo <- create_reads(exo_file)
 set1 <- create_reads(set_file1)
@@ -24,7 +24,7 @@ set2 <- create_reads(set_file2)
 scc_exo <- strand_cross_corr(exo,chrom.sizes = sizes,parallel = TRUE,shift = 1:300)
 scc_set1 <- strand_cross_corr(set1,chrom.sizes = sizes,parallel = TRUE,shift = 1:300) 
 scc_set2 <- strand_cross_corr(set2,chrom.sizes = sizes,parallel = TRUE,shift = 1:300) 
-
+ 
 scc_exo[,seq := "ChIP-exo"]
 scc_set1[,seq := "ChIP-Seq (SE)"]
 scc_set2[,seq := "ChIP-Seq (SE)"]
@@ -34,25 +34,37 @@ dt <- rbind(scc_exo[, repl := 1],
             scc_set1[,repl := 1],
             scc_set2[,repl := 2])
 
+r1 <- brewer.pal(name = "Set1",5)
+r2 <- brewer.pal(name = "Dark2",5)
+r <- c(r2[c(1,3:5)],r1[2:3])
+
+r <- brewer.pal(name = "Set1",3)
+
 pdf(file ="figs/for_paper/scc_ctcf.pdf",width = 8,height = 5)
 ggplot(dt,
-  aes(shift , cross.corr,colour = seq,linetype = as.factor(repl)))+
[…]

### 2016-07-05 rscripts/TBP_generate_sequence.R

@@ -8,33 +8,38 @@ library(parallel)
 
 load_all("~/Desktop/Docs/Code/ChIPexoQual")
 
-data_dir1 <- "/p/keles/ChIPexo/volume4/venters_data/sortbam"
-files <- list.files(data_dir1,
-       pattern = "sort.bam",full.names = TRUE,include.dirs = TRUE)
+data_dir <- "data/ChIPexo_QC_runs"
+files <- list.files(data_dir,full.names = TRUE)
 
+files <- files[grep("K562",files)]
+files <- files[grep("TBP",files)]
 files <- files[grep("bai",files,invert = TRUE)]
 files <- files[grep("TBP",files)]
+files <- files[-7]
 
-exo <- lapply(files,create_exo_experiment,parallel = TRUE)
+load_file <- function(x){
+  load(x)
+  return(ext_stats[["stats"]])
+}
 
-exo_reads <- lapply(exo,reads)
-rl_summary <- lapply(exo_reads,function(x)table(width(x)))
-rl <- 45
+files <- files[grep("samp",files,invert = TRUE)]
 
+stats <- mclapply(files,load_file,mc.cores = 22)
+names(stats) <- gsub(".RData","",basename(files))
 
-stats <- lapply(exo,summary_stats)
-### filter regions
+rl <- 50
 
 ## 1 - wider regions
-stats <- lapply(stats,function(x)x[between(width,3 * rl,1e3)])
+stats <- lapply(stats,function(x)x[between(width,rl,2e3)])
 
-## most of the regions 
-## > stats <- lapply(stats,function(x)x[between(width,2 * rl,1e3)])
-## > sapply(stats,nrow)
-## [1] 32326  7447  5500
-## > stats <- lapply(stats,function(x)x[between(width,3 * rl,1e3)])
-## > sapply(stats,nrow)
-## [1] 2146 2578 1832
+## 2 - removed single - stranded regions
+stats <- lapply(stats,function(x)x[f > 0 & r > 0])
+
+## 3 - deeper regions
+stats <- lapply(stats,function(x){
+  mm <- x[,median(depth)]
+  return(x[depth > mm])})
+
 
 regions <- lapply(stats,function(x)
                   ChIPUtils::dt2gr(x[,2:4,with = FALSE]))
@@ -43,66 +48,9 @@ regions <- lapply(regions,sort)
 library(BSgenome.Hsapiens.UCSC.hg19)
 
 sequences <- mclapply(regions,function(x)
[…]

### 2016-07-05 rscripts/carroll_mice_strand_imbalance.R

@@ -9,102 +9,17 @@ library(GenomicAlignments)
 library(parallel)
 library(viridis)
 
-dr <- "/p/keles/ChIPexo/volume3/CarrollData/mouse"
-files <- list.files(dr)
+dr <- "data/ChIPexo_QC_runs"
+files <- list.files(dr,full.names = TRUE)
+files <- files[grepl("carr",files) & grepl("mouse",files)]
 
-files <- files[grep("bam",files)]
-files <- files[grep("bai",files,invert = TRUE)]
+load_stat <- function(x){
+  load(x)
+  ext_stats$stats}
 
-files <- file.path(dr,files)
-
-reads <- mclapply(files,create_reads,mc.cores = 3)
-names(reads) <- basename(files)
-
-
-create_regions <- function(reads,lower,rangesOnly=TRUE)
-{
-  stopifnot(class(reads) == "GRanges")
-  stopifnot(lower > 0)
-  sreads <- split(reads,seqnames(reads))
-  sreads <- as.list(sreads)
-  cover <- mcmapply(function(x,y)coverage(x)[[y]],sreads,names(sreads),mc.cores = detectCores())
-  islands <- lapply(cover,slice,lower = lower, rangesOnly = rangesOnly)
-  islands <- mapply(function(x,y)GRanges(seqnames = y,ranges = x),islands,names(islands),SIMPLIFY =FALSE)
-  islands <- do.call(rbind,lapply(islands,gr2dt))
-  return(islands)
-}
-
-gr <- lapply(reads,function(x){
-  byChr <- mapply(rbind,readsF(x),readsR(x),SIMPLIFY =FALSE)
-  out <- do.call(rbind,byChr)
-  return(dt2gr(out))
-})
-
-regs <- lapply(gr,create_regions,lower  = 1)
-
-build_stats <- function(region,reads)
-{
-  ## fix formats and stuff
-  region <- dt2gr(region)
-  
-  ov <- findOverlaps(region,reads)
-  reads <- gr2dt(reads)
-  w <- width(region)    
-  region <- gr2dt(region)
-  region[ , width := w]
-  region[, match := paste0(seqnames,":",start,"_",end)]
-  reads[  subjectHits(ov), match := region[queryHits(ov), (match)] ]  
[…]

### 2016-07-05 rscripts/local_NSC_example_Rep1.R

@@ -11,11 +11,12 @@ library(viridis)
 library(grid)
 library(gridExtra)
 
-dr <- "/p/keles/ChIPexo/volume3/CarrollData/mouse"
+dr <- "/p/keles/ChIPexo/volume4/carroll_data/mouse"
 files <- list.files(dr)
 
 files <- files[grep("bam",files)]
 files <- files[grep("bai",files,invert = TRUE)]
+files <- files[grep("sort",files)]
 
 files <- file.path(dr,files)
 
@@ -130,8 +131,8 @@ p <- ggplot(stats , aes( ave_reads,cover_rate))+stat_binhex(bins = 50)+
   scale_fill_gradientn(colours = r,trans = "log10",
     labels = trans_format("log10",math_format(10^.x)))+xlim(0,4)+ylim(0,1)+
   theme_bw()+theme(legend.position = "top",plot.title = element_text(hjust = 0))+
-  xlab("Average read coverage (ARC)")+
-  ylab("Unique read coverage rate (URCR)")
+  xlab("Average Read Coefficient (ARC)")+
+  ylab("Unique Read Coefficient (URC)")
 print(p)
 print( p  %+% stats[ npos > 10])  ## npos > 10 
 print( p  %+% stats[ npos > 30]) ## npos > 30
@@ -203,12 +204,12 @@ make_plots <- function(arc,urcr,p,regs,reads,data)
   plots[[2]] <- p +
     geom_point(data = point,aes(x = ave_reads,y = cover_rate),fill = "orange",colour = "orange", size = 4,shape =16)
 
-  plots[[3]] <- ggplot(loc_cc,aes(shift,cross.corr))+
-    geom_point(size = 2,shape = 1)+
-    geom_line(linetype = 2)+
-    theme_bw()+theme(plot.title = element_text(hjust = 0))+ylab("local Strand Cross Correlation")+
-    geom_smooth(method = "loess",se = FALSE,size = 2)+
-    geom_abline(slope = 0,intercept = 0,linetype = 3,size = 1,colour = "grey")
+  ## plots[[3]] <- ggplot(loc_cc,aes(shift,cross.corr))+
+  ##   geom_point(size = 2,shape = 1)+
+  ##   geom_line(linetype = 2)+
+  ##   theme_bw()+theme(plot.title = element_text(hjust = 0))+ylab("local Strand Cross Correlation")+
[…]

### 2016-07-05 rscripts/paper_analysis/QC_pipeline_eval.R

@@ -127,8 +127,12 @@ DT <- coeffs[grepl("K562",name)][!grepl("mei",name)]
 
 
 DT <- merge(DT,meta,by = "name") %>%
-  filter(name != "chipnexus_K562_TBP_Rep1") %>%
-  filter(!grepl("venters_TBP_K562_Rep1",name))
+    filter(name != "chipnexus_K562_TBP_Rep1")
+
+
+  ## filter(!grepl("venters_TBP_K562_Rep1",name))
+
+
 
 figs_dir <- "figs/for_paper"
 
@@ -194,6 +198,8 @@ dev.off()
 
 
 
+
+
 pdf(file = file.path(figs_dir,"K562_TBP_sample.pdf"),width = 6,height = 4)
 p1 <- ggplot(DT[term == "npos"],aes(samp,estimate,
                   colour = paste0(protocol,repl)))+
@@ -205,7 +211,7 @@ p1 <- ggplot(DT[term == "npos"],aes(samp,estimate,
                    axis.title.x = element_blank(),
                    plot.title = element_text(hjust = 0),
                    strip.background = element_blank())+
-  scale_color_manual(values = r[c(4,4,7)])+ggtitle("C")+  
+  scale_color_manual(values = r[c(4,4,4,7)])+ggtitle("C")+  
   geom_abline(slope = 0 ,intercept = 1,linetype = 2 ,colour = "darkgrey")+
   ylab("Adjusted Average Read Coverage")
 p2 <- ggplot(DT[term == "width"],aes(samp,-estimate,
@@ -218,9 +224,151 @@ p2 <- ggplot(DT[term == "width"],aes(samp,-estimate,
                    plot.title = element_text(hjust = 0),
                    strip.background = element_blank())+
   ylim(-1,5)+ylab("Average Read Coverage Bias")+
-  scale_color_manual(values = r[c(4,4,7)])+ggtitle("D")+
+  scale_color_manual(values = r[c(4,4,4,7)])+ggtitle("D")+
   geom_abline(slope = 0 ,intercept = 0,linetype = 2 ,colour = "darkgrey")
 p1
 p2
 dev.off()
 
+DT <- DT %>% filter(samp != "complete")
+
+
+DT <- DT %>% select(term,estimate,protocol,repl,samp,lab)
+coeff <- group_by(DT,term,protocol,repl,samp)
+out <- summarize(coeff,
[…]

### 2016-07-05 rscripts/paper_analysis/methods_comparison_binding_events.R

@@ -287,3 +287,32 @@ reso <- mapply(function(x,y)copy(x)[,dataset := y],reso,edsn,SIMPLIFY = FALSE)
 reso <- do.call(rbind,reso)
 
 save(reso,file = "data/Resolution_method_with_peaks.RData")
+
+pdf(file = "figs/gem_dpeak_reso/nr_sites.pdf",height = 12,width = 4)
+ggplot(melt(dcast.data.table(reso[dataset == 931], value.var = "reso",
+     formula = peak + dataset ~ method,
+     fun.aggregate = length)
+),aes( variable,peak , fill = factor(value)))+geom_tile()+
+  theme_bw()+
+  theme(legend.position = "top",axis.text.x = element_text(angle = 90))+
+  scale_fill_brewer(palette = "Pastel1",name = "nsites")
+dev.off()
+
+reso2 <- dcast.data.table(reso[dataset == 931], value.var = "reso",
+     formula = peak + dataset ~ method,
+     fun.aggregate = length)
+aux <- strsplit(reso2[,(peak)],"-")
+dt <- data.table(sq = "U00096",
+   start = sapply(aux,function(x)as.numeric(x[1])),
+   end = sapply(aux,function(x)as.numeric(x[2])))
+
+dt[start == 0, start := 1]
+
+write.table(dt,
+ file = "/p/keles/ChIPexo/volume6/K12/other_methods/dpeak_test/peak_to_check_dpeak.txt",
+ quote = FALSE,row.names = FALSE,col.names = FALSE,sep = "\t")
+
+u <- dt[, paste0(sq,":",start,"-",end)]
+write.table(u,
+ file = "/p/keles/ChIPexo/volume6/K12/other_methods/dpeak_test/peak_to_check_gem.txt",
+ quote = FALSE,row.names = FALSE,col.names = FALSE,sep = "\t")

### 2016-07-05 rscripts/paper_analysis/sig70_dpeak_motif_binding.R

@@ -19,7 +19,7 @@ read_files <- do.call(c,lapply(idx,function(x)read_files[x]))
 read_files <- read_files[grep("sort",read_files)]
 read_files <- read_files[grep("bai",read_files,invert = TRUE)]
 
-fdr <- "FDR1/"
+fdr <- "FDR5/"
 peak_dir <- "/p/keles/ChIPexo/volume6/K12/downstream/ChIPexo"
 peak_files <- list.files(peak_dir,recursive = TRUE)
 idx <- lapply(edsn,grep,peak_files)
@@ -33,7 +33,7 @@ peaks <- lapply(peaks,function(x)x[,V1 := "NC_010473"])
 
 mc <- detectCores()
 motif_file <- "/p/keles/ChIPexo/volume6/K12/motif_dpeak/Sig70_motifs.RData"
-motif_file_par <- "/p/keles/ChIPexo/volume6/K12/motif_dpeak/Sig70_motifs_par.RData"
+motif_file_par <- "/p/keles/ChIPexo/volume6/K12/motif_dpeak/Sig70_motifs_FDR5.RData"
 
 load(motif_file_par) ## dp_motifs
 
@@ -58,7 +58,7 @@ dpeakEst <- mapply(dpeak_call,file.path(peak_dir,peak_files),
 
 save(dpeakEst,file = "dpeak_sites_motif.RData")
 
-out_dir <- "/p/keles/ChIPexo/volume6/K12/downstream/ChIPexo/motif_init/FDR1"
+out_dir <- "/p/keles/ChIPexo/volume6/K12/downstream/ChIPexo/motif_init/FDR5"
 dir.create(out_dir,recursive = TRUE)
 
 out_files <- gsub("_peaks.txt","_sites.txt",basename(peak_files))

### 2016-07-05 rscripts/paper_analysis/sig70_dpeak_motif_parallel.R

@@ -16,7 +16,7 @@ mc <- 22
 ## read_files <- read_files[grep("sort",read_files)]
 ## read_files <- read_files[grep("bai",read_files,invert = TRUE)]
 
-fdr <- "FDR1/"
+fdr <- "FDR5/"
 peak_dir <- "/p/keles/ChIPexo/volume6/K12/downstream/ChIPexo"
 peak_files <- list.files(peak_dir,recursive = TRUE)
 idx <- lapply(edsn,grep,peak_files)
@@ -46,4 +46,4 @@ dp_motifs <- lapply(peak_tmp,dpeakMotif,refGenome=Ecoli,memeArgument = memeArg,f
 
 names(dp_motifs) <- basename(peak_files)
 
-save(dp_motifs,file = file.path(motif_dir,"Sig70_motifs_par.RData"))
+save(dp_motifs,file = file.path(motif_dir,"Sig70_motifs_FDR5.RData"))

### 2016-07-05 rscripts/peak_pair_assumption_new.R

@@ -3,7 +3,7 @@ rm(list = ls())
 
 library(mosaics)
 library(ChIPUtils)
-
+library(data.table)
 library(dpeak)
 
 exo_file <- "/p/keles/ChIPexo/volume3/LandickData/ChIPexo/edsn931_042814_qc.sorted.bam"

### 2016-07-05 rscripts/peak_pair_assumption_old_data.R

@@ -207,7 +207,9 @@ balance1 <- sapply( fit.exo1@fragSet, function(x)
 	} else {
 		return( NA )
 	} )
-balance1 <- balance1[ peak.R1[,5] > 3000 ]
+nsites1 <- sapply(fit.exo1@optMu,length)
+
+balance1 <- balance1[ peak.R1[,5] > 3000  ]
 
 balance2 <- sapply( fit.exo2@fragSet, function(x) 
 	if ( !is.na(x[1,1]) ) {
@@ -234,12 +236,16 @@ dt[ ,seq := plyr::mapvalues(seq , from = c("exo","set"),
        to = c("ChIP-exo","ChIP-Seq (SE)"))]
 
 pdf(file = "figs/for_paper/forward_strand_ratio_comp_old.pdf",width = 4 ,height  = 4)
-ggplot(dt , aes(fratio, colour = seq))+geom_density()+
+ggplot(dt , aes(fratio, colour = seq))+stat_density(geom = "line")+
   theme_bw()+scale_color_manual(values = brewer.pal(3,"Set1")[c(1,3)],name = "")+
   theme(legend.position = "top",plot.title = element_text(hjust = 0))+
   xlab("Ratio of forward strand reads")+
   ylab("Density")+geom_vline(xintercept = .5,linetype = 2)+ggtitle("B")
 dev.off()
+
+
+
+
 
 
 
@@ -368,3 +374,56 @@ plot( density(sigma.SET), xlim=c(0,100),
 	xlab="sigma estimate", main="ChIP-seq (SET)",
 	cex.axis=1.5, cex.lab=1.5, cex.main=1.5 )
 dev.off()
+
+
+
+# strand balance 1 site
+
+
+balance1 <- sapply( fit.exo1@fragSet, function(x) 
+	if ( !is.na(x[1,1]) ) {
+		return( mean( x[,3]=="F" ) )
+	} else {
+		return( NA )
+	} )
+nsites1 <- sapply(fit.exo1@optMu,length)
+
+balance1 <- balance1[ peak.R1[,5] > 3000 & nsites1  == 1 ]
+
+balance2 <- sapply( fit.exo2@fragSet, function(x) 
+	if ( !is.na(x[1,1]) ) {
+		return( mean( x[,3]=="F" ) )
+	} else {
+		return( NA )
+	} )
+nsites2 <- sapply(fit.exo2@optMu,length)
[…]

### 2016-07-05 rscripts/scripts/npos_width_vs_depth_regression.R

@@ -39,6 +39,8 @@ if("--help" %in% args){
 
 stopifnot(length(args) == 4)
 
+u1 <- proc.time()
+
 library(GenomicAlignments)
 library(parallel)
 library(devtools)
@@ -47,14 +49,25 @@ library(broom)
 
 load_all("~/Desktop/Docs/Code/ChIPexoQual")
 
+u2 <- proc.time()
+
 infile <- args[1]
 outfile <- args[2]
 M <- as.numeric(args[3])
 N <- as.numeric(args[4])
 
+timefile <- file.path("data/times",
+                      gsub(".RData","_time.txt",basename(infile)))
+M <- 1e3
+N <- 1e4
+
+
+
 mc <- detectCores()
 
 load(infile)
+
+u3 <- proc.time()
 
 stats <- ext_stats[[1]]
 
@@ -69,5 +82,13 @@ samp <- function(stats,nreg)
 
 coeff <- mclapply(1:N,function(i)samp(stats,M),mc.cores = mc)
 coeff <- do.call(rbind,coeff)
+
+u4 <- proc.time()
+
+u <- list(u2 - u1,u3 - u2,u4 - u3)
+
+write.table(Reduce(rbind,u),file = timefile,quote = FALSE,
+            append = TRUE,col.names = FALSE,row.names = FALSE,
+            sep = "\t")
 
 save(coeff,file = outfile)

### 2016-07-05 rscripts/test_ideas/dpeak_evaluation_SET.R

@@ -11,8 +11,10 @@ files <- list.files(dr,recursive = TRUE)
 files <- files[grep("rif_treatment",files)]
 files <- files[grep("sort",files)]
 files <- files[grep("bai",files,invert = TRUE)]
-files <- files[grep("PET",files,invert = TRUE)] ## only keep SET
+## files <- files[grep("PET",files,invert = TRUE)] ## only keep SET
 files <- files[grep("Input",files,invert = TRUE)] ## only chip
+files <- files[c(4,8,12)]
+
 
 annot_dir <- "/p/keles/ChIPexo/volume6/K12/annotations"
 afiles <- list.files(annot_dir)
@@ -25,127 +27,268 @@ annots <- reduce(c(annots[[1]],annots[[2]]))
 width(annots) <- 1
 annots <- sort(annots)
 
-size <- 300
+size <- 250
 anchors <- annots[which(diff(mid(annots)) > size)]
 center <- mid(anchors)
-start(anchors) <- center - size * 2
-end(anchors) <- center + size * 2
+start(anchors) <- center -2* size 
+end(anchors) <- center + 2*size 
 start(anchors[1]) <- 1
 anchors <- anchors[countOverlaps(anchors,annots) > 1]
+anchors <- reduce(anchors)
 
-dt <- ChIPUtils::gr2dt(GRanges(seqnames = "U00096",anchors))
+peaks <- ChIPUtils::gr2dt(GRanges(seqnames = "U00096",anchors))
 
 temp_peakfile <- tempfile(pattern = "peak",fileext = ".txt")
 
-write.table(dt,file = temp_peakfile,row.names = FALSE,col.names = FALSE,quote = FALSE)
+write.table(peaks,file = temp_peakfile,row.names = FALSE,col.names = FALSE,quote = FALSE)
 
-## reads <- lapply(reads,function(x){
-##   x <- resize(x,1)
-##   return(x)})
+## reads chunk
+load_reads <- function(file,PET)
+{
+  if(PET){
+    out <- readGAlignmentPairs(file,param = NULL)
+    out1 <- left(out)
+    out2 <- right(out)
+    sqnms <- seqnames(out1)
+    st <- start(out1)
+    en <- end(out2)
+    out <- GRanges(seqnames =sqnms,
+            ranges =  IRanges(start = st , end = en),strand = strand(out1))
+  }else{
[…]

### 2016-07-05 scripts/tbp_hg19_exo_nexus.sh

@@ -1,32 +1,27 @@
 #1/bin/sh
 
 basedir=/p/keles/ChIPexo/volume4/tbp_analysis
-seqexodir=$basedir/sequences/chipexo
-seqnexusdir=$basedir/sequences/chipnexus
+seqdir=$basedir/sequences
 
 motifdr=/p/keles/MEME/motif_databases
 db=JASPAR_CORE_2009.meme
 motifdb=$motifdr/$db
 
-fimoexodir=$basedir/fimo/chipexo
-fimonexusdir=$basedir/fimo/chipnexus
+seq1=$seqdir/venters_TBP_K562_Rep1sequences.s
+seq2=$seqdir/venters_TBP_K562_Rep2sequences.s
+seq3=$seqdir/venters_TBP_K562_Rep3sequences.s
+seq4=$seqdir/chipnexus_K562_TBP_Rep1sequences.s
+seq5=$seqdir/chipnexus_K562_TBP_Rep2sequences.s
+
+fimodir=$basedir/fimo
 
 ## motif1
-
 motif=MA0108.1
 
-fimo -oc $fimoexodir/Rep1_TBP1 --motif $motif $motifdb $seqexodir/TBP_K562_Rep1.sequences.fna
-fimo -oc $fimoexodir/Rep2_TBP1 --motif $motif $motifdb $seqexodir/TBP_K562_Rep2.sequences.fna
-fimo -oc $fimoexodir/Rep3_TBP1 --motif $motif $motifdb $seqexodir/TBP_K562_Rep3.sequences.fna
-fimo -oc $fimonexusdir/Rep1_TBP1 --motif $motif $motifdb $seqnexusdir/ChIPnexus_K562_TBP_rep1.sequences.fna
-fimo -oc $fimonexusdir/Rep2_TBP1 --motif $motif $motifdb $seqnexusdir/ChIPnexus_K562_TBP_rep2.sequences.fna
+fimo -oc $fimodir/venters_Rep1_jaspar --thresh 1e-2 --motif $motif $motifdb $seq1 &
+fimo -oc $fimodir/venters_Rep2_jaspar --thresh 1e-2 --motif $motif $motifdb $seq2 &
+fimo -oc $fimodir/venters_Rep3_jaspar --thresh 1e-2 --motif $motif $motifdb $seq3 &
[…]

### 2016-07-06 paper/chipexo_QC_vs2.tex
corrected first part of comments, until half of results section

@@
 \begin{fmbox}
-\dochead{Draft}
+\dochead{Research}
 
@@
 ]{\inits{DC}\fnm{Dongjun} \snm{Chung} } 
-\author[
-   addressref={aff3},
-   email={ong@cs.wisc.edu}
-]{\inits{IO}\fnm{Irene} \snm{Ong}}
 \author[
@@
 \begin{abstract}
-    ChIP-exo/nexus experiments present modifications on the commonly
-    used ChIP-seq protocol for high resolution mapping of
-    transcription factor binding sites. Although many aspects of the
-    ChIP-exo data analysis are similar to those of ChIP-seq, these
-    high throughput experiments present a number of unique quality
-    control and analysis challenges. We develop a statistical quality
-    control pipeline and accompanying \texttt{R} package, \pname{}, to
-    enable exploration and analysis of ChIP-exo and related
-    experiments. \pname{} evaluates a number of key issues including
-    strand imbalance, library complexity, and signal enrichment of
-    data from ChIP-exo/nexus experiments. Assessment of these
-    properties are facilitated through diagnostic plots and summary
-    statistics calculated over regions of the genome with varying
-    levels of coverage.
+ChIP-exo/nexus experiments present modifications on the commonly used ChIP-seq protocol for high resolution mapping of transcription factor binding sites.
+Although many aspects of the ChIP-exo data analysis are similar to those of ChIP-seq, these high throughput experiments pose a number of unique quality control and analysis challenges.
+We develop a statistical quality control pipeline and accompanying \texttt{R} package, \pname{}, to enable exploration and analysis of ChIP-exo and related experiments.
+\pname{} evaluates a number of key issues including strand imbalance, library complexity, and signal enrichment of data.
+Assessment of these features are facilitated through diagnostic plots and summary statistics calculated over regions of the genome with varying levels of coverage.
+
+We evaluated our QC pipeline with both large collections of public ChIP-exo/nexus data and multiple, new ChIP-exo datasets from \textit{E.
+Coli}.
+\pname{} analysis of these datasets resulted in guidelines for using these QC metrics across a wide range of sequencing depths and further insights for modelling ChIP-exo data.Finally, although ChIP-exo experiments have been compared to ChIP-seq experiments with single-end (SE) sequencing, we provide, for the first time, comparisons with paired-end (PE) ChIP-seq experiments.
+We illustrate that, at fixed sequencing depths, ChIP-exo provides higher sensitivity, specificity, and spatial resolution than PE ChIP-seq and both significantly outperform their SE ChIP-seq counterpart.
+Furthermore, we show that for binding events separated by 200-400 bp ChIP-exo exhibits a significantly higher sensitivity.
+
+%    \SK{The last sentence does not really make much sense? }, \RW{I
+%      meant that when we have a peak composed by more than one binding
+%      event, as the mean distance between binding events increases
+%      then ChIP-exo is more likely to identify all the binding events,
+%      I used sensitivity because we defined it as the proportion of
+%      identified binding events in a given peak. I think it depends if
+%      we decide to add Fig8 in the main article or in the
+%      supplement.}
+
 %\pname pipeline explores and quantifies these aspects by partitioning the experiment reads into a collection of regions, calculating a series of summary statistics for each region, providing visualizations and calculating measures to globally asses t quality of a ChIP-exo experiment.
-
-    We evaluated our QC pipeline with both large collections of public
-    ChIP-exo/nexus data and multiple, new ChIP-exo datasets from
-    \textit{E. Coli}.  \pname{} analysis of these datasets resulted in
-    guidelines for using these QC metrics across a wide range of
-    sequencing depths and further insights for modeling ChIP-exo
-    data. %In addition, we demonstrate that the
-    %ChIP-exo QC pipeline it is also applicable to ChIP-nexus data,
-    %showing that those experiments present higher quality than
-    %ChIP-exo experiments under similar conditions.
-    Finally, although ChIP-exo experiments have been compared to
-    ChIP-seq experiments with single-end (SE) sequencing, we provide,
-    for the first time, comparisons with paired-end (PE) ChIP-seq
-    experiments.
-    %We compared ChIP-exo with Paired End (PE) and Single End (SE)
-    %ChIP-Seq and found the following characteristics: First, although
-    %often assumed in ChIP-exo data analysis methods, the ``peak pair''
-    %assumptions does not hold locally in actual ChIP-exo data. Second,
-    %we for the first time compared PE ChIP-Seq with ChIP-exo and found
-%    that 
-    We illustrate that, at fixed sequencing depths, ChIP-exo provides
-    higher sensitivity, specificity, and spatial resolution than PE
-    ChIP-seq and both significantly outperform their SE ChIP-seq
-    counterpart. Furthermore, we show that ChIP-exo and PE ChIP-seq
-    are comparable in sensitivity for closely located binding events,
-    but as the average distance between binding events increases,
-    ChIP-exo exhibits higher sensitivity than PE ChIP-Seq.
-
-    \SK{The last sentence does not really make much sense? }, \RW{I
-      meant that when we have a peak composed by more than one binding
-      event, as the mean distance between binding events increases
-      then ChIP-exo is more likely to identify all the binding events,
-      I used sensitivity because we defined it as the proportion of
-      identified binding events in a given peak. I think it depends if
-      we decide to add Fig8 in the main article or in the
-      supplement.}
-
+%In addition, we demonstrate that the ChIP-exo QC pipeline it is also applicable to ChIP-nexus data, showing that those experiments present higher quality than ChIP-exo experiments under similar conditions.
+%We compared ChIP-exo with Paired End (PE) and Single End (SE) ChIP-Seq and found the following characteristics: First, although often assumed in ChIP-exo data analysis methods, the ``peak pair'' assumptions does not hold locally in actual ChIP-exo data.
+Second, we for the first time compared PE ChIP-Seq with ChIP-exo and found that
+
+% ChIP-exo and PE ChIP-seq are comparable in sensitivity for closely located binding events, but as the average distance between binding events increases, ChIP-exo exhibits higher sensitivity than PE ChIP-seq.
 \end{abstract}
@@
 \kwd{Transcription Factor}  
-  \kwd{Binding Site Identification with High-Resolution}
 \kwd{Deconvolution}
+\kwd{Binding Site Identification with High-Resolution.}
 \end{keyword}
@@
 
-Chromatin Immunoprecipitation followed by exonuclease digestion and
-next generation sequencing (ChIP-exo) is currently one of the
-state-of-the-art high throughput assays for profiling protein-DNA
-interactions at or close to single base-pair resolution
-\cite{exo1}. It presents a powerful alternative to popularly used
-ChIP-seq (Chromatin immunoprecipitation coupled with next generation
-sequencing) assay. ChIP-exo experiments first capture millions of DNA
-fragments ($150$ - $250$ bps in length) that the protein under study
-interacts with using a protein-specific antibody and random
-fragmentation of DNA. Then, $\lambda$-exonuclease ($\lambda$-exo) is
-deployed to trim the $5^{\prime}$ end of each DNA fragment to each
-protein-DNA interaction boundary. This step is unique to ChIP-exo and
-aims to achieve significantly higher spatial resolution compared to
-ChIP-seq. Finally, high throughput sequencing of a small region ($36$
-to $100$ bps) at the $5^{\prime}$ end of each fragment generates
-millions of reads.  Similarly, ChIP-nexus (Chromatin
-Immunoprecipitation followed by exonuclease digestion, unique barcode,
-single ligation and next generation ligation) \cite{chipnexus} is a
-further modification on the ChIP-exo protocol. ChIP-nexus aims to
-overcome limitations of ChIP-exo by yielding high complexity libraries
-with numbers of cells comparable to that of ChIP-seq experiments. This
-is achieved by reducing the numbers of ligations in the standard
-ChIP-exo protocol from two to one, and adding unique, randomized
-barcodes to adaptors to enable monitoring of
-overamplification. Figure~\ref{fig:1}A illustrates the differences between
-different ChIP-based protocols: ChIP-exo, single-end (SE) ChIP-seq,
-paired-end (PE) ChIP-seq, ChIP-nexus. The $5^{\prime}$ ends of a
-ChIP-exo/nexus experiment are clustered more tightly around the
-binding sites of the protein than in a ChIP-seq experiment. In a PE
-ChIP-seq experiment, both ends are sequenced as opposed to only the
-$5^{\prime}$ end in a SE ChIP-seq.
-
-% \RW{This modification makes a library generated
-%   using the ChIP-nexus protocol to be larger than one generated with
-%   ChIP-exo, which }
-%where both sequencing adaptors are ligated at the
-%end of the ChIP fragments.
+Chromatin Immunoprecipitation followed by exonuclease digestion and next generation sequencing (ChIP-exo) is currently one of the state-of-the-art high throughput assays for profiling protein-DNA interactions at or close to single base-pair resolution \cite{exo1}.
+It presents a powerful alternative to popular ChIP-seq (Chromatin Immunoprecipitation coupled with next generation sequencing) assay.
+ChIP-exo experiments first capture millions of DNA fragments ($150$ - $250$ bps in length) that the protein under study interacts with using a protein-specific antibody and random fragmentation of DNA.
+Then, $\lambda$-exonuclease ($\lambda$-exo) is deployed to trim the $5^{\prime}$ end of each DNA fragment to each protein-DNA interaction boundary.
+This step is unique to ChIP-exo and aims to achieve significantly higher spatial resolution compared to ChIP-seq.
+Finally, high throughput sequencing of a small region ($36$ to $100$ bps) at the $5^{\prime}$ end of each fragment generates millions of reads.
+Similarly, ChIP-nexus (Chromatin Immunoprecipitation followed by exonuclease digestion, unique barcode, single ligation and next generation ligation) \cite{chipnexus} is a further modification on the ChIP-exo protocol.
+ChIP-nexus aims to overcome limitations of ChIP-exo by yielding high complexity libraries with numbers of cells comparable to that of ChIP-seq experiments.
+This is achieved by reducing the numbers of ligations in the standard ChIP-exo protocol from two to one, and adding unique, randomized barcodes to adaptors to enable monitoring of overamplification.
+Figure~\ref{fig:1}A illustrates the differences between different ChIP-based protocols: ChIP-exo, single-end (SE) ChIP-seq, paired-end (PE) ChIP-seq, ChIP-nexus.
+The $5^{\prime}$ ends of a ChIP-exo/nexus experiment are clustered more tightly around the binding sites of the protein than in a ChIP-seq experiment.
+In a PE ChIP-seq experiment, both ends are sequenced as opposed to only the $5^{\prime}$ end in a SE ChIP-seq.
+
+% \RW{This modification makes a library generated using the ChIP-nexus protocol to be larger than one generated with ChIP-exo, which where both sequencing adaptors are ligated at the end of the ChIP fragments.
 Then, after exonuclease digestion, DNA self-circularization with circLigase, and restriction enzyme cutting between the two adaptors, the final library is amplified.
@@
 
-\SK{Note to Rene: ChIP-nexus paper has a good description of what can
-  go wrong with ChIP-exo; It seems like most of the read imbalance
-  could also be due to ligation inefficiency. Although this probably
-  does not explain why we have reads only from one strand in some
-  regions. This is more likely explained by over digestion in one
-  strand or single stranded TF-DNA interaction. Think a bit more on
-  these issues and understand how the ligation could give rise to
-  strand imbalance, we can ask Bob to have a closer look at the
-  section where we have this discussion.}
-
-% \RW{1 - I think it may come from the starting material in each side. I
-%   think that at least the number of fragments on each side must
-%   preserve that if there is more starting reads in one side, then this
-%   must be the case after the amplification.}
-
-% \RW{2 - There is not control about how the enzyme digests the
-%   fragments, as far as we know it starts on the $5\prime$ side and
-%   stops until it finds the TF or it runs out of energy to keep
-%   digesting. I think Bob mentioned that it may be possible for the
-%   enzyme to pass though the TF if it is left digesting for a longer
-%   time, but if that's the case then the question may be why it
-%   preffers one strand over the other, which may lead to the GC-bias
-%   but as far as I can tell is very hard to observe that because that
-%   bias appear also in ChIP-seq (and it seeems that also appear in the
-%   circularization step from ChIP-nexus).}
-
-% \RW{3 - I think that the regions formed by fragments in one strand,
-%   may be copies of few undigested fragments, in a good experiment the
-%   fragments are expected to be clustered around the binding sites,
-%   therefore it is less likely to observe this fragments scattered in
-%   the genome. In a bad experiment is the opposite case, so the
-%   fragments are not clustered so if the enzyme doesn't have a strong
-%   preference to how is attracted to the fragments then the remaining
-%   ones can be seen as random positions around the genome, this
-%   positions are copied several times during the amplification and that
-%   may be the reason why we observe something as scattered. Perhaps if
-%   the enzyme was left for longer time, then the starting material that
-%   later was amplified into this regions would have been digested and
-%   some of this regions wouldn't appear (I am thinking about the deeply
-%   sequenced TBP experiments.)}
-
-Although ChIP-exo/nexus protocols are being adopted by research
-community, features of ChIP-exo data, especially those pertaining to
-data quality, have not been investigated much. The key features of
-ChIP-exo/nexus that separate them from ChIP-seq are broadly as
-follows. First, DNA libraries generated by the ChIP-exo protocol are
-expected to be less complex than the libraries generated by ChIP-seq
-\cite{exo_review}) because digestion by $\lambda$-exo is expected to
-restrict the space of genomic positions that sequencing reads can map
-to, to small local regions around the actual binding sites. Therefore,
-in high quality and especially deeply sequenced ChIP-exo datasets, it
-is possible to observe large numbers of reads accumulating at a small
-number of bases due to actual signal rather than overamplification
-bias as commonly observed in ChIP-seq experiments.  Second, although
-we expect approximately the same numbers of reads from both DNA
-strands at a given binding site, there may be locally more reads in
-one strand than in the other, owing to $\lambda$-exo efficiency,
-ligation efficiency, or other factors.  This is a key point with
-implications on the statistical analysis of ChIP-exo
-data. Specifically, currently available ChIP-exo specific statistical
-analysis methods (e.g., Mace \cite{mace}, CexoR \cite{cexor}, and
-Peakzilla \cite{peakzilla}) rely on the existence of peak-pairs formed
-by forward and reverse strand reads at the binding site.  Finally,
-most of current widely used ChIP-seq quality control (QC) guidelines
-\cite{encode_qc} may not be directly applicable to ChIP-exo data.
-
-To address these challenges, we develop a suite of diagnostic plots
-and summary statistics and implement them in a versatile \texttt{R
-  package} named \pname{}.  We apply this pipeline on a large
-collection of public and newly generated ChIP-exo/nexus data. We
-validate implications of the QC pipeline by evaluation of the samples
-for features that capture high signal to noise such as occurrences of
-motifs recognized by the profiled DNA interacting protein. Our
-analyses of this large collection of data revealed that the so-called
-ChIP-exo peak-pair assumption is subject to violations. To further
-address this and provide a platform where ChIP-exo and ChIP-seq
-experiments can be evaluated with comparable methods, we assess
-performances of recently developed methods suitable for ChIP-exo
-analysis, including dpeak \cite{dpeak} and GEM \cite{gem}. We observe
-that dPeak performs as good or better than the available ChIP-exo
-methods and provides a platform where PE and SE ChIP-seq can be
-compared with their ChIP-exo counterpart. Our comparisons of PE
-ChIP-seq with ChIP-exo interestingly highlights that while ChIP-exo
-outperforms PE ChIP-seq in terms of resolution and detection power,
-both are significantly better than SE ChIP-seq.
-
-% while there are not established QC pipelines for ChIP-exo; previous
-% ChIP-exo analyses used ChIP-Seq samples to compare the resolution
-% between experiments \cite{exo1,exoillumina,exo2}.
-
-%to interrogate these
-%biases in a ChIP-exo experiment and globally asses the enrichment and
-%library complexity of a ChIP-exo sample and a procedure to distinguish
-%low complexity libraries from deeply sequenced experiments. 
-%This
-%aspect is unique to ChIP-exo, since the exonuclease enzyme it is
-%expected to digested the reads that are not bound to a transcription
-%factor, therefore the number of bases where a ChIP-exo fragment could
-%be potentially aligned is reduced. That way, in a high quality
-%ChIP-exo experiment is possible to observe a large amount of reads to
-%be aligned to unique position due to genomic signal instead of PCR
-%artifacts.
-
-%In order to obtain the potential benefits of ChIP-exo on protein
-%binding site identification, it is critical to use algorithms that
-%could fully utilize information available in ChIP-exo data. Rhee and
-%Pugh, 2011 \nocite{exo1} discussed that reads in the forward and reverse
-%strand might construct peak pairs around bound proteins, of which
-%heights were implicitly assumed to be symmetric. Based on this
-%rationale, they used the ``peak pair method'' that predicts the
-%midpoint of two modes of peak pairs as potential binding
-%sites. Recently developed ChIP-exo data analysis methods, such as Mace
-%(Wang et al, 20114\nocite{mace}), CexoR (Madrigal, 2015\nocite{cexor})
-%and Peakzilla (Bardet et al., 2013\nocite{peakzilla}), are also based
-%on this peak pair assumption. However, appropriateness of such
-%assumption was not fully evaluated in the literature yet. Furthermore,
-%it is still unknown which factors could affect protein binding site
-%identification using ChIP-exo data. In order to address this problem,
-%we investigated various aspects of ChIP-exo data by contrasting them
-%with their respective ChIP-Seq experiments.
-
-%Currently, research on statistical methods for ChIP-exo data is still
-%in its very early stage. Although many methods have been proposed to
-%identify protein binding sites from ChIP-Seq data (reviewed by
-%Wilbanks and Facciotti, 2012\nocite{evaluation} and Pepke and Wold,
-%2009 \nocite{computation}), such as MACS (Zhang et al.,
-%2008\nocite{macs}), CisGenome (Ji et al., 2008\nocite{cisgenome}) and
-%MOSAiCS (Kuan et al., 2009\nocite{mosaics}), these approaches might
-%not fully utilize potentials of ChIP-exo data for high resolution
-%identification of protein binding sites. Specifically these approaches
-%reveal protein binding sites only in lower resolution, i.e., at an
-%interval of hundreds to thousands of base pairs. Furthermore, they
-%implicitly assume that there is only one ``mode'' or ``predicted
-%binding location'' per this wide genomic interval. More recently,
-%deconvolution algorithms such as Deconvolution (Lun et al.,
-%2009\cite{csdeconv}), GEM (Guo et al., 2012\nocite{gem}, an improved
-%version of Guo et al., 2010\nocite{gps} ) and PICS (Zhang et al.,
-%2010\nocite{pics}) have been proposed to identify binding sites in
-%higher resolution using ChIP-Seq data. However, most of them are still
-%not tailored for ChIP-exo and PE and SE ChIP-Seq data in a unified
-%framework and as a result, currently available methods are not
-%appropriate for fair comparison between ChIP-exo and ChIP-Seq. To
-%address these limitations, we developed and utilized an improved
-%version of dPeak (Chung et al., 2013\nocite{dpeak}), a high resolution
-%binding site identification (deconvolution) algorithm that we
-%previously developed for PE and SE ChIP-Seq data, so that it can also
-%handle ChIP-exo data. The dPeak algorithm implements a probabilistic
-%model that accurately describes the ChIP-exo and ChIP-Seq data
-%generation process.
-
-%Some of the key findings in this work are as follows. First, we
-%demonstrate that the ``peak pair'' assumption of Rhee and Pugh,
-%2013\nocite{exo2} does not hold well in real ChIP-exo data. Second, we
-%found that when analyzing ChIP-exo data and the Input control is not
-%available, it is useful to adjust for GC content and mappability
-%biases to improve peak calling and binding site identification. Third,
-%we evaluated several methods to identify binding events and dPeak
-%performs competitively respect to GEM and MACE when analyzing ChIP-exo
-%data. Finally, when comparable number of reads is used for both
-%ChIP-exo and ChIP-Seq , dPeak coupled with ChIP-exo data provides
-%resolution comparable to PE ChIP-Seq and both significantly improve
-%the resolution of protein binding identification compared to SE-based
-%analysis with any of the available methods.
+% \SK{Note to Rene: ChIP-nexus paper has a good description of what can go wrong with ChIP-exo; It seems like most of the read imbalance could also be due to ligation inefficiency.
+Although this probably does not explain why we have reads only from one strand in some regions.
+This is more likely explained by over digestion in one strand or single stranded TF-DNA interaction.
+Think a bit more on these issues and understand how the ligation could give rise to strand imbalance, we can ask Bob to have a closer look at the section where we have this discussion.}
+
+% \RW{1 - I think it may come from the starting material in each side.
+I think that at least the number of fragments on each side must preserve that if there is more starting reads in one side, then this must be the case after the amplification.}
+% \RW{2 - There is not control about how the enzyme digests the fragments, as far as we know it starts on the $5\prime$ side and stops until it finds the TF or it runs out of energy to keep digesting.
+I think Bob mentioned that it may be possible for the enzyme to pass though the TF if it is left digesting for a longer time, but if that's the case then the question may be why it preffers one strand over the other, which may lead to the GC-bias but as far as I can tell is very hard to observe that because that bias appear also in ChIP-seq (and it seeems that also appear in the circularization step from ChIP-nexus).}
+
+% \RW{3 - I think that the regions formed by fragments in one strand, may be copies of few undigested fragments, in a good experiment the    fragments are expected to be clustered around the binding sites, therefore it is less likely to observe this fragments scattered in the genome.
+In a bad experiment is the opposite case, so the fragments are not clustered so if the enzyme doesn't have a strong preference to how is attracted to the fragments then the remaining ones can be seen as random positions around the genome, this positions are copied several times during the amplification and that may be the reason why we observe something as scattered.
+Perhaps if the enzyme was left for longer time, then the starting material that later was amplified into this regions would have been digested and some of this regions wouldn't appear (I am thinking about the deeply sequenced TBP experiments.)}
+
+Although ChIP-exo/nexus protocols are being adopted by research community, features of ChIP-exo data, especially those pertaining to data quality, have not been investigated much.
+The key features of ChIP-exo/nexus that separate them from ChIP-seq are broadly as follows.
+First, DNA libraries generated by the ChIP-exo protocol are expected to be less complex than the libraries generated by ChIP-seq \cite{exo_review}) because digestion by $\lambda$-exo is expected to restrict the space of genomic positions that sequencing reads can map to, to small local regions around the actual binding sites.
+Therefore, in high quality and especially deeply sequenced ChIP-exo datasets, it is possible to observe large numbers of reads accumulating at a small number of bases due to actual signal rather than overamplification bias as commonly observed in ChIP-seq experiments.
+Second, although we expect approximately the same numbers of reads from both DNA strands at a given binding site, there may be locally more reads in one strand than in the other, owing to $\lambda$-exo efficiency, ligation efficiency, or other factors.
+This is a key point with implications on the statistical analysis of ChIP-exo data.
+Specifically, currently available ChIP-exo specific statistical analysis methods (e.g., Mace \cite{mace}, CexoR \cite{cexor}, and Peakzilla \cite{peakzilla}) rely on the existence of peak-pairs formed by forward and reverse strand reads at the binding site.
+Finally, most of current widely used ChIP-seq quality control (QC) guidelines \cite{encode_qc} may not be directly applicable to ChIP-exo data.
+
+To address these challenges, we develop a suite of diagnostic plots and summary statistics and implement them in a versatile \texttt{R ackage} named \pname{}.
+We apply this pipeline on a large collection of public and newly generated ChIP-exo/nexus data.
+We validate the QC pipeline by evaluation of the samples for features that capture high signal to noise such as occurrences of motifs recognized by the profiled DNA interacting protein.
+Our analyses of this large collection of data reveal that the ChIP-exo peak-pair assumption, underlying most of the ChIP-exo analysis pipelines is subject to violations.
+To further address this and provide a platform where ChIP-exo and ChIP-seq experiments can be evaluated with comparable methods, we assess performances of recently developed methods suitable for ChIP-exo analysis, including dPeak \cite{dpeak} and GEM \cite{gem}.
+We observe that dPeak performs as good or better than the available ChIP-exo methods and provides a platform where PE and SE ChIP-seq can be compared with their ChIP-exo counterpart.
+Our comparisons of PE ChIP-seq with ChIP-exo interestingly highlights that while ChIP-exo outperforms PE ChIP-seq in terms of resolution and detection power, both are significantly better than SE ChIP-seq.
+
+% while there are not established QC pipelines for ChIP-exo; previous ChIP-exo analyses used ChIP-Seq samples to compare the resolution  between experiments \cite{exo1,exoillumina,exo2}.
+
+%to interrogate these biases in a ChIP-exo experiment and globally asses the enrichment and library complexity of a ChIP-exo sample and a procedure to distinguish low complexity libraries from deeply sequenced experiments.
+This aspect is unique to ChIP-exo, since the exonuclease enzyme it is expected to digested the reads that are not bound to a transcription factor, therefore the number of bases where a ChIP-exo fragment could be potentially aligned is reduced.
+That way, in a high quality ChIP-exo experiment is possible to observe a large amount of reads to be aligned to unique position due to genomic signal instead of PCR artifacts.
+
+%In order to obtain the potential benefits of ChIP-exo on protein binding site identification, it is critical to use algorithms that could fully utilize information available in ChIP-exo data.
+Rhee and Pugh, 2011 \nocite{exo1} discussed that reads in the forward and reverse strand might construct peak pairs around bound proteins, of which heights were implicitly assumed to be symmetric.
+Based on this rationale, they used the ``peak pair method'' that predicts the midpoint of two modes of peak pairs as potential binding sites.
+Recently developed ChIP-exo data analysis methods, such as Mace (Wang et al, 20114\nocite{mace}), CexoR (Madrigal, 2015\nocite{cexor}) and Peakzilla (Bardet et al., 2013\nocite{peakzilla}), are also based on this peak pair assumption.
+However, appropriateness of such assumption was not fully evaluated in the literature yet.
+Furthermore, it is still unknown which factors could affect protein binding site identification using ChIP-exo data.
+In order to address this problem, we investigated various aspects of ChIP-exo data by contrasting them with their respective ChIP-Seq experiments.
+
+%Currently, research on statistical methods for ChIP-exo data is still in its very early stage.
+Although many methods have been proposed to %identify protein binding sites from ChIP-Seq data (reviewed by Wilbanks and Facciotti, 2012\nocite{evaluation} and Pepke and Wold, 2009 \nocite{computation}), such as MACS (Zhang et al., 2008\nocite{macs}), CisGenome (Ji et al., 2008\nocite{cisgenome}) and MOSAiCS (Kuan et al., 2009\nocite{mosaics}), these approaches might not fully utilize potentials of ChIP-exo data for high resolution identification of protein binding sites.
+Specifically these approaches reveal protein binding sites only in lower resolution, i.e., at an interval of hundreds to thousands of base pairs.
+Furthermore, they implicitly assume that there is only one ``mode'' or ``predicted binding location'' per this wide genomic interval.
+More recently, deconvolution algorithms such as Deconvolution (Lun et al., 2009\cite{csdeconv}), GEM (Guo et al., 2012\nocite{gem}, an improved version of Guo et al., 2010\nocite{gps} ) and PICS (Zhang et al., 2010\nocite{pics}) have been proposed to identify binding sites in higher resolution using ChIP-Seq data.
+However, most of them are still not tailored for ChIP-exo and PE and SE ChIP-Seq data in a unified framework and as a result, currently available methods are not appropriate for fair comparison between ChIP-exo and ChIP-Seq.
+To address these limitations, we developed and utilized an improved version of dPeak (Chung et al., 2013\nocite{dpeak}), a high resolution binding site identification (deconvolution) algorithm that we previously developed for PE and SE ChIP-Seq data, so that it can also handle ChIP-exo data.
+The dPeak algorithm implements a probabilistic model that accurately describes the ChIP-exo and ChIP-Seq data generation process.
+
+%Some of the key findings in this work are as follows.
+First, we demonstrate that the ``peak pair'' assumption of Rhee and Pugh, 2013\nocite{exo2} does not hold well in real ChIP-exo data.
+Second, we found that when analyzing ChIP-exo data and the Input control is not available, it is useful to adjust for GC content and mappability biases to improve peak calling and binding site identification.
+Third, we evaluated several methods to identify binding events and dPeak performs competitively respect to GEM and MACE when analyzing ChIP-exo data.
+Finally, when comparable number of reads is used for both ChIP-exo and ChIP-Seq , dPeak coupled with ChIP-exo data provides resolution comparable to PE ChIP-Seq and both significantly improve the resolution of protein binding identification compared to SE-based analysis with any of the available methods.
 
@@
 
-\subsection*{Publicly available ChIP-exo/nexus and novel \textit{E. coli} ChIP-seq/exo datasets}
-We utilized a rich collection of publicly available ChIP-exo/nexus
-data from multiple organisms to build and evaluate our quality control
-pipeline (Table~\ref{tab:qc}). These include: CTCF factor in human
-HeLa cells \cite{exo1}; ER factor in human MCF-7 cells
-\cite{exoillumina}; GR factor in IMR90, K562, and U2OS human cells
-\cite{starick15}; TBP factor in human K562 cells \cite{venters13}.
-ChIP-nexus data included experiments from \cite{chipnexus} profiling
-TBP in human K562 cells, MyC and Max in \textit{D. melanogaster} S2
-cells, and Twist and Dorsal in \textit{D. melanogaster} embryo.
-
-In order to have a setting where we can compare SE and PE ChIP-seq
-with their ChIP-exo counterpart, we profiled $\sig$ under a variety of
-conditions in \textit{E. coli} with both ChIP-exo and PE ChIP-seq
-(Table~\ref{tab:qc_sig}). Collectively, we generated $\sig$ factor
-ChIP-exo, PE and SE ChIP-seq experiments under aerobic ($+O_2$) and
-anaerobic ($-O_2$) conditions in glucose minimal media, for simplicity
-we named these experiments as E1, P1 and S1 respectively. Similarly,
-we generated $\sig$ factor ChIP-exo, PE and SE (generated \textit{in
-  silico}) ChIP-seq experiments in \textit{E. coli} under aerobic
-($+O_2$) conditions with and without rifampicin treatment.  We refer
-to these latter set of experiments as time point 20 minutes and time
-point 0 minutes, we also named these experiments E2, P2 and S2 for
-ChIP-exo, PE ChIP-seq and SE ChIP-seq respectively.
-% \SK{Check for
-%   consistency with other parts of the paper and correct, ok to use
-%   different naming convention}.
-We specifically used these experimental data for comparisons of
-ChIP-exo and PE ChIP-seq assays in identifying closely spaced binding
-events and their resolution, i.e., physical proximity of the predicted
-events to the actual binding sites.  This comparison benefits from
-using $\sig$ which is a transcription initiation factor of
-housekeeping genes in \textit{E. coli}. Many \textit{E. coli}
-promoters contain multiple transcription start sites (TSS). These TSSs
-are often closely spaced, i.e., within 10 $\sim$ 150 bps of each
-other, and are considered to be multiple ``switches'' that
-differentially regulate gene expression under diverse growth
-conditions \cite{regulondb}.
+\subsection*{Publicly available ChIP-exo/nexus and novel \textit{E. Coli} ChIP-seq/exo datasets}
+
+We utilized a rich collection of publicly available ChIP-exo/nexus data from multiple organisms to build and evaluate our quality control pipeline (Table~\ref{tab:qc}).
+These include: CTCF factor in human HeLa cells \cite{exo1}; ER factor in human MCF-7 cells \cite{exoillumina}; GR factor in IMR90, K562, and U2OS human cells \cite{starick15}; TBP factor in human K562 cells \cite{venters13}.
+ChIP-nexus data included experiments from \cite{chipnexus} profiling TBP in human K562 cells, MyC and Max in \textit{D.
+melanogaster} S2 cells, and Twist and Dorsal in \textit{D.
+melanogaster} embryo.
+
+In order to have a setting where we can compare SE and PE ChIP-seq with their ChIP-exo counterpart, we profiled $\sig$ under a variety of conditions in \textit{E.
+coli} with both ChIP-exo and PE ChIP-seq (Table~\ref{tab:qc_sig}).
+Collectively, we generated $\sig$ factor ChIP-exo, PE and SE ChIP-seq experiments under aerobic ($+O_2$) and anaerobic ($-O_2$) conditions in glucose minimal media.
+For simplicity, we named these experiments as E1, P1, and S1, respectively.
+Similarly, we generated $\sig$ factor ChIP-exo, PE and SE (generated \textit{in   silico}) ChIP-seq experiments in \textit{E.
+Coli} under aerobic ($+O_2$) conditions with and without rifampicin treatment.
+We refer to these latter set of experiments as time point 20 minutes and time point 0 minutes.
+We also named these experiments E2, P2, and S2 for ChIP-exo, PE ChIP-seq and SE ChIP-seq, respectively.
+We specifically used these experimental data for comparisons of ChIP-exo and PE ChIP-seq assays in identifying closely spaced binding events and their resolution, i.e., physical proximity of the predicted events to the actual binding sites.
+This comparison benefits from using $\sig$ which is a transcription initiation factor of housekeeping genes in \textit{E.
+coli}.
+Many \textit{E.
+Coli} promoters contain multiple transcription start sites (TSS).
+These TSSs are often closely spaced, i.e., within 10 $\sim$ 150 bps of each other, and are considered to be multiple ``switches'' that differentially regulate gene expression under diverse growth conditions \cite{regulondb}.
+
+% \SK{Check for consistency with other parts of the paper and correct, ok to use different naming convention}.
 
@@
 
- % \SK{Maybe add a column as sample
- %  number to this table to make referencing easy}
-\textit{Read distributions within signal and background regions.} We
-first compared ChIP-seq and ChIP-exo in terms of data features that
-are well studied in ChIP-seq studies.  Our $\sig$ ChIP-seq and
-ChIP-exo samples from \textit{E. coli} are especially well suited for
-this task since they are all deeply sequenced compared to the genome
-size of \textit{E. coli}.  Figure~\ref{fig:1} summarizes this
-comparison for one biological replicate of ChIP-exo and ChIP-seq
-experiments from the same biological conditions (samples EB1-1 and
-P1-1 from Table~\ref{tab:qc_sig}). Comparisons with other paired
-\textit{E. coli} ChIP-seq and ChIP-exo samples led to similar
-conclusions (Supplementary Figure~\ref{sfig:comp1}).
-%compared various factors that could affect binding site
-%identification between ChIP-exo and ChIP-Seq data by using the
-%ChIP-exo reads from the first biological sample and first replicate
-%grown under aerobic condition (first line from Table \ref{tab:qc_sig})
-%and a SE ChIP-Seq replicate grown under the same conditions. 
-We summarized the extended raw read counts within $150$ bps
-non-overlapping intervals, i.e., bins, interrogating the genome.
-%In order
-%to compare distribution of signal and background between ChIP-exo and
-%ChIP-Seq data, we counted the number of extended reads mapping to a
-%partition of the genome into non-overlapping bins. 
-Figure~\ref{fig:1}B depicts that, as expected, ChIP read counts
-from ChIP-exo and ChIP-seq are linearly correlated especially at high
-read counts. This indicates that signals for potential binding sites
-are well reproducible between ChIP-exo and ChIP-seq data.  In
-contrast, there is a clear difference among the two data types for
-bins with low read counts, highlighting expected differences in the
-background read distributions of the two data types.  \SK{***After
-  seeing these types of plots from other samples (e..g., ER in MCF7,
-  it is unlikely that what we are seeing from $\sig$ samples regarding
-  background reads (the next few sentences) is typical. For CTCF and
-  $\sig$, the ChIP-exo samples have a wider dynamic range, but this
-  could be due to differences in sequencing depth. *** Specifically,
-  background reads from ChIP-seq are almost uniformly distributed over
-  background (non-enriched) regions. In comparison, background regions
-  in ChIP-exo show larger variation but have overall lower read
-  counts. Furthermore, there are considerably large numbers of bins
-  with zero read counts in ChIP-exo and non-zero read counts in
-  ChIP-seq.
-%and the
-%ChIP tag counts in there regions were significantly larger than
-%zero. In contrast, in ChIP-exo data, there was larger variation in
-%ChIP tag counts among background regions and ChIP tag counts were much
-%lower in these regions compared to ChIP-Seq data. There were also
-%large proportion of regions without any read in ChIP-exo data. 
-  Overall, these observations indicate that a much smaller portion of
-  the genome is expected to be background in ChIP-exo compared to
-  ChIP-seq and methods that specifically model the background read
-  distribution might benefit from acknowledging this.} \RW{Figure 2B
-  from the chip-nexus paper makes a similar comparison but specific to
-  peaks, i.e. our argument about both ChIP-seq and ChIP-exo is
-  correlated in high signal region still holds.}
-
-\textit{Peak-pair assumption.} We next evaluated the peak-pair
-assumption, i.e., a cluster of reads in the forward strand is usually
-paired with a cluster of reads in the reverse strand that is located
-on the right-hand-side of the binding site, that is commonly utilized
-in designing statistical analysis methods for ChIP-exo data
-\cite{mace, cexor, peakzilla}.  We considered the set of peaks
-identified in both the ChIP-seq and ChIP-exo samples as high quality
-peaks (Materials and Methods) and calculated the proportion of forward
-strand reads in these regions (Figure~\ref{fig:1}C). This plot
-reveals a higher level of strand imbalance for ChIP-exo compared to
-ChIP-seq.
-% In order to evaluate this assumption, we reviewed the proportion of
-% reads in the forward strand in high quality ChIP-exo peaks such as
-% at least one binding site is predicted in both ChIP-exo and ChIP-Seq
-% data.  We found that strands of reads were much less balanced in
-% ChIP-exo data than in ChIP-Seq data in these regions with potential
-% binding sites (Fig. \ref{fig:comp}B) and this indicates that the
-% peak pair assumption might not hold in ChIP-exo data.
-Potential reasons for this observation include ligation efficiency,
-efficiency of $\lambda$-exo digestion, and single-stranded protein-DNA
-interactions. Overall, such an imbalance is prevalent in 85\% of the
-ChIP-exo samples used in this paper \RW{(Supplement
-  Figures~\ref{sfig:comp1} to \ref{sfig:comp3})}. \RW{Note: The only ChIP-exo
-  experiment were it seems that ChIP-exo is not more imbalanced than
-  ChIP-seq is Carroll's ER dataset.}
-%the enzyme digestion or the strength at which the protein binds to the
-%DNA: Although it is expected for the exonuclease enzyme to digest the
-%ChIP fragments starting by their $5\prime$ ends and stopping when
-%finding a binding event, it may not be able to reach and digest every
-%fragment in the ChIP sample. On the other hand, if the protein is not
-%bound to both DNA strand, the enzyme may completely digest the
-%fragment. Finally both effect are increased by the PCR amplification.
-
-\textit{Mappability and GC-content bias.} We next evaluated ChIP-exo
-data of CTCF in HeLa cells \cite{exo1} to investigate biases inherent
-to next generation sequencing experiments with eukaryotic genomes.
-Figures \ref{fig:1}E and \ref{fig:1}F display the bin-level
-average read counts against mappability and GC-content. Each data
-point is obtained by averaging the read counts across bins with the
-same mappability of GC-content.  These biases, increasing linear trend
-with mappability and non-linear trend with GC-content, are similar to
-those observed in ChIP-seq datasets \cite{benjamini2011,
-  mosaics,quest}.
-%In Figure \ref{fig:comp}C it is shown that the ChIP-exo tag
-%counts linearly increases with the mappability score and in Figure
-%\ref{fig:comp}D it is shown that for GC - content below 0.6, the mean
-%ChIP tag count increases and for GC - content greater than 0.6 it
-%shows a decreasing trend. Benjamini and Speed,
-%2011\nocite{benjamini2011} and Kuan et al., 2011\nocite{mosaics}
-%studied the presence of the mappability and GC - content biases in
-%ChIP-Seq's background. It is not surprising to see these biases also
-%present in ChIP-exo data, since ChIP-Seq and ChIP-exo signal seems to
-%be linearly correlated for enriched regions (Figure
-%\ref{fig:comp}A). Rozowsky et al., 2009\nocite{peakseq} and Valouev et
-%al., 2008\nocite{quest} provide in depth analysis of the mappability
-%and GC - content biases for ChIP-Seq respectively. 
-This observation  indicates that analysis of  ChIP-exo data
-should benefit from methods that take into account  apparent
-sequencing biases such as mappability and GC content, mostly when an
-input  control sample is not available.
-
-\subsection*{Application of ENCODE ChIP-seq quality metrics to
-  ChIP-exo and ChIP-nexus data}
-
-ENCODE consortium established empirical and widely used QC metrics on
-ChIP-seq data \cite{encode_qc}. Currently, these constitute the
-state-of-the-art QC pipelines for these high throughput experiments.
-We evaluated how these metrics, namely Normalized Strand
-Cross-Correlation (NSC), Relative Strand Cross-Correlation (RSC), and
-PCR Bottleneck Coefficient (PBC) defined at
-\url{https://genome.ucsc.edu/ENCODE/qualityMetrics.html}
-\cite{encode_qc}, behave on ChIP-exo/nexus data in Tables~\ref{tab:qc}
-and \ref{tab:qc_sig}.
-
-%We continued our exploration by investigating whether the current
-%state-of-the-art QC pipelines for ChIP-Seq are suitable for ChIP-exo
-%and ChIP-nexus. In Tables \ref{tab:qc_sig} and \ref{tab:qc} we
-%calculated a collection of the commonly used ChIP-Seq QC metrics using
-%the ChIP-exo and ChIP-nexus experiments instead: Normalized Strand
-%Cross-Correlation (NSC), Relative Strand Cross-Correlation and PCR
-%Bottleneck Coefficient (PBC) defined as in
-%\url{https://genome.ucsc.edu/ENCODE/qualityMetrics.html}\nocite{encode_qc}.
+\textit{Read distributions within signal and background regions.} We first compared ChIP-seq and ChIP-exo in terms of data features that
+are well studied in ChIP-seq studies.
+Our $\sig$ ChIP-seq and ChIP-exo samples from \textit{E.
+Coli} are especially well suited for this task since they are all deeply sequenced compared to the genome size of \textit{E.
+Coli}.
+Figure~\ref{fig:1}(B-E) summarizes this comparison for one biological replicate of ChIP-exo and ChIP-seq experiments from the same biological conditions (samples EB1-1 and P1-1 from Table~\ref{tab:qc_sig}).
+Comparisons with other paired \textit{E.
+Coli} ChIP-seq and ChIP-exo samples led to similar conclusions (Supplementary Figure~\ref{sfig:comp1}).
+We summarized the extended raw read counts within $150$ bps non-overlapping intervals, i.e., bins, interrogating the genome.
+Figure~\ref{fig:1}B depicts that, as observed by others, ChIP read counts from ChIP-exo and ChIP-seq are linearly correlated especially at high read counts.
+This indicates that signals for potential binding sites are well reproducible between ChIP-exo and ChIP-seq data.
+In contrast, there is a clear difference among the two data types for bins with low read counts, highlighting potential differences in the background read distributions of those data types.
+
+%\SK{***After seeing these types of plots from other samples (e..g., ER in MCF7, it is unlikely that what we are seeing from $\sig$ samples regarding background reads (the next few sentences) is typical.
+For CTCF and $\sig$, the ChIP-exo samples have a wider dynamic range, but this could be due to differences in sequencing depth.
+*** Specifically, background reads from ChIP-seq are almost uniformly distributed over   background (non-enriched) regions.
+In comparison, background regions in ChIP-exo show larger variation but have overall lower read counts.
+Furthermore, there are considerably large numbers of bins with zero read counts in ChIP-exo and non-zero read counts in ChIP-seq.
+The genome is expected to be background in ChIP-exo compared to ChIP-seq and methods that specifically model the background read distribution might benefit from acknowledging this.}
+%\RW{Figure 2B from the chip-nexus paper makes a similar comparison but specific to peaks, i.e. our argument about both ChIP-seq and ChIP-exo is correlated in high signal region still holds.}
+
+\textit{Peak-pair assumption.} We next evaluated the peak-pair assumption, i.e., a cluster of reads in the forward strand is usually paired with a cluster of reads in the reverse strand that is located on the right-hand-side of the binding site, that is commonly utilized in designing statistical analysis methods for ChIP-exo data \cite{mace, cexor, peakzilla}.
+We considered the set of peaks identified in both the ChIP-seq and ChIP-exo samples as high quality peaks (Materials and Methods) and calculated the proportion of forward strand reads in these regions (Figure~\ref{fig:1}C).
+This plot reveals a higher level of strand imbalance for ChIP-exo compared to ChIP-seq.
+Potential reasons for this observation include ligation efficiency, efficiency of $\lambda$-exo digestion, and single-stranded protein-DNA interactions.
+Overall, such an imbalance is prevalent in 85\% of the ChIP-exo samples used in this paper.
+
+% \RW{(Supplement Figures~\ref{sfig:comp1} to \ref{sfig:comp3})}. \RW{Note: The only ChIP-exo experiment were it seems that ChIP-exo is not more imbalanced than ChIP-seq is Carroll's ER dataset.}
+
+\textit{Mappability and GC-content bias.} We next evaluated ChIP-exo data of CTCF in HeLa cells \cite{exo1} to investigate biases inherent to next generation sequencing experiments with eukaryotic genomes.
+Figures \ref{fig:1}E and \ref{fig:1}F display the bin-level average read counts against mappability and GC-content.
+Each data point is obtained by averaging the read counts across bins with the same mappability of GC-content.
+These biases, increasing linear trend with mappability and non-linear trend with GC-content, are similar to those observed in ChIP-seq datasets \cite{benjamini2011,mosaics,quest}.
+This observation  indicates that analysis of  ChIP-exo data should benefit from methods that take into account  apparent sequencing biases such as mappability and GC content, mostly when an input  control sample is not available to account for variability in the background read distribution.
+
+% \SK{Maybe add a column as sample number to this table to make referencing easy}
+
+%compared various factors that could affect binding site identification between ChIP-exo and ChIP-Seq data by using the ChIP-exo reads from the first biological sample and first replicate grown under aerobic condition (first line from Table \ref{tab:qc_sig}) and a SE ChIP-Seq replicate grown under the same conditions.
+
+%In order to compare distribution of signal and background between ChIP-exo and ChIP-Seq data, we counted the number of extended reads mapping to a partition of the genome into non-overlapping bins.
+
+% In order to evaluate this assumption, we reviewed the proportion of reads in the forward strand in high quality ChIP-exo peaks such as at least one binding site is predicted in both ChIP-exo and ChIP-Seq data.
+We found that strands of reads were much less balanced in ChIP-exo data than in ChIP-Seq data in these regions with potential binding sites (Fig.
+\ref{fig:comp}B) and this indicates that the peak pair assumption might not hold in ChIP-exo data.
+
+%the enzyme digestion or the strength at which the protein binds to the DNA: Although it is expected for the exonuclease enzyme to digest the %ChIP fragments starting by their $5\prime$ ends and stopping when finding a binding event, it may not be able to reach and digest every %fragment in the ChIP sample.
+On the other hand, if the protein is not bound to both DNA strand, the enzyme may completely digest the %fragment.
+Finally both effect are increased by the PCR amplification.
+
+%In Figure \ref{fig:comp}C it is shown that the ChIP-exo tag counts linearly increases with the mappability score and in Figure \ref{fig:comp}D it is shown that for GC - content below 0.6, the mean ChIP tag count increases and for GC - content greater than 0.6 it shows a decreasing trend.
+Benjamini and Speed, 2011\nocite{benjamini2011} and Kuan et al., 2011\nocite{mosaics} studied the presence of the mappability and GC - content biases in ChIP-Seq's background.
+It is not surprising to see these biases also present in ChIP-exo data, since ChIP-Seq and ChIP-exo signal seems to be linearly correlated for enriched regions (Figure \ref{fig:comp}A).
+Rozowsky et al., 2009\nocite{peakseq} and Valouev et al., 2008\nocite{quest} provide in depth analysis of the mappability and GC - content biases for ChIP-Seq respectively.
+
+\subsection*{Application of ENCODE ChIP-seq quality metrics to ChIP-exo and ChIP-nexus data}
+
+ENCODE consortium established empirical and widely used QC metrics on ChIP-seq data \cite{encode_qc}.
+Currently, these constitute the state-of-the-art QC pipelines for these high throughput experiments.
+We evaluated how these metrics, namely Normalized Strand Cross-Correlation (NSC), Relative Strand Cross-Correlation (RSC), and PCR Bottleneck Coefficient (PBC) defined at \url{https://genome.ucsc.edu/ENCODE/qualityMetrics.html} \cite{encode_qc}, behave on ChIP-exo/nexus data in Tables~\ref{tab:qc} and \ref{tab:qc_sig}.
 
@@
     &  & & & 3 & 22,421,72  & 72.04 & 1.1975 & 0.1068 \\
-    & \multirow{3}{1em}{Human} & \multirow{3}{1em}{GR} & IMR90 & & 47,443,803 & 8.86  & 1.3678 & 0.2978 \\
-    & &  & K562 & & 116,518,000   & 4.11 & 1.0441 &0.0504 \\
-    & &  & U2OS & & 3,255,111 &  10.05 & 1.0288 & 0.7714 \\
+    & \multirow{3}{1em}{Human} & \multirow{3}{1em}{GR} & IMR90 & 1& 47,443,803 & 8.86  & 1.3678 & 0.2978 \\
+    & &  & K562 & 1& 116,518,000   & 4.11 & 1.0441 &0.0504 \\
+    & &  & U2OS & 1& 3,255,111 &  10.05 & 1.0288 & 0.7714 \\
     & \multirow{3}{1em}{Human} & \multirow{3}{1em}{TBP} & \multirow{3}{1em}{K562} &1 & 61,046,382 &  12.01  & 1.1119
@@
     \hline
-    ChIP-seq PE & Exp. $+O_2$ & No Rif. &  & 1 & 27,665,432 & 4.01 & 
+    ChIP-seq PE & Exp. $+O_2$ & No Rif. & 1 & 1 & 27,665,432 & 4.01 & 
     3.9582 & 0.3869 \\
-    (P1) & Exp. $-O_2$ & No Rif. &  & 2 & 44,707,340 & 3.56 & 
+    (P1) & Exp. $-O_2$ & No Rif. & 1 & 2 & 44,707,340 & 3.56 & 
     3.3045 & 0.3134 \\
     \hline
-    ChIP-seq SE & Exp. $+O_2$ & No Rif. &  & 1 & 7,456,068 & 3.27 & 
+    ChIP-seq SE & Exp. $+O_2$ & No Rif. & 1 & 1 & 7,456,068 & 3.27 & 
     2.3863 & 0.5629 \\
-    (S1) & Exp. $-O_2$ & No Rif. &  & 2 & 11,467,086 & 2.91 & 
+    (S1) & Exp. $-O_2$ & No Rif. & 1 & 2 & 11,467,086 & 2.91 & 
     2.1362 & 0.5452 \\
@@
 
-DNA libraries generated by ChIP-exo and ChIP-nexus protocols are
-expected to be less complex than the libraries generated by ChIP-seq
-because the numbers of positions to which the reads can align to are
-reduced due to the exonuclease digestion.
-%considerable amounts of reads are being mapped to specific positions.
-This affects the interpretation of the PBC, which is defined as the
-ratio of the number of genomic positions to which exactly one read
-maps to the number of genomic positions to which at least one read
-maps.  For ChIP-seq samples, low PBC values (e.g., $\le 0.5$) indicate
-high levels of PCR amplification bias, i.e., PC bottleneck, unless the
-sequencing depth is high enough to sature all targets of the factor
-profiled.  In contrast, for ChIP-exo/nexus, exonuclease digestion will
-lead to reads with same exact $5^{\prime}$ end even before the PCR
-amplification step.
-%because of the exonuclease digestion,
-% several reads that their $5\prime$ end was digested to
-%the same position before the amplification step.
-We note  that  the PBC values are especially low
-for deeply sequenced ChIP-exo and ChIP-nexus
-samples; however, this does not automatically indicate severe 
-bottlenecking as suggested by standard ChIP-seq guidelines.
-%, which by following the
-%ChIP-Seq guidelines it would indicate that those experiment show
-%severe bottlenecking problems, which may incorrectly suggest that the
-%positions with a large amount of aligned reads are being caused by PCR
-%amplification rather than observed genomic signal.
-
-The Strand Cross-Correlation (SCC), introduced by \cite{strandcc}, is the most commonly used quality metric in assessing 
-ChIP-seq enrichment quality. It aims to quantify  how well the reads mapped to each
-strand are clustered around the locations of the protein-DNA interaction sites by calculating the Pearson correlation among forward and backward strands reads by shifting them across a range that covers both the read length of the experiment and the expected average fragment length.
+DNA libraries generated by ChIP-exo and ChIP-nexus protocols are expected to be less complex than the libraries generated by ChIP-seq because the numbers of positions to which the reads can align to are reduced due to the exonuclease digestion.
+This affects the interpretation of the PBC, which is defined as the ratio of the number of genomic positions to which exactly one read maps to the number of genomic positions to which at least one read maps.
+For ChIP-seq samples, low PBC values (e.g., $\le 0.5$) indicate high levels of PCR amplification bias, i.e., PC bottleneck, unless the sequencing depth is high enough to sature all targets of the factor profiled.
+In contrast, for ChIP-exo/nexus, exonuclease digestion will lead to reads with same exact $5^{\prime}$ end even before the PCR amplification step.
+We note  that  the PBC values are especially low for deeply sequenced ChIP-exo and ChIP-nexus samples; however, this does not automatically indicate severe  bottlenecking as suggested by standard ChIP-seq guidelines.
+
+The Strand Cross-Correlation (SCC), introduced by \cite{strandcc}, is the most commonly used quality metric in assessing ChIP-seq enrichment quality.
+It aims to quantify  how well the reads mapped to each strand are clustered around the locations of the protein-DNA interaction sites by calculating the Pearson correlation among forward and backward strands reads by shifting them across a range that covers both the read length of the experiment and the expected average fragment length.
 Typical SCC profiles exhibit two local maxima: at the average fragment length and the read length.
 In high quality experiments with clear ChIP enrichment, the average fragment length maximum coincides with the global maximum.
-%, and usually it is expected to observed two local
-%maxima, one when the profiles are shifted by the average read length
-%and another when the profiles are shifted by the unobserved fragment
-%length. 
-%In a high quality ChIP-Seq dataset the last one is also the
-%SCC global maxima. 
-In an idealized ChIP-exo experiment where the DNA fragments are
-digested to the boundaries of the protein-DNA interaction sites, we
-would expect the SCC profile to maximize at the motif length
-indicating clustering of the forward and reverse strand reads around
-the binding site. Figure~\ref{fig:1}D displays the SCC curves for the
-CTCF HeLa samples where the ChIP-exo curve shows local maxima at the
-motif and read lengths, while the SE ChIP-seq curves have a local
-maxima at the read length and a global maxima at the average fragment
-length. SCC profiles for other samples are available in Supplementary
-Figures \ref{sfig:scc1} to \ref{sfig:scc8}. The read length and motif
-length maxima are often in close proximity of each other; as a result,
-this renders QC metrics such as the Normalized Strand
-Cross-Correlation (NSC) or the Relative Strand Cross-Correlation (RSC)
-harder to interpret; however, the profile itself seems informative
-about the enrichment signal in ChIP-exo/nexus experiments \SK{this
-  seems true at least based on my initial look at the plots, but we
-  should be cautious in reporting this}. \RW{i agree, and additionally
-  Carroll et al, 2014 kinda agrees with us that the SCC-related
-  measure may not be useful since the fragment length coincides with
-  read length}
-
- % %http://journal.frontiersin.org/article/10.3389/fgene.2014.00075/full#h5
- %  the paragraph is a little bit above conclusions
-
-\subsection*{ChIP-exo quality control pipeline}
+In an idealized ChIP-exo experiment where the DNA fragments are digested to the boundaries of the protein-DNA interaction sites, we would expect the SCC profile to maximize at the motif length indicating clustering of the forward and reverse strand reads around the binding site.
+Figure~\ref{fig:1}D displays the SCC curves for the CTCF HeLa samples where the ChIP-exo curve shows local maxima at the motif and read lengths, while the SE ChIP-seq curves have a local maxima at the read length and a global maxima at the average fragment length.
+SCC profiles for other samples are available in Supplementary Figures \ref{sfig:scc1} to \ref{sfig:scc8}.
+The read length and motif length maxima are often in close proximity of each other; as a result, this renders QC metrics such as the Normalized Strand Cross-Correlation (NSC) or the Relative Strand Cross-Correlation (RSC) harder to interpret; however, the profile itself seems informative about the enrichment signal in ChIP-exo/nexus experiments.
+
+% \SK{this seems true at least based on my initial look at the plots, but we should be cautious in reporting this}. 
+% \RW{i agree, and additionally Carroll et al, 2014 kinda agrees with us that the SCC-related measure may not be useful since the fragment length coincides with read length}
+
+% http://journal.frontiersin.org/article/10.3389/fgene.2014.00075/full#h5 the paragraph is a little bit above conclusions
+
+% We continued our exploration by investigating whether the current state-of-the-art QC pipelines for ChIP-Seq are suitable for ChIP-exo %and ChIP-nexus.
+In Tables \ref{tab:qc_sig} and \ref{tab:qc} we calculated a collection of the commonly used ChIP-Seq QC metrics using the ChIP-exo and ChIP-nexus experiments instead: Normalized Strand Cross-Correlation (NSC), Relative Strand Cross-Correlation and PCR Bottleneck Coefficient (PBC) defined as in \url{https://genome.ucsc.edu/ENCODE/qualityMetrics.html}\nocite{encode_qc}.
+
+% considerable amounts of reads are being mapped to specific positions.
+
+% because of the exonuclease digestion, several reads that their $5\prime$ end was digested to the same position before the amplification step.
+
+%, which by following the ChIP-Seq guidelines it would indicate that those experiment show severe bottlenecking problems, which may incorrectly suggest that the positions with a large amount of aligned reads are being caused by PCR amplification rather than observed genomic signal.
+
+%, and usually it is expected to observed two local maxima, one when the profiles are shifted by the average read length and another when the profiles are shifted by the unobserved fragment length.
+In a high quality ChIP-Seq dataset the last one is also the SCC global maxima.
+
+\subsection*{ChIP-exo quality control pipeline \pname{}}
 
@@
 
-\subsubsection*{Saturation analysis with ChIP-exo and ChIP-seq: Systematic comparison of ChIP-seq vs ChIP-exo under
+\subsubsection*{Saturation analysis with ChIP-exo and ChIP-seq: Systematic comparison of ChIP-seq vs. ChIP-exo under
   varying sequencing depths}
@@
 identified whereas Figure~\ref{fig:7}B depicts the number of
-identified binding events, i.e., each candidate region can harbor
+identified binding events, i.e., each candidate region can harbour
 multiple binding events.  In terms of the number of binding events,
@@
 comparison between ChIP-exo and PE ChIP-seq datasets using our
-\textit{E. coli} $\sig$ samples.  This comparison revealed that PE
+\textit{E. Coli} $\sig$ samples.  This comparison revealed that PE
 ChIP-seq compares much more competitively with ChIP-exo compared to SE
@@
 
-\subsubsection*{\textit{E. coli} ChIP-exo and ChIP-seq samples}
+\subsubsection*{\textit{E. Coli} ChIP-exo and ChIP-seq samples}
 
@@
 ($OD_{600} = 0.3$ using a Perkin Elmer Lambda 25 $UV/Vis$
-Spectrophotometer). WT \textit{E. coli} K-12 MG1655 was used for the
+Spectrophotometer). WT \textit{E. Coli} K-12 MG1655 was used for the
 E1, P1 and S1 experiments and WT \textit{E. coli} K-12 RL3000 was used
@@
 
-\RW{ \textit{Library preparation and sequencing.} For ChIP-Seq
+\RW{ \textit{Library preparation and sequencing.} For ChIP-seq
   experiments, $10$ ng of immunoprecipitated and purified DNA
@@
   submitted to the University of Wisconsin-Madison DNA Sequencing
-  Facility for ChIP-Seq library preparation. Samples were sheared to
+  Facility for ChIP-seq library preparation. Samples were sheared to
   $200$ - $500$ nt during the IP process to facilitate library
@@
 
-We partitioned the \textit{E. coli} genome into non-overlaping
+We partitioned the \textit{E. coli} genome into non-overlapping
 intervals, i.e., bins, of length 150 bps and counted the number of
@@
   get shorter we are observing regions that behave more like ChIP-seq
-  in than ChIP-exo. in fig10A we can see that the mean distance
+  in than ChIP-exo. in 0A we can see that the mean distance
   between binding sites is below 100 for the majority of the peaks.}

### 2016-07-07 paper/chipexo_QC_vs2.tex
edited the results part and figures that correspond to the first results part

@@
 The read length and motif length maxima are often in close proximity of each other; as a result, this renders QC metrics such as the Normalized Strand Cross-Correlation (NSC) or the Relative Strand Cross-Correlation (RSC) harder to interpret; however, the profile itself seems informative about the enrichment signal in ChIP-exo/nexus experiments.
-
 % \SK{this seems true at least based on my initial look at the plots, but we should be cautious in reporting this}. 
 % \RW{i agree, and additionally Carroll et al, 2014 kinda agrees with us that the SCC-related measure may not be useful since the fragment length coincides with read length}
-
 % http://journal.frontiersin.org/article/10.3389/fgene.2014.00075/full#h5 the paragraph is a little bit above conclusions 
-
 % We continued our exploration by investigating whether the current state-of-the-art QC pipelines for ChIP-Seq are suitable for ChIP-exo %and ChIP-nexus.
 In Tables \ref{tab:qc_sig} and \ref{tab:qc} we calculated a collection of the commonly used ChIP-Seq QC metrics using the ChIP-exo and ChIP-nexus experiments instead: Normalized Strand Cross-Correlation (NSC), Relative Strand Cross-Correlation and PCR Bottleneck Coefficient (PBC) defined as in \url{https://genome.ucsc.edu/ENCODE/qualityMetrics.html}\nocite{encode_qc}.
-
 % considerable amounts of reads are being mapped to specific positions.
-
 % because of the exonuclease digestion, several reads that their $5\prime$ end was digested to the same position before the amplification step.
-
 %, which by following the ChIP-Seq guidelines it would indicate that those experiment show severe bottlenecking problems, which may incorrectly suggest that the positions with a large amount of aligned reads are being caused by PCR amplification rather than observed genomic signal.
-
 %, and usually it is expected to observed two local maxima, one when the profiles are shifted by the average read length and another when the profiles are shifted by the unobserved fragment length.
@@
 
-We first present the overall pipeline and then discuss individual
-components with a case study using ChIP-exo data of FoxA1 from
-\cite{exoillumina} and ChIP-nexus data from \cite{chipnexus}.
-Figure~\ref{fig:2} summarizes the 4-step pipeline.  Given
-aligned reads from a ChIP-exo/nexus sample, the first step partitions
-the reference genome into islands by keeping the non-digested ChIP-exo
-regions.  In step 2, the total number of extended reads (by the
-experiment's read length) \SK{check: extended or not? What do we
-  extend to in ChIP-exo?} \RW{right now, we don't extend. we use the
-  coverage plot with the experiment's read length to partition the
-  coverage into the regions. In the package I left it as a parameter
-  but I don't is going to change a lot. For example, the ChIP-exo
-  datasets read length is ~45 and it's regions seems to be composed by
-  scattered reads in both sides. So, I think that using a higher value
-  will only made the regions wider but will not increase the number of
-  unique position or the depth by a lot} overlapping each island
-($D_i$) and the number of unique island positions with at least one
-aligned read re recorded ($(U_i)$. Then, three summary statistics
-$\text{ARC}_i$, $\text{URC}_i$, and $\text{FSR}_i$ are computed for
-each region $i$.  $\text{ARC}_i$ denotes the \textit{average read
-  coefficient} and is defined as the ratio of the \# of reads in
-island $i$ ($D_i$) to the width of the island $i$ ($W_i$);
-$\text{URC}_i$, \textit{unique read coefficient}, quantifies the
-inverse of the effective coverage and is defined as the ratio of the
-\# of genomic positions with at least one aligned read within island
-$i$ ($U_i$) to the \# of reads in island $i$ ($D_i$); and
-$\text{FSR}_i$ denotes the proportion of forward strand reads.
+To address the limitations of available analytical exploration approaches discussed above, we developed \pname{}.
+We first present the overall pipeline and then discuss individual components with a case study using ChIP-exo data of FoxA1 from \cite{exoillumina} and ChIP-nexus data from \cite{chipnexus}.
+Figure~\ref{fig:2} summarizes the 4-step pipeline.
+Given aligned reads from a ChIP-exo/nexus sample, the first step partitions the reference genome into islands by keeping the non-digested ChIP-exo regions.
+In step 2, the total number of reads overlapping each island ($D_i$) and the number of island positions with at least one aligned read are recorded $(U_i)$.
+Then, three summary statistics $\text{ARC}_i$, $\text{URC}_i$, and $\text{FSR}_i$ are computed for each region $i$.
+$\text{ARC}_i$ denotes the \textit{average read coefficient} and is defined as the ratio of the \# of reads in island $i$ ($D_i$) to the width of the island $i$ ($W_i$); $\text{URC}_i$, \textit{unique read coefficient}, quantifies the inverse of the effective coverage and is defined as the ratio of the \# of genomic positions with at least one aligned read within island $i$ ($U_i$) to the \# of reads in island $i$ ($D_i$); and $\text{FSR}_i$ denotes the proportion of forward strand reads.
+Step 3 of the pipeline generates several diagnostic plots aimed at quantifying ChIP enrichment and strand imbalance and step 4 generates quantitative summaries of these diagnostic plots.
+
+% (extended by the experiment's read length)
+% \SK{check: extended or not? What do we extend to in ChIP-exo?}
+% \RW{right now, we don't extend.
+we use the coverage plot with the experiment's read length to partition the coverage into the regions.
+In the package I left it as a parameter but I don't is going to change a lot.
+For example, the ChIP-exo datasets read length is ~45 and it's regions seems to be composed by scattered reads in both sides.
+So, I think that using a higher value   will only made the regions wider but will not increase the number of unique position or the depth by a lot}
 %\begin{align*}
@@
 %\end{align*}
-Step 3 of the pipeline generates several diagnostic plots aimed at
-quantifying ChIP enrichment and strand imbalance and step 4 generates
-quantitative summaries of these diagnostic plots.
-
-Figure~\ref{fig:2}A presents the typical behavior of the URC vs. ARC
-plot for a high quality ChIP-exo sample. In general, the plot depicts
-two strong arms.  High URC values correspond to regions with reads
-concentrated on a small number of positions.  The left arm, with low
-ARC and varying URC values, corresponds to background islands, regions
-that are usually composed of scattered reads that were not digested
-during the exonuclease step. The right arm where the URC decreases as
-the ARC increases corresponds to regions that are usually ChIP
-enriched. \RW{Finally we quantify the shape of the ARC vs. URC plot by
-  the use of two estimated parameters: $\beta_1$ which
-  represents the average number of reads aligned to the unique
-  positions in large depth regions and $\beta_2$ which
-  represents the overall change in depth as the width varies across a
-  large set of regions.}  \RW{Figures~\ref{fig:2}B and \ref{fig:2}C
-  presents the typical behavior of the Region Composition and Forward
-  Strand Ratio distribution plots respectively, both quantify the
-  strand imbalance as part of the QC pipeline. The former depicts how
-  quickly the islands exclusively composed by fragments on a single
-  strand are filtered out as islands with higher depths are
-  observed. In a high quality sample, the proportion of islands with
-  reads from only one strand is expected to decrease rapidly as we
-  consider higher depth regions. In contrast, this proportion remains
-  approximately constant in lower quality samples. The latter plot
-  illustrates how quickly the quantiles of the FSR approaches to 0.5,
-  the expected FSR value in high quality samples. Even though not
-  every region in a ChIP-exo experiment is perfectly balanced, the
-  most enriched regions are expected to be composed by approximately
-  the same quantity of reads in both strands.}
-
-%and as the URCR decreases the library complexity does it as well, on
-%the other hand 
+Figure~\ref{fig:2}A presents the typical behaviour of the URC vs.
+ARC plot for a high quality ChIP-exo sample.
+In general, the plot depicts two strong arms.
+High URC values correspond to regions with reads concentrated on a small number of positions.
+The left arm, with low ARC and varying URC values, corresponds to background islands, regions that are usually composed of scattered reads that were not digested during the exonuclease step.
+The right arm where the URC decreases as the ARC increases corresponds to regions that are usually ChIP enriched.
+We quantify the shape of the ARC vs.
+URC plot by the use of two estimated parameters: $\beta_1$ which represents the average number of reads aligned to the unique positions in large depth regions and $\beta_2$ which represents the overall change in depth as the width varies across a large set of regions.
+These parameters are estimated by sampling experiments on the original samples.
+We provide further details on them later in the paper where we apply the pipeline to a large collection of ChIP-exo/nexus experiments.
+Figures~\ref{fig:2}B and \ref{fig:2}C   present the typical behaviour of the \textit{Region Composition} and \textit{Forward Strand Ratio distribution} plots, both quantify the strand imbalance as part of the QC pipeline respectively, the former depicts how quickly the islands exclusively composed by fragments on a single strand are filtered out as islands with higher depths are observed.
+In a high quality sample, the proportion of islands with reads from only one strand is expected to decrease rapidly as we consider higher depth regions.
+In contrast, this proportion remains approximately constant in lower quality samples.
+The latter plot illustrates how quickly the quantiles of the FSR approaches to 0.5, the expected FSR value in high quality samples.
+Even though not every region in a ChIP-exo experiment is perfectly balanced, the most enriched regions are expected to be composed by approximately the same quantity of reads in both strands.
+
+%and as the URCR decreases the library complexity does it as well, on the other hand 
 % \SK{Not sure what this next sentence is referring to: Finally we
@@
 %   the depth changes.}
-
 %median, since in a high quality sample it is
@@
 %strands.
-
 \subsubsection*{Application and validation of \pname{} with the FoxA1
@@
 
-We next use Fox A1 ChIP-exo datasets, with three biological replicates
-at comparable sequencing depths from mouse liver cells, to illustrate
-the QC pipeline.  Figure~\ref{fig:3}A presents URC vs. ARC plots
-for all three replicates. The first and third replicates exhibit a
-defined decreasing trend in URC as the ARC increases. This indicates
-that these samples exhibit a higher ChIP enrichment than the second
-replicate. On the other hand, the overall URC level from the first two
-replicates is higher than that of the third replicate level,
-elucidating that the libraries for the first two replicates are more
-complex than that of the third replicate.
-
-%As we discussed above,  background fragments are often digested by
-%the exonuclease enzyme in ChIP-exo experiments; therefore, the balance between the enrichment
-%and library complexity of an experiment is one of the key factors determining
-%the sample's data quality.
-%Using the Fox A1 in mouse liver cell lines generated by Serandour et
-%al., 2013\nocite{exoillumina} and these two quantities, we explored
-%the relationship between library complexity and experiment
-%enrichment.
-
+We next use FoxA1 ChIP-exo datasets, with three biological replicates at comparable sequencing depths from mouse liver cells, to illustrate the QC pipeline.
+Figure~\ref{fig:3}A presents URC vs.
+ARC plots for all three replicates.
+The first and third replicates exhibit a defined decreasing trend in URC as the ARC increases.
+This indicates that these samples exhibit a higher ChIP enrichment than the second replicate.
+On the other hand, the overall URC level from the first two replicates is higher than that of the third replicate level,elucidating that the libraries for the first two replicates are more complex than that of the third replicate.
+%\SK{I have a hard time interpreting B and C specifically, need to include something about their implications for replicates 2 and 3.}
+
+Figure~\ref{fig:3}B and \ref{fig:3}C display the Read Composition and FSR distribution plots, which highlight specific problems with replicates 2 and 3.
+Figure~\ref{fig:3}B, exhibits apparent decreasing trends in the proportions of regions formed by fragments in one exclusive strand.
+High quality experiments tend to show exponential decay in the proportion of single   stranded regions, while for the lower quality experiments, the trend may be linear or even constant (Supplement   Figure~\ref{sfig:qcsamp3}).
+FSR distributions of both of replicates 2 and 3 are more spread around their respective medians (Figure~\ref{fig:3}C).
+The approximation of the 0.1 and 0..9 quantiles indicate the aforementioned lower enrichment in the   second replicate and the low complexity in the third one.
+
+Overall, we conclude that replicate 1 is higher quality than both of replicates 2 and 3.
+We validate this observation with a motif analysis on the candidate binding regions identified from these replicates.
+A conservative approach to identify high quality binding regions (Materials and Methods) identify 7014, 1855 and 2187 regions for replicates 1, 2 and 3, respectively.
+The lower number of enriched regions from replicate 2 is consistent with the lower ChIP enrichment pattern in the $UCR$ vs.
+$ARC$ diagnostic plot.
+Scanning of these regions for the occurrence of FoxA1 sequence motif with the FIMO tool \cite{fimo} indicates that the first replicate outperforms the other two in terms of percentage of candidate regions with the FoxA1 motif  (Figure~\ref{fig:4}A).
+% \SK{Could we say binding events? or is the peak analysis on the whole binding region, i.e., peak? If so, we may want o    change "candidate regions" to peaks?} 
+% \RW{I used a set of high quality regions, i.e.
+using the ChIP-exo pipeline I filtered them by npos, FSR and depth.
+Figure~\ref{sfig:venn} show the overlap between these regions and the peaks (roughly speaking these regions are almost a proper subset of the peaks)}.
+
+Furthermore, Figure~\ref{fig:4}B displays the average normalized read coverage around the actual motif locations in the candidate binding regions.
+These coverage plots reveal that the ChIP signal is more defined for the first and third replicates than the second one, indicating overall strength of the ChIP enrichment in these samples compared to the second replicate.
+Figure~\ref{fig:4}C compares the FIMO scores among the three replicates, not-surprisingly confirming that the first replicate exhibits the highest quality.
+
+%As we discussed above,  background fragments are often digested by the exonuclease enzyme in ChIP-exo experiments; therefore, the balance between the enrichment and library complexity of an experiment is one of the key factors determining the sample's data quality.
+Using the Fox A1 in mouse liver cell lines generated by Serandour et al., 2013\nocite{exoillumina} and these two quantities, we explored the relationship between library complexity and experiment enrichment.
 %\textbf{Analysis of strand imbalance in FoxA1 ChIP-exo data}
-
-%The strand imbalance assessment is based in the observation that the
-%enriched regions usually are composed of a higher quantity of reads,
-%therefore we examined the FSR as the regions with lower depth are
-%being filtered out. This indicator is of particular importance, as it
-%evaluates the ``peak pair'' assumption that the original ChIP-exo
-%paper suggested and multiple ChIP-exo data analysis methods rely
-%on. For every ChIP-exo experiment, we calculated the global FSR and
-%noticed that for all experiments is roughly $0.5$, which means there
-%are approximately the same amount of reads in both strands.
-
-\SK{I have a hard time interpreting B and C specifically, need to
-  include something about their implications for replicates 2 and 3.}
-
-Figure~\ref{fig:3}B and \ref{fig:3}C display the Read Composition and
-FSR distribution plots, which highlight specific problems with
-replicates 2 and 3. \RW{In Figure~\ref{fig:3}B, we can observe
-  decreasing trends in the proportions of regions formed by fragments
-  in one exclusive strand. In a high quality experiment its is
-  expected to observe an exponential decay in the proportion of single
-  stranded regions, while in lower quality experiments the trend may
-  be linear of even constant (Supplement
-  Figure~\ref{sfig:qcsamp3}). In the FSR distribution plot, both
-  replicate 2 and 3 FSR's distributions are more spread around their
-  respective medians. The approximation of the 0.1 and 0..9 quantiles
-  to the median indicated the aforementioned lower enrichment in the
-  second replicate and the low complexity in the third one.
- }
- % In particular, \RW{both replicate 2 and replicate 3 FSR's
- %  distributions are more spread around the median. The approximation
- %  of the 0.1 and 0.9 quantiles to the median is indicative of the
- %  aforementioned lower enrichment of the second replicate, since there
- %  are less regions with a higher concentration of reads, this makes
- %  the FSR's distribution to be dominated by non-enriched regions. On
- %  the other hand, the slow approximation of the quantiles to the FSR
- %  median in the third replicate is indicative of low library
- %  complexity. In Figure~\ref{fig:3}C, we can observe an decreasing
- %  trend in the proportion of regions formed by fragments in one
- %  exclusive strand. In a high quality experiment it is expected to
- %  observe an exponential trend, while in lower quality experiments the
- %  trend may be linear or even constant (Supplement
- %  Figure~\ref{sfig:qcsamp3}).}
-
-%In order to study the strand imbalance of the FoxA1 factor ChIP-exo
-%experiments in mouse liver generated by Serandour et al.,
-%2013\nocite{exoillumina} we considered the visualizations shown in
-%Figure \ref{fig:strand}: Figure \ref{fig:strand}A presents the FSR's
-%behavior as the lower depth regions are being filtered out, while
-%Figure \ref{fig:strand}B) shows the proportion of regions that are
-%composed by reads from both strands against the regions formed by
-%reads in exclusively one strand. In a high quality dataset, it is
-%expected for all quantiles to quickly converging towards the median
-%(in panel A) or the regions formed by reads in one strand (either
-%forward or backward) to be formed by few fragments (in panel
-%B). 
-%Additionally, it is worth mentioning that the background in a ChIP-exo
-%experiment is expected to be formed by the undigested ChIP fragments
-%that are not bound to a protein, hence it is for the reads in both
-%strand to be imbalanced in a background region.
-
- Overall, we conclude that replicate 1 is higher quality than both of
- replicates 2 and 3.  We validate this observation with a motif
- analysis on the candidate binding regions identified from these
- replicates \SK{Could we say binding events? or is the peak analysis
-   on the whole binding region, i.e., peak? If so, we may want o
-   change "candidate regions" to peaks?} \RW{I used a set of high
-   quality regions, i.e. using the ChIP-exo pipeline I filtered them
-   by npos, FSR and depth. Figure~\ref{sfig:venn} show the overlap
-   between these regions and the peaks (roughly speaking these regions
-   are almost a proper subset of the peaks)}. Figure~\ref{fig:4}A
- summarizes the total numbers of regions identified from each
- replicate. The lower number of enriched regions from replicate 2 is
- consistent with the lower ChIP enrichment pattern in the $UCR$
- vs. $ARC$ diagnostic plot.  Scanning of these regions for the
- occurrence of FoxA1 sequence motif with the FIMO tool \cite{fimo}
- indicates that the first replicate outperforms the other two in terms
- of percentage of candidate regions with the FoxA1 motif
- (Figure~\ref{fig:4}B).
-
-%Considering the regions and summary statistics we obtained a set of
-%candidate sites for each replicate, the total number of candidate
-%sites is shown in Figure \ref{fig:enrich}B.
-%For this sites, we
-%extracted the sequence and searched for the FoxA1 motif using
-%\textbf{FIMO} \cite{fimo}. Figure \ref{fig:enrich}C shows the fraction
-%of candidate site where a motif was found. The first replicate
-%outperforms the rest in both number of candidate sites and proportion
-%of sites with motif; on the other hand we can see that the second
-%replicate candidate sites are more likely to contain the motif
-%sequence, but there more candidate sites in the third replicate than
-%in the second. 
-Furthermore, Figure~\ref{fig:4}C displays the average normalized
-read coverage around the actual motif locations in the candidate
-binding regions.  These coverage plots reveal that the ChIP signal is
-more defined for the first and third replicates than the second one,
-indicating overall strength of the ChIP enrichment in these samples
-compared to the second replicate.
+%The strand imbalance assessment is based in the observation that the enriched regions usually are composed of a higher quantity of reads, therefore we examined the FSR as the regions with lower depth are being filtered out.
+This indicator is of particular importance, as it evaluates the ``peak pair'' assumption that the original ChIP-exo paper suggested and multiple ChIP-exo data analysis methods rely on.
+For every ChIP-exo experiment, we calculated the global FSR and noticed that for all experiments is roughly $0.5$, which means there are approximately the same amount of reads in both strands.
+% In particular, \RW{both replicate 2 and replicate 3 FSR's distributions are more spread around the median.
+The approximation of the 0.1 and 0.9 quantiles to the median is indicative of the aforementioned lower enrichment of the second replicate, since there are less regions with a higher concentration of reads, this makes the FSR's distribution to be dominated by non-enriched regions.
+On the other hand, the slow approximation of the quantiles to the FSR median in the third replicate is indicative of low library complexity.
+In Figure~\ref{fig:3}C, we can observe an decreasing trend in the proportion of regions formed by fragments in one exclusive strand.
+In a high quality experiment it is expected to observe an exponential trend, while in lower quality experiments the trend may be linear or even constant (Supplement Figure~\ref{sfig:qcsamp3}).}
+%In order to study the strand imbalance of the FoxA1 factor ChIP-exo experiments in mouse liver generated by Serandour et al., %2013\nocite{exoillumina} we considered the visualizations shown in Figure \ref{fig:strand}: Figure \ref{fig:strand}A presents the FSR's behavior as the lower depth regions are being filtered out, while Figure \ref{fig:strand}B) shows the proportion of regions that are composed by reads from both strands against the regions formed by reads in exclusively one strand.
+In a high quality dataset, it is expected for all quantiles to quickly converging towards the median (in panel A) or the regions formed by reads in one strand (either forward or backward) to be formed by few fragments (in panel B).
+% Additionally, it is worth mentioning that the background in a ChIP-exo experiment is expected to be formed by the undigested ChIP fragments that are not bound to a protein, hence it is for the reads in both strand to be imbalanced in a background region.
+% Figure~\ref{fig:4}A summarizes the total numbers of regions identified from each  replicate. 
+%Considering the regions and summary statistics we obtained a set of candidate sites for each replicate, the total number of candidate sites is shown in Figure \ref{fig:enrich}B.
+For this sites, we extracted the sequence and searched for the FoxA1 motif using \textbf{FIMO} \cite{fimo}.
+Figure \ref{fig:enrich}C shows the fraction of candidate site where a motif was found.
+The first replicate outperforms the rest in both number of candidate sites and proportion of sites with motif; on the other hand we can see that the second replicate candidate sites are more likely to contain the motif sequence, but there more candidate sites in the third replicate than in the second.
 % however,  overall, the signal pattern is reproducible in all of the three biological replicates.
-Figure~\ref{fig:4}D further highlights the overall quality of the
-identified motif sequences for each replicate and suggest that
-libraries with high library complexity (replicates 1 and \RW{2})
-capture binding sites with better motif
-matches. \RW{Figure~\ref{fig:4}E compares the top FIMO scores among
-  the three replicates, not-surprisingly confirming that the first
-  explicate exhibits the highest quality. Finally, Figure~\ref{fig:4}F
-  compares the Strand Imbalance by considering the regions overlapping
-  with their respective sets of ChIP-exo peaks. Not surprisingly, for
-  the replicate 2 we can see that the Imbalance distribution differ
-  among both classes, which is caused by the aforementioned lower
-  enrichment of that sample; for the remaining replicates we can see
-  that the Imbalance is more similar when regions composed by more
-  reads are considered. This suggests that enriched regions are more
-  likely to be balanced, therefore this suggest that the ``peak-pair''
-  assumption may not hold in non-enriched regions.}
-
- % Hence, in a low enriched sample the
- %  ``peak-pair'' assumption does not hold globally.}
-% \SK{I also don't follow the discussion for panel C: Figure
-% \ref{fig:4}C compares the strand imbalance when ChIP-exo islands 
-% overlap with high quality peaks. It is noticeable that for regions
-% composed by a large amount of reads, it is harder to distinguish their
-% peaks by considering only the strand imbalance, hence in a high
-% quality ChIP-exo experiment the background is expected to show a
-% higher imbalance that the enriched regions. In conclusion, Figure
-% \ref{fig:4} shows that the global FSR does not accurately
-% represent a ChIP-exo experiment's strand imbalance locally, hence the
-% ``peak pair'' assumption does not completely hold in every ChIP-exo
-% enriched region.}
-% \SK{Worth thinking a bit more about the interpretation.}
-
-%In summary,  the diagnostics plots of the QC pipeline highlights that  observing a defined
-%decreasing trend in $\mbox{URCR}$ as $\mbox{ARC}$ increases implies
-%ChIP enrichment and multiple samples can be compared with respect to the degree of this trend. 
-%Second, for experiments that are not deeply
-%sequenced, low $\mbox{URCR}$ values represent low complexity. 
-%decreases the amount of sites of motif found but it doesn't modify the
-%read distribution around the motif. Finally, libraries with higher
-%complexity returns sequences with a better motif match. \SK{We do not have enough analysis with motif validation to conclude these.}
-
-\subsubsection*{High sequencing depth may confound low-complexity
-  library issues}
+% Figure~\ref{fig:4}D further highlights the overall quality of the identified motif sequences for each replicate and suggest that libraries with high library complexity (replicates 1 and \RW{2}) capture binding sites with better motif matches.
+% Finally, Figure~\ref{fig:4}F compares the Strand Imbalance by considering the regions overlapping with their respective sets of ChIP-exo peaks.
+Not surprisingly, for the replicate 2 we can see that the Imbalance distribution differ among both classes, which is caused by the aforementioned lower enrichment of that sample; for the remaining replicates we can see that the Imbalance is more similar when regions composed by more reads are considered.
+This suggests that enriched regions are more  likely to be balanced, therefore this suggest that the ``peak-pair'' assumption may not hold in non-enriched regions.
+%In summary,  the diagnostics plots of the QC pipeline highlights that  observing a defined decreasing trend in $\mbox{URCR}$ as $\mbox{ARC}$ increases implies ChIP enrichment and multiple samples can be compared with respect to the degree of this trend.
+Second, for experiments that are not deeply sequenced, low $\mbox{URCR}$ values represent low complexity.
+decreases the amount of sites of motif found but it doesn't modify the read distribution around the motif.
+Finally, libraries with higher complexity returns sequences with a better motif match.
+\SK{We do not have enough analysis with motif validation to conclude these.}
+\subsubsection*{High sequencing depth may confound low-complexity library issues}
+
 We next evaluated every sample listed in Tables~\ref{tab:qc} and
@@
 
-\RW{The enumeration of some figures is a little bit off, I am going to
-  fix it after we finish the text}
-  % \caption{\textbf{Description of ChIP-exo, SE ChIP-Seq and PE
-  %     ChIP-Seq:} A TF is bound to the forward (red) and backward
-  %   (blue) strand of the DNA. Then, is sonicated: For ChIP-exo a
-  %   exonuclease enzyme (orange hexagon) trims the $5\prime$ ends of
-  %   each DNA fragment to a fixed distance from the bound protein,
-  %   finally is subjected to Immunoprecipitation and amplification. For
-  %   both ChIP-exo and SE ChIP-Seq an adapter is ligated (green
-  %   triangles) at the $5\prime$ ends, while for PE ChIP-Seq is ligated
-  %   to both ends.}
 \begin{figure}[h!]
   \centering
-  \includegraphics[width =
-  .45 \textwidth]{figures/fig1/chip_explanation_withEnzyme_nexus.pdf}
-  \includegraphics[width = .45\textwidth,page = 3]{figures/fig1/ChIPseqPET_ChIPexo_tagCount_comparison.pdf}
-  \includegraphics[width =.3\textwidth]{figures/fig1/forward_strand_ratio_comp_old.pdf}
-\includegraphics[width = .6 \textwidth]{figures/fig1/scc_ctcf2.pdf}
-  \includegraphics[width = .4\textwidth,page = 1]{figures/fig1/eukaryotic_bias_CTCF.pdf}
-  \includegraphics[width = .4\textwidth,page = 2]{figures/fig1/eukaryotic_bias_CTCF.pdf}
+  \includegraphics[width = .95\textwidth]{figures/fig1/fig1.pdf}
+%  \includegraphics[width =
+%  .45 \textwidth]{figures/fig1/chip_explanation_withEnzyme_nexus.pdf}
+%  \includegraphics[width = .45\textwidth,page = 3]{figures/fig1/ChIPseqPET_ChIPexo_tagCount_comparison.pdf}
+%  \includegraphics[width =.3\textwidth]{figures/fig1/forward_strand_ratio_comp_old.pdf}
+%\includegraphics[width = .6 \textwidth]{figures/fig1/scc_ctcf2.pdf}
+%  \includegraphics[width = .4\textwidth,page = 1]{figures/fig1/eukaryotic_bias_CTCF.pdf}
+%  \includegraphics[width = .4\textwidth,page = 2]{figures/fig1/eukaryotic_bias_CTCF.pdf}
   \caption{A) A TF is bound to the forward (red) and backward (blue)
@@
   \centering
-  \includegraphics[width = \textwidth]{figures/fig2/coverage_diagram4.pdf}
+  \includegraphics[width = .95\textwidth]{figures/fig2/fig2.pdf}
   \caption{\RW{The ChIP-exo reads are partitioned by keeping the
@@
   \centering
-  \includegraphics[width = .8\textwidth,page = 1]{figures/fig3/Carroll_FoxA1_mouse_enrichment.pdf}
-  \includegraphics[width = .8\textwidth,page = 3]{figures/fig3/Carroll_FoxA1_mouse_strand_imbalance.pdf} 
-  \includegraphics[width = .8\textwidth,page = 1]{figures/fig3/Carroll_FoxA1_mouse_strand_imbalance.pdf} 
+%  \includegraphics[width = .8\textwidth,page = 1]{figures/fig3/Carroll_FoxA1_mouse_enrichment.pdf}
+%  \includegraphics[width = .8\textwidth,page = 3]{figures/fig3/Carroll_FoxA1_mouse_strand_imbalance.pdf} 
+%  \includegraphics[width = .8\textwidth,page = 1]{figures/fig3/Carroll_FoxA1_mouse_strand_imbalance.pdf} 
+    \includegraphics[width = .95\textwidth]{figures/fig3/fig3.pdf}
   \caption{\RW{Diagnostic plots generated by \pname{}. Comparison of
@@
   \centering  
-  \includegraphics[width = .23 \textwidth, page = 1]{figures/fig4/FoxA1_barplots.pdf}
+%  \includegraphics[width = .23 \textwidth, page = 1]{figures/fig4/FoxA1_barplots.pdf}
   \includegraphics[width = .23 \textwidth, page = 3]{figures/fig4/FoxA1_barplots.pdf}
   \includegraphics[width =.4\textwidth]{figures/fig4/FoxA1_profiles_around_motif.pdf}
-  \includegraphics[width = .25\textwidth,page = 2]{figures/fig4/FoxA1_matched_motif_sequence.pdf}
-  \includegraphics[width = .25\textwidth,page = 3]{figures/fig4/FoxA1_matched_motif_sequence.pdf}
-  \includegraphics[width = .25\textwidth,page = 1]{figures/fig4/FoxA1_matched_motif_sequence.pdf}  
+%  \includegraphics[width = .25\textwidth,page = 2]{figures/fig4/FoxA1_matched_motif_sequence.pdf}
+%  \includegraphics[width = .25\textwidth,page = 3]{figures/fig4/FoxA1_matched_motif_sequence.pdf}
+%  \includegraphics[width = .25\textwidth,page = 1]{figures/fig4/FoxA1_matched_motif_sequence.pdf}  
   \includegraphics[width = .8\textwidth,page = 1]{figures/fig4/FoxA1_fimo_score.pdf}
-  \includegraphics[width = .8\textwidth,page = 2]{figures/fig4/strand_imbalance_FoxA1.pdf}
+  \includegraphics[width = .5\textwidth]{figures/fig4/TBP_scores_comparison.pdf}
+%  \includegraphics[width = .8\textwidth,page = 2]{figures/fig4/strand_imbalance_FoxA1.pdf}
   \caption{Using the mouse FoxA1 experiment from \cite{exoillumina}
@@
   % \includegraphics[width = .45  \textwidth,page = 2]{figures/fig5/K562_TBP_sample.pdf}
-\includegraphics[width = .5\textwidth]{figures/fig5/TBP_scores_comparison.pdf}
 \caption{\RW{Comparison of all estimated A) $\beta_1$ and B) $\beta_2$
